### 2020-12-19 documents/marlin-paper.Rmd
updated version of paper working up until MPA generation

---
title: "The Role of MPAs in Bycatch Reduction"
author: "Dan Ovando"
date: "12/11/2020"
output: 
  bookdown::pdf_document2: default
  bookdown::html_document2: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(here)
library(bookdown)
```

Marine protected areas (MPAs) are a growing part of the marine conservation toolbox, with 7.4% of the world’s oceans currently protected and, and bodies such as the IUCN calling for 30% protection by 2030. Many of the largest of these MPAs are being implemented in the form of “blue-water” MPAs largely covering pelagic ecosystems, characterized by highly mobile species and fishing fleets using a diverse mix of static and dynamic ocean features. These systems operate very differently from the coastal areas where modern MPA design practices have been developed. Pelagic systems are also characterized by combinations of relatively abundant species targeted by fishing, such as most tuna stocks, and highly threatened species caught as incidental “by-catch” in targeted fishing operations, such as oceanic whitetip sharks. Blue-water MPAs are increasingly being proposed specifically to benefit these by-catch populations while minimizing losses or where possible providing benefits to fisheries.

Pelagic ecosystems present both opportunities and challenges for MPAs. The highly productive and mobile nature of targeted fisheries in these systems may mean that highly targeted MPAs can be designed that provide effective conservation at minimal fishery costs. Alternatively, if individual by-catch species do not share common habitat preferences or fishery vulnerability, protection in one area targeted to benefit one species may negatively impact another through effort displacement. Here, we use a case study motivated by the upcoming expansion of blue-water MPAs throughout the Federated States of Micronesia to assess the sensitivity of by-catch populations and targeted fisheries to MPA design and key life-history variables.

The outcome of MPAs for fishing and conservation are highly sensitive to a wide range of economic and ecological factors. In order to constrain the possible parameter space, we tune a simulation model based around best available knowledge of the species and fleets operating in the areas of the western central pacific ocean (WCPO) while allowing key unknowns to vary, in particular the nature of adult and larval movement, the degree of habitat overlap between different species, the economic incentives facing fishing fleets, and the design strategy used in MPA placement XX. Most conventional MPA design models are either static, single species, or both. We present a dynamic, multi-species simulation model that allows us to cohesively assess the impact of MPAs on a suite of by-catch and targeted species collectively affected by multiple fishing fleets. Our goal here is not to provide tactical recommendations on the design of blue-water MPAs throughout the WCPO, but rather to provide a strategic evaluation of the sensitivity of by-catch populations and targeted fisheries to key unknowns and design choices in MPA design.

We show 
  - something about how when species are homogeneous in space, conservation gains are basically guaranteed except for constant catch scenearios, which need to be considered in tuna-type scenarios
  - When we introduce heterogeneity in stock structure roughly reflective of distributions in the WCPO we do / do not begin to see tradeoffs in conservation outcomes among bycatch species
    - quantify if there is a concave function to this
  - Specific design paradigms are / are not more / less resilient to this issue
  - Compare tradeoffs between MPA designs that minimze internal bycatch tradeoffs (there have to be none if MPA = 100%) and fishery outcomes

We show that MPAs increased biomass of by-catch species in XX% of simulations, and decreased them in XX% of simulations. On average a 1% increase in by-catch species resulted in a XX% change in fishery catches… Outcomes were most sensitive to XX and XX, and least sensitive to XX and XX. We do not suggest that these same results would hold for other systems, but rather demonstrate the importance / resilience of MPAs to key uncertain variables. In particular for pelagic ecosystems, our work demonstrates the importance of considering the collective impacts of space, vulnerability, life history, and fleet dynamics in MPA design for conservation.

## Figures

Are there or are there not tradeoffs in conservation and fisheries yields

What are the explicit considerations of

Interesting considerations we do biodiversity MPAs we define biodiversity as system diversity and we put all the species in there, threatened bycatch species become the the focus and stated objective is ramp up the value of bycatch species . There might be instances where just the biodiversity base actually gets you a better outcome than . There might be overlap and not.

Dig into that mismatch

We can explore that mismatch between is the issue q and is the issue of spatial habitat

### Fig.1  Motivating figure - basically something like the shortfin mako example

### Fig.2  Case study figure for WCPO: shows degree of overlap in species abundance in space and time, as well as a platform for some examples of MPA networks designed around different paradigms

### Fig.3 Tradeoff plots across species, though will need to put some thought into visualizing across species. But something to convey the idea of whether there are tradeoffs among species in the case study of the WCPO high seas style system. Interestingly, you know that if you closed 100% all benefit, so doing this along a gradient is interesting, are there tradeoffs at low sizes that disappear with higher? Actually you might expect a dome-shaped tradeoff, where there are no tradeoffs at low size since little displacement, largest tradeoffs in the middle where you maximize displacement but still haven’t maximized protection, and then no tradeoffs at the extreme where everything is protected. If you have a small enough number of species, say less than 10, then can do color / shape by species, mpa size on x axis, change in biomass on y axis, and connected lines for design paradigms. If not too busy, would be good to have a counterfactual where everything is the same except species are all homogeneously distributed

### Fig 4. Similar idea but, now bringing fishing in there. Maybe collapsing down a few axes, so you do like percent increase in total bycatch by total fishing, or some measure of collective benefit / loss, as a function of MPA size. Maybe highlight inflection points that seem to minimize fishery losses while also minimizing tradeoffs with bycatch

### Fig 5 + Figures showing a few select “what if the world were different”, i.e. how important is movement rate, DD form, etc.

# Scraps

What are the considerations of protection you need to make good decisions. It’s not always obvious how to design things to get the outcome you want. If we could make the case that something like closing the high seas is insufficient conservation management without conservation management inside EEZs / what is the

Need to do an optimization here, calculate the marginal value of added cells.

What’s this about?
Pointing out a problem, which to make impactful would require making the case that there is a problem with current design strategies. e.g.. “MPAs are often designed like this (simple dispersive models with constant habitat, or MARXSAN style things), and doing that does/does not create problems
A tactical evaluation. Picking a handful of high profile MPAs, parameterizing all the knowns as well as possible, iterating over unknowns, and see what the range of outcomes are. Cleaner, more impactful, but also opens it up (depending on the results) to a details fight about very specific modeling choices
How plausible are losses? Maybe we drill down on this a bit. From a purely bycatch perspective, the usual questions of fisheries tradeoffs aren’t on the table (but can add to the discussion considerations of how much target fishing you give up). Given that then, while it’s possible that “targeted” interventions would work better (ignoring costs/enforcement, etc), there’s really no reason not to do MPAs. Unless, you actually end up concentrating fishing pressure outside on populations. So, we consider how plausible is it for that to really happen. Issue here is that seems likely to be pretty hand-wavey. Most likely that the vast majority of simulations will show positive impacts, but have no way of assigning probabilities to any one state of nature.

The Role of Blue-Water Marine Protected Areas in Bycatch Reduction

What’s the question: How effective are Blue-Water MPAs in reducing fishing mortality of important pelagic bycatch species?

Another angle. There’s going to be a lot of unknowns in designing BWMPA. How  much does not fully understanding a few key life history (e.g. movement) matter, and how much does the role that the species plays in the broader fleets portfolio matter

Could simulate habitat from a multivariate normal, where you use the sigma to control how similar rasters are across species

Marine protected areas are increasingly proposed to protect biodiversity in pelagic systems. However, maby paradigms of mpa design come from coastal systems. Here we ask, how likely are blue-water MPAs to provide conservation benefits to bycatch species, and what are the key uncertainties underlying these outcomes? We begin with a general discussion of the ways in which traits of open-ocean systems might influence the conservation performance of BWMPAs. Given the massive parameter space, we then constrain ourselves to a styulized version of the western central pacitic ocean to ask, how feasible is it to provide broad protection of bycatch species through common MPA design paradigms in the context of this system. Our results demonstrate the critical importance of XX and XX in designing MPAs in pelagic systems, and highlight the potential risks of not accounting for XX. 
Introductionhere

Bycatch is a significant driver of population declines for many pelagic species
Sharks, seabirds, mammals, etc
MPAs increasingly turned to as a solution for conservation problems, etc. 
MPA science is largely based on dynamics of coastal systems. Open-ocean ecosystems have very different dynamics that require dedicated research
Burgess et al. (2018) estimated the reductions in fishing mortality needed to halt declines for a number of bycatch species
Here we ask, what are the key drivers of bycatch conservation outcomes from MPAs?
There are all of these species; you reduce fishing mortality. You can either do fisheries management, or you can do MPAs, let’s focus in on the MPA side. This could reduce fishing mortality, or it could do, and it depends on the dynamics of the species and the fleets. 
When can MPAs reduce fishing mortality enough for bycatch species? And where can you go wrong?

It doesn’t make any sense to ask “do MPAs help bycatch”: the question is obviously it depends. To quickly make this case, we’ll assume that a species is in need of recovery. We then do the simple mako example to show that effects of spatial closure depend on the overlap of habitat and the assumption of fleet dynamics, and provide some empirical evidence that in particular effort redistribution is not unlikely.

Determining the effect of an MPA network on bycatch then will also depend on trying to get close to a tactical model. But, this creates a problem for blue water systems in particular. Can make the argument that given the centrality of relatively static habitat features in coastal MPAs “protect the key habitats” might not be a bad strategy from a conservation strategy. But in highly mobile pelagic systems there (may be) a lot more variables in play. Where should researchers direct their attention?

To  reduce the parameter space, we constrain ourselves to a system where 1) we know the species involved 2) we have some sense of the relative depletion and value of the different species and 3) some basic life history of the species. We then vary commonly unknown parameters, particularly preferred habitats, spatial catchability, movement rates etc, and pair these simulations with different MPA designs, and examine the conservation outcomes for bycatch species.

What i don’t like about this

There isn’t an “answer” to this: baring something like “changing the type of density dependence never had any effect greater than XX%”, for anything that on occasion has an impact the answer of “how much” will almost always depend on what else is going on.

So, it seems like there are two real angles of attack here

Pointing out a problem, which to make impactful would require making the case that there is a problem with current design strategies. I.e. “MPAs are often designed like this (simple dispersive models with constant habitat, or MARXSAN style things), and doing that does/does not create problems
A tactical evaluation. Picking a handful of high profile MPAs, parameterizing all the knownns as well as possible, iterating over unknowns, and see what the range of outcomes are. Cleaner, more impactful, but also opens it up (depending on the results) to a details fight about very specific modeling choices

Methods

Collect life history for key bycatch species and target species, cite Bradley et al. pulling together species distribution and threats etc. 
Introduce marlin
Parameterize marlin runs with “knowns” from life history of bycatch and target species
Highlight key “unknowns” that we will iterate over, e.g. fleet models, density dependent movement, timing of density dependence etc
Simulate effects of MPAs given reasonably realistic situations but iterating over key unknowns
Evaluate the range of outcomes and key drivers of MPA effects
I’m thinking we try and avoid a really specific case study but hard not to accidentally do one without making everything up….

Results
Range of likely changes in fishing-induced mortality for the species selected
Key drivers of uncertainty in those outcomes, focusing on things that switch states from reduction to increase in mortality
How common are conservation tradeoffs between bycatch species?
Fishery impacts

Discussion
What does this say about BWMPAs and bycatch?
Not going to get into design decisions here, depending how common tradeoffs are I think that will be the main focus
Highlight any species that frequently induced tradeoffs
Tee up future research
MPAs in context of other management
MPAs as risk reduction strategy
Etc.

### 2020-12-19 make-marlin-paper.R

@@ -1,16 +1,6 @@
+# Assesing Tradeoffs in the use of MPAs for Bycatch Reduction
 
 # setup -------------------------------------------------------------------
-# how sensitive is the outcome of an MPA to different uncertainties.
-#
-# i.e. put it somewhere, but suppose you don't know movement exactly
-#
-# shared, discrete, shared habitat
-# price information
-# ecology of the species
-# mpa modeling design
-# static modeling
-#
-#
 
 set.seed(42)
 library(tidyverse)
@@ -22,6 +12,14 @@ library(here)
 library(furrr)
 
 library(ggridges)
+
+library(googledrive)
+
+library(data.table)
+
+library(gamm4)
+
+library(sf)
 
 options(dplyr.summarise.inform = FALSE)
 
@@ -62,40 +60,271 @@ tune_type <- "explt"
 
 plan(multisession, workers = workers)
 
-# create core scenarios ---------------------------------------------------
 
-# use this space to develop the core building block of scenarios you're going to use
+# create distributions for species ----------------------------------------
+# create an spatial abundance distribution for each of the species in play
+# for now, pull in species from marlin team database
+# for now, requires authentication for emlab shared marlin data drive
 
+# this is broken for now
+# a = googledrive::drive_find("wcpfc_monthly.csv",n_max = 1)
+#
+# a = googledrive::drive_get("emlab/projects/current-projects/blue-prosperity-coalition/broader-research/mpa-bycatch/data/wcpfc_monthly.csv")
+#
+# drive_download(
+#   "wcpfc_monthly.csv",
+#   path = here("data","wcpfc_monthly.csv"),
+#   overwrite = TRUE
+# )
+#
+
+wcpfc_data <- read_csv(here("data","wcpfc_monthly.csv"))
+
+unique(wcpfc_data$species_commonname)
+
+top_bycatch <- wcpfc_data %>%
+  filter(catch > 0) %>%
+  group_by(species_commonname) %>%
+  count() %>%
+  arrange(desc(n)) %>%
+  ungroup() %>%
+  slice(1:4)
+
+wcpfc_data <- wcpfc_data %>%
+  filter(species_commonname %in% top_bycatch$species_commonname)
+
+
+bycatch_ll_cpue <- wcpfc_data %>%
[…]

### 2020-12-19 make-marlin-sensitivity-paper.R

@@ -1,0 +1,387 @@
+
+# setup -------------------------------------------------------------------
+# how sensitive is the outcome of an MPA to different uncertainties.
+#
+# i.e. put it somewhere, but suppose you don't know movement exactly
+#
+# shared, discrete, shared habitat
+# price information
+# ecology of the species
+# mpa modeling design
+# static modeling
+#
+#
+
+set.seed(42)
+library(tidyverse)
+
+library(marlin)
+
+library(here)
+
+library(furrr)
+
+library(ggridges)
+
+options(dplyr.summarise.inform = FALSE)
+
+
+foos <- list.files(here("R"))
+
+walk(foos, ~ source(here("R", .x)))
+
+results_name <- "v0.5"
+
+results_path <- here("results", results_name)
+
+if (!dir.exists(results_path)){
+  dir.create(results_path, recursive = TRUE)
+
+  dir.create(file.path(results_path,"sims"))
+}
+
+draws <- 300
+
+resolution <- 20 # resolution is in squared patches, so 20 implies a 20X20 system, i.e. 400 patches
+
+years <- 40
+
+seasons <- 1
+
+time_step <- 1 / seasons
+
+workers <- 3
+
+steps <- years * seasons
+
+time_steps <- seq(0,years - 1, by = time_step)
+
+theme_set(marlin::theme_marlin())
+
+tune_type <- "explt"
+
+plan(multisession, workers = workers)
+
+# create core scenarios ---------------------------------------------------
+
+# use this space to develop the core building block of scenarios you're going to use
+
+fauna_frame <- tibble(
+  scientific_name =  c(
+    "Thunnus obesus",
+    "Katsuwonus pelamis",
+    "Thunnus albacares",
+    "Kajikia audax",
+    "Carcharhinus longimanus"
+  ),
+  xid = list(1:draws)) %>%
+  unnest(cols = xid)
+
+
+hab1 <- expand_grid(x = 1:resolution, y = 1:resolution) %>%
+  mutate(habitat =  .2 * x) %>%
+  pivot_wider(names_from = y, values_from = habitat) %>%
+  select(-x) %>%
+  as.matrix()
+
+
+hab2 <- expand_grid(x = 1:resolution, y = 1:resolution) %>%
+  mutate(habitat =  dnorm(x, 9,5)) %>%
+  pivot_wider(names_from = y, values_from = habitat) %>%
+  select(-x) %>%
+  as.matrix()
+
+
+generate_traits <- function(scientific_name){
+
[…]

### 2021-02-26 README.Rmd

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Something about MPAs and Bycatch

Repo for manuscript exploring the ways in which habitat overlap, fleet spillover, and MPA design strategies can interact to affect conservation and fishery outcomes of MPAs, focused on WCPO-ish system.

### 2021-03-26 README.Rmd
passing to squid

@@
 Repo for manuscript exploring the ways in which habitat overlap, fleet spillover, and MPA design strategies can interact to affect conservation and fishery outcomes of MPAs, focused on WCPO-ish system.
+
+starginw thi hastings botsford weakstock species etc. can you hedge around this or not?Build intuition around BWMPAs for weakstocks
+
+Covariance matrix across species: perfectly correlated to totally uncorrelated and everything in between
+
+Clean up by setting Hastings and Botsford as a perfect template for this. PNAS? Isolate the life history
+
+It is clear that if species are mutually exclusive, there is no potential for overlap protection... but where does it fall when overlap is something in between
+
+Hmmm you could use virtual species.. or I guess that's even too complicated
+
+What if you just picked an "environmental" variable, and generated a random matrix
+
+For each species, you then assign a beta to that variable, where when they are all the same the covariance across the species is high.
+When they are all near zero there is no correlation, when they are opposite signs they are negatively correlated?
+
+Suppose you had two variables actually.
+One specifies kind of the "default" distributino for all species.
+The other is an environmental variable, and species can have coefficients that on average range from all 1, evenly split 1 and -1, and near zero
+
+Ah ok what about this.
+YOu define a set of benchmark species, where those benchmark species are defined by their relationship to a habitat.
+FOr the remainder of the species, their habitat is on average with noise on a gradient from -1 to 1, where one means they have the same as the benchmark, 0 means they are uncorrlated with the benchmark species, and -1 they are negatively correlated.
+
+```{r}
+
+library(marlin)
+library(tidyverse)
+
+theme_set(marlin::theme_marlin())
+
+resolution <- 20 # resolution is in squared patches, so 20 implies a 20X20 system, i.e. 400 patches
+
+seasons <- 1
+
+years <- 50
+
+tune_type <- "depletion"
+
+steps <- years * seasons
+
+# for now make up some habitat
+
+base_hab <- expand_grid(x = 1:resolution, y = 1:resolution) %>%
+  mutate(habitat =  .05 * x * y) %>% 
+  mutate(habitat = as.numeric(scale(habitat)))
+
+hab_scenes <- expand_grid(scene = c("friends", "enemies","frenemies"), species = fruit[1:10]) %>% 
+  group_by(scene) %>% 
+  nest()
+
+generate_hab <- function(species, homogifier = 1,scene,base_hab,sigma_hab = 1, beta_hab = 1){
+
+  sigma_hab = .2
+
+  sigma_block <- 0.1
+
+  homogifier <- 0
+
+  species <- hab_scenes$data[[1]]
+
+  block_a <- species$species[1:5]
+
+  block_b <- species$species[6:10]
+
+  block_a_betas <- rnorm(length(block_a),1,sigma_block)
+
+  block_b_betas <- rnorm(length(block_b), homogifier, sigma_block)
+
+  species$betas <- c(block_a_betas, block_b_betas)
+
+  species$habitat <- vector(mode = "list", length = nrow(species))
+
+  for (i in 1:nrow(species)){
+
+    base_hab$realized_hab <- rnorm(nrow(base_hab), base_hab$habitat * species$betas[i], sigma_hab)
+
+    species$habitat[[i]] <- base_hab
+
+  }
+
+  (species$habitat[[1]] %>% 
+    ggplot(aes(x,y,fill = realized_hab))+ 
+    geom_tile() + 
+    scale_fill_viridis_c()) + 
+      (species$habitat[[10]] %>% 
+    ggplot(aes(x,y,fill = realized_hab))+ 
+    geom_tile() + 
+    scale_fill_viridis_c())
+
+}
+
+```

### 2021-03-26 documents/marlin-paper.Rmd
passing to squid

@@
 ```
+
+what dooes the overlap tradeoff curve look like?
+
+Suppose you close off 30% of an area in an MPA
+
+Then plot conservation outcome on y-axis, degree of overlap between target and bycatch species on the x axis.
+When overlap is one, conservation benefit should be maxed.
+Then overlap is perfeclty anticorrelated, if effort concentrates, conservation should be at some negative minimum.
+If concentration doesn't happen, then conservation effect should be zero.
+If overlap is uncorrelated...
+unclear?
+
+If there is perfect overlap, then improvement in coservation outcome should be a straight positive function of area protected. If there is no overlap, then
 

### 2021-03-26 make-marlin-paper.R

@@ -2,6 +2,14 @@
 
 # setup -------------------------------------------------------------------
 
+# idea: define a centroid of habitat and a CV around that for each species and use that to tune the scenarios..
+# like manually creats random centroids that avoid each other... then increase the CV, such that there really is no key habitat when they... no I don't like that
+#
+# what if you randomly sample the centroid for each species, where when the cv of the sampling is low, most species have the same centroid, and as that icnreases you get more and more different centroids for each species.... that's a better idea.. centroid by gindinf the index of center cell as a default starting position
+#
+# iterate over cv of the centroid,
+# and then cv around the centroid
+
 set.seed(42)
 library(tidyverse)
 
@@ -21,6 +29,8 @@ library(gamm4)
 
 library(sf)
 
+library(mvtnorm)
+
 options(dplyr.summarise.inform = FALSE)
 
 
@@ -37,6 +47,8 @@ if (!dir.exists(results_path)){
 
   dir.create(file.path(results_path,"sims"))
 }
+
+run_experiments <- TRUE
 
 safety_stop <- FALSE
 
@@ -339,72 +351,235 @@ fauna_frame <- tibble(
 # pull from marlin_inputs, and randomizing others. That way you can easily
 # stick one "one" result, or generate a bunch of iterations varying in key unknowns
 
+# so you need to break the experiment in two. First you generate blocks of critters, where that can be combinations of
+# habitat heterogeneity, larval / adult habitat overlap, seasonal movement etc. Create that
+# Then, join that to a frame of MPA simulations, where yousimulate MPA sizes with different prioritizations
 
-create_critters <- function(sciname, habitat,seasons = 1, marlin_inputs){
 
-  # sciname <- marlin_inputs$scientific_name[[1]]
+# for now let's leave the life history stuff out of it... and then once you have a sense of timing for this decide on how many iterations to do of each
[…]

### 2021-05-05 documents/marlin-paper.Rmd
converting to makefile, major restructure, back to main

--- 2020-12-19 documents/marlin-paper.Rmd
@@
 title: "The Role of MPAs in Bycatch Reduction"
-author: "Dan Ovando"
-date: "12/11/2020"
+author:
+    - Daniel Ovando
+    - Darcy Bradley
+    - Echelle Burns
+    - Lennon Thomas
+date: "`r Sys.Date()`"
 output: 
@@
   bookdown::html_document2: default
+editor_options: 
+  markdown: 
+    wrap: 72
+bibliography: references.bib
+params:
+  run_name: ["v0.5"]
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE)
-library(tidyverse)
-library(here)
-library(bookdown)
+knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, include = FALSE)
+
+source(here::here("scripts", "00_setup.R"))
+
+theme_set(marlin::theme_marlin())
+
 ```
 
-Marine protected areas (MPAs) are a growing part of the marine conservation toolbox, with 7.4% of the world’s oceans currently protected and, and bodies such as the IUCN calling for 30% protection by 2030.
-Many of the largest of these MPAs are being implemented in the form of “blue-water” MPAs largely covering pelagic ecosystems, characterized by highly mobile species and fishing fleets using a diverse mix of static and dynamic ocean features.
-These systems operate very differently from the coastal areas where modern MPA design practices have been developed.
-Pelagic systems are also characterized by combinations of relatively abundant species targeted by fishing, such as most tuna stocks, and highly threatened species caught as incidental “by-catch” in targeted fishing operations, such as oceanic whitetip sharks.
-Blue-water MPAs are increasingly being proposed specifically to benefit these by-catch populations while minimizing losses or where possible providing benefits to fisheries.
+```{r load-results}
+experiments <- read_rds(file = file.path(results_path, "experiements.rds"))
+
+critter_lookup <- read_rds(file = file.path(results_path, 'critter-lookup.rds'))
+
+rough_habitat <- read_rds(file = file.path(results_path, "rough-habitat.rds"))
+
+a = critter_lookup %>%
+  left_join(rough_habitat %>% select(species_commonname, bycatch) %>% unique)
+
+scenes <- experiments %>%
+  select(xid, data) %>%
+  unnest(cols = data) %>%
+  select(xid:seasonal_movement) %>%
+  unique()
+
+mpa_results <-  experiments %>%
+  select(xid, experiment) %>%
+  unnest(cols = experiment) %>%
+  select(xid,target_fauna, mpa_size, results) %>%
+  unnest(cols = results) %>%
+  mutate(target_fauna = map_chr(target_fauna, ~paste(.x, collapse = ","))) %>%
+  left_join(a, by = c("critter" = "species_sciname")) %>%
+  left_join(scenes, by = "xid") %>%
+  mutate(bycatch = ifelse(bycatch == TRUE, "Bycatch Species","Target Species"))
+
+mpa_strategy <- c(
+  `carcharhinus longimanus` = "MPAs for OWT Shark",
+  `prionace glauca,carcharhinus longimanus,isurus oxyrinchus,carcharhinus falciformis` = "MPAs for All Bycatch"
+)
+
+ssb_traj_plot <- mpa_results %>%
+  filter(variable == "ssb") %>%
+  ggplot(aes(
+    mpa_size,
+    percent_change_ssb0,
+    color = species_commonname,
+    group = interaction(xid,species_commonname,bycatch)
+  )) +
+  geom_hline(aes(yintercept = 0)) +
+  geom_line(size = 1.5, alpha = 0.25) +
+  facet_grid( bycatch~ target_fauna, labeller = labeller(target_fauna = mpa_strategy), scales = "free_y") +
+  scale_x_continuous(name = "MPA Size", labels = scales::percent) +
+  scale_y_continuous(name = "% of SSB0 Change", labels = scales::percent) +
+  scale_color_discrete(name = "Species") +
+  scale_linetype(name = '') +
+  theme(axis.text = element_text(size = 8),
+        axis.title = element_text(size = 14),
+        legend.text = element_text(size = 12))
+
+mpa_strategy <- c(
+  `carcharhinus longimanus` = "MPAs for OWT Shark",
+  `prionace glauca,carcharhinus longimanus,isurus oxyrinchus,carcharhinus falciformis` = "MPAs for All Bycatch"
+)
+
+catch_traj_plot <- mpa_results %>%
+  filter(variable == "c") %>%
+  ggplot(aes(
+    mpa_size,
+    percent_change,
+    color = species_commonname,
+    group = interaction(xid,species_commonname,bycatch)
+  )) +
+  geom_hline(aes(yintercept = 0)) +
+  geom_line(size = 1, alpha = 0.25) +
+  facet_wrap(bycatch ~ target_fauna, labeller = labeller(target_fauna = mpa_strategy)) +
+  scale_x_continuous(name = "MPA Size", labels = scales::percent) +
+  scale_y_continuous(name = "% Change in Catch", labels = scales::percent) +
+  scale_color_discrete(name = "Species") +
+  scale_linetype(name = '') +
+  theme(axis.text = element_text(size = 8),
+        axis.title = element_text(size = 14),
+        legend.text = element_text(size = 12))
+
+mpa_strategy <- c("MPAs for OWT Shark" = "carcharhinus longimanus" ,
+                  "MPAs for All Bycatch" = "prionace glauca,carcharhinus longimanus,isurus oxyrinchus,carcharhinus falciformis")
+
+tradeoff_frontier_plot <- mpa_results %>%
+  mutate(target_fauna = as.factor(target_fauna)) %>%
+  mutate(target_fauna = fct_recode(target_fauna,!!!mpa_strategy)) %>%
+  select(mpa_size,
+         percent_change,
+         species_commonname,
+         target_fauna,
+         variable,
+         xid) %>%
+  filter(variable %in% c("c", "ssb")) %>%
+  pivot_wider(names_from = "variable", values_from = percent_change) %>%
+  ggplot(aes(ssb, c, shape = target_fauna, linetype = target_fauna)) +
+  geom_vline(aes(xintercept = 0)) +
+  geom_hline(aes(yintercept = 0)) +
+  # geom_line() +
+  geom_point(aes(color = mpa_size), size = 2) +
+  facet_wrap(~ species_commonname, scales = "free") +
+  scale_y_continuous(name = "% Change in Catch", labels = scales::percent) +
+  scale_x_continuous(name = "% of SSB0 Change", labels = scales::percent) +
+  theme(
+    axis.text = element_text(size = 8),
+    axis.title = element_text(size = 14),
+    legend.text = element_text(size = 12),
+    legend.position = "top",
+    legend.direction = "horizontal"
+  ) +
+  scale_colour_viridis_b(
+    name = "MPA Size",
+    labels = scales::percent,
+    guide = guide_colorbar(barwidth = unit(9, "lines"))
+  ) +
+  scale_shape(name = "") +
+  scale_linetype(name = "")
+
+a %>% knitr::kable() %>% kableExtra::kable_paper()
+
+# what is interesting
+#
+#
+# proportion stocks benefiting / losing as a functino of MPA size and spatial homogeneity
+#
+
+any_ssb_loss_plot <- mpa_results %>%
+  mutate(target_fauna = as.factor(target_fauna)) %>%
+  mutate(target_fauna = fct_recode(target_fauna,!!!mpa_strategy)) %>%
+  filter(variable == "ssb") %>%
+  group_by(mpa_size, sigma_centroid, seasonal_movement, target_fauna) %>%
+  summarise(percent_improve = mean(percent_change >= 0),
+            percent_lose = 1 - percent_improve) %>%
+  ggplot(aes(mpa_size, percent_lose, color = factor(sigma_centroid))) +
+  geom_line(size = 1.5) +
+  geom_hline(aes(yintercept = 0)) +
+  scale_color_viridis_d(name = "Heterogeneity in Adult Habitat") +
+  scale_x_continuous(labels = scales::percent, name = 'Area in MPA') +
+  scale_y_continuous(labels = scales::percent, name = "Sims where MPA reduces SSB") +
+  facet_grid(target_fauna~seasonal_movement, labeller = label_both) +
+  theme(legend.position = "top")
+
+signif_ssb_loss_plot <- mpa_results %>%
+  mutate(target_fauna = as.factor(target_fauna)) %>%
+  mutate(target_fauna = fct_recode(target_fauna,!!!mpa_strategy)) %>%
+  filter(variable == "ssb") %>%
+  group_by(mpa_size, sigma_centroid, seasonal_movement, target_fauna) %>%
+  summarise(percent_improve = mean(percent_change >= .1),
+            percent_lose = mean(percent_change <= -.1)) %>%
+  ggplot(aes(mpa_size, percent_lose, color = factor(sigma_centroid))) +
+  geom_line(size = 1.5) +
+  geom_hline(aes(yintercept = 0)) +
+  scale_color_viridis_d(name = "Heterogeneity in Adult Habitat") +
+  scale_x_continuous(labels = scales::percent, name = 'Area in MPA') +
+  scale_y_continuous(labels = scales::percent, name = "Sims where SSB Decreaes by more than 10%") +
+  facet_grid(target_fauna~seasonal_movement, labeller = label_both) +
+  theme(legend.position = "top")
+
+signif_ssb_gain_plot <- mpa_results %>%
+  mutate(target_fauna = as.factor(target_fauna)) %>%
+  mutate(target_fauna = fct_recode(target_fauna,!!!mpa_strategy)) %>%
+  filter(variable == "ssb") %>%
+  group_by(mpa_size, sigma_centroid, seasonal_movement, target_fauna) %>%
+  summarise(percent_improve = mean(percent_change >= .1),
+            percent_lose = mean(percent_change <= -.1)) %>%
+  ggplot(aes(mpa_size, percent_improve, color = factor(sigma_centroid))) +
+  geom_line(size = 1.5) +
+  geom_hline(aes(yintercept = 0)) +
+  scale_color_viridis_d(name = "Heterogeneity in Adult Habitat") +
+  scale_x_continuous(labels = scales::percent, name = 'Area in MPA') +
+  scale_y_continuous(labels = scales::percent, name = "Sims where MPAs increase by more htan 10%") +
+  facet_grid(target_fauna~seasonal_movement, labeller = label_both) +
+  theme(legend.position = "top")
+
+mean_ssb_traj_plot <- mpa_results %>%
+  mutate(target_fauna = as.factor(target_fauna)) %>%
+  mutate(target_fauna = fct_recode(target_fauna,!!!mpa_strategy)) %>%
+  filter(variable == "ssb") %>%
+  group_by(mpa_size, sigma_centroid, seasonal_movement, target_fauna) %>%
+  summarise(net_percent_change = weighted.mean(percent_change,ssb0)) %>%
+  ggplot(aes(mpa_size, net_percent_change, color = factor(sigma_centroid))) +
+  geom_line(size = 1.5) +
+  geom_hline(aes(yintercept = 0)) +
+  scale_color_viridis_d(name = "Heterogeneity in Adult Habitat") +
+  scale_x_continuous(labels = scales::percent, name = 'Area in MPA') +
+  scale_y_continuous(labels = scales::percent, name = "% Change SSB") +
+  facet_grid(target_fauna~seasonal_movement, labeller = label_both) +
+  theme(legend.position = "top")
+
+total_ssb_traj_plot <- mpa_results %>%
+  mutate(target_fauna = as.factor(target_fauna)) %>%
+  mutate(target_fauna = fct_recode(target_fauna,!!!mpa_strategy)) %>%
+  filter(variable == "ssb") %>%
+  group_by(mpa_size, sigma_centroid, seasonal_movement, target_fauna) %>%
+  summarise(net_change = sum(abs_change)) %>%
+  ggplot(aes(mpa_size, net_change, color = factor(sigma_centroid))) +
+  geom_line(size = 1.5) +
+  geom_hline(aes(yintercept = 0)) +
+  scale_color_viridis_d(name = "Heterogeneity in Adult Habitat") +
+  scale_x_continuous(labels = scales::percent, name = 'Area in MPA') +
+  scale_y_continuous(name = "Net Change in SSB") +
+  facet_grid(target_fauna~seasonal_movement, labeller = label_both) +
+  theme(legend.position = "top")
+
+signif_bycatch_ssb_loss_plot <- mpa_results %>%
+  filter(bycatch == "Bycatch Species") %>%
+  mutate(target_fauna = as.factor(target_fauna)) %>%
+  mutate(target_fauna = fct_recode(target_fauna,!!!mpa_strategy)) %>%
+  filter(variable == "ssb") %>%
+  group_by(mpa_size, sigma_centroid, seasonal_movement, target_fauna) %>%
+  summarise(percent_improve = mean(percent_change >= 0),
+            percent_lose = 1 - percent_improve) %>%
+  ggplot(aes(mpa_size, percent_lose, color = factor(sigma_centroid))) +
+  geom_line(size = 1.5) +
+  geom_hline(aes(yintercept = 0)) +
+  scale_color_viridis_d(name = "Heterogeneity in Adult Habitat") +
+  scale_x_continuous(labels = scales::percent, name = 'Area in MPA') +
+  scale_y_continuous(labels = scales::percent, name = "Sims where MPA reduces SSB") +
+  facet_grid(target_fauna~seasonal_movement, labeller = label_both) +
+  theme(legend.position = "top")
+
+catch_loss_plot <- mpa_results %>%
+  mutate(target_fauna = as.factor(target_fauna)) %>%
+  mutate(target_fauna = fct_recode(target_fauna,!!!mpa_strategy)) %>%
+  filter(variable == "c") %>%
+  group_by(mpa_size, sigma_centroid, seasonal_movement, target_fauna) %>%
+  summarise(percent_improve = mean(percent_change >= 0),
+            percent_lose = 1 - percent_improve) %>%
+  ggplot(aes(mpa_size, percent_lose, color = factor(sigma_centroid))) +
+  geom_line(size = 1.5) +
+  geom_hline(aes(yintercept = 0)) +
+  scale_color_viridis_d(name = "Heterogeneity in Adult Habitat") +
+  scale_x_continuous(labels = scales::percent, name = 'Area in MPA') +
+  scale_y_continuous(labels = scales::percent, name = "Sims where MPA reduces Catch") +
+  facet_grid(target_fauna~seasonal_movement, labeller = label_both) +
+  theme(legend.position = "top")
+
+signif_catch_loss_plot <- mpa_results %>%
+  mutate(target_fauna = as.factor(target_fauna)) %>%
+  mutate(target_fauna = fct_recode(target_fauna,!!!mpa_strategy)) %>%
+  filter(variable == "c") %>%
+  group_by(mpa_size, sigma_centroid, seasonal_movement, target_fauna) %>%
+  summarise(percent_improve = mean(percent_change >= .1),
+            percent_lose = mean(percent_change <= -.1)) %>%
+  ggplot(aes(mpa_size, percent_lose, color = factor(sigma_centroid))) +
+  geom_line(size = 1.5) +
+  geom_hline(aes(yintercept = 0)) +
+  scale_color_viridis_d(name = "Heterogeneity in Adult Habitat") +
+  scale_x_continuous(labels = scales::percent, name = 'Area in MPA') +
+  scale_y_continuous(labels = scales::percent, name = "Sims where MPA reduces Catch by more than 10%") +
+  facet_grid(target_fauna~seasonal_movement, labeller = label_both) +
+  theme(legend.position = "top")
+
+signif_catch_gain_plot <- mpa_results %>%
+  mutate(target_fauna = as.factor(target_fauna)) %>%
+  mutate(target_fauna = fct_recode(target_fauna,!!!mpa_strategy)) %>%
+  filter(variable == "c") %>%
+  group_by(mpa_size, sigma_centroid, seasonal_movement, target_fauna) %>%
+  summarise(percent_improve = mean(percent_change > .1),
+            percent_lose = mean(percent_change < -.1)) %>%
+  ggplot(aes(mpa_size, percent_improve, color = factor(sigma_centroid))) +
+  geom_line(size = 1.5) +
+  geom_hline(aes(yintercept = 0)) +
+  scale_color_viridis_d(name = "Heterogeneity in Adult Habitat") +
+  scale_x_continuous(labels = scales::percent, name = 'Area in MPA') +
+  scale_y_continuous(labels = scales::percent, name = "Sims where MPA incease Catch by >10%") +
+  facet_grid(target_fauna~seasonal_movement, labeller = label_both) +
+  theme(legend.position = "top")
+
+```
+
+# Introduction
+
+Marine protected areas (MPAs) are a growing part of the marine
+conservation toolbox, with 7.4% of the world's oceans currently
+protected and, and bodies such as the IUCN calling for 30% protection by
+2030. Many of the largest of these MPAs are being implemented in the
+form of "blue-water" MPAs largely covering pelagic ecosystems,
+characterized by highly mobile species and fishing fleets using a
+diverse mix of static and dynamic ocean features. These systems operate
+very differently from the coastal areas where modern MPA design
+practices have been developed. Pelagic systems are also characterized by
+combinations of relatively abundant species targeted by fishing, such as
+most tuna stocks, and highly threatened species caught as incidental
+"by-catch" in targeted fishing operations, such as oceanic whitetip
+sharks. Blue-water MPAs are increasingly being proposed specifically to
+benefit these by-catch populations while minimizing losses or where
+possible providing benefits to fisheries.
+
+While the effect of MPAs on fisheries yields is a continuously debated
+topic, the conservation impacts of MPAs have received less attention of
+late, due in part perhaps to the perceived obvious nature of the
+problem: if you prevent fishing of a species within an area, there will
+be more of that species in that area, and so the conservation benefits
+of MPAs will increase with MPA size. \@hastings2017 clearly illustrates
+this idea. Using a simple two-species two-patch analytical model, they
+derive conditions under which MPAs can solve the "weak-stock" problem in
+which a less productive species is incidentally exploited by a fleet
+targeting a more valuable and productive stock.
+
+However, as @hastings2017 point out, their results depend on a very
+specific set of assumptions, namely that 1) adults are perfectly
+sedentary 2) the target and bycatch species overlap perfectly in terms
+of habitat xx....The authors suggest that the impact of these
+assumptions should be tested in scenarios under which they are unlikely
+to apply. We build on \@hastings2017 ..xx sort of... theirs is about
+yield and we don't talk about that much yet... though we can xx to
+consider the role of MPAs in protecting bycatch species in pelagic
+ecosystems.
+
+The core models underpinning MPA science have largely been based on
+stylized versions of near-shore reef-style xx what to call that, like
+coral / rocky / kelp forest, etc xx systems, where multiple species
+often share common habitat, and adults have reasonable site fidelity
+while larvae are broadly dispersed. Pelagic ecosystems have vastly
+different dynamics than coastal seas [@game2009; @lubchenco2003],
+characterized by highly migratory species specialized to take advantage
+of highly dynamic ocean features such as eddies currents while also
+utilizing static features such as seamounts. Pelagic ecosystems also
+support both coastal domestic fishing industries along with highly
+mobile international high-seas fishing operations. Given these dynamics,
+it remains unclear how many of the lessons from coastal MPA science are
+applicable to pelagic ecosystems [\@ovando2021XX].
 
 Pelagic ecosystems present both opportunities and challenges for MPAs.
-The highly productive and mobile nature of targeted fisheries in these systems may mean that highly targeted MPAs can be designed that provide effective conservation at minimal fishery costs.
-Alternatively, if individual by-catch species do not share common habitat preferences or fishery vulnerability, protection in one area targeted to benefit one species may negatively impact another through effort displacement.
-Here, we use a case study motivated by the upcoming expansion of blue-water MPAs throughout the Federated States of Micronesia to assess the sensitivity of by-catch populations and targeted fisheries to MPA design and key life-history variables.
+The highly productive and mobile nature of targeted fisheries in these
+systems may mean that highly targeted MPAs can be designed that provide
+effective conservation at minimal fishery costs. Strategic protection of
+key aggregation locations may enhance depleted pelagic fisheries [xx
+that Indian ocean study I can't remember right now]. xx big question
+here: how focused do we want this paper to be? is it just about
+conservation tradeoffs or lack thereof with bycatch species, or is about
+the net conservation and fishery effects of BWMPAs. I'm inclined the
+former to tee-up the later xx. Alternatively, if individual bycatch
+species do not share common habitat preferences or fishery
+vulnerability, protection in one area targeted to benefit one species
+may negatively impact another through effort displacement [e.g.
+@abbott2012].
 
-The outcome of MPAs for fishing and conservation are highly sensitive to a wide range of economic and ecological factors.
-In order to constrain the possible parameter space, we tune a simulation model based around best available knowledge of the species and fleets operating in the areas of the western central pacific ocean (WCPO) while allowing key unknowns to vary, in particular the nature of adult and larval movement, the degree of habitat overlap between different species, the economic incentives facing fishing fleets, and the design strategy used in MPA placement XX.
-Most conventional MPA design models are either static, single species, or both.
-We present a dynamic, multi-species simulation model that allows us to cohesively assess the impact of MPAs on a suite of by-catch and targeted species collectively affected by multiple fishing fleets.
-Our goal here is not to provide tactical recommendations on the design of blue-water MPAs throughout the WCPO, but rather to provide a strategic evaluation of the sensitivity of by-catch populations and targeted fisheries to key unknowns and design choices in MPA design.
+Here, we use a case study motivated by the global expansion of
+blue-water MPAs to assess the sensitivity of bycatch populations and
+targeted fisheries to MPA design and key life-history variables... and
+what else?
 
-We show 
-  - something about how when species are homogeneous in space, conservation gains are basically guaranteed except for constant catch scenearios, which need to be considered in tuna-type scenarios
-  - When we introduce heterogeneity in stock structure roughly reflective of distributions in the WCPO we do / do not begin to see tradeoffs in conservation outcomes among bycatch species
-    - quantify if there is a concave function to this
-  - Specific design paradigms are / are not more / less resilient to this issue
-  - Compare tradeoffs between MPA designs that minimze internal bycatch tradeoffs (there have to be none if MPA = 100%) and fishery outcomes
+-   Case study of WCPO?
 
-We show that MPAs increased biomass of by-catch species in XX% of simulations, and decreased them in XX% of simulations.
-On average a 1% increase in by-catch species resulted in a XX% change in fishery catches… Outcomes were most sensitive to XX and XX, and least sensitive to XX and XX.
-We do not suggest that these same results would hold for other systems, but rather demonstrate the importance / resilience of MPAs to key uncertain variables.
-In particular for pelagic ecosystems, our work demonstrates the importance of considering the collective impacts of space, vulnerability, life history, and fleet dynamics in MPA design for conservation.
+-   Comparison to input / output controls
 
-## Figures
+-   Comparison to DOM?
 
-Are there or are there not tradeoffs in conservation and fisheries yields
+Our goal here is not to provide tactical recommendations on the design
+of blue-water MPAs, but rather to provide a strategic evaluation of the
+sensitivity of bycatch populations and targeted fisheries to key
+unknowns and design choices in MPA planning.
+
+<!-- The outcome of MPAs for fishing and conservation are highly sensitive to -->
+
+<!-- a wide range of economic and ecological factors. In order to constrain -->
+
+<!-- the possible parameter space, we tune a simulation model based around -->
+
+<!-- best available knowledge of the species and fleets operating in the -->
+
+<!-- areas of the western central pacific ocean (WCPO) while allowing key -->
+
+<!-- unknowns to vary, in particular the nature of adult and larval movement, -->
+
+<!-- the degree of habitat overlap between different species, the economic -->
+
+<!-- incentives facing fishing fleets, and the design strategy used in MPA -->
+
+<!-- placement XX. Most conventional MPA design models are either static, -->
+
+<!-- single species, or both. We present a dynamic, multi-species simulation -->
+
+<!-- model that allows us to cohesively assess the impact of MPAs on a suite -->
+
+<!-- of by-catch and targeted species collectively affected by multiple -->
+
+<!-- fishing fleets.  -->
+
+<!-- We show - something about how when species are homogeneous in space, -->
+
+<!-- conservation gains are basically guaranteed except for constant catch -->
+
+<!-- scenearios, which need to be considered in tuna-type scenarios - When we -->
+
+<!-- introduce heterogeneity in stock structure roughly reflective of -->
+
+<!-- distributions in the WCPO we do / do not begin to see tradeoffs in -->
+
+<!-- conservation outcomes among bycatch species - quantify if there is a -->
+
+<!-- concave function to this - Specific design paradigms are / are not more -->
+
+<!-- / less resilient to this issue - Compare tradeoffs between MPA designs -->
+
+<!-- that minimze internal bycatch tradeoffs (there have to be none if MPA = -->
+
+<!-- 100%) and fishery outcomes -->
+
+<!-- We show that MPAs increased biomass of by-catch species in XX% of -->
+
+<!-- simulations, and decreased them in XX% of simulations. On average a 1% -->
+
+<!-- increase in by-catch species resulted in a XX% change in fishery -->
+
+<!-- catches... Outcomes were most sensitive to XX and XX, and least -->
+
+<!-- sensitive to XX and XX. We do not suggest that these same results would -->
+
+<!-- hold for other systems, but rather demonstrate the importance / -->
+
+<!-- resilience of MPAs to key uncertain variables. In particular for pelagic -->
+
+<!-- ecosystems, our work demonstrates the importance of considering the -->
+
+<!-- collective impacts of space, vulnerability, life history, and fleet -->
+
+<!-- dynamics in MPA design for conservation. -->
+
+# Methods
+
+@hastings2017 employed a simply model to derive conditions under which
+MPAs would benefit bycatch species. However they also pointed out that
+even small increases in model complexity quickly make analytical
+solutions to the effects of MPAs intractable. As such, we based our
+analysis on a spatially explicit, multi-species, multi-fleet,
+bio-economic model we call
+[`marlin`](https://github.com/DanOvando/marlin). Models of MPAs span a
+wide range of complexity, from simple analytical tools to end-to-end
+ecosystem models [@fulton2015]. `marlin` follows a "models of
+intermediate complexity" (MICE) philosophy [@plagányi2014] targeted
+specifically around the role of technical interactions across species.
+Many MPA models model the impacts of MPAs on a single-species level [@ovando2021XX].
+This can be useful for for example examining the ways in which species-specific life history traits interact with fleet dynamics to influence MPA outcomes.
+However, bycatch problems are driven by technical interactions, in which fishing effort for a target species incidentally captures another species xx how do we want to talk about this in here?
+Like usually incidentally, but clearly sometimes people try and catch sharks too xx.
+`marlin` accounts for this by allowing the user to simulate an arbitrary number of fleets capable of catching an arbitrary number of species in a spatial area.
+Different fleets can place different economic value on different species, while also having species-specific selectivity functions.
+This allows us to simulate bycatch dynamics of MPAs in pelagic ecosystems.
+
+## `marlin`
+
+Lots of equations.
+
+## Simulation Experiments
+
+Given the complexity of the operating model used here, a complete sensitivity analysis across all parameters is not practical.
+Instead, we focus on the role of XX key attributes of pelagic ecosystems: the degree of spatial overlap between populations, the presence of seasonal migrations, and the potential for ontogenetic shifts in habitat use.
+
+### Habitat Overlap
+
+We simulate habitat overlap through two parameters: the standard deviation of the habitat centroid for each species $\sigma_c$, and the the dispersion of the species around its centroid $\sigma_d$.
+When $\sigma_c$ is low, all species share a common habitat.
+As $\sigma_c$ inscreases, the variance in the centroid of each of the simulated species increases.
+
+We simulate seasonal movement by breaking the year into two seasons. We then set the adult habitat in season two as the negative inverse of the habitat in season one
+
+We simulate ontogenetic movement by allowing for separate habitat layers for recruits and adults.
+We then simulate a scenario where recruitment habitat is the negative inverse of adult habitat, such that recruits primarily arive in one area of the system, and then move to another area as they grow.
+
+etc.
+
+## comparison to...
+
+For each scenario, could then compare that MPA to the amount of yield
+lost / bycatch gained by constraining effort / catch...
+
+# Results
+
+## Conservation Outcomes
+
+```{r traj-plot, include=TRUE, fig.cap="Trajectoris of percent change in spawning stock biomass as a function of MPA size. Color indicates separate species, each line is a different simulation."}
+
+ssb_traj_plot + 
+  guides(color = FALSE)
+
+```
+```{r ssb-plot,fig.cap="Percent of simulations that resulted in any decrease in spawning stock biomass as a function of seasonal movement, habitat heterogeneity, and MPA design.", include=TRUE}
+any_ssb_loss_plot
+```
+
+```{r signif-ssb-plot, fig.cap= "Percent of simulations that resulted in decreases of 10% or more in spawning stock biomass as a function of seasonal movement, habitat heterogeneity, and MPA design.", include=TRUE}
+signif_ssb_loss_plot
+```
+
+## Yield Outcomes
+
+```{r catch-plot,fig.cap="Percent of simulations that resulted in any decrease in catch as a function of seasonal movement, habitat heterogeneity, and MPA design.", include=TRUE}
+catch_loss_plot
+```
+
+```{r signif-catch-plot, fig.cap= "Percent of simulations that resulted in decreases of 10% or more in catch as a function of seasonal movement, habitat heterogeneity, and MPA design.", include=TRUE}
+signif_catch_loss_plot
+```
+
+## Value of Information
+
+I think this would be one of the easiest and interesting: how good does you habitat knowledge have to be to minimize tradeoffs?
+
+## Tradeoffs
+
+```{r tradeoff-plot, fig.cap= "Tradeoff frontier of catch and spawning stock biomass as a function of MPA size and MPA design. Each point represents a different simulation", include=TRUE}
+tradeoff_frontier_plot
+```
+
+## Sensitivy to...
+
+Summary (hopefully statistical) of sensitivity of outcome to different parameters? But hard to know if there's really the sample size for that
+
+## Case Study Of...
+
+# Discussion
+
+XX stress a key difference: may other papers on this are static, in that
+they just some up displacement on different layers, rather than
+addressing dynamic respones over tim xx
+
+We show....
+  - sensitivity of both conservation and fishery outcomes to pelagic realism and design choices
+  - compared to targeted / dynamic intervention... 
+  - value of information would be great...
+
+**what is novel????**
+
+Contrast to coastal MPA theoretical work
+  - @hastings2017
+  - @ovando2021
+  - @white2010
+  - not quite sure what to do about points raised in @fulton2011a
+    - to what extent has this been done...
+
+The paragraph from @fulton2011a that broke me...
+
+"
+Spatial management was insufficient as the dominant management lever because it proves quite difficult to strike the balance between protecting areas that are of a size sufficient to produce any conservation benefit and providing sufficient fishing grounds to sustain an econom- ically viable industry (which is not possible in many fisheries if closures on the order of 70–80% or more are imposed, Fulton et al.
+2007).
+More- over, Savina etal.
+(2008) found that effort displaced by closures potentially undermines any benefits coming from the zoning.
+For instance, strong improvements in diversity (by nearly 20%) under heavy fishing pressure occurred if effort outside a marine protected area (MPA) was controlled.
+However, without effort control, there was a significant (>20%) drop in overall diversity.
+Fulton et al.
+(2007), Savina et al.
+(2008), Smith et al.
+(2010) and unpublished work associated with Horne et al.
+(2010) all found that while spatial management appears to be an effective means of satisfying conservation objectives con- cerning habitat and restoring age-structures of less mobile species, benefits from closures for individual species depend on its trophic position, degree of mobility, geographical distribution and extent of ontogenetic shifts in habitat use and depth (highly mobile or prey species do not benefit in the way that more site attached or predatory species do)."
+
+Lots of studies have shown that lots of things matter, can we provide guidance on which things are priorities?
+
+Where does this fit in context of @smith2021?
+One important difference is that they point out differences in effectiveness, where effectiveness is still almost always positive, since they just measure leatherbacks.
+So theri fig 5, there are a few cases where leatherback catches actually go up, but it's rare.
+So, extending to multiple species and keeping track off all of them shows it's even harder
+
+How does this contrast with the the large and growing dynamic ocean management lit?
+  - By definition, with perfect information dynamic $\geq$ static spatial controls
+  - DOM literature is mostly static, tracks displacemento onto static abundance / risk layers
+  - Doesn't allow for dynamic impacts, for example fleet adjustments based on updated economic surface, or short-term costs for long-term benefits (can overstate costs?)
+  - Dynamic model allows for delta between dynamic policies, e.g. Fmsy vs DOM
+    - analytical models a la hastings have value but do / don't stand up to realism
+
+Presentation of the model as a MICE model for MSP
+  - reasonable parameterizable
+  - flexible
+  - fast
+
+Focus of debate has been on fisheries and yields.
+As MPAs move to pelagic systems though, we show that conservation outcomes also depend on correctly identifying ecological state / design decisions etc.
+
+# References
+
+::: {#refs}
+:::
+
+## Figure Ideas
+
+Are there or are there not tradeoffs in conservation and fisheries
+yields
 
@@
 
-Interesting considerations we do biodiversity MPAs we define biodiversity as system diversity and we put all the species in there, threatened bycatch species become the the focus and stated objective is ramp up the value of bycatch species .
-There might be instances where just the biodiversity base actually gets you a better outcome than .
+Interesting considerations we do biodiversity MPAs we define
+biodiversity as system diversity and we put all the species in there,
+threatened bycatch species become the the focus and stated objective is
+ramp up the value of bycatch species . There might be instances where
+just the biodiversity base actually gets you a better outcome than .
 There might be overlap and not.
@@
 
-We can explore that mismatch between is the issue q and is the issue of spatial habitat
+We can explore that mismatch between is the issue q and is the issue of
+spatial habitat
 
@@
 Interestingly, you know that if you closed 100% all benefit, so doing this along a gradient is interesting, are there tradeoffs at low sizes that disappear with higher?
-Actually you might expect a dome-shaped tradeoff, where there are no tradeoffs at low size since little displacement, largest tradeoffs in the middle where you maximize displacement but still haven’t maximized protection, and then no tradeoffs at the extreme where everything is protected.
+Actually you might expect a dome-shaped tradeoff, where there are no tradeoffs at low size since little displacement, largest tradeoffs in the middle where you maximize displacement but still haven't maximized protection, and then no tradeoffs at the extreme where everything is protected.
 If you have a small enough number of species, say less than 10, then can do color / shape by species, mpa size on x axis, change in biomass on y axis, and connected lines for design paradigms.
@@
 
-### Fig 5 + Figures showing a few select “what if the world were different”, i.e. how important is movement rate, DD form, etc.
+### Fig 5 + Figures showing a few select "what if the world were different", i.e. how important is movement rate, DD form, etc.
 
@@
 
-What are the considerations of protection you need to make good decisions.
-It’s not always obvious how to design things to get the outcome you want.
-If we could make the case that something like closing the high seas is insufficient conservation management without conservation management inside EEZs / what is the
+What are the considerations of protection you need to make good
+decisions. It's not always obvious how to design things to get the
+outcome you want. If we could make the case that something like closing
+the high seas is insufficient conservation management without
+conservation management inside EEZs / what is the
 
-Need to do an optimization here, calculate the marginal value of added cells.
+Need to do an optimization here, calculate the marginal value of added
+cells.
 
-What’s this about?
-Pointing out a problem, which to make impactful would require making the case that there is a problem with current design strategies.
-e.g..
-“MPAs are often designed like this (simple dispersive models with constant habitat, or MARXSAN style things), and doing that does/does not create problems
-A tactical evaluation.
-Picking a handful of high profile MPAs, parameterizing all the knowns as well as possible, iterating over unknowns, and see what the range of outcomes are.
-Cleaner, more impactful, but also opens it up (depending on the results) to a details fight about very specific modeling choices
-How plausible are losses?
-Maybe we drill down on this a bit.
-From a purely bycatch perspective, the usual questions of fisheries tradeoffs aren’t on the table (but can add to the discussion considerations of how much target fishing you give up).
-Given that then, while it’s possible that “targeted” interventions would work better (ignoring costs/enforcement, etc), there’s really no reason not to do MPAs.
-Unless, you actually end up concentrating fishing pressure outside on populations.
-So, we consider how plausible is it for that to really happen.
-Issue here is that seems likely to be pretty hand-wavey.
-Most likely that the vast majority of simulations will show positive impacts, but have no way of assigning probabilities to any one state of nature.
+What's this about? Pointing out a problem, which to make impactful would
+require making the case that there is a problem with current design
+strategies. e.g.. "MPAs are often designed like this (simple dispersive
+models with constant habitat, or MARXSAN style things), and doing that
+does/does not create problems A tactical evaluation. Picking a handful
+of high profile MPAs, parameterizing all the knowns as well as possible,
+iterating over unknowns, and see what the range of outcomes are.
+Cleaner, more impactful, but also opens it up (depending on the results)
+to a details fight about very specific modeling choices How plausible
+are losses? Maybe we drill down on this a bit. From a purely bycatch
+perspective, the usual questions of fisheries tradeoffs aren't on the
+table (but can add to the discussion considerations of how much target
+fishing you give up). Given that then, while it's possible that
+"targeted" interventions would work better (ignoring costs/enforcement,
+etc), there's really no reason not to do MPAs. Unless, you actually end
+up concentrating fishing pressure outside on populations. So, we
+consider how plausible is it for that to really happen. Issue here is
+that seems likely to be pretty hand-wavey. Most likely that the vast
+majority of simulations will show positive impacts, but have no way of
+assigning probabilities to any one state of nature.
 
@@
 
-What’s the question: How effective are Blue-Water MPAs in reducing fishing mortality of important pelagic bycatch species?
+What's the question: How effective are Blue-Water MPAs in reducing
+fishing mortality of important pelagic bycatch species?
 
-Another angle.
-There’s going to be a lot of unknowns in designing BWMPA.
+Another angle. There's going to be a lot of unknowns in designing BWMPA.
 How much does not fully understanding a few key life history (e.g.
-movement) matter, and how much does the role that the species plays in the broader fleets portfolio matter
+movement) matter, and how much does the role that the species plays in
+the broader fleets portfolio matter
 
-Could simulate habitat from a multivariate normal, where you use the sigma to control how similar rasters are across species
+Could simulate habitat from a multivariate normal, where you use the
+sigma to control how similar rasters are across species
 
-Marine protected areas are increasingly proposed to protect biodiversity in pelagic systems.
-However, maby paradigms of mpa design come from coastal systems.
-Here we ask, how likely are blue-water MPAs to provide conservation benefits to bycatch species, and what are the key uncertainties underlying these outcomes?
-We begin with a general discussion of the ways in which traits of open-ocean systems might influence the conservation performance of BWMPAs.
-Given the massive parameter space, we then constrain ourselves to a styulized version of the western central pacitic ocean to ask, how feasible is it to provide broad protection of bycatch species through common MPA design paradigms in the context of this system.
-Our results demonstrate the critical importance of XX and XX in designing MPAs in pelagic systems, and highlight the potential risks of not accounting for XX.
-Introductionhere
+Marine protected areas are increasingly proposed to protect biodiversity
+in pelagic systems. However, maby paradigms of mpa design come from
+coastal systems. Here we ask, how likely are blue-water MPAs to provide
+conservation benefits to bycatch species, and what are the key
+uncertainties underlying these outcomes? We begin with a general
+discussion of the ways in which traits of open-ocean systems might
+influence the conservation performance of BWMPAs. Given the massive
+parameter space, we then constrain ourselves to a styulized version of
+the western central pacitic ocean to ask, how feasible is it to provide
+broad protection of bycatch species through common MPA design paradigms
+in the context of this system. Our results demonstrate the critical
+importance of XX and XX in designing MPAs in pelagic systems, and
+highlight the potential risks of not accounting for XX. Introductionhere
 
-Bycatch is a significant driver of population declines for many pelagic species
-Sharks, seabirds, mammals, etc
-MPAs increasingly turned to as a solution for conservation problems, etc. 
-MPA science is largely based on dynamics of coastal systems. Open-ocean ecosystems have very different dynamics that require dedicated research
-Burgess et al. (2018) estimated the reductions in fishing mortality needed to halt declines for a number of bycatch species
-Here we ask, what are the key drivers of bycatch conservation outcomes from MPAs?
-There are all of these species; you reduce fishing mortality.
-You can either do fisheries management, or you can do MPAs, let’s focus in on the MPA side.
-This could reduce fishing mortality, or it could do, and it depends on the dynamics of the species and the fleets.
-When can MPAs reduce fishing mortality enough for bycatch species? And where can you go wrong?
+Bycatch is a significant driver of population declines for many pelagic
+species Sharks, seabirds, mammals, etc MPAs increasingly turned to as a
+solution for conservation problems, etc. MPA science is largely based on
+dynamics of coastal systems. Open-ocean ecosystems have very different
+dynamics that require dedicated research Burgess et al. (2018) estimated
+the reductions in fishing mortality needed to halt declines for a number
+of bycatch species Here we ask, what are the key drivers of bycatch
+conservation outcomes from MPAs? There are all of these species; you
+reduce fishing mortality. You can either do fisheries management, or you
+can do MPAs, let's focus in on the MPA side. This could reduce fishing
+mortality, or it could do, and it depends on the dynamics of the species
+and the fleets. When can MPAs reduce fishing mortality enough for
+bycatch species? And where can you go wrong?
 
-It doesn’t make any sense to ask “do MPAs help bycatch”: the question is obviously it depends.
-To quickly make this case, we’ll assume that a species is in need of recovery.
-We then do the simple mako example to show that effects of spatial closure depend on the overlap of habitat and the assumption of fleet dynamics, and provide some empirical evidence that in particular effort redistribution is not unlikely.
+It doesn't make any sense to ask "do MPAs help bycatch": the question is
+obviously it depends. To quickly make this case, we'll assume that a
+species is in need of recovery. We then do the simple mako example to
+show that effects of spatial closure depend on the overlap of habitat
+and the assumption of fleet dynamics, and provide some empirical
+evidence that in particular effort redistribution is not unlikely.
 
-Determining the effect of an MPA network on bycatch then will also depend on trying to get close to a tactical model.
-But, this creates a problem for blue water systems in particular.
-Can make the argument that given the centrality of relatively static habitat features in coastal MPAs “protect the key habitats” might not be a bad strategy from a conservation strategy.
-But in highly mobile pelagic systems there (may be) a lot more variables in play.
-Where should researchers direct their attention?
+Determining the effect of an MPA network on bycatch then will also
+depend on trying to get close to a tactical model. But, this creates a
+problem for blue water systems in particular. Can make the argument that
+given the centrality of relatively static habitat features in coastal
+MPAs "protect the key habitats" might not be a bad strategy from a
+conservation strategy. But in highly mobile pelagic systems there (may
+be) a lot more variables in play. Where should researchers direct their
+attention?
 
-To  reduce the parameter space, we constrain ourselves to a system where 1) we know the species involved 2) we have some sense of the relative depletion and value of the different species and 3) some basic life history of the species.
-We then vary commonly unknown parameters, particularly preferred habitats, spatial catchability, movement rates etc, and pair these simulations with different MPA designs, and examine the conservation outcomes for bycatch species.
+To reduce the parameter space, we constrain ourselves to a system where
+1) we know the species involved 2) we have some sense of the relative
+depletion and value of the different species and 3) some basic life
+history of the species. We then vary commonly unknown parameters,
+particularly preferred habitats, spatial catchability, movement rates
+etc, and pair these simulations with different MPA designs, and examine
+the conservation outcomes for bycatch species.
 
-What i don’t like about this
+What i don't like about this
 
-There isn’t an “answer” to this: baring something like “changing the type of density dependence never had any effect greater than XX%”, for anything that on occasion has an impact the answer of “how much” will almost always depend on what else is going on.
+There isn't an "answer" to this: baring something like "changing the
+type of density dependence never had any effect greater than XX%", for
+anything that on occasion has an impact the answer of "how much" will
+almost always depend on what else is going on.
 
@@
 
-Pointing out a problem, which to make impactful would require making the case that there is a problem with current design strategies.
-I.e.
-“MPAs are often designed like this (simple dispersive models with constant habitat, or MARXSAN style things), and doing that does/does not create problems
-A tactical evaluation.
-Picking a handful of high profile MPAs, parameterizing all the knownns as well as possible, iterating over unknowns, and see what the range of outcomes are.
-Cleaner, more impactful, but also opens it up (depending on the results) to a details fight about very specific modeling choices
+Pointing out a problem, which to make impactful would require making the
+case that there is a problem with current design strategies. I.e. "MPAs
+are often designed like this (simple dispersive models with constant
+habitat, or MARXSAN style things), and doing that does/does not create
+problems A tactical evaluation. Picking a handful of high profile MPAs,
+parameterizing all the knownns as well as possible, iterating over
+unknowns, and see what the range of outcomes are. Cleaner, more
+impactful, but also opens it up (depending on the results) to a details
+fight about very specific modeling choices
 
@@
 
-Collect life history for key bycatch species and target species, cite Bradley et al. pulling together species distribution and threats etc. 
-Introduce marlin
-Parameterize marlin runs with “knowns” from life history of bycatch and target species
-Highlight key “unknowns” that we will iterate over, e.g. fleet models, density dependent movement, timing of density dependence etc
-Simulate effects of MPAs given reasonably realistic situations but iterating over key unknowns
-Evaluate the range of outcomes and key drivers of MPA effects
-I’m thinking we try and avoid a really specific case study but hard not to accidentally do one without making everything up….
+Collect life history for key bycatch species and target species, cite
+Bradley et al. pulling together species distribution and threats etc.
+Introduce marlin Parameterize marlin runs with "knowns" from life
+history of bycatch and target species Highlight key "unknowns" that we
+will iterate over, e.g. fleet models, density dependent movement, timing
+of density dependence etc Simulate effects of MPAs given reasonably
+realistic situations but iterating over key unknowns Evaluate the range
+of outcomes and key drivers of MPA effects I'm thinking we try and avoid
+a really specific case study but hard not to accidentally do one without
+making everything up....
 
-Results
-Range of likely changes in fishing-induced mortality for the species selected
-Key drivers of uncertainty in those outcomes, focusing on things that switch states from reduction to increase in mortality
-How common are conservation tradeoffs between bycatch species?
-Fishery impacts
+Results Range of likely changes in fishing-induced mortality for the
+species selected Key drivers of uncertainty in those outcomes, focusing
+on things that switch states from reduction to increase in mortality How
+common are conservation tradeoffs between bycatch species? Fishery
+impacts
 
-Discussion
-What does this say about BWMPAs and bycatch?
-Not going to get into design decisions here, depending how common tradeoffs are I think that will be the main focus
-Highlight any species that frequently induced tradeoffs
-Tee up future research
-MPAs in context of other management
-MPAs as risk reduction strategy
-Etc.
+Discussion What does this say about BWMPAs and bycatch? Not going to get
+into design decisions here, depending how common tradeoffs are I think
+that will be the main focus Highlight any species that frequently
+induced tradeoffs Tee up future research MPAs in context of other
+management MPAs as risk reduction strategy Etc.

### 2021-05-05 documents/marlin-paper-notes.Rmd
Habitat retool (#2)
* converting to makefile, major restructure, back to main

---
title: "marlin-paper-notes"
author: "Dan Ovando"
date: "4/14/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

what dooes the overlap tradeoff curve look like?

Suppose you close off 30% of an area in an MPA

Then plot conservation outcome on y-axis, degree of overlap between target and bycatch species on the x axis. When overlap is one, conservation benefit should be maxed. Then overlap is perfeclty anticorrelated, if effort concentrates, conservation should be at some negative minimum. If concentration doesn't happen, then conservation effect should be zero. If overlap is uncorrelated... unclear?

MPA literature has been dominated by single-species model covering homogemous habitat without any technical interactions across multiple fleets and species.

@hastings2017 assumes that adults are stationary, larvae are perfectly mixed, age structure can be ignored, density dependence at time of settlement. Same harvest rate outside reserve.

static spatial, dynamic spatial , input / output controls, gear

If there is perfect overlap, then improvement in coservation outcome should be a straight positive function of area protected. If there is no overlap, then

# what's the angle?

Smith et al and references therein make the basic case pretty damn clear: when species move a lot and when habitats overlap, static MPAs don't work as well as dynamic. There are other nuances in there but that kind of summarizes a lot of it

so what are the specific angles this paper can take

- When do static closures go from "less effective" to "not effective"

Those papers point out "less effective" since max species = 2 to 4. But, we're seeing here that non-trivial case of net loss? But is it really non-trivial? like 15% of simulations with more than 10%? though that number is a lot bigger if you're basically asking "did it help"

So that's certainly an angle that's unexplored here.

- Yet another comparison to active fisheries management

Tying more directly to fisheries management, compare to say a TAE, in terms of units bycatch saved per unit of target catch lost

Could also do the dynamic comparison, though that one is a bit boring to me since it's sort of self evident, so question is the wedge?

Could also ask something about how much would gear efficiency have to change, though it feels like that would be so detail derived it would be hard to make general

- Value of information

- A more tactical evaluation

Lean in a bit more to an area and specific existing or proposed MPAs with best estimates of bycatch risk / target fleets in space.

The cleanest to what we have now, but also the most contentious / work. If we go with actual MPAs, really need to dial in the species and fleets, which we're pretty close to do be honest, but also the habitats etc.

Though there is a more general question like "what's the best outcome you could get for bycatch with static closures in a place that looks like the WCPO". Or rather what design gets you the best outcome. Now you'd have to put a constraint on that, say total catches can't decline by more than xx%?

Hmm I like this one sort of

- incentives piece around bycatch
incentive to fish and incentive to protect

So this isn't the odd rare turtle, these are valuable sharks

### 2021-05-05 R/compile-fleet.R

@@ -124,6 +124,16 @@ compile_fleet <- function(fauna, tune_type = "explt") {
           p_explt = .99,
           sel_unit = "length"
 
+        ),
+        "xiphias gladius" = Metier$new(
+          critter = fauna$`xiphias gladius`,
+          price = 7.47,
+          sel_form = "logistic",
+          sel_start = 150,
+          sel_delta = 0.01,
+          catchability = .1,
+          p_explt = .99,
+          sel_unit = "length"
         )
       ),
       base_effort = resolution
@@ -250,6 +260,16 @@ compile_fleet <- function(fauna, tune_type = "explt") {
           p_explt = 0.01,
           sel_unit = "length"
 
+        ),
+        "xiphias gladius" = Metier$new(
+          critter = fauna$`xiphias gladius`,
+          price = 7.47,
+          sel_form = "logistic",
+          sel_start = 999,
+          sel_delta = 0.01,
+          catchability = .1,
+          p_explt = .01,
+          sel_unit = "length"
         )
       ),
       base_effort = resolution

### 2021-05-07 documents/marlin-paper-notes.Rmd
testing on ubuntu

@@
 So this isn't the odd rare turtle, these are valuable sharks
+
+## optimization
+
+similar SIR approach... why not just get it running first
+
+Yields: sum of revenue of target species: or if you just say revenue, then you can manipulate prices to get desired impact: if you say that bycatch literally has no value then just set price to 0, or negative.
+
+biodiversity: B/K of all species (for now): average or sum? Sum I think, since you'd rather have all at 0.5 than half at 0 and half at 1
+
+yield_weight: do you have MSY for all species?
+I don't think so, and gets crazy complicated to think about MSY once you start messing with technical interactions etc.
+So you just need something to weight the marginal value of both.
+
+So, for every step, you'll have N candidate cells, each with a marginal yield effect and a marginal biodiversity effect.
+You then need to rank each cell by the objective function.
+Need some function to convert yield and biomass into comparable units then.
+So just center and scale?
+alpha then controls the relative weight of a standard deviation increase in yield relative to a standard deviation increase in biodiversity?
+Issue there: once you're down to the last cell this won't work.
+Dividing by max caps both at 1, but no limit to the loss there...
+Rescale to be positive?
+yield - min(yield) + 1e-3 will ensure the min value is 1e-3, then scaling by max of that vector will make the max 1, so now everything has the same range.
+Oh, or just use scaes::rescale to put both on range of 0 to 1.
+
+Is that going to have weird behavior rescaling at every iteration like that?
+i.e.
+the effective value of alpha will change depending on the scale of the biomass and yield.
+Setting alpha to 1 will maximize yields, setting it to 0 will maximize biodiversity, what does setting to 0.5 mean?
+value the best yield cell as much as the best biomass cell.
+
+Seems good enough for now. So doing it this way, you only have to iterate over the weight, no different objective function shapes
+
+So what are you optimizing?
+Location and size or just location?
+In theory, you can save some computation time by stopping the optimization when the objective function starts to consistently decline.
+But, since there might be multiplie maxima here, that could get tricky, and wouldn't allow you to look at tradeoff wedges.
+So...
+
+- for each weight
+- for each range of MPA sizes
+- find the optimal MPA configuration. 
+- Not too bad!
+
+## what are you comparing optimization to
+
+OK, you have some optimized results, so what.
+On the one hand, you can examine to what extent given this case study, tradeoffs across species or across yields are inevitable or just gotta find the right network.
+
+From there though, can also compare "optimal" to "realistic"
+
+Design strategies
+1. Assign protection by SSB / SSB0 weighted abundance in space (protect the highest concentrations of the most threatened species)
+2. Assign protection by bycatch rates (distribution of biomass bycatch / biomass target)
+3. Assign protection by targeting fishing effort
+4. Assign protection by avoiding fishing effort
+
+Knowledge states
+1. Degrees of accuracy over the thing being measured (i.e. error in the design strategy)
+
+2. Number of species considered (e.g. design around only 2 bycatch species instead of all)
+
+3. über naive: just close off random patches up to a specified size
+
+We then plot these points in objective function space and compare to the efficiency frontier / consider frequency and magnitude of conservation and economic tradeoofs
+
+The question all this is answering: how sensitive are conservation and economic outcomes to design strategy / uncertainty when there are technical interactions and heterogeneity in habitat?
+The first stage is teeing up the idea that beyond conservation / yield tradeoffs, as you increase habitat heterogeneity with technical interactions you increase the potential for conservation tradeoffs.
+But, the degree is going to be super context dependent, so to dive in a bit we then focus on this case study of the particular species / habitat overlap in the WCPO.
+Further work might reveal more generalities in this, but htis is basically getting a sense for the potential magnitude of the challenge.
+
+OK, this is actually coming together! You should run the same design / knowledge states for the broader range of simulated habitats. That will make the point nice and clearly.
+
+But, that begs the question of, is it an inevitable tradeoff, or just a case of poor design.
+That requires optimization.
+But, optimization is very computationally intensive.
+To examine this point then, we repeated the exercies, but with a specific case study where we hold habitat as constant, but redo the exercise conditional on constant habitat.
+We can then see to what extent tradeoofs persist, and ideally come up with some way of comparing the case study habitats to the metric of habitat heterogeneity in the full simulations.
+Case study also demonstrates that its a real thing: not making up the idea of different habitat use
+
+This might actually be a cool paper. Proc B?

### 2021-05-07 R/optimize-mpa.R

@@ -1,0 +1,265 @@
+#' Optimize MPA network
+#'
+#' @param fauna fauna object
+#' @param fleets fleet object
+#' @param alpha the weight given to conservation relative to economics (range 0 to 1)
+#' @param max_prop_mpa max proportion of cells to place in MPA
+#' @param resolution the resolution of the simulated system
+#' @param prop_sampled the proportion of cells to sample in each SIR iteration
+#' @param workers number of workers for parallel process
+#'
+#' @return a list with results of MPA optimization
+#' @export
+#'
+optimize_mpa <-
+  function(fauna,
+           fleets,
+           alpha = 0.33,
+           max_prop_mpa = 1,
+           resolution,
+           prop_sampled = .2,
+           workers = 6) {
+    workers <- 1
+    future::plan(future::multisession, workers = workers)
+
+    on.exit(future::plan(future::sequential))
+
+    fauna <- casestudy$fauna[[1]]
+
+    fleets <- casestudy$fleet[[1]]
+
+    alpha <- 0
+
+    percent_mpa <- 0.3
+
+    prop_sampled <- 0.2
+
+    max_prop_mpa <-  1
+
+    patches <- resolution ^ 2
+
+    samps <- round(prop_sampled * patches)
+
+    mpas <- expand_grid(x = 1:resolution, y = 1:resolution) %>%
+      mutate(patch = 1:nrow(.)) %>%
+      mutate(mpa = FALSE)
+
+    max_patches_protected <- round(patches * max_prop_mpa)
+
+    candidate_patches <- 1:patches
+
+    patch_value <- tibble(patch = 1:patches, obj_value = 0.5)
+
+    results <- vector(mode = "list", length = max_patches_protected)
+
+    mpa_network <-
+      vector(mode = "list", length = max_patches_protected)
+
+    calc_objective_function <-
+      function(candidate_patch, fauna, fleets, mpas) {
+        tmp_mpas <- mpas
+
+        tmp_mpas$mpa[tmp_mpas$patch %in% candidate_patch] <- TRUE
+
+        sim_mpa <- simmar(
+          fauna = fauna,
+          fleets = fleets,
+          years = years,
+          mpas = list(locations = tmp_mpas,
+                      mpa_year = 1)
+        )
+
+        res <-
+          sim_mpa[[length(sim_mpa)]] # for now, just calculate in the final timestep
+
+        biodiv <-
[…]

### 2021-05-07 make-marlin-paper.R

@@ -1,1202 +1,0 @@
-# Assessing Tradeoffs in the use of MPAs for Bycatch Reduction
-
-#something about relative magnitude given finite resouces to collect data?
-#
-# what have the existing MPAs done for sharks
-# setup -------------------------------------------------------------------
-
-# idea: define a centroid of habitat and a CV around that for each species and use that to tune the scenarios..
-# like manually creats random centroids that avoid each other... then increase the CV, such that there really is no key habitat when they... no I don't like that
-#
-# what if you randomly sample the centroid for each species, where when the cv of the sampling is low, most species have the same centroid, and as that icnreases you get more and more different centroids for each species.... that's a better idea.. centroid by gindinf the index of center cell as a default starting position
-#
-# iterate over cv of the centroid,
-# and then cv around the centroid
-
-set.seed(42)
-library(tidyverse)
-
-library(marlin)
-
-library(here)
-
-library(furrr)
-
-library(ggridges)
-
-library(googledrive)
-
-library(data.table)
-
-library(gamm4)
-
-library(sf)
-
-library(mvtnorm)
-
-library(patchwork)
-
-options(dplyr.summarise.inform = FALSE)
-
-
-foos <- list.files(here("R"))
-
-walk(foos, ~ source(here("R", .x)))
-
-results_name <- "v0.5"
-
-results_path <- here("results", results_name)
-
-if (!dir.exists(results_path)){
-  dir.create(results_path, recursive = TRUE)
-
-  dir.create(file.path(results_path,"sims"))
-}
-
-run_experiments <- FALSE
-
-safety_stop <- FALSE
-
-draws <- 20
-
-resolution <- 20 # resolution is in squared patches, so 20 implies a 20X20 system, i.e. 400 patches
-
-years <- 40
-
-seasons <- 2
-
-time_step <- 1 / seasons
-
-workers <- round(parallel::detectCores() - 8)
-
-steps <- years * seasons
-
-time_steps <- seq(0,years - 1, by = time_step)
-
-theme_set(marlin::theme_marlin())
-
-tune_type <- "explt"
-
-plan(multicore, workers = workers)
-
-
[…]

### 2021-05-07 scripts/00_setup.R

@@ -11,7 +11,7 @@ library(ggridges)
 
 library(googledrive)
 
-library(data.table)
+# library(data.table)
 
 library(gamm4)
 
@@ -38,15 +38,17 @@ if (!dir.exists(results_path)){
   dir.create(file.path(results_path,"sims"))
 }
 
-run_experiments <- TRUE
+run_experiments <- FALSE
 
-run_casestudy <- TRUE
+run_casestudy <- FALSE
 
 safety_stop <- FALSE
 
+optimize_casestudy <- TRUE
+
 draws <- 20
 
-resolution <- 20 # resolution is in squared patches, so 20 implies a 20X20 system, i.e. 400 patches
+resolution <- 10 # resolution is in squared patches, so 20 implies a 20X20 system, i.e. 400 patches
 
 years <- 40
 

### 2021-05-17 R/optimize-mpa.R
first real version of results

@@ -14,27 +14,28 @@
 optimize_mpa <-
   function(fauna,
            fleets,
+           starting_conditions = NA,
            alpha = 0.33,
            max_prop_mpa = 1,
            resolution,
            prop_sampled = .2,
            workers = 6) {
-    workers <- 1
+    # workers <- 6
     future::plan(future::multisession, workers = workers)
 
     on.exit(future::plan(future::sequential))
 
-    fauna <- casestudy$fauna[[1]]
-
-    fleets <- casestudy$fleet[[1]]
-
-    alpha <- 0
-
-    percent_mpa <- 0.3
-
-    prop_sampled <- 0.2
-
-    max_prop_mpa <-  1
+    # fauna <- casestudy$fauna[[1]]
+    #
+    # fleets <- casestudy$fleet[[1]]
+    #
+    # alpha <- 0
+    #
+    # percent_mpa <- 0.3
+    #
+    # prop_sampled <- 0.2
+    #
+    # max_prop_mpa <-  1
 
     patches <- resolution ^ 2
 
@@ -56,7 +57,7 @@ optimize_mpa <-
       vector(mode = "list", length = max_patches_protected)
 
     calc_objective_function <-
-      function(candidate_patch, fauna, fleets, mpas) {
+      function(candidate_patch, fauna, fleets, mpas,starting_conditions) {
         tmp_mpas <- mpas
 
         tmp_mpas$mpa[tmp_mpas$patch %in% candidate_patch] <- TRUE
@@ -66,7 +67,8 @@ optimize_mpa <-
           fleets = fleets,
           years = years,
           mpas = list(locations = tmp_mpas,
-                      mpa_year = 1)
+                      mpa_year = 1),
+          initial_conditions = starting_conditions[[1]]
         )
 
         res <-
@@ -98,7 +100,8 @@ optimize_mpa <-
           fauna = fauna,
           fleets = fleets,
           mpas = mpas,
-          # .options = furrr_options(seed = 42),
+          starting_conditions = starting_conditions,
+          .options = furrr_options(seed = 42),
           .progress = FALSE
         )
       # Sys.time() - a
@@ -131,7 +134,8 @@ optimize_mpa <-
         fleets = fleets,
         years = years,
         mpas = list(locations = mpas,
-                    mpa_year = 1)
+                    mpa_year = 1),
+        initial_conditions = starting_conditions[[1]]
       )
 
       res <-
@@ -172,7 +176,8 @@ optimize_mpa <-
 
[…]

### 2021-05-17 scripts/00_setup.R

@@ -40,11 +40,11 @@ if (!dir.exists(results_path)){
 
 run_experiments <- FALSE
 
-run_casestudy <- FALSE
-
-safety_stop <- FALSE
+run_casestudy <- TRUE
 
 optimize_casestudy <- TRUE
+
+safety_stop <- FALSE
 
 draws <- 20
 
@@ -65,9 +65,6 @@ time_steps <- seq(0,years - 1, by = time_step)
 theme_set(marlin::theme_marlin())
 
 tune_type <- "explt"
-
-plan(multicore, workers = workers)
-
 
 # load data ---------------------------------------------------------------
 

### 2021-05-19 R/create-critters.R
case study and experiments done

@@ -4,7 +4,8 @@ create_critters <-
            seasons = 1,
            marlin_inputs,
            seasonal_movement = FALSE,
-           ontogenetic_shift = FALSE) {
+           ontogenetic_shift = FALSE,
+           random_rec = FALSE) {
     # sciname <- marlin_inputs$scientific_name[[1]]
     #
     #
@@ -31,9 +32,13 @@ create_critters <-
       hab <- list(hab, hab)
     }
 
-    # rec_form <-  sample(c(0, 1, 2, 3), 1, replace = TRUE)
+    if (random_rec){
+    rec_form <-  sample(c(0, 1), 1, replace = TRUE)
 
-    rec_form <- 1 # sample(c(0, 1, 2, 3), 1, replace = TRUE)
+    } else {
+      rec_form <- 1
+    }
+    # rec_form <- 1 # sample(c(0, 1, 2, 3), 1, replace = TRUE)
 
 
     if (ontogenetic_shift) {

### 2021-05-19 scripts/00_setup.R

@@ -40,9 +40,9 @@ if (!dir.exists(results_path)){
 
 run_experiments <- FALSE
 
-run_casestudy <- TRUE
+run_casestudy <- FALSE
 
-optimize_casestudy <- TRUE
+optimize_casestudy <- FALSE
 
 safety_stop <- FALSE
 

### 2021-05-19 scripts/01_run_experiments.R

@@ -1,298 +1,5 @@
 source(file.path("scripts", "00_setup.R"))
 
-wcpfc_data <- read_csv(here("data", "wcpfc_monthly.csv"))
-
-unique(wcpfc_data$species_commonname)
-
-top_bycatch <- wcpfc_data %>%
-  filter(catch > 0) %>%
-  group_by(species_commonname) %>%
-  count() %>%
-  arrange(desc(n)) %>%
-  ungroup() %>%
-  slice(1:4)
-
-wcpfc_data <- wcpfc_data %>%
-  filter(species_commonname %in% top_bycatch$species_commonname)
-
-marlin_inputs$bycatch <-
-  marlin_inputs$scientific_name %in% unique(tolower(wcpfc_data$species_sciname))
-
-
-bycatch_ll_cpue <- wcpfc_data %>%
-  filter(gear == "longline") %>%
-  sf::st_as_sf(
-    coords = c("longitude", "latitude"),
-    crs = "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0",
-    remove = FALSE
-  ) %>%
-  recenter_vector(center = -80) %>%
-  mutate(cpue = catch / effort) %>%
-  group_by(species_commonname, longitude, latitude, year) %>%
-  summarise(cpue = sum(cpue)) %>%
-  group_by(species_commonname, longitude, latitude) %>%
-  summarise(cpue = mean(cpue)) %>%
-  group_by(species_commonname) %>%
-  mutate(scaled_cpue = scale(cpue),
-         bycatch = TRUE) %>%
-  ungroup()
-
-
-bycatch_ll_cpue_plot <- bycatch_ll_cpue %>%
-  ungroup() %>%
-  ggplot(aes(color = scaled_cpue)) +
-  geom_sf(size = 2) +
-  facet_wrap( ~ species_commonname) +
-  scale_color_viridis_c(trans = "log10")
-
-
-target_ll_catch <- wcpfc_data %>%
-  select(year, longitude, latitude, contains("target_catch")) %>%
-  group_by(year, longitude, latitude) %>%
-  select(contains("_longline") &
-           contains("tonnes") |
-           (contains("_SKIPJACK") & contains("tonnes"))) %>%
-  select(contains("TUNA") | contains("MARLIN")) %>%
-  pivot_longer(contains("target_"),
-               names_to = "species_commonname",
-               values_to = "catch") %>%
-  mutate(species_commonname = tolower(as.character(
-    str_extract_all(
[…]

### 2021-06-10 documents/03_marlin_paper.Rmd
first draft of paper. 01_ is completely broken to some accidental deleting and undoing, so waiting until home to revert to backup version from 2021-06-09 at 6:28AM

---
title: "Solutions to the potential for unintended conservation consequences of MPAs"
author:
    - Daniel Ovando
    - Darcy Bradley
    - Lennon Thomas
    - Echelle Burns
date: "`r Sys.Date()`"
output: 
  bookdown::pdf_document2: default
  bookdown::html_document2: default
bibliography: references.bib
params:
  run_name: ["v1.0"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, include = TRUE)

source(here::here("scripts", "00_setup.R"))

theme_set(marlin::theme_marlin())

```

```{r load-results}

results_files <- list.files(file.path(results_path))

rdses <- results_files[str_detect(results_files,'.rds')]

assign(str_remove(rdses[8],".rds"), read_rds(file.path(results_path,rdses[2])))

walk(rdses,~assign(str_remove(.x,".rds"), read_rds(file.path(results_path,.x)), envir = .GlobalEnv)) #load and assign all .rds objects in the appropriate results folder

rdatas <- results_files[str_detect(results_files,'.Rdata')]

walk(rdatas, ~load(file.path(results_path,.x), envir = .GlobalEnv))

# a bit of housekeeping to avoid saving a massive object
tmp <- raw_experiment_results  %>%
  map( ~ .x$obj)

group_var <- "placement_strategy"

total_results <- total_experiment_results

rm(total_experiment_results)

results <- placement_experiment %>%
  mutate(temp = raw_experiment_results) %>%
  group_by_at(colnames(.)[!colnames(.) %in% c("temp", "prop_mpa")]) %>%
  nest() %>%
  ungroup() %>%
  mutate(placement_id = 1:nrow(.)) %>%
  unnest(cols = data) %>%
  mutate(xid = map(raw_experiment_results, names)) %>%
  unnest(cols = c(temp, xid)) %>%
  mutate(obj = map(temp, "obj"),
         xid = as.numeric(xid)) %>%
  unnest(cols = obj)

if (any(results$econ[results$prop_mpa == 1] > 0)) {
  stop("something has gone wrong: yields present with 100% mpa")
}

warning("move placement experiement id back above run_experiments")

# total_experiment <- placement_experiment %>%
#   left_join(experiments, by = 'xid')

# results <- results %>%
#   left_join(placement_experiment)

# compare to BAU

results <- results %>%
  group_by(placement_id, xid, critter) %>%
  mutate(bau_biodiv = biodiv[prop_mpa == 0],
         bau_econ = econ[prop_mpa == 0])

excols <-
  experiments %>%
  select(xid, data) %>%
  unnest(cols = data) %>%
  select(-habitat,-critter, -scientific_name, -centroid) %>%
  ungroup() %>%
  unique()

```

# Introduction

Marine protected areas (MPAs) are a growing part of the marine
conservation toolbox, with 7.4% of the world's oceans currently
protected and, and bodies such as the IUCN calling for 30% protection by
2030. Many of the largest of these MPAs are being implemented in the
form of "blue-water" MPAs largely covering pelagic ecosystems,
characterized by highly mobile species and fishing fleets using a
diverse mix of static and dynamic ocean features. These systems operate
very differently from the coastal areas where modern MPA design
practices have been developed. Pelagic systems are also characterized by
combinations of relatively abundant species targeted by fishing, such as
most tuna stocks, and highly threatened species caught as incidental
"bycatch" in targeted fishing operations, such as oceanic whitetip
sharks. Blue-water MPAs are increasingly being proposed specifically to
benefit these by-catch populations while minimizing losses or where
possible providing benefits to fisheries.

While the effect of MPAs on fisheries yields is a continuously debated
topic, the conservation impacts of MPAs have received less attention of
late, due in part perhaps to the perceived obvious nature of the
problem: if you prevent fishing of a species within an area, there will
be more of that species in that area, and so the conservation benefits
of MPAs will increase with MPA size. \@hastings2017 clearly illustrates
this idea. Using a simple two-species two-patch analytical model, they
derive conditions under which MPAs can solve the "weak-stock" problem in
which a less productive species is incidentally exploited by a fleet
targeting a more valuable and productive stock.

However, as @hastings2017 point out, their results depend on a very
specific set of assumptions, namely that 1) adults are perfectly
sedentary 2) the target and bycatch species overlap perfectly in terms
of habitat xx....The authors suggest that the impact of these
assumptions should be tested in scenarios under which they are unlikely
to apply. We build on \@hastings2017 to consider the role of MPAs in reducing bycatch reductino while managing tradeoffs with fisheries under a more realistic model reflecting the dynamics of pelagic ecosystems and fisheries.

The core models underpinning MPA science have largely been based on
stylized versions of near-shore reef-style xx what to call that, like
coral / rocky / kelp forest, etc xx systems, where multiple species
often share common habitat, and adults have reasonable site fidelity
while larvae are broadly dispersed. Pelagic ecosystems have vastly
different dynamics than coastal seas [@game2009; @lubchenco2003],
characterized by highly migratory species specialized to take advantage
of highly dynamic ocean features such as eddies currents while also
utilizing static features such as seamounts. Pelagic ecosystems also
support both coastal domestic fishing industries along with highly
mobile international high-seas fishing operations. Given these dynamics,
it remains unclear how many of the lessons from coastal MPA science are
applicable to pelagic ecosystems [\@ovando2021XX].

Pelagic ecosystems present both opportunities and challenges for MPAs.
The highly productive and mobile nature of targeted fisheries in these
systems may mean that highly targeted MPAs can be designed that provide
effective conservation at minimal fishery costs. Strategic protection of
key aggregation locations may enhance depleted pelagic fisheries [xx
that Indian ocean study I can't remember right now]. xx big question
here: how focused do we want this paper to be? is it just about
conservation tradeoffs or lack thereof with bycatch species, or is about
the net conservation and fishery effects of BWMPAs. I'm inclined the
former to tee-up the later xx.

However, the dynamic nature of pelagic systems also creates a risk of unintended conservation consequences of MPAs. When an MPA is implemented on former fishing grounds, fishing fleets have two choices: exit the fishery or move to new fishing grounds. When MPAs redistribute effort away from one area and into another, they can accidentally increase fishing effort on species found within the new fishing grounds [@abbott2012]. This problem of policy "leakage" is well known in many environmental fields [@citations] but the general properties of this problem have received relatively little attention in the theoretical or empirical MPA literature [see @abbott2012;@savina2008;fulton2011a;@baum2003 for notable exceptions]. We fill this gap by

- Assessing the conditions under which this may be a problem using a generalized simulation environment
  - General insights about design strategy, effect of varying degrees of habitat overlap, etc, tradeoffs with fisheries

- Given that "it depends" nature of the problem, assess the specific potential for this problem in a case study
  - specific insights about design strategy, effect of varying degrees of habitat overlap, etc, tradeoffs with fisheries

Our goal here is not to provide tactical recommendations on the design
of blue-water MPAs, but rather to provide a strategic evaluation of the
sensitivity of bycatch populations and targeted fisheries to key
unknowns and design choices in MPA planning.

<!-- The outcome of MPAs for fishing and conservation are highly sensitive to -->

<!-- a wide range of economic and ecological factors. In order to constrain -->

<!-- the possible parameter space, we tune a simulation model based around -->

<!-- best available knowledge of the species and fleets operating in the -->

<!-- areas of the western central pacific ocean (WCPO) while allowing key -->

<!-- unknowns to vary, in particular the nature of adult and larval movement, -->

<!-- the degree of habitat overlap between different species, the economic -->

<!-- incentives facing fishing fleets, and the design strategy used in MPA -->

<!-- placement XX. Most conventional MPA design models are either static, -->

<!-- single species, or both. We present a dynamic, multi-species simulation -->

<!-- model that allows us to cohesively assess the impact of MPAs on a suite -->

<!-- of by-catch and targeted species collectively affected by multiple -->

<!-- fishing fleets.  -->

<!-- We show - something about how when species are homogeneous in space, -->

<!-- conservation gains are basically guaranteed except for constant catch -->

<!-- scenearios, which need to be considered in tuna-type scenarios - When we -->

<!-- introduce heterogeneity in stock structure roughly reflective of -->

<!-- distributions in the WCPO we do / do not begin to see tradeoffs in -->

<!-- conservation outcomes among bycatch species - quantify if there is a -->

<!-- concave function to this - Specific design paradigms are / are not more -->

<!-- / less resilient to this issue - Compare tradeoffs between MPA designs -->

<!-- that minimze internal bycatch tradeoffs (there have to be none if MPA = -->

<!-- 100%) and fishery outcomes -->

<!-- We show that MPAs increased biomass of by-catch species in XX% of -->

<!-- simulations, and decreased them in XX% of simulations. On average a 1% -->

<!-- increase in by-catch species resulted in a XX% change in fishery -->

<!-- catches... Outcomes were most sensitive to XX and XX, and least -->

<!-- sensitive to XX and XX. We do not suggest that these same results would -->

<!-- hold for other systems, but rather demonstrate the importance / -->

<!-- resilience of MPAs to key uncertain variables. In particular for pelagic -->

<!-- ecosystems, our work demonstrates the importance of considering the -->

<!-- collective impacts of space, vulnerability, life history, and fleet -->

<!-- dynamics in MPA design for conservation. -->

# Methods

@hastings2017 employed a simply model to derive conditions under which
MPAs would benefit bycatch species. However they also pointed out that
even small increases in model complexity quickly make analytical
solutions to the effects of MPAs intractable. As such, we based our numerical
analysis on a spatially explicit, multi-species, multi-fleet,
bio-economic model we call
[`marlin`](https://github.com/DanOvando/marlin). Models of MPAs span a
wide range of complexity, from simple analytical tools to end-to-end
ecosystem models [@fulton2015]. `marlin` follows a "models of
intermediate complexity" (MICE) philosophy [@plagányi2014] targeted
specifically around the role of technical interactions across species.
Many MPA models simulate the impacts of MPAs on a single-species level [@ovando2021XX]. This can be useful for for example examining the ways in which species-specific life history traits interact with fleet dynamics to influence MPA outcomes. However, bycatch problems are driven by technical interactions, in which fishing effort for a target species incidentally captures another species xx how do we want to talk about this in here? Like usually incidentally, but clearly sometimes people try and catch sharks too xx. `marlin` accounts for this by allowing the user to simulate an arbitrary number of fleets capable of catching an arbitrary number of species in a spatial area. Different fleets can place different economic value on different species, while also having species-specific selectivity functions. This allows us to simulate bycatch dynamics of MPAs in pelagic ecosystems.

We first ran a series of simulation exeriments designed to explore broad sensitivities of bycatch and fisheries to key factors.

We then ran a focused case study intended to reflect some of the traits of the tuna fisheries of the WCPO, holding constant...

## `marlin`

Lots of equations. Pretty standard age-structured population models, extra bells and whistles that need explaining

- Design of movement and habitat
  - Movement is a product of distance and habitat, such that critters will all else equal move to the closest patches with the best habitat

- Ontogenetic shifts

- Nature of fleet tuning (shared effort different q per species)

-

## Simulation Experiments

Given the complexity of the operating model used here, a complete sensitivity analysis across all parameters is not practical. Instead, we focus on the role of XX key attributes of pelagic ecosystems: the degree of spatial overlap between populations, the presence of seasonal migrations, and the potential for ontogenetic shifts in habitat use, and degree of fishing pressure on each species.

## Species & Fleets

Each simulation considers `n_distinct(marlin_inputs$scientific_name)` species comprised of the top 12 species by landings reported in the tuna fisheries of the Western and Central Pacific Ocean (WCPO). Rough Life history parameters are pulled from `FishLife` [@thorson2020]. For teleost fishes, we assumed that fecundity was proportional to weight-at-age times the maturity ogive. For cartilaginous species, we adopt a fixed fecundity relationship of XX pups per mature individual XX check on what I actually implemented here XX. The unfished spawning stock biomass for each of the included species was tuned in the model to reflect best available estimates from stock assessments.

Each simulation includes a longline and a purse-seine fleet. Each species can have a separate price (USD/kg) and selectivity ogive for each fleet, as dictated by the literature.

Fleets exert a given amount of effort in a patch each time step. The catch resulting from that effort is given by the Baranov [@baranov18XX] equatino

$$C_{f,s,p,t} = \frac{q_{f,s}E_{f,p,t}}{q_{f,s}E_{f,p,t} + m_s} \times \sum_{a=0}^{A}B_{s,p,t,a} \times (1 - e^{-(q_{f,s}E_{f,p,t}S_a + m_s)}) $$

Where $C$ is catch by fleet $f$ of species $s$ in patch $p$ in timestep *t*. $E$ is total effort by that fleet in a patch and time step, $m$ is natrual mortality for a given species, $B$ is biomass at age *a*, and $S$ is selectivity.

$q$ is the catchability coefficient for fleet $f$ and species $s$.

Fleets distribute themselves in space according to revenue per unit effort

$$R_{f,p,t} = \frac{\sum_{s=1}^{S}v_{f,s}C_{f,s,p,t}}{E_{f,p,t}}$$
Where $v$ is the price per kg of each species. Bycatch in the model is defined through the price. An unregulated species with no commercial value would have $v = 0$, whereas a regulated and prohibited species might actually have a negative price.

Effort per fleet is set a a constant of 1 per patch, and the model then tunes $q_{f,s}$ such that given all the other parameters of the model the fishing mortality rate for each species matches the best available level for that species. This is what allows the model to capture bycatch dynamics: A fleet can distribute itself purely based on revenues from target species, but through the $q$ parameter still exert fishing mortality on other species, even if that species has a price of zero, if those bycatch species inhabit the same patches as the more valuable target species.

### Species Overlap

We simulate degrees of species overlap using two parameters, $\sigma_c$ and $\sigma_h$. All species start with the assumption that the centroid of their preferred habitat occurs in the center-most patch of teh two dimensional system simulated by marlin. For our simulations, we then generated a random deviation from that default centroid for each species using $\sigma_c$ in the manner

$$centroid_s \sim normal(center,\sigma_c)$$
where *s* denotes species, and *center* is the coordinate of the central cell. In this way, the larger $\sigma_c$ grows, the more variation there is betwen the centroid of the preferred habitat for each species.

We then control the spread of a creatures preferred habitat around that centroid per a Gaussian dispersal kernal in the manner

$$habitat_{p,s} = \frac{1}{\sqrt{2\pi\sigma_h^2}}e^{-\frac{d_p^2}{2\sigma_h^2}}$$

Where $d$ is the euclidian distance from the $centroid_s$ to patch $p$ and $\sigma_h$ controls the dispersal of the habitat around that range.

### Ontogenetic Shifts

`marlin` allows for different habitats to be used by adults vs. recruits. For scenarios with ontogenetic shifts ,we assign recruits to a habitat grid that is the inverse of $habitat_{p,s}$. In this way organisms enter the model living on one area, and then transition to adult habitat at older life stages.

XX this isn't very interesting at this point since movement is so high so I vote we get rid of it for now XX.

### Seasonal Migration

For this paper `marlin` is set to run with two seasons per year. When seasonal movement is turned on, animals use one habitat matrix in season one, and the inverse of that habitat matrix in season 2.

### MPA Design Strategies

MPA design is governed by four general parameters: the number of cells to place inside an MPAs, and the design strategy used to place the MPAs, the species considered in the MPA planning process, and the accuracy of the information used in the planning process.

We consider `r n_distinct(experiment_obj$placement_strategy)` different MPA placement strategies.

For the "depletion" strategy, MPA locations are prioritized based in proportion to the spawning stock biomass (SSB) of the most depleted species. So, patches with high amounts of highly depleted species are prioritized over patches with high biomass of lightly depleted species.

For the "rate" strategy, MPAs are placed in based on the depletion-weighted catch relative to the total catch in a patch. So, patches where with high rates of catch of depleted species relative to totcal catch are prioritized.

For the "avoid_fishing" strategy, MPAs locations are prioritized inversely to fishery catches. So, patches with low catches are prioritized over patches with high catches.

For the "target_fishing" strategy, MPAs locations are prioritized proportionally to fishery catches. So, patches with high total catches are prioritized over patches with high catches.

For the "area" strategy, we simply assign patches in sequential order until the desired total MPA size is reached.

The information used to inform each MPA design strategy (except area) can include a multiplicative observation error term $\sigma_{obs}$ where

$$information \sim lnorm(data,\sigma_{obs}) $$

For example, for the depletion scenario, $\sigma_{obs}$ controls the accuracy of the spatial depletion data used to prioritize MPA placement.

## Case Study

In order to constrain the parameter space somewhat, we also ran a series of case study experiments. Rathere than interating over different habitat configurations and movement dynamics, we fix habitat for each species based on the results of @bradley2021XX, and set ontogenetic movement and seasonal migrations to FALSE.

### MPA Network Optimization

For the case study section, as a point of comparison we calculate an optimized MPA network along with the "rule of thumb" networks based on the design strategies outlined in section XX. The optimization routine is based on an objective function of the form

$$obj = \alpha \times scale(\sum_{s=1}^{S}\frac{ssb_s}{ssb0_s}) + (1 - \alpha) \times scale(\sum_{s=1}^{S}R_s) $$

where $\alpha$ controls the relative weight assigned to a "conservation" objective based on depletion, relative to an "economic" objective based on revenues. The 'scale' function rescales both objective components to the range of 0 to 1 to facilitate interpretation of $\alpha$ xx come up with a better version of this XX

The optimization routine itself works through a sampling-importance-resampling algorithm. In the first iteration of the algorithm, the marginal contribution of each individual patch to the objective function is calculated. The patch with the highest objective function value is then placed in an MPA. For the remaining iterations, we sample *X* cells from the remaining non-MPA cells with probability proportional to their marginal value to the objective function in the last iteration. We then calculate the marginal value of each of these cells when individually added to the MPA network begun in the prior step. The cell with the higheset marginal value is then added to the MPA network, and the marginal values of the non-selected but sampled cells are updated based on the values in that iteration. The process then continues until the desired MPA size is reached.

## Assessing Performance

Each model run started at equilibrium conditions and was then run for `r years` years (`r seasons * years` seasons. We assessed performance by comparing relevant metrics (e.g. depletion and revenue) under a given scenario with MPAs relative to the values of those same metrics in the simulated counterfactual world identical in every way save for the presencee of MPAs.

# Results

## Experiments: How severe and undere what conditions are unintended conservation consequences? (clearer name for these)

Make clear that a goal of all this complexity was to see if there were any clear emergent properties (this thing matters way more than this thing). We highlight some of those, but given the complexity we found we present the complexity of interacting ways in which different things can effect the conservation outcomes of MPAs

XX after looking through these should widdle it down to only things you actually examine, annoying / no point to having a bunch of unexplained minor variation in the mix relatede to things like ontogenetic shifts) xx

xx the problem here. These things all interact in so many complex ways that it's hard to look at either graphically or statistically, and maybe for the best. i.e. the answer isn't "if habitat overlap is less than XX% you might have a problem, otherwise you're fine", but rather that a bunch of things can interact to cause potential problems in complex ways. but that's hard to explain xx.

Big message: even constrained by the same initial depletion and core biology (e.g. movement rates), a wide range of outcomes are possible for each species for a given MPA size, dependent on the design strategy, habitat overlap, migrations, quality of information, etc. Across the board as MPA size increases the ratio of positive to negative outcomes increases, until all species achieve their maximum conservation potential in every iteration at 100% MPA. HOwever, every species experienced a set of simulations, roughyl xx%, that resulted in net losses in SSB/SSB0 relative to BAU. Losses were smaller than potential gaines, but reached meaningful levels in some species, and interestingly peaked in many cases right before max MPA (reflecting concentration of fishing in small amounts of habitat in habited by the species).Important to stress in here that there are no probabilities in this, no way of knowing which state is more probable than others, just given the range of things simulated more ways for a bigger positive effect. Also make clear though, biggish losses (10% or more) possible, but rare.

```{r exp-results, fig.cap="Results of simulation experiments on change in SSB/SSB0 (y-axis) as a function of MPA size (x-axis). Color indicates number of simulations in log-10 space that fell into a particular effect size at a given MPA size"}
tmp <- results %>%
  left_join(excols, by = "xid") %>%
  ggplot(aes(prop_mpa, biodiv - bau_biodiv)) +
  geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
  geom_bin2d() +
  facet_wrap( ~ critter) +
  scale_x_continuous(labels = scales::percent, name = "% MPA") +
  scale_y_continuous(name = "Change in SSB/SSB0") + 
  scale_fill_viridis_c(trans = "log10")

tmp

```

Moving to expressing things are percentage of species with some / significant loss

key result: Assuming perfect information, degree of habitat matters buts its messy. Both the depletion and target_fishing strategies are capable of completely avoiding losses if all species share the same habitat. As habitat non-overlap (xx better phrase) inceases, this gets harder to do, and both incur some losses. All of the other placement strategies exhhibit some losses, even with perfect overlap, at some MPA size.

```{r hab-effect, fig.cap="Precent of species with some degree of loss in SSB as a function of MPA size, MPA design strategy, and degree of habitat non-overlap (sigma_c), come up with a better name for this. Perfect information, no seasonal movement, all species considered in planning."}
fig_3 + 
  scale_y_continuous(labels = scales::percent, name = "% of Species with Loss in SSB/SSB0") +
  scale_color_discrete(name = "Habitat Non-Overlap") + 
  theme(legend.position = "top")
```

moving away from perfect information though, things get messier: now just looking at the same plot but with all the other options turned on things get messy quickly.

```{r mess-effect, fig.cap="Precent of species with some degree of loss in SSB as a function of MPA size, MPA design strategy, and degree of habitat non-overlap (sigma_c), come up with a better name for this. Rows indicate presence or absence of seasonal movement."}

tmp <- total_results %>%
  ggplot(aes(prop_mpa, loss, color = factor(sigma_centroid))) +
  geom_jitter(alpha = 0.1) +
  geom_smooth() +
  facet_grid(seasonal_movement ~ placement_strategy) + 
    scale_color_discrete(name = "Habitat Non-Overlap") + 
    scale_y_continuous(labels = scales::percent, name = "% of Species with Loss in SSB/SSB0")+
  theme(legend.position = "top")

tmp
```

same thing but now for SSB loss greater than or equal to 10%. Maybe should make all figure this? Basically, pretty common to cause some technical loss, but it's usually pretty marginal and effectively zero, so no benefit but not really much loss either. Significant losses can happen but less common, though throughout need to be really careful about being clear that this nothign about the probability of any one of these states of nature.

```{r signif-mess-effect, fig.cap="Percent of species with more than 10% loss in SSB/SSB0 as a function of MPA size, MPA design strategy, and degree of habitat non-overlap, come up with a better name for this. Rows indicate presence or absence of seasonal movement."}

tmp <- total_results %>%
  ggplot(aes(prop_mpa, signif_loss, color = factor(sigma_centroid))) +
  geom_jitter(alpha = 0.1) +
  geom_smooth() +
  facet_grid(seasonal_movement ~ placement_strategy) + 
    scale_color_discrete(name = "Habitat Non-Overlap") + 
      scale_y_continuous(labels = scales::percent, name = "% of Species with more than 10% Loss in SSB/SSB0")+

  theme(legend.position = "top")

tmp
```

Huh the seasonal migration thing is counterintuitive, but now that I think about it maybe makes sense, since with migrations every species gets some protection and some fishing, rather than some getting a lot of protection and others a lot more fishing.

My prior: that the magnitude of the benefits will be smaller with seasonal migration. So, you get a sort of tradeoff where on the plus side seasonal migrations means less change of unitended consequences, but less overall benefits as well

Yep that mostly checks.

```{r signif-mess-biodiv-effect, fig.cap="Percent of species with more than 10% loss in SSB/SSB0 as a function of MPA size, MPA design strategy, and degree of habitat non-overlap, come up with a better name for this. Rows indicate presence or absence of seasonal movement."}

tmp <- total_results %>%
  ggplot(aes(prop_mpa, delta_biodiv, color = factor(sigma_centroid))) +
  geom_jitter(alpha = 0.1) +
  geom_smooth() +
  facet_grid(seasonal_movement ~ placement_strategy) + 
    scale_color_discrete(name = "Habitat Non-Overlap") + 
      scale_y_continuous(name = "Total change in SSB/SSB0")+

  theme(legend.position = "top")

tmp

```

The other thing to consider is quality of information... weird results so far.

Poorer information decreases benefits in some cases...

```{r mess-biodiv-voi-effect, fig.cap="Precent of species with some degree of loss in SSB as a function of MPA size, MPA design strategy, degree of observation error. Rows indicate presence or absence of seasonal movement."}

tmp <- total_results %>%
  ggplot(aes(prop_mpa, delta_biodiv, color = factor(placement_error))) +
  geom_jitter(alpha = 0.1) +
  geom_smooth() +
  facet_grid(seasonal_movement ~ placement_strategy) + 
    scale_color_discrete(name = "Placement Error") + 
    scale_y_continuous(name = "Change in SSB/SSB0")+
  theme(legend.position = "top")

tmp
```

But has a confusing effect on unintended consequences. I guess makes some sense if unintended consequences are a consequence of very targeted interventions, and the greater observation error becomes, the close MPA placement gets to random, and the closer it gets to random, the less likely you are to start hammering one species while benefiting another, rather all just get some weaker protection? Kind of interesting if so.

```{r mess-loss-voi-effect, fig.cap="Precent of species with some degree of loss in SSB as a function of MPA size, MPA design strategy, degree of observation error. Rows indicate presence or absence of seasonal movement."}

tmp <- total_results %>%
  ggplot(aes(prop_mpa, loss, color = factor(placement_error))) +
  geom_jitter(alpha = 0.1) +
  geom_smooth() +
  facet_grid(seasonal_movement ~ placement_strategy) + 
    scale_color_discrete(name = "Placement Error") + 
    scale_y_continuous(labels = scales::percent, name = "% with loss in SSB/SSB0")+
  theme(legend.position = "top")

tmp
```

## Case Study

The general experiments conducted show two general insights

1. Even conditional on fishing exploitation and general life history strategy, the effects of MPAs on a group of species with technical interactions is highly variable and dependent on habitat distribution, design decisions, and data quality. xx make this more explicit.

2. Some degree of unintended conservation consequences occured across nearly all of our simulations below MPA coverage near 100%, though small losses were much more common than large. Different design strategies are better suited to avoiding unintended consequences.

There two general results beg some important questions. First, are unintended consequences inevitable given the kinds of technical interactions explored here, or can optimized MPA networks reduce or eliminate these unintended consequences. Second, one clear solution to the problem of unintended consequences is to simply make the MPAs large: if MPAs cover 100% of the area there can be no direct unintended conservation consequences. The potential for negative impacts on economic fishery outcomes increases as MPAs reach large size though. What then are the tradeoffs between reducing the potential for unintended conservation consequences and fishery revenues?

To address these questions, we ran a more limited "case study" set of simulations, based on the dynamics of the WCPO tuna fisheries. Rather than iterating over different movement dynamics and habitat configurations, we fix habitat layers on the values predicted by @bradley2021, and xx other things we fix xx.

For this subset of simulations then, along with the various "rule of thumb" MPA design strategies we ran an optimization routine (see section XX) across a range of objective function weightings, in the manner of @rassweiler2014. This allowed us to assess 1) the degree to which unintended conservation consequences of MPAs can be reduced by optimized design and 2) the magnitude of tradeoffs between conservation outcomes and fishery revenues under both rule-of-thumb and optimized MPA networks.

key results: optimization helps (Nobel prize please). Can reduce percent of species with conservation loss down to 25% pretty quickly if you prioritize conservation. Rule of thumb designs like targeting fishing can get there with perfect information and when considering all animals, but can be pretty far from optimal with imperfect information or when only considering a subset of animals in the planning process (that's what all the different points represent). Some rules of thumb are more sensitive to these decision than others. E.g. rate can get you optimal, but goes off the rails pretty quickly with bad information.

There is also a pretty clear breakpoint in results between basically trying for conservation or trying for economics, and not a lot of wiggle room in between (the two general bands of lines). xx check to what extent this depends on the way you scale the two objective function components xx.

```{r, fig.cap="Percent of species with some conservation loss as a function of MPA size. Points are draws from rule-of-thumb scenarios, lines are results of optimization routine across a gradient of conservation weighting."}
total_experiment_obj %>%
  filter(prop_mpa > 0) %>%
  ggplot(aes(prop_mpa, loss)) +
    geom_point() +
  geom_smooth(data = total_opt_obj, aes(p_protected / 100, loss, color = alpha, group = alpha), se = FALSE) +
  scale_x_continuous(labels = scales::percent, name = "Area in MPA") +
  scale_y_continuous(labels = scales::percent, name = "% Experiments With Conservation Loss") +
  facet_wrap(~placement_strategy) +
  scale_color_viridis_c(name = "Conservation Weighting") + 
  theme(legend.position = "top")
```

Key results: Even with optimization, you can basically either be in a world with 50% or more of species having some level of unintended conservation consequence, but no economic loss / some economic gain, or you can be in a world with lower unintended consequences, but economic losses. But, even rule of thumb approaches can get you to some places with relatively little unintended consequences at relatively minimal economic loss.

```{r, fig.cap="Tradeoff between unitendedn consequences and fisheries revenue. Filled polygons in background represent rule-of-thumb results, with redder colors indicating higher density of points. Points are results of optimization, with color indicating conservation weighting in optimization. Black diagonal line shows 1:1 relationship"}

cs_fig_11

```

Key results: Presence of and magnitude of unintended consequences pretty sensitive to design descisions. On a macro scale though far less so. In terms of change in total SSB/SSB0 (metric of biodiversity, i.e. maxed out when all species are unfished) and change in revenue, there is basically a pretty simple linear tradeoff between biodiversity and revenue, with a slope less than negative 1, meaning that 1% increase in biodiversity comes at a greater than 1% loss in fishery revenue. This stacks up pretty nicely with @rassweiler2014., where value of optimization was low when overfishing of valuable species isn't prevalent, as is the case here. So, design and optimization can help a lot with the more "micro" level outcomes (e.g. making sure that biodiversity gaines aren't coming from massive gains in one species to detriment of other), but on a macro level pretty insensitive in this case study.

```{r, fig.cap="Tradeoff frontier between revenue and biodiversity xx across different rule odf thumb design strategies. Points indicate role of thumb results, lines are results of optimized networks across a gradient of conservation weighting."}

cs_fig_10 + 
  theme(legend.position = "top")
```

# Discussion

**what is the story**

- Unintended conservation consequences of MPAs resulting from technical interactions have been discussed in the literature, but almost exclusively in the context of highly tactical modeling or localed empirics. We conduct one of the first large-scale assessments of the potential of this problem across a range of scenarios

- We find that these unintended consequences can be common outside of "canonical" MPA models *a la* hastings etc.

- But, good news is magnitude of unintended consequences were usually small, and total benefits almost always outweigh individual species losses (again no probability of states though)

- Some design strategies are better at avoiding unintended consequences than others, but sensitive to data quality. Optimization can help, but even with optimization unintended consequences can happen.

- Under the levels of fishing mortality in the WCPO, optimization and design can help reduce tradeoffs between fisheries revenue and unintended consequences than others, but macro-level tradeoffs between biodiversity and fishery revenue are very insensitive to design choices, with a pretty clear "increasing biodiversity through MPAs reduces fishery revenues". 
  -  xx This is clearly hyper-sensitive to the one thing we don't iterate over much here: how overfished are the target species. At the moment "they're not". But, this runs fast enough, is it worth running a hypothetical case study where they are just to illustrate how things change?

# Conclusions

- Unintended conservation outcomes of MPAs totally possible outside of "consider a homogenous fish blob" models

- But, usually small, and net conservation benefits outweigh conservation costs

- Should pay particular attention to potential though when species near tipping points are of concern: hard to imagine those not being part of the prioritization but something to consider. 
- Other interesting things.

# References

::: {#refs}
:::

# Scraps

Thinking through the “naive” MPA design strategies to compare against the optimized outcomes. Thoughts at the moment below. These can be run in a few seconds each, so we cab be be pretty liberal in coming up with combinations if there are other ideas people want to try, but also if some of these seem pointless can drop as well!
6:11
Design strategies
1. Assign protection by SSB / SSB0 weighted abundance in space (protect the highest concentrations of the most threatened species)
2. Assign protection by bycatch rates (distribution of biomass bycatch / biomass target)
3. Assign protection by targeting fishing effort
4. Assign protection by avoiding fishing effort
Knowledge states
1. Degrees of accuracy over the thing being measured (i.e. error in the design strategy)
2. Number of species considered (e.g. design aronnd only 2 bycatch species instead of all)
3. über naive: just close off random patches up to a specified size
We then plot these points in objective function space and compare to the efficiency frontier / consider frequency and magnitude of conservation and economic tradeoofs
The question all this is answering: how sensitive are conservation and economic outcomes to design strategy / uncertainty when there are technical interactions and heterogeneity in habitat? The first stage is teeing up the idea that beyond conservation / yield tradeoffs, as you increase habitat heterogeneity with technical interactions you increase the potential for conservation tradeoffs. But, the degree is going to be super context dependent, so to dive in a bit we then focus on this case study of the particular species / habitat overlap in the WCPO. Further work might reveal more generalities in this, but htis is basically getting a sense for the potential magnitude of the challenge.

XX stress a key difference: may other papers on this are static, in that
they just some up displacement on different layers, rather than
addressing dynamic respones over tim xx

We show....
  - sensitivity of both conservation and fishery outcomes to pelagic realism and design choices
  - compared to targeted / dynamic intervention... 
  - value of information would be great...

Contrast to coastal MPA theoretical work
  - @hastings2017
  - @ovando2021
  - @white2010
  - not quite sure what to do about points raised in @fulton2011a
    - to what extent has this been done...

The paragraph from @fulton2011a that broke me...

"
Spatial management was insufficient as the dominant management lever because it proves quite difficult to strike the balance between protecting areas that are of a size sufficient to produce any conservation benefit and providing sufficient fishing grounds to sustain an econom- ically viable industry (which is not possible in many fisheries if closures on the order of 70–80% or more are imposed, Fulton et al. 2007). More- over, Savina etal. (2008) found that effort displaced by closures potentially undermines any benefits coming from the zoning. For instance, strong improvements in diversity (by nearly 20%) under heavy fishing pressure occurred if effort outside a marine protected area (MPA) was controlled. However, without effort control, there was a significant (>20%) drop in overall diversity. Fulton et al. (2007), Savina et al. (2008), Smith et al. (2010) and unpublished work associated with Horne et al. (2010) all found that while spatial management appears to be an effective means of satisfying conservation objectives con- cerning habitat and restoring age-structures of less mobile species, benefits from closures for individual species depend on its trophic position, degree of mobility, geographical distribution and extent of ontogenetic shifts in habitat use and depth (highly mobile or prey species do not benefit in the way that more site attached or predatory species do)."

Lots of studies have shown that lots of things matter, can we provide guidance on which things are priorities?

Where does this fit in context of @smith2021? One important difference is that they point out differences in effectiveness, where effectiveness is still almost always positive, since they just measure leatherbacks. So their fig 5, there are a few cases where leatherback catches actually go up, but it's rare. So, extending to multiple species and keeping track off all of them shows it's even harder

How does this contrast with the the large and growing dynamic ocean management lit?
  - By definition, with perfect information dynamic $\geq$ static spatial controls
  - DOM literature is mostly static, tracks displacement onto static abundance / risk layers
  - Doesn't allow for dynamic impacts, for example fleet adjustments based on updated economic surface, or short-term costs for long-term benefits (can overstate costs?)
  - Dynamic model allows for delta between dynamic policies, e.g. Fmsy vs DOM
    - analytical models a la hastings have value but do / don't stand up to realism

Presentation of the model as a MICE model for MSP
  - reasonable parameterizable
  - flexible
  - fast

Focus of debate has been on fisheries and yields. As MPAs move to pelagic systems though, we show that conservation outcomes also depend on correctly identifying ecological state / design decisions etc.

What are the considerations of protection you need to make good
decisions. It's not always obvious how to design things to get the
outcome you want. If we could make the case that something like closing
the high seas is insufficient conservation management without
conservation management inside EEZs / what is the

Need to do an optimization here, calculate the marginal value of added
cells.

What's this about? Pointing out a problem, which to make impactful would
require making the case that there is a problem with current design
strategies. e.g.. "MPAs are often designed like this (simple dispersive
models with constant habitat, or MARXSAN style things), and doing that
does/does not create problems A tactical evaluation. Picking a handful
of high profile MPAs, parameterizing all the knowns as well as possible,
iterating over unknowns, and see what the range of outcomes are.
Cleaner, more impactful, but also opens it up (depending on the results)
to a details fight about very specific modeling choices How plausible
are losses? Maybe we drill down on this a bit. From a purely bycatch
perspective, the usual questions of fisheries tradeoffs aren't on the
table (but can add to the discussion considerations of how much target
fishing you give up). Given that then, while it's possible that
"targeted" interventions would work better (ignoring costs/enforcement,
etc), there's really no reason not to do MPAs. Unless, you actually end
up concentrating fishing pressure outside on populations. So, we
consider how plausible is it for that to really happen. Issue here is
that seems likely to be pretty hand-wavey. Most likely that the vast
majority of simulations will show positive impacts, but have no way of
assigning probabilities to any one state of nature.

The Role of Blue-Water Marine Protected Areas in Bycatch Reduction

What's the question: How effective are Blue-Water MPAs in reducing
fishing mortality of important pelagic bycatch species?

Another angle. There's going to be a lot of unknowns in designing BWMPA.
How much does not fully understanding a few key life history (e.g.
movement) matter, and how much does the role that the species plays in
the broader fleets portfolio matter

Could simulate habitat from a multivariate normal, where you use the
sigma to control how similar rasters are across species

Marine protected areas are increasingly proposed to protect biodiversity
in pelagic systems. However, maby paradigms of mpa design come from
coastal systems. Here we ask, how likely are blue-water MPAs to provide
conservation benefits to bycatch species, and what are the key
uncertainties underlying these outcomes? We begin with a general
discussion of the ways in which traits of open-ocean systems might
influence the conservation performance of BWMPAs. Given the massive
parameter space, we then constrain ourselves to a styulized version of
the western central pacitic ocean to ask, how feasible is it to provide
broad protection of bycatch species through common MPA design paradigms
in the context of this system. Our results demonstrate the critical
importance of XX and XX in designing MPAs in pelagic systems, and
highlight the potential risks of not accounting for XX. Introductionhere

Bycatch is a significant driver of population declines for many pelagic
species Sharks, seabirds, mammals, etc MPAs increasingly turned to as a
solution for conservation problems, etc. MPA science is largely based on
dynamics of coastal systems. Open-ocean ecosystems have very different
dynamics that require dedicated research Burgess et al. (2018) estimated
the reductions in fishing mortality needed to halt declines for a number
of bycatch species Here we ask, what are the key drivers of bycatch
conservation outcomes from MPAs? There are all of these species; you
reduce fishing mortality. You can either do fisheries management, or you
can do MPAs, let's focus in on the MPA side. This could reduce fishing
mortality, or it could do, and it depends on the dynamics of the species
and the fleets. When can MPAs reduce fishing mortality enough for
bycatch species? And where can you go wrong?

It doesn't make any sense to ask "do MPAs help bycatch": the question is
obviously it depends. To quickly make this case, we'll assume that a
species is in need of recovery. We then do the simple mako example to
show that effects of spatial closure depend on the overlap of habitat
and the assumption of fleet dynamics, and provide some empirical
evidence that in particular effort redistribution is not unlikely.

Determining the effect of an MPA network on bycatch then will also
depend on trying to get close to a tactical model. But, this creates a
problem for blue water systems in particular. Can make the argument that
given the centrality of relatively static habitat features in coastal
MPAs "protect the key habitats" might not be a bad strategy from a
conservation strategy. But in highly mobile pelagic systems there (may
be) a lot more variables in play. Where should researchers direct their
attention?

To reduce the parameter space, we constrain ourselves to a system where
1) we know the species involved 2) we have some sense of the relative
depletion and value of the different species and 3) some basic life
history of the species. We then vary commonly unknown parameters,
particularly preferred habitats, spatial catchability, movement rates
etc, and pair these simulations with different MPA designs, and examine
the conservation outcomes for bycatch species.

What i don't like about this

There isn't an "answer" to this: baring something like "changing the
type of density dependence never had any effect greater than XX%", for
anything that on occasion has an impact the answer of "how much" will
almost always depend on what else is going on.

So, it seems like there are two real angles of attack here

Pointing out a problem, which to make impactful would require making the
case that there is a problem with current design strategies. I.e. "MPAs
are often designed like this (simple dispersive models with constant
habitat, or MARXSAN style things), and doing that does/does not create
problems A tactical evaluation. Picking a handful of high profile MPAs,
parameterizing all the knownns as well as possible, iterating over
unknowns, and see what the range of outcomes are. Cleaner, more
impactful, but also opens it up (depending on the results) to a details
fight about very specific modeling choices

Methods

Collect life history for key bycatch species and target species, cite
Bradley et al. pulling together species distribution and threats etc.
Introduce marlin Parameterize marlin runs with "knowns" from life
history of bycatch and target species Highlight key "unknowns" that we
will iterate over, e.g. fleet models, density dependent movement, timing
of density dependence etc Simulate effects of MPAs given reasonably
realistic situations but iterating over key unknowns Evaluate the range
of outcomes and key drivers of MPA effects I'm thinking we try and avoid
a really specific case study but hard not to accidentally do one without
making everything up....

Results Range of likely changes in fishing-induced mortality for the
species selected Key drivers of uncertainty in those outcomes, focusing
on things that switch states from reduction to increase in mortality How
common are conservation tradeoffs between bycatch species? Fishery
impacts

Discussion What does this say about BWMPAs and bycatch? Not going to get
into design decisions here, depending how common tradeoffs are I think
that will be the main focus Highlight any species that frequently
induced tradeoffs Tee up future research MPAs in context of other
management MPAs as risk reduction strategy Etc.

### 2021-06-10 scripts/03_process_results.R

@@ -1,285 +1,0 @@
-source(file.path("scripts", "00_setup.R"))
-
-
-experiments <- read_rds(file = file.path(results_path, "experiements.rds"))
-
-critter_lookup <- read_rds(file = file.path(results_path, 'critter-lookup.rds'))
-
-rough_habitat <- read_rds(file = file.path(results_path, "rough-habitat.rds"))
-
-a = critter_lookup %>%
-  left_join(rough_habitat %>% select(species_commonname, bycatch) %>% unique)
-
-
-scenes <- experiments %>%
-  select(xid, data) %>%
-  unnest(cols = data) %>%
-  select(xid:seasonal_movement) %>%
-  unique()
-
-
-mpa_results <-  experiments %>%
-  select(xid, experiment) %>%
-  unnest(cols = experiment) %>%
-  select(xid,target_fauna, mpa_size, results) %>%
-  unnest(cols = results) %>%
-  mutate(target_fauna = map_chr(target_fauna, ~paste(.x, collapse = ","))) %>%
-  left_join(a, by = c("critter" = "species_sciname")) %>%
-  left_join(scenes, by = "xid") %>%
-  mutate(bycatch = ifelse(bycatch == TRUE, "Bycatch Species","Target Species"))
-
-mpa_strategy <- c(
-  `carcharhinus longimanus` = "MPAs for OWT Shark",
-  `prionace glauca,carcharhinus longimanus,isurus oxyrinchus,carcharhinus falciformis` = "MPAs for All Bycatch"
-)
-
-
-
-mpa_results %>%
-  filter(variable == "ssb") %>%
-  ggplot(aes(
-    mpa_size,
-    percent_change_ssb0,
-    color = species_commonname,
-    group = interaction(xid,species_commonname,bycatch)
-  )) +
-  geom_hline(aes(yintercept = 0)) +
-  geom_line(size = 1.5, alpha = 0.25) +
-  facet_grid( bycatch~ target_fauna, labeller = labeller(target_fauna = mpa_strategy), scales = "free_y") +
-  scale_x_continuous(name = "MPA Size", labels = scales::percent) +
-  scale_y_continuous(name = "% of SSB0 Change", labels = scales::percent) +
-  scale_color_discrete(name = "Species") +
-  scale_linetype(name = '') +
-  theme(axis.text = element_text(size = 8),
-        axis.title = element_text(size = 14),
-        legend.text = element_text(size = 12))
-
-
-
-mpa_strategy <- c(
[…]

### 2021-06-15 documents/03_marlin_paper.Rmd
recovered weird script loss, merging back to main

@@
 Perfect information, no seasonal movement, all species considered in planning."}
-fig_3 + 
-  scale_y_continuous(labels = scales::percent, name = "% of Species with Loss in SSB/SSB0") +
+
+tmp <- total_results %>%
+  filter(seasonal_movement == FALSE,
+         ontogenetic_shift == FALSE,
+         placement_error == min(placement_error),
+         prop_critters_considered == max(prop_critters_considered)) %>% 
+  ggplot(aes(prop_mpa, loss, color = factor(sigma_centroid))) +
+  geom_jitter() +
+  facet_wrap( ~ placement_strategy) + 
     scale_color_discrete(name = "Habitat Non-Overlap") + 
+    scale_y_continuous(labels = scales::percent, name = "% of Species with Loss in SSB/SSB0")+
   theme(legend.position = "top")
+
+tmp
+
 ```

### 2021-06-15 documents/marlin-paper-notes.Rmd
first real version of results
optimization branch now working for first full set of results (#3)
* testing on ubuntu

* first real version of results

* case study and experiments done

@@
 This might actually be a cool paper. Proc B?
+
+Idea: what ig instead of local or global, you have a parameter that says how many metapopulations there are.
+1 says global density dependence, patches says one per patch.
+you then divide up the patches in equal aquares based on that, and calculate density dependence on that level.
+
+## 2020-06-15 check in call
+
+Framing - MPAs and bycatch for large MPAs- what is the impact of MPAs on bycatch in the pelagic
+
+We don't have the right tools in the toolkit - better presentation of marlin - conventional models. We've packaged this up in a neat little package called marlin
+
+Hone in on what are the research questions
+
+- How big to get a big effect
+
+- How significant of a risk are unintended consequences
+
+- How do benefits and risk map onto economic tradeoffs
+
+- Potential for unintended consequences is real
+
+- But, since has to reach 0 at 100%, hard for effects to be that negative: either MPA is small so small effect, or MPA starts to get large enough to produce some protection too. 
+  - But that's clearly context dependent, wouldn't hold if you have a critter that only lives in one patch and you don't protect it
+
+- Some design paradigms are better than others but even good designs struggle as information quality goes down
+
+- Optimization can help, but can't remove tradeoffs
+
+- Unintended consequences are a real issue that need to be considered any time you have effort displacement and non-perfectly overlapping species, but problem isn't severe enough given reasonable parameters as an argument for doing nothing say
+
+- Add initial status as a thing to vary over - largely depends on their starting status
+
+- Build intuition over life history and
+
+Hi Dan! Thanks for sending the marlin paper outline. Just a few thoughts in advance of our meeting.
+I think the message can be much simpler than what is currently written.
+The key question is simply: can large offshore MPAs effectively protect species threatened by fishing?
+We can bring in stats on how overfishing drives extinction risk in many threatened species (e.g.
+sharks).
+Then we can make a case for what is different about the pelagic environment from the nearshore environment, and how models and empirics currently focus on the nearshore.
+We can then isolate the parameters of interest in the offshore...which brings me to my second framing point,
+There is a tension between 'we are just extending hastings et al.
+to explicitly relax assumptions made for nearshore environments to understand their implications on mpa impacts on pelagic bycatch' and 'here is a new method that is needed to explore this question, I.e.
+marlin).' We should discuss this and figure out which case we want to make or if there is a real compromise there.
+And finally,
+These decisions will help us decide which specific research questions we are ultimately interested in.
+Currently, because we don't have clear research questions, it is almost impossible to figure out how to best position our results (which is why I think the figures are a bit all over the place).
+I would suggest bullet-ing out the 3-5 most exciting findings from the work, and then we can use those to inform our conversation about the framing of the paper and presentation of results.
+
+LENNON
+
+Walk through some of the figures and unpack them a bit

### 2021-06-15 R/create-fauna.R

@@ -6,7 +6,8 @@ create_fauna <- function(scientific_name,
                          rec_form,
                          seasons,
                          init_explt,
-                         explt_type = "f"){
+                         explt_type = "f",
+                         recruit_habitat = NA){
 
   tmp <-  create_critter(
     scientific_name = scientific_name,
@@ -18,6 +19,7 @@ create_fauna <- function(scientific_name,
     seasons = seasons,
     init_explt = init_explt,
     explt_type = "explt_type",
+    recruit_habitat = recruit_habitat
   )
 
 

### 2021-06-15 scripts/00_setup.R

@@ -11,7 +11,7 @@ library(ggridges)
 
 library(googledrive)
 
-library(data.table)
+# library(data.table)
 
 library(gamm4)
 
@@ -38,15 +38,17 @@ if (!dir.exists(results_path)){
   dir.create(file.path(results_path,"sims"))
 }
 
-run_experiments <- TRUE
+run_experiments <- FALSE
 
-run_casestudy <- TRUE
+run_casestudy <- FALSE
+
+optimize_casestudy <- FALSE
 
 safety_stop <- FALSE
 
 draws <- 20
 
-resolution <- 20 # resolution is in squared patches, so 20 implies a 20X20 system, i.e. 400 patches
+resolution <- 10 # resolution is in squared patches, so 20 implies a 20X20 system, i.e. 400 patches
 
 years <- 40
 
@@ -63,9 +65,6 @@ time_steps <- seq(0,years - 1, by = time_step)
 theme_set(marlin::theme_marlin())
 
 tune_type <- "explt"
-
-plan(multicore, workers = workers)
-
 
 # load data ---------------------------------------------------------------
 

### 2021-06-27 R/create_experiment_critters.R
passing to squid

@@ -1,0 +1,85 @@
+create_experiment_critters <-
+  function(sciname,
+           habitat,
+           seasons = 1,
+           marlin_inputs,
+           adult_movement_sigma = 10,
+           recruit_movement_sigma = 20,
+           seasonal_movement = FALSE,
+           ontogenetic_shift = FALSE,
+           f_v_m,
+           rec_form) {
+    # sciname <- marlin_inputs$scientific_name[[1]]
+    #
+    #
+
+    # habitat <- experiments$habitat[[1]]
+    tmp_inputs <- marlin_inputs %>%
+      filter(scientific_name == sciname)
+
+    hab <- habitat %>%
+      pivot_wider(names_from = y, values_from = habitat) %>%
+      select(-x) %>%
+      as.matrix()
+
+    # seasonal_movement <- sample(c(1, 0), 1, replace = TRUE)
+
+    # uniform_rec_hab <- sample(c(1, 0), 1, replace = TRUE)
+
+
+    # seasonal_movement <- 0
+    if (seasonal_movement) {
+      hab <- list(hab,-hab - (min(-hab)))
+
+    } else {
+      hab <- list(hab, hab)
+    }
+
+    # if (random_rec){
+    #   rec_form <-  sample(c(0, 1), 1, replace = TRUE)
+    #
+    # } else {
+    #   rec_form <- 2
+    # }
+    # rec_form <- 1 # sample(c(0, 1, 2, 3), 1, replace = TRUE)
+
+
+    if (ontogenetic_shift) {
+      recruit_habitat <-
+        -hab[[1]] - min(-hab[[1]]) # place recruits in different places than adults
+
+      # set recruitment form to allow for recruit habitat
+      #
+      rec_form <- 0
+    } else {
+      recruit_habitat <- hab[[1]]
+    }
+
+    critter <- marlin::create_critter(
+      scientific_name = sciname,
+      seasonal_habitat = hab,
+      recruit_habitat = recruit_habitat,
+      adult_movement = 0,
+      adult_movement_sigma = adult_movement_sigma,
+      recruit_movement_sigma = recruit_movement_sigma,
+      rec_form = rec_form,
+      fec_form = ifelse(str_detect(sciname,"carcharhinus"),"pups","weight"),
+      pups = 6,
[…]

### 2021-06-27 scripts/00_setup.R

@@ -21,6 +21,8 @@ library(mvtnorm)
 
 library(patchwork)
 
+library(progress)
+
 options(dplyr.summarise.inform = FALSE)
 
 
@@ -28,7 +30,7 @@ foos <- list.files(here("R"))
 
 walk(foos, ~ source(here("R", .x)))
 
-results_name <- "v1.0"
+results_name <- "test"
 
 results_path <- here("results", results_name)
 
@@ -38,7 +40,7 @@ if (!dir.exists(results_path)){
   dir.create(file.path(results_path,"sims"))
 }
 
-run_experiments <- FALSE
+run_experiments <- TRUE
 
 run_casestudy <- FALSE
 
@@ -64,7 +66,7 @@ time_steps <- seq(0,years - 1, by = time_step)
 
 theme_set(marlin::theme_marlin())
 
-tune_type <- "explt"
+tune_type <- "v1.0"
 
 # load data ---------------------------------------------------------------
 

### 2021-06-27 scripts/01_run_experiments.R

@@ -1,21 +1,22 @@
 source(file.path("scripts", "00_setup.R"))
 
 
-# for now let's leave the life history stuff out of it... and then once you have a sense of timing for this decide on how many iterations to do of each
-experiments <-
-  expand_grid(
-    sigma_centroid = seq(.25 * resolution, resolution ^ 2 / 2, length.out = 5),
-    sigma_hab = c(20, 5),
-    ontogenetic_shift = c(TRUE, FALSE),
-    seasonal_movement = c(TRUE, FALSE)
+# generate state experiments. This is a somewhat tricky process where the actual generated values are created for many variables in create_experiment_critters
+
+n_states <- 250
+
+state_experiments <-
+  tibble(
+    sigma_centroid = runif(n_states,.1 * resolution, resolution ^ 2 / 2),
+    sigma_hab = runif(n_states,.1 * resolution, resolution ^ 2 / 8)
   ) %>%
-  mutate(xid = 1:nrow(.))
+  mutate(state_id = 1:nrow(.))
 
 
 # function to assign habitat for each species based on
 
 critters <-
-  tibble(scientific_name = unique(marlin_inputs$scientific_name)) %>%
+  tibble(scientific_name = c("katsuwonus pelamis","kajikia audax","carcharhinus longimanus")) %>%
   mutate(centroid = NA)
 
 
@@ -90,7 +91,7 @@ create_critter_habitats <-
   }
 
 
-experiments <- experiments %>%
+state_experiments <- state_experiments %>%
   mutate(
     habitats = map2(
       sigma_centroid,
@@ -100,23 +101,32 @@ experiments <- experiments %>%
       resolution = resolution
     )
   ) %>%
-  unnest(cols = habitats)
+  unnest(cols = habitats) %>%
+  mutate(seasonal_movement = sample(c(TRUE, FALSE), nrow(.), replace = TRUE),
+         f_v_m = runif(nrow(.), .1,3),
+         adult_movement_sigma = runif(nrow(.), min = 0.05 * resolution, max = 3 * resolution),
+         recruit_movement_sigma = runif(nrow(.), min = 0.05 * resolution, max = 3 * resolution),
+         rec_form = sample(c(0,1,2,3), nrow(.), replace = TRUE)
+         )
 
-experiments$habitat[[3]] %>%
+state_experiments$habitat[[sample(1:n_states,1)]] %>%
   ggplot(aes(x, y, fill = habitat)) +
[…]

### 2021-07-12 documents/03_marlin_paper.Rmd
passing to squid
checked that revamped version works

@@
 
+** something to piont out: increased catches are a paradox, either come from rebuilding stock or concentrated fishing, but trajectory can tell you something.
+not possible for MPAs to increase catches without decreasing catches first if they are rebuilding?
+**
 # Introduction
+
+Empirically, how do you differentiate increase in catch from fishing harder vs increases in catch from rebuilding: rebuilding has to have a decrease at first
 

### 2021-07-12 R/compile_experiment_fleet.R

@@ -5,7 +5,7 @@ compile_experiment_fleet <- function(fauna, tune_type = "explt") {
         critter = fauna$`katsuwonus pelamis`,
         price = 1.36,
         sel_form = "logistic",
-        sel_start = fauna$`katsuwonus pelamis`$length_50_mature *  sample(c(0.25, 1, 1.75), 1, replace = TRUE),
+        sel_start = fauna$`katsuwonus pelamis`$length_50_mature *  sample(c(0.8), 1, replace = TRUE),
         sel_delta = 0.01,
         catchability = .1,
         p_explt = 1,
@@ -16,7 +16,7 @@ compile_experiment_fleet <- function(fauna, tune_type = "explt") {
         critter = fauna$`kajikia audax`,
         price = 5.16,
         sel_form = "logistic",
-        sel_start = fauna$`kajikia audax`$length_50_mature *  sample(c(0.25, 1, 1.75), 1, replace = TRUE),
+        sel_start = fauna$`kajikia audax`$length_50_mature *  sample(c(0.8), 1, replace = TRUE),
         sel_delta = 0.01,
         catchability = .1,
         p_explt = 1,
@@ -28,7 +28,7 @@ compile_experiment_fleet <- function(fauna, tune_type = "explt") {
         price = 1.89,
         sel_form = "logistic",
         sel_start = fauna$`carcharhinus longimanus`$length_50_mature *
-          sample(c(0.25, 1, 1.75), 1, replace = TRUE),
+          sample(c(0.8), 1, replace = TRUE),
         sel_delta = 0.01,
         catchability = .1,
         p_explt = 1,

### 2021-07-12 scripts/02_run_case_study.R

@@ -116,7 +116,8 @@ casestudy <- casestudy %>%
       sciname = scientific_name,
       habitat = habitat,
       ontogenetic_shift = FALSE,
-      seasonal_movement = FALSE
+      seasonal_movement = FALSE,
+      random_rec = TRUE
     ),
     create_critters,
     marlin_inputs = marlin_inputs,
@@ -211,9 +212,10 @@ if (optimize_casestudy == TRUE){
 
   fleets <- casestudy$fleet[[1]]
 
+  a <- Sys.time()
   optimized_networks <- optimized_networks %>%
     mutate(
-      network = purrr::pmap(
+      network = pmap(
         list(alpha = alpha),
         optimize_mpa,
         fauna = casestudy$fauna[[1]],
@@ -222,8 +224,10 @@ if (optimize_casestudy == TRUE){
         prop_sampled = 0.2,
         starting_conditions = starting_conditions,
         resolution = resolution,
-      workers = 8
+        workers = 8
       ))
+
+  Sys.time() - a
 
   write_rds(optimized_networks, file = file.path(results_path,"optimized_networks.rds"))
 

### 2021-07-14 R/compile_experiment_fleet.R
first real cut of experiments

@@ -5,7 +5,7 @@ compile_experiment_fleet <- function(fauna, tune_type = "explt") {
         critter = fauna$`katsuwonus pelamis`,
         price = 1.36,
         sel_form = "logistic",
-        sel_start = fauna$`katsuwonus pelamis`$length_50_mature *  sample(c(0.8), 1, replace = TRUE),
+        sel_start = fauna$`katsuwonus pelamis`$length_50_mature *  sample(c(0.6), 1, replace = TRUE),
         sel_delta = 0.01,
         catchability = .1,
         p_explt = 1,
@@ -16,7 +16,7 @@ compile_experiment_fleet <- function(fauna, tune_type = "explt") {
         critter = fauna$`kajikia audax`,
         price = 5.16,
         sel_form = "logistic",
-        sel_start = fauna$`kajikia audax`$length_50_mature *  sample(c(0.8), 1, replace = TRUE),
+        sel_start = fauna$`kajikia audax`$length_50_mature *  sample(c(0.7), 1, replace = TRUE),
         sel_delta = 0.01,
         catchability = .1,
         p_explt = 1,
@@ -28,7 +28,7 @@ compile_experiment_fleet <- function(fauna, tune_type = "explt") {
         price = 1.89,
         sel_form = "logistic",
         sel_start = fauna$`carcharhinus longimanus`$length_50_mature *
-          sample(c(0.8), 1, replace = TRUE),
+          sample(c(0.9), 1, replace = TRUE),
         sel_delta = 0.01,
         catchability = .1,
         p_explt = 1,

### 2021-07-14 R/optimize-mpa.R

@@ -7,6 +7,8 @@
 #' @param resolution the resolution of the simulated system
 #' @param prop_sampled the proportion of cells to sample in each SIR iteration
 #' @param workers number of workers for parallel process
+#' @param starting_conditions starting conditions of the simmar object
+#' @param objective one of max_ssb to maximize spawning stock biomass or min_loss to prioritize not having any losses
 #'
 #' @return a list with results of MPA optimization
 #' @export
@@ -19,7 +21,8 @@ optimize_mpa <-
            max_prop_mpa = 1,
            resolution,
            prop_sampled = .2,
-           workers = 6) {
+           workers = 6,
+           objective = "max_ssb") {
     # workers <- 6
     future::plan(future::multisession, workers = workers)
 
@@ -74,8 +77,22 @@ optimize_mpa <-
         res <-
           sim_mpa[[length(sim_mpa)]] # for now, just calculate in the final timestep
 
+        if (objective == "max_ssb"){
         biodiv <-
           sum(map_dbl(res, ~ sum(.x$ssb_p_a) / .x$ssb0)) # calculate biodiversity component of objective function
+        } else if (objective == "min_loss") {
+
+        biodiv_mpa <-
+          (map_dbl(res, ~ sum(.x$ssb_p_a) / .x$ssb0)) # calculate biodiversity component of objective function
+
+        biodiv_sq <-
+          map_dbl(starting_conditions[[1]], ~ sum(.x$ssb_p_a) / .x$ssb0) # calculate biodiversity component of objective function
+
+
+        biodiv <- sum((biodiv_mpa - biodiv_sq) >= 0)
+
+        }
+
 
         # econ <- sum(map_dbl(res, ~sum(.x$c_p_a))) #  calculate econ component of objective function
 
@@ -109,7 +126,7 @@ optimize_mpa <-
       marginal_values$patch <- candidate_patches # assign patches
 
       marginal_values$obj_value <-
-        alpha * scales::rescale(marginal_values$biodiv) + (1 - alpha) * scales::rescale(marginal_values$econ) # calculate objective function
+        alpha * scales::rescale(marginal_values$biodiv) + (1 - alpha) * scales::rescale(marginal_values$econ) # calculate objective function. Rescaling means that alpha dictates the weithing of a given percent rank of biodiversity relative to a relative percent rank of economics
 
       marginal_values <- marginal_values %>%
[…]

### 2021-07-14 scripts/01_run_experiments.R

@@ -3,283 +3,326 @@ source(file.path("scripts", "00_setup.R"))
 
 # generate state experiments. This is a somewhat tricky process where the actual generated values are created for many variables in create_experiment_critters
 
-n_states <- 250
+n_states <- 500
 
-state_experiments <-
-  tibble(
-    sigma_centroid = runif(n_states,.1 * resolution, resolution ^ 2 / 2),
-    sigma_hab = runif(n_states,.1 * resolution, resolution ^ 2 / 8)
-  ) %>%
-  mutate(state_id = 1:nrow(.))
-
-
-# function to assign habitat for each species based on
-
-critters <-
-  tibble(scientific_name = c("katsuwonus pelamis","kajikia audax","carcharhinus longimanus")) %>%
-  mutate(centroid = NA)
+if (run_experiments) {
+  state_experiments <-
+    tibble(
+      sigma_centroid = runif(n_states, .1 * resolution, resolution ^ 2 / 2),
+      sigma_hab = runif(n_states, .1 * resolution, resolution ^ 2 / 8)
+    ) %>%
+    mutate(state_id = 1:nrow(.))
 
 
-create_critter_habitats <-
-  function(sigma_centroid,
-           sigma_hab = 0.2,
-           base_centroid = c(resolution / 2, resolution / 2) ,
-           critters,
-           resolution) {
-    # base_centroid <- c(10,10)
-    # sigma_hab = .2
+  # function to assign habitat for each species based on
 
-    # sigma_block <- 0.1
-
-    # bycatch_corr <- -1
-
-
-    base_layer <-
-      tidyr::expand_grid(x = 1:resolution, y = 1:resolution)
+  critters <-
+    tibble(
+      scientific_name = c(
+        "katsuwonus pelamis",
+        "kajikia audax",
+        "carcharhinus longimanus"
+      )
+    ) %>%
+    mutate(centroid = NA)
 
 
-    centroid_index <-
-      which(base_layer$x == base_centroid[1] &
-              base_layer$y == base_centroid[2])
+  create_critter_habitats <-
+    function(sigma_centroid,
+             sigma_hab = 0.2,
+             base_centroid = c(resolution / 2, resolution / 2) ,
+             critters,
+             resolution) {
+      # base_centroid <- c(10,10)
+      # sigma_hab = .2
+
+      # sigma_block <- 0.1
+
+      # bycatch_corr <- -1
 
 
[…]

### 2021-07-22 documents/03_marlin_paper.Rmd
first real cut of experiments
Retool experiments (#4)
* redesigned experiments around simpler structure

* first real cut of experiments

--- 2021-06-10 documents/03_marlin_paper.Rmd
@@
 params:
-  run_name: ["v1.0"]
+  run_name: ["v0.5"]
 ---
@@
 
-assign(str_remove(rdses[8],".rds"), read_rds(file.path(results_path,rdses[2])))
-
 walk(rdses,~assign(str_remove(.x,".rds"), read_rds(file.path(results_path,.x)), envir = .GlobalEnv)) #load and assign all .rds objects in the appropriate results folder
@@
 
-# a bit of housekeeping to avoid saving a massive object
-tmp <- raw_experiment_results  %>%
-  map( ~ .x$obj)
-
 group_var <- "placement_strategy"
 
-total_results <- total_experiment_results
-
-rm(total_experiment_results)
-
-results <- placement_experiment %>%
+results <- placement_experiments %>%
   mutate(temp = raw_experiment_results) %>%
-  group_by_at(colnames(.)[!colnames(.) %in% c("temp", "prop_mpa")]) %>%
-  nest() %>%
-  ungroup() %>%
-  mutate(placement_id = 1:nrow(.)) %>%
-  unnest(cols = data) %>%
-  mutate(xid = map(raw_experiment_results, names)) %>%
-  unnest(cols = c(temp, xid)) %>%
+  mutate(state_id = map(raw_experiment_results, names)) %>%
+  unnest(cols = c(temp, state_id)) %>%
   mutate(obj = map(temp, "obj"),
-         xid = as.numeric(xid)) %>%
-  unnest(cols = obj)
+         state_id = as.numeric(state_id)) %>%
+  unnest(cols = obj) %>%
+  mutate(
+    critter = fct_recode(
+      critter,
+      "Shark" = "carcharhinus longimanus",
+      "Tuna" = "katsuwonus pelamis",
+      "Marlin" = "kajikia audax"
+    )
+  )
 
@@
 
-warning("move placement experiement id back above run_experiments")
-
-# total_experiment <- placement_experiment %>%
-#   left_join(experiments, by = 'xid')
-
-# results <- results %>%
-#   left_join(placement_experiment)
-
 # compare to BAU
@@
 results <- results %>%
-  group_by(placement_id, xid, critter) %>%
+  group_by(placement_id, state_id, critter) %>%
   mutate(bau_biodiv = biodiv[prop_mpa == 0],
@@
 excols <-
-  experiments %>%
-  select(xid, data) %>%
+  state_experiments %>%
+  select(state_id, data) %>%
   unnest(cols = data) %>%
-  select(-habitat,-critter, -scientific_name, -centroid) %>%
+  select(state_id:sigma_hab) %>%
   ungroup() %>%
@@
 
+total_results <- results %>%
+  group_by(
+    placement_id,
+    placement_strategy,
+    prop_critters_considered,
+    placement_error,
+    state_id,
+    prop_mpa
+  ) %>%
+  summarise(
+    loss = mean(biodiv < bau_biodiv),
+    n_loss = sum(biodiv < bau_biodiv),
+    signif_loss = mean(biodiv < (0.9 * bau_biodiv)),
+    total_loss = sum(biodiv - bau_biodiv),
+    econ = sum(econ),
+    biodiv = sum(biodiv),
+    bau_econ = sum(bau_econ),
+    bau_biodiv = sum(bau_biodiv)
+  ) %>%
+  ungroup() %>%
+  mutate(delta_biodiv = biodiv / bau_biodiv - 1,
+         abs_delta_biodiv = biodiv - bau_biodiv,
+         delta_econ = econ / bau_econ - 1) %>%
+  left_join(excols, by = "state_id")
 ```
 
+** something to point out: increased catches are a paradox, either come from rebuilding stock or concentrated fishing, but trajectory can tell you something.
+not possible for MPAs to increase catches without decreasing catches first if they are rebuilding?
+**
+
+Seasonal movemetn is kind of interesting, basically just dilutes the extremes
+
+less resilient species are both more in need of protection and easier to accidentally fish harder
+
 # Introduction
+
+Empirically, how do you differentiate increase in catch from fishing harder vs increases in catch from rebuilding: rebuilding has to have a decrease at first
 
@@
 This problem of policy "leakage" is well known in many environmental fields [@citations] but the general properties of this problem have received relatively little attention in the theoretical or empirical MPA literature [see @abbott2012;@savina2008;fulton2011a;@baum2003 for notable exceptions].
-We fill this gap by
+We fill this gap by examining the general range of conditions...
 
@@
 
-@hastings2017 employed a simply model to derive conditions under which
+@hastings2017 employed a simple model to derive conditions under which
 MPAs would benefit bycatch species. However they also pointed out that
@@
 specifically around the role of technical interactions across species.
-Many MPA models simulate the impacts of MPAs on a single-species level [@ovando2021XX].
+Many MPA models simulate the impacts of MPAs on a single-species level [@2021].
 This can be useful for for example examining the ways in which species-specific life history traits interact with fleet dynamics to influence MPA outcomes.
@@
 
-Given the complexity of the operating model used here, a complete sensitivity analysis across all parameters is not practical.
-Instead, we focus on the role of XX key attributes of pelagic ecosystems: the degree of spatial overlap between populations, the presence of seasonal migrations, and the potential for ontogenetic shifts in habitat use, and degree of fishing pressure on each species.
+Given the complexity of the operating model used here, a complete sensitivity analysis across all parameters is not practical. Instead, we focus on the role of life history, fishing pressure,
 
-## Species & Fleets
+### Species & Fleets
 
@@
 Where $d$ is the euclidian distance from the $centroid_s$ to patch $p$ and $\sigma_h$ controls the dispersal of the habitat around that range.
-
-### Ontogenetic Shifts
-
-`marlin` allows for different habitats to be used by adults vs.
-recruits.
-For scenarios with ontogenetic shifts ,we assign recruits to a habitat grid that is the inverse of $habitat_{p,s}$.
-In this way organisms enter the model living on one area, and then transition to adult habitat at older life stages.
-
-XX this isn't very interesting at this point since movement is so high so I vote we get rid of it for now XX.
 
@@
 
-## Experiments: How severe and undere what conditions are unintended conservation consequences? (clearer name for these)
+## Simulation Experiments
+
+We focused first on the range of impacts that MPAs can have on mature biomass across the range of variables explored.
+To illustrate, we pulled out two example simulations, one where MPAs provide relatively large benefits across all three species, and a second where MPAs actually produced substantial losses in some species....
+
+```{r fig-info, fig.cap="Illustration of ways in which variables can interact to affect MPA outcomes."}
+
+```
+
+Looking at the full `r n_distinct(total_results$placement_id) * n_distinct(total_results$state_id)` experiments run, the results were highly variable, with MPAs capable of producing substantial conservation gains to substantial conservation losses and everything in between.
+We do not focus into the individual effects of particular variables as we found them to be highly dependent on the exact question asked and the set of data included, and since we cannot ascribe any actual likelihood to any individual state of nature simulated here.
+
+However, three important features do emerge.
+The first is that life history plays an important role in the potential for unintended conservation consequences from MPAs.
+All else equal, MPAs were much less likely to produce losses in spawning biomass for our tuna-like species than the marlin or shark-like species.
+Given that within the broad life history traits we do vary both adult and larval movement, as well as the timing of density dependence, we believe this result is mostly due to the high resilience (high steepness, low age at maturity, fast growth) of the tuna-like species relative to the marlin and shark-like species.
+
+The second is that design matters...
+
+The third is that extreme outcomes (both positive and negative), particularly at lower MPA sizes, are largely driven by the presence of small range sizes interacting with seasonal migrations.
+Across species simulations without seasonal migrations and small ranges were much more likely to produce extreme results than those with seasonal migrations and/or larger ranges.
+Presence of seasonal migrations acts as a dilutant: As a given static MPA only provides protection, or displaces fishing effort, onto a particular area in only half of the year, both the benefits and costs of MPAs are smaller.
+When seasonal migrations are not present and the species range is small, MPAs are more capable of providing either large benefits or large losses.
 
@@
 Also make clear though, biggish losses (10% or more) possible, but rare.
-
-```{r exp-results, fig.cap="Results of simulation experiments on change in SSB/SSB0 (y-axis) as a function of MPA size (x-axis).
-Color indicates number of simulations in log-10 space that fell into a particular effect size at a given MPA size"}
-tmp <- results %>%
-  left_join(excols, by = "xid") %>%
-  ggplot(aes(prop_mpa, biodiv - bau_biodiv)) +
-  geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
-  geom_bin2d() +
-  facet_wrap( ~ critter) +
-  scale_x_continuous(labels = scales::percent, name = "% MPA") +
-  scale_y_continuous(name = "Change in SSB/SSB0") + 
-  scale_fill_viridis_c(trans = "log10")
-
-tmp
-
-```
-
-Moving to expressing things are percentage of species with some / significant loss
-
-key result: Assuming perfect information, degree of habitat matters buts its messy.
-Both the depletion and target_fishing strategies are capable of completely avoiding losses if all species share the same habitat.
-As habitat non-overlap (xx better phrase) inceases, this gets harder to do, and both incur some losses.
-All of the other placement strategies exhhibit some losses, even with perfect overlap, at some MPA size.
-
-```{r hab-effect, fig.cap="Precent of species with some degree of loss in SSB as a function of MPA size, MPA design strategy, and degree of habitat non-overlap (sigma_c), come up with a better name for this.
-Perfect information, no seasonal movement, all species considered in planning."}
-fig_3 + 
-  scale_y_continuous(labels = scales::percent, name = "% of Species with Loss in SSB/SSB0") +
-  scale_color_discrete(name = "Habitat Non-Overlap") + 
-  theme(legend.position = "top")
-```
-
-moving away from perfect information though, things get messier: now just looking at the same plot but with all the other options turned on things get messy quickly.
-
-```{r mess-effect, fig.cap="Precent of species with some degree of loss in SSB as a function of MPA size, MPA design strategy, and degree of habitat non-overlap (sigma_c), come up with a better name for this.
-Rows indicate presence or absence of seasonal movement."}
-
-tmp <- total_results %>%
-  ggplot(aes(prop_mpa, loss, color = factor(sigma_centroid))) +
-  geom_jitter(alpha = 0.1) +
-  geom_smooth() +
-  facet_grid(seasonal_movement ~ placement_strategy) + 
-    scale_color_discrete(name = "Habitat Non-Overlap") + 
-    scale_y_continuous(labels = scales::percent, name = "% of Species with Loss in SSB/SSB0")+
-  theme(legend.position = "top")
-
-tmp
-```
-
-same thing but now for SSB loss greater than or equal to 10%.
-Maybe should make all figure this?
-Basically, pretty common to cause some technical loss, but it's usually pretty marginal and effectively zero, so no benefit but not really much loss either.
-Significant losses can happen but less common, though throughout need to be really careful about being clear that this nothign about the probability of any one of these states of nature.
-
-```{r signif-mess-effect, fig.cap="Percent of species with more than 10% loss in SSB/SSB0 as a function of MPA size, MPA design strategy, and degree of habitat non-overlap, come up with a better name for this.
-Rows indicate presence or absence of seasonal movement."}
-
-tmp <- total_results %>%
-  ggplot(aes(prop_mpa, signif_loss, color = factor(sigma_centroid))) +
-  geom_jitter(alpha = 0.1) +
-  geom_smooth() +
-  facet_grid(seasonal_movement ~ placement_strategy) + 
-    scale_color_discrete(name = "Habitat Non-Overlap") + 
-      scale_y_continuous(labels = scales::percent, name = "% of Species with more than 10% Loss in SSB/SSB0")+
-
-  theme(legend.position = "top")
-
-tmp
-```
-
-Huh the seasonal migration thing is counterintuitive, but now that I think about it maybe makes sense, since with migrations every species gets some protection and some fishing, rather than some getting a lot of protection and others a lot more fishing.
-
-My prior: that the magnitude of the benefits will be smaller with seasonal migration.
-So, you get a sort of tradeoff where on the plus side seasonal migrations means less change of unitended consequences, but less overall benefits as well
-
-Yep that mostly checks.
-
-```{r signif-mess-biodiv-effect, fig.cap="Percent of species with more than 10% loss in SSB/SSB0 as a function of MPA size, MPA design strategy, and degree of habitat non-overlap, come up with a better name for this.
-Rows indicate presence or absence of seasonal movement."}
-
-tmp <- total_results %>%
-  ggplot(aes(prop_mpa, delta_biodiv, color = factor(sigma_centroid))) +
-  geom_jitter(alpha = 0.1) +
-  geom_smooth() +
-  facet_grid(seasonal_movement ~ placement_strategy) + 
-    scale_color_discrete(name = "Habitat Non-Overlap") + 
-      scale_y_continuous(name = "Total change in SSB/SSB0")+
-
-  theme(legend.position = "top")
-
-tmp
-
-```
-
-The other thing to consider is quality of information... weird results so far.
-
-Poorer information decreases benefits in some cases...
-
-```{r mess-biodiv-voi-effect, fig.cap="Precent of species with some degree of loss in SSB as a function of MPA size, MPA design strategy, degree of observation error.
-Rows indicate presence or absence of seasonal movement."}
-
-tmp <- total_results %>%
-  ggplot(aes(prop_mpa, delta_biodiv, color = factor(placement_error))) +
-  geom_jitter(alpha = 0.1) +
-  geom_smooth() +
-  facet_grid(seasonal_movement ~ placement_strategy) + 
-    scale_color_discrete(name = "Placement Error") + 
-    scale_y_continuous(name = "Change in SSB/SSB0")+
-  theme(legend.position = "top")
-
-tmp
-```
-
-But has a confusing effect on unintended consequences.
-I guess makes some sense if unintended consequences are a consequence of very targeted interventions, and the greater observation error becomes, the close MPA placement gets to random, and the closer it gets to random, the less likely you are to start hammering one species while benefiting another, rather all just get some weaker protection?
-Kind of interesting if so.
-
-```{r mess-loss-voi-effect, fig.cap="Precent of species with some degree of loss in SSB as a function of MPA size, MPA design strategy, degree of observation error.
-Rows indicate presence or absence of seasonal movement."}
-
-tmp <- total_results %>%
-  ggplot(aes(prop_mpa, loss, color = factor(placement_error))) +
-  geom_jitter(alpha = 0.1) +
-  geom_smooth() +
-  facet_grid(seasonal_movement ~ placement_strategy) + 
-    scale_color_discrete(name = "Placement Error") + 
-    scale_y_continuous(labels = scales::percent, name = "% with loss in SSB/SSB0")+
-  theme(legend.position = "top")
-
-tmp
-```
 
@@
 Points are draws from rule-of-thumb scenarios, lines are results of optimization routine across a gradient of conservation weighting."}
-total_experiment_obj %>%
-  filter(prop_mpa > 0) %>%
-  ggplot(aes(prop_mpa, loss)) +
-    geom_point() +
-  geom_smooth(data = total_opt_obj, aes(p_protected / 100, loss, color = alpha, group = alpha), se = FALSE) +
-  scale_x_continuous(labels = scales::percent, name = "Area in MPA") +
-  scale_y_continuous(labels = scales::percent, name = "% Experiments With Conservation Loss") +
-  facet_wrap(~placement_strategy) +
-  scale_color_viridis_c(name = "Conservation Weighting") + 
-  theme(legend.position = "top")
+
 ```
@@
 Black diagonal line shows 1:1 relationship"}
-
-cs_fig_11
 
@@
 
-cs_fig_10 + 
-  theme(legend.position = "top")
 ```

### 2021-09-15 documents/03_marlin_paper.Rmd
pre visual commit

@@
 
-Empirically, how do you differentiate increase in catch from fishing harder vs increases in catch from rebuilding: rebuilding has to have a decrease at first
+<!-- Empirically, how do you differentiate increase in catch from fishing harder vs increases in catch from rebuilding: rebuilding has to have a decrease at first -->
 
 Marine protected areas (MPAs) are a growing part of the marine
-conservation toolbox, with 7.4% of the world's oceans currently
-protected and, and bodies such as the IUCN calling for 30% protection by
+conservation toolbox, with 5.9% of the world's oceans currently in implemented MPAs [@grorud-colvert2021], and bodies such as the IUCN calling for 30% protection by
 2030. Many of the largest of these MPAs are being implemented in the
-form of "blue-water" MPAs largely covering pelagic ecosystems,
-characterized by highly mobile species and fishing fleets using a
-diverse mix of static and dynamic ocean features. These systems operate
+form of large MPAs primarily covering pelagic ecosystems. These pelagic systems operate
 very differently from the coastal areas where modern MPA design
-practices have been developed. Pelagic systems are also characterized by
+practices have been developed, characterized by highly mobile species and fishing fleets using a
+diverse mix of static and dynamic ocean features. These ecosystems
 combinations of relatively abundant species targeted by fishing, such as

### 2021-09-15 documents/03_marlin_paper.Rmd
safety stop before open access idea

@@
 
-** something to point out: increased catches are a paradox, either come from rebuilding stock or concentrated fishing, but trajectory can tell you something.
-not possible for MPAs to increase catches without decreasing catches first if they are rebuilding?
-**
+\*\* something to point out: increased catches are a paradox, either
+come from rebuilding stock or concentrated fishing, but trajectory can
+tell you something. not possible for MPAs to increase catches without
+decreasing catches first if they are rebuilding? \*\*
 
-Seasonal movemetn is kind of interesting, basically just dilutes the extremes
+Seasonal movemetn is kind of interesting, basically just dilutes the
+extremes
 
-less resilient species are both more in need of protection and easier to accidentally fish harder
+less resilient species are both more in need of protection and easier to
+accidentally fish harder
 
@@
 Marine protected areas (MPAs) are a growing part of the marine
-conservation toolbox, with 5.9% of the world's oceans currently in implemented MPAs [@grorud-colvert2021], and bodies such as the IUCN calling for 30% protection by
-2030. Many of the largest of these MPAs are being implemented in the
-form of large MPAs primarily covering pelagic ecosystems. These pelagic systems operate
-very differently from the coastal areas where modern MPA design
-practices have been developed, characterized by highly mobile species and fishing fleets using a
-diverse mix of static and dynamic ocean features. These ecosystems
-combinations of relatively abundant species targeted by fishing, such as
-most tuna stocks, and highly threatened species caught as incidental
-"bycatch" in targeted fishing operations, such as oceanic whitetip
-sharks. Blue-water MPAs are increasingly being proposed specifically to
-benefit these by-catch populations while minimizing losses or where
-possible providing benefits to fisheries.
+conservation toolbox, with 5.9% of the world's oceans currently in
+implemented MPAs [@grorud-colvert2021], and bodies such as the IUCN
+calling for 30% protection by 2030. The modern science of MPAs was
+largely developed around coastal ecosystems, often characterized by
+relatively static features such as coral reefs [citation]. However,
+large MPAs are increasingly being implemented and considered across
+pelagic ecosystems. These pelagic systems are characterized by highly
+mobile species and fishing fleets using a diverse mix of static and
+dynamic ocean features [@game2009; @lubchenco2003]. These ecosystems are
+also home to combinations of relatively abundant species targeted by
+fishing, such as many tuna stocks [@pons2016], and highly threatened
+species caught as incidental "bycatch" in targeted fishing operations,
+such as oceanic whitetip sharks [@pacoureau2021]. Large pelagic MPAs are
+increasingly being proposed specifically to benefit these by-catch
+populations while minimizing losses or where possible providing benefits
+to fisheries [citation].
 
-While the effect of MPAs on fisheries yields is a continuously debated
-topic, the conservation impacts of MPAs have received less attention of
-late, due in part perhaps to the perceived obvious nature of the
-problem: if you prevent fishing of a species within an area, there will
-be more of that species in that area, and so the conservation benefits
-of MPAs will increase with MPA size. \@hastings2017 clearly illustrates
-this idea. Using a simple two-species two-patch analytical model, they
-derive conditions under which MPAs can solve the "weak-stock" problem in
-which a less productive species is incidentally exploited by a fleet
-targeting a more valuable and productive stock.
+In this paper we evaluate the potential conservation and fishery effects
+of MPAs within the pelagic context. We focus in particular on the
+potential for pelagic MPAs to provide conservation benefits to bycatch
+species. While the effect of MPAs on fisheries yields is a continuously
+debated topic, the conservation impacts of MPAs are seemingly simpler,
+albeit challenging to detect in many cases [@ovando2021a], due in part
+perhaps to the perceived obvious nature of their benefits: if you
+prevent fishing of a species within an area, there will be more of that
+species in that area, and so the conservation benefits of MPAs will
+increase with MPA size. \@hastings2017 clearly illustrates this idea.
+Using a simple two-species two-patch analytical model, they derive
+conditions under which MPAs can solve the "weak-stock" problem in which
+a less productive species is caught as btcatch by a fleet targeting a
+more valuable and productive stock.
 
@@
 sedentary 2) the target and bycatch species overlap perfectly in terms
-of habitat xx....The authors suggest that the impact of these
-assumptions should be tested in scenarios under which they are unlikely
-to apply.
-We build on \@hastings2017 to consider the role of MPAs in reducing bycatch reductino while managing tradeoffs with fisheries under a more realistic model reflecting the dynamics of pelagic ecosystems and fisheries.
-
-The core models underpinning MPA science have largely been based on
-stylized versions of near-shore reef-style xx what to call that, like
-coral / rocky / kelp forest, etc xx systems, where multiple species
-often share common habitat, and adults have reasonable site fidelity
-while larvae are broadly dispersed. Pelagic ecosystems have vastly
-different dynamics than coastal seas [@game2009; @lubchenco2003],
-characterized by highly migratory species specialized to take advantage
-of highly dynamic ocean features such as eddies currents while also
-utilizing static features such as seamounts. Pelagic ecosystems also
-support both coastal domestic fishing industries along with highly
-mobile international high-seas fishing operations. Given these dynamics,
-it remains unclear how many of the lessons from coastal MPA science are
-applicable to pelagic ecosystems [\@ovando2021XX].
+of habitat and 3) recruits are perfectly mixed throughout the system.
+Pelagic species are likely to violate one or more fo these assumptions.
+We build on \@hastings2017 to consider the role of MPAs in reducing
+bycatch while managing tradeoffs with fisheries under a more resolved
+model reflecting the dynamics of pelagic ecosystems and fisheries.
 
@@
 key aggregation locations may enhance depleted pelagic fisheries [xx
-that Indian ocean study I can't remember right now]. xx big question
-here: how focused do we want this paper to be? is it just about
-conservation tradeoffs or lack thereof with bycatch species, or is about
-the net conservation and fishery effects of BWMPAs. I'm inclined the
-former to tee-up the later xx.
+that Indian ocean study I can't remember right now]. However, the
+dynamic nature of pelagic systems also creates a risk of unintended
+conservation consequences of MPAs. When an MPA is implemented on former
+fishing grounds, fishing fleets have two choices: exit the fishery or
+move to new fishing grounds. When MPAs redistribute effort away from one
+area and into another, they can accidentally increase fishing effort on
+species found within the new fishing grounds [@abbott2012]. This problem
+of policy "leakage" is well known in many environmental fields
+[@citations] but the general properties of this problem have received
+relatively little attention in the theoretical or empirical MPA
+literature [see
+@abbott2012;@savina2008;[fulton2011a;\@baum2003](mailto:fulton2011a;@baum2003){.email}
+for notable exceptions].
 
-However, the dynamic nature of pelagic systems also creates a risk of unintended conservation consequences of MPAs.
-When an MPA is implemented on former fishing grounds, fishing fleets have two choices: exit the fishery or move to new fishing grounds.
-When MPAs redistribute effort away from one area and into another, they can accidentally increase fishing effort on species found within the new fishing grounds [@abbott2012].
-This problem of policy "leakage" is well known in many environmental fields [@citations] but the general properties of this problem have received relatively little attention in the theoretical or empirical MPA literature [see @abbott2012;@savina2008;fulton2011a;@baum2003 for notable exceptions].
-We fill this gap by examining the general range of conditions...
+This study addresses this gap using a novel bio-economic simulation
+model, capable of reflecting several of the key challenges of pelagic
+ecosystems, namely high and variable movement rates, mobile fishing
+fleets, and heterogeneous species distributions. We first provide a set
+of stylized simulations intended to illustrate the potential and
+challenges of achieving conservation objectives through pelagic MPAs,
+and highlight particularly important drivers and the magnitude and
+direction of MPA effects. We then present a more focused case study
+based on the Western and Central Pacific Ocean, examining the potential
+outcomes of pelagic MPAs on bycatch and target species.
 
-- Assessing the conditions under which this may be a problem using a generalized simulation environment
-  - General insights about design strategy, effect of varying degrees of habitat overlap, etc, tradeoffs with fisheries
+<!-- We fill this gap by examining the general range of conditions… -->
 
-- Given that "it depends" nature of the problem, assess the specific potential for this problem in a case study
-  - specific insights about design strategy, effect of varying degrees of habitat overlap, etc, tradeoffs with fisheries
+\<!-- \# Assessing the conditions under which this may be a problem
+using a generalized simulation environment
+
+General insights about design strategy, effect of varying degrees of
+habitat overlap, etc, tradeoffs with fisheries
+
+-   Given that "it depends" nature of the problem, assess the specific
+    potential for this problem in a case study
+
+    -   specific insights about design strategy, effect of varying
+        degrees of habitat overlap, etc, tradeoffs with fisheries
 
@@
 sensitivity of bycatch populations and targeted fisheries to key
-unknowns and design choices in MPA planning.
+unknowns and design choices in MPA planning. --\>
 
@@
 
-@hastings2017 employed a simple model to derive conditions under which
-MPAs would benefit bycatch species. However they also pointed out that
-even small increases in model complexity quickly make analytical
-solutions to the effects of MPAs intractable. As such, we based our numerical
-analysis on a spatially explicit, multi-species, multi-fleet,
-bio-economic model we call
+@hastings2017 employed an analytically tractable model to derive
+conditions under which MPAs would benefit bycatch species. However they
+also pointed out that even small increases in model complexity quickly
+make analytical solutions to the effects of MPAs intractable. As such,
+we employ a numerical simulation strategy in the form of a spatially
+explicit, multi-species, multi-fleet, bio-economic model we call
 [`marlin`](https://github.com/DanOvando/marlin). Models of MPAs span a
@@
 specifically around the role of technical interactions across species.
-Many MPA models simulate the impacts of MPAs on a single-species level [@2021].
-This can be useful for for example examining the ways in which species-specific life history traits interact with fleet dynamics to influence MPA outcomes.
-However, bycatch problems are driven by technical interactions, in which fishing effort for a target species incidentally captures another species xx how do we want to talk about this in here?
-Like usually incidentally, but clearly sometimes people try and catch sharks too xx.
-`marlin` accounts for this by allowing the user to simulate an arbitrary number of fleets capable of catching an arbitrary number of species in a spatial area.
-Different fleets can place different economic value on different species, while also having species-specific selectivity functions.
-This allows us to simulate bycatch dynamics of MPAs in pelagic ecosystems.
-
-We first ran a series of simulation exeriments designed to explore broad sensitivities of bycatch and fisheries to key factors.
-
-We then ran a focused case study intended to reflect some of the traits of the tuna fisheries of the WCPO, holding constant...
+Many MPA models simulate the impacts of MPAs on a single-species level
+[@ovando2021a]. This can be useful for for example examining the ways in
+which species-specific life history traits interact with fleet dynamics
+to influence MPA outcomes. However, bycatch problems are driven by
+technical interactions, in which fishing effort for a target species
+incidentally captures another species. `marlin` accounts for this by
+allowing the user to simulate an arbitrary number of fleets capable of
+catching an arbitrary number of species in a spatial area. Different
+fleets can place different economic value on different species, while
+also having species-specific selectivity functions. This allows us to
+simulate bycatch dynamics of MPAs in pelagic ecosystems.
 
@@
 
-Lots of equations. Pretty standard age-structured population models, extra bells and whistles that need explaining
+Lots of equations. Pretty standard age-structured population models,
+extra bells and whistles that need explaining
 
 -   Design of movement and habitat
-  - Movement is a product of distance and habitat, such that critters will all else equal move to the closest patches with the best habitat
+
+    -   Movement is a product of distance and habitat, such that
+        critters will all else equal move to the closest patches with
+        the best habitat
 
@@
 
--
+Fleets exert a given amount of effort in a patch each time step. The
+catch resulting from that effort is given by the Baranov [@baranov1918]
+equation
 
-## Simulation Experiments
+$$C_{f,s,p,t} = \frac{q_{f,s}E_{f,p,t}}{q_{f,s}E_{f,p,t} + m_s} \times \sum_{a=0}^{A}B_{s,p,t,a} \times (1 - e^{-(q_{f,s}E_{f,p,t}S_a + m_s)}) $$
 
-Given the complexity of the operating model used here, a complete sensitivity analysis across all parameters is not practical. Instead, we focus on the role of life history, fishing pressure,
+Where $C$ is catch by fleet $f$ of species $s$ in patch $p$ in timestep
+*t*. $E$ is total effort by that fleet in a patch and time step, $m$ is
+natrual mortality for a given species, $B$ is biomass at age *a*, and
+$S$ is selectivity.
+
+$q$ is the catchability coefficient for fleet $f$ and species $s$.
+
+Fleets distribute themselves in space according to revenue per unit
+effort
+
+$$R_{f,p,t} = \frac{\sum_{s=1}^{S}v_{f,s}C_{f,s,p,t}}{E_{f,p,t}}$$ Where
+$v$ is the price per kg of each species. Bycatch in the model is defined
+through the price. An unregulated species with no commercial value would
+have $v = 0$, whereas a regulated and prohibited species might actually
+have a negative price.
+
+Effort per fleet is set a a constant of 1 per patch, and the model then
+tunes $q_{f,s}$ such that given all the other parameters of the model
+the fishing mortality rate for each species matches the best available
+level for that species. This is what allows the model to capture bycatch
+dynamics: A fleet can distribute itself purely based on revenues from
+target species, but through the $q$ parameter still exert fishing
+mortality on other species, even if that species has a price of zero, if
+those bycatch species inhabit the same patches as the more valuable
+target species.
+
+###
+
+### MPA Design Strategies
+
+MPA design is governed by four general parameters: the number of cells
+to place inside an MPAs, and the design strategy used to place the MPAs,
+the species considered in the MPA planning process, and the accuracy of
+the information used in the planning process.
+
+We consider `r n_distinct(experiment_obj$placement_strategy)` different
+MPA placement strategies.
+
+For the "depletion" strategy, MPA locations are prioritized based in
+proportion to the spawning stock biomass (SSB) of the most depleted
+species. So, patches with high amounts of highly depleted species are
+prioritized over patches with high biomass of lightly depleted species.
+
+For the "rate" strategy, MPAs are placed in based on the
+depletion-weighted catch relative to the total catch in a patch. So,
+patches where with high rates of catch of depleted species relative to
+totcal catch are prioritized.
+
+For the "avoid_fishing" strategy, MPAs locations are prioritized
+inversely to fishery catches. So, patches with low catches are
+prioritized over patches with high catches.
+
+For the "target_fishing" strategy, MPAs locations are prioritized
+proportionally to fishery catches. So, patches with high total catches
+are prioritized over patches with high catches.
+
+For the "area" strategy, we simply assign patches in sequential order
+until the desired total MPA size is reached.
+
+The information used to inform each MPA design strategy (except area)
+can include a multiplicative observation error term $\sigma_{obs}$ where
+
+$$information \sim lnorm(data,\sigma_{obs}) $$
+
+For example, for the depletion scenario, $\sigma_{obs}$ controls the
+accuracy of the spatial depletion data used to prioritize MPA placement.
+
+## Illustrative Experiments
+
+Given the complexity of the operating model used here, a complete
+sensitivity analysis across all parameters is not practical. Instead, we
+focus on the role of life history, fishing pressure, and MPA design.
 
@@
 
-Each simulation considers `n_distinct(marlin_inputs$scientific_name)` species comprised of the top 12 species by landings reported in the tuna fisheries of the Western and Central Pacific Ocean (WCPO).
-Rough Life history parameters are pulled from `FishLife` [@thorson2020].
-For teleost fishes, we assumed that fecundity was proportional to weight-at-age times the maturity ogive.
-For cartilaginous species, we adopt a fixed fecundity relationship of XX pups per mature individual XX check on what I actually implemented here XX.
-The unfished spawning stock biomass for each of the included species was tuned in the model to reflect best available estimates from stock assessments.
+Each simulation considers `n_distinct(marlin_inputs$scientific_name)`
+species comprised of the top 12 species by landings reported in the tuna
+fisheries of the Western and Central Pacific Ocean (WCPO). Rough Life
+history parameters are pulled from `FishLife` [@thorson2020]. For
+teleost fishes, we assumed that fecundity was proportional to
+weight-at-age times the maturity ogive. For cartilaginous species, we
+adopt a fixed fecundity relationship of XX pups per mature individual XX
+check on what I actually implemented here XX. The unfished spawning
+stock biomass for each of the included species was tuned in the model to
+reflect best available estimates from stock assessments.
 
-Each simulation includes a longline and a purse-seine fleet. Each species can have a separate price (USD/kg) and selectivity ogive for each fleet, as dictated by the literature.
-
-Fleets exert a given amount of effort in a patch each time step. The catch resulting from that effort is given by the Baranov [@baranov18XX] equatino
-
-$$C_{f,s,p,t} = \frac{q_{f,s}E_{f,p,t}}{q_{f,s}E_{f,p,t} + m_s} \times \sum_{a=0}^{A}B_{s,p,t,a} \times (1 - e^{-(q_{f,s}E_{f,p,t}S_a + m_s)}) $$
-
-Where $C$ is catch by fleet $f$ of species $s$ in patch $p$ in timestep *t*.
-$E$ is total effort by that fleet in a patch and time step, $m$ is natrual mortality for a given species, $B$ is biomass at age *a*, and $S$ is selectivity.
-
-$q$ is the catchability coefficient for fleet $f$ and species $s$.
-
-Fleets distribute themselves in space according to revenue per unit effort
-
-$$R_{f,p,t} = \frac{\sum_{s=1}^{S}v_{f,s}C_{f,s,p,t}}{E_{f,p,t}}$$
-Where $v$ is the price per kg of each species.
-Bycatch in the model is defined through the price.
-An unregulated species with no commercial value would have $v = 0$, whereas a regulated and prohibited species might actually have a negative price.
-
-Effort per fleet is set a a constant of 1 per patch, and the model then tunes $q_{f,s}$ such that given all the other parameters of the model the fishing mortality rate for each species matches the best available level for that species.
-This is what allows the model to capture bycatch dynamics: A fleet can distribute itself purely based on revenues from target species, but through the $q$ parameter still exert fishing mortality on other species, even if that species has a price of zero, if those bycatch species inhabit the same patches as the more valuable target species.
+Each simulation includes a longline and a purse-seine fleet. Each
+species can have a separate price (USD/kg) and selectivity ogive for
+each fleet, as dictated by the literature.
 
@@
 
-We simulate degrees of species overlap using two parameters, $\sigma_c$ and $\sigma_h$.
-All species start with the assumption that the centroid of their preferred habitat occurs in the center-most patch of teh two dimensional system simulated by marlin.
-For our simulations, we then generated a random deviation from that default centroid for each species using $\sigma_c$ in the manner
+We simulate degrees of species overlap using two parameters, $\sigma_c$
+and $\sigma_h$. All species start with the assumption that the centroid
+of their preferred habitat occurs in the center-most patch of teh two
+dimensional system simulated by marlin. For our simulations, we then
+generated a random deviation from that default centroid for each species
+using $\sigma_c$ in the manner
 
-$$centroid_s \sim normal(center,\sigma_c)$$
-where *s* denotes species, and *center* is the coordinate of the central cell.
-In this way, the larger $\sigma_c$ grows, the more variation there is betwen the centroid of the preferred habitat for each species.
+$$centroid_s \sim normal(center,\sigma_c)$$ where *s* denotes species,
+and *center* is the coordinate of the central cell. In this way, the
+larger $\sigma_c$ grows, the more variation there is betwen the centroid
+of the preferred habitat for each species.
 
-We then control the spread of a creatures preferred habitat around that centroid per a Gaussian dispersal kernal in the manner
+We then control the spread of a creatures preferred habitat around that
+centroid per a Gaussian dispersal kernal in the manner
 
@@
 
-Where $d$ is the euclidian distance from the $centroid_s$ to patch $p$ and $\sigma_h$ controls the dispersal of the habitat around that range.
+Where $d$ is the euclidian distance from the $centroid_s$ to patch $p$
+and $\sigma_h$ controls the dispersal of the habitat around that range.
 
@@
 
-For this paper `marlin` is set to run with two seasons per year. When seasonal movement is turned on, animals use one habitat matrix in season one, and the inverse of that habitat matrix in season 2.
+For this paper `marlin` is set to run with two seasons per year. When
+seasonal movement is turned on, animals use one habitat matrix in season
+one, and the inverse of that habitat matrix in season 2.
 
-### MPA Design Strategies
+## WCPO Case Study
 
-MPA design is governed by four general parameters: the number of cells to place inside an MPAs, and the design strategy used to place the MPAs, the species considered in the MPA planning process, and the accuracy of the information used in the planning process.
-
-We consider `r n_distinct(experiment_obj$placement_strategy)` different MPA placement strategies.
-
-For the "depletion" strategy, MPA locations are prioritized based in proportion to the spawning stock biomass (SSB) of the most depleted species.
-So, patches with high amounts of highly depleted species are prioritized over patches with high biomass of lightly depleted species.
-
-For the "rate" strategy, MPAs are placed in based on the depletion-weighted catch relative to the total catch in a patch.
-So, patches where with high rates of catch of depleted species relative to totcal catch are prioritized.
-
-For the "avoid_fishing" strategy, MPAs locations are prioritized inversely to fishery catches. So, patches with low catches are prioritized over patches with high catches.
-
-For the "target_fishing" strategy, MPAs locations are prioritized proportionally to fishery catches. So, patches with high total catches are prioritized over patches with high catches.
-
-For the "area" strategy, we simply assign patches in sequential order until the desired total MPA size is reached.
-
-The information used to inform each MPA design strategy (except area) can include a multiplicative observation error term $\sigma_{obs}$ where
-
-$$information \sim lnorm(data,\sigma_{obs}) $$
-
-For example, for the depletion scenario, $\sigma_{obs}$ controls the accuracy of the spatial depletion data used to prioritize MPA placement.
-
-## Case Study
-
-In order to constrain the parameter space somewhat, we also ran a series of case study experiments.
-Rathere than interating over different habitat configurations and movement dynamics, we fix habitat for each species based on the results of @bradley2021XX, and set ontogenetic movement and seasonal migrations to FALSE.
+In order to constrain the parameter space somewhat, we also ran a series
+of case study experiments. Rather than iterating over different habitat
+configurations and movement dynamics, we defined habitat based on
+reported catch per unit effort (CPUE) from the WCPO. XX more methods
+here XX
 
@@
 
-For the case study section, as a point of comparison we calculate an optimized MPA network along with the "rule of thumb" networks based on the design strategies outlined in section XX.
-The optimization routine is based on an objective function of the form
+For the case study section, as a point of comparison we calculate an
+optimized MPA network along with the "rule of thumb" networks based on
+the design strategies outlined in section XX. The optimization routine
+is based on an objective function of the form
 
@@
 
-where $\alpha$ controls the relative weight assigned to a "conservation" objective based on depletion, relative to an "economic" objective based on revenues.
-The 'scale' function rescales both objective components to the range of 0 to 1 to facilitate interpretation of $\alpha$ xx come up with a better version of this XX
+where $\alpha$ controls the relative weight assigned to a "conservation"
+objective based on depletion, relative to an "economic" objective based
+on revenues. The 'scale' function rescales both objective components to
+the range of 0 to 1 to facilitate interpretation of $\alpha$ xx come up
+with a better version of this XX
 
-The optimization routine itself works through a sampling-importance-resampling algorithm.
-In the first iteration of the algorithm, the marginal contribution of each individual patch to the objective function is calculated.
-The patch with the highest objective function value is then placed in an MPA.
-For the remaining iterations, we sample *X* cells from the remaining non-MPA cells with probability proportional to their marginal value to the objective function in the last iteration.
-We then calculate the marginal value of each of these cells when individually added to the MPA network begun in the prior step.
-The cell with the higheset marginal value is then added to the MPA network, and the marginal values of the non-selected but sampled cells are updated based on the values in that iteration.
-The process then continues until the desired MPA size is reached.
+The optimization routine itself works through a
+sampling-importance-resampling algorithm. In the first iteration of the
+algorithm, the marginal contribution of each individual patch to the
+objective function is calculated. The patch with the highest objective
+function value is then placed in an MPA. For the remaining iterations,
+we sample *X* cells from the remaining non-MPA cells with probability
+proportional to their marginal value to the objective function in the
+last iteration. We then calculate the marginal value of each of these
+cells when individually added to the MPA network begun in the prior
+step. The cell with the higheset marginal value is then added to the MPA
+network, and the marginal values of the non-selected but sampled cells
+are updated based on the values in that iteration. The process then
+continues until the desired MPA size is reached.
 
@@
 
-Each model run started at equilibrium conditions and was then run for `r years` years (`r seasons * years` seasons.
-We assessed performance by comparing relevant metrics (e.g.
-depletion and revenue) under a given scenario with MPAs relative to the values of those same metrics in the simulated counterfactual world identical in every way save for the presencee of MPAs.
+Each model run started at equilibrium conditions and was then run for
+`r years` years (`r seasons * years` seasons. We assessed performance by
+comparing relevant metrics (e.g. depletion and revenue) under a given
+scenario with MPAs relative to the values of those same metrics in the
+simulated counterfactual world identical in every way save for the
+presence of MPAs.
 
@@
 
-## Simulation Experiments
+## Illustrative Experiments
 
-We focused first on the range of impacts that MPAs can have on mature biomass across the range of variables explored.
-To illustrate, we pulled out two example simulations, one where MPAs provide relatively large benefits across all three species, and a second where MPAs actually produced substantial losses in some species....
+We focused first on the range of impacts that MPAs can have on mature
+biomass across the range of variables explored. To illustrate, we pulled
+out two example simulations, one where MPAs provide relatively large
+benefits across all three species, and a second where MPAs actually
+produced substantial losses in some species....
 
@@
 
-Looking at the full `r n_distinct(total_results$placement_id) * n_distinct(total_results$state_id)` experiments run, the results were highly variable, with MPAs capable of producing substantial conservation gains to substantial conservation losses and everything in between.
-We do not focus into the individual effects of particular variables as we found them to be highly dependent on the exact question asked and the set of data included, and since we cannot ascribe any actual likelihood to any individual state of nature simulated here.
+Looking at the full
+`r n_distinct(total_results$placement_id) * n_distinct(total_results$state_id)`
+experiments run, the results were highly variable, with MPAs capable of
+producing substantial conservation gains to substantial conservation
+losses and everything in between. We do not focus into the individual
+effects of particular variables as we found them to be highly dependent
+on the exact question asked and the set of data included, and since we
+cannot ascribe any actual likelihood to any individual state of nature
+simulated here.
 
-However, three important features do emerge.
-The first is that life history plays an important role in the potential for unintended conservation consequences from MPAs.
-All else equal, MPAs were much less likely to produce losses in spawning biomass for our tuna-like species than the marlin or shark-like species.
-Given that within the broad life history traits we do vary both adult and larval movement, as well as the timing of density dependence, we believe this result is mostly due to the high resilience (high steepness, low age at maturity, fast growth) of the tuna-like species relative to the marlin and shark-like species.
+However, three important features do emerge. The first is that life
+history plays an important role in the potential for unintended
+conservation consequences from MPAs. All else equal, MPAs were much less
+likely to produce losses in spawning biomass for our tuna-like species
+than the marlin or shark-like species. Given that within the broad life
+history traits we do vary adult and larval movement, fishing mortality,
+as well as the timing of density dependence, we believe this result is
+mostly due to the high resilience (high steepness, low age at maturity,
+fast growth) of the tuna-like species relative to the marlin and
+shark-like species.
 
@@
 
-The third is that extreme outcomes (both positive and negative), particularly at lower MPA sizes, are largely driven by the presence of small range sizes interacting with seasonal migrations.
-Across species simulations without seasonal migrations and small ranges were much more likely to produce extreme results than those with seasonal migrations and/or larger ranges.
-Presence of seasonal migrations acts as a dilutant: As a given static MPA only provides protection, or displaces fishing effort, onto a particular area in only half of the year, both the benefits and costs of MPAs are smaller.
-When seasonal migrations are not present and the species range is small, MPAs are more capable of providing either large benefits or large losses.
+The third is that extreme outcomes (both positive and negative),
+particularly at lower MPA sizes, are largely driven by the presence of
+small range sizes interacting with seasonal migrations. Across species,
+simulations without seasonal migrations and small ranges were much more
+likely to produce extreme results than those with seasonal migrations
+and/or larger ranges. Presence of seasonal migrations acts as a
+dilutant: As a given static MPA only provides protection, or displaces
+fishing effort, onto a particular area in only half of the year, both
+the benefits and costs of MPAs are smaller. When seasonal migrations are
+not present and the species range is small, MPAs are more capable of
+providing either large benefits or large losses.
 
-Make clear that a goal of all this complexity was to see if there were any clear emergent properties (this thing matters way more than this thing).
-We highlight some of those, but given the complexity we found we present the complexity of interacting ways in which different things can effect the conservation outcomes of MPAs
+Make clear that a goal of all this complexity was to see if there were
+any clear emergent properties (this thing matters way more than this
+thing). We highlight some of those, but given the complexity we found we
+present the complexity of interacting ways in which different things can
+effect the conservation outcomes of MPAs
 
-XX after looking through these should widdle it down to only things you actually examine, annoying / no point to having a bunch of unexplained minor variation in the mix relatede to things like ontogenetic shifts) xx
+XX after looking through these should widdle it down to only things you
+actually examine, annoying / no point to having a bunch of unexplained
+minor variation in the mix relatede to things like ontogenetic shifts)
+xx
 
-xx the problem here.
-These things all interact in so many complex ways that it's hard to look at either graphically or statistically, and maybe for the best.
-i.e.
-the answer isn't "if habitat overlap is less than XX% you might have a problem, otherwise you're fine", but rather that a bunch of things can interact to cause potential problems in complex ways.
-but that's hard to explain xx.
+xx the problem here. These things all interact in so many complex ways
+that it's hard to look at either graphically or statistically, and maybe
+for the best. i.e. the answer isn't "if habitat overlap is less than XX%
+you might have a problem, otherwise you're fine", but rather that a
+bunch of things can interact to cause potential problems in complex
+ways. but that's hard to explain xx.
 
-Big message: even constrained by the same initial depletion and core biology (e.g.
-movement rates), a wide range of outcomes are possible for each species for a given MPA size, dependent on the design strategy, habitat overlap, migrations, quality of information, etc.
-Across the board as MPA size increases the ratio of positive to negative outcomes increases, until all species achieve their maximum conservation potential in every iteration at 100% MPA.
-HOwever, every species experienced a set of simulations, roughyl xx%, that resulted in net losses in SSB/SSB0 relative to BAU.
-Losses were smaller than potential gaines, but reached meaningful levels in some species, and interestingly peaked in many cases right before max MPA (reflecting concentration of fishing in small amounts of habitat in habited by the species).Important to stress in here that there are no probabilities in this, no way of knowing which state is more probable than others, just given the range of things simulated more ways for a bigger positive effect.
-Also make clear though, biggish losses (10% or more) possible, but rare.
+Big message: even constrained by the same initial depletion and core
+biology (e.g. movement rates), a wide range of outcomes are possible for
+each species for a given MPA size, dependent on the design strategy,
+habitat overlap, migrations, quality of information, etc. Across the
+board as MPA size increases the ratio of positive to negative outcomes
+increases, until all species achieve their maximum conservation
+potential in every iteration at 100% MPA. HOwever, every species
+experienced a set of simulations, roughyl xx%, that resulted in net
+losses in SSB/SSB0 relative to BAU. Losses were smaller than potential
+gaines, but reached meaningful levels in some species, and interestingly
+peaked in many cases right before max MPA (reflecting concentration of
+fishing in small amounts of habitat in habited by the species).Important
+to stress in here that there are no probabilities in this, no way of
+knowing which state is more probable than others, just given the range
+of things simulated more ways for a bigger positive effect. Also make
+clear though, biggish losses (10% or more) possible, but rare.
 
@@
 
-1.
-Even conditional on fishing exploitation and general life history strategy, the effects of MPAs on a group of species with technical interactions is highly variable and dependent on habitat distribution, design decisions, and data quality.
-xx make this more explicit.
+1.  Even conditional on fishing exploitation and general life history
+    strategy, the effects of MPAs on a group of species with technical
+    interactions is highly variable and dependent on habitat
+    distribution, design decisions, and data quality. xx make this more
+    explicit.
 
-2.
-Some degree of unintended conservation consequences occured across nearly all of our simulations below MPA coverage near 100%, though small losses were much more common than large.
-Different design strategies are better suited to avoiding unintended consequences.
+2.  Some degree of unintended conservation consequences occured across
+    nearly all of our simulations below MPA coverage near 100%, though
+    small losses were much more common than large. Different design
+    strategies are better suited to avoiding unintended consequences.
 
-There two general results beg some important questions.
-First, are unintended consequences inevitable given the kinds of technical interactions explored here, or can optimized MPA networks reduce or eliminate these unintended consequences.
-Second, one clear solution to the problem of unintended consequences is to simply make the MPAs large: if MPAs cover 100% of the area there can be no direct unintended conservation consequences.
-The potential for negative impacts on economic fishery outcomes increases as MPAs reach large size though.
-What then are the tradeoffs between reducing the potential for unintended conservation consequences and fishery revenues?
+There two general results beg some important questions. First, are
+unintended consequences inevitable given the kinds of technical
+interactions explored here, or can optimized MPA networks reduce or
+eliminate these unintended consequences. Second, one clear solution to
+the problem of unintended consequences is to simply make the MPAs large:
+if MPAs cover 100% of the area there can be no direct unintended
+conservation consequences. The potential for negative impacts on
+economic fishery outcomes increases as MPAs reach large size though.
+What then are the tradeoffs between reducing the potential for
+unintended conservation consequences and fishery revenues?
 
-To address these questions, we ran a more limited "case study" set of simulations, based on the dynamics of the WCPO tuna fisheries.
-Rather than iterating over different movement dynamics and habitat configurations, we fix habitat layers on the values predicted by @bradley2021, and xx other things we fix xx.
+To address these questions, we ran a more limited "case study" set of
+simulations, based on the dynamics of the WCPO tuna fisheries. Rather
+than iterating over different movement dynamics and habitat
+configurations, we fix habitat layers on the values predicted by
+@bradley2021, and xx other things we fix xx.
 
-For this subset of simulations then, along with the various "rule of thumb" MPA design strategies we ran an optimization routine (see section XX) across a range of objective function weightings, in the manner of @rassweiler2014.
-This allowed us to assess 1) the degree to which unintended conservation consequences of MPAs can be reduced by optimized design and 2) the magnitude of tradeoffs between conservation outcomes and fishery revenues under both rule-of-thumb and optimized MPA networks.
+For this subset of simulations then, along with the various "rule of
+thumb" MPA design strategies we ran an optimization routine (see section
+XX) across a range of objective function weightings, in the manner of
+@rassweiler2014. This allowed us to assess 1) the degree to which
+unintended conservation consequences of MPAs can be reduced by optimized
+design and 2) the magnitude of tradeoffs between conservation outcomes
+and fishery revenues under both rule-of-thumb and optimized MPA
+networks.
 
-key results: optimization helps (Nobel prize please).
-Can reduce percent of species with conservation loss down to 25% pretty quickly if you prioritize conservation.
-Rule of thumb designs like targeting fishing can get there with perfect information and when considering all animals, but can be pretty far from optimal with imperfect information or when only considering a subset of animals in the planning process (that's what all the different points represent).
-Some rules of thumb are more sensitive to these decision than others.
-E.g.
-rate can get you optimal, but goes off the rails pretty quickly with bad information.
+key results: optimization helps (Nobel prize please). Can reduce percent
+of species with conservation loss down to 25% pretty quickly if you
+prioritize conservation. Rule of thumb designs like targeting fishing
+can get there with perfect information and when considering all animals,
+but can be pretty far from optimal with imperfect information or when
+only considering a subset of animals in the planning process (that's
+what all the different points represent). Some rules of thumb are more
+sensitive to these decision than others. E.g. rate can get you optimal,
+but goes off the rails pretty quickly with bad information.
 
-There is also a pretty clear breakpoint in results between basically trying for conservation or trying for economics, and not a lot of wiggle room in between (the two general bands of lines).
-xx check to what extent this depends on the way you scale the two objective function components xx.
+There is also a pretty clear breakpoint in results between basically
+trying for conservation or trying for economics, and not a lot of wiggle
+room in between (the two general bands of lines). xx check to what
+extent this depends on the way you scale the two objective function
+components xx.
 
@@
 
-Key results: Even with optimization, you can basically either be in a world with 50% or more of species having some level of unintended conservation consequence, but no economic loss / some economic gain, or you can be in a world with lower unintended consequences, but economic losses.
-But, even rule of thumb approaches can get you to some places with relatively little unintended consequences at relatively minimal economic loss.
+Key results: Even with optimization, you can basically either be in a
+world with 50% or more of species having some level of unintended
+conservation consequence, but no economic loss / some economic gain, or
+you can be in a world with lower unintended consequences, but economic
+losses. But, even rule of thumb approaches can get you to some places
+with relatively little unintended consequences at relatively minimal
+economic loss.
 
@@
 
-Key results: Presence of and magnitude of unintended consequences pretty sensitive to design descisions.
-On a macro scale though far less so.
-In terms of change in total SSB/SSB0 (metric of biodiversity, i.e.
-maxed out when all species are unfished) and change in revenue, there is basically a pretty simple linear tradeoff between biodiversity and revenue, with a slope less than negative 1, meaning that 1% increase in biodiversity comes at a greater than 1% loss in fishery revenue.
-This stacks up pretty nicely with @rassweiler2014., where value of optimization was low when overfishing of valuable species isn't prevalent, as is the case here.
-So, design and optimization can help a lot with the more "micro" level outcomes (e.g.
-making sure that biodiversity gaines aren't coming from massive gains in one species to detriment of other), but on a macro level pretty insensitive in this case study.
+Key results: Presence of and magnitude of unintended consequences pretty
+sensitive to design descisions. On a macro scale though far less so. In
+terms of change in total SSB/SSB0 (metric of biodiversity, i.e. maxed
+out when all species are unfished) and change in revenue, there is
+basically a pretty simple linear tradeoff between biodiversity and
+revenue, with a slope less than negative 1, meaning that 1% increase in
+biodiversity comes at a greater than 1% loss in fishery revenue. This
+stacks up pretty nicely with @rassweiler2014., where value of
+optimization was low when overfishing of valuable species isn't
+prevalent, as is the case here. So, design and optimization can help a
+lot with the more "micro" level outcomes (e.g. making sure that
+biodiversity gaines aren't coming from massive gains in one species to
+detriment of other), but on a macro level pretty insensitive in this
+case study.
 
@@
 
+motivated by but not limited to pelagic ecosystems
+
 **what is the story**
 
-- Unintended conservation consequences of MPAs resulting from technical interactions have been discussed in the literature, but almost exclusively in the context of highly tactical modeling or localed empirics.
-We conduct one of the first large-scale assessments of the potential of this problem across a range of scenarios
+-   Unintended conservation consequences of MPAs resulting from
+    technical interactions have been discussed in the literature, but
+    almost exclusively in the context of highly tactical modeling or
+    localed empirics. We conduct one of the first large-scale
+    assessments of the potential of this problem across a range of
+    scenarios
 
-- We find that these unintended consequences can be common outside of "canonical" MPA models *a la* hastings etc.
+-   We find that these unintended consequences can be common outside of
+    "canonical" MPA models *a la* hastings etc.
 
-- But, good news is magnitude of unintended consequences were usually small, and total benefits almost always outweigh individual species losses (again no probability of states though)
+-   But, good news is magnitude of unintended consequences were usually
+    small, and total benefits almost always outweigh individual species
+    losses (again no probability of states though)
 
-- Some design strategies are better at avoiding unintended consequences than others, but sensitive to data quality.
-Optimization can help, but even with optimization unintended consequences can happen.
+-   Some design strategies are better at avoiding unintended
+    consequences than others, but sensitive to data quality.
+    Optimization can help, but even with optimization unintended
+    consequences can happen.
 
-- Under the levels of fishing mortality in the WCPO, optimization and design can help reduce tradeoffs between fisheries revenue and unintended consequences than others, but macro-level tradeoffs between biodiversity and fishery revenue are very insensitive to design choices, with a pretty clear "increasing biodiversity through MPAs reduces fishery revenues".
-  -  xx This is clearly hyper-sensitive to the one thing we don't iterate over much here: how overfished are the target species.
-At the moment "they're not".
-But, this runs fast enough, is it worth running a hypothetical case study where they are just to illustrate how things change?
+-   Under the levels of fishing mortality in the WCPO, optimization and
+    design can help reduce tradeoffs between fisheries revenue and
+    unintended consequences than others, but macro-level tradeoffs
+    between biodiversity and fishery revenue are very insensitive to
+    design choices, with a pretty clear "increasing biodiversity through
+    MPAs reduces fishery revenues".
+
+    -   xx This is clearly hyper-sensitive to the one thing we don't
+        iterate over much here: how overfished are the target species.
+        At the moment "they're not". But, this runs fast enough, is it
+        worth running a hypothetical case study where they are just to
+        illustrate how things change?
 
@@
 
-- Unintended conservation outcomes of MPAs totally possible outside of "consider a homogenous fish blob" models
+-   Unintended conservation outcomes of MPAs totally possible outside of
+    "consider a homogenous fish blob" models
 
-- But, usually small, and net conservation benefits outweigh conservation costs
+-   But, usually small, and net conservation benefits outweigh
+    conservation costs
 
-- Should pay particular attention to potential though when species near tipping points are of concern: hard to imagine those not being part of the prioritization but something to consider. 
+-   Should pay particular attention to potential though when species
+    near tipping points are of concern: hard to imagine those not being
+    part of the prioritization but something to consider.
+
 -   Other interesting things.
@@
 
-Thinking through the “naive” MPA design strategies to compare against the optimized outcomes.
-Thoughts at the moment below.
-These can be run in a few seconds each, so we cab be be pretty liberal in coming up with combinations if there are other ideas people want to try, but also if some of these seem pointless can drop as well!
-6:11
-Design strategies
-1. Assign protection by SSB / SSB0 weighted abundance in space (protect the highest concentrations of the most threatened species)
-2. Assign protection by bycatch rates (distribution of biomass bycatch / biomass target)
-3. Assign protection by targeting fishing effort
-4. Assign protection by avoiding fishing effort
-Knowledge states
-1. Degrees of accuracy over the thing being measured (i.e. error in the design strategy)
-2. Number of species considered (e.g. design aronnd only 2 bycatch species instead of all)
-3. über naive: just close off random patches up to a specified size
-We then plot these points in objective function space and compare to the efficiency frontier / consider frequency and magnitude of conservation and economic tradeoofs
-The question all this is answering: how sensitive are conservation and economic outcomes to design strategy / uncertainty when there are technical interactions and heterogeneity in habitat?
-The first stage is teeing up the idea that beyond conservation / yield tradeoffs, as you increase habitat heterogeneity with technical interactions you increase the potential for conservation tradeoffs.
-But, the degree is going to be super context dependent, so to dive in a bit we then focus on this case study of the particular species / habitat overlap in the WCPO.
-Further work might reveal more generalities in this, but htis is basically getting a sense for the potential magnitude of the challenge.
+Thinking through the "naive" MPA design strategies to compare against
+the optimized outcomes. Thoughts at the moment below. These can be run
+in a few seconds each, so we cab be be pretty liberal in coming up with
+combinations if there are other ideas people want to try, but also if
+some of these seem pointless can drop as well! 6:11 Design strategies 1.
+Assign protection by SSB / SSB0 weighted abundance in space (protect the
+highest concentrations of the most threatened species) 2. Assign
+protection by bycatch rates (distribution of biomass bycatch / biomass
+target) 3. Assign protection by targeting fishing effort 4. Assign
+protection by avoiding fishing effort Knowledge states 1. Degrees of
+accuracy over the thing being measured (i.e. error in the design
+strategy) 2. Number of species considered (e.g. design aronnd only 2
+bycatch species instead of all) 3. über naive: just close off random
+patches up to a specified size We then plot these points in objective
+function space and compare to the efficiency frontier / consider
+frequency and magnitude of conservation and economic tradeoofs The
+question all this is answering: how sensitive are conservation and
+economic outcomes to design strategy / uncertainty when there are
+technical interactions and heterogeneity in habitat? The first stage is
+teeing up the idea that beyond conservation / yield tradeoffs, as you
+increase habitat heterogeneity with technical interactions you increase
+the potential for conservation tradeoffs. But, the degree is going to be
+super context dependent, so to dive in a bit we then focus on this case
+study of the particular species / habitat overlap in the WCPO. Further
+work might reveal more generalities in this, but htis is basically
+getting a sense for the potential magnitude of the challenge.
 
@@
 
-We show....
-  - sensitivity of both conservation and fishery outcomes to pelagic realism and design choices
-  - compared to targeted / dynamic intervention... 
-  - value of information would be great...
+We show.... - sensitivity of both conservation and fishery outcomes to
+pelagic realism and design choices - compared to targeted / dynamic
+intervention... - value of information would be great...
 
-Contrast to coastal MPA theoretical work
-  - @hastings2017
-  - @ovando2021
-  - @white2010
-  - not quite sure what to do about points raised in @fulton2011a
-    - to what extent has this been done...
+Contrast to coastal MPA theoretical work - @hastings2017 - @ovando2021 -
+@white2010 - not quite sure what to do about points raised in
+@fulton2011a - to what extent has this been done...
 
@@
 
-"
-Spatial management was insufficient as the dominant management lever because it proves quite difficult to strike the balance between protecting areas that are of a size sufficient to produce any conservation benefit and providing sufficient fishing grounds to sustain an econom- ically viable industry (which is not possible in many fisheries if closures on the order of 70–80% or more are imposed, Fulton et al.
-2007).
-More- over, Savina etal.
-(2008) found that effort displaced by closures potentially undermines any benefits coming from the zoning.
-For instance, strong improvements in diversity (by nearly 20%) under heavy fishing pressure occurred if effort outside a marine protected area (MPA) was controlled.
-However, without effort control, there was a significant (>20%) drop in overall diversity.
-Fulton et al.
-(2007), Savina et al.
-(2008), Smith et al.
-(2010) and unpublished work associated with Horne et al.
-(2010) all found that while spatial management appears to be an effective means of satisfying conservation objectives con- cerning habitat and restoring age-structures of less mobile species, benefits from closures for individual species depend on its trophic position, degree of mobility, geographical distribution and extent of ontogenetic shifts in habitat use and depth (highly mobile or prey species do not benefit in the way that more site attached or predatory species do)."
+" Spatial management was insufficient as the dominant management lever
+because it proves quite difficult to strike the balance between
+protecting areas that are of a size sufficient to produce any
+conservation benefit and providing sufficient fishing grounds to sustain
+an econom- ically viable industry (which is not possible in many
+fisheries if closures on the order of 70--80% or more are imposed,
+Fulton et al. 2007). More- over, Savina etal. (2008) found that effort
+displaced by closures potentially undermines any benefits coming from
+the zoning. For instance, strong improvements in diversity (by nearly
+20%) under heavy fishing pressure occurred if effort outside a marine
+protected area (MPA) was controlled. However, without effort control,
+there was a significant (\>20%) drop in overall diversity. Fulton et al.
+(2007), Savina et al. (2008), Smith et al. (2010) and unpublished work
+associated with Horne et al. (2010) all found that while spatial
+management appears to be an effective means of satisfying conservation
+objectives con- cerning habitat and restoring age-structures of less
+mobile species, benefits from closures for individual species depend on
+its trophic position, degree of mobility, geographical distribution and
+extent of ontogenetic shifts in habitat use and depth (highly mobile or
+prey species do not benefit in the way that more site attached or
+predatory species do)."
 
-Lots of studies have shown that lots of things matter, can we provide guidance on which things are priorities?
+Lots of studies have shown that lots of things matter, can we provide
+guidance on which things are priorities?
 
-Where does this fit in context of @smith2021?
-One important difference is that they point out differences in effectiveness, where effectiveness is still almost always positive, since they just measure leatherbacks.
-So their fig 5, there are a few cases where leatherback catches actually go up, but it's rare.
-So, extending to multiple species and keeping track off all of them shows it's even harder
+Where does this fit in context of @smith2021? One important difference
+is that they point out differences in effectiveness, where effectiveness
+is still almost always positive, since they just measure leatherbacks.
+So their fig 5, there are a few cases where leatherback catches actually
+go up, but it's rare. So, extending to multiple species and keeping
+track off all of them shows it's even harder
 
-How does this contrast with the the large and growing dynamic ocean management lit?
-  - By definition, with perfect information dynamic $\geq$ static spatial controls
-  - DOM literature is mostly static, tracks displacement onto static abundance / risk layers
-  - Doesn't allow for dynamic impacts, for example fleet adjustments based on updated economic surface, or short-term costs for long-term benefits (can overstate costs?)
-  - Dynamic model allows for delta between dynamic policies, e.g. Fmsy vs DOM
-    - analytical models a la hastings have value but do / don't stand up to realism
+How does this contrast with the the large and growing dynamic ocean
+management lit? - By definition, with perfect information dynamic $\geq$
+static spatial controls - DOM literature is mostly static, tracks
+displacement onto static abundance / risk layers - Doesn't allow for
+dynamic impacts, for example fleet adjustments based on updated economic
+surface, or short-term costs for long-term benefits (can overstate
+costs?) - Dynamic model allows for delta between dynamic policies, e.g.
+Fmsy vs DOM - analytical models a la hastings have value but do / don't
+stand up to realism
 
-Presentation of the model as a MICE model for MSP
-  - reasonable parameterizable
-  - flexible
-  - fast
+Presentation of the model as a MICE model for MSP - reasonable
+parameterizable - flexible - fast
 
-Focus of debate has been on fisheries and yields.
-As MPAs move to pelagic systems though, we show that conservation outcomes also depend on correctly identifying ecological state / design decisions etc.
+Focus of debate has been on fisheries and yields. As MPAs move to
+pelagic systems though, we show that conservation outcomes also depend
+on correctly identifying ecological state / design decisions etc.
 

### 2021-09-15 documents/marlin-paper-notes.Rmd
safety stop before open access idea

@@
 Walk through some of the figures and unpack them a bit
+
+# comparing catches to CPUE to abundance
+
+```{r}
+
+rfmo <- "wcpfc"
+mats <- list.files(path = here("data", "tuna-catch"))
+
+# mats <- mats[str_detect(mats, "wcpfc_")]
+
+mats <- mats[str_detect(mats, paste0(rfmo,"_")) & str_detect(mats, "tonnes") & str_detect(mats, "monthly") & str_detect(mats, "2017") & str_detect(mats, "5grid")]
+
+get_layer <- function(file) {
+  # file <- mats[7]
+
+  file_comps <- str_split(file, "_", simplify = TRUE)
+
+  species <- paste(file_comps[,7],file_comps[,8])
+
+  species <- str_replace(species, ".tif", "")
+# 
+  hab <- stars::read_stars(here("data", "tuna-catch", file)) %>%
+    sf::st_as_sf() %>%
+    recenter_vector(center= -200)
+
+#   plot(hab)
+
+  centers <- sf::st_centroid(hab)
+
+  centers <- sf::st_coordinates(centers) %>% 
+    as.data.frame()
+
+  centers$catch = as.data.frame(hab)[,1]
+
+  x_binsize <- (max(centers$X) - min(centers$X) + 1) / resolution
+
+  y_binsize <- (max(centers$Y) - min(centers$Y) + 1) / resolution
+
+  if (x_binsize < 1 | y_binsize < 1) {
+    stop("resolution is too high for input data")
+  }
+
+  habitat <- centers %>%
+    mutate(X = X - min(X),
+           Y = Y - min(Y)) %>%
+    mutate(rough_x = floor(X / x_binsize),
+           rough_y = floor(Y / y_binsize)) %>%
+    dplyr::select(contains("rough_"), catch)
+
+  habitat <- habitat %>%
+    group_by(rough_x, rough_y) %>%
+    summarise(hab = sum(catch)) %>%
+    mutate(xy = rough_x * rough_y)
+
+  mod <-
+    gamm4::gamm4(hab ~ s(rough_x) + s(rough_y) + s(xy), data = habitat)
+
+  habitat_frame <-
+    expand_grid(rough_x = 1:resolution,
+                rough_y = 1:resolution) %>%
+    mutate(xy = rough_x * rough_y)
+
+interp_habitat <- predict(mod$gam, newdata = habitat_frame)
+
+interp_habitat <- pmax(0, interp_habitat / max(interp_habitat))
+
+habitat_frame$habitat <- interp_habitat
+
+habitat_frame %>%
+  ggplot(aes(rough_x, rough_y, fill = habitat)) +
+  geom_tile()
+
+  hab <- habitat_frame %>%
+    dplyr::select(contains("rough"), habitat) %>%
+    rename(x = rough_x, y = rough_y) %>%
+    ungroup()
+
+  out <- tibble(scientific_name = species, habitat = list(hab))
+
+}
+
+mats <- map(mats, safely(get_layer))
+
+mats_worked <- map(mats, "error") %>% map_lgl(is.null)
+
+mats <- mats[mats_worked]
+
+mats <- map(mats, "result")
+
+catch_mats <- mats %>%
+  bind_rows() %>% 
+  unnest(cols = habitat) %>% 
+  group_by(scientific_name, x, y) %>% 
+  summarise(habitat = sum(habitat)) %>% 
+  ungroup() %>% 
+  nest(habitat = c(-scientific_name)) %>% 
+  ungroup()
+
+catch_habitat <- catch_mats %>%
+  bind_rows() %>%
+  unnest(cols = habitat) %>%
+  ggplot(aes(x, y, fill = habitat)) +
+  geom_tile() +
+  facet_wrap( ~ scientific_name) +
+  scale_fill_viridis_c()
+
+catch_habitat
+
+```
+
+Compare to the SDM stuff
+
+```{r}
+
+mats <- list.files(path = here("data", "matrices"))
+
+mats <- mats[str_detect(mats, "longline_wcpfc")]
+
+get_layer <- function(file) {
+  # file <- mats[3]
+
+  file_comps <- str_split(file, "_", simplify = TRUE)
+
+  species <- str_replace(file_comps[, 1], "-", " ")
+
+  fleet <- file_comps[, 2]
+
+  hab <-
+    read_csv(here("data", "matrices", file),
+             col_names = FALSE,
+             skip = 1) %>%
+    rename(lat = X1) %>%
+    mutate(lat = 1:nrow(.)) %>%
+    pivot_longer(
+      -lat,
+      names_to = "lon",
+      values_to = "hab",
+      names_prefix = "X"
+    ) %>%
+    mutate(hab = tidyr::replace_na(hab, 0),
+           lon = as.numeric(lon) - 1)
+
+  # hab %>%
+  #   ggplot(aes(lon, lat, fill = hab)) +
+  #   geom_tile()
+  #
+
+  x_binsize <- (max(hab$lon) - min(hab$lon) + 1) / resolution
+
+  y_binsize <- (max(hab$lat) - min(hab$lat) + 1) / resolution
+
+  if (x_binsize < 1 | y_binsize < 1) {
+    stop("resolution is too high for input data")
+  }
+
+  hab <- hab %>%
+    mutate(X = lon - min(lon),
+           Y = lat - min(lat)) %>%
+    mutate(rough_x = floor(X / x_binsize),
+           rough_y = floor(Y / y_binsize)) %>%
+    select(contains("rough_"), hab)
+  #
+  #   hab %>%
+  #     ggplot(aes(rough_x, rough_y, fill = hab)) +
+  #     geom_tile()
+
+  hab <- hab %>%
+    group_by(rough_x, rough_y) %>%
+    summarise(hab = sum(hab)) %>%
+    mutate(xy = rough_x * rough_y)
+
+  mod <-
+    gamm4::gamm4(hab ~ s(rough_x) + s(rough_y) + s(xy), data = hab)
+
+  hab$interp_cpue <- as.numeric(predict(mod$gam))
+
+  hab$habitat <- pmax(0, hab$interp_cpue / max(hab$interp_cpue))
+
+  if (sqrt(nrow(hab)) != resolution) {
+    stop("habitat layer does not match simulation resolution")
+  }
+  hab <- hab %>%
+    select(contains("rough"), habitat) %>%
+    rename(x = rough_x, y = rough_y) %>%
+    ungroup()
+
+  # hab %>%
+  #   ggplot(aes(rough_x, rough_y, fill = habitat)) +
+  #   geom_tile()
+
+  out <- tibble(scientific_name = species, habitat = list(hab))
+
+}
+
+mats <- map(mats, get_layer)
+
+sdm_habitat <- mats %>%
+  bind_rows() %>%
+  unnest(cols = habitat) %>%
+  filter(scientific_name %in% unique(catch_mats$scientific_name)) %>% 
+  ggplot(aes(x, y, fill = habitat)) +
+  geom_tile() +
+  facet_wrap( ~ scientific_name) +
+  scale_fill_viridis_c()
+
+sdm_habitat
+
+catch_habitat / sdm_habitat
+```

### 2021-09-15 R/optimize-mpa.R

@@ -77,11 +77,6 @@ optimize_mpa <-
         res <-
           sim_mpa[[length(sim_mpa)]] # for now, just calculate in the final timestep
 
-        if (objective == "max_ssb"){
-        biodiv <-
-          sum(map_dbl(res, ~ sum(.x$ssb_p_a) / .x$ssb0)) # calculate biodiversity component of objective function
-        } else if (objective == "min_loss") {
-
         biodiv_mpa <-
           (map_dbl(res, ~ sum(.x$ssb_p_a) / .x$ssb0)) # calculate biodiversity component of objective function
 
@@ -89,15 +84,36 @@ optimize_mpa <-
           map_dbl(starting_conditions[[1]], ~ sum(.x$ssb_p_a) / .x$ssb0) # calculate biodiversity component of objective function
 
 
+        delta_biodiv <- biodiv_mpa - biodiv_sq
+
+        if (objective == "max_ssb"){
+        biodiv <-
+          sum(map_dbl(res, ~ sum(.x$ssb_p_a) / .x$ssb0)) # calculate biodiversity component of objective function
+        } else if (objective == "min_loss") {
+
+
         biodiv <- sum((biodiv_mpa - biodiv_sq) >= 0)
 
         }
 
-
         # econ <- sum(map_dbl(res, ~sum(.x$c_p_a))) #  calculate econ component of objective function
 
-        econ <-
-          sum(map_dbl(res, ~ sum(.x$r_p_a_fl, na.rm = TRUE))) #  calculate econ component of objective function, currently revenues across all fleets and species
+        econ_sq <-
+          (map_dbl(starting_conditions[[1]], ~ sum(.x$r_p_a_fl, na.rm = TRUE)))
+
+        econ_mpa <-
+          (map_dbl(res, ~ sum(.x$r_p_a_fl, na.rm = TRUE))) #  calculate econ component of objective function, currently revenues across all fleets and species
+
+        delta_econ <- econ_mpa - econ_sq
+
+        econ <- delta_econ
+
+        econ[delta_econ > 0 & delta_biodiv < 0] <- 0
+
+        econ <- sum(econ)
+
+        # econ <-
+        #   sum(map_dbl(res, ~ sum(.x$r_p_a_fl, na.rm = TRUE))) #  calculate econ component of objective function, currently revenues across all fleets and species
 
         out <- tibble(biodiv = biodiv, econ = econ)
 
[…]

### 2021-09-15 scripts/00_setup.R

@@ -34,7 +34,7 @@ foos <- list.files(here("R"))
 
 walk(foos, ~ source(here("R", .x)))
 
-results_name <- "tester"
+results_name <- "v0.9"
 
 results_path <- here("results", results_name)
 
@@ -46,11 +46,11 @@ if (!dir.exists(results_path)){
 
 run_experiments <- TRUE
 
-experiment_workers <- 4
+experiment_workers <- 8
 
 run_casestudy <- FALSE
 
-optimize_casestudy <- FALSE
+optimize_casestudy <- TRUE
 
 safety_stop <- FALSE
 

### 2021-09-20 documents/marlin-paper-notes.Rmd
safety stop, things working but tricky to think through optimization

@@
 ```
+
+# RPUE vs REvenue vs PPUE
+
+So there are clearly some differences between the outcomes with the three fleet models.
+
+when the fleet distributes itself based on RPUE, as you start to fish things down the "best" patches start to look worse relative to the worst patches.
+This is because as you start to allocate more and more effort towards the high revenye patches, ther marginal benefit goes down, since the marginal increase in f per unit increase in E starts to decrease.
+So, the best revenye patch is 53 times better than the worst, but the best RPUE patch is only 7 time bigger than the worst.
+
+So RPUE is closer to PPUE, since it captures the effect that the demonitator starts to grow faster than the numerator at high E.
+Whereas with just revenue, as you increase effort the value of that patch just keeps going up, assuming a constant supply of fish.
+So you're including the "stock effect" sort of, and then with PPUE you add in the fact that not only do you get diminishing returns in E, but it costs you money to exert E.
+
+The revenue model says that even if it takes you infinite effort to get 101 units, and 1 effort to get 100 units, go to 101 units.
+
+So that bets two questions. 1 is there any logical reason for revenue in a real world setting where it would be distributed by revenue, not by RPUE?
+
+A fleet's objective function is maximize profits, either individually or in total.
+Let's think about in total then.
+The fleet will pick the the total amount of effort that maximizes profits, conditional on distributing those profits to mazimize profits per unit effort, right?
+
+The other is, did the solving the increased revenus through overfishing by mPA get solved by the open access or by the fleet model?
+GOod test, set alpha = 0, set fleet to RPUE, but also open access, see if that fixes it.
+
+If it doesn't another thing to consider it to modify the objective function to maximize RPUE or something like that to try and reflect costs
+
+the fundamental issues is that the fleet model and the optimization can have two different objectives.
+When you allocate around RPUE and optimize around revenue, the fleet is "suboptimal" since it allocqtes a bit less effort to the best patches, since their RPUE is a little lower thanks to the baranov.
+
+When you allocate by revenue you are least being internally consistent, i.e.
+the fleet is just maximizing revenue, and then you optimize around that.
+So, that seems like the best option until you figure out how to deal with profits / costs intelligently.
+
+Which is a strategy: shift the optimization to be around profits not revenue, which because you are assuming constant effort unless you include an exponent on effort, which causes a whole new set of problems since minute differences in the spatial distribution can change the outcomes, then optimizing around RPUE should be the same as optimizing around profits?
+So changing the optimization to profits should do nothing.
+becuase it's just going to be the same thing, except minus cost * constant effort.
+
+The only reason anyof this really matters is it if has a big impact on the bycatch implications.
+So, go back and run the experiments.
+If the revenue vs rpue this is intrinsic, then you need to consider it.
+If it's just a trait ot he WCPO scenario, then interestinh and flag and use this to discuss the critical need for better fleet modeling with Jim.
+
+Fishing the line is the culprit i think.
+in order for unintended consequences to occue, species have to live in different places.
+when you disteirbute by revenue, fishing the line is much more pronouned: people stick by the MPA edge.
+But, when you do RPUE, the near-MPA area starts to lose its appeal after a bit and people wander to new areas, causing problems.

### 2021-09-20 R/compile-fleet.R

@@ -136,7 +136,7 @@ compile_fleet <- function(fauna, tune_type = "explt") {
           sel_unit = "length"
         )
       ),
-      base_effort = resolution
+      base_effort = resolution^2
     ),
     "purseseine" = create_fleet(
       list(
@@ -272,7 +272,7 @@ compile_fleet <- function(fauna, tune_type = "explt") {
           sel_unit = "length"
         )
       ),
-      base_effort = resolution
+      base_effort = resolution^2
     )
   )
 

### 2021-10-02 documents/marlin-paper-notes.Rmd
holy cow. things are back working again

@@
 But, when you do RPUE, the near-MPA area starts to lose its appeal after a bit and people wander to new areas, causing problems.
+
+So, I think the solution is
+
+1. store profit per fleet per patch in simmar so you can stop having to wrangle it
+
+2. optimize around profits
+
+3. Distribute fleet based on profits
+
+4. Turn off open access
+
+This means that if you assume 0 profits at start, then you can only go down profit wise from overfished, but could in theory go up from rebuilding.
+
+So, that means that you're capturing the assumption that there's a reason why things are underfished now, but still allows for rebulding benefits
+
+In order for this to have any spatial effect you need to have an exponent on effort cost or something like that.
+Otherwise, maximizing revenue is the exact same thing as maximizing profits - a constant (base_effort * cost_per_unit_effort).
+
+All of this serves two purposes. 1st to make things internally consistent for the optimization and the fleet model. 2nd to stop the "optimal" thing being to corral the fleet into catching more. \
+
+An alternative would be to introduce a port and include a distance from fleet thing?
+But I think you can rationalize the exponent based on competition, gear competition (e.g.
+need to pull in longline and reset a bunch), and hold capacity?
+though does that really hold in the real world?
+I think it would have to be things like longer soak times needed?
+
+The other thing that matter here is how you are thinking about the market.
+If it's one mobile fleet  then it makes sense ot pool revenues and costs everywhere.
+But, if it were a bunch of independent fleets in each patch, that get some displaced into them, then the decreased profit margins in indiivual patches start to matter
+
+OK the only thing clear in this is that you do nede to actually run the damn counterfactual, it doesn't take long soj ust do it.
+
+sigh ok calling it.
+But for tomorrow, the key is to go back into simmar, and calculate profits by species, by first tallying fleet wide profits, and then dividing by the proportino of revenue by species
+
+Within the simulation model, if the spatial allocation was RPUE, then effort in the current time step would be exactly proportional to RPUE in the prior time step.
+
+You don't see that in the data.
+You do see a strong correlation with revenue in the prior time step.
+Really hard to say if that is causal, i.e.
+are fleets chasing revenue or is it a site effect or what, but it's at elast a piece of evidence.
+So, I think you go with revenue in the model, and use the experiments to make the case in the discussion the discussion that any tactical modeling needs to understand how the fleet actually responds and distributes itself.
+
+Choice depends a lot on the unts in question. IF you think that every unit of effort is discrete and per vessel, i.e. vessel I does or does not fish in that patch, then RPUE woul be the metric

### 2021-10-02 R/compile_experiment_fleet.R

@@ -1,4 +1,5 @@
-compile_experiment_fleet <- function(fauna, state, tune_type = "explt") {
+compile_experiment_fleet <- function(fauna, state, tune_type = "explt",
+                                     effort_cost_exponent = 1.3) {
 
 
   if (all(state$spatial_q == TRUE)) {
@@ -57,7 +58,7 @@ compile_experiment_fleet <- function(fauna, state, tune_type = "explt") {
       ),
       "carcharhinus longimanus" = Metier$new(
         critter = fauna$`carcharhinus longimanus`,
-        price = 1.89,
+        price = 0,
         sel_form = "logistic",
         sel_start = fauna$`carcharhinus longimanus`$length_50_mature *
           sample(c(0.9), 1, replace = TRUE),
@@ -68,8 +69,10 @@ compile_experiment_fleet <- function(fauna, state, tune_type = "explt") {
         spatial_catchability = shark_spatial_q
       )
     ),
-    spatial_allocation = state$spatial_allocation[[1]],
-    base_effort = resolution^2
+    base_effort = resolution^2,
+    cost_per_unit_effort = 10,
+    effort_cost_exponent = effort_cost_exponent,
+    spatial_allocation = state$spatial_allocation[1]
   ))
 
   # a <- Sys.time()

### 2021-10-02 R/run_mpa_experiment.R

@@ -16,52 +16,8 @@ run_mpa_experiment <-
 
     # set up open access
 
-    starting <- starting_conditions[[length(starting_conditions)]]
+    # starting <- starting_conditions[[length(starting_conditions)]]
 
-    revenues <-  map_df(starting, ~.x$r_p_a_fl %>%
-                          reshape2::melt() %>%
-                          group_by(Var3) %>%
-                          summarise(revenue = sum(value, na.rm = TRUE))) %>%
-      group_by(Var3) %>%
-      rename(fleet = Var3) %>%
-      summarise(revenue = sum(revenue))
-
-    effort <- map_df(starting, ~.x$e_p_fl) %>%
-      ungroup() %>%
-      mutate(patch = 1:nrow(.)) %>%
-      pivot_longer(-patch, names_to = "fleet", values_to = "effort") %>%
-      group_by(fleet) %>%
-      summarise(effort = sum(effort) / length(fauna))
-
-
-    profits <- revenues %>%
-      left_join(effort, by = "fleet") %>%
-      mutate(cost = revenue / effort) %>%
-      mutate(profit = revenue - cost * effort)
-
-    max_rev <- map_dbl(fauna, "ssb0")
-
-    prices = pluck(fleets, 1,1) %>%
-      map_dbl("price")
-
-    max_p <- sum(max_rev * prices[names(max_rev)])
-
-    profits <- profits %>%
-      mutate(theta = log((1 + max_delta)) / (max_p))
-
-    fleets$longline$cost_per_unit_effort <- profits$cost[profits$fleet == "longline"]
-
-    fleets$longline$profit_sensitivity <- profits$theta[profits$fleet == "longline"]
-
-    fleets$longline$fleet_model <- "open access"
-
-    # fleets$purseseine$cost_per_unit_effort <- profits$cost[profits$fleet == "purseseine"]
-    #
-    # fleets$purseseine$profit_sensitivity <- profits$theta[profits$fleet == "purseseine"]
-
-    # fleets$purseseine$fleet_model <- "open access"
-
-    # assign objective score to each cell
 
     critters_considered <- sample(names(fauna), round(prop_critters_considered * n_distinct(names(fauna))), replace = FALSE)
 
@@ -69,7 +25,7 @@ run_mpa_experiment <-
 
       # place MPAs in proportion to depletion-weighted spawning stock biomass
       depletion <-
[…]

### 2021-10-02 scripts/01_run_experiments.R

@@ -3,7 +3,7 @@ source(file.path("scripts", "00_setup.R"))
 
 # generate state experiments. This is a somewhat tricky process where the actual generated values are created for many variables in create_experiment_critters
 
-n_states <- 250
+n_states <- 500
 
 if (run_experiments) {
   state_experiments <-
@@ -11,7 +11,7 @@ if (run_experiments) {
       sigma_centroid = runif(n_states, .1 * resolution, resolution ^ 2 / 2),
       sigma_hab = runif(n_states, .1 * resolution, resolution ^ 2 / 8),
       spatial_q = sample(c(FALSE,TRUE), n_states, replace = TRUE, prob = c(3,1)),
-      spatial_allocation = sample(c("revenue","rpue"), n_states, replace = TRUE)
+      spatial_allocation = sample(c("ppue","rpue", "revenue",'profit'), n_states, replace = TRUE)
     ) %>%
     mutate(state_id = 1:nrow(.))
 
@@ -176,7 +176,6 @@ if (run_experiments) {
     mutate(fleet = map2(fauna,data, compile_experiment_fleet))
 
   # add in starting conditions
-
   init_condit <- function(fauna, fleets, years = 100) {
     starting_trajectory <-
       simmar(fauna = fauna,
@@ -192,13 +191,75 @@ if (run_experiments) {
     proc_starting_conditions <-
       process_marlin(starting_conditions, keep_age = FALSE)
 
-    out <- list(starting_conditions = starting_conditions,
+    out <- list(starting_conditions = starting_conditions[[1]],
                 proc_starting_conditions = proc_starting_conditions)
 
   }
 
   state_experiments <- state_experiments %>%
     mutate(tmp = map2(fauna, fleet, init_condit))
+
+  state_experiments$starting_conditions <-
+    map(state_experiments$tmp, "starting_conditions")
+
+  assign_costs <- function(init_condit, fleets, fauna){
+
+    starting <- init_condit
+
+    revenues <-  map_df(starting, ~.x$r_p_a_fl %>%
+                          reshape2::melt() %>%
+                          group_by(Var3) %>%
+                          summarise(revenue = sum(value, na.rm = TRUE))) %>%
+      group_by(Var3) %>%
[…]

### 2021-10-02 scripts/02_run_case_study.R

@@ -197,13 +197,77 @@ fleets <- casestudy$fleet[[1]]
 starting_trajectory <- simmar(fauna = fauna, fleets = fleets, years = 200)
 
 
-starting_conditions <- starting_trajectory[length(starting_trajectory)]
+starting <- starting_trajectory[[length(starting_trajectory)]]
 
+revenues <-  map_df(starting, ~.x$r_p_a_fl %>%
+                      reshape2::melt() %>%
+                      group_by(Var3) %>%
+                      summarise(revenue = sum(value, na.rm = TRUE))) %>%
+  group_by(Var3) %>%
+  rename(fleet = Var3) %>%
+  summarise(revenue = sum(revenue))
+
+# effort <- map_df(starting, ~.x$e_p_fl) %>%
+#   ungroup() %>%
+#   mutate(patch = 1:nrow(.)) %>%
+#   pivot_longer(-patch, names_to = "fleet", values_to = "effort") %>%
+#   group_by(fleet) %>%
+#   summarise(e2 = sum((effort / length(fauna))^1),
+#             ee = sum(effort / length(fauna)))
+
+effort <- map_df(starting[1], ~.x$e_p_fl) %>%
+  ungroup() %>%
+  mutate(patch = 1:nrow(.)) %>%
+  pivot_longer(-patch, names_to = "fleet", values_to = "effort") %>%
+  group_by(fleet) %>%
+  summarise(e2 = sum(((effort + 1)^fleets[[1]]$effort_cost_exponent)),
+            check = sum(effort))
+
+
+# revenues$revenue <- 339613.9
+
+profits <- revenues %>%
+  left_join(effort, by = "fleet") %>%
+  mutate(cost = revenue / e2) %>%
+  mutate(profit = revenue - cost * e2)
+
+max_rev <- map_dbl(fauna, "ssb0")
+
+prices = pluck(fleets, 1,1) %>%
+  map_dbl("price")
+
+max_p <- sum(max_rev * prices[names(max_rev)])
+
+profits <- profits %>%
+  mutate(theta = log((1 + max_delta)) / (max_p))
+
+fleets$longline$cost_per_unit_effort <- profits$cost[profits$fleet == "longline"]
+
+fleets$longline$profit_sensitivity <- profits$theta[profits$fleet == "longline"]
+
+fleets$purseseine$spatial_allocation <- "profit"
+
+fleets$purseseine$cost_per_unit_effort <- profits$cost[profits$fleet == "purseseine"]
+
[…]

### 2021-10-04 documents/03_marlin_paper.Rmd
testing reverse optimization

@@
 
-group_var <- "placement_strategy"
-
-results <- placement_experiments %>%
-  mutate(temp = raw_experiment_results) %>%
-  mutate(state_id = map(raw_experiment_results, names)) %>%
-  unnest(cols = c(temp, state_id)) %>%
-  mutate(obj = map(temp, "obj"),
-         state_id = as.numeric(state_id)) %>%
-  unnest(cols = obj) %>%
-  mutate(
-    critter = fct_recode(
-      critter,
-      "Shark" = "carcharhinus longimanus",
-      "Tuna" = "katsuwonus pelamis",
-      "Marlin" = "kajikia audax"
-    )
-  )
-
-if (any(results$econ[results$prop_mpa == 1] > 0)) {
-  stop("something has gone wrong: yields present with 100% mpa")
-}
-
-# compare to BAU
-
-results <- results %>%
-  group_by(placement_id, state_id, critter) %>%
-  mutate(bau_biodiv = biodiv[prop_mpa == 0],
-         bau_econ = econ[prop_mpa == 0])
-
-excols <-
-  state_experiments %>%
-  select(state_id, data) %>%
-  unnest(cols = data) %>%
-  select(state_id:sigma_hab) %>%
-  ungroup() %>%
-  unique()
-
-total_results <- results %>%
-  group_by(
-    placement_id,
-    placement_strategy,
-    prop_critters_considered,
-    placement_error,
-    state_id,
-    prop_mpa
-  ) %>%
-  summarise(
-    loss = mean(biodiv < bau_biodiv),
-    n_loss = sum(biodiv < bau_biodiv),
-    signif_loss = mean(biodiv < (0.9 * bau_biodiv)),
-    total_loss = sum(biodiv - bau_biodiv),
-    econ = sum(econ),
-    biodiv = sum(biodiv),
-    bau_econ = sum(bau_econ),
-    bau_biodiv = sum(bau_biodiv)
-  ) %>%
-  ungroup() %>%
-  mutate(delta_biodiv = biodiv / bau_biodiv - 1,
-         abs_delta_biodiv = biodiv - bau_biodiv,
-         delta_econ = econ / bau_econ - 1) %>%
-  left_join(excols, by = "state_id")
 ```
@@
 
+### Fleet Dynamics
+
+The key feature of `marlin` for the purposes of this project is the
+ability to capture technical interactions across species. While marlin
+does not simulate species interactions, fishing fleets in marlin are
+capable of targeting and affecting multiple species simultaneously. This
+allows us to capture for example the potential impacts of displacement
+of effort from one area into another across multiple species.
+
+A fishing fleet or *metier* is defined in `marlin` by a total pool of
+effort that can have different catchability, prices, and selectivity for
+individual species within a simulation. For example, both a longline and
+purse seine fleet may capture bigeye (*Thunnus obesus*) and skipjack
+(*Katswamus pelamus*), but the longline fleet is much more likely to
+catch bigeye than skipjack, and *vice versa*. The fishing fleet then
+generates catches, revenues, and profits from individual species that it
+targets, and makes decisions based on its total objective function
+across all species.
+
+Revenues *R* for fleet *f* in time step *t* are given by
+
+$$
+ R_{t,p,f} = \sum_s^S \sum_a^A \Pi_{s,f}\frac{\alpha_{a,f}q_{s,f}E_{t,p,f}}{F_{t,p,s,a}} \times Y_{t,p,s,a}
+$$
+
+where *p* is patch, *s* is species, *a* is age, $\alpha$ is selectivity
+at age, *q* is catchability , *F* is total fishing mortality, and *Y*
+are total catches.
+
+$$
+Y_{t,p,s,a} = \frac{F_{t,p,s,a}}{z_{t,p,s,a}} \times b_{t,p,s,a} \times (1 - e^{-z_{t,p,s,a}}) 
+$$
+
+where *F* is total fishing mortality, given by
+
+$$
+F_{t,p,s,a} = \sum_f^F\alpha_{a,f}q_{s,f}E_{t,p,f}
+$$
+
+and
+
+$$
+ z_{t,p,s,a} = F_{t,p,s,a} + m_{s,a}
+$$
+
+Where *m* is natural mortality.
+
+This means that each species experiences a total mortality at age in a
+time step in a patch, individual fractions of which are portioned off as
+catches and subsequently revenues for each fleet. Given revenues, we
+then calculate profits \$\\phi\$ per
+
+$$
+\phi_{t,p,f} = R_{t,p,f} - c_f(E_{t,p,f} + 1)_{t,p,f}
+^{1.3}$$
+
+where *c* is cost per unit effort and, the 1.3 exponent is in the manner
+of @costello2016a . The addition of one to the total effort is to
+constrain the value to be greater than 1, preventing undesired behavior
+where fishing effort actually gets cheaper if total effort is less than
+1.
+
+The important thing to note here then is that the revenues and profits
+for a given fleet are a function of the prices and catchability of
+individual species, but that mortality itself is independent of price.
+So, fleet that has selectivity and catchability greater than zero can
+exert fishing mortality on a species even if it obtains no value from
+that species.
+
+The fleet then decides how to allocate its effort in space based on one
+of four possible spatial allocation strategies. The ideal free
+distribution (IFD) (citation) is the standard method for distributing
+fishing fleets in spaces. However, the IFD presents a number of
+complications for our study. First, in order to simplify the assessment
+of the impacts of different fishing mortality rates we simulate
+scenarios with a fixed total pool of effort. As MPAs change the spatial
+distribution of profits though, the total amount of effort required to
+satisfy the IFD may not equal the total amount of effort required by the
+model. Second, even if we were to implement the IFD for a single fleet,
+the model is capable of tracking multiple fleets at once. This means
+that the IFD for a single fleet would be a game-theoretic outcome of the
+actions of each of the individual fleet conditional on the actions of
+all other fleets. While possible to implement this would slow down our
+model runs to the point of practically preventing large-scale evaluation
+of the effects of MPAs on bycatch.
+
+As such we explored a series of "next best" fleet distribution
+algorithms. While not the IFD in any individual time step, over time
+they start to approximate the IFD (check this) as each assumes that
+fleets base their decisions for the current time step on the outcomes in
+the prior timestep, meaning that the impacts of the actions of other
+fishing fleets are somewhat internalized. The four possible fleet
+distribution algorithms are
+
+1.  Revenue per unit effort (RPUE): The fleet distributes itself in
+    space based on revenue per unit effort
+2.  Revenue: The fleet distributes itself in space based on revenue
+3.  Profit per unit effort (PPUE): The fleet distributes itself in space
+    based on profit per unit effort
+4.  Profit: The fleet distributes itself in space based on profit
+
+### Population Model
+
 Lots of equations. Pretty standard age-structured population models,
@@
 
-## Illustrative Experiments
+## Generalized Experiments
 
@@
 
-## WCPO Case Study
+## Case Study Application
 
@@
 
-## Illustrative Experiments
+what we found was a series of neccesary but not sufficient conditions
+
+make the tautology clear: illustrating a point, not proving a thing since clearly depends on if you assume displacement, and then degree of non-overlap etc.
+
+## Generalized Experiments
 
@@
 benefits across all three species, and a second where MPAs actually
-produced substantial losses in some species....
+produced substantial losses in some species. In the "best case"
+scenario, each of the three species shares a common habitat
+distribution, and so protection of the central habitat of one species
+results in protection for all species. In the "worst case" scenario,
+each of the species have markedly different centroids. An MPA designed
+primarily around the tuna and marlin populations then ends up displacing
+effort into the shark habitat, producing benefits for the first two
+species but driving the shark population down.
 
@@
 
+knitr::include_graphics(file.path(results_path,"fig_info.pdf"))
 ```
 
+the point is not once answer
+
 Looking at the full
-`r n_distinct(total_results$placement_id) * n_distinct(total_results$state_id)`
+`r n_distinct(total_experiment_results$placement_id) * n_distinct(total_experiment_results$state_id)`
 experiments run, the results were highly variable, with MPAs capable of
@@
 
-Make clear that a goal of all this complexity was to see if there were
-any clear emergent properties (this thing matters way more than this
-thing). We highlight some of those, but given the complexity we found we
-present the complexity of interacting ways in which different things can
-effect the conservation outcomes of MPAs
-
-XX after looking through these should widdle it down to only things you
-actually examine, annoying / no point to having a bunch of unexplained
-minor variation in the mix relatede to things like ontogenetic shifts)
-xx
-
-xx the problem here. These things all interact in so many complex ways
-that it's hard to look at either graphically or statistically, and maybe
-for the best. i.e. the answer isn't "if habitat overlap is less than XX%
-you might have a problem, otherwise you're fine", but rather that a
-bunch of things can interact to cause potential problems in complex
-ways. but that's hard to explain xx.
-
-Big message: even constrained by the same initial depletion and core
+\\Big message: even constrained by the same initial depletion and core
 biology (e.g. movement rates), a wide range of outcomes are possible for
@@
 than iterating over different movement dynamics and habitat
-configurations, we fix habitat layers on the values predicted by
-@bradley2021, and xx other things we fix xx.
+configurations, we fix habitat layers based on reported CPUE for the
+species in question, and xx other things we fix xx.
 

### 2021-10-04 R/optimize_mpa.R

@@ -142,7 +142,7 @@ optimize_mpa <-
           fauna = fauna,
           fleets = fleets,
           mpas = mpas,
-          starting_conditions = starting,
+          starting_conditions = starting_conditions,
           .options = furrr_options(seed = 42),
           .progress = FALSE
         )

### 2021-10-04 scripts/01_run_experiments.R

@@ -281,7 +281,7 @@ if (run_experiments) {
     ),
     prop_mpa = seq(0, 1, by = 0.05),
     prop_critters_considered = 1,
-    placement_error = c(0)
+    placement_error = c(0,.2)
   ) %>%
     group_by_at(colnames(.)[!colnames(.) %in% c("temp", "prop_mpa")]) %>%
     nest() %>%
@@ -483,6 +483,72 @@ if (any(results$econ[results$prop_mpa == 1] > 0)) {
   stop("something has gone wrong: yields present with 100% mpa")
 }
 
+
+states <- state_experiments %>%
+  select(state_id, data) %>%
+  unnest(cols= data) %>%
+  mutate(
+    critter = fct_recode(
+      scientific_name,
+      "Shark" = "carcharhinus longimanus",
+      "Tuna" = "katsuwonus pelamis",
+      "Marlin" = "kajikia audax"
+    )
+  )
+
+
+
+calc_centroid_distance <- function(tmp_state){
+
+  distance_matrix <-
+    expand_grid(x = 1:resolution, y = 1:resolution) %>%
+    dist() %>%
+    as.matrix()
+
+  centroid_distances <- expand_grid(x = tmp_state$centroid, y = tmp_state$centroid) %>%
+    rowwise() %>%
+    mutate(distance = distance_matrix[x,y])
+
+  cd <- sum(centroid_distances$distance)
+
+}
+
+calc_sigma_habitat <- function(tmp_state){
+
+
+  habitat <- tmp_state %>%
+    select(scientific_name, contains("sigma_"), habitat) %>%
+    unnest(cols = habitat) %>%
+    mutate(
+      scientific_name = fct_recode(
+        scientific_name,
+        "Shark" = "carcharhinus longimanus",
+        "Tuna" = "katsuwonus pelamis",
+        "Marlin" = "kajikia audax"
+      )
+    ) %>%
+    select(x,y,scientific_name,habitat) %>%
+    pivot_wider(names_from = scientific_name, values_from = habitat) %>%
+    select(-x,-y) %>%
+    as.matrix()
+
+
+  sigma_habitat <- sd(sqrt(habitat))
+
+}
+
+
+
+states <- states %>%
+  group_by(state_id) %>%
+  nest() %>%
+  mutate(centroid_distance = map_dbl(data,calc_centroid_distance),
+         sigma_habitat = map_dbl(data,calc_sigma_habitat)) %>%
+  unnest(cols = data)
+
+
[…]

### 2021-10-04 scripts/02_run_case_study.R

@@ -277,7 +277,7 @@ if (run_casestudy == TRUE){
       placement_strategy = c("depletion", "rate", "avoid_fishing", "target_fishing", "area"),
       prop_mpa = seq(0,.99, by = 0.01),
       prop_critters_considered = 1,
-      placement_error = 0
+      placement_error = c(0,.2)
     )
 
   a <- Sys.time()
@@ -311,7 +311,6 @@ write_rds(case_study_experiments, file = file.path(results_path, "case_study_exp
 
 
 # optimize network --------------------------------------------------------
-
 if (optimize_casestudy == TRUE){
 
   optimized_networks <- tibble(alpha = seq(0,1, length.out = 20))
@@ -326,8 +325,8 @@ if (optimize_casestudy == TRUE){
         optimize_mpa,
         fauna = fauna,
         fleets = fleets,
-        max_prop_mpa = .95,
-        prop_sampled = 0.2,
+        max_prop_mpa = 1,
+        prop_sampled = 0.25,
         starting_conditions = starting_conditions,
         resolution = resolution,
         workers = experiment_workers
@@ -558,7 +557,8 @@ bau_econ <-
 #  calculate econ component of objective function, currently revenues across all fleets and species
 
 bau <- bau_econ %>%
-  left_join(bau_biodiv, by = "critter")
+  left_join(bau_biodiv, by = "critter") %>%
+  mutate(bau_econ = bau_econ / 1e6)
 
 
 # calculate objective function for experiments
@@ -566,7 +566,8 @@ bau <- bau_econ %>%
 experiment_obj <- case_study_experiments %>%
   mutate(obj = map(results, "obj")) %>%
   select(-results) %>%
-  unnest(cols = obj)
+  unnest(cols = obj) %>%
+  mutate(econ = econ / 1e6)
 
 experiment_obj %>%
   ggplot(aes(biodiv, econ, color = placement_strategy)) +
@@ -589,21 +590,16 @@ experiment_obj %>%
 experiment_obj %>%
   ggplot(aes(prop_mpa, econ, color = placement_strategy)) +
   geom_point() +
-  facet_wrap(~critter, scales = "free_y")
+  facet_wrap( ~ critter, scales = "free_y")
 
-
-
-experiment_obj %>%
[…]

### 2021-10-07 documents/03_marlin_paper.Rmd
adding in hyperallometry

@@
 title: "Solutions to the potential for unintended conservation consequences of MPAs"
+subtitle: Conditions for success and unintended consequences in pelagic MPA design
 author:
@@
 producing substantial conservation gains to substantial conservation
-losses and everything in between. We do not focus into the individual
+losses and everything in between.
+
+Design matters size matters species matters habitat matters
+```{r}
+
+compact_experiment_results %>%
+  filter(prop_mpa < 1) %>%
+  # filter(f_v_m > 1.5) %>%
+  group_by(state_id, placement_id) %>%
+  ggplot(aes(
+    prop_mpa,
+    pmin(2,biodiv / bau_biodiv -1),
+    color =  sigma_habitat,
+    group = interaction(placement_id, state_id)
+  )) +
+  geom_line(alpha = 0.25) +
+  geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
+  facet_grid(critter ~ placement_strategy) +
+  scale_color_viridis_c(name = "Habitat Difference", guide = guide_colorbar(frame.colour = "black",ticks.colour = "black")) +
+  scale_y_continuous(labels = scales::percent,"Change in SSB/SSB0 relative to BAU") +
+  scale_x_continuous(labels = scales::label_percent(accuracy = 1), name = "MPA Size") +
+  theme(legend.position = "top")
+
+```
+
+```{r}
+
+critter_results <- compact_experiment_results %>%
+  group_by(state_id, placement_id) %>%
+  mutate(real_sigma_centroid = sd(centroid)) %>%
+  mutate(delta_biodiv = pmin(2,biodiv / bau_biodiv - 1),
+         loss = biodiv < bau_biodiv,
+         rec_form = as.factor(rec_form)) %>% 
+  mutate(critter = fct_relevel(critter,c("Tuna", "Marlin","Shark")),
+         rec_form = fct_recode(rec_form, "Global DD, Rec. By Habitat" = "0", "Local DD, local Recruitment" = "1", "Local DD, Disperse Recruits" = "2", "Disperse Larvae, Poset-Settlement DD" = "3"))
+
+reg_2 <-   rstanarm::stan_glm(
+  loss ~   f_v_m  + critter + seasonal_movement + adult_movement_sigma + rec_form + recruit_movement_sigma + prop_mpa + sigma_habitat + spatial_allocation,
+  family = binomial(),
+  data = critter_results %>% filter(between(prop_mpa,.2,.4), placement_strategy == "area"),
+  cores = 4,
+  chains = 4
+)
+
+fig_11 <- plot(reg_2)
+
+fig_11
+
+```
+
+We do not focus into the individual
 effects of particular variables as we found them to be highly dependent
@@
 The second is that design matters...
+
+```{r}
+
+```
 

### 2021-10-07 scripts/00_setup.R

@@ -34,7 +34,7 @@ foos <- list.files(here("R"))
 
 walk(foos, ~ source(here("R", .x)))
 
-results_name <- "v1.0"
+results_name <- "v1.1"
 
 results_path <- here("results", results_name)
 
@@ -46,7 +46,7 @@ if (!dir.exists(results_path)){
 
 run_experiments <- FALSE
 
-experiment_workers <- 8
+experiment_workers <- 10
 
 run_casestudy <- FALSE
 

### 2021-10-09 documents/03_marlin_paper.Rmd
OK, looking good. Knits with all base figure versions. back in the morning

--- 2021-10-04 documents/03_marlin_paper.Rmd
@@
 params:
-  run_name: ["v0.5"]
+  run_name: ["v1.1"]
 ---
@@
 
+sci_to_com <- taxize::sci2comm(unique(marlin_inputs$scientific_name)) %>% 
+  bind_rows(.id = "critter") %>% 
+  pivot_longer(everything(),names_to = "critter",values_to = "common_name") %>% 
+  mutate(common_name = stringr::str_to_title(common_name))
+
 ```
@@
 ```
+
+c
 
@@
 
-Marine protected areas (MPAs) are a growing part of the marine
+in the Marine protected areas (MPAs) are a growing part of the marine
 conservation toolbox, with 5.9% of the world's oceans currently in
@@
 
-what we found was a series of neccesary but not sufficient conditions
-
-make the tautology clear: illustrating a point, not proving a thing since clearly depends on if you assume displacement, and then degree of non-overlap etc.
+Our simulation experiments allow us to examine the wide range of effects
+that MPAs can have on groups of species depending on combinations of
+ecological, economic, and design parameters. The goal of this exercise
+is not to assign any probability to the potential impacts of MPAs; we
+have not provided any explicit probabilites to the alternative states
+modeled here, and so it may be that a state that was only modeled on a
+few occasions in our simulations is in fact highly common in the real
+world.
+
+What these experiments are designed to do is illustrate the range of
+outcomes that are possible from MPAs under the types of conditions
+evaluated here, and evaluate which drivers are potentially the most
+influential in affecting particular outcomes, such as unintended
+conservation consequences. Our case study of a WCPO-esque system is
+intended then to reduce the parameter space enough that we can examine
+the relative tradeoffs and value of optimization for economic and
+conservation objectives in a pelagic system.
+
+One clear caveat across all of these results is that we have made two
+very strong assumptions around the behavior of the fishing fleet. The
+first is that we assume that effort is displaced, not removed, by the
+creation of MPAs. This is a widely debated question around MPAs, and the
+real impacts of MPAs on fishing effort will clearly be setting
+dependent. For example, large-scale pelagic fishing fleets already have
+high rates of vessel movement built into their economic models, and so
+may be more likely to move rather than exit the fishery in response to
+MPAs. In contrast, smaller scale coastal fishing operations may simply
+exit the fishery if their local fishing grounds are placed in MPAs.
+There is clear evidence of effort displacement in many fisheries though
+(citation), and moreover a motivation for this study was to evaluate the
+potential unintended consequences of MPAs from effort displacement on
+target and bycatch species. If MPAs do not produce increases in fishing
+effort outside their borders, then the conservation effects of MPAs have
+to be positive for all species concerned.
+
+The second is that we do not account for evolution of total effort. Our
+assumption is that while fleets can make economic decisions based around
+where to allocate their effort in space, the total pool of effort they
+exert is fixed (we would note that the underling `marlin` model does
+have capacity for more realistic fleet models such as open-access
+dynamics). This is clearly an incorrect assumption over the long term,
+but allows us to focus on the impacts of the other drivers considered
+here. A more through consideration of the impacts of fleet model
+structure on MPA outcomes is generally needed, and is critical for any
+tactical efforts at MPA modeling.
+
+<!-- what we found was a series of neccesary but not sufficient conditions  -->
+
+<!-- make the tautology clear: illustrating a point, not proving a thing since clearly depends on if you assume displacement, and then degree of non-overlap etc.  -->
 
@@
 biomass across the range of variables explored. To illustrate, we pulled
-out two example simulations, one where MPAs provide relatively large
+out two example simulations, one where MPAs provided relatively large
 benefits across all three species, and a second where MPAs actually
@@
 results in protection for all species. In the "worst case" scenario,
-each of the species have markedly different centroids. An MPA designed
-primarily around the tuna and marlin populations then ends up displacing
-effort into the shark habitat, producing benefits for the first two
-species but driving the shark population down.
+each of the species have . An MPA designed primarily around the tuna and
+marlin populations then ends up displacing effort into the shark
+habitat, producing benefits for the first two species but driving the
+shark population down XX (Fig.\@ref(fig:fig-info)).
 
@@
 
-the point is not once answer
+```{r}
+
+critter_results <- compact_experiment_results %>%
+  filter(bau_biodiv > 0, prop_mpa > 0) %>% 
+  group_by(state_id, placement_id) %>%
+  mutate(real_sigma_centroid = sd(centroid)) %>%
+  mutate(delta_biodiv = pmin(1,biodiv - bau_biodiv),
+         loss = biodiv < bau_biodiv,
+         real_loss = biodiv < (0.9 * bau_biodiv),
+         rec_form = as.factor(rec_form)) %>% 
+  mutate(critter = fct_relevel(critter,c("Tuna", "Marlin","Shark")),
+         rec_form = fct_recode(rec_form, "Global DD, Rec. By Habitat" = "0", "Local DD, local Recruitment" = "1", "Local DD, Disperse Recruits" = "2", "Disperse Larvae, Poset-Settlement DD" = "3"))
+
+```
 
@@
 producing substantial conservation gains to substantial conservation
-losses and everything in between. We do not focus into the individual
-effects of particular variables as we found them to be highly dependent
-on the exact question asked and the set of data included, and since we
-cannot ascribe any actual likelihood to any individual state of nature
-simulated here.
-
-However, three important features do emerge. The first is that life
-history plays an important role in the potential for unintended
-conservation consequences from MPAs. All else equal, MPAs were much less
-likely to produce losses in spawning biomass for our tuna-like species
-than the marlin or shark-like species. Given that within the broad life
-history traits we do vary adult and larval movement, fishing mortality,
-as well as the timing of density dependence, we believe this result is
-mostly due to the high resilience (high steepness, low age at maturity,
-fast growth) of the tuna-like species relative to the marlin and
-shark-like species.
-
-The second is that design matters...
-
-The third is that extreme outcomes (both positive and negative),
-particularly at lower MPA sizes, are largely driven by the presence of
-small range sizes interacting with seasonal migrations. Across species,
-simulations without seasonal migrations and small ranges were much more
-likely to produce extreme results than those with seasonal migrations
-and/or larger ranges. Presence of seasonal migrations acts as a
-dilutant: As a given static MPA only provides protection, or displaces
-fishing effort, onto a particular area in only half of the year, both
-the benefits and costs of MPAs are smaller. When seasonal migrations are
-not present and the species range is small, MPAs are more capable of
-providing either large benefits or large losses.
-
-\\Big message: even constrained by the same initial depletion and core
-biology (e.g. movement rates), a wide range of outcomes are possible for
-each species for a given MPA size, dependent on the design strategy,
-habitat overlap, migrations, quality of information, etc. Across the
-board as MPA size increases the ratio of positive to negative outcomes
-increases, until all species achieve their maximum conservation
-potential in every iteration at 100% MPA. HOwever, every species
-experienced a set of simulations, roughyl xx%, that resulted in net
-losses in SSB/SSB0 relative to BAU. Losses were smaller than potential
-gaines, but reached meaningful levels in some species, and interestingly
-peaked in many cases right before max MPA (reflecting concentration of
-fishing in small amounts of habitat in habited by the species).Important
-to stress in here that there are no probabilities in this, no way of
-knowing which state is more probable than others, just given the range
-of things simulated more ways for a bigger positive effect. Also make
-clear though, biggish losses (10% or more) possible, but rare.
-
-## Case Study
-
-The general experiments conducted show two general insights
-
-1.  Even conditional on fishing exploitation and general life history
-    strategy, the effects of MPAs on a group of species with technical
-    interactions is highly variable and dependent on habitat
-    distribution, design decisions, and data quality. xx make this more
-    explicit.
-
-2.  Some degree of unintended conservation consequences occured across
-    nearly all of our simulations below MPA coverage near 100%, though
-    small losses were much more common than large. Different design
-    strategies are better suited to avoiding unintended consequences.
+losses and everything in between. Across our simulations, gains were
+much more common than losses, with losses of any magnitude occurring in
+`r scales::percent(mean(critter_results$loss))` of simulations with
+MPAs, but losses of 10% or greater of SSB in only
+`r scales::percent(mean(critter_results$real_loss))` of simulations.
+However we stress again that we can assign no probabilities to any of
+these outcomes in the real world; this simply tells us that under the
+range of conditions, selected to represent plausible states of nature,
+there are many more ways for MPAs to provide a conservation effect
+greater than or equal to 0 than there are for them to provide a net loss
+for some species.
+
+The question of interest then are which of the considered drivers appear
+to be the most influential in affecting the magnitude and direction of
+conservation effects of MPAs. Fishing mortality is the most obvious
+answer; if fishing mortality is zero in the absence of an MPA then
+clearly a no-take MPA will have no effect. As such we do not focus on
+that metric here, focusing instead on other key drivers identified by
+our simulation. Four visually prominent factors were MPA size, species,
+design strategy, and degree of habitat difference. The degree to which
+MPAs of a given size impact spawning biomass depends on a host of
+factors. The potential for conservation losses are intuitively convex in
+MPA size: effects of any kind are low when MPA size is low, and
+conservation losses must be zero when MPA size is 100% (which this graph
+doesnot quite reach), with peak losses in our simulations occuring
+between MPA sizes of XX and XX. Life history plays a clear role our
+stylized with shark and marlin populations being much more likely to
+experience net conservation losses as a function of MPA placement than
+the more resilient tunas. High degree of habitat differences can lead to
+both to the potential for large conservation gains or losses depending
+on whether a more isolated species is located primarily inside or
+outside of MPAs (Fig.\@ref(fig:fig-exp-biodiv)).
+
+```{r fig-exp-biodiv}
+
+fig_exp_biodiv <- compact_experiment_results %>%
+  filter(prop_mpa < 1) %>%
+  # filter(f_v_m > 1.5) %>%
+  group_by(state_id, placement_id) %>%
+  ggplot(aes(
+    prop_mpa,
+    pmin(1,biodiv - bau_biodiv),
+    color =  sigma_habitat,
+    group = interaction(placement_id, state_id)
+  )) +
+  geom_line(alpha = 0.5) +
+  geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
+  facet_grid(critter ~ placement_strategy) +
+  scale_color_viridis_c(name = "Habitat Difference", guide = guide_colorbar(frame.colour = "black",ticks.colour = "black", barwidth = unit ("15", "lines"))) +
+  scale_y_continuous("Change in SSB/SSB0 relative to BAU") +
+  scale_x_continuous(labels = scales::label_percent(accuracy = 1), name = "MPA Size") +
+  theme(legend.position = "top")
+
+# 
+# compact_experiment_results %>%
+#   filter(prop_mpa < 1) %>%
+#   # filter(f_v_m > 1.5) %>%
+#   group_by(state_id, placement_id) %>%
+#   ggplot(aes(
+#     prop_mpa,
+#     econ - bau_econ,
+#     color =  f_v_m,
+#     group = interaction(placement_id, state_id)
+#   )) +
+#   geom_line(alpha = 0.5) +
+#   geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
+#   facet_grid(critter ~ placement_strategy, scales = "free_y") +
+#   scale_color_viridis_c(name = "Habitat Difference", guide = guide_colorbar(frame.colour = "black",ticks.colour = "black", barwidth = unit ("15", "lines"))) +
+#   scale_y_continuous("Change in SSB/SSB0 relative to BAU") +
+#   scale_x_continuous(labels = scales::label_percent(accuracy = 1), name = "MPA Size") +
+#   theme(legend.position = "top")
+
+fig_exp_biodiv
+```
+
+We can somewhat quantitatively evaluate the relative impact of these
+drivers using a regression, given that by construction variation in each
+of evaluated drivers are independent in our simulations. We fit a
+binomial regression to the presence or absence of of a net conservation
+loss per species as a function of the simulated experimental variables.
+The factors that according to this model most increase the probability
+of a net conservation loss occurring include habitat differences, a
+placement strategy of avoiding fishing areas, and the life history of
+the species (shark or marlin relative to tuna). The factors that most
+decrease the probability of a conservation loss from an MPA are MPA
+size, placement strategy based on local depletion (xx might remove this
+strategy as unrealistic xx), and somewhat interestingly fishing
+mortality rate (Fig.\@ref(fig:fig-loss-drivers)-A).
+
+However, this result is somewhat misleading (xx so should we just
+remove?xx). Fig.\@ref(fig:fig-exp-biodiv) shows the clearly bimodal
+nature of for example habitat differences. This analysis helps inform us
+that habitat differences are a necessary but not sufficient condition
+for MPA leakage induced conservation losses. The regression based
+results require assuming that each of the effects are independent and
+linear in nature. In fact each of these factors can interact in complex
+and non-linear ways that may not be apparent visually or through this
+type of regression analysis. To address this, we ran a random forest
+using the `ranger` package in R (citaitons) and extracted the impurity
+corrected importance scores for each of the variables. While this does
+not tell us the expected marginal effect of individual drivers, it does
+tell us which variables are critical if our goal is to accurately
+predict the potential for unintended consequences or magnitude of
+conservation gains in MPAs. Note here that variables that intuitively
+should have an impact but have no visually or statistically (in a linear
+model) clear impact on the probability of experiencing a conservation
+loss such as the dispersal distances of adults or larvae have importance
+scores comparable in magnitude to habitat differences when considered in
+the context of nonlinear effects and interactions with other variables
+(Fig.\@ref(fig:fig-loss-drivers)-B).
+
+```{r fig-loss-drivers}
+
+# loss_reg <-   rstanarm::stan_glm(
+#   loss ~   f_v_m  + critter + seasonal_movement + adult_movement_sigma + rec_form + recruit_movement_sigma + prop_mpa + sigma_habitat + spatial_allocation + placement_strategy + hyper,
+#   family = binomial(),
+#   data = critter_results %>% filter(between(prop_mpa,.2,.4)),
+#   cores = 4,
+#   chains = 4
+# )
+
+loss_reg <-   glm(
+  loss ~   f_v_m  + critter + seasonal_movement + adult_movement_sigma + rec_form + recruit_movement_sigma + prop_mpa + sigma_habitat + spatial_allocation + placement_strategy + hyper,
+  family = binomial(),
+  data = critter_results %>% filter(between(prop_mpa,.2,.4))
+)
+
+loss_reg_coefs <- broom::tidy(loss_reg) %>%
+  mutate(lower = estimate - 1.96 * std.error,
+         upper = estimate + 1.96 * std.error) %>%
+  mutate(
+    term = str_replace(term, "placement_strategy", "Placement Strategy: "),
+    term = str_replace(term, "rec_form", "DD Timing: "),
+    term = str_replace(term, "spatial_allocation", "Fleet Allocation: "),
+    term = str_replace(term, "hyper", "Hyperallometry present"),
+    term = str_replace(term, "prop_mpa", "MPA size"),
+    term = str_replace(term, "sigma_habitat", "Habitat differences"),
+     term = str_remove_all(term, "critter")
+  )
+
+loss_coef_plot <- loss_reg_coefs %>%
+  ggplot() +
+  geom_hline(aes(yintercept = 0)) +
+  geom_pointrange(aes(
+    forcats::fct_reorder(term, estimate),
+    estimate,
+    ymin = lower,
+    ymax = upper
+  )) +
+  scale_y_continuous(name = "Log-odds Ratio Effect") +
+  scale_x_discrete(name = '') +
+  coord_flip()
+
+tmp_critter <-
+  critter_results %>% filter(between(prop_mpa, .2, .4)) %>%
+  select(-habitat) %>% 
+  mutate(fctr_loss = factor(loss))
+# 
+# tmp_rcp <- recipes::recipe(loss ~ ., data = tmp_critter) %>% 
+#   themis::step_upsample(fctr_loss, over_ratio = 0.75) %>% 
+#   recipes::prep(training = tmp_critter, retain = TRUE) %>% 
+#   recipes::juice()
+
+loss_rf <-   ranger::ranger(
+  loss ~   f_v_m  + critter + seasonal_movement + adult_movement_sigma + rec_form + recruit_movement_sigma + prop_mpa + sigma_habitat + spatial_allocation + placement_strategy + hyper,
+  data = tmp_critter,
+  importance = "impurity"
+)
+
+loss_importance <- data.frame(importance = loss_rf$variable.importance %>% sort()) %>% 
+  rownames_to_column(var = "term")
+
+loss_importance_plot <- loss_importance %>% 
+  ggplot(aes(forcats::fct_reorder(term, importance), importance)) + 
+  geom_col() +
+  scale_y_continuous(name = "Variable Importance") +
+  scale_x_discrete(name = '') +
+  coord_flip()
+
+fig_loss_drivers <- loss_coef_plot + loss_importance_plot
+
+fig_loss_drivers
+```
+
+```{r fig-delta-drivers, eval = FALSE}
+
+delta_reg <-   rstanarm::stan_glm(
+  delta_biodiv ~   f_v_m  + critter + seasonal_movement + adult_movement_sigma + rec_form + recruit_movement_sigma + prop_mpa + sigma_habitat + spatial_allocation + hyper,
+  data = critter_results %>% filter(between(prop_mpa,.2,.4)),
+  cores = 4,
+  chains = 4
+)
+
+tmp_critter <-
+  critter_results %>% filter(between(prop_mpa, 0.2, 0.4)) %>%
+  select(-habitat) %>% 
+  mutate(fctr_loss = factor(loss))
+# 
+# tmp_rcp <- recipes::recipe(loss ~ ., data = tmp_critter) %>% 
+#   themis::step_upsample(fctr_loss, over_ratio = 0.75) %>% 
+#   recipes::prep(training = tmp_critter, retain = TRUE) %>% 
+#   recipes::juice()
+
+delta_rf <-   ranger::ranger(
+  delta_biodiv ~   f_v_m  + critter + seasonal_movement + adult_movement_sigma + rec_form + recruit_movement_sigma + prop_mpa + sigma_habitat + spatial_allocation + placement_strategy + hyper,
+  data = tmp_critter,
+  importance = "impurity"
+)
+
+delta_rf$variable.importance %>% sort()
+
+plot(delta_reg)
+
+```
+
+## WCPO Case Study
+
+The generalized experiments help us understand the potential impacts of
+the evaluated drivers, but also highlight the challenge in making
+definitive statements about the effect of a given driver on the outcomes
+of pelagic(?) MPAs, given the complex ways in which all of these factors
+interact, even within a relatively simple model such as this one. In
+order to help illustrate how addressing the kinds of spatial dynamics
+and technical interactions included in our model can affect and benefit
+marine spatial planning, we then ran a case study intended to loosely
+reflect the dynamics of WCPO tuna, billfish, and shark fisheries and
+populations.
+
+This case study allows us to focus in on important questions around the
+relative performance of "rule of thumb" MPA design strategies relative
+to optimized networks. We ran a spatial optimization routine (see
+section XX) across a range of objective function weightings, in the
+manner of @rassweiler2014. This allowed us to assess 1) the potential
+risk of unintended conservation consequences from MPAs in this WCPO
+bycatch study and the 2) assess value of optimization in this WCPO case
+study in achieving economic and conservation objectives.
+
+Given the set of parameters simulated here, we find little intrinsic
+risk for large-scale unintended losses in total spawning biomass caused
+by MPA effort displacement in the WCPO case study. Each of the design
+scenarios produced at least some small loss in biomass for one or more
+species at some MPA size. However, these losses were generally small and
+isolated to a subset of species within a small MPA size range, with the
+exception of the "Avoid_fishing" MPA placement strategy that resulted in
+net losses for most species at large MPA sizes (though simply minimal
+impact at smaller sizes). All design strategies except "Avoid_fishing"
+and "Rate" resulted in relatively rapid declines in profits for the
+three main tuna fisheries (skipjack, yellowfin, and bigeye). Both
+"Avoid_fishing" and "Rate" mainlined profits close to business as usual
+up to a large MPA size, but did so at the expense of conservation
+benefits (Fig.\@ref(fig:fig-cs-critters)).
+
+```{r fig-cs-critters, fig.cap = "Simulated trajectories of changes in profits and spawning stock biomass (SSB) for selected species caught in the WCPO.
+Each column represents a different MPA design strategy, see Table.XX for definitions.
+Optimized refers to an MPA network optimized to maximize total SSB across all species."}
+
+opt_comparison <- opt_obj %>%
+  filter(alpha == max(alpha)) %>%
+  mutate(placement_error = 0,
+         prop_mpa = p_protected / 100,
+         placement_strategy = "Optimized") %>%
+  select(critter, placement_error, delta_biodiv, delta_econ, prop_mpa,placement_strategy)
+
+capitalize <- function(string) {
+  substr(string, 1, 1) <- toupper(substr(string, 1, 1))
+  string
+}
+
+ experiment_obj %>% 
+   select(colnames(opt_comparison)) %>% 
+  bind_rows(opt_comparison) %>% 
+   mutate(placement_strategy = fct_relevel(placement_strategy,"Optimized")) %>% 
+   left_join(sci_to_com, by = "critter") %>% 
+  filter(placement_error == min(placement_error)) %>% 
+   rename("Change in SSB" = delta_biodiv, "Change in Profits" = delta_econ) %>% 
+  pivot_longer(c("Change in SSB", "Change in Profits"), names_to = "metric", values_to = "value") %>% 
+  ggplot(aes(prop_mpa, value, color = common_name)) + 
+   scale_color_discrete(name = '') +
+   geom_hline(aes(yintercept = 0), linetype = 2)+
+  geom_line(show.legend = TRUE) + 
+   scale_y_continuous(name = '') +
+   scale_x_continuous(name = "MPA Size",labels = scales::percent, guide = guide_axis(n.dodge = 2)) +
+  facet_grid(metric~placement_strategy, scales = "free_y", switch = "y", labeller = labeller(placement_strategy = capitalize)) + 
+   theme(strip.text.x = element_text(size = 8), 
+         legend.position = "top", 
+         legend.text = element_text(size = 8))
+
+```
+
+Viewed as totals, we found a small parameter space under the "Area" and
+"Avoid_fishing" design strategies where MPAs were capable of causing a
+net decrease in total spawning stock biomass across all evaluated
+species. However, the efficiency frontier indicates that these points
+are not Pareto efficient, and MPA design configurations exist that
+achieve the same levels of economic performance while producing a net
+gain in spawning biomass. Optimal MPA design was able to slow the
+tradeoff between total profits and bycatch spawning biomass, but for any
+conservation weighting any increases in total biomass of bycatch species
+came at a decrease in total profits. However we also found limited need
+for optimized networks, with the "Depletion" design scenario essentially
+following the Pareto frontier (Fig.\@ref(fig:fig-cs-totals)).
+
+```{r fig-cs-totals, fig.cap = "Efficiency frontiers for changes in total profits and total bycatch spawning stock biomass (SSB).
+Color gradient line shows optimized pareto frontier, with color showing weighting of conservation (total spawning stock biomass) in the objective function.
+Red points show outcomes of 'rule of thumb' MPA design strategies.
+"}
+ total_opt_obj %>%
+  group_by(alpha) %>%
+  filter(obj_fun == max(obj_fun)) %>%
+  ungroup() %>%
+  ggplot() +
+  geom_line(aes(total_bycatch_ssb_change, delta_econ, color = alpha), size = 1.5) +
+  geom_point(
+    data = total_experiment_obj,
+    aes(total_bycatch_ssb_change, delta_econ),
+    size = 2,
+    alpha = 0.1,
+    fill = "red",
+    shape = 21
+  ) +
+  geom_vline(aes(xintercept = 0), linetype = 2) +
+  geom_hline(aes(yintercept = 0), linetype = 2) +
+  geom_abline(intercept = 0, slope = -1) +
+  facet_wrap(~placement_strategy, scales = "free_y", labeller = labeller(placement_strategy = capitalize)) + 
+  scale_color_viridis_c(name = "Conservation Weighting in Optimization",
+                        guide = guide_colorbar(frame.colour = "black", ticks.colour = "black", barwidth = unit("15", "lines"))) +
+  scale_x_continuous(name = "Change in Total Bycatch SSB") +
+  scale_y_continuous(name = "Change in Total Profits ") +
+  theme(axis.text.x = element_text(size = 8),
+        legend.position = "top")
+
+```
+
+```{r, eval = FALSE}
+
+ total_opt_obj %>%
+  group_by(alpha) %>%
+  filter(obj_fun == max(obj_fun)) %>%
+  ungroup() %>%
+  ggplot() +
+  geom_line(aes(delta_biodiv, delta_econ, color = alpha), size = 2) +
+  geom_point(
+    data = total_experiment_obj,
+    aes(delta_biodiv, delta_econ),
+    size = 2,
+    alpha = 0.1,
+    fill = "red",
+    shape = 21
+  ) +
+  geom_vline(aes(xintercept = 0), linetype = 2) +
+  geom_hline(aes(yintercept = 0), linetype = 2) +
+  geom_abline(intercept = 0, slope = -1) +
+  facet_wrap( ~ placement_strategy) +
+  scale_color_viridis_c(name = "Conservation Weighting in Optimization") +
+  scale_x_continuous(name = "Change in Total Biodiversity Relative to BAU ") +
+  scale_y_continuous(name = "Change in Total Profits Relative to BAU ") +
+  theme(axis.text.x = element_text(size = 8),
+        legend.position = "top")
+
+```
+
+# Discussion
+
+MPAs can be a powerful tool for providing conservation benefits across
+large swaths of species and space. Under the appropriate circumstances
+they can also provide benefits to food security and economic outcomes.
+However, while extremely simple in concept, predicting the likely
+effects of MPAs across groups of species and fishing fleets becomes
+increasingly complicated as the systems being studied deviate from
+assumptions of homogenously distributed species with perfectly selective
+fleets. In this paper we used a new bio-economic model to demonstrate
+how a range of variables often omitted in MPA modeling can act to
+influence the outcomes of MPAs on conservation and economic objectives.
+We demonstrate that unintended conservation losses are possible, but
+under a range of specific conditions, and under most simulated cases did
+not occur. MPAs were capable of causing small losses in biomass for the
+WCPO case study, but only for a subset of design strategies and species,
+most of which could be overcome by optimized design.
+
+Our results indicate that as suggested by theory (citation) and
+empirical evidence (citation), MPAs can indeed have unintended
+consequences for conservation if they result in effort displacement
+rather than attrition. However, our model found these to be a minority
+of the simulations tested. This does in no way suggests that leakage may
+not a problem for specific MPA. But, it also indicates that across the
+range of plausible parameter combinations that we explored it was much
+more common for MPAs to provide net conservation benefits even if effort
+is displaced. This suggests then that we should of course not dismiss
+the conservation benefits of MPAs in the presence of effort
+displacement, but also that users must be aware of the potential for
+unintended consequences that can indeed occur. While many different
+factors interact in complex ways to affect the potential for unintended
+conservation consequences, the most obvious is the degree of difference
+in the habitats utilized by the species within an ecosystem in which an
+MPA is to be placed. As a starting point, it is critical that
+communities engaged in MSP processes create a localized assessment of
+the degree of habitat overlap across potentially affected species to
+assess the potential risk of unintended consequences.
+
+The MPA literature is highly dominated by modeling studies (citation).
+What then is the added value of the model and results presented here?
+Our model contains many features missing from many general MPA models,
+including size structure, seasonal movement, larval and adult habitats
+and dispersal, and critically technical interactions xx clearer term for
+this xx across species. A reader might reasonably ask about a range of
+other factors not included, such as time-varying parameters and species
+interactions. One could of course continue to add features to the model
+until it reaches the complexity of an end-to-end model such as Atlantis
+(citation). While such models have tremendous value, our goal here is to
+present a model that can help provide closer to tactical advice for
+users without the data or capacity for such high resolution models. One
+justification for simpler MPA models is often computational
+tractability: tracking the full age structure of multiple species and
+the economic outcomes of multiple fleets dynamically across a high
+resolution spatial grid can quickly become computationally expensive.
+Faster implementations completely written in languages such as C++
+however become inaccessible to many users. Our model seeks to fill a
+middle ground here, with the core operations run in C++ while providing
+a relatively user-friendly front end in the open-source and commonly
+used R programming language. This allows users to ask the sorts of
+questions addressed in this paper that require a more realistic model,
+without requiring inordinate amounts of data, programming expertise, or
+computational time.
+
+Caveats
+
+Note that lack of importance of density dependence timing is conditional
+on age structure, and matters more if you're doing surplus production
+modeling (cite letter)
+
+Mention hyperallometry and distinction to what that study did
+
+Perfect information
+
+Fleet model
+
+**what is the story**
+
+-   Unintended conservation consequences of MPAs resulting from
+    technical interactions have been discussed in the literature, but
+    almost exclusively in the context of highly tactical modeling or
+    localed empirics. We conduct one of the first large-scale
+    assessments of the potential of this problem across a range of
+    scenarios
+
+-   We find that these unintended consequences can be common outside of
+    "canonical" MPA models *a la* hastings etc.
+
+-   But, good news is magnitude of unintended consequences were usually
+    small, and total benefits almost always outweigh individual species
+    losses (again no probability of states though)
+
+-   Some design strategies are better at avoiding unintended
+    consequences than others, but sensitive to data quality.
+    Optimization can help, but even with optimization unintended
+    consequences can happen.
+
+-   Under the levels of fishing mortality in the WCPO, optimization and
+    design can help reduce tradeoffs between fisheries revenue and
+    unintended consequences than others, but macro-level tradeoffs
+    between biodiversity and fishery revenue are very insensitive to
+    design choices, with a pretty clear "increasing biodiversity through
+    MPAs reduces fishery revenues".
+
+    -   xx This is clearly hyper-sensitive to the one thing we don't
+        iterate over much here: how overfished are the target species.
+        At the moment "they're not". But, this runs fast enough, is it
+        worth running a hypothetical case study where they are just to
+        illustrate how things change?
+
+# Conclusions
+
+-   Unintended conservation outcomes of MPAs totally possible outside of
+    "consider a homogenous fish blob" models
+
+-   But, usually small, and net conservation benefits outweigh
+    conservation costs
+
+-   Should pay particular attention to potential though when species
+    near tipping points are of concern: hard to imagine those not being
+    part of the prioritization but something to consider.
+
+-   Other interesting things.
+
+# References
+
+::: {#refs}
+:::
+
+# Scraps
 
@@
 
-For this subset of simulations then, along with the various "rule of
-thumb" MPA design strategies we ran an optimization routine (see section
-XX) across a range of objective function weightings, in the manner of
-@rassweiler2014. This allowed us to assess 1) the degree to which
-unintended conservation consequences of MPAs can be reduced by optimized
-design and 2) the magnitude of tradeoffs between conservation outcomes
-and fishery revenues under both rule-of-thumb and optimized MPA
-networks.
-
-key results: optimization helps (Nobel prize please). Can reduce percent
-of species with conservation loss down to 25% pretty quickly if you
-prioritize conservation. Rule of thumb designs like targeting fishing
-can get there with perfect information and when considering all animals,
-but can be pretty far from optimal with imperfect information or when
-only considering a subset of animals in the planning process (that's
-what all the different points represent). Some rules of thumb are more
-sensitive to these decision than others. E.g. rate can get you optimal,
-but goes off the rails pretty quickly with bad information.
-
-There is also a pretty clear breakpoint in results between basically
-trying for conservation or trying for economics, and not a lot of wiggle
-room in between (the two general bands of lines). xx check to what
-extent this depends on the way you scale the two objective function
-components xx.
-
-```{r, fig.cap="Percent of species with some conservation loss as a function of MPA size.
-Points are draws from rule-of-thumb scenarios, lines are results of optimization routine across a gradient of conservation weighting."}
-
-```
-
-Key results: Even with optimization, you can basically either be in a
-world with 50% or more of species having some level of unintended
-conservation consequence, but no economic loss / some economic gain, or
-you can be in a world with lower unintended consequences, but economic
-losses. But, even rule of thumb approaches can get you to some places
-with relatively little unintended consequences at relatively minimal
-economic loss.
-
-```{r, fig.cap="Tradeoff between unitendedn consequences and fisheries revenue.
-Filled polygons in background represent rule-of-thumb results, with redder colors indicating higher density of points.
-Points are results of optimization, with color indicating conservation weighting in optimization.
-Black diagonal line shows 1:1 relationship"}
-
-```
-
-Key results: Presence of and magnitude of unintended consequences pretty
-sensitive to design descisions. On a macro scale though far less so. In
-terms of change in total SSB/SSB0 (metric of biodiversity, i.e. maxed
-out when all species are unfished) and change in revenue, there is
-basically a pretty simple linear tradeoff between biodiversity and
-revenue, with a slope less than negative 1, meaning that 1% increase in
-biodiversity comes at a greater than 1% loss in fishery revenue. This
-stacks up pretty nicely with @rassweiler2014., where value of
-optimization was low when overfishing of valuable species isn't
-prevalent, as is the case here. So, design and optimization can help a
-lot with the more "micro" level outcomes (e.g. making sure that
-biodiversity gaines aren't coming from massive gains in one species to
-detriment of other), but on a macro level pretty insensitive in this
-case study.
-
-```{r, fig.cap="Tradeoff frontier between revenue and biodiversity xx across different rule odf thumb design strategies.
-Points indicate role of thumb results, lines are results of optimized networks across a gradient of conservation weighting."}
-
-```
-
-# Discussion
-
-motivated by but not limited to pelagic ecosystems
-
-**what is the story**
-
--   Unintended conservation consequences of MPAs resulting from
-    technical interactions have been discussed in the literature, but
-    almost exclusively in the context of highly tactical modeling or
-    localed empirics. We conduct one of the first large-scale
-    assessments of the potential of this problem across a range of
-    scenarios
-
--   We find that these unintended consequences can be common outside of
-    "canonical" MPA models *a la* hastings etc.
-
--   But, good news is magnitude of unintended consequences were usually
-    small, and total benefits almost always outweigh individual species
-    losses (again no probability of states though)
-
--   Some design strategies are better at avoiding unintended
-    consequences than others, but sensitive to data quality.
-    Optimization can help, but even with optimization unintended
-    consequences can happen.
-
--   Under the levels of fishing mortality in the WCPO, optimization and
-    design can help reduce tradeoffs between fisheries revenue and
-    unintended consequences than others, but macro-level tradeoffs
-    between biodiversity and fishery revenue are very insensitive to
-    design choices, with a pretty clear "increasing biodiversity through
-    MPAs reduces fishery revenues".
-
-    -   xx This is clearly hyper-sensitive to the one thing we don't
-        iterate over much here: how overfished are the target species.
-        At the moment "they're not". But, this runs fast enough, is it
-        worth running a hypothetical case study where they are just to
-        illustrate how things change?
-
-# Conclusions
-
--   Unintended conservation outcomes of MPAs totally possible outside of
-    "consider a homogenous fish blob" models
-
--   But, usually small, and net conservation benefits outweigh
-    conservation costs
-
--   Should pay particular attention to potential though when species
-    near tipping points are of concern: hard to imagine those not being
-    part of the prioritization but something to consider.
-
--   Other interesting things.
-
-# References
-
-::: {#refs}
-:::
-
-# Scraps
+The general experiments conducted show two general insights
+
+1.  Even conditional on fishing exploitation and general life history
+    strategy, the effects of MPAs on a group of species with technical
+    interactions is highly variable and dependent on habitat
+    distribution, design decisions, and data quality. xx make this more
+    explicit.
+
+2.  Some degree of unintended conservation consequences occured across
+    nearly all of our simulations below MPA coverage near 100%, though
+    small losses were much more common than large. Different design
+    strategies are better suited to avoiding unintended consequences.
 
@@
 management MPAs as risk reduction strategy Etc.
+
+We do not focus into the individual effects of particular variables as
+we found them to be highly dependent on the exact question asked and the
+set of data included, and since we cannot ascribe any actual likelihood
+to any individual state of nature simulated here.
+
+However, three important features do emerge. The first is that life
+history plays an important role in the potential for unintended
+conservation consequences from MPAs. All else equal, MPAs were much less
+likely to produce losses in spawning biomass for our tuna-like species
+than the marlin or shark-like species. Given that within the broad life
+history traits we do vary adult and larval movement, fishing mortality,
+as well as the timing of density dependence, we believe this result is
+mostly due to the high resilience (high steepness, low age at maturity,
+fast growth) of the tuna-like species relative to the marlin and
+shark-like species.
+
+The second is that design matters...
+
+```{r}
+
+```
+
+The third is that extreme outcomes (both positive and negative),
+particularly at lower MPA sizes, are largely driven by the presence of
+small range sizes interacting with seasonal migrations. Across species,
+simulations without seasonal migrations and small ranges were much more
+likely to produce extreme results than those with seasonal migrations
+and/or larger ranges. Presence of seasonal migrations acts as a
+dilutant: As a given static MPA only provides protection, or displaces
+fishing effort, onto a particular area in only half of the year, both
+the benefits and costs of MPAs are smaller. When seasonal migrations are
+not present and the species range is small, MPAs are more capable of
+providing either large benefits or large losses.
+
+\\Big message: even constrained by the same initial depletion and core
+biology (e.g. movement rates), a wide range of outcomes are possible for
+each species for a given MPA size, dependent on the design strategy,
+habitat overlap, migrations, quality of information, etc. Across the
+board as MPA size increases the ratio of positive to negative outcomes
+increases, until all species achieve their maximum conservation
+potential in every iteration at 100% MPA. HOwever, every species
+experienced a set of simulations, roughyl xx%, that resulted in net
+losses in SSB/SSB0 relative to BAU. Losses were smaller than potential
+gaines, but reached meaningful levels in some species, and interestingly
+peaked in many cases right before max MPA (reflecting concentration of
+fishing in small amounts of habitat in habited by the species).Important
+to stress in here that there are no probabilities in this, no way of
+knowing which state is more probable than others, just given the range
+of things simulated more ways for a bigger positive effect. Also make
+clear though, biggish losses (10% or more) possible, but rare.

### 2021-10-09 R/optimize_mpa.R

@@ -202,10 +202,14 @@ optimize_mpa <-
 
         econ <- profit_mpa
 
+        yield_mpa <-  (map_dbl(res, ~ sum(.x$c_p_fl, na.rm = TRUE)))
+
+
       out <-
         list(
           biodiv = tibble(critter = names(biodiv), ssb_v_ssb0 = biodiv),
-          econ = tibble(critter = names(econ), econ = econ)
+          econ = tibble(critter = names(econ), econ = econ),
+          yield = tibble(critter = names(yield_mpa), yield = yield_mpa)
         )
       # keep just biomass and catch for now: you can always use the MPA layer at that step to recreate the whole sim if you need
 
@@ -249,17 +253,18 @@ optimize_mpa <-
       econ_sq <-
         (map_dbl(starting_conditions, ~ sum(.x$r_p_a_fl, na.rm = TRUE)))
 
-      profit_mpa <-  (map_dbl(res, ~ sum(.x$prof_p_fl, na.rm = TRUE)))
+      profit_sq <-  (map_dbl(res, ~ sum(.x$prof_p_fl, na.rm = TRUE)))
 
-      econ <- profit_mpa
+      econ <- profit_sq
 
-    # econ <-
-    #   (map_dbl(res, ~ sum(.x$r_p_a_fl, na.rm = TRUE))) #  calculate econ component of objective function, currently revenues across all fleets and species
+      yield <-  (map_dbl(res, ~ sum(.x$c_p_fl, na.rm = TRUE)))
 
     out <-
       list(
         biodiv = tibble(critter = names(biodiv), ssb_v_ssb0 = biodiv),
-        econ = tibble(critter = names(econ), econ = econ)
+        econ = tibble(critter = names(econ), econ = econ),
+        yield = tibble(critter = names(yield), yield = yield)
+
       )
 
     baseline_outcome <- tibble(p_protected = 0, out = list(out))
@@ -272,63 +277,23 @@ optimize_mpa <-
         out = results
       )) %>%
       mutate(bio = map(out, "biodiv"),
-             econ = map(out, "econ")) %>%
-      unnest(cols = c(bio, econ), names_repair = "universal") %>%
-      select(-`critter...5`) %>%
+             econ = map(out, "econ"),
+             yield = map(out, "yield")) %>%
+      unnest(cols = c(bio, econ, yield), names_repair = "universal") %>%
[…]

### 2021-10-09 scripts/00_setup.R

@@ -46,11 +46,11 @@ if (!dir.exists(results_path)){
 
 run_experiments <- FALSE
 
-experiment_workers <- 10
+experiment_workers <- 1
 
-run_casestudy <- FALSE
+run_casestudy <- TRUE
 
-optimize_casestudy <- FALSE
+optimize_casestudy <- TRUE
 
 safety_stop <- FALSE
 

### 2021-10-09 scripts/01_run_experiments.R

@@ -3,7 +3,7 @@ source(file.path("scripts", "00_setup.R"))
 
 # generate state experiments. This is a somewhat tricky process where the actual generated values are created for many variables in create_experiment_critters
 
-n_states <- 500
+n_states <- 600
 
 if (run_experiments) {
   state_experiments <-
@@ -124,6 +124,7 @@ if (run_experiments) {
         min = 0.01 * resolution,
         max = .5 * resolution
       ),
+      hyper = sample(c(1,1.5),nrow(.), replace = TRUE),
       rec_form = sample(c(0, 1, 2, 3), nrow(.), replace = TRUE)
     )
 
@@ -143,7 +144,8 @@ if (run_experiments) {
           f_v_m = f_v_m,
           adult_movement_sigma = adult_movement_sigma,
           recruit_movement_sigma = recruit_movement_sigma,
-          rec_form = rec_form
+          rec_form = rec_form,
+          hyper = hyper
         ),
         create_experiment_critters,
         marlin_inputs = marlin_inputs,
@@ -939,48 +941,19 @@ fig_9
 
 critter_results <- results %>%
   group_by(state_id, placement_id) %>%
-  mutate(delta_biodiv = pmin(2,biodiv / bau_biodiv - 1),
+  mutate(delta_biodiv = pmin(2,biodiv / bau_biodiv),
          loss = biodiv < bau_biodiv,
          real_loss = biodiv <= (0.9 * bau_biodiv))
+
+critter_results %>%
+  ggplot(aes(delta_biodiv)) +
+  geom_histogram() +
+  scale_x_continuous(labels = scales::percent) +
+  geom_vline(xintercept = 0.9)
 
 
 mean(critter_results$loss, na.rm = TRUE)
 
-tree_1 <-   rpart::rpart(
-  real_loss ~ f_v_m  + critter + seasonal_movement + adult_movement_sigma + rec_form + recruit_movement_sigma + placement_strategy + prop_mpa + sigma_habitat + spatial_allocation,
-  data = critter_results %>% filter(prop_mpa > .2, prop_mpa < 0.4),
-  method = "class"
-)
-#
-rpart.plot(tree_1, type = 5)
-
-
-reg_1 <-   rstanarm::stan_glm(
[…]

### 2021-10-09 scripts/02_run_case_study.R

@@ -276,7 +276,8 @@ if (run_casestudy == TRUE){
       placement_strategy = c("depletion", "rate", "avoid_fishing", "target_fishing", "area"),
       prop_mpa = seq(0,1, by = 0.01),
       prop_critters_considered = 1,
-      placement_error = c(0,.2)
+      placement_error = c(0),
+      mpa_response = c("stay")
     )
 
   a <- Sys.time()
@@ -285,7 +286,8 @@ if (run_casestudy == TRUE){
     mutate(results = future_pmap(list(placement_strategy = placement_strategy,
                                prop_mpa = prop_mpa,
                                prop_critters_considered = prop_critters_considered,
-                               placement_error = placement_error),
+                               placement_error = placement_error,
+                               mpa_response = mpa_response),
                           run_mpa_experiment,
                           starting_conditions = starting_conditions,
                           proc_starting_conditions = proc_starting_conditions,
@@ -567,7 +569,15 @@ experiment_obj <- case_study_experiments %>%
   select(-results) %>%
   unnest(cols = obj) %>%
   mutate(econ = econ / 1e6) %>%
-  filter(prop_mpa < 1)
+  filter(prop_mpa < 1) %>%
+  mutate(bycatch = !(str_detect(critter, "thunnus") | str_detect(critter,"katsuwonus")))
+
+
+experiment_obj <- experiment_obj %>%
+  left_join(bau, by = "critter") %>%
+  mutate(delta_biodiv = biodiv - bau_biodiv,
+         delta_econ = econ - bau_econ)
+
 
 experiment_obj %>%
   ggplot(aes(biodiv, econ, color = placement_strategy)) +
@@ -599,7 +609,9 @@ opt_obj <- optimized_networks %>%
   mutate(obj = map(network,"outcomes")) %>%
   select(-network) %>%
   unnest(cols = obj) %>%
-  mutate(econ = econ / 1e6)
+  mutate(econ = econ / 1e6) %>%
+  mutate(bycatch = !(str_detect(critter, "thunnus") | str_detect(critter,"katsuwonus")))
+
 
 obj_objective_value <- optimized_networks %>%
[…]

### 2021-10-12 documents/03_marlin_paper.Rmd
finally back to main

@@
 ---
-title: "Solutions to the potential for unintended conservation consequences of MPAs"
-subtitle: Conditions for success and unintended consequences in pelagic MPA design
+title: "Opportunities and precautions for the use of MPAs in bycatch reduction"
 author:
@@
 output: 
+  bookdown::word_document2:
+    reference_docx: template.docx
+  bookdown::html_document2: default
   bookdown::pdf_document2: default
-  bookdown::html_document2: default
+header-includes:
+  - \usepackage{setspace}\doublespacing
+  - \usepackage{lineno}\linenumbers
 bibliography: references.bib
@@
 
+capitalize <- function(string) {
+  substr(string, 1, 1) <- toupper(substr(string, 1, 1))
+  string
+}
+
 ```
@@
 
-c
-
-\*\* something to point out: increased catches are a paradox, either
-come from rebuilding stock or concentrated fishing, but trajectory can
-tell you something. not possible for MPAs to increase catches without
-decreasing catches first if they are rebuilding? \*\*
-
-Seasonal movemetn is kind of interesting, basically just dilutes the
-extremes
-
-less resilient species are both more in need of protection and easier to
-accidentally fish harder
+<!-- \*\* something to point out: increased catches are a paradox, either -->
+<!-- come from rebuilding stock or concentrated fishing, but trajectory can -->
+<!-- tell you something. not possible for MPAs to increase catches without -->
+<!-- decreasing catches first if they are rebuilding? \*\* -->
+
+<!-- Seasonal movement is kind of interesting, basically just dilutes the -->
+<!-- extremes -->
+
+<!-- less resilient species are both more in need of protection and easier to -->
+<!-- accidentally fish harder -->
 
@@
 
+<!-- MAKE IT MORE ABOUT BYCATCH-->
+
 <!-- Empirically, how do you differentiate increase in catch from fishing harder vs increases in catch from rebuilding: rebuilding has to have a decrease at first -->
 
-in the Marine protected areas (MPAs) are a growing part of the marine
+Marine protected areas (MPAs) are a growing part of the marine
 conservation toolbox, with 5.9% of the world's oceans currently in
 implemented MPAs [@grorud-colvert2021], and bodies such as the IUCN
-calling for 30% protection by 2030. The modern science of MPAs was
+calling for 30% protection by 2030. One of the potential benefits of
+MPAs is their ability to provide protection to broad segments of an
+ecosystem at once, rather than requiring targeted management based on
+individual species that may be unfeasible in many parts of the world
+lacking in data or fisheries management capacity. Their ability to
+provide protection across a range of species with one policy has led them to be suggested as a tool to reduce problems around weak stock or "bycatch"
+species [@hastings2017]. We use the term "bycatch" here to generally refer to the
+capture of species within a fishery primarily targeting a different set
+of species.
+
+While the effect of MPAs on fisheries yields is a continuously debated
+topic, the conservation impacts of MPAs are seemingly simpler, albeit
+challenging to detect in many cases [@ovando2021a], due in part perhaps
+to the perceived obvious nature of their benefits: if you prevent
+fishing of a species within an area, there will be more of that species
+in that area, and so the conservation benefits of MPAs will increase
+with MPA size. @hastings2017 clearly illustrates this idea. Using a two-species two-patch analytical model, they derive conditions
+under which MPAs can solve the "weak-stock" problem in which a less
+productive species is caught as bycatch by a fleet targeting a more
+valuable and productive stock.
+
+However, as @hastings2017 point out, their results depend on a very
+specific set of assumptions, namely that 1) adults are perfectly
+sedentary 2) the target and bycatch species overlap perfectly in terms
+of habitat and 3) recruits are perfectly mixed throughout the system. In
+this paper We build on @hastings2017 to consider the role of MPAs in
+reducing bycatch while managing tradeoffs with fisheries under a more
+resolved model.
+
+Specifically, the potential benefits of MPAs as a bycatch management are
+complicated by two key issues: effort displacement and technical
+interactions. One of the primary questions around MPAs is to what extent
+do MPAs remove versus displace fishing effort. If MPAs remove fishing
+effort, then their direct effects on target and bycatch species can only
+be positive. If however MPAs displace fishing effort evaluating their
+potential impacts becomes more complicated. The net conservation impact
+of MPAs becomes a balancing act between the effects of reduced fishing
+inside protected borders and increased fishing intensity outside MPAs.
+The outcomes of this displacement can be further complicated by
+technical interactions, i.e. the degree to which fishing effort for one
+species affects another, particularly when species have markedly
+different habitat distributions. For example, @abbott2012 demonstrated
+that an MPA designed to protect red king crab (*Paralithodes
+camtschaticus*) caused an unintended increase in bycatch of Pacific
+halibut (*Hippoglossus stenolepis*) by the displaced bottom-trawl
+fishery. Policy leakage is not a unique feature to MPAs, e.g. XX, but
+the potential impacts of MPAs on target and bycatch species taking into
+account the potential for these leakage effects has received relativelly
+little generalized attention in the literature [see @abbott2012;
+@savina2008; @fulton2011a; @baum2003 for notable exceptions].
+
+This study addresses this gap using a novel bio-economic simulation
+model that allows for efficient evaluation of the effects of MPAs on
+target and bycatch species across a range of ecological and economic
+dynamics, critically accounting for the potential impacts of effort
+displacement and technical interactions. We use this model first to
+explore general conditions that lead to positive and negative outcomes
+for target and bycatch species from MPAs. We then present a case study
+tuned to reflect the dynamics of the tuna, billfish, and shark
+populations and fisheries of the Western and Central Pacific Ocean
+(WCPO) to explore the value of optimal MPA design in managing tradeoffs
+between maximizing bycatch conservation and minimizing economic impacts.
+We chose the WCPO as a case study as the modern science of MPAs was
 largely developed around coastal ecosystems, often characterized by
@@
 species caught as incidental "bycatch" in targeted fishing operations,
-such as oceanic whitetip sharks [@pacoureau2021]. Large pelagic MPAs are
-increasingly being proposed specifically to benefit these by-catch
-populations while minimizing losses or where possible providing benefits
-to fisheries [citation].
-
-In this paper we evaluate the potential conservation and fishery effects
-of MPAs within the pelagic context. We focus in particular on the
-potential for pelagic MPAs to provide conservation benefits to bycatch
-species. While the effect of MPAs on fisheries yields is a continuously
-debated topic, the conservation impacts of MPAs are seemingly simpler,
-albeit challenging to detect in many cases [@ovando2021a], due in part
-perhaps to the perceived obvious nature of their benefits: if you
-prevent fishing of a species within an area, there will be more of that
-species in that area, and so the conservation benefits of MPAs will
-increase with MPA size. \@hastings2017 clearly illustrates this idea.
-Using a simple two-species two-patch analytical model, they derive
-conditions under which MPAs can solve the "weak-stock" problem in which
-a less productive species is caught as btcatch by a fleet targeting a
-more valuable and productive stock.
-
-However, as @hastings2017 point out, their results depend on a very
-specific set of assumptions, namely that 1) adults are perfectly
-sedentary 2) the target and bycatch species overlap perfectly in terms
-of habitat and 3) recruits are perfectly mixed throughout the system.
-Pelagic species are likely to violate one or more fo these assumptions.
-We build on \@hastings2017 to consider the role of MPAs in reducing
-bycatch while managing tradeoffs with fisheries under a more resolved
-model reflecting the dynamics of pelagic ecosystems and fisheries.
-
-Pelagic ecosystems present both opportunities and challenges for MPAs.
-The highly productive and mobile nature of targeted fisheries in these
-systems may mean that highly targeted MPAs can be designed that provide
-effective conservation at minimal fishery costs. Strategic protection of
-key aggregation locations may enhance depleted pelagic fisheries [xx
-that Indian ocean study I can't remember right now]. However, the
-dynamic nature of pelagic systems also creates a risk of unintended
-conservation consequences of MPAs. When an MPA is implemented on former
-fishing grounds, fishing fleets have two choices: exit the fishery or
-move to new fishing grounds. When MPAs redistribute effort away from one
-area and into another, they can accidentally increase fishing effort on
-species found within the new fishing grounds [@abbott2012]. This problem
-of policy "leakage" is well known in many environmental fields
-[@citations] but the general properties of this problem have received
-relatively little attention in the theoretical or empirical MPA
-literature [see
-@abbott2012;@savina2008;[fulton2011a;\@baum2003](mailto:fulton2011a;@baum2003){.email}
-for notable exceptions].
-
-This study addresses this gap using a novel bio-economic simulation
-model, capable of reflecting several of the key challenges of pelagic
-ecosystems, namely high and variable movement rates, mobile fishing
-fleets, and heterogeneous species distributions. We first provide a set
-of stylized simulations intended to illustrate the potential and
-challenges of achieving conservation objectives through pelagic MPAs,
-and highlight particularly important drivers and the magnitude and
-direction of MPA effects. We then present a more focused case study
-based on the Western and Central Pacific Ocean, examining the potential
-outcomes of pelagic MPAs on bycatch and target species.
-
-<!-- We fill this gap by examining the general range of conditions… -->
-
-\<!-- \# Assessing the conditions under which this may be a problem
-using a generalized simulation environment
-
-General insights about design strategy, effect of varying degrees of
-habitat overlap, etc, tradeoffs with fisheries
-
--   Given that "it depends" nature of the problem, assess the specific
-    potential for this problem in a case study
-
-    -   specific insights about design strategy, effect of varying
-        degrees of habitat overlap, etc, tradeoffs with fisheries
-
-Our goal here is not to provide tactical recommendations on the design
-of blue-water MPAs, but rather to provide a strategic evaluation of the
-sensitivity of bycatch populations and targeted fisheries to key
-unknowns and design choices in MPA planning. --\>
-
-<!-- The outcome of MPAs for fishing and conservation are highly sensitive to -->
-
-<!-- a wide range of economic and ecological factors. In order to constrain -->
-
-<!-- the possible parameter space, we tune a simulation model based around -->
-
-<!-- best available knowledge of the species and fleets operating in the -->
-
-<!-- areas of the western central pacific ocean (WCPO) while allowing key -->
-
-<!-- unknowns to vary, in particular the nature of adult and larval movement, -->
-
-<!-- the degree of habitat overlap between different species, the economic -->
-
-<!-- incentives facing fishing fleets, and the design strategy used in MPA -->
-
-<!-- placement XX. Most conventional MPA design models are either static, -->
-
-<!-- single species, or both. We present a dynamic, multi-species simulation -->
-
-<!-- model that allows us to cohesively assess the impact of MPAs on a suite -->
-
-<!-- of by-catch and targeted species collectively affected by multiple -->
-
-<!-- fishing fleets.  -->
-
-<!-- We show - something about how when species are homogeneous in space, -->
-
-<!-- conservation gains are basically guaranteed except for constant catch -->
-
-<!-- scenearios, which need to be considered in tuna-type scenarios - When we -->
-
-<!-- introduce heterogeneity in stock structure roughly reflective of -->
-
-<!-- distributions in the WCPO we do / do not begin to see tradeoffs in -->
-
-<!-- conservation outcomes among bycatch species - quantify if there is a -->
-
-<!-- concave function to this - Specific design paradigms are / are not more -->
-
-<!-- / less resilient to this issue - Compare tradeoffs between MPA designs -->
-
-<!-- that minimze internal bycatch tradeoffs (there have to be none if MPA = -->
-
-<!-- 100%) and fishery outcomes -->
-
-<!-- We show that MPAs increased biomass of by-catch species in XX% of -->
-
-<!-- simulations, and decreased them in XX% of simulations. On average a 1% -->
-
-<!-- increase in by-catch species resulted in a XX% change in fishery -->
-
-<!-- catches... Outcomes were most sensitive to XX and XX, and least -->
-
-<!-- sensitive to XX and XX. We do not suggest that these same results would -->
-
-<!-- hold for other systems, but rather demonstrate the importance / -->
-
-<!-- resilience of MPAs to key uncertain variables. In particular for pelagic -->
-
-<!-- ecosystems, our work demonstrates the importance of considering the -->
-
-<!-- collective impacts of space, vulnerability, life history, and fleet -->
-
-<!-- dynamics in MPA design for conservation. -->
+such as oceanic whitetip sharks [@pacoureau2021]. These dynamics,
+combined with the increased policy interest in large MPAs covering
+expanses of blue-water ecosystems, make pelagic systems such as the WCPO
+an ideal case study in bycatch management through MPAs. The goal of this
+paper is to provide a strategic evaluation of the sensitivity of bycatch
+populations and targeted fisheries to key unknowns and design choices in
+MPA planning.
 
@@
 
+Below we provide a summary of the `marlin` model, as well as details of
+the generalized simulation exercises and case study analyses.
+
 ## `marlin`
 
-### Fleet Dynamics
+#### Fleet Dynamics
 
@@
 ability to capture technical interactions across species. While marlin
-does not simulate species interactions, fishing fleets in marlin are
+does not simulate species interactions, fishing fleets in `marlin` are
 capable of targeting and affecting multiple species simultaneously. This
@@
 (*Katswamus pelamus*), but the longline fleet is much more likely to
-catch bigeye than skipjack, and *vice versa*. The fishing fleet then
+catch bigeye than skipjack, and *vice versa*. Each fishing fleet
 generates catches, revenues, and profits from individual species that it
@@
 $$
- R_{t,p,f} = \sum_s^S \sum_a^A \Pi_{s,f}\frac{\alpha_{a,f}q_{s,f}E_{t,p,f}}{F_{t,p,s,a}} \times Y_{t,p,s,a}
+ R_{t,p,f} = \sum_s^S \sum_a^A \Pi_{s,f}\frac{\alpha_{a,s,f}q_{s,f}E_{t,p,f}}{F_{t,p,s,a}} \times Y_{t,p,s,a}
 $$
@@
 where *p* is patch, *s* is species, *a* is age, $\alpha$ is selectivity
-at age, *q* is catchability , *F* is total fishing mortality, and *Y*
-are total catches.
+at age, *q* is catchability , *F* is total fishing mortality, $\Pi$ is
+price per unit weight, and *Y* is total catches.
 
@@
 $$
-F_{t,p,s,a} = \sum_f^F\alpha_{a,f}q_{s,f}E_{t,p,f}
+F_{t,p,s,a} = \sum_f^F\alpha_{a,s,f}q_{s,f}E_{t,p,f}
 $$
 
-and
+Selectivity $\alpha$ is modeled through a logistic form per
+
+```{=tex}
+\begin{equation}
+\alpha_{a,s,f}=\frac{1}{(1 + e^{-log(19)\times\frac{l_{a,s} - l^{sel}_{s,f}}{\delta^{sel}_{s,f}}})}
+  (\#eq:sel)
+\end{equation}
+```
+where $l$ is length at age, $l^{sel}$ is the length at 50% selectivity,
+and $\delta$ is the difference between the length at 50% selectivity and
+the length at 95% selectivity.
+
+*z* is total mortality given by
 
@@
 catches and subsequently revenues for each fleet. Given revenues, we
-then calculate profits \$\\phi\$ per
+then calculate profits $\phi$ per
 
@@
 \phi_{t,p,f} = R_{t,p,f} - c_f(E_{t,p,f} + 1)_{t,p,f}
-^{1.3}$$
+^{1.3}
+$$
 
@@
 constrain the value to be greater than 1, preventing undesired behavior
-where fishing effort actually gets cheaper if total effort is less than
-1.
-
-The important thing to note here then is that the revenues and profits
-for a given fleet are a function of the prices and catchability of
-individual species, but that mortality itself is independent of price.
-So, fleet that has selectivity and catchability greater than zero can
-exert fishing mortality on a species even if it obtains no value from
-that species.
-
-The fleet then decides how to allocate its effort in space based on one
-of four possible spatial allocation strategies. The ideal free
-distribution (IFD) (citation) is the standard method for distributing
-fishing fleets in spaces. However, the IFD presents a number of
-complications for our study. First, in order to simplify the assessment
-of the impacts of different fishing mortality rates we simulate
-scenarios with a fixed total pool of effort. As MPAs change the spatial
-distribution of profits though, the total amount of effort required to
-satisfy the IFD may not equal the total amount of effort required by the
-model. Second, even if we were to implement the IFD for a single fleet,
-the model is capable of tracking multiple fleets at once. This means
-that the IFD for a single fleet would be a game-theoretic outcome of the
-actions of each of the individual fleet conditional on the actions of
-all other fleets. While possible to implement this would slow down our
-model runs to the point of practically preventing large-scale evaluation
-of the effects of MPAs on bycatch.
+where fishing effort actually gets less expensive if total effort is
+less than 1.
+
+The important feature here then is that the revenues and profits for a
+given fleet are a function of the prices and catchability of individual
+species, but that mortality itself is independent of price. So, a fleet
+that has selectivity and catchability greater than zero can exert
+fishing mortality on a species even if it obtains no value from that
+species.
+
+Each fishing fleet then decides how to allocate its effort in space
+based on one of four possible spatial allocation strategies. The ideal
+free distribution (IFD) is the standard method for distributing fishing
+fleets in spaces (see @gillis2003 and references therein). However, the
+IFD presents a number of complications for our study. First, in order to
+simplify the assessment of the impacts of different fishing mortality
+rates we simulate scenarios with a fixed total pool of effort. As MPAs
+change the spatial distribution of profits though, the total amount of
+effort required to satisfy the IFD may not equal the total amount of
+effort required by the model. Second, even if we were to implement the
+IFD for a single fleet, the model is capable of tracking multiple fleets
+at once. This means that the IFD for a single fleet would be a
+game-theoretic outcome of the actions of each of the individual fleet
+conditional on the actions of all other fleets. While possible to
+implement this would slow down our model runs to the point of
+practically preventing large-scale evaluation of the effects of MPAs on
+bycatch.
 
@@
 algorithms. While not the IFD in any individual time step, over time
-they start to approximate the IFD (check this) as each assumes that
-fleets base their decisions for the current time step on the outcomes in
-the prior timestep, meaning that the impacts of the actions of other
-fishing fleets are somewhat internalized. The four possible fleet
-distribution algorithms are
+they start to approximate the IFD, as each assumes that fleets base
+their decisions for the current time step on the outcomes in the prior
+timestep, meaning that the impacts of the actions of other fishing
+fleets are somewhat internalized. The four possible fleet distribution
+algorithms are
 
@@
 
-### Population Model
-
-Lots of equations. Pretty standard age-structured population models,
-extra bells and whistles that need explaining
-
--   Design of movement and habitat
-
-    -   Movement is a product of distance and habitat, such that
-        critters will all else equal move to the closest patches with
-        the best habitat
-
--   Ontogenetic shifts
-
--   Nature of fleet tuning (shared effort different q per species)
-
-Fleets exert a given amount of effort in a patch each time step. The
-catch resulting from that effort is given by the Baranov [@baranov1918]
-equation
-
-$$C_{f,s,p,t} = \frac{q_{f,s}E_{f,p,t}}{q_{f,s}E_{f,p,t} + m_s} \times \sum_{a=0}^{A}B_{s,p,t,a} \times (1 - e^{-(q_{f,s}E_{f,p,t}S_a + m_s)}) $$
-
-Where $C$ is catch by fleet $f$ of species $s$ in patch $p$ in timestep
-*t*. $E$ is total effort by that fleet in a patch and time step, $m$ is
-natrual mortality for a given species, $B$ is biomass at age *a*, and
-$S$ is selectivity.
-
-$q$ is the catchability coefficient for fleet $f$ and species $s$.
-
-Fleets distribute themselves in space according to revenue per unit
-effort
-
-$$R_{f,p,t} = \frac{\sum_{s=1}^{S}v_{f,s}C_{f,s,p,t}}{E_{f,p,t}}$$ Where
-$v$ is the price per kg of each species. Bycatch in the model is defined
-through the price. An unregulated species with no commercial value would
-have $v = 0$, whereas a regulated and prohibited species might actually
-have a negative price.
-
-Effort per fleet is set a a constant of 1 per patch, and the model then
-tunes $q_{f,s}$ such that given all the other parameters of the model
-the fishing mortality rate for each species matches the best available
-level for that species. This is what allows the model to capture bycatch
-dynamics: A fleet can distribute itself purely based on revenues from
-target species, but through the $q$ parameter still exert fishing
-mortality on other species, even if that species has a price of zero, if
-those bycatch species inhabit the same patches as the more valuable
-target species.
-
-###
-
-### MPA Design Strategies
-
-MPA design is governed by four general parameters: the number of cells
-to place inside an MPAs, and the design strategy used to place the MPAs,
-the species considered in the MPA planning process, and the accuracy of
-the information used in the planning process.
+Revenue based spatial distribution is not likely to be very realistic .
+However, due to the complexity of parameterizing cost functions fleet
+dynamics are often evaluated based on yield or revenue alone, and so we
+include those scenarios here to evaluate the potential implications of
+this choice. The decision on whether to allocate the fleet based on
+absolute or relative (per unit effort) is more complex. When effort
+represents the actions of separate and individual fishing actors (e.g.
+independent fishing vessels), a per-unit-effort strategy may be more
+realistic, in which fishers distribute themselves based on the expected
+catch of their individual efforts [@hilborn1987]. Conversely, a system
+defined by a sole owner seeking to maximize total profits might be
+better represented by a fleet model based on total profits
+
+#### Population Model
+
+The underlying population model used is an age structured single-species
+model in the manner of @ovando2021a.
+
+For the population model, numbers *n* at time *t* for age *a* are given
+by
+
+```{=tex}
+\begin{equation}
+n_{t,p,s,a}=\begin{cases}
+      = BH(ssb_{t-1,p,s,a}) & \text{if $a = 1$}\\
+     = n_{t-1,p,s,a-1}e^{-(z_{t-1,p,s,a-1}}, & \text{if $1< a < max(age)$}\\
+     =  n_{t-1,p,s,a}e^{-(z_{t-1,p,s,a})} + n_{t-1,a-1}e^{-(z_{t-1,p,s,a-1})}, & \text{if $a = max(a)$}
+  \end{cases}
+  (\#eq:pop)
+\end{equation}
+```
+where *BH* is the Beverton-Holt recruitment function, *ssb* is
+spawning-stock-biomass.
+
+where $l_a$ is the mean length at at age, $l_sel$ is the length at which
+on average 50% of individuals are selected by the fishery, and
+$\delta_{sel}$ are the additional units of length at which on average
+95% of fish are selected by the fishery.
+
+*ssb* is calculated by converting age to mean length, calculating weight
+at age, maturity at age, and then calculating spawning stock biomass as
+the sum of spawning potential at age in a given time step.
+
+```{=tex}
+\begin{equation}
+  l_{a,s} = l\infty_s\left(1 - e^{-k_s(a - a0_{s})}\right)
+  (\#eq:len)
+\end{equation}
+```
+Biomass *b* at age is then given by
+
+```{=tex}
+\begin{equation}
+  b_{a,s} = wa_s \times l_{a,s}^{wb_s}
+  (\#eq:weight)
+\end{equation}
+```
+and maturity *mat* is calculated as
+
+```{=tex}
+\begin{equation}
+  mat_{a,s} = \frac{1}{(1 + e^{-log(19)\times\frac{l_{a,s} - lmat_s}{{\delta}mat_s}}}
+  (\#eq:mat)
+\end{equation}
+```
+where $l_{mat}$ is the length at which on average 50% of individuals are
+sexual maturity, and $\delta_{mat}$ is the units of length beyond
+$l_{mat}$ at which on average 95% of fish are sexually mature.
+
+Spawning stock biomass at time *t* is then calculated as
+
+```{=tex}
+\begin{equation}
+  ssb_{t,p,s} = \sum_{a=1}^Aw_{a,s,t}^{\gamma_s} mat_{t,a} n_{t,p,s,a}
+    (\#eq:ssb)
+\end{equation}
+```
+##### Recruitment {.unnumbered}
+
+Recruitment follows Beverton-Holt dynamics. We do however allow for
+three variants in the timing of density dependence:
+
+1.  Global density dependence: Density dependence is a function of the
+    sum of spawning biomass across all patches, and recruits are then
+    distributed according to habitat quality
+
+```{=tex}
+\begin{equation}
+n_{t,p,s,a = 1} = \left(\frac{0.8{\times}\sum_{p=1}^P{r0_{p}}\times{h}\times\sum_{p=1}^P{ssb_{t-1,p}}}{0.2\times{\sum_{p=1}^Pssb0_p}\times(1 - h)+(h - 0.2)\times{\sum_{p=1}^Pssb_{t-1,p}}}\right)\times {hab_p}
+  (\#eq:dd2)
+\end{equation}
+```
+where *hab* is a vector of recruitment habitat quality by patch degined
+by
+
+$$hab_p = r0_p / \sum_p^P{r0_p}$$ 2. Local density dependence: Density
+dependence occurs independently in each patch and recruits are retained
+in their home patch.
+
+```{=tex}
+\begin{equation}
+  n_{t,p,s,a = 1,p} = \left(\frac{0.8{\times}r0_{s,p}\times{h_s}\times{ssb_{t-1,p,s}}}{0.2\times{ssb0_{p,s}}\times(1 - h_s)+(h_s - 0.2)\times{ssb_{t-1,p}}}\right)
+  (\#eq:dd1)
+\end{equation}
+```
+3.  Local density dependence then disperse: Density dependence occurs
+    independently in each patch and recruits are then dispersed.
+
+```{=tex}
+\begin{equation}
+  n_{t,p,s,a = 1,p} = \left(\frac{0.8{\times}r0_{s,p}\times{h_s}\times{ssb_{t-1,p,s}}}{0.2\times{ssb0_{p,s}}\times(1 - h_s)+(h_s - 0.2)\times{ssb_{t-1,p}}}\right)\times \boldsymbol{d^l} \times \epsilon_t
+  (\#eq:dd3)
+\end{equation}
+```
+where **d^l^** is the recruitment movement matrix
+
+4.  Post-dispersal density dependence: Larvae are distributed throughout
+    the system, and then density dependence occurs based on the density
+    of adult biomass at the destination patch.
+
+```{=tex}
+\begin{equation}
+  larv_{t,p} = ssb_{t-1} \times\boldsymbol{d^l}
+  (\#eq:larvmove)
+\end{equation}
+```
+```{=tex}
+\begin{equation}
+n_{t,a = 1,p} = \left(\frac{0.8{\times}r0_{p}\times{h}\times{larv_{t,p}}}{0.2\times{ssb0_p}\times(1 - h)+(h - 0.2)\times{larv_{t,p}}}\right) \times \epsilon_t
+  (\#eq:dd4)
+\end{equation}
+```
+##### Dispersal {.unnumbered}
+
+Dispersal in the model is broken into two components: adults and
+recruits Both assume a Gaussian dispersal kernel of the form
+
+```{=tex}
+\begin{equation}
+m_{s,p_i,p_j}=\frac{1}{\sqrt{2\pi\sigma_{s,p_i}^2}}e^{-\frac{d_{p_i,p_j}^2}{2\sigma_{s,p_i}^2}}
+  (\#eq:move)
+\end{equation}
+```
+where *i* is the source patch, *j* is the destination patch, *d* is the
+distance between patches *i* and *j* , and $\sigma_s*$is the movement
+rate, in units of patches.
+
+Along with distance, net movement is defined by habitat quality. Habitat
+quality defines the quailty for adults or larvae of a given patch. The
+net probability of movement from patch *i* to patch *j* is defined then
+by the product of distance and habitat quality, normalized to sum to one
+from each patch, such that all else being equal organisms will move the
+most to the closest patch with the best habitat. We allow for seasonal
+movement in the model, where years are broken into two seasons. Movement
+in season two is defined as the opposite of movement in season one,
+intended to reflect the behavior of seasonal migrations.
+
+## Rule-of-thumb MPA Design Strategies
+
+MPA design is governed by two general parameters: the number of cells to
+place inside an MPAs, and the design strategy used to place the MPAs.
 
@@
 
-For the "depletion" strategy, MPA locations are prioritized based in
-proportion to the spawning stock biomass (SSB) of the most depleted
-species. So, patches with high amounts of highly depleted species are
-prioritized over patches with high biomass of lightly depleted species.
-
-For the "rate" strategy, MPAs are placed in based on the
-depletion-weighted catch relative to the total catch in a patch. So,
-patches where with high rates of catch of depleted species relative to
-totcal catch are prioritized.
-
-For the "avoid_fishing" strategy, MPAs locations are prioritized
-inversely to fishery catches. So, patches with low catches are
-prioritized over patches with high catches.
-
-For the "target_fishing" strategy, MPAs locations are prioritized
-proportionally to fishery catches. So, patches with high total catches
-are prioritized over patches with high catches.
-
-For the "area" strategy, we simply assign patches in sequential order
-until the desired total MPA size is reached.
-
-The information used to inform each MPA design strategy (except area)
-can include a multiplicative observation error term $\sigma_{obs}$ where
-
-$$information \sim lnorm(data,\sigma_{obs}) $$
-
-For example, for the depletion scenario, $\sigma_{obs}$ controls the
-accuracy of the spatial depletion data used to prioritize MPA placement.
+1.  Depletion: MPA locations are prioritized in proportion to the
+    spawning stock biomass of the most depleted species pre-MPA. So,
+    patches with high spawning biomass of highly depleted species are
+    prioritized over patches with high spawning biomass of lightly
+    depleted species.
+
+2.  Rate: MPAs are placed in based on the pre-MPA depletion-weighted
+    catch relative to the total catch in a patch. So, patches with high
+    rates of catch of depleted species relative to total catch are
+    prioritized.
+
+3.  Avoid_fishing: MPAs locations are prioritized inversely to fishery
+    catches. So, patches with low catches are prioritized over patches
+    with high catches.
+
+4.  Target_fishing: MPAs locations are prioritized proportionally to
+    fishery catches. So, patches with high total catches are prioritized
+    over patches with low catches.
+
+5.  Area: For the "area" strategy, we simply assign patches in
+    sequential order until the desired total MPA size is reached.
+
+<!-- The information used to inform each MPA design strategy (except area) -->
+
+<!-- can include a multiplicative observation error term $\sigma_{obs}$ where -->
+
+<!-- $$information \sim lnorm(data,\sigma_{obs}) $$ -->
+
+<!-- For example, for the depletion scenario, $\sigma_{obs}$ controls the -->
+
+<!-- accuracy of the spatial depletion data used to prioritize MPA placement. -->
+
+## MPA Optimization
+
+Our MPA optimization algorithm seems to create an MPA network that
+maximizes an objective function, given by
+
+$$obj_p = \Phi \times scale(\sum_f^F \boldsymbol{\phi_{f,p}}) + (1 - \Phi)  scale(\boldsymbol{\sum_s^S\frac{ssb_s}{ssb0_{s,p}}})$$
+
+Where *scale* rescales the values to be between 0 and 1 such that the
+weight parameter is comparable between vectors. The algorithm works by
+first assigning every patch in the system to an MPA. For the first
+round, the algorithm then calculates the equilibrium objective function
+value resulting from sequentially removing each individual patch from
+the MPA network. The cell whose removal from the MPA produces the
+highest objective function value is then removed from the MPA network,
+and the marginal value of the objective function each of the remaining
+cells from the last steps are then stored. For the next steps, we
+sampled 25% of the remaining MPA cells with probability proportional to
+their marginal contribution to the objective function in the prior
+iteration, selecting the cell whose removal produces the highest total
+objective function as the cell to remove from the network, and updateing
+the marginal value of the sampled-but-not-selected cells. This in
+essence is a sampling-importance-resampling algorithm. The algorithm
+continues until no more cells are left in the MPA. This then provides
+for a given objective weighting value $\Phi$ the near-optimal MPA
+network at every MPA size from 0 to 100%. We say "near" as they is some
+path dependency here, in that for example the network that maximizes the
+objective function with 40% of the MPAs does not necessarily contain the
+optimal network at 30%. However, the reversed nature of the algorithm,
+starting at 100% MPA coverage and working better, produces a more stable
+network with regards to this problem than starting at 0% and growing to
+100%.
 
@@
 
-### Species & Fleets
-
-Each simulation considers `n_distinct(marlin_inputs$scientific_name)`
-species comprised of the top 12 species by landings reported in the tuna
-fisheries of the Western and Central Pacific Ocean (WCPO). Rough Life
-history parameters are pulled from `FishLife` [@thorson2020]. For
-teleost fishes, we assumed that fecundity was proportional to
-weight-at-age times the maturity ogive. For cartilaginous species, we
-adopt a fixed fecundity relationship of XX pups per mature individual XX
-check on what I actually implemented here XX. The unfished spawning
-stock biomass for each of the included species was tuned in the model to
-reflect best available estimates from stock assessments.
-
-Each simulation includes a longline and a purse-seine fleet. Each
-species can have a separate price (USD/kg) and selectivity ogive for
-each fleet, as dictated by the literature.
-
-### Species Overlap
+Each iteration uses one of three species: A stylized tuna modeled after
+skipjack tuna (), a stylized marlin modeled after (), and a stylized
+shark modeled after (). Each species has all life history parameters
+pulled from FishLife (@thorson2020), with the exception of steepness, set
+to XX, and fecundity. For the tuna and marlin species, fecundity is
+assumed to scale with weight modified by the hyperallometric exponent.
+For the shark species, we assume a constant value of six pups per year
+at all ages greater than the age at maturity. We assume that tunas have
+an unfished biomass of XX, marlins XX, and sharks XX, to reflect a
+large, medium, and small population.
+
+For the generalized experiments we only include one fishing fleet. We
+asign a price per KG of \$XX for tuna, intended to reflect a low value
+high volume fishery, \$XX for marlin for a higher value low volume
+fishery, and \$0 for sharks to reflect at true bycatch species. We
+assign a selectivity of 60% of length at maturity for skipjack to allow
+for some overfishing when F/M is high enough, 70% for mariln, and 90%
+for shark.
+
+We ran `r nrow(state_experiments)` state experiments, varing the state
+of nature experienced by a simulation. The candidate states include
+
+| State                 | Distribution          | Description                                     |
+|-----------------------|-----------------------|-------------------------------------------------|
+| $\sigma_{centroid}$   |                       | Standard deviation of habitat centroid          |
+| $\sigma_{habitat}$    |                       | Standard deviation around centroid              |
+| Spatial catchability  |                       | Catchability varies in sapce                    |
+| Spatial allocation    |                       | Fleet allocation strategy                       |
+| F/M                   |                       | Fishing mortality relative to natural mortality |
+| Adult movement rate   |                       |                                                 |
+| Recruit movement rate |                       |                                                 |
+| Hyperallometry        |                       | Presence of hyperallometric fec                 |
+| DD timing             |                       | Timing of density dependence                    |
+| Seasonal Movement     | $\subset[TRUE,FALSE]$ | Seasonal Movement Present                       |
+
+For each of these `r nrow(state_experiments)` state experiments, we then
+apply `r nrow(placement_experiments)` placement experiments defined by
+an MPA size, an MPA placement strategy, and ...XX
 
@@
 
-### Seasonal Migration
-
-For this paper `marlin` is set to run with two seasons per year. When
-seasonal movement is turned on, animals use one habitat matrix in season
-one, and the inverse of that habitat matrix in season 2.
-
-## Case Study Application
-
-In order to constrain the parameter space somewhat, we also ran a series
-of case study experiments. Rather than iterating over different habitat
-configurations and movement dynamics, we defined habitat based on
-reported catch per unit effort (CPUE) from the WCPO. XX more methods
-here XX
-
-### MPA Network Optimization
-
-For the case study section, as a point of comparison we calculate an
-optimized MPA network along with the "rule of thumb" networks based on
-the design strategies outlined in section XX. The optimization routine
-is based on an objective function of the form
-
-$$obj = \alpha \times scale(\sum_{s=1}^{S}\frac{ssb_s}{ssb0_s}) + (1 - \alpha) \times scale(\sum_{s=1}^{S}R_s) $$
-
-where $\alpha$ controls the relative weight assigned to a "conservation"
-objective based on depletion, relative to an "economic" objective based
-on revenues. The 'scale' function rescales both objective components to
-the range of 0 to 1 to facilitate interpretation of $\alpha$ xx come up
-with a better version of this XX
-
-The optimization routine itself works through a
-sampling-importance-resampling algorithm. In the first iteration of the
-algorithm, the marginal contribution of each individual patch to the
-objective function is calculated. The patch with the highest objective
-function value is then placed in an MPA. For the remaining iterations,
-we sample *X* cells from the remaining non-MPA cells with probability
-proportional to their marginal value to the objective function in the
-last iteration. We then calculate the marginal value of each of these
-cells when individually added to the MPA network begun in the prior
-step. The cell with the higheset marginal value is then added to the MPA
-network, and the marginal values of the non-selected but sampled cells
-are updated based on the values in that iteration. The process then
-continues until the desired MPA size is reached.
+## WCPO Case Study Application
+
+The purpose of the case study is to assess the value of optimization.
+Given the time constraints of optimization, rather than iterating over
+the `r nrow(state_experiments)`, we fix the values from that table to
+values that best reflect the state of
+`r n_distinct(experiment_obj$critter)` species reported as caught in the
+WCPO.
+
+Sigh. lots of details to add in here. This is the worst.
+
+```{r, echo = FALSE}
+knitr::kable(sci_to_com %>%
+               rename("Scientific Name" = "critter", "Common Name" = common_name))
+```
 
@@
 `r years` years (`r seasons * years` seasons. We assessed performance by
-comparing relevant metrics (e.g. depletion and revenue) under a given
+comparing relevant metrics (e.g. depletion and profits) under a given
 scenario with MPAs relative to the values of those same metrics in the
@@
 
+Degree of habitat difference was calculated as
+
+xx will work out the notation for this xx
+
+$$
+ Habitat Difference = sd(sqrt(hab))
+$$
+
+In order to weight habitat difference by cells where animals occur, not
+where they are not.
+
 # Results
@@
 Our simulation experiments allow us to examine the wide range of effects
-that MPAs can have on groups of species depending on combinations of
-ecological, economic, and design parameters. The goal of this exercise
-is not to assign any probability to the potential impacts of MPAs; we
-have not provided any explicit probabilites to the alternative states
-modeled here, and so it may be that a state that was only modeled on a
-few occasions in our simulations is in fact highly common in the real
-world.
+that MPAs can have on groups of target and bycatch species depending on
+combinations of ecological, economic, and design parameters. The goal of
+this exercise is not to assign any probability to the potential impacts
+of MPAs; we have not provided any explicit probabilities to the
+alternative states modeled here, and so it may be that a state that was
+only modeled on a few occasions in our simulations is in fact highly
+common in the real world.
 
@@
 evaluated here, and evaluate which drivers are potentially the most
-influential in affecting particular outcomes, such as unintended
-conservation consequences. Our case study of a WCPO-esque system is
-intended then to reduce the parameter space enough that we can examine
-the relative tradeoffs and value of optimization for economic and
-conservation objectives in a pelagic system.
+influential in affecting particular outcomes with regards to bycatch.
+Our case study of a WCPO-esque system is intended then to reduce the
+parameter space enough that we can examine the relative tradeoffs and
+value of optimization for economic and conservation objectives in a
+pelagic system.
 
@@
 (citation), and moreover a motivation for this study was to evaluate the
-potential unintended consequences of MPAs from effort displacement on
-target and bycatch species. If MPAs do not produce increases in fishing
-effort outside their borders, then the conservation effects of MPAs have
-to be positive for all species concerned.
+performance of MPAs in reducing bycatch given the potential for effort
+displacement and technical interactions. If MPAs do not produce
+increases in fishing effort outside their borders, then the direct
+conservation effects of MPAs have to be positive for all species
+concerned.
 
@@
 We focused first on the range of impacts that MPAs can have on mature
-biomass across the range of variables explored. To illustrate, we pulled
-out two example simulations, one where MPAs provided relatively large
-benefits across all three species, and a second where MPAs actually
-produced substantial losses in some species. In the "best case"
-scenario, each of the three species shares a common habitat
-distribution, and so protection of the central habitat of one species
-results in protection for all species. In the "worst case" scenario,
-each of the species have . An MPA designed primarily around the tuna and
-marlin populations then ends up displacing effort into the shark
-habitat, producing benefits for the first two species but driving the
-shark population down XX (Fig.\@ref(fig:fig-info)).
-
-```{r fig-info, fig.cap="Illustration of ways in which variables can interact to affect MPA outcomes."}
-
-knitr::include_graphics(file.path(results_path,"fig_info.pdf"))
+biomass of target and bycatch species across the range of variables
+explored. To illustrate, we pulled out two example simulations, one
+where MPAs provided relatively large benefits across all three species,
+and a second where MPAs actually produced substantial losses in some
+species. In the "best case" scenario, each of the three species shares a
+common habitat distribution, and so protection of the central habitat of
+one species results in protection for all species. In the "worst case"
+scenario, an MPA designed primarily around the tuna fishery displaces
+effort onto the marlin and shark habitats, increasing the tuna
+population while driving the other two species down
+(Fig.\@ref(fig:fig-info)).
+
+```{r fig-info, fig.cap="Illustrative example demonstrating how degrees of overlap can influence "}
+
+experiment_summary <- total_experiment_results %>%
+  filter(prop_mpa < 0.5,
+         !placement_strategy %in% c("area", "avoid_fishing"),
+         placement_strategy == "rate",
+         placement_error == min(placement_error)) %>%
+  group_by(placement_id, state_id) %>%
+  summarise(
+    max_loss = max(signif_loss),
+    min_loss = max(n_loss),
+    total_gain = sum(abs_delta_biodiv)
+  ) %>%
+  ungroup()
+
+worst_case <- experiment_summary %>%
+  # filter(max_loss == max(max_loss)) %>%
+  arrange(total_gain) %>%
+  # filter(total_gain == min(total_gain)) %>%
+  slice(2)
+
+worst_state <- state_experiments %>%
+  filter(state_id == worst_case$state_id)
+
+worst_result <- compact_experiment_results %>%
+  filter(placement_id == worst_case$placement_id,
+         state_id == worst_case$state_id,
+         prop_mpa <0.5)
+
+worst_result_plot <- worst_result %>%
+  ggplot() +
+  geom_text(data = tibble(
+    x = .1,
+    y = max(worst_result$biodiv - worst_result$bau_biodiv)
+  ),
+  aes(x = x, y = y, label = "Worst Case"),
+  size = 4) +
+  geom_hline(yintercept = 0, linetype = 2) +
+  geom_line(aes(prop_mpa, biodiv - bau_biodiv, color = critter),
+            show.legend = FALSE, size = 1.25) +
+  scale_x_continuous(labels = scales::percent, name = "MPA Size") +
+  scale_y_continuous(
+    name = ''
+  )
+
+worst_habitat <- worst_state %>%
+  select(state_id, data) %>%
+  unnest(cols = data) %>%
+  select(scientific_name, contains("sigma_"), habitat) %>%
+  unnest(cols = habitat) %>%
+  mutate(
+    scientific_name = fct_recode(
+      scientific_name,
+      "Shark" = "carcharhinus longimanus",
+      "Tuna" = "katsuwonus pelamis",
+      "Marlin" = "kajikia audax"
+    )
+  )
+
+worst_habitat_plot <- worst_habitat %>%
+  ggplot(aes(x, y, z = (habitat), color = scientific_name)) +
+  geom_contour(show.legend = FALSE) +
+  scale_x_continuous(expand = c(0, 0)) +
+  scale_y_continuous(expand = c(0, 0)) +
+  theme(
+    axis.text.x = element_blank(),
+    axis.text.y = element_blank(),
+    axis.title.y = element_blank(),
+    axis.title.x = element_blank(),
+    axis.ticks.y = element_blank(),
+    axis.ticks.x = element_blank(),
+    panel.border = element_blank()
+  )
+
+best_case <- experiment_summary %>%
+  filter(min_loss == min(min_loss, na.rm = TRUE)) %>%
+  filter(total_gain == max(total_gain, na.rm = TRUE)) %>%
+  slice(1)
+
+best_state <- state_experiments %>%
+  filter(state_id == best_case$state_id)
+
+best_result <- compact_experiment_results %>%
+  filter(placement_id == best_case$placement_id,
+         state_id == best_case$state_id,
+         prop_mpa < 0.5)
+
+best_result_plot <- best_result %>%
+  ggplot() +
+  geom_hline(yintercept = 0, linetype = 2) +
+  geom_line(aes(prop_mpa, biodiv - bau_biodiv, color = critter), size = 1.25) +
+  geom_text(data = tibble(
+    x = .1,
+    y = max(best_result$biodiv - best_result$bau_biodiv)
+  ),
+  aes(x = x, y = y, label = "Best Case"),
+  size = 4) +
+  scale_x_continuous(labels = scales::percent, name = "") +
+  scale_y_continuous("Change in SSB/SSB0") +
+  theme(legend.position = "top") +
+  scale_color_discrete(name = element_blank()) +
+  labs(title = "A) ")
+
+best_habitat <- best_state %>%
+  select(state_id, data) %>%
+  unnest(cols = data) %>%
+  select(scientific_name, contains("sigma_"), habitat) %>%
+  unnest(cols = habitat) %>%
+  mutate(
+    scientific_name = fct_recode(
+      scientific_name,
+      "Shark" = "carcharhinus longimanus",
+      "Tuna" = "katsuwonus pelamis",
+      "Marlin" = "kajikia audax"
+    )
+  )
+
+best_habitat_plot <- best_habitat %>%
+  ggplot(aes(x, y, z = (habitat), color = scientific_name)) +
+  geom_contour(show.legend = FALSE) +
+  labs(title=  "B)", subtitle = "Habitat Distribution") +
+  scale_x_continuous(expand = c(0, 0)) +
+  scale_y_continuous(expand = c(0, 0)) +
+  theme(
+    axis.text.x = element_blank(),
+    axis.text.y = element_blank(),
+    axis.title.y = element_blank(),
+    axis.title.x = element_blank(),
+    axis.ticks.y = element_blank(),
+    axis.ticks.x = element_blank(),
+    panel.border = element_blank()
+  ) # +
+  # facet_wrap(~scientific_name)
+
+fig_info <- (best_result_plot / worst_result_plot) |
+  (best_habitat_plot / worst_habitat_plot)
+
+fig_info
+
 ```
@@
   mutate(critter = fct_relevel(critter,c("Tuna", "Marlin","Shark")),
-         rec_form = fct_recode(rec_form, "Global DD, Rec. By Habitat" = "0", "Local DD, local Recruitment" = "1", "Local DD, Disperse Recruits" = "2", "Disperse Larvae, Poset-Settlement DD" = "3"))
+         rec_form = fct_recode(rec_form, "Global" = "0", "Local" = "1", "Local Then Disperse" = "2", "Disperse Then Local" = "3"))
 
@@
 `r n_distinct(total_experiment_results$placement_id) * n_distinct(total_experiment_results$state_id)`
-experiments run, the results were highly variable, with MPAs capable of
-producing substantial conservation gains to substantial conservation
-losses and everything in between. Across our simulations, gains were
-much more common than losses, with losses of any magnitude occurring in
-`r scales::percent(mean(critter_results$loss))` of simulations with
-MPAs, but losses of 10% or greater of SSB in only
+experiments run, the results were highly variable. Across our
+simulations, gains were much more common than losses, with losses of any
+magnitude occurring in `r scales::percent(mean(critter_results$loss))`
+of simulations with MPAs, but losses of 10% or greater of SSB in only
 `r scales::percent(mean(critter_results$real_loss))` of simulations.
@@
 MPAs of a given size impact spawning biomass depends on a host of
-factors. The potential for conservation losses are intuitively convex in
+factors. Any potential for conservation losses are intuitively convex in
 MPA size: effects of any kind are low when MPA size is low, and
 conservation losses must be zero when MPA size is 100% (which this graph
-doesnot quite reach), with peak losses in our simulations occuring
+does not quite reach), with peak losses in our simulations occurring
 between MPA sizes of XX and XX. Life history plays a clear role our
@@
 
-```{r fig-exp-biodiv}
+```{r fig-exp-biodiv, fig.cap = "Trajectories of changes in SSB/SSB0 relative to BAU for each species across rule-of-thumb MPA design scenarios. Color indicates degree of habitat difference."}
 
@@
   geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
-  facet_grid(critter ~ placement_strategy) +
+  facet_grid(critter ~ placement_strategy, labeller = labeller(placement_strategy = capitalize)) +
   scale_color_viridis_c(name = "Habitat Difference", guide = guide_colorbar(frame.colour = "black",ticks.colour = "black", barwidth = unit ("15", "lines"))) +
@@
   theme(legend.position = "top")
-
-# 
-# compact_experiment_results %>%
-#   filter(prop_mpa < 1) %>%
-#   # filter(f_v_m > 1.5) %>%
-#   group_by(state_id, placement_id) %>%
-#   ggplot(aes(
-#     prop_mpa,
-#     econ - bau_econ,
-#     color =  f_v_m,
-#     group = interaction(placement_id, state_id)
-#   )) +
-#   geom_line(alpha = 0.5) +
-#   geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
-#   facet_grid(critter ~ placement_strategy, scales = "free_y") +
-#   scale_color_viridis_c(name = "Habitat Difference", guide = guide_colorbar(frame.colour = "black",ticks.colour = "black", barwidth = unit ("15", "lines"))) +
-#   scale_y_continuous("Change in SSB/SSB0 relative to BAU") +
-#   scale_x_continuous(labels = scales::label_percent(accuracy = 1), name = "MPA Size") +
-#   theme(legend.position = "top")
 
@@
 decrease the probability of a conservation loss from an MPA are MPA
-size, placement strategy based on local depletion (xx might remove this
-strategy as unrealistic xx), and somewhat interestingly fishing
-mortality rate (Fig.\@ref(fig:fig-loss-drivers)-A).
+size, placement strategy based on local depletion, and somewhat
+interestingly fishing mortality rate
+(Fig.\@ref(fig:fig-loss-drivers)-A).
 
@@
 that habitat differences are a necessary but not sufficient condition
-for MPA leakage induced conservation losses. The regression based
-results require assuming that each of the effects are independent and
-linear in nature. In fact each of these factors can interact in complex
-and non-linear ways that may not be apparent visually or through this
-type of regression analysis. To address this, we ran a random forest
-using the `ranger` package in R (citaitons) and extracted the impurity
-corrected importance scores for each of the variables. While this does
-not tell us the expected marginal effect of individual drivers, it does
-tell us which variables are critical if our goal is to accurately
-predict the potential for unintended consequences or magnitude of
-conservation gains in MPAs. Note here that variables that intuitively
-should have an impact but have no visually or statistically (in a linear
-model) clear impact on the probability of experiencing a conservation
-loss such as the dispersal distances of adults or larvae have importance
-scores comparable in magnitude to habitat differences when considered in
-the context of nonlinear effects and interactions with other variables
+for conservation losses induced by effort displacement. The regression
+based results require assuming that each of the effects are independent
+and linear in nature. In fact each of these factors can interact in
+complex and non-linear ways that may not be apparent visually or through
+this type of regression analysis. To address this, we ran a random
+forest using the `ranger` package (citation) in R (citation) and
+extracted the impurity corrected importance scores for each of the
+variables. While this does not tell us the expected marginal effect of
+individual drivers, it does tell us which variables are important if our
+goal is to accurately predict the potential for unintended consequences
+or magnitude of conservation gains in MPAs. Note here that variables
+that intuitively should have an impact but have no visually or
+statistically (in a linear model) clear impact on the probability of
+experiencing a conservation loss such as the dispersal distances of
+adults or larvae have importance scores comparable in magnitude to
+habitat differences when considered in the context of nonlinear effects
+and interactions with other variables
 (Fig.\@ref(fig:fig-loss-drivers)-B).
 
-```{r fig-loss-drivers}
-
-# loss_reg <-   rstanarm::stan_glm(
-#   loss ~   f_v_m  + critter + seasonal_movement + adult_movement_sigma + rec_form + recruit_movement_sigma + prop_mpa + sigma_habitat + spatial_allocation + placement_strategy + hyper,
-#   family = binomial(),
-#   data = critter_results %>% filter(between(prop_mpa,.2,.4)),
-#   cores = 4,
-#   chains = 4
-# )
+```{r fig-loss-drivers, fig.cap = "Things"}
 
@@
   mutate(fctr_loss = factor(loss))
-# 
-# tmp_rcp <- recipes::recipe(loss ~ ., data = tmp_critter) %>% 
-#   themis::step_upsample(fctr_loss, over_ratio = 0.75) %>% 
-#   recipes::prep(training = tmp_critter, retain = TRUE) %>% 
-#   recipes::juice()
 
@@
 
-The generalized experiments help us understand the potential impacts of
-the evaluated drivers, but also highlight the challenge in making
-definitive statements about the effect of a given driver on the outcomes
-of pelagic(?) MPAs, given the complex ways in which all of these factors
-interact, even within a relatively simple model such as this one. In
-order to help illustrate how addressing the kinds of spatial dynamics
-and technical interactions included in our model can affect and benefit
-marine spatial planning, we then ran a case study intended to loosely
-reflect the dynamics of WCPO tuna, billfish, and shark fisheries and
-populations.
-
-This case study allows us to focus in on important questions around the
+<!-- The generalized experiments help us understand the potential impacts of -->
+
+<!-- the evaluated drivers, but also highlight the challenge in making -->
+
+<!-- definitive statements about the effect of a given driver on the conservation impacts of MPAs, given the complex ways in which all of these factors -->
+
+<!-- interact, even within a relatively simple model such as this one. In -->
+
+<!-- order to help illustrate how addressing the kinds of spatial dynamics -->
+
+<!-- and technical interactions included in our model can affect and benefit -->
+
+<!-- marine spatial planning, we then ran a case study intended to loosely -->
+
+<!-- reflect the dynamics of WCPO tuna, billfish, and shark fisheries and -->
+
+<!-- populations. -->
+
+Our case study allows us to focus in on important questions around the
 relative performance of "rule of thumb" MPA design strategies relative
-to optimized networks. We ran a spatial optimization routine (see
-section XX) across a range of objective function weightings, in the
-manner of @rassweiler2014. This allowed us to assess 1) the potential
-risk of unintended conservation consequences from MPAs in this WCPO
-bycatch study and the 2) assess value of optimization in this WCPO case
-study in achieving economic and conservation objectives.
+to optimized networks. In particular, as any leakage induced losses from
+MPAs must be 0 if MPA size is 100%, an important question then is within
+this case study what tradeoffs are present between increasing the
+probability of a successful conservation outcome by increasing MPA size
+and the increasing probability of negative economic impacts with MPA
+size (xx phrase better xx).
 
@@
 risk for large-scale unintended losses in total spawning biomass caused
-by MPA effort displacement in the WCPO case study. Each of the design
-scenarios produced at least some small loss in biomass for one or more
-species at some MPA size. However, these losses were generally small and
-isolated to a subset of species within a small MPA size range, with the
-exception of the "Avoid_fishing" MPA placement strategy that resulted in
-net losses for most species at large MPA sizes (though simply minimal
-impact at smaller sizes). All design strategies except "Avoid_fishing"
-and "Rate" resulted in relatively rapid declines in profits for the
-three main tuna fisheries (skipjack, yellowfin, and bigeye). Both
-"Avoid_fishing" and "Rate" mainlined profits close to business as usual
-up to a large MPA size, but did so at the expense of conservation
-benefits (Fig.\@ref(fig:fig-cs-critters)).
+by MPA effort displacement for the species included in our WCPO case
+study. Each of the design scenarios produced at least some small loss in
+biomass for one or more species at some MPA size. However, these losses
+were generally small and isolated to a subset of species within a small
+MPA size range, with the exception of the "Avoid_fishing" MPA placement
+strategy that resulted in net losses for most species at large MPA sizes
+(though simply minimal impact at smaller sizes). All design strategies
+except "Avoid_fishing" and "Rate" resulted in relatively rapid declines
+in profits for the three main tuna fisheries (skipjack, yellowfin, and
+bigeye). Both "Avoid_fishing" and "Rate" maintained profits close to
+business as usual up to a large MPA size, but did so at the expense of
+conservation benefits (Fig.\@ref(fig:fig-cs-critters)). Results for
+yield rather than profits are qualitatively similar, see SI.
 
 ```{r fig-cs-critters, fig.cap = "Simulated trajectories of changes in profits and spawning stock biomass (SSB) for selected species caught in the WCPO.
-Each column represents a different MPA design strategy, see Table.XX for definitions.
-Optimized refers to an MPA network optimized to maximize total SSB across all species."}
+Each column represents a different MPA design strategy.
+'Optimized' refers to an MPA network optimized to maximize total SSB across all species."}
 
@@
          placement_strategy = "Optimized") %>%
-  select(critter, placement_error, delta_biodiv, delta_econ, prop_mpa,placement_strategy)
-
-capitalize <- function(string) {
-  substr(string, 1, 1) <- toupper(substr(string, 1, 1))
-  string
-}
+  select(critter, placement_error, delta_biodiv, delta_econ, delta_yield, prop_mpa,placement_strategy)
 
@@
 for optimized networks, with the "Depletion" design scenario essentially
-following the Pareto frontier (Fig.\@ref(fig:fig-cs-totals)).
+following the Pareto efficiency frontier (Fig.\@ref(fig:fig-cs-totals)).
 
@@
 they can also provide benefits to food security and economic outcomes.
-However, while extremely simple in concept, predicting the likely
-effects of MPAs across groups of species and fishing fleets becomes
-increasingly complicated as the systems being studied deviate from
-assumptions of homogenously distributed species with perfectly selective
-fleets. In this paper we used a new bio-economic model to demonstrate
-how a range of variables often omitted in MPA modeling can act to
-influence the outcomes of MPAs on conservation and economic objectives.
-We demonstrate that unintended conservation losses are possible, but
-under a range of specific conditions, and under most simulated cases did
-not occur. MPAs were capable of causing small losses in biomass for the
-WCPO case study, but only for a subset of design strategies and species,
-most of which could be overcome by optimized design.
+However, while spatial closures are simple in concept, predicting the
+likely effects of MPAs across groups of species and fishing fleets
+becomes increasingly complicated as the systems being studied deviate
+from assumptions of homogenously distributed species with perfectly
+selective fleets. In this paper we used a new bio-economic model to
+demonstrate how a range of variables often omitted in MPA modeling can
+act to influence the outcomes of MPAs on conservation and economic
+objectives, particularly with regards to protecting bycatch species. We
+demonstrate that unintended conservation losses to both target and
+bycatch species are possible, but only under a range of specific
+conditions, and under most simulated cases did not occur. MPAs were
+capable of causing small losses in biomass for the WCPO case study, but
+only for a subset of design strategies and species, most of which could
+be overcome by optimized design.
 
@@
 rather than attrition. However, our model found these to be a minority
-of the simulations tested. This does in no way suggests that leakage may
-not a problem for specific MPA. But, it also indicates that across the
-range of plausible parameter combinations that we explored it was much
-more common for MPAs to provide net conservation benefits even if effort
-is displaced. This suggests then that we should of course not dismiss
-the conservation benefits of MPAs in the presence of effort
-displacement, but also that users must be aware of the potential for
-unintended consequences that can indeed occur. While many different
-factors interact in complex ways to affect the potential for unintended
+of the simulations tested. This does not suggest that leakage may not a
+problem for specific MPA. But, it also indicates that across the range
+of plausible parameter combinations that we explored it was much more
+common for MPAs to provide net conservation benefits even if effort is
+displaced. This suggests then that we should of course not dismiss the
+conservation benefits of MPAs in the presence of effort displacement,
+but also that users must be aware of the potential for unintended
+consequences that can indeed occur. While many different factors
+interact in complex ways to affect the potential for unintended
 conservation consequences, the most obvious is the degree of difference
@@
 MPA is to be placed. As a starting point, it is critical that
-communities engaged in MSP processes create a localized assessment of
+communities engaged in marine spatial planning create a localized assessment of
 the degree of habitat overlap across potentially affected species to
@@
 
-Caveats
-
-Note that lack of importance of density dependence timing is conditional
-on age structure, and matters more if you're doing surplus production
-modeling (cite letter)
-
-Mention hyperallometry and distinction to what that study did
-
-Perfect information
-
-Fleet model
-
-**what is the story**
-
--   Unintended conservation consequences of MPAs resulting from
-    technical interactions have been discussed in the literature, but
-    almost exclusively in the context of highly tactical modeling or
-    localed empirics. We conduct one of the first large-scale
-    assessments of the potential of this problem across a range of
-    scenarios
-
--   We find that these unintended consequences can be common outside of
-    "canonical" MPA models *a la* hastings etc.
-
--   But, good news is magnitude of unintended consequences were usually
-    small, and total benefits almost always outweigh individual species
-    losses (again no probability of states though)
-
--   Some design strategies are better at avoiding unintended
-    consequences than others, but sensitive to data quality.
-    Optimization can help, but even with optimization unintended
-    consequences can happen.
-
--   Under the levels of fishing mortality in the WCPO, optimization and
-    design can help reduce tradeoffs between fisheries revenue and
-    unintended consequences than others, but macro-level tradeoffs
-    between biodiversity and fishery revenue are very insensitive to
-    design choices, with a pretty clear "increasing biodiversity through
-    MPAs reduces fishery revenues".
-
-    -   xx This is clearly hyper-sensitive to the one thing we don't
-        iterate over much here: how overfished are the target species.
-        At the moment "they're not". But, this runs fast enough, is it
-        worth running a hypothetical case study where they are just to
-        illustrate how things change?
+Several results from our model may seem counterintuitive based on
+existing literature. We found little effect of the timing and spatial
+structure of density dependence in our model. Other studies have pointed
+out that decisions about whether density dependence is local or global
+can have major implications for fishery outcomes of MPAs (citations).
+The difference here is that our model is fully age structured. As such,
+yields are separated from recruitment, the location of density dependent
+processes in our model, by growth, natural mortality, and adult movment.
+In contrast biomass dynamics models assume that density dependence
+occurs on a net "growth" parameter combining both recruitment and
+biomass growth, and that this growth is instantly available to the
+fishery. As such we suspect that the timing of density dependence plays
+a much bigger role in the expected outcomes of MPAs when biomass
+dynamics models are used as opposed to when it is included as an effect
+on recruitment in an age structured model.
+
+@marshall2021 pointed out that many fisheries models assume that
+fecundity is simply proportional to spawning biomass, when in fact there
+is ample empirical evidence that fecundity follows a hyperallometric
+pattern, i.e. that fecundity increases faster than weight. In that study
+the authors found that the presence of hyperallometric growth
+substantially amplified the predicted benefits of spatiao-temporal
+closures such as MPAs, relative to an assumption of allometric
+fecundity. In contrast, we find that the presence of hyperallometric
+fecundity, even at levels greater than the mean values reported in
+@marshall2021, had little effect on the expected conservation outcomes
+of MPAs (Fig.\@ref(fig:fig-loss-drivers), see SI for additional
+analyses). The key distinction here is that @marshall2021 evaluated the
+effect of MPAs under different assumptions of management, e.g.
+management based on spawning potential ratio values calculated assuming
+allmetric fecundity when in fact the relationship was hyperallometric.
+Our study holds effort constant, and so reflects more the isolated
+effect of hyperallometric on MPA outcomes independent of management
+actions.
+
+@rassweiler2012 and @rassweiler2014 both demonstrated the potential
+importance of optimized MPA planning over more "rule of thumb" MPA
+designs. In contrast, we found little value of optimization, with the
+"Depletion" MPA allocation strategy more or less following the Pareto
+efficiency frontier (Fig.\@ref(fig:fig-cs-totals)). As pointed out in
+@rassweiler2014, the value of optimization (at least if the objective
+function depends on both fishery yields and fish biomass) depends on the
+degree of overfishing present. In our WCPO case study, best available
+evidence is that none of the major tuna stocks that provide the bulk of
+yield, revenues, and profits to the WCPO fleets are overfished or
+experiencing overfishing (citations, need to update with assessments
+since pons is out of date now). Our results then are consistent with
+@rassweiler2014, and evaluations of our model on stocks experiencing
+overfishing will likely find much higher value of optimization.
+
+Value of information and fleet dynamics are two critical areas for
+future study in this realm. Our study assumes perfect information in the
+design of MPAs. We feel this is a useful benchmark for isolating the
+potential impacts of key ecological and economic drivers of MPA effects.
+However it is clearly not a realistic assumption, particularly for
+pelagic ecosystems where species inhabit a complex, dynamic, and hard to
+observe mosaic of habitats. This is likely to be particularly true for
+rarer bycatch species such as sharks that may be the impetus behind many
+pelagic MPA designs. Further research is needed to examine how the types
+of dynamics explored here are affected by varying degrees of information
+quality.
+
+We focused our efforts on modeling the spatial distribution of a fishing
+fleet conditional on a fixed pool of fishing effort. As discussed
+previously, the outcomes of MPAs depend critically on the reaction of
+fishing fleets to new spatial protections. In reality neither perfect
+attrition nor perfect redistribution are likely to be realistic
+scenarios. Instead individual fishing operations will adjust their
+behavior according to their own objective functions. More realistic
+models of total effort could include simple open-access dynamics in the
+manner of @cabral2019, to more complex dynamics reflecting combinations
+of policies such as effort controls, spatial closures, and rights based
+fisheries management.
+
+As demonstrated here there though there is a possibly endless list of
+refinements and realism that could be added to simulation based studies.
+While efforts such as this can help illustrate which drivers may be the
+most important in particular circumstances in many cases as demonstrated
+here the effect of specific variables are likely to be so context
+dependent as to defy easy classification of "important" or "not". We
+feel these models are better viewed as a foundation for empirical
+action. In terms of planning, models such as this can be used as tool
+for answering "what if" rather "what will be". Helping community members
+visualize the potential impacts of alternative states of nature, such as
+debates around life history or habitat distribution, can help set
+expectations and guide data collection. Models such as this can also
+serve as hypotheses for expected effects that can ideally be confronted
+with data as MPAs are actually implemented.
 
@@
 
--   Unintended conservation outcomes of MPAs totally possible outside of
-    "consider a homogenous fish blob" models
-
--   But, usually small, and net conservation benefits outweigh
-    conservation costs
-
--   Should pay particular attention to potential though when species
-    near tipping points are of concern: hard to imagine those not being
-    part of the prioritization but something to consider.
-
--   Other interesting things.
+The bluntness of MPAs as a tool is both their strength and weakness.
+Targeted fisheries management is often focused on the assessment and
+management of individual stocks. While this has worked well in many
+cases [@hilborn2014a; @hilborn2020; @worm2009], it is an unrealistic and perhaps undesirable 
+model to replicate in many parts of the world, where data limitations
+and fishery complexity make this sort of highly targeted management
+untenable. In contrast, MPAs can provide protection to entire portions
+of an ecosystem with minimal need for targeted interventions. As a
+result and as shown by @hastings2017, MPAs can provide an effective
+solution to bycatch problems under the right circumstances. We extend
+the spirit of that analysis here, but expanding the model used to assess
+the impact of MPAs on bycatch under broader range of scenarios. As
+predicted by @hastings2017, as conditions deviate from the stylized
+nature of their model the effects of MPAs on bycatch become more
+nuanced. While we similarly find a large parameters space where MPAs
+provide clear benefits to the conservation status of both target and
+bycatch species, we also identify a smaller subset of enabling
+conditions, such as degrees of habitat overlap, that can lead to MPAs
+producing net reductions in bycatch biomass. As increasing numbers of
+MPA networks are implemented around the world, tools such as this can
+help communities both design and evaluate the performance of MPA
+networks with the goal of better outcomes for marine ecosystems and the
+human communities that depend on them.
 
@@
 :::
+
+# SI
+
+```{r fig-cs-yield-critters, fig.cap = "Simulated trajectories of changes in yields and spawning stock biomass (SSB) for selected species caught in the WCPO.
+Each column represents a different MPA design strategies.
+Optimized refers to an MPA network optimized to maximize total SSB across all species."}
+
+opt_comparison <- opt_obj %>%
+  filter(alpha == max(alpha)) %>%
+  mutate(placement_error = 0,
+         prop_mpa = p_protected / 100,
+         placement_strategy = "Optimized") %>%
+  select(critter, placement_error, delta_biodiv, delta_econ, delta_yield, prop_mpa,placement_strategy)
+
+ experiment_obj %>% 
+   select(colnames(opt_comparison)) %>% 
+  bind_rows(opt_comparison) %>% 
+   mutate(placement_strategy = fct_relevel(placement_strategy,"Optimized")) %>% 
+   left_join(sci_to_com, by = "critter") %>% 
+  filter(placement_error == min(placement_error)) %>% 
+   rename("Change in SSB" = delta_biodiv, "Change in Yield" = delta_yield) %>% 
+  pivot_longer(c("Change in SSB", "Change in Yield"), names_to = "metric", values_to = "value") %>% 
+  ggplot(aes(prop_mpa, value, color = common_name)) + 
+   scale_color_discrete(name = '') +
+   geom_hline(aes(yintercept = 0), linetype = 2)+
+  geom_line(show.legend = TRUE) + 
+   scale_y_continuous(name = '') +
+   scale_x_continuous(name = "MPA Size",labels = scales::percent, guide = guide_axis(n.dodge = 2)) +
+  facet_grid(metric~placement_strategy, scales = "free_y", switch = "y", labeller = labeller(placement_strategy = capitalize)) + 
+   theme(strip.text.x = element_text(size = 8), 
+         legend.position = "top", 
+         legend.text = element_text(size = 8))
+
+```
 

### 2021-10-12 scripts/00_setup.R

@@ -46,11 +46,11 @@ if (!dir.exists(results_path)){
 
 run_experiments <- FALSE
 
-experiment_workers <- 1
+experiment_workers <- 10
 
-run_casestudy <- TRUE
+run_casestudy <- FALSE
 
-optimize_casestudy <- TRUE
+optimize_casestudy <- FALSE
 
 safety_stop <- FALSE
 

### 2021-10-12 scripts/02_run_case_study.R

@@ -557,9 +557,17 @@ bau_econ <-
   pivot_longer(everything(), names_to = "critter",values_to = "bau_econ")
 #  calculate econ component of objective function, currently revenues across all fleets and species
 
+bau_yield <-
+  (map_df(starting_conditions, ~ sum(.x$c_p_fl, na.rm = TRUE))) %>%
+  pivot_longer(everything(), names_to = "critter",values_to = "bau_yield")
+#  calculate yield component of objective function, currently revenues across all fleets and species
+
+
 bau <- bau_econ %>%
   left_join(bau_biodiv, by = "critter") %>%
-  mutate(bau_econ = bau_econ / 1e6)
+  left_join(bau_yield, by = "critter") %>%
+  mutate(bau_econ = bau_econ / 1e6,
+         bau_yield = bau_yield / 1e6)
 
 
 # calculate objective function for experiments
@@ -568,7 +576,8 @@ experiment_obj <- case_study_experiments %>%
   mutate(obj = map(results, "obj")) %>%
   select(-results) %>%
   unnest(cols = obj) %>%
-  mutate(econ = econ / 1e6) %>%
+  mutate(econ = econ / 1e6,
+         yield = yield / 1e6) %>%
   filter(prop_mpa < 1) %>%
   mutate(bycatch = !(str_detect(critter, "thunnus") | str_detect(critter,"katsuwonus")))
 
@@ -576,7 +585,8 @@ experiment_obj <- case_study_experiments %>%
 experiment_obj <- experiment_obj %>%
   left_join(bau, by = "critter") %>%
   mutate(delta_biodiv = biodiv - bau_biodiv,
-         delta_econ = econ - bau_econ)
+         delta_econ = econ - bau_econ,
+         delta_yield = yield - bau_yield)
 
 
 experiment_obj %>%
@@ -609,7 +619,8 @@ opt_obj <- optimized_networks %>%
   mutate(obj = map(network,"outcomes")) %>%
   select(-network) %>%
   unnest(cols = obj) %>%
-  mutate(econ = econ / 1e6) %>%
+  mutate(econ = econ / 1e6,
+         yield = yield / 1e6) %>%
   mutate(bycatch = !(str_detect(critter, "thunnus") | str_detect(critter,"katsuwonus")))
 
 
[…]

### 2021-10-12 README.Rmd
Cpue cs (#5)
* interesting. open access now works

* safety stop, things working but tricky to think through optimization

* testing reverse optimization

* adding in hyperallometry

* revised version focused on bycatch

@@
 ```
+
+# idea for fleet revamp
+
+I think the most straightforward route is just to build a bit more realism in.
+
+1. Burn the model in
+
+2. Calculate costs such that profit per unit effort is 0, where costs are calculated based on f, not effort.
+
+3.
+Redistribute effort, and going forward use open access dynamics.
+So, if you set that up right, there will now be a tradeoff between increased catch from effort concentration but also increased cost from f approaching 1.
+
+Worth a shot
+
+```{r}
+
+q <- 0.1
+
+m <- 0.2
+
+effort <- seq(0,100, by = 1)
+
+f <- effort * q
+
+b <- 100
+
+price <- 10
+
+catch <- f / (f + m) * b * (1 - exp(-(f + m)))
+
+biomass <- b * (exp(-(f + m)))
+
+revenue <- catch * price
+
+cost = effort^2
+
+plot(revenue - cost)
+
+plot(revenue)
+
+f <- 10 * q
+
+r1 <- 2 * (f / (f + m) * b * (1 - exp(-(f+ m))) * price - 10^2)
+
+f <- 20 * q
+
+r2 <- (f / (f + m) * b * (1 - exp(-(f+ m))) * price - 20^2)
+
+r1
+
+r2
+
+```
+ Yeah that checks out. So, the idea is what? Run the population to equilibrium, and then find a cost coefficient such that system-wide profits are zero
+
+```{r}
+
+q <- 0.1
+
+m <- 0.2
+
+effort <-10
+
+f <- effort * q
+
+b <- 100
+
+price <- 10
+
+cost <- 10
+
+max_delta <- 0.25
+
+cost = (f / (f + m) * b * (1 - exp(-(f+ m))) * price) / effort^2
+
+p1 <- (f / (f + m) * b * (1 - exp(-(f+ m))) * price) - cost * effort^2
+
+theta = log((effort * (1 + max_delta)) / effort) / (f / (f + m) * b * (1 - exp(-(f+ m))) * price)
+
+e2 <- effort * exp(theta * (f / (f + m) * b * (1 - exp(-(f+ m))) * price))
+
+e2 / effort - 1
+
+e2 <- effort * exp(theta * (10000))
+
+e2 / effort
+
+# so what if you tuned it so that if you caught all of ssb0 it would go up by 25%
+```
+
+ What about a sigmoid form, so that it's symetric in increase vs decrease?
+
+```{r}
+
+p = seq(-10,10)
+
+x = 1.25 * (1 / (1 + exp(-p)))
+
+plot(p,x)
+
+```

### 2021-10-12 R/compile_experiment_fleet.R

@@ -1,4 +1,5 @@
-compile_experiment_fleet <- function(fauna, state, tune_type = "explt") {
+compile_experiment_fleet <- function(fauna, state, tune_type = "explt",
+                                     effort_cost_exponent = 1.3) {
 
 
   if (all(state$spatial_q == TRUE)) {
@@ -57,7 +58,7 @@ compile_experiment_fleet <- function(fauna, state, tune_type = "explt") {
       ),
       "carcharhinus longimanus" = Metier$new(
         critter = fauna$`carcharhinus longimanus`,
-        price = 1.89,
+        price = 0,
         sel_form = "logistic",
         sel_start = fauna$`carcharhinus longimanus`$length_50_mature *
           sample(c(0.9), 1, replace = TRUE),
@@ -68,9 +69,11 @@ compile_experiment_fleet <- function(fauna, state, tune_type = "explt") {
         spatial_catchability = shark_spatial_q
       )
     ),
-    base_effort = resolution
+    base_effort = resolution^2,
+    cost_per_unit_effort = 10,
+    effort_cost_exponent = effort_cost_exponent,
+    spatial_allocation = state$spatial_allocation[1]
   ))
-
 
   # a <- Sys.time()
 

### 2021-10-12 R/create_experiment_critters.R

@@ -8,6 +8,7 @@ create_experiment_critters <-
            seasonal_movement = FALSE,
            ontogenetic_shift = FALSE,
            f_v_m,
+           hyper = 1,
            rec_form) {
     # sciname <- marlin_inputs$scientific_name[[1]]
     #
@@ -67,6 +68,7 @@ create_experiment_critters <-
       pups = 6,
       # weight_a = ifelse(str_detect(sciname,"carcharhinus"),2,NA),
       seasons = seasons,
+      fec_expo = hyper,
       steepness =  ifelse(is.nan(
         mean(tmp_inputs$steepness, na.rm = TRUE)
       ), 0.8, mean(tmp_inputs$steepness, na.rm = TRUE)),

### 2021-10-12 R/create_fauna.R

@@ -1,0 +1,27 @@
+create_fauna <- function(scientific_name,
+                         seasonal_habitat,
+                         season_blocks,
+                         adult_movement,
+                         adult_movmement_sigma,
+                         rec_form,
+                         seasons,
+                         init_explt,
+                         explt_type = "f",
+                         recruit_habitat = NA){
+
+  tmp <-  create_critter(
+    scientific_name = scientific_name,
+    seasonal_habitat = seasonal_habitat,
+    season_blocks = season_blocks,
+    adult_movement = adult_movement,
+    adult_movement_sigma = adult_movement_sigma,
+    rec_form = rec_form,
+    seasons = seasons,
+    init_explt = init_explt,
+    explt_type = "explt_type",
+    recruit_habitat = recruit_habitat
+  )
+
+
+
+}

### 2021-10-12 R/optimize-mpa.R

@@ -1,288 +1,0 @@
-#' Optimize MPA network
-#'
-#' @param fauna fauna object
-#' @param fleets fleet object
-#' @param alpha the weight given to conservation relative to economics (range 0 to 1)
-#' @param max_prop_mpa max proportion of cells to place in MPA
-#' @param resolution the resolution of the simulated system
-#' @param prop_sampled the proportion of cells to sample in each SIR iteration
-#' @param workers number of workers for parallel process
-#' @param starting_conditions starting conditions of the simmar object
-#' @param objective one of max_ssb to maximize spawning stock biomass or min_loss to prioritize not having any losses
-#'
-#' @return a list with results of MPA optimization
-#' @export
-#'
-optimize_mpa <-
-  function(fauna,
-           fleets,
-           starting_conditions = NA,
-           alpha = 0.33,
-           max_prop_mpa = 1,
-           resolution,
-           prop_sampled = .2,
-           workers = 6,
-           objective = "max_ssb") {
-    # workers <- 6
-    future::plan(future::multisession, workers = workers)
-
-    on.exit(future::plan(future::sequential))
-
-    # fauna <- casestudy$fauna[[1]]
-    #
-    # fleets <- casestudy$fleet[[1]]
-    #
-    # alpha <- 0
-    #
-    # percent_mpa <- 0.3
-    #
-    # prop_sampled <- 0.2
-    #
-    # max_prop_mpa <-  1
-
-    patches <- resolution ^ 2
-
-    samps <- round(prop_sampled * patches)
-
-    mpas <- expand_grid(x = 1:resolution, y = 1:resolution) %>%
-      mutate(patch = 1:nrow(.)) %>%
-      mutate(mpa = FALSE)
-
-    max_patches_protected <- round(patches * max_prop_mpa)
-
-    candidate_patches <- 1:patches
-
-    patch_value <- tibble(patch = 1:patches, obj_value = 0.5)
-
-    results <- vector(mode = "list", length = max_patches_protected)
-
-    mpa_network <-
-      vector(mode = "list", length = max_patches_protected)
-
-    calc_objective_function <-
-      function(candidate_patch, fauna, fleets, mpas,starting_conditions) {
-        tmp_mpas <- mpas
-
-        tmp_mpas$mpa[tmp_mpas$patch %in% candidate_patch] <- TRUE
-
-        sim_mpa <- simmar(
-          fauna = fauna,
-          fleets = fleets,
-          years = years,
[…]

### 2021-10-12 R/run-mpa-experiment.R

@@ -1,164 +1,0 @@
-run_mpa_experiment <-
-  function(placement_strategy = "depletion",
-           starting_conditions,
-           proc_starting_conditions,
-           prop_mpa = 0.3,
-           prop_species = 1,
-           fauna,
-           fleets,
-           placement_error = 0,
-           prop_critters_considered,
-           random_mpas = FALSE,
-           resolution) {
-
-
-    n_mpa <- round(prop_mpa * resolution^2)
-
-    # assign objective score to each cell
-
-    critters_considered <- sample(names(fauna), round(prop_critters_considered * n_distinct(names(fauna))), replace = FALSE)
-
-    if (placement_strategy == "depletion"){
-
-      # place MPAs in proportion to depletion-weighted spawning stock biomass
-      depletion <-
-        (1 - (map_df(starting_conditions[[1]], ~ sum(.x$ssb_p_a) / .x$ssb0))) %>%  # depletion of each species
-        pivot_longer(everything(), names_to = "critter", values_to = "weight")
-
-      priorities <- proc_starting_conditions$fauna %>%
-        left_join(depletion, by = "critter") %>%
-        filter(critter %in% critters_considered) %>%
-        group_by(critter) %>%
-        mutate(patch_weight = ssb / sum(ssb) * weight) %>%
-        group_by(patch) %>%
-        summarise(patch_weight = sum(patch_weight)) %>%
-        ungroup() %>%
-        mutate(patch_weight = patch_weight * rlnorm(nrow(.), 0, placement_error)) %>%
-        arrange(desc(patch_weight))
-
-      # browser()
-      # priorities %>%
-      #   ggplot(aes(patch_weight, patch_weight2)) +
-      #   geom_point()
-
-    } else if (placement_strategy == "rate"){
-
-      # place in proportino to depletion weighted catch relative to total catch. So, cells in which most of the catch comes from really depleted species, higher priority
-
-      depletion <-
-        (1 - (map_df(starting_conditions[[1]], ~ sum(.x$ssb_p_a) / .x$ssb0))) %>%  # depletion of each species
-        pivot_longer(everything(), names_to = "critter", values_to = "weight")
-
-      priorities <- proc_starting_conditions$fauna %>%
-        left_join(depletion, by = "critter") %>%
[…]

### 2021-10-12 R/run_mpa_experiment.R

@@ -1,0 +1,188 @@
+run_mpa_experiment <-
+  function(placement_strategy = "depletion",
+           starting_conditions,
+           proc_starting_conditions,
+           prop_mpa = 0.3,
+           prop_species = 1,
+           fauna,
+           fleets,
+           placement_error = 0,
+           mpa_response = "stay",
+           prop_critters_considered,
+           random_mpas = FALSE,
+           max_delta = 1,
+           resolution) {
+
+
+    fleets <-  purrr::modify_in(fleets, list(2, "mpa_response"), ~ mpa_response)
+
+    n_mpa <- round(prop_mpa * resolution^2)
+
+    # set up open access
+
+    # starting <- starting_conditions[[length(starting_conditions)]]
+
+
+    critters_considered <- sample(names(fauna), round(prop_critters_considered * n_distinct(names(fauna))), replace = FALSE)
+
+    if (placement_strategy == "depletion"){
+
+      # place MPAs in proportion to depletion-weighted spawning stock biomass
+      depletion <-
+        (1 - (map_df(starting_conditions, ~ sum(.x$ssb_p_a) / .x$ssb0))) %>%  # depletion of each species
+        pivot_longer(everything(), names_to = "critter", values_to = "weight")
+
+      priorities <- proc_starting_conditions$fauna %>%
+        left_join(depletion, by = "critter") %>%
+        filter(critter %in% critters_considered) %>%
+        group_by(critter) %>%
+        mutate(patch_weight = ssb / sum(ssb) * weight) %>%
+        group_by(patch) %>%
+        summarise(patch_weight = sum(patch_weight)) %>%
+        ungroup() %>%
+        mutate(patch_weight = patch_weight * rlnorm(nrow(.), 0, placement_error)) %>%
+        arrange(desc(patch_weight))
+
+      # browser()
+      # priorities %>%
+      #   ggplot(aes(patch_weight, patch_weight2)) +
+      #   geom_point()
+
+    } else if (placement_strategy == "rate"){
+
+      # place in proportino to depletion weighted catch relative to total catch. So, cells in which most of the catch comes from really depleted species, higher priority
+
+      depletion <-
+        (1 - (map_df(starting_conditions, ~ sum(.x$ssb_p_a) / .x$ssb0))) %>%  # depletion of each species
[…]

### 2021-10-12 scripts/00_setup.R

@@ -34,7 +34,7 @@ foos <- list.files(here("R"))
 
 walk(foos, ~ source(here("R", .x)))
 
-results_name <- "tester"
+results_name <- "v1.1"
 
 results_path <- here("results", results_name)
 
@@ -44,9 +44,9 @@ if (!dir.exists(results_path)){
   dir.create(file.path(results_path,"sims"))
 }
 
-run_experiments <- TRUE
+run_experiments <- FALSE
 
-experiment_workers <- 4
+experiment_workers <- 10
 
 run_casestudy <- FALSE
 
@@ -64,6 +64,8 @@ seasons <- 2
 
 time_step <- 1 / seasons
 
+max_delta <- 1
+
 workers <- round(parallel::detectCores() - 8)
 
 steps <- years * seasons
@@ -72,7 +74,6 @@ time_steps <- seq(0,years - 1, by = time_step)
 
 theme_set(marlin::theme_marlin())
 
-tune_type <- "v1.0"
 
 # load data ---------------------------------------------------------------
 

### 2021-12-02 documents/03_marlin_paper.Rmd
back to main

@@
   bookdown::word_document2:
+    number_sections: false
     reference_docx: template.docx
   bookdown::html_document2: default
-  bookdown::pdf_document2: default
+  bookdown::pdf_document2:
+    number_sections: false
 header-includes:
@@
   run_name: ["v1.1"]
+csl: proceedings-of-the-royal-society-b
 ---
@@
 
-walk(rdatas, ~load(file.path(results_path,.x), envir = .GlobalEnv))
+walk(rdatas, ~load(file.path(results_path,.x), envir = .GlobalEnv)) # load results into global environment
 
@@
 <!-- accidentally fish harder -->
+
+# Abstract
+
+Marine protected areas (MPAs) are a growing part of the marine
+conservation toolbox.
+One of the goals of global MPA expansion is to protect and restore biodiversity through conservation of threatened species often caught incidentally in fisheries, such as marine mammals, turtles, seabirds, and sharks.
+However, the ability of MPAs to provide population-level conservation benefits to bycatch species depends in part on the ways in which fishing fleets react to MPA placement, and the degree to which habitat and movement dynamics of different species within an ecosystem overlap.
+In this paper we assess the impacts of MPAs on target and bycatch species and fisheries when faced with fleet redistribution and heterogeneity in life history and habitat use using a novel bio-economic model.
+We show that while in the majority (xx%) of simulated cases MPAs provided net conservation benefits to all species, in xx% of simulations MPAs resulted in net losses in biomass for 1 or more species.
+While factors interact in complex and nonlinear ways, placement strategy, adult and larval movement, and degree of habitat heterogeneity were among the most important factors in determining whether MPAs produced a net conservation loss.
+We then used our model to assess the potential impacts of MPAs in the Western and Central Pacific Ocean on tuna, billfish, and shark populations and fisheries.
+MPAs produced net conservation gains for all species in nearly all simulations, but did so at the expense of both yields and profits, even when MPA design was optimized.
+Our results demonstrate the importance of considering fleet behavior, movement dynamics, and habitat heterogeneity in assessing the potential impacts of MPAs in a multi-species and multi-fleet context.
+XX some punchier closing sentence XX.
+
+# To Do
+
+- Produce results that design MPAs around only a subset of species, to see how sensitive results are to say not including some species in the analysis
+
+- Expand more on the WCPO case study and make the paper a bit more *about* that: make clearer how the model could be used in a real-world setting
+
+- Link better to Will White and company's work on MPA modeling etc
 
@@
 
-Specifically, the potential benefits of MPAs as a bycatch management are
+The potential benefits of MPAs as a bycatch management are
 complicated by two key issues: effort displacement and technical
@@
 effort, then their direct effects on target and bycatch species can only
-be positive. If however MPAs displace fishing effort evaluating their
+be positive. If however MPAs displace fishing effort, evaluating their
 potential impacts becomes more complicated. The net conservation impact
@@
 halibut (*Hippoglossus stenolepis*) by the displaced bottom-trawl
-fishery. Policy leakage is not a unique feature to MPAs, e.g. XX, but
+fishery. Policy leakage is not a unique feature to MPAs, e.g. XX include non-marine examples XX, but
 the potential impacts of MPAs on target and bycatch species taking into
-account the potential for these leakage effects has received relativelly
+account the potential for these leakage effects has received relatively
 little generalized attention in the literature [see @abbott2012;
@@
 
-This study addresses this gap using a novel bio-economic simulation
+We address this gap by providing a strategic evaluation of the sensitivity of bycatch
+populations and targeted fisheries to key unknowns and design choices in
+MPA planning. We used a novel bio-economic simulation
 model that allows for efficient evaluation of the effects of MPAs on
@@
 dynamics, critically accounting for the potential impacts of effort
-displacement and technical interactions. We use this model first to
+displacement and technical interactions.  We used this model first to
 explore general conditions that lead to positive and negative outcomes
@@
 between maximizing bycatch conservation and minimizing economic impacts.
+
 We chose the WCPO as a case study as the modern science of MPAs was
@@
 large MPAs are increasingly being implemented and considered across
-pelagic ecosystems. These pelagic systems are characterized by highly
+pelagic ecosystems [@hilborn2021c]. These pelagic systems are characterized by highly
 mobile species and fishing fleets using a diverse mix of static and
@@
 expanses of blue-water ecosystems, make pelagic systems such as the WCPO
-an ideal case study in bycatch management through MPAs. The goal of this
-paper is to provide a strategic evaluation of the sensitivity of bycatch
-populations and targeted fisheries to key unknowns and design choices in
-MPA planning.
+an ideal case study in bycatch management through MPAs.
 
@@
 Along with distance, net movement is defined by habitat quality. Habitat
-quality defines the quailty for adults or larvae of a given patch. The
+quality defines the quality for adults or larvae of a given patch. The
 net probability of movement from patch *i* to patch *j* is defined then
@@
 iteration, selecting the cell whose removal produces the highest total
-objective function as the cell to remove from the network, and updateing
+objective function as the cell to remove from the network, and updating
 the marginal value of the sampled-but-not-selected cells. This in
@@
 For the generalized experiments we only include one fishing fleet. We
-asign a price per KG of \$XX for tuna, intended to reflect a low value
+assign a price per KG of \$XX for tuna, intended to reflect a low value
 high volume fishery, \$XX for marlin for a higher value low volume
@@
 assign a selectivity of 60% of length at maturity for skipjack to allow
-for some overfishing when F/M is high enough, 70% for mariln, and 90%
+for some overfishing when F/M is high enough, 70% for marlin, and 90%
 for shark.
 
-We ran `r nrow(state_experiments)` state experiments, varing the state
+We ran `r nrow(state_experiments)` state experiments, varying the state
 of nature experienced by a simulation. The candidate states include
@@
 | $\sigma_{habitat}$    |                       | Standard deviation around centroid              |
-| Spatial catchability  |                       | Catchability varies in sapce                    |
+| Spatial catchability  |                       | Catchability varies in space                    |
 | Spatial allocation    |                       | Fleet allocation strategy                       |
@@
 and $\sigma_h$. All species start with the assumption that the centroid
-of their preferred habitat occurs in the center-most patch of teh two
+of their preferred habitat occurs in the center-most patch of the two
 dimensional system simulated by marlin. For our simulations, we then
@@
 and *center* is the coordinate of the central cell. In this way, the
-larger $\sigma_c$ grows, the more variation there is betwen the centroid
+larger $\sigma_c$ grows, the more variation there is between the centroid
 of the preferred habitat for each species.
@@
 We then control the spread of a creatures preferred habitat around that
-centroid per a Gaussian dispersal kernal in the manner
+centroid per a Gaussian dispersal kernel in the manner
 
@@
 
-Where $d$ is the euclidian distance from the $centroid_s$ to patch $p$
+Where $d$ is the euclidean distance from the $centroid_s$ to patch $p$
 and $\sigma_h$ controls the dispersal of the habitat around that range.
@@
 first is that we assume that effort is displaced, not removed, by the
-creation of MPAs. This is a widely debated question around MPAs, and the
+creation of MPAs. This is a widely debated question around MPAs [@hilborn2021c], and the
 real impacts of MPAs on fishing effort will clearly be setting
@@
 
-<!-- what we found was a series of neccesary but not sufficient conditions  -->
+<!-- what we found was a series of necessary but not sufficient conditions  -->
 
@@
 `r n_distinct(total_experiment_results$placement_id) * n_distinct(total_experiment_results$state_id)`
-experiments run, the results were highly variable. Across our
-simulations, gains were much more common than losses, with losses of any
+experiments run gains were much more common than losses, with losses of any
 magnitude occurring in `r scales::percent(mean(critter_results$loss))`
@@
 conservation losses must be zero when MPA size is 100% (which this graph
-does not quite reach), with peak losses in our simulations occurring
+does not quite reach). Peak losses in our simulations occurring
 between MPA sizes of XX and XX. Life history plays a clear role our
@@
 this type of regression analysis. To address this, we ran a random
-forest using the `ranger` package (citation) in R (citation) and
+forest using the `ranger` package [@wright2017] in R [@rcoreteam2021] and
 extracted the impurity corrected importance scores for each of the
@@
 becomes increasingly complicated as the systems being studied deviate
-from assumptions of homogenously distributed species with perfectly
+from assumptions of homogeneously distributed species with perfectly
 selective fleets. In this paper we used a new bio-economic model to
@@
 
-The MPA literature is highly dominated by modeling studies (citation).
+The MPA literature is highly dominated by modeling studies [@oleary2016].
 What then is the added value of the model and results presented here?
@@
 interactions. One could of course continue to add features to the model
-until it reaches the complexity of an end-to-end model such as Atlantis
-(citation). While such models have tremendous value, our goal here is to
+until it reaches the complexity of an end-to-end model such as Atlantis [@fulton2011a]. While such models have tremendous value, our goal here is to
 present a model that can help provide closer to tactical advice for
@@
 yields are separated from recruitment, the location of density dependent
-processes in our model, by growth, natural mortality, and adult movment.
+processes in our model, by growth, natural mortality, fishery selectivity, and adult movement.
 In contrast biomass dynamics models assume that density dependence
@@
 the authors found that the presence of hyperallometric growth
-substantially amplified the predicted benefits of spatiao-temporal
+substantially amplified the predicted benefits of spatio-temporal
 closures such as MPAs, relative to an assumption of allometric
@@
 management based on spawning potential ratio values calculated assuming
-allmetric fecundity when in fact the relationship was hyperallometric.
+allometric fecundity when in fact the relationship was hyperallometric.
 Our study holds effort constant, and so reflects more the isolated
@@
 overfishing will likely find much higher value of optimization.
+As a demonstration of this, we ran a sensitivity analyses dramatically increasing the fishing mortality rates of all the tuna species in our WCPO case study.
+We found that indeed the value of optimization on profits increased by XX% (calculate area of the ribbon between the two curves as a measure of value of optimization).
 
 Value of information and fleet dynamics are two critical areas for
-future study in this realm. Our study assumes perfect information in the
-design of MPAs. We feel this is a useful benchmark for isolating the
+future study. Our study assumes perfect information in the
+design of MPAs. This is a useful benchmark for isolating the
 potential impacts of key ecological and economic drivers of MPA effects.
@@
 We focused our efforts on modeling the spatial distribution of a fishing
-fleet conditional on a fixed pool of fishing effort. As discussed
-previously, the outcomes of MPAs depend critically on the reaction of
+fleet conditional on a fixed pool of fishing effort. The outcomes of MPAs depend critically on the reaction of
 fishing fleets to new spatial protections. In reality neither perfect
@@
 
-As demonstrated here there though there is a possibly endless list of
+There is a possibly endless list of
 refinements and realism that could be added to simulation based studies.
@@
 here the effect of specific variables are likely to be so context
-dependent as to defy easy classification of "important" or "not". We
-feel these models are better viewed as a foundation for empirical
+dependent as to defy easy classification of "important" or "not". These models are better viewed as a foundation for empirical
 action. In terms of planning, models such as this can be used as tool
@@
 
+\renewcommand{\thesection}{Appendix S\arabic{section}}
+
+\renewcommand{\thefigure}{S\arabic{figure}}
+
+\setcounter{figure}{0}
+
+\renewcommand{\thetable}{S\arabic{table}}
+
+\setcounter{table}{0}
+
+\renewcommand{\theequation}{S\arabic{equation}}
+
+\setcounter{equation}{0}
+
+```{r fig-cs-yield-totals, fig.cap = "Efficiency frontiers for changes in total yield and total bycatch spawning stock biomass (SSB).
+Color gradient line shows optimized pareto frontier, with color showing weighting of conservation (total spawning stock biomass) in the objective function.
+Red points show outcomes of 'rule of thumb' MPA design strategies.
+"}
+total_opt_obj %>%
+  group_by(alpha) %>%
+  filter(obj_fun == max(obj_fun)) %>%
+  ungroup() %>%
+  ggplot() +
+  geom_line(aes(delta_biodiv, delta_yield, color = alpha), size = 2) +
+  geom_point(
+    data = total_experiment_obj,
+    aes(delta_biodiv, delta_yield),
+    size = 2,
+    alpha = 0.1,
+    fill = "red",
+    shape = 21
+  ) +
+  geom_vline(aes(xintercept = 0), linetype = 2) +
+  geom_hline(aes(yintercept = 0), linetype = 2) +
+  geom_abline(intercept = 0, slope = -1) +
+  facet_wrap( ~ placement_strategy) +
+  scale_color_viridis_c(name = "Conservation Weighting in Optimization") +
+  scale_x_continuous(name = "Change in Total Biodiversity Relative to BAU ") +
+  scale_y_continuous(name = "Change in Total Yield Relative to BAU ") +
+  theme(axis.text.x = element_text(size = 8),
+        legend.position = "top")
+
+```
+
 ```{r fig-cs-yield-critters, fig.cap = "Simulated trajectories of changes in yields and spawning stock biomass (SSB) for selected species caught in the WCPO.
@@
 
+```
+
+```{r highf_prof_frontier,fig.cap = "Efficiency frontiers for changes in total profits and total bycatch spawning stock biomass (SSB) when tunas are heavily overfished.
+Color gradient line shows optimized pareto frontier, with color showing weighting of conservation (total spawning stock biomass) in the objective function.
+Red points show outcomes of 'rule of thumb' MPA design strategies."}
+hf_cs_fig_10 <- total_opt_obj_highf %>%
+  group_by(alpha) %>%
+  filter(obj_fun == max(obj_fun)) %>%
+  ungroup() %>%
+  ggplot() +
+  geom_line(aes(delta_biodiv, delta_econ, color = alpha), size = 2) +
+  geom_point(
+    data = total_experiment_obj,
+    aes(delta_biodiv, delta_econ),
+    size = 2,
+    alpha = 0.1,
+    fill = "red",
+    shape = 21
+  ) +
+  geom_vline(aes(xintercept = 0), linetype = 2) +
+  geom_hline(aes(yintercept = 0), linetype = 2) +
+  geom_abline(intercept = 0, slope = -1) +
+  facet_wrap( ~ placement_strategy) +
+  scale_color_viridis_c(name = "Conservation Weighting in Optimization") +
+  scale_x_continuous(name = "Change in Total Biodiversity Relative to BAU ") +
+  scale_y_continuous(name = "Change in Total Profits Relative to BAU ") +
+  theme(axis.text.x = element_text(size = 8),
+        legend.position = "top")
+
+hf_cs_fig_10
+```
+
+```{r highf_yield_frontier, fig.cap = "Efficiency frontiers for changes in total yield and total bycatch spawning stock biomass (SSB) when tunas are heavily overfished.
+Color gradient line shows optimized pareto frontier, with color showing weighting of conservation (total spawning stock biomass) in the objective function.
+Red points show outcomes of 'rule of thumb' MPA design strategies.
+"}
+hf_cs_fig_10_yield <- total_opt_obj_highf %>%
+  group_by(alpha) %>%
+  filter(obj_fun == max(obj_fun)) %>%
+  ungroup() %>%
+  ggplot() +
+  geom_line(aes(delta_biodiv, delta_yield, color = alpha), size = 2) +
+  geom_point(
+    data = total_experiment_obj,
+    aes(delta_biodiv, delta_yield),
+    size = 2,
+    alpha = 0.1,
+    fill = "red",
+    shape = 21
+  ) +
+  geom_vline(aes(xintercept = 0), linetype = 2) +
+  geom_hline(aes(yintercept = 0), linetype = 2) +
+  geom_abline(intercept = 0, slope = -1) +
+  facet_wrap( ~ placement_strategy) +
+  scale_color_viridis_c(name = "Conservation Weighting in Optimization") +
+  scale_x_continuous(name = "Change in Total Biodiversity Relative to BAU ") +
+  scale_y_continuous(name = "Change in Total Yield Relative to BAU ") +
+  theme(axis.text.x = element_text(size = 8),
+        legend.position = "top")
+hf_cs_fig_10_yield
 ```
@@
 
-Contrast to coastal MPA theoretical work - @hastings2017 - @ovando2021 -
-@white2010 - not quite sure what to do about points raised in
+Contrast to coastal MPA theoretical work - @hastings2017 - @ovando2021a -
+@white2010c - not quite sure what to do about points raised in
 @fulton2011a - to what extent has this been done...

### 2021-12-02 R/run_mpa_experiment.R

@@ -3,12 +3,11 @@ run_mpa_experiment <-
            starting_conditions,
            proc_starting_conditions,
            prop_mpa = 0.3,
-           prop_species = 1,
            fauna,
            fleets,
            placement_error = 0,
            mpa_response = "stay",
-           prop_critters_considered,
+           critters_considered = NA,
            random_mpas = FALSE,
            max_delta = 1,
            resolution) {
@@ -22,8 +21,11 @@ run_mpa_experiment <-
 
     # starting <- starting_conditions[[length(starting_conditions)]]
 
+    if (is.na(critters_considered)){
+      n_critters_considered <-  length(fauna)
+    }
 
-    critters_considered <- sample(names(fauna), round(prop_critters_considered * n_distinct(names(fauna))), replace = FALSE)
+    critters_considered <- sample(names(fauna), n_critters_considered, replace = FALSE)
 
     if (placement_strategy == "depletion"){
 

### 2021-12-02 scripts/00_setup.R

@@ -15,6 +15,8 @@ library(googledrive)
 
 library(gamm4)
 
+library(rgdal)
+
 library(sf)
 
 library(mvtnorm)
@@ -46,7 +48,7 @@ if (!dir.exists(results_path)){
 
 run_experiments <- FALSE
 
-experiment_workers <- 10
+experiment_workers <- 8
 
 run_casestudy <- FALSE
 

### 2021-12-02 scripts/01_run_experiments.R

@@ -272,6 +272,7 @@ if (run_experiments) {
   state_experiments <- state_experiments %>%
     select(-tmp)
 
+  stop()
 
   placement_experiments <- expand_grid(
     placement_strategy = c(
@@ -282,8 +283,8 @@ if (run_experiments) {
       "area"
     ),
     prop_mpa = seq(0, 1, by = 0.05),
-    prop_critters_considered = 1,
-    placement_error = c(0,.2)
+    critters_considered = seq(1,n_distinct(names(state_experiments$fauna[[1]])), by = 1),
+    placement_error = c(0)
   ) %>%
     group_by_at(colnames(.)[!colnames(.) %in% c("temp", "prop_mpa")]) %>%
     nest() %>%
@@ -420,7 +421,7 @@ if (run_experiments) {
           run_mpa_experiment,
           placement_strategy = placement_experiments$placement_strategy[p],
           prop_mpa = placement_experiments$prop_mpa[p],
-          prop_critters_considered = placement_experiments$prop_critters_considered[p],
+          critters_considered = placement_experiments$critters_considered[p],
           placement_error = placement_experiments$placement_error[p],
           resolution = resolution,
           .options = furrr_options(seed = 42),
@@ -535,7 +536,27 @@ calc_sigma_habitat <- function(tmp_state){
     as.matrix()
 
 
+  # habitat <- tmp_state %>%
+  #   select(scientific_name, contains("sigma_"), habitat) %>%
+  #   unnest(cols = habitat) %>%
+  #   mutate(
+  #     scientific_name = fct_recode(
+  #       scientific_name,
+  #       "Shark" = "carcharhinus longimanus",
+  #       "Tuna" = "katsuwonus pelamis",
+  #       "Marlin" = "kajikia audax"
+  #     )
+  #   ) %>%
+  #   group_by(x,y) %>%
+  #   summarise(habprod = prod(habitat))
+  #
+  # browser()
+
+  # sigma_habitat <- sum(sqrt(habitat$habprod))
+
   sigma_habitat <- sd(sqrt(habitat))
+
+
 
 }
 
@@ -577,7 +598,7 @@ total_results <- results %>%
   group_by(
     placement_id,
     placement_strategy,
-    prop_critters_considered,
+    critters_considered,
     placement_error,
     state_id,
     prop_mpa
@@ -605,7 +626,7 @@ total_results %>%
 # total_results %>%
 #   ggplot(
 #     aes(
[…]

### 2021-12-02 scripts/02_run_case_study.R

@@ -275,7 +275,7 @@ if (run_casestudy == TRUE){
     expand_grid(
       placement_strategy = c("depletion", "rate", "avoid_fishing", "target_fishing", "area"),
       prop_mpa = seq(0,1, by = 0.01),
-      prop_critters_considered = 1,
+      critters_considered = length(fauna),
       placement_error = c(0),
       mpa_response = c("stay")
     )
@@ -285,7 +285,7 @@ if (run_casestudy == TRUE){
     ungroup() %>%
     mutate(results = future_pmap(list(placement_strategy = placement_strategy,
                                prop_mpa = prop_mpa,
-                               prop_critters_considered = prop_critters_considered,
+                               critters_considered = critters_considered,
                                placement_error = placement_error,
                                mpa_response = mpa_response),
                           run_mpa_experiment,
@@ -708,7 +708,7 @@ cs_fig_5 <- experiment_obj %>%
 total_experiment_obj <- experiment_obj %>%
   group_by(placement_strategy,
            prop_mpa,
-           prop_critters_considered,
+           critters_considered,
            placement_error) %>%
   summarise(
     loss = mean(biodiv <= bau_biodiv),
@@ -833,6 +833,32 @@ cs_fig_10 <- total_opt_obj %>%
   scale_color_viridis_c(name = "Conservation Weighting in Optimization") +
   scale_x_continuous(name = "Change in Total Biodiversity Relative to BAU ") +
   scale_y_continuous(name = "Change in Total Profits Relative to BAU ") +
+  theme(axis.text.x = element_text(size = 8),
+        legend.position = "top")
+
+
+
+cs_fig_10_yield <- total_opt_obj %>%
+  group_by(alpha) %>%
+  filter(obj_fun == max(obj_fun)) %>%
+  ungroup() %>%
+  ggplot() +
+  geom_line(aes(delta_biodiv, delta_yield, color = alpha), size = 2) +
+  geom_point(
+    data = total_experiment_obj,
+    aes(delta_biodiv, delta_yield),
+    size = 2,
+    alpha = 0.1,
+    fill = "red",
+    shape = 21
+  ) +
+  geom_vline(aes(xintercept = 0), linetype = 2) +
+  geom_hline(aes(yintercept = 0), linetype = 2) +
+  geom_abline(intercept = 0, slope = -1) +
+  facet_wrap( ~ placement_strategy) +
+  scale_color_viridis_c(name = "Conservation Weighting in Optimization") +
[…]

### 2022-08-17 documents/03_marlin_paper.Rmd
haev functional results with DAT movement

@@
 output: 
+  bookdown::pdf_document2:
+    number_sections: false
+  bookdown::html_document2: default
   bookdown::word_document2:
@@
     reference_docx: template.docx
-  bookdown::html_document2: default
-  bookdown::pdf_document2:
-    number_sections: false
 header-includes:
@@
   mutate(common_name = stringr::str_to_title(common_name))
-
-capitalize <- function(string) {
-  substr(string, 1, 1) <- toupper(substr(string, 1, 1))
-  string
-}
 
@@
 - Link better to Will White and company's work on MPA modeling etc
+
+- Include partial dependency plots for variable importance plot
 
@@
     x = .1,
-    y = max(worst_result$biodiv - worst_result$bau_biodiv)
+    y = max(worst_result$biodiv - worst_result$bau_biodiv) * .9
   ),
@@
             show.legend = FALSE, size = 1.25) +
-  scale_x_continuous(labels = scales::percent, name = "MPA Size") +
+  scale_x_continuous(labels = scales::label_percent(accuracy = 1), name = "MPA Size") +
   scale_y_continuous(
@@
     x = .1,
-    y = max(best_result$biodiv - best_result$bau_biodiv)
+    y = max(best_result$biodiv - best_result$bau_biodiv) * .9
   ),
@@
   size = 4) +
-  scale_x_continuous(labels = scales::percent, name = "") +
+  scale_x_continuous(labels = scales::label_percent(accuracy = 1), name = "") +
   scale_y_continuous("Change in SSB/SSB0") +

### 2022-08-17 R/create_critters.R

@@ -54,11 +54,11 @@ create_critters <-
 
     critter <- marlin::create_critter(
       scientific_name = sciname,
-      seasonal_habitat = hab,
+      base_habitat = hab,
       recruit_habitat = recruit_habitat,
       adult_movement = 0,
-      adult_movement_sigma = 4,
-      recruit_movement_sigma = 2,
+      adult_diffusion = 4,
+      recruit_diffusion = 2,
       rec_form = rec_form,
       fec_form = ifelse(str_detect(sciname,"carcharhinus"),"pups","weight"),
       pups = 6,
@@ -74,17 +74,5 @@ create_critters <-
       ), 1e4, mean(tmp_inputs$ssb0 * 1000, na.rm = TRUE))
     )
 
-
-    # # later sub in a lookup table for this
-    # trait_frame <- tibble(
-    #   seasonal_habitat = list(habitat$data[[which(habitat$species_sciname == sciname)]]),
-    #   adult_movement = 1,
-    #   adult_movement_sigma = runif(1,min = 1, max = 30),
-    #   seasons = seasons,
-    #   init_explt = sample(c(.05,.1,.2), 1, replace = TRUE),
-    #   rec_form = sample(c(0,1,2,3),1, replace = TRUE),
-    #   fec_form = c("power"),
-    #   weight_a = NA
-    # )
 
   }

### 2022-08-17 scripts/00_setup.R

@@ -36,7 +36,7 @@ foos <- list.files(here("R"))
 
 walk(foos, ~ source(here("R", .x)))
 
-results_name <- "v1.1"
+results_name <- "v1.2"
 
 results_path <- here("results", results_name)
 
@@ -46,7 +46,7 @@ if (!dir.exists(results_path)){
   dir.create(file.path(results_path,"sims"))
 }
 
-run_experiments <- FALSE
+run_experiments <- TRUE
 
 experiment_workers <- 8
 
@@ -54,7 +54,7 @@ run_casestudy <- FALSE
 
 optimize_casestudy <- FALSE
 
-safety_stop <- FALSE
+safety_stop <- TRUE
 
 draws <- 20
 

### 2022-08-17 scripts/01_run_experiments.R

@@ -3,7 +3,7 @@ source(file.path("scripts", "00_setup.R"))
 
 # generate state experiments. This is a somewhat tricky process where the actual generated values are created for many variables in create_experiment_critters
 
-n_states <- 600
+n_states <- 5
 
 if (run_experiments) {
   state_experiments <-
@@ -35,13 +35,6 @@ if (run_experiments) {
              base_centroid = c(resolution / 2, resolution / 2) ,
              critters,
              resolution) {
-      # base_centroid <- c(10,10)
-      # sigma_hab = .2
-
-      # sigma_block <- 0.1
-
-      # bycatch_corr <- -1
-
 
       base_layer <-
         tidyr::expand_grid(x = 1:resolution, y = 1:resolution)
@@ -70,31 +63,13 @@ if (run_experiments) {
           mutate(habitat = dnorm(distance, 0, sigma_hab)) %>%
           select(x, y, habitat)
 
-
-
-
         tmp$habitat <- tmp$habitat - min(tmp$habitat)
 
-        # tmp %>%
-        #   ggplot(aes(x, y, fill = habitat)) +
-        #   geom_tile() +
-        #   scale_fill_viridis_c()
-        #
 
         critters$habitat[[i]] <-  tmp
 
 
       }
-
-
-      # (critters$habitat[[1]] %>%
-      #     ggplot(aes(x,y,fill = habitat))+
-      #     geom_tile() +
-      #     scale_fill_viridis_c()) +
-      #   (critters$habitat[[10]] %>%
-      #      ggplot(aes(x,y,fill = habitat))+
-      #      geom_tile() +
-      #      scale_fill_viridis_c())
 
       return(critters)
     }
@@ -114,12 +89,12 @@ if (run_experiments) {
     mutate(
       seasonal_movement = sample(c(TRUE,FALSE), nrow(.), replace = TRUE),
       f_v_m = sample(c(0.4,0.8,1.6), nrow(.), replace = TRUE),
-      adult_movement_sigma = runif(
+      adult_diffusion = runif(
         nrow(.),
         min = 0.01 * resolution,
         max = .5 * resolution
       ),
-      recruit_movement_sigma = runif(
+      recruit_diffusion = runif(
         nrow(.),
         min = 0.01 * resolution,
         max = .5 * resolution
[…]

### 2022-10-06 marlin-paper.qmd
revamp with coral experiment

--- 2021-10-12 documents/03_marlin_paper.Rmd
@@
 ---
-title: "Opportunities and precautions for the use of MPAs in bycatch reduction"
+title: "Efficient Simulation of Economic and Ecological Dynamics for Marine Spatial Planning"
 author:
@@
 date: "`r Sys.Date()`"
-output: 
-  bookdown::word_document2:
-    reference_docx: template.docx
-  bookdown::html_document2: default
-  bookdown::pdf_document2: default
-header-includes:
-  - \usepackage{setspace}\doublespacing
-  - \usepackage{lineno}\linenumbers
+format:
+  pdf:
+    toc: true
+  html: default
+  docx: default
+execute:
+  echo: false
+  message: false
+  warning: false
 bibliography: references.bib
+params:
   params:
@@
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, include = TRUE)
+```{r setup}
+#| echo: false
+#| message: false
 
@@
 
-capitalize <- function(string) {
-  substr(string, 1, 1) <- toupper(substr(string, 1, 1))
-  string
-}
-
-```
-
-```{r load-results}
-
 results_files <- list.files(file.path(results_path))
@@
 
-walk(rdatas, ~load(file.path(results_path,.x), envir = .GlobalEnv))
-
-```
-
-<!-- \*\* something to point out: increased catches are a paradox, either -->
-<!-- come from rebuilding stock or concentrated fishing, but trajectory can -->
-<!-- tell you something. not possible for MPAs to increase catches without -->
-<!-- decreasing catches first if they are rebuilding? \*\* -->
-
-<!-- Seasonal movement is kind of interesting, basically just dilutes the -->
-<!-- extremes -->
-
-<!-- less resilient species are both more in need of protection and easier to -->
-<!-- accidentally fish harder -->
-
-# Introduction
-
-<!-- MAKE IT MORE ABOUT BYCATCH-->
-
-<!-- Empirically, how do you differentiate increase in catch from fishing harder vs increases in catch from rebuilding: rebuilding has to have a decrease at first -->
-
-Marine protected areas (MPAs) are a growing part of the marine
-conservation toolbox, with 5.9% of the world's oceans currently in
-implemented MPAs [@grorud-colvert2021], and bodies such as the IUCN
-calling for 30% protection by 2030. One of the potential benefits of
-MPAs is their ability to provide protection to broad segments of an
-ecosystem at once, rather than requiring targeted management based on
-individual species that may be unfeasible in many parts of the world
-lacking in data or fisheries management capacity. Their ability to
-provide protection across a range of species with one policy has led them to be suggested as a tool to reduce problems around weak stock or "bycatch"
-species [@hastings2017]. We use the term "bycatch" here to generally refer to the
-capture of species within a fishery primarily targeting a different set
-of species.
-
-While the effect of MPAs on fisheries yields is a continuously debated
-topic, the conservation impacts of MPAs are seemingly simpler, albeit
-challenging to detect in many cases [@ovando2021a], due in part perhaps
-to the perceived obvious nature of their benefits: if you prevent
-fishing of a species within an area, there will be more of that species
-in that area, and so the conservation benefits of MPAs will increase
-with MPA size. @hastings2017 clearly illustrates this idea. Using a two-species two-patch analytical model, they derive conditions
-under which MPAs can solve the "weak-stock" problem in which a less
-productive species is caught as bycatch by a fleet targeting a more
-valuable and productive stock.
-
-However, as @hastings2017 point out, their results depend on a very
-specific set of assumptions, namely that 1) adults are perfectly
-sedentary 2) the target and bycatch species overlap perfectly in terms
-of habitat and 3) recruits are perfectly mixed throughout the system. In
-this paper We build on @hastings2017 to consider the role of MPAs in
-reducing bycatch while managing tradeoffs with fisheries under a more
-resolved model.
-
-Specifically, the potential benefits of MPAs as a bycatch management are
-complicated by two key issues: effort displacement and technical
-interactions. One of the primary questions around MPAs is to what extent
-do MPAs remove versus displace fishing effort. If MPAs remove fishing
-effort, then their direct effects on target and bycatch species can only
-be positive. If however MPAs displace fishing effort evaluating their
-potential impacts becomes more complicated. The net conservation impact
-of MPAs becomes a balancing act between the effects of reduced fishing
-inside protected borders and increased fishing intensity outside MPAs.
-The outcomes of this displacement can be further complicated by
-technical interactions, i.e. the degree to which fishing effort for one
-species affects another, particularly when species have markedly
-different habitat distributions. For example, @abbott2012 demonstrated
-that an MPA designed to protect red king crab (*Paralithodes
-camtschaticus*) caused an unintended increase in bycatch of Pacific
-halibut (*Hippoglossus stenolepis*) by the displaced bottom-trawl
-fishery. Policy leakage is not a unique feature to MPAs, e.g. XX, but
-the potential impacts of MPAs on target and bycatch species taking into
-account the potential for these leakage effects has received relativelly
-little generalized attention in the literature [see @abbott2012;
-@savina2008; @fulton2011a; @baum2003 for notable exceptions].
-
-This study addresses this gap using a novel bio-economic simulation
-model that allows for efficient evaluation of the effects of MPAs on
-target and bycatch species across a range of ecological and economic
-dynamics, critically accounting for the potential impacts of effort
-displacement and technical interactions. We use this model first to
-explore general conditions that lead to positive and negative outcomes
-for target and bycatch species from MPAs. We then present a case study
-tuned to reflect the dynamics of the tuna, billfish, and shark
-populations and fisheries of the Western and Central Pacific Ocean
-(WCPO) to explore the value of optimal MPA design in managing tradeoffs
-between maximizing bycatch conservation and minimizing economic impacts.
-We chose the WCPO as a case study as the modern science of MPAs was
-largely developed around coastal ecosystems, often characterized by
-relatively static features such as coral reefs [citation]. However,
-large MPAs are increasingly being implemented and considered across
-pelagic ecosystems. These pelagic systems are characterized by highly
-mobile species and fishing fleets using a diverse mix of static and
-dynamic ocean features [@game2009; @lubchenco2003]. These ecosystems are
-also home to combinations of relatively abundant species targeted by
-fishing, such as many tuna stocks [@pons2016], and highly threatened
-species caught as incidental "bycatch" in targeted fishing operations,
-such as oceanic whitetip sharks [@pacoureau2021]. These dynamics,
-combined with the increased policy interest in large MPAs covering
-expanses of blue-water ecosystems, make pelagic systems such as the WCPO
-an ideal case study in bycatch management through MPAs. The goal of this
-paper is to provide a strategic evaluation of the sensitivity of bycatch
-populations and targeted fisheries to key unknowns and design choices in
-MPA planning.
-
-# Methods
-
-@hastings2017 employed an analytically tractable model to derive
-conditions under which MPAs would benefit bycatch species. However they
-also pointed out that even small increases in model complexity quickly
-make analytical solutions to the effects of MPAs intractable. As such,
-we employ a numerical simulation strategy in the form of a spatially
-explicit, multi-species, multi-fleet, bio-economic model we call
-[`marlin`](https://github.com/DanOvando/marlin). Models of MPAs span a
-wide range of complexity, from simple analytical tools to end-to-end
-ecosystem models [@fulton2015]. `marlin` follows a "models of
-intermediate complexity" (MICE) philosophy [@plagányi2014] targeted
-specifically around the role of technical interactions across species.
-Many MPA models simulate the impacts of MPAs on a single-species level
-[@ovando2021a]. This can be useful for for example examining the ways in
-which species-specific life history traits interact with fleet dynamics
-to influence MPA outcomes. However, bycatch problems are driven by
-technical interactions, in which fishing effort for a target species
-incidentally captures another species. `marlin` accounts for this by
-allowing the user to simulate an arbitrary number of fleets capable of
-catching an arbitrary number of species in a spatial area. Different
-fleets can place different economic value on different species, while
-also having species-specific selectivity functions. This allows us to
-simulate bycatch dynamics of MPAs in pelagic ecosystems.
+walk(rdatas, ~load(file.path(results_path,.x), envir = .GlobalEnv)) # load results into global environment
+
+```
+
+need to contrast with SEAPODYM
+
+## Introduction
+
+Communities around the world are increasingly looking to spatial
+strategies for managing marine ecosystems. For example, the "30x30"
+movement calls for implementation of marine protected areas, a form of
+spatial management, across 30% of the world's oceans by 2030. While the
+objectives of individual spatial management strategies vary widely,
+spatial management strategies are often supported based on their ability
+to provide protection to broad swaths of marine ecosystems with one
+management action, with potentially fewer costs and complications than
+for example assessing, setting, and enforcing catch quotas for
+individual species.
+
+However, the design of spatial management strategies faces numerous
+challenges. Marine ecosystems are highly dynamic and opaque, and
+extractive entities such as fishing fleets respond to changes in marine
+ecosystems in highly complex ways. These dynamics post challenges even
+for single-species assessment and management strategies [@field2006a],
+which are only amplified when we consider management policies
+specifically designed around spatial features in the oceans. For
+example, policies such as MPAs designed for both conservation and food
+security must consider factors such as the optimal size and placement of
+a protected area network given a wide array of life history,
+exploitation levels, fleet dynamics, and preferred habitats, all of
+which may change over time, particularly given the impacts of climate
+change.
+
+These challenges are amplified by the fact that our knowledge regarding
+the design and implementation of spatial management strategies such as
+MPAs is largely based on theoretical studies; the size of protected
+areas [@ovando2021a] and time required[@nickols2019] to see substantial
+effects from many MPAs means that we have relatively little empirical
+evidence as to their large-scale effects under a range of
+social-ecological settings [@mcclanahan2021].
+
+Many models have been developed to support the theory and design of
+spatial management strategies [@fulton2015]. However, they have tended
+to fall into one of either highly complex and tactical models designed
+for use in a specific location, or to be highly stylized representations
+intended to provide theoretical insights with less relevance for highly
+specific design questions. This complex landscape presents a challenge
+to stakeholders charged with marine spatial planning exercises lacking
+accesible models suited for more tactical applications, and a barrier to
+the scientific community, wherein comparisons across spatial management
+simulation studies are clouded by discrepancies in underlying model
+structures beyond differences in the core phenomenon in question.
+
+In part to help address this challenge, we present here a novel
+bio-economic modeling framework called `marlin` that allows for
+efficient simulation of spatio-temporal biological and economic
+dynamics. `marlin` allows users to simulation the impacts spatial
+policies such as MPAs across a range of species targeted by various
+fleets across heterogeneous and dynamic seascapes. Here we present the
+core methods and functionality of this framework, and demonstrate its
+use in two xx case-study applications. In the first, we show how
+`marlin` can be used to assess critical drivers for positive outcomes
+and unintended consequences of MPAs as a bycatch-mitigation strategy. In
+the second, we present a more tactical application, demonstrating how
+`marlin` can be used to compare and design MPA networks to achieve
+societal objectives.
+
+-   MPAs increasingly part of plan
+
+-   A model based field - hard to do experiments at the time and scale
+    needed
+
+-   many models out there, but challenges
+
+    -   highly stylized
+
+    -   highly tactical
+
+-   We present a general tool for modeling spatial dynamics of fished
+    populations and fishing fleets, with a focus on MPAs
+
+-   We present two case studies showing the utility of this approach
+
+    -   habitat heterogeneity and bycatch
+
+    -   MPA optimization
+
+-   Basis for improved "what if" testing of MPA scenario
+
+## Methods
+
+Models of MPAs span a wide range of complexity, from simple analytical
+tools to end-to-end ecosystem models [@fulton2015]. `marlin` follows a
+"models of intermediate complexity" (MICE) philosophy [@plagányi2014]
+targeted specifically around the role of technical interactions across
+species. Many MPA models simulate the impacts of MPAs on a
+single-species level [@ovando2021a]. This can be useful for for example
+examining the ways in which species-specific life history traits
+interact with fleet dynamics to influence MPA outcomes. However, bycatch
+problems are driven by technical interactions, in which fishing effort
+for a target species incidentally captures another species. `marlin`
+accounts for this by allowing the user to simulate an arbitrary number
+of fleets capable of catching an arbitrary number of species in a
+spatial area. Different fleets can place different economic value on
+different species, while also having species-specific selectivity
+functions. This allows us to simulate bycatch dynamics of MPAs in
+pelagic ecosystems.
 
@@
 
-## `marlin`
-
-#### Fleet Dynamics
+### Fleet Dynamics
 
@@
 
-#### Population Model
+### Population Model
 
@@
 ```
-##### Recruitment {.unnumbered}
+#### Recruitment {.unnumbered}
 
@@
 ```
-##### Dispersal {.unnumbered}
-
-Dispersal in the model is broken into two components: adults and
-recruits Both assume a Gaussian dispersal kernel of the form
-
-```{=tex}
-\begin{equation}
-m_{s,p_i,p_j}=\frac{1}{\sqrt{2\pi\sigma_{s,p_i}^2}}e^{-\frac{d_{p_i,p_j}^2}{2\sigma_{s,p_i}^2}}
-  (\#eq:move)
-\end{equation}
-```
-where *i* is the source patch, *j* is the destination patch, *d* is the
-distance between patches *i* and *j* , and $\sigma_s*$is the movement
-rate, in units of patches.
-
-Along with distance, net movement is defined by habitat quality. Habitat
-quality defines the quailty for adults or larvae of a given patch. The
-net probability of movement from patch *i* to patch *j* is defined then
-by the product of distance and habitat quality, normalized to sum to one
-from each patch, such that all else being equal organisms will move the
-most to the closest patch with the best habitat. We allow for seasonal
-movement in the model, where years are broken into two seasons. Movement
-in season two is defined as the opposite of movement in season one,
-intended to reflect the behavior of seasonal migrations.
-
-## Rule-of-thumb MPA Design Strategies
+#### Dispersal
+
+`marlin`'s movement dynamics are based on a continuous-time Markov
+chain, as described in @thorson2021a. Within this framework, the model
+allows for movement to be broken down into three components of
+*diffusion*, *advection* (follow currents), and *taxis* (active movement
+towards say a preferred habitat).
+
+xx movement equations xx
+
+While this approach to movement simulation is not commonly seen in the
+marine modeling literature, it has numerous advantages that warrant its
+seeming complexity. First, when only diffusion is present, the model
+will generalize to the familiar dynamics of a Gaussian dispersal kernel
+xx check this xx. However, the time component *t* in equation xx allows
+the movement model to be run on arbitrary time steps while preserving
+its behavior.
+
+Second, the *advection* component allows users to input useful features
+such as the outputs of ROMs models.xxx
+
+Third, the *taxis* model allow for more active habitat choices by
+species, allowing us to simulate both preferences of species in space
+and time efficiently.
+
+In this way then, `marlin` allows users to run anything from a simple
+Gaussian dispersal kernel up to a system governed by species that
+passively diffuse around a dynamic thermal range
+
+Lastly, this form has the advantage that the parameters are directly
+estimable from real data. So, if provided with say CPUE data and a
+tagging study, users can estimates the movement parameters in the same
+manner as @thorson2021, and then pass their estimated parameters to
+`marlin` for simulation.
+
+### Rule-of-thumb MPA Design Strategies
 
@@
 <!-- accuracy of the spatial depletion data used to prioritize MPA placement. -->
+
+## Reference Points
+
+A note here WHy no reference points? The problem: Fmsy and MSY are
+selectivity dependent, and once we allow for different fleets with
+different selectivities that critical can move in space, that movement
+affects the net selectivity and fishing mortality on the population.
+
+Bmsy is also affected, since if selectivity is very low, you might need
+more biomass of immature individuals relative to less biomass of more
+mature individuals; a biomass of X can be more or less reproductively
+productive depending on the size / age makeup of that biomass.
+
+SSBmsy is slightly more conserved, as over the long term the key things
+is the flow of recruits, so we do provide an option for that
+
+There are many different strategies for estimating reference points
+conditional on selectivity and regimes, see references. We leave it to
+users to find relevant reference point as required by their methods on
+their own, using the sort of MSE examples available in the package.
 
@@
 
+Two features of marine ecosystems that are often omitted from
+theoretical and simulation-based considerations of marine spatial
+planning are *technical interactions* and *species on the move*. May
+canonical MPA models, informed by planning efforts centered around
+relatively static features such as coral reefs, have assumed that
+species inhabit a fixed and often uniform (e.g. @hastings2017). In
+reality, the habitat of many species are heterogeneous in space, and can
+experience both seasonal and directional changes, e.g. climate-change
+mediated range shifts.
+
+This reality creates a basic challenge to MPA design, in that given the
+need to protect an area less than the total range of a species, the more
+complex and dynamics the habitat of the species the more challenging it
+will be to identify an area that provides effective protection today and
+tomorrow. However, *technical interactions* interact with this spatial
+complexity to create another and potentially more difficult challenge.
+One of the key questions around spatial management strategies such as
+MPAs is to what extent do they *remove* vs *displace* threats.
+
+The potential benefits of MPAs as a bycatch management are complicated
+by two key issues: effort displacement and technical interactions. One
+of the primary questions around MPAs is to what extent do MPAs remove
+versus displace fishing effort. If MPAs remove fishing effort, then
+their direct effects on target and bycatch species can only be positive.
+If however MPAs displace fishing effort, evaluating their potential
+impacts becomes more complicated. The net conservation impact of MPAs
+becomes a balancing act between the effects of reduced fishing inside
+protected borders and increased fishing intensity outside MPAs. The
+outcomes of this displacement can be further complicated by technical
+interactions, i.e. the degree to which fishing effort for one species
+affects another, particularly when species have markedly different
+habitat distributions. For example, @abbott2012 demonstrated that an MPA
+designed to protect red king crab (*Paralithodes camtschaticus*) caused
+an unintended increase in bycatch of Pacific halibut (*Hippoglossus
+stenolepis*) by the displaced bottom-trawl fishery.
+
 Given the complexity of the operating model used here, a complete
@@
 shark modeled after (). Each species has all life history parameters
-pulled from FishLife (@thorson2020), with the exception of steepness, set
-to XX, and fecundity. For the tuna and marlin species, fecundity is
+pulled from FishLife (@thorson2020), with the exception of steepness,
+set to XX, and fecundity. For the tuna and marlin species, fecundity is
 assumed to scale with weight modified by the hyperallometric exponent.
@@
 and *center* is the coordinate of the central cell. In this way, the
-larger $\sigma_c$ grows, the more variation there is betwen the centroid
-of the preferred habitat for each species.
+larger $\sigma_c$ grows, the more variation there is between the
+centroid of the preferred habitat for each species.
 
 We then control the spread of a creatures preferred habitat around that
-centroid per a Gaussian dispersal kernal in the manner
+centroid per a Gaussian dispersal kernel in the manner
 
@@
 
-Where $d$ is the euclidian distance from the $centroid_s$ to patch $p$
+Where $d$ is the euclidean distance from the $centroid_s$ to patch $p$
 and $\sigma_h$ controls the dispersal of the habitat around that range.
-
-## WCPO Case Study Application
-
-The purpose of the case study is to assess the value of optimization.
-Given the time constraints of optimization, rather than iterating over
-the `r nrow(state_experiments)`, we fix the values from that table to
-values that best reflect the state of
-`r n_distinct(experiment_obj$critter)` species reported as caught in the
-WCPO.
-
-Sigh. lots of details to add in here. This is the worst.
-
-```{r, echo = FALSE}
-knitr::kable(sci_to_com %>%
-               rename("Scientific Name" = "critter", "Common Name" = common_name))
-```
-
-## Assessing Performance
-
-Each model run started at equilibrium conditions and was then run for
-`r years` years (`r seasons * years` seasons. We assessed performance by
-comparing relevant metrics (e.g. depletion and profits) under a given
-scenario with MPAs relative to the values of those same metrics in the
-simulated counterfactual world identical in every way save for the
-presence of MPAs.
-
-Degree of habitat difference was calculated as
-
-xx will work out the notation for this xx
-
-$$
- Habitat Difference = sd(sqrt(hab))
-$$
-
-In order to weight habitat difference by cells where animals occur, not
-where they are not.
-
-# Results
-
-Our simulation experiments allow us to examine the wide range of effects
-that MPAs can have on groups of target and bycatch species depending on
-combinations of ecological, economic, and design parameters. The goal of
-this exercise is not to assign any probability to the potential impacts
-of MPAs; we have not provided any explicit probabilities to the
-alternative states modeled here, and so it may be that a state that was
-only modeled on a few occasions in our simulations is in fact highly
-common in the real world.
-
-What these experiments are designed to do is illustrate the range of
-outcomes that are possible from MPAs under the types of conditions
-evaluated here, and evaluate which drivers are potentially the most
-influential in affecting particular outcomes with regards to bycatch.
-Our case study of a WCPO-esque system is intended then to reduce the
-parameter space enough that we can examine the relative tradeoffs and
-value of optimization for economic and conservation objectives in a
-pelagic system.
-
-One clear caveat across all of these results is that we have made two
-very strong assumptions around the behavior of the fishing fleet. The
-first is that we assume that effort is displaced, not removed, by the
-creation of MPAs. This is a widely debated question around MPAs, and the
-real impacts of MPAs on fishing effort will clearly be setting
-dependent. For example, large-scale pelagic fishing fleets already have
-high rates of vessel movement built into their economic models, and so
-may be more likely to move rather than exit the fishery in response to
-MPAs. In contrast, smaller scale coastal fishing operations may simply
-exit the fishery if their local fishing grounds are placed in MPAs.
-There is clear evidence of effort displacement in many fisheries though
-(citation), and moreover a motivation for this study was to evaluate the
-performance of MPAs in reducing bycatch given the potential for effort
-displacement and technical interactions. If MPAs do not produce
-increases in fishing effort outside their borders, then the direct
-conservation effects of MPAs have to be positive for all species
-concerned.
-
-The second is that we do not account for evolution of total effort. Our
-assumption is that while fleets can make economic decisions based around
-where to allocate their effort in space, the total pool of effort they
-exert is fixed (we would note that the underling `marlin` model does
-have capacity for more realistic fleet models such as open-access
-dynamics). This is clearly an incorrect assumption over the long term,
-but allows us to focus on the impacts of the other drivers considered
-here. A more through consideration of the impacts of fleet model
-structure on MPA outcomes is generally needed, and is critical for any
-tactical efforts at MPA modeling.
-
-<!-- what we found was a series of neccesary but not sufficient conditions  -->
-
-<!-- make the tautology clear: illustrating a point, not proving a thing since clearly depends on if you assume displacement, and then degree of non-overlap etc.  -->
-
-## Generalized Experiments
-
-We focused first on the range of impacts that MPAs can have on mature
-biomass of target and bycatch species across the range of variables
-explored. To illustrate, we pulled out two example simulations, one
-where MPAs provided relatively large benefits across all three species,
-and a second where MPAs actually produced substantial losses in some
-species. In the "best case" scenario, each of the three species shares a
-common habitat distribution, and so protection of the central habitat of
-one species results in protection for all species. In the "worst case"
-scenario, an MPA designed primarily around the tuna fishery displaces
-effort onto the marlin and shark habitats, increasing the tuna
-population while driving the other two species down
-(Fig.\@ref(fig:fig-info)).
-
-```{r fig-info, fig.cap="Illustrative example demonstrating how degrees of overlap can influence "}
-
-experiment_summary <- total_experiment_results %>%
-  filter(prop_mpa < 0.5,
-         !placement_strategy %in% c("area", "avoid_fishing"),
-         placement_strategy == "rate",
-         placement_error == min(placement_error)) %>%
-  group_by(placement_id, state_id) %>%
-  summarise(
-    max_loss = max(signif_loss),
-    min_loss = max(n_loss),
-    total_gain = sum(abs_delta_biodiv)
-  ) %>%
-  ungroup()
-
-worst_case <- experiment_summary %>%
-  # filter(max_loss == max(max_loss)) %>%
-  arrange(total_gain) %>%
-  # filter(total_gain == min(total_gain)) %>%
-  slice(2)
-
-worst_state <- state_experiments %>%
-  filter(state_id == worst_case$state_id)
-
-worst_result <- compact_experiment_results %>%
-  filter(placement_id == worst_case$placement_id,
-         state_id == worst_case$state_id,
-         prop_mpa <0.5)
-
-worst_result_plot <- worst_result %>%
-  ggplot() +
-  geom_text(data = tibble(
-    x = .1,
-    y = max(worst_result$biodiv - worst_result$bau_biodiv)
-  ),
-  aes(x = x, y = y, label = "Worst Case"),
-  size = 4) +
-  geom_hline(yintercept = 0, linetype = 2) +
-  geom_line(aes(prop_mpa, biodiv - bau_biodiv, color = critter),
-            show.legend = FALSE, size = 1.25) +
-  scale_x_continuous(labels = scales::percent, name = "MPA Size") +
-  scale_y_continuous(
-    name = ''
-  )
-
-worst_habitat <- worst_state %>%
-  select(state_id, data) %>%
-  unnest(cols = data) %>%
-  select(scientific_name, contains("sigma_"), habitat) %>%
-  unnest(cols = habitat) %>%
-  mutate(
-    scientific_name = fct_recode(
-      scientific_name,
-      "Shark" = "carcharhinus longimanus",
-      "Tuna" = "katsuwonus pelamis",
-      "Marlin" = "kajikia audax"
-    )
-  )
-
-worst_habitat_plot <- worst_habitat %>%
-  ggplot(aes(x, y, z = (habitat), color = scientific_name)) +
-  geom_contour(show.legend = FALSE) +
-  scale_x_continuous(expand = c(0, 0)) +
-  scale_y_continuous(expand = c(0, 0)) +
-  theme(
-    axis.text.x = element_blank(),
-    axis.text.y = element_blank(),
-    axis.title.y = element_blank(),
-    axis.title.x = element_blank(),
-    axis.ticks.y = element_blank(),
-    axis.ticks.x = element_blank(),
-    panel.border = element_blank()
-  )
-
-best_case <- experiment_summary %>%
-  filter(min_loss == min(min_loss, na.rm = TRUE)) %>%
-  filter(total_gain == max(total_gain, na.rm = TRUE)) %>%
-  slice(1)
-
-best_state <- state_experiments %>%
-  filter(state_id == best_case$state_id)
-
-best_result <- compact_experiment_results %>%
-  filter(placement_id == best_case$placement_id,
-         state_id == best_case$state_id,
-         prop_mpa < 0.5)
-
-best_result_plot <- best_result %>%
-  ggplot() +
-  geom_hline(yintercept = 0, linetype = 2) +
-  geom_line(aes(prop_mpa, biodiv - bau_biodiv, color = critter), size = 1.25) +
-  geom_text(data = tibble(
-    x = .1,
-    y = max(best_result$biodiv - best_result$bau_biodiv)
-  ),
-  aes(x = x, y = y, label = "Best Case"),
-  size = 4) +
-  scale_x_continuous(labels = scales::percent, name = "") +
-  scale_y_continuous("Change in SSB/SSB0") +
-  theme(legend.position = "top") +
-  scale_color_discrete(name = element_blank()) +
-  labs(title = "A) ")
-
-best_habitat <- best_state %>%
-  select(state_id, data) %>%
-  unnest(cols = data) %>%
-  select(scientific_name, contains("sigma_"), habitat) %>%
-  unnest(cols = habitat) %>%
-  mutate(
-    scientific_name = fct_recode(
-      scientific_name,
-      "Shark" = "carcharhinus longimanus",
-      "Tuna" = "katsuwonus pelamis",
-      "Marlin" = "kajikia audax"
-    )
-  )
-
-best_habitat_plot <- best_habitat %>%
-  ggplot(aes(x, y, z = (habitat), color = scientific_name)) +
-  geom_contour(show.legend = FALSE) +
-  labs(title=  "B)", subtitle = "Habitat Distribution") +
-  scale_x_continuous(expand = c(0, 0)) +
-  scale_y_continuous(expand = c(0, 0)) +
-  theme(
-    axis.text.x = element_blank(),
-    axis.text.y = element_blank(),
-    axis.title.y = element_blank(),
-    axis.title.x = element_blank(),
-    axis.ticks.y = element_blank(),
-    axis.ticks.x = element_blank(),
-    panel.border = element_blank()
-  ) # +
-  # facet_wrap(~scientific_name)
-
-fig_info <- (best_result_plot / worst_result_plot) |
-  (best_habitat_plot / worst_habitat_plot)
-
-fig_info
-
-```
-
-```{r}
-
-critter_results <- compact_experiment_results %>%
-  filter(bau_biodiv > 0, prop_mpa > 0) %>% 
-  group_by(state_id, placement_id) %>%
-  mutate(real_sigma_centroid = sd(centroid)) %>%
-  mutate(delta_biodiv = pmin(1,biodiv - bau_biodiv),
-         loss = biodiv < bau_biodiv,
-         real_loss = biodiv < (0.9 * bau_biodiv),
-         rec_form = as.factor(rec_form)) %>% 
-  mutate(critter = fct_relevel(critter,c("Tuna", "Marlin","Shark")),
-         rec_form = fct_recode(rec_form, "Global" = "0", "Local" = "1", "Local Then Disperse" = "2", "Disperse Then Local" = "3"))
-
-```
-
-Looking at the full
-`r n_distinct(total_experiment_results$placement_id) * n_distinct(total_experiment_results$state_id)`
-experiments run, the results were highly variable. Across our
-simulations, gains were much more common than losses, with losses of any
-magnitude occurring in `r scales::percent(mean(critter_results$loss))`
-of simulations with MPAs, but losses of 10% or greater of SSB in only
-`r scales::percent(mean(critter_results$real_loss))` of simulations.
-However we stress again that we can assign no probabilities to any of
-these outcomes in the real world; this simply tells us that under the
-range of conditions, selected to represent plausible states of nature,
-there are many more ways for MPAs to provide a conservation effect
-greater than or equal to 0 than there are for them to provide a net loss
-for some species.
-
-The question of interest then are which of the considered drivers appear
-to be the most influential in affecting the magnitude and direction of
-conservation effects of MPAs. Fishing mortality is the most obvious
-answer; if fishing mortality is zero in the absence of an MPA then
-clearly a no-take MPA will have no effect. As such we do not focus on
-that metric here, focusing instead on other key drivers identified by
-our simulation. Four visually prominent factors were MPA size, species,
-design strategy, and degree of habitat difference. The degree to which
-MPAs of a given size impact spawning biomass depends on a host of
-factors. Any potential for conservation losses are intuitively convex in
-MPA size: effects of any kind are low when MPA size is low, and
-conservation losses must be zero when MPA size is 100% (which this graph
-does not quite reach), with peak losses in our simulations occurring
-between MPA sizes of XX and XX. Life history plays a clear role our
-stylized with shark and marlin populations being much more likely to
-experience net conservation losses as a function of MPA placement than
-the more resilient tunas. High degree of habitat differences can lead to
-both to the potential for large conservation gains or losses depending
-on whether a more isolated species is located primarily inside or
-outside of MPAs (Fig.\@ref(fig:fig-exp-biodiv)).
-
-```{r fig-exp-biodiv, fig.cap = "Trajectories of changes in SSB/SSB0 relative to BAU for each species across rule-of-thumb MPA design scenarios. Color indicates degree of habitat difference."}
-
-fig_exp_biodiv <- compact_experiment_results %>%
-  filter(prop_mpa < 1) %>%
-  # filter(f_v_m > 1.5) %>%
-  group_by(state_id, placement_id) %>%
-  ggplot(aes(
-    prop_mpa,
-    pmin(1,biodiv - bau_biodiv),
-    color =  sigma_habitat,
-    group = interaction(placement_id, state_id)
-  )) +
-  geom_line(alpha = 0.5) +
-  geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
-  facet_grid(critter ~ placement_strategy, labeller = labeller(placement_strategy = capitalize)) +
-  scale_color_viridis_c(name = "Habitat Difference", guide = guide_colorbar(frame.colour = "black",ticks.colour = "black", barwidth = unit ("15", "lines"))) +
-  scale_y_continuous("Change in SSB/SSB0 relative to BAU") +
-  scale_x_continuous(labels = scales::label_percent(accuracy = 1), name = "MPA Size") +
-  theme(legend.position = "top")
-
-fig_exp_biodiv
-```
-
-We can somewhat quantitatively evaluate the relative impact of these
-drivers using a regression, given that by construction variation in each
-of evaluated drivers are independent in our simulations. We fit a
-binomial regression to the presence or absence of of a net conservation
-loss per species as a function of the simulated experimental variables.
-The factors that according to this model most increase the probability
-of a net conservation loss occurring include habitat differences, a
-placement strategy of avoiding fishing areas, and the life history of
-the species (shark or marlin relative to tuna). The factors that most
-decrease the probability of a conservation loss from an MPA are MPA
-size, placement strategy based on local depletion, and somewhat
-interestingly fishing mortality rate
-(Fig.\@ref(fig:fig-loss-drivers)-A).
-
-However, this result is somewhat misleading (xx so should we just
-remove?xx). Fig.\@ref(fig:fig-exp-biodiv) shows the clearly bimodal
-nature of for example habitat differences. This analysis helps inform us
-that habitat differences are a necessary but not sufficient condition
-for conservation losses induced by effort displacement. The regression
-based results require assuming that each of the effects are independent
-and linear in nature. In fact each of these factors can interact in
-complex and non-linear ways that may not be apparent visually or through
-this type of regression analysis. To address this, we ran a random
-forest using the `ranger` package (citation) in R (citation) and
-extracted the impurity corrected importance scores for each of the
-variables. While this does not tell us the expected marginal effect of
-individual drivers, it does tell us which variables are important if our
-goal is to accurately predict the potential for unintended consequences
-or magnitude of conservation gains in MPAs. Note here that variables
-that intuitively should have an impact but have no visually or
-statistically (in a linear model) clear impact on the probability of
-experiencing a conservation loss such as the dispersal distances of
-adults or larvae have importance scores comparable in magnitude to
-habitat differences when considered in the context of nonlinear effects
-and interactions with other variables
-(Fig.\@ref(fig:fig-loss-drivers)-B).
-
-```{r fig-loss-drivers, fig.cap = "Things"}
-
-loss_reg <-   glm(
-  loss ~   f_v_m  + critter + seasonal_movement + adult_movement_sigma + rec_form + recruit_movement_sigma + prop_mpa + sigma_habitat + spatial_allocation + placement_strategy + hyper,
-  family = binomial(),
-  data = critter_results %>% filter(between(prop_mpa,.2,.4))
-)
-
-loss_reg_coefs <- broom::tidy(loss_reg) %>%
-  mutate(lower = estimate - 1.96 * std.error,
-         upper = estimate + 1.96 * std.error) %>%
-  mutate(
-    term = str_replace(term, "placement_strategy", "Placement Strategy: "),
-    term = str_replace(term, "rec_form", "DD Timing: "),
-    term = str_replace(term, "spatial_allocation", "Fleet Allocation: "),
-    term = str_replace(term, "hyper", "Hyperallometry present"),
-    term = str_replace(term, "prop_mpa", "MPA size"),
-    term = str_replace(term, "sigma_habitat", "Habitat differences"),
-     term = str_remove_all(term, "critter")
-  )
-
-loss_coef_plot <- loss_reg_coefs %>%
-  ggplot() +
-  geom_hline(aes(yintercept = 0)) +
-  geom_pointrange(aes(
-    forcats::fct_reorder(term, estimate),
-    estimate,
-    ymin = lower,
-    ymax = upper
-  )) +
-  scale_y_continuous(name = "Log-odds Ratio Effect") +
-  scale_x_discrete(name = '') +
-  coord_flip()
-
-tmp_critter <-
-  critter_results %>% filter(between(prop_mpa, .2, .4)) %>%
-  select(-habitat) %>% 
-  mutate(fctr_loss = factor(loss))
-
-loss_rf <-   ranger::ranger(
-  loss ~   f_v_m  + critter + seasonal_movement + adult_movement_sigma + rec_form + recruit_movement_sigma + prop_mpa + sigma_habitat + spatial_allocation + placement_strategy + hyper,
-  data = tmp_critter,
-  importance = "impurity"
-)
-
-loss_importance <- data.frame(importance = loss_rf$variable.importance %>% sort()) %>% 
-  rownames_to_column(var = "term")
-
-loss_importance_plot <- loss_importance %>% 
-  ggplot(aes(forcats::fct_reorder(term, importance), importance)) + 
-  geom_col() +
-  scale_y_continuous(name = "Variable Importance") +
-  scale_x_discrete(name = '') +
-  coord_flip()
-
-fig_loss_drivers <- loss_coef_plot + loss_importance_plot
-
-fig_loss_drivers
-```
-
-```{r fig-delta-drivers, eval = FALSE}
-
-delta_reg <-   rstanarm::stan_glm(
-  delta_biodiv ~   f_v_m  + critter + seasonal_movement + adult_movement_sigma + rec_form + recruit_movement_sigma + prop_mpa + sigma_habitat + spatial_allocation + hyper,
-  data = critter_results %>% filter(between(prop_mpa,.2,.4)),
-  cores = 4,
-  chains = 4
-)
-
-tmp_critter <-
-  critter_results %>% filter(between(prop_mpa, 0.2, 0.4)) %>%
-  select(-habitat) %>% 
-  mutate(fctr_loss = factor(loss))
-# 
-# tmp_rcp <- recipes::recipe(loss ~ ., data = tmp_critter) %>% 
-#   themis::step_upsample(fctr_loss, over_ratio = 0.75) %>% 
-#   recipes::prep(training = tmp_critter, retain = TRUE) %>% 
-#   recipes::juice()
-
-delta_rf <-   ranger::ranger(
-  delta_biodiv ~   f_v_m  + critter + seasonal_movement + adult_movement_sigma + rec_form + recruit_movement_sigma + prop_mpa + sigma_habitat + spatial_allocation + placement_strategy + hyper,
-  data = tmp_critter,
-  importance = "impurity"
-)
-
-delta_rf$variable.importance %>% sort()
-
-plot(delta_reg)
-
-```
-
-## WCPO Case Study
-
-<!-- The generalized experiments help us understand the potential impacts of -->
-
-<!-- the evaluated drivers, but also highlight the challenge in making -->
-
-<!-- definitive statements about the effect of a given driver on the conservation impacts of MPAs, given the complex ways in which all of these factors -->
-
-<!-- interact, even within a relatively simple model such as this one. In -->
-
-<!-- order to help illustrate how addressing the kinds of spatial dynamics -->
-
-<!-- and technical interactions included in our model can affect and benefit -->
-
-<!-- marine spatial planning, we then ran a case study intended to loosely -->
-
-<!-- reflect the dynamics of WCPO tuna, billfish, and shark fisheries and -->
-
-<!-- populations. -->
-
-Our case study allows us to focus in on important questions around the
-relative performance of "rule of thumb" MPA design strategies relative
-to optimized networks. In particular, as any leakage induced losses from
-MPAs must be 0 if MPA size is 100%, an important question then is within
-this case study what tradeoffs are present between increasing the
-probability of a successful conservation outcome by increasing MPA size
-and the increasing probability of negative economic impacts with MPA
-size (xx phrase better xx).
-
-Given the set of parameters simulated here, we find little intrinsic
-risk for large-scale unintended losses in total spawning biomass caused
-by MPA effort displacement for the species included in our WCPO case
-study. Each of the design scenarios produced at least some small loss in
-biomass for one or more species at some MPA size. However, these losses
-were generally small and isolated to a subset of species within a small
-MPA size range, with the exception of the "Avoid_fishing" MPA placement
-strategy that resulted in net losses for most species at large MPA sizes
-(though simply minimal impact at smaller sizes). All design strategies
-except "Avoid_fishing" and "Rate" resulted in relatively rapid declines
-in profits for the three main tuna fisheries (skipjack, yellowfin, and
-bigeye). Both "Avoid_fishing" and "Rate" maintained profits close to
-business as usual up to a large MPA size, but did so at the expense of
-conservation benefits (Fig.\@ref(fig:fig-cs-critters)). Results for
-yield rather than profits are qualitatively similar, see SI.
-
-```{r fig-cs-critters, fig.cap = "Simulated trajectories of changes in profits and spawning stock biomass (SSB) for selected species caught in the WCPO.
-Each column represents a different MPA design strategy.
-'Optimized' refers to an MPA network optimized to maximize total SSB across all species."}
-
-opt_comparison <- opt_obj %>%
-  filter(alpha == max(alpha)) %>%
-  mutate(placement_error = 0,
-         prop_mpa = p_protected / 100,
-         placement_strategy = "Optimized") %>%
-  select(critter, placement_error, delta_biodiv, delta_econ, delta_yield, prop_mpa,placement_strategy)
-
- experiment_obj %>% 
-   select(colnames(opt_comparison)) %>% 
-  bind_rows(opt_comparison) %>% 
-   mutate(placement_strategy = fct_relevel(placement_strategy,"Optimized")) %>% 
-   left_join(sci_to_com, by = "critter") %>% 
-  filter(placement_error == min(placement_error)) %>% 
-   rename("Change in SSB" = delta_biodiv, "Change in Profits" = delta_econ) %>% 
-  pivot_longer(c("Change in SSB", "Change in Profits"), names_to = "metric", values_to = "value") %>% 
-  ggplot(aes(prop_mpa, value, color = common_name)) + 
-   scale_color_discrete(name = '') +
-   geom_hline(aes(yintercept = 0), linetype = 2)+
-  geom_line(show.legend = TRUE) + 
-   scale_y_continuous(name = '') +
-   scale_x_continuous(name = "MPA Size",labels = scales::percent, guide = guide_axis(n.dodge = 2)) +
-  facet_grid(metric~placement_strategy, scales = "free_y", switch = "y", labeller = labeller(placement_strategy = capitalize)) + 
-   theme(strip.text.x = element_text(size = 8), 
-         legend.position = "top", 
-         legend.text = element_text(size = 8))
-
-```
-
-Viewed as totals, we found a small parameter space under the "Area" and
-"Avoid_fishing" design strategies where MPAs were capable of causing a
-net decrease in total spawning stock biomass across all evaluated
-species. However, the efficiency frontier indicates that these points
-are not Pareto efficient, and MPA design configurations exist that
-achieve the same levels of economic performance while producing a net
-gain in spawning biomass. Optimal MPA design was able to slow the
-tradeoff between total profits and bycatch spawning biomass, but for any
-conservation weighting any increases in total biomass of bycatch species
-came at a decrease in total profits. However we also found limited need
-for optimized networks, with the "Depletion" design scenario essentially
-following the Pareto efficiency frontier (Fig.\@ref(fig:fig-cs-totals)).
-
-```{r fig-cs-totals, fig.cap = "Efficiency frontiers for changes in total profits and total bycatch spawning stock biomass (SSB).
-Color gradient line shows optimized pareto frontier, with color showing weighting of conservation (total spawning stock biomass) in the objective function.
-Red points show outcomes of 'rule of thumb' MPA design strategies.
-"}
- total_opt_obj %>%
-  group_by(alpha) %>%
-  filter(obj_fun == max(obj_fun)) %>%
-  ungroup() %>%
-  ggplot() +
-  geom_line(aes(total_bycatch_ssb_change, delta_econ, color = alpha), size = 1.5) +
-  geom_point(
-    data = total_experiment_obj,
-    aes(total_bycatch_ssb_change, delta_econ),
-    size = 2,
-    alpha = 0.1,
-    fill = "red",
-    shape = 21
-  ) +
-  geom_vline(aes(xintercept = 0), linetype = 2) +
-  geom_hline(aes(yintercept = 0), linetype = 2) +
-  geom_abline(intercept = 0, slope = -1) +
-  facet_wrap(~placement_strategy, scales = "free_y", labeller = labeller(placement_strategy = capitalize)) + 
-  scale_color_viridis_c(name = "Conservation Weighting in Optimization",
-                        guide = guide_colorbar(frame.colour = "black", ticks.colour = "black", barwidth = unit("15", "lines"))) +
-  scale_x_continuous(name = "Change in Total Bycatch SSB") +
-  scale_y_continuous(name = "Change in Total Profits ") +
-  theme(axis.text.x = element_text(size = 8),
-        legend.position = "top")
-
-```
-
-```{r, eval = FALSE}
-
- total_opt_obj %>%
-  group_by(alpha) %>%
-  filter(obj_fun == max(obj_fun)) %>%
-  ungroup() %>%
-  ggplot() +
-  geom_line(aes(delta_biodiv, delta_econ, color = alpha), size = 2) +
-  geom_point(
-    data = total_experiment_obj,
-    aes(delta_biodiv, delta_econ),
-    size = 2,
-    alpha = 0.1,
-    fill = "red",
-    shape = 21
-  ) +
-  geom_vline(aes(xintercept = 0), linetype = 2) +
-  geom_hline(aes(yintercept = 0), linetype = 2) +
-  geom_abline(intercept = 0, slope = -1) +
-  facet_wrap( ~ placement_strategy) +
-  scale_color_viridis_c(name = "Conservation Weighting in Optimization") +
-  scale_x_continuous(name = "Change in Total Biodiversity Relative to BAU ") +
-  scale_y_continuous(name = "Change in Total Profits Relative to BAU ") +
-  theme(axis.text.x = element_text(size = 8),
-        legend.position = "top")
-
-```
-
-# Discussion
-
-MPAs can be a powerful tool for providing conservation benefits across
-large swaths of species and space. Under the appropriate circumstances
-they can also provide benefits to food security and economic outcomes.
-However, while spatial closures are simple in concept, predicting the
-likely effects of MPAs across groups of species and fishing fleets
-becomes increasingly complicated as the systems being studied deviate
-from assumptions of homogenously distributed species with perfectly
-selective fleets. In this paper we used a new bio-economic model to
-demonstrate how a range of variables often omitted in MPA modeling can
-act to influence the outcomes of MPAs on conservation and economic
-objectives, particularly with regards to protecting bycatch species. We
-demonstrate that unintended conservation losses to both target and
-bycatch species are possible, but only under a range of specific
-conditions, and under most simulated cases did not occur. MPAs were
-capable of causing small losses in biomass for the WCPO case study, but
-only for a subset of design strategies and species, most of which could
-be overcome by optimized design.
-
-Our results indicate that as suggested by theory (citation) and
-empirical evidence (citation), MPAs can indeed have unintended
-consequences for conservation if they result in effort displacement
-rather than attrition. However, our model found these to be a minority
-of the simulations tested. This does not suggest that leakage may not a
-problem for specific MPA. But, it also indicates that across the range
-of plausible parameter combinations that we explored it was much more
-common for MPAs to provide net conservation benefits even if effort is
-displaced. This suggests then that we should of course not dismiss the
-conservation benefits of MPAs in the presence of effort displacement,
-but also that users must be aware of the potential for unintended
-consequences that can indeed occur. While many different factors
-interact in complex ways to affect the potential for unintended
-conservation consequences, the most obvious is the degree of difference
-in the habitats utilized by the species within an ecosystem in which an
-MPA is to be placed. As a starting point, it is critical that
-communities engaged in marine spatial planning create a localized assessment of
-the degree of habitat overlap across potentially affected species to
-assess the potential risk of unintended consequences.
-
-The MPA literature is highly dominated by modeling studies (citation).
-What then is the added value of the model and results presented here?
-Our model contains many features missing from many general MPA models,
-including size structure, seasonal movement, larval and adult habitats
-and dispersal, and critically technical interactions xx clearer term for
-this xx across species. A reader might reasonably ask about a range of
-other factors not included, such as time-varying parameters and species
-interactions. One could of course continue to add features to the model
-until it reaches the complexity of an end-to-end model such as Atlantis
-(citation). While such models have tremendous value, our goal here is to
-present a model that can help provide closer to tactical advice for
-users without the data or capacity for such high resolution models. One
-justification for simpler MPA models is often computational
-tractability: tracking the full age structure of multiple species and
-the economic outcomes of multiple fleets dynamically across a high
-resolution spatial grid can quickly become computationally expensive.
-Faster implementations completely written in languages such as C++
-however become inaccessible to many users. Our model seeks to fill a
-middle ground here, with the core operations run in C++ while providing
-a relatively user-friendly front end in the open-source and commonly
-used R programming language. This allows users to ask the sorts of
-questions addressed in this paper that require a more realistic model,
-without requiring inordinate amounts of data, programming expertise, or
-computational time.
-
-Several results from our model may seem counterintuitive based on
-existing literature. We found little effect of the timing and spatial
-structure of density dependence in our model. Other studies have pointed
-out that decisions about whether density dependence is local or global
-can have major implications for fishery outcomes of MPAs (citations).
-The difference here is that our model is fully age structured. As such,
-yields are separated from recruitment, the location of density dependent
-processes in our model, by growth, natural mortality, and adult movment.
-In contrast biomass dynamics models assume that density dependence
-occurs on a net "growth" parameter combining both recruitment and
-biomass growth, and that this growth is instantly available to the
-fishery. As such we suspect that the timing of density dependence plays
-a much bigger role in the expected outcomes of MPAs when biomass
-dynamics models are used as opposed to when it is included as an effect
-on recruitment in an age structured model.
-
-@marshall2021 pointed out that many fisheries models assume that
-fecundity is simply proportional to spawning biomass, when in fact there
-is ample empirical evidence that fecundity follows a hyperallometric
-pattern, i.e. that fecundity increases faster than weight. In that study
-the authors found that the presence of hyperallometric growth
-substantially amplified the predicted benefits of spatiao-temporal
-closures such as MPAs, relative to an assumption of allometric
-fecundity. In contrast, we find that the presence of hyperallometric
-fecundity, even at levels greater than the mean values reported in
-@marshall2021, had little effect on the expected conservation outcomes
-of MPAs (Fig.\@ref(fig:fig-loss-drivers), see SI for additional
-analyses). The key distinction here is that @marshall2021 evaluated the
-effect of MPAs under different assumptions of management, e.g.
-management based on spawning potential ratio values calculated assuming
-allmetric fecundity when in fact the relationship was hyperallometric.
-Our study holds effort constant, and so reflects more the isolated
-effect of hyperallometric on MPA outcomes independent of management
-actions.
-
-@rassweiler2012 and @rassweiler2014 both demonstrated the potential
-importance of optimized MPA planning over more "rule of thumb" MPA
-designs. In contrast, we found little value of optimization, with the
-"Depletion" MPA allocation strategy more or less following the Pareto
-efficiency frontier (Fig.\@ref(fig:fig-cs-totals)). As pointed out in
-@rassweiler2014, the value of optimization (at least if the objective
-function depends on both fishery yields and fish biomass) depends on the
-degree of overfishing present. In our WCPO case study, best available
-evidence is that none of the major tuna stocks that provide the bulk of
-yield, revenues, and profits to the WCPO fleets are overfished or
-experiencing overfishing (citations, need to update with assessments
-since pons is out of date now). Our results then are consistent with
-@rassweiler2014, and evaluations of our model on stocks experiencing
-overfishing will likely find much higher value of optimization.
-
-Value of information and fleet dynamics are two critical areas for
-future study in this realm. Our study assumes perfect information in the
-design of MPAs. We feel this is a useful benchmark for isolating the
-potential impacts of key ecological and economic drivers of MPA effects.
-However it is clearly not a realistic assumption, particularly for
-pelagic ecosystems where species inhabit a complex, dynamic, and hard to
-observe mosaic of habitats. This is likely to be particularly true for
-rarer bycatch species such as sharks that may be the impetus behind many
-pelagic MPA designs. Further research is needed to examine how the types
-of dynamics explored here are affected by varying degrees of information
-quality.
-
-We focused our efforts on modeling the spatial distribution of a fishing
-fleet conditional on a fixed pool of fishing effort. As discussed
-previously, the outcomes of MPAs depend critically on the reaction of
-fishing fleets to new spatial protections. In reality neither perfect
-attrition nor perfect redistribution are likely to be realistic
-scenarios. Instead individual fishing operations will adjust their
-behavior according to their own objective functions. More realistic
-models of total effort could include simple open-access dynamics in the
-manner of @cabral2019, to more complex dynamics reflecting combinations
-of policies such as effort controls, spatial closures, and rights based
-fisheries management.
-
-As demonstrated here there though there is a possibly endless list of
-refinements and realism that could be added to simulation based studies.
-While efforts such as this can help illustrate which drivers may be the
-most important in particular circumstances in many cases as demonstrated
-here the effect of specific variables are likely to be so context
-dependent as to defy easy classification of "important" or "not". We
-feel these models are better viewed as a foundation for empirical
-action. In terms of planning, models such as this can be used as tool
-for answering "what if" rather "what will be". Helping community members
-visualize the potential impacts of alternative states of nature, such as
-debates around life history or habitat distribution, can help set
-expectations and guide data collection. Models such as this can also
-serve as hypotheses for expected effects that can ideally be confronted
-with data as MPAs are actually implemented.
-
-# Conclusions
-
-The bluntness of MPAs as a tool is both their strength and weakness.
-Targeted fisheries management is often focused on the assessment and
-management of individual stocks. While this has worked well in many
-cases [@hilborn2014a; @hilborn2020; @worm2009], it is an unrealistic and perhaps undesirable 
-model to replicate in many parts of the world, where data limitations
-and fishery complexity make this sort of highly targeted management
-untenable. In contrast, MPAs can provide protection to entire portions
-of an ecosystem with minimal need for targeted interventions. As a
-result and as shown by @hastings2017, MPAs can provide an effective
-solution to bycatch problems under the right circumstances. We extend
-the spirit of that analysis here, but expanding the model used to assess
-the impact of MPAs on bycatch under broader range of scenarios. As
-predicted by @hastings2017, as conditions deviate from the stylized
-nature of their model the effects of MPAs on bycatch become more
-nuanced. While we similarly find a large parameters space where MPAs
-provide clear benefits to the conservation status of both target and
-bycatch species, we also identify a smaller subset of enabling
-conditions, such as degrees of habitat overlap, that can lead to MPAs
-producing net reductions in bycatch biomass. As increasing numbers of
-MPA networks are implemented around the world, tools such as this can
-help communities both design and evaluate the performance of MPA
-networks with the goal of better outcomes for marine ecosystems and the
-human communities that depend on them.
-
-# References
-
-::: {#refs}
-:::
-
-# SI
-
-```{r fig-cs-yield-critters, fig.cap = "Simulated trajectories of changes in yields and spawning stock biomass (SSB) for selected species caught in the WCPO.
-Each column represents a different MPA design strategies.
-Optimized refers to an MPA network optimized to maximize total SSB across all species."}
-
-opt_comparison <- opt_obj %>%
-  filter(alpha == max(alpha)) %>%
-  mutate(placement_error = 0,
-         prop_mpa = p_protected / 100,
-         placement_strategy = "Optimized") %>%
-  select(critter, placement_error, delta_biodiv, delta_econ, delta_yield, prop_mpa,placement_strategy)
-
- experiment_obj %>% 
-   select(colnames(opt_comparison)) %>% 
-  bind_rows(opt_comparison) %>% 
-   mutate(placement_strategy = fct_relevel(placement_strategy,"Optimized")) %>% 
-   left_join(sci_to_com, by = "critter") %>% 
-  filter(placement_error == min(placement_error)) %>% 
-   rename("Change in SSB" = delta_biodiv, "Change in Yield" = delta_yield) %>% 
-  pivot_longer(c("Change in SSB", "Change in Yield"), names_to = "metric", values_to = "value") %>% 
-  ggplot(aes(prop_mpa, value, color = common_name)) + 
-   scale_color_discrete(name = '') +
-   geom_hline(aes(yintercept = 0), linetype = 2)+
-  geom_line(show.legend = TRUE) + 
-   scale_y_continuous(name = '') +
-   scale_x_continuous(name = "MPA Size",labels = scales::percent, guide = guide_axis(n.dodge = 2)) +
-  facet_grid(metric~placement_strategy, scales = "free_y", switch = "y", labeller = labeller(placement_strategy = capitalize)) + 
-   theme(strip.text.x = element_text(size = 8), 
-         legend.position = "top", 
-         legend.text = element_text(size = 8))
-
-```
-
-# Scraps
-
-There two general results beg some important questions. First, are
-unintended consequences inevitable given the kinds of technical
-interactions explored here, or can optimized MPA networks reduce or
-eliminate these unintended consequences. Second, one clear solution to
-the problem of unintended consequences is to simply make the MPAs large:
-if MPAs cover 100% of the area there can be no direct unintended
-conservation consequences. The potential for negative impacts on
-economic fishery outcomes increases as MPAs reach large size though.
-What then are the tradeoffs between reducing the potential for
-unintended conservation consequences and fishery revenues?
-
-To address these questions, we ran a more limited "case study" set of
-simulations, based on the dynamics of the WCPO tuna fisheries. Rather
-than iterating over different movement dynamics and habitat
-configurations, we fix habitat layers based on reported CPUE for the
-species in question, and xx other things we fix xx.
-
-The general experiments conducted show two general insights
-
-1.  Even conditional on fishing exploitation and general life history
-    strategy, the effects of MPAs on a group of species with technical
-    interactions is highly variable and dependent on habitat
-    distribution, design decisions, and data quality. xx make this more
-    explicit.
-
-2.  Some degree of unintended conservation consequences occured across
-    nearly all of our simulations below MPA coverage near 100%, though
-    small losses were much more common than large. Different design
-    strategies are better suited to avoiding unintended consequences.
-
-Thinking through the "naive" MPA design strategies to compare against
-the optimized outcomes. Thoughts at the moment below. These can be run
-in a few seconds each, so we cab be be pretty liberal in coming up with
-combinations if there are other ideas people want to try, but also if
-some of these seem pointless can drop as well! 6:11 Design strategies 1.
-Assign protection by SSB / SSB0 weighted abundance in space (protect the
-highest concentrations of the most threatened species) 2. Assign
-protection by bycatch rates (distribution of biomass bycatch / biomass
-target) 3. Assign protection by targeting fishing effort 4. Assign
-protection by avoiding fishing effort Knowledge states 1. Degrees of
-accuracy over the thing being measured (i.e. error in the design
-strategy) 2. Number of species considered (e.g. design aronnd only 2
-bycatch species instead of all) 3. über naive: just close off random
-patches up to a specified size We then plot these points in objective
-function space and compare to the efficiency frontier / consider
-frequency and magnitude of conservation and economic tradeoofs The
-question all this is answering: how sensitive are conservation and
-economic outcomes to design strategy / uncertainty when there are
-technical interactions and heterogeneity in habitat? The first stage is
-teeing up the idea that beyond conservation / yield tradeoffs, as you
-increase habitat heterogeneity with technical interactions you increase
-the potential for conservation tradeoffs. But, the degree is going to be
-super context dependent, so to dive in a bit we then focus on this case
-study of the particular species / habitat overlap in the WCPO. Further
-work might reveal more generalities in this, but htis is basically
-getting a sense for the potential magnitude of the challenge.
-
-XX stress a key difference: may other papers on this are static, in that
-they just some up displacement on different layers, rather than
-addressing dynamic respones over tim xx
-
-We show.... - sensitivity of both conservation and fishery outcomes to
-pelagic realism and design choices - compared to targeted / dynamic
-intervention... - value of information would be great...
-
-Contrast to coastal MPA theoretical work - @hastings2017 - @ovando2021 -
-@white2010 - not quite sure what to do about points raised in
-@fulton2011a - to what extent has this been done...
-
-The paragraph from @fulton2011a that broke me...
-
-" Spatial management was insufficient as the dominant management lever
-because it proves quite difficult to strike the balance between
-protecting areas that are of a size sufficient to produce any
-conservation benefit and providing sufficient fishing grounds to sustain
-an econom- ically viable industry (which is not possible in many
-fisheries if closures on the order of 70--80% or more are imposed,
-Fulton et al. 2007). More- over, Savina etal. (2008) found that effort
-displaced by closures potentially undermines any benefits coming from
-the zoning. For instance, strong improvements in diversity (by nearly
-20%) under heavy fishing pressure occurred if effort outside a marine
-protected area (MPA) was controlled. However, without effort control,
-there was a significant (\>20%) drop in overall diversity. Fulton et al.
-(2007), Savina et al. (2008), Smith et al. (2010) and unpublished work
-associated with Horne et al. (2010) all found that while spatial
-management appears to be an effective means of satisfying conservation
-objectives con- cerning habitat and restoring age-structures of less
-mobile species, benefits from closures for individual species depend on
-its trophic position, degree of mobility, geographical distribution and
-extent of ontogenetic shifts in habitat use and depth (highly mobile or
-prey species do not benefit in the way that more site attached or
-predatory species do)."
-
-Lots of studies have shown that lots of things matter, can we provide
-guidance on which things are priorities?
-
-Where does this fit in context of @smith2021? One important difference
-is that they point out differences in effectiveness, where effectiveness
-is still almost always positive, since they just measure leatherbacks.
-So their fig 5, there are a few cases where leatherback catches actually
-go up, but it's rare. So, extending to multiple species and keeping
-track off all of them shows it's even harder
-
-How does this contrast with the the large and growing dynamic ocean
-management lit? - By definition, with perfect information dynamic $\geq$
-static spatial controls - DOM literature is mostly static, tracks
-displacement onto static abundance / risk layers - Doesn't allow for
-dynamic impacts, for example fleet adjustments based on updated economic
-surface, or short-term costs for long-term benefits (can overstate
-costs?) - Dynamic model allows for delta between dynamic policies, e.g.
-Fmsy vs DOM - analytical models a la hastings have value but do / don't
-stand up to realism
-
-Presentation of the model as a MICE model for MSP - reasonable
-parameterizable - flexible - fast
-
-Focus of debate has been on fisheries and yields. As MPAs move to
-pelagic systems though, we show that conservation outcomes also depend
-on correctly identifying ecological state / design decisions etc.
-
-What are the considerations of protection you need to make good
-decisions. It's not always obvious how to design things to get the
-outcome you want. If we could make the case that something like closing
-the high seas is insufficient conservation management without
-conservation management inside EEZs / what is the
-
-Need to do an optimization here, calculate the marginal value of added
-cells.
-
-What's this about? Pointing out a problem, which to make impactful would
-require making the case that there is a problem with current design
-strategies. e.g.. "MPAs are often designed like this (simple dispersive
-models with constant habitat, or MARXSAN style things), and doing that
-does/does not create problems A tactical evaluation. Picking a handful
-of high profile MPAs, parameterizing all the knowns as well as possible,
-iterating over unknowns, and see what the range of outcomes are.
-Cleaner, more impactful, but also opens it up (depending on the results)
-to a details fight about very specific modeling choices How plausible
-are losses? Maybe we drill down on this a bit. From a purely bycatch
-perspective, the usual questions of fisheries tradeoffs aren't on the
-table (but can add to the discussion considerations of how much target
-fishing you give up). Given that then, while it's possible that
-"targeted" interventions would work better (ignoring costs/enforcement,
-etc), there's really no reason not to do MPAs. Unless, you actually end
-up concentrating fishing pressure outside on populations. So, we
-consider how plausible is it for that to really happen. Issue here is
-that seems likely to be pretty hand-wavey. Most likely that the vast
-majority of simulations will show positive impacts, but have no way of
-assigning probabilities to any one state of nature.
-
-The Role of Blue-Water Marine Protected Areas in Bycatch Reduction
-
-What's the question: How effective are Blue-Water MPAs in reducing
-fishing mortality of important pelagic bycatch species?
-
-Another angle. There's going to be a lot of unknowns in designing BWMPA.
-How much does not fully understanding a few key life history (e.g.
-movement) matter, and how much does the role that the species plays in
-the broader fleets portfolio matter
-
-Could simulate habitat from a multivariate normal, where you use the
-sigma to control how similar rasters are across species
-
-Marine protected areas are increasingly proposed to protect biodiversity
-in pelagic systems. However, maby paradigms of mpa design come from
-coastal systems. Here we ask, how likely are blue-water MPAs to provide
-conservation benefits to bycatch species, and what are the key
-uncertainties underlying these outcomes? We begin with a general
-discussion of the ways in which traits of open-ocean systems might
-influence the conservation performance of BWMPAs. Given the massive
-parameter space, we then constrain ourselves to a styulized version of
-the western central pacitic ocean to ask, how feasible is it to provide
-broad protection of bycatch species through common MPA design paradigms
-in the context of this system. Our results demonstrate the critical
-importance of XX and XX in designing MPAs in pelagic systems, and
-highlight the potential risks of not accounting for XX. Introductionhere
-
-Bycatch is a significant driver of population declines for many pelagic
-species Sharks, seabirds, mammals, etc MPAs increasingly turned to as a
-solution for conservation problems, etc. MPA science is largely based on
-dynamics of coastal systems. Open-ocean ecosystems have very different
-dynamics that require dedicated research Burgess et al. (2018) estimated
-the reductions in fishing mortality needed to halt declines for a number
-of bycatch species Here we ask, what are the key drivers of bycatch
-conservation outcomes from MPAs? There are all of these species; you
-reduce fishing mortality. You can either do fisheries management, or you
-can do MPAs, let's focus in on the MPA side. This could reduce fishing
-mortality, or it could do, and it depends on the dynamics of the species
-and the fleets. When can MPAs reduce fishing mortality enough for
-bycatch species? And where can you go wrong?
-
-It doesn't make any sense to ask "do MPAs help bycatch": the question is
-obviously it depends. To quickly make this case, we'll assume that a
-species is in need of recovery. We then do the simple mako example to
-show that effects of spatial closure depend on the overlap of habitat
-and the assumption of fleet dynamics, and provide some empirical
-evidence that in particular effort redistribution is not unlikely.
-
-Determining the effect of an MPA network on bycatch then will also
-depend on trying to get close to a tactical model. But, this creates a
-problem for blue water systems in particular. Can make the argument that
-given the centrality of relatively static habitat features in coastal
-MPAs "protect the key habitats" might not be a bad strategy from a
-conservation strategy. But in highly mobile pelagic systems there (may
-be) a lot more variables in play. Where should researchers direct their
-attention?
-
-To reduce the parameter space, we constrain ourselves to a system where
-1) we know the species involved 2) we have some sense of the relative
-depletion and value of the different species and 3) some basic life
-history of the species. We then vary commonly unknown parameters,
-particularly preferred habitats, spatial catchability, movement rates
-etc, and pair these simulations with different MPA designs, and examine
-the conservation outcomes for bycatch species.
-
-What i don't like about this
-
-There isn't an "answer" to this: baring something like "changing the
-type of density dependence never had any effect greater than XX%", for
-anything that on occasion has an impact the answer of "how much" will
-almost always depend on what else is going on.
-
-So, it seems like there are two real angles of attack here
-
-Pointing out a problem, which to make impactful would require making the
-case that there is a problem with current design strategies. I.e. "MPAs
-are often designed like this (simple dispersive models with constant
-habitat, or MARXSAN style things), and doing that does/does not create
-problems A tactical evaluation. Picking a handful of high profile MPAs,
-parameterizing all the knownns as well as possible, iterating over
-unknowns, and see what the range of outcomes are. Cleaner, more
-impactful, but also opens it up (depending on the results) to a details
-fight about very specific modeling choices
-
-Methods
-
-Collect life history for key bycatch species and target species, cite
-Bradley et al. pulling together species distribution and threats etc.
-Introduce marlin Parameterize marlin runs with "knowns" from life
-history of bycatch and target species Highlight key "unknowns" that we
-will iterate over, e.g. fleet models, density dependent movement, timing
-of density dependence etc Simulate effects of MPAs given reasonably
-realistic situations but iterating over key unknowns Evaluate the range
-of outcomes and key drivers of MPA effects I'm thinking we try and avoid
-a really specific case study but hard not to accidentally do one without
-making everything up....
-
-Results Range of likely changes in fishing-induced mortality for the
-species selected Key drivers of uncertainty in those outcomes, focusing
-on things that switch states from reduction to increase in mortality How
-common are conservation tradeoffs between bycatch species? Fishery
-impacts
-
-Discussion What does this say about BWMPAs and bycatch? Not going to get
-into design decisions here, depending how common tradeoffs are I think
-that will be the main focus Highlight any species that frequently
-induced tradeoffs Tee up future research MPAs in context of other
-management MPAs as risk reduction strategy Etc.
-
-We do not focus into the individual effects of particular variables as
-we found them to be highly dependent on the exact question asked and the
-set of data included, and since we cannot ascribe any actual likelihood
-to any individual state of nature simulated here.
-
-However, three important features do emerge. The first is that life
-history plays an important role in the potential for unintended
-conservation consequences from MPAs. All else equal, MPAs were much less
-likely to produce losses in spawning biomass for our tuna-like species
-than the marlin or shark-like species. Given that within the broad life
-history traits we do vary adult and larval movement, fishing mortality,
-as well as the timing of density dependence, we believe this result is
-mostly due to the high resilience (high steepness, low age at maturity,
-fast growth) of the tuna-like species relative to the marlin and
-shark-like species.
-
-The second is that design matters...
-
-```{r}
-
-```
-
-The third is that extreme outcomes (both positive and negative),
-particularly at lower MPA sizes, are largely driven by the presence of
-small range sizes interacting with seasonal migrations. Across species,
-simulations without seasonal migrations and small ranges were much more
-likely to produce extreme results than those with seasonal migrations
-and/or larger ranges. Presence of seasonal migrations acts as a
-dilutant: As a given static MPA only provides protection, or displaces
-fishing effort, onto a particular area in only half of the year, both
-the benefits and costs of MPAs are smaller. When seasonal migrations are
-not present and the species range is small, MPAs are more capable of
-providing either large benefits or large losses.
-
-\\Big message: even constrained by the same initial depletion and core
-biology (e.g. movement rates), a wide range of outcomes are possible for
-each species for a given MPA size, dependent on the design strategy,
-habitat overlap, migrations, quality of information, etc. Across the
-board as MPA size increases the ratio of positive to negative outcomes
-increases, until all species achieve their maximum conservation
-potential in every iteration at 100% MPA. HOwever, every species
-experienced a set of simulations, roughyl xx%, that resulted in net
-losses in SSB/SSB0 relative to BAU. Losses were smaller than potential
-gaines, but reached meaningful levels in some species, and interestingly
-peaked in many cases right before max MPA (reflecting concentration of
-fishing in small amounts of habitat in habited by the species).Important
-to stress in here that there are no probabilities in this, no way of
-knowing which state is more probable than others, just given the range
-of things simulated more ways for a bigger positive effect. Also make
-clear though, biggish losses (10% or more) possible, but rare.

### 2022-10-06 R/compile_experiment_fleet.R

@@ -1,81 +1,85 @@
-compile_experiment_fleet <- function(fauna, state, tune_type = "explt",
-                                     effort_cost_exponent = 1.3) {
+compile_experiment_fleet <-
+  function(fauna,
+           state,
+           tune_type = "explt",
+           effort_cost_exponent = 1.3) {
+    if (all(state$spatial_q == TRUE)) {
+      tuna_spatial_q <-
+        state$habitat[state$scientific_name == "katsuwonus pelamis"][[1]] %>%
+        pivot_wider(names_from = y, values_from = habitat) %>%
+        select(-x) %>%
+        as.matrix()
 
+      marlin_spatial_q <-
+        state$habitat[state$scientific_name == "kajikia audax"][[1]] %>%
+        pivot_wider(names_from = y, values_from = habitat) %>%
+        select(-x) %>%
+        as.matrix()
 
-  if (all(state$spatial_q == TRUE)) {
+      shark_spatial_q <-
+        state$habitat[state$scientific_name == "carcharhinus longimanus"][[1]] %>%
+        pivot_wider(names_from = y, values_from = habitat) %>%
+        select(-x) %>%
+        as.matrix()
 
+    } else {
+      tuna_spatial_q <- NA
 
-    tuna_spatial_q <-  state$habitat[state$scientific_name =="katsuwonus pelamis"][[1]] %>%
-      pivot_wider(names_from = y, values_from = habitat) %>%
-      select(-x) %>%
-      as.matrix()
+      marlin_spatial_q <- NA
 
-    marlin_spatial_q <-  state$habitat[state$scientific_name =="kajikia audax"][[1]] %>%
-      pivot_wider(names_from = y, values_from = habitat) %>%
-      select(-x) %>%
-      as.matrix()
+      shark_spatial_q <- NA
 
-   shark_spatial_q <-  state$habitat[state$scientific_name =="carcharhinus longimanus"][[1]] %>%
-      pivot_wider(names_from = y, values_from = habitat) %>%
-      select(-x) %>%
-      as.matrix()
+    }
 
-  } else {
+    fleets <- list(
+      "longline" = create_fleet(
+        list(
+          "katsuwonus pelamis" = Metier$new(
+            critter = fauna$`katsuwonus pelamis`,
+            price = 1.36,
+            sel_form = "logistic",
+            sel_start = fauna$`katsuwonus pelamis`$length_50_mature *  sample(c(0.6), 1, replace = TRUE),
+            sel_delta = 0.01,
[…]

### 2022-10-06 R/compile_fleet.R

@@ -1,8 +1,6 @@
 compile_fleet <- function(fauna, tune_type = "explt",  effort_cost_exponent = 1.3,
                           spatial_allocation = "rpue") {
 
-  # browser()
-  # fauna <- wtf
 
   fleets <- list(
     "longline" = create_fleet(
@@ -49,17 +47,6 @@ compile_fleet <- function(fauna, tune_type = "explt",  effort_cost_exponent = 1.
           sel_unit = "length"
 
         ),
-        # "kajikia audax" = Metier$new(
-        #   critter = fauna$`kajikia audax`,
-        #   price = 5.16,
-        #   sel_form = "logistic",
-        #   sel_start = 175,
-        #   sel_delta = 0.01,
-        #   catchability = .1,
-        #   p_explt = .99,
-        #   sel_unit = "length"
-        #
-        # ),
         "carcharhinus longimanus" = Metier$new(
           critter = fauna$`carcharhinus longimanus`,
           price = 1.89,
@@ -71,28 +58,6 @@ compile_fleet <- function(fauna, tune_type = "explt",  effort_cost_exponent = 1.
           sel_unit = "length"
 
         ),
-        # "istiompax indica" = Metier$new(
-        #   critter = fauna$`istiompax indica`,
-        #   price = 2.74,
-        #   sel_form = "logistic",
-        #   sel_start = 100,
-        #   sel_delta = 0.01,
-        #   catchability = .1,
-        #   p_explt = .99,
-        #   sel_unit = "length"
-        #
-        # ),
-        # "makaira mazara" = Metier$new(
-        #   critter = fauna$`makaira mazara`,
-        #   price = 3.74,
-        #   sel_form = "logistic",
-        #   sel_start = 150,
-        #   sel_delta = 0.01,
-        #   catchability = .1,
-        #   p_explt = .95,
-        #   sel_unit = "length"
-        #
-        # ),
         "carcharhinus falciformis" = Metier$new(
           critter = fauna$`carcharhinus falciformis`,
           price = 2.16,
@@ -139,7 +104,8 @@ compile_fleet <- function(fauna, tune_type = "explt",  effort_cost_exponent = 1.
       ),
       effort_cost_exponent = effort_cost_exponent,
       base_effort = resolution^2,
-      spatial_allocation = spatial_allocation
+      spatial_allocation = spatial_allocation,
+      resolution = resolution
     ),
     "purseseine" = create_fleet(
       list(
[…]

### 2022-10-06 R/create_critters.R

@@ -56,7 +56,6 @@ create_critters <-
       scientific_name = sciname,
       base_habitat = hab,
       recruit_habitat = recruit_habitat,
-      adult_movement = 0,
       adult_diffusion = 4,
       recruit_diffusion = 2,
       rec_form = rec_form,

### 2022-10-06 R/optimize_mpa.R

@@ -72,8 +72,8 @@ optimize_mpa <-
           fauna = fauna,
           fleets = fleets,
           years = years,
-          mpas = list(locations = tmp_mpas,
-                      mpa_year = 1),
+          manager = list(mpas = list(locations = tmp_mpas,
+                      mpa_year = 1)),
           initial_conditions = starting_conditions
         )
 
@@ -175,8 +175,8 @@ optimize_mpa <-
         fauna = fauna,
         fleets = fleets,
         years = years,
-        mpas = list(locations = mpas,
-                    mpa_year = 1),
+        manager = list(mpas = list(locations = mpas,
+                    mpa_year = 1)),
         initial_conditions = starting_conditions
       )
 

### 2022-10-06 R/run_mpa_experiment.R

@@ -1,5 +1,6 @@
 run_mpa_experiment <-
   function(placement_strategy = "depletion",
+           fleet_model = "constant effort",
            starting_conditions,
            proc_starting_conditions,
            prop_mpa = 0.3,
@@ -10,16 +11,22 @@ run_mpa_experiment <-
            critters_considered = NA,
            random_mpas = FALSE,
            max_delta = 1,
-           resolution) {
+           resolution,
+           years = 50) {
+
+    options(dplyr.summarise.inform = FALSE)
+
 
     fleets <-  purrr::modify_in(fleets, list(1, "mpa_response"), ~ mpa_response)
+
+
+    fleets <-  purrr::modify_in(fleets, list(1, "fleet_model"), ~ fleet_model)
 
     n_mpa <- round(prop_mpa * resolution^2)
 
     # set up open access
 
-    # starting <- starting_conditions[[length(starting_conditions)]]
-
+    #
     if (is.na(critters_considered)){
       critters_considered <-  length(fauna)
     }
@@ -30,57 +37,80 @@ run_mpa_experiment <-
 
       # place MPAs in proportion to depletion-weighted spawning stock biomass
       depletion <-
-        (1 - (map_df(starting_conditions, ~ sum(.x$ssb_p_a) / .x$ssb0))) %>%  # depletion of each species
-        pivot_longer(everything(), names_to = "critter", values_to = "weight")
+        map_df(starting_conditions[1], ~ (1 - (map_df(.x, ~ sum(.x$ssb_p_a) / .x$ssb0))) %>%  # depletion of each species
+        pivot_longer(everything(), names_to = "critter", values_to = "weight"), .id = "step") %>%
+        select(-step)
 
       priorities <- proc_starting_conditions$fauna %>%
         left_join(depletion, by = "critter") %>%
         filter(critter %in% critters_considered) %>%
-        group_by(critter) %>%
+        group_by(critter, step) %>%
         mutate(patch_weight = ssb / sum(ssb) * weight) %>%
-        group_by(patch) %>%
+        group_by(patch, step) %>%
         summarise(patch_weight = sum(patch_weight)) %>%
         ungroup() %>%
         mutate(patch_weight = patch_weight * rlnorm(nrow(.), 0, placement_error)) %>%
         arrange(desc(patch_weight))
 
-      # browser()
-      # priorities %>%
-      #   ggplot(aes(patch_weight, patch_weight2)) +
[…]

### 2022-10-06 scripts/01_run_experiments.R

@@ -3,15 +3,18 @@ source(file.path("scripts", "00_setup.R"))
 
 # generate state experiments. This is a somewhat tricky process where the actual generated values are created for many variables in create_experiment_critters
 
-n_states <- 5
+n_states <- 2
+
 
 if (run_experiments) {
   state_experiments <-
     tibble(
-      sigma_centroid = runif(n_states, .1 * resolution, resolution ^ 2 / 2),
-      sigma_hab = runif(n_states, .1 * resolution, resolution ^ 2 / 8),
+      mpa_response = sample(c("stay","leave"), n_states, replace = TRUE),
+      sigma_centroid = runif(n_states, .1 * resolution, resolution ^ 2 / 20),
+      sigma_hab = runif(n_states, .1 * resolution, resolution ^ 2 / 20),
       spatial_q = sample(c(FALSE,TRUE), n_states, replace = TRUE, prob = c(3,1)),
-      spatial_allocation = sample(c("ppue","rpue", "revenue",'profit'), n_states, replace = TRUE)
+      spatial_allocation = sample(c("ppue","rpue"), n_states, replace = TRUE),
+      fleet_model = sample(c("open access", "constant effort"), n_states, replace = TRUE)
     ) %>%
     mutate(state_id = 1:nrow(.))
 
@@ -60,7 +63,7 @@ if (run_experiments) {
           mutate(c_x = base_layer$x[critters$centroid[i]],
                  c_y = base_layer$y[critters$centroid[i]]) %>%
           mutate(distance = sqrt((c_x - x) ^ 2 + (c_y - y) ^ 2)) %>%
-          mutate(habitat = dnorm(distance, 0, sigma_hab)) %>%
+          mutate(habitat = resolution * dnorm(distance, 0, sigma_hab)) %>%
           select(x, y, habitat)
 
         tmp$habitat <- tmp$habitat - min(tmp$habitat)
@@ -87,27 +90,29 @@ if (run_experiments) {
     ) %>%
     unnest(cols = habitats) %>%
     mutate(
-      seasonal_movement = sample(c(TRUE,FALSE), nrow(.), replace = TRUE),
+      seasonal_movement = sample(c(FALSE,TRUE), nrow(.), replace = TRUE),
+      spawning_aggregation = sample(c(TRUE,FALSE), nrow(.), replace = TRUE),
+      spawning_season = sample(1:seasons, nrow(.), replace = TRUE),
       f_v_m = sample(c(0.4,0.8,1.6), nrow(.), replace = TRUE),
[…]

### 2022-10-06 scripts/02_run_case_study.R

@@ -54,13 +54,10 @@ for (s in seq_along(unique(species))){
 }
 
 get_layer <- function(file) {
-  # file <- mats[3]
 
   file_comps <- str_split(file, "_", simplify = TRUE)
 
   species <- str_replace(file_comps[, 1], "-", " ")
-
-  # fleet <- file_comps[, 2]
 
 
   hab <-
@@ -274,7 +271,7 @@ if (run_casestudy == TRUE){
   case_study_experiments <-
     expand_grid(
       placement_strategy = c("depletion", "rate", "avoid_fishing", "target_fishing", "area"),
-      prop_mpa = seq(0,1, by = 0.01),
+      prop_mpa = seq(0,1, by = 0.33),
       critters_considered = length(fauna),
       placement_error = c(0),
       mpa_response = c("stay")
@@ -314,7 +311,7 @@ write_rds(case_study_experiments, file = file.path(results_path, "case_study_exp
 # optimize network --------------------------------------------------------
 if (optimize_casestudy == TRUE){
 
-  optimized_networks <- tibble(alpha = seq(0,1, length.out = 20))
+  optimized_networks <- tibble(alpha = seq(0,1, length.out = 5))
 
   # optimized_networks <- tibble(alpha = c(0,1))
 

### 2022-10-10 marlin-paper.qmd
adding deep water snapper to coral reef example

@@
 need to contrast with SEAPODYM
+
+<!-- Single column, double line spaced -->
+<!-- Within the word count (no more than 7000-8000 words for Standard Articles, 3000-4000 words for Applications and Practical Tools) -->
+<!-- Continuous line and page numbering throughout -->
+<!-- Clearly defined manuscript structure as standard: Author details, Abstract (must be numbered 1-4 according to Manuscript Specifications), Keywords, Introduction, Materials and Methods, Results, Discussion, Figures and Tables with captions -->
+<!-- Figures and Tables can be embedded within the text where referenced to facilitate reviewing -->
+<!--         Author Contributions (if >1 author – see Authorship Policy). -->
+<!--         Statement of where you intend to archive your data -->
 
@@
 | State                 | Distribution          | Description                                     |
-|-----------------------|-----------------------|-------------------------------------------------|
+|------------------|------------------|------------------------------------|
 | $\sigma_{centroid}$   |                       | Standard deviation of habitat centroid          |

### 2022-10-10 R/compile_fleet.R

@@ -1,5 +1,5 @@
-compile_fleet <- function(fauna, tune_type = "explt",  effort_cost_exponent = 1.3,
-                          spatial_allocation = "rpue") {
+compile_fleet <- function(fauna, tune_type = "explt",  effort_cost_exponent = 1,
+                          spatial_allocation = "ppue", fleet_model = "constant effort") {
 
 
   fleets <- list(
@@ -105,7 +105,8 @@ compile_fleet <- function(fauna, tune_type = "explt",  effort_cost_exponent = 1.
       effort_cost_exponent = effort_cost_exponent,
       base_effort = resolution^2,
       spatial_allocation = spatial_allocation,
-      resolution = resolution
+      resolution = resolution,
+      fleet_model = fleet_model
     ),
     "purseseine" = create_fleet(
       list(
@@ -208,6 +209,7 @@ compile_fleet <- function(fauna, tune_type = "explt",  effort_cost_exponent = 1.
           sel_unit = "length"
         )
       ),
+      fleet_model = fleet_model,
       effort_cost_exponent = effort_cost_exponent,
       base_effort = resolution^2,
       spatial_allocation = spatial_allocation,

### 2022-10-10 R/create_experiment_critters.R

@@ -70,7 +70,6 @@ create_experiment_critters <-
     if (ontogenetic_shift) {
       recruit_habitat <-
         -hab[[spawning_seasons[[1]]]] - min(-hab[[spawning_seasons[[1]]]]) # place recruits in different places than adults
-
       # set recruitment form to allow for recruit habitat
       #
       rec_form <- 0

### 2022-10-10 R/run_mpa_experiment.R

@@ -19,15 +19,19 @@ run_mpa_experiment <-
 
     fleets <-  purrr::modify_in(fleets, list(1, "mpa_response"), ~ mpa_response)
 
+    if (!is.na(fleet_model)) {
+      fleets <-
+        purrr::modify_in(fleets, list(1, "fleet_model"), ~ fleet_model)
 
-    fleets <-  purrr::modify_in(fleets, list(1, "fleet_model"), ~ fleet_model)
+
+    }
 
     n_mpa <- round(prop_mpa * resolution^2)
 
     # set up open access
 
     #
-    if (is.na(critters_considered)){
+    if (all(is.na(critters_considered))){
       critters_considered <-  length(fauna)
     }
 
@@ -191,10 +195,14 @@ run_mpa_experiment <-
     }
 
 
+    fleet_econ <- map_df(mpa_sim[[length(mpa_sim)]], ~ colSums(.x$prof_p_fl, na.rm = TRUE), .id = "critter")
+
     econ <-
       (map_df(mpa_sim[[length(mpa_sim)]], ~ sum(.x$prof_p_fl, na.rm = TRUE))) %>%
       pivot_longer(everything(), names_to = "critter",values_to = "econ")
     #  calculate econ component of objective function, currently revenues across all fleets and species
+
+    fleet_yield <- (map_df(mpa_sim[[length(mpa_sim)]], ~ colSums(.x$c_p_fl, na.rm = TRUE), .id = "critter"))
 
     yield <-
       (map_df(mpa_sim[[length(mpa_sim)]], ~ sum(.x$c_p_fl, na.rm = TRUE))) %>%
@@ -203,7 +211,8 @@ run_mpa_experiment <-
 
     objective_outcomes <- biodiv %>%
       left_join(econ, by = "critter") %>%
-      left_join(yield, by = "critter")
+      left_join(yield, by = "critter") %>%
+      left_join(fleet_yield, by = "critter")
 
     outcomes <- list()
 

### 2022-10-10 scripts/01_run_experiments.R

@@ -3,7 +3,7 @@ source(file.path("scripts", "00_setup.R"))
 
 # generate state experiments. This is a somewhat tricky process where the actual generated values are created for many variables in create_experiment_critters
 
-n_states <- 2
+n_states <- 250
 
 
 if (run_experiments) {
@@ -77,6 +77,8 @@ if (run_experiments) {
       return(critters)
     }
 
+  port_locations <- tibble(x = c(1, resolution), y = c(1, resolution)) # coordinates to test impact of highly disparate ports
+
 
   state_experiments <- state_experiments %>%
     mutate(
@@ -93,7 +95,7 @@ if (run_experiments) {
       seasonal_movement = sample(c(FALSE,TRUE), nrow(.), replace = TRUE),
       spawning_aggregation = sample(c(TRUE,FALSE), nrow(.), replace = TRUE),
       spawning_season = sample(1:seasons, nrow(.), replace = TRUE),
-      f_v_m = sample(c(0.4,0.8,1.6), nrow(.), replace = TRUE),
+      f_v_m = runif(nrow(.), 0.2,1.6),
       adult_diffusion = runif(
         nrow(.),
         min = 0,
@@ -104,10 +106,11 @@ if (run_experiments) {
         min = 0,
         max = resolution
       ),
-      hyper = sample(c(1,1.5),nrow(.), replace = TRUE),
+      hyper = runif(nrow(.), 1,2),
       density_dependence = sample(c("global_habitat", "local_habitat", "pre_dispersal", "post_dispersal"), nrow(.), replace = TRUE)
     ) %>%
-    mutate(spawning_season = ifelse(spawning_aggregation, spawning_season, NA))
+    mutate(spawning_season = ifelse(spawning_aggregation, spawning_season, NA),
+           ontogenetic_shift = sample(c(FALSE,TRUE), nrow(.), replace = TRUE))
 
   state_experiments$habitat[[sample(1:n_states, 1)]] %>%
     ggplot(aes(x, y, fill = habitat)) +
@@ -127,7 +130,8 @@ if (run_experiments) {
           adult_diffusion = adult_diffusion,
           recruit_diffusion = recruit_diffusion,
           density_dependence = density_dependence,
-          hyper = hyper
+          hyper = hyper,
+          ontogenetic_shift = ontogenetic_shift
         ),
[…]

### 2022-10-12 Untitled.qmd
v0.1

---
title: "Untitled"
format: pdf
editor: visual
crossref:
  fig-prefix: Fig.
  tbl-prefix: Tbl
---

## Quarto

Quarto enables you to weave together content and executable code into a
finished document. To learn more about Quarto see <https://quarto.org>.

## Running Code

When you click the **Render** button a document will be generated that
includes both content and the output of embedded code. You can embed
code like this:

```{r}
1 + 1
```

You can add options to executable code like this, see @fig-test

```{r}
#| label: fig-test
#| fig-cap: "Hello"
plot(1:10)
```

The `echo: false` option disables the printing of code (only output is
displayed).

### 2022-10-12 marlin-paper.qmd
v0.1

--- 2022-10-06 marlin-paper.qmd
@@
 ---
-title: "Efficient Simulation of Economic and Ecological Dynamics for Marine Spatial Planning"
+title: "Efficient Simulation of Economic and Ecological Dynamics in Aquatic Systems"
 author:
@@
     - Darcy Bradley
+    - Echelle Burns
     - Lennon Thomas
-    - Echelle Burns
+    - James Thorson
 date: "`r Sys.Date()`"
@@
     toc: true
+    fig-pos: 'H'
   html: default
@@
 params:
-  params:
-  run_name: ["v1.1"]
+  run_name: ["v1.2"]
+abstract: "Very rough cut.
+Communities around the world are increasingly looking to spatial strategies for managing marine ecosystems.
+However, the design of spatial management strategies faces numerous challenges.
+Marine ecosystems are highly dynamic and opaque, and extractive entities such as fishing fleets respond to changes in marine ecosystems in highly complex ways.
+We present a framework for efficient simulation of the bio-economic dynamics of marine systems, designed to support both management and research.
+We demonstrate the model's capabilities by focusing on three case studies on the conservation and food security impacts of marine protected areas (MPAs)."
+header-includes:
+  - \usepackage{setspace}\doublespacing
+  - \usepackage{lineno}\linenumbers
+
 ---
@@
 #| message: false
+#| include: false
 
@@
 
-need to contrast with SEAPODYM
+<!-- Single column, double line spaced -->
+
+<!-- Within the word count (no more than 7000-8000 words for Standard Articles, 3000-4000 words for Applications and Practical Tools) -->
+
+<!-- Continuous line and page numbering throughout -->
+
+<!-- Clearly defined manuscript structure as standard: Author details, Abstract (must be numbered 1-4 according to Manuscript Specifications), Keywords, Introduction, Materials and Methods, Results, Discussion, Figures and Tables with captions -->
+
+<!-- Figures and Tables can be embedded within the text where referenced to facilitate reviewing -->
+
+<!--         Author Contributions (if >1 author – see Authorship Policy). -->
+
+<!--         Statement of where you intend to archive your data -->
 
@@
 Many models have been developed to support the theory and design of
-spatial management strategies [@fulton2015]. However, they have tended
+spatial management strategies (see @fulton2015 for a review of existing methods). However, they have tended
 to fall into one of either highly complex and tactical models designed
@@
 to stakeholders charged with marine spatial planning exercises lacking
-accesible models suited for more tactical applications, and a barrier to
-the scientific community, wherein comparisons across spatial management
-simulation studies are clouded by discrepancies in underlying model
-structures beyond differences in the core phenomenon in question.
+accessible models suited for more tactical applications, and a barrier
+to the scientific community, wherein comparisons across spatial
+management simulation studies are clouded by discrepancies in underlying
+model structures beyond differences in the core phenomenon in question.
 
@@
 core methods and functionality of this framework, and demonstrate its
-use in two xx case-study applications. In the first, we show how
+use in three xx case-study applications. In the first, we show how
 `marlin` can be used to assess critical drivers for positive outcomes
@@
 
--   MPAs increasingly part of plan
-
--   A model based field - hard to do experiments at the time and scale
-    needed
-
--   many models out there, but challenges
-
-    -   highly stylized
-
-    -   highly tactical
-
--   We present a general tool for modeling spatial dynamics of fished
-    populations and fishing fleets, with a focus on MPAs
-
--   We present two case studies showing the utility of this approach
-
-    -   habitat heterogeneity and bycatch
-
-    -   MPA optimization
-
--   Basis for improved "what if" testing of MPA scenario
-
 ## Methods
 
-Models of MPAs span a wide range of complexity, from simple analytical
-tools to end-to-end ecosystem models [@fulton2015]. `marlin` follows a
-"models of intermediate complexity" (MICE) philosophy [@plagányi2014]
-targeted specifically around the role of technical interactions across
-species. Many MPA models simulate the impacts of MPAs on a
-single-species level [@ovando2021a]. This can be useful for for example
-examining the ways in which species-specific life history traits
-interact with fleet dynamics to influence MPA outcomes. However, bycatch
-problems are driven by technical interactions, in which fishing effort
-for a target species incidentally captures another species. `marlin`
-accounts for this by allowing the user to simulate an arbitrary number
-of fleets capable of catching an arbitrary number of species in a
-spatial area. Different fleets can place different economic value on
-different species, while also having species-specific selectivity
-functions. This allows us to simulate bycatch dynamics of MPAs in
-pelagic ecosystems.
+`marlin` follows a "models of intermediate complexity" (MICE) philosophy [@plagányi2014]
+targeted specifically towards capturing heterogeneous habitats, movement dynamics, and technical interactions across species.
+In its general form, `marlin` simulates the behavior of populations of spatially explicit biologically independent age-structured animal populations affected by one or more fishing fleets in time steps specified by the user (generally annual or finer).
+While `marlin`
+does not simulate species interactions, fishing fleets in `marlin` are
+capable of targeting and affecting multiple species simultaneously.
+This allows the model to simulate processes such as fisheries bycatch or effort displacement in a way that accounts for the ways that changes in fleet behavior might affect multiple species in the system.
+The ability to track the impacts of fishing fleets across multiple species simultaneously is particularly important as very few fishing fleets are truly single species.
 
@@
 the generalized simulation exercises and case study analyses.
+We note that `marlin` is a complex package with many options to control detailed behaviors.
+Given that these specific features will evolve over time, we do not describe in detail every option of the model here.
+Rather, we focus on the equations and methods describing the core foundations of our framework for efficiently simulating bio-economic dynamics in aquatic systems.
 
@@
 
-The key feature of `marlin` for the purposes of this project is the
-ability to capture technical interactions across species. While marlin
-does not simulate species interactions, fishing fleets in `marlin` are
-capable of targeting and affecting multiple species simultaneously. This
-allows us to capture for example the potential impacts of displacement
-of effort from one area into another across multiple species.
-
-A fishing fleet or *metier* is defined in `marlin` by a total pool of
+A fishing fleet is defined in `marlin` by a total pool of
 effort that can have different catchability, prices, and selectivity for
-individual species within a simulation. For example, both a longline and
+individual species within a simulation (each of which we denote as a *metier*. For example, both a longline and
 purse seine fleet may capture bigeye (*Thunnus obesus*) and skipjack
-(*Katswamus pelamus*), but the longline fleet is much more likely to
+(*Katswamus pelamus*) tunas, but the longline fleet is much more likely to
 catch bigeye than skipjack, and *vice versa*. Each fishing fleet
@@
 $$
- R_{t,p,f} = \sum_s^S \sum_a^A \Pi_{s,f}\frac{\alpha_{a,s,f}q_{s,f}E_{t,p,f}}{F_{t,p,s,a}} \times Y_{t,p,s,a}
-$$
+ R_{t,p,f} = \sum_s^S \sum_a^A \Pi_{s,f}\frac{\alpha_{a,s,f}q_{s,f}E_{t,p,f}}{u_{t,p,s,a}} \times Y_{t,p,s,a}
+$$ {#eq-revenue}
 
 where *p* is patch, *s* is species, *a* is age, $\alpha$ is selectivity
-at age, *q* is catchability , *F* is total fishing mortality, $\Pi$ is
+at age, *q* is catchability , *u* is total fishing mortality, $\Pi$ is
 price per unit weight, and *Y* is total catches.
@@
 $$
-Y_{t,p,s,a} = \frac{F_{t,p,s,a}}{z_{t,p,s,a}} \times b_{t,p,s,a} \times (1 - e^{-z_{t,p,s,a}}) 
-$$
+Y_{t,p,s,a} = \frac{u_{t,p,s,a}}{z_{t,p,s,a}} \times b_{t,p,s,a} \times (1 - e^{-z_{t,p,s,a}})
+$$ {#eq-yield}
 
-where *F* is total fishing mortality, given by
+where *b* is biomass and *u* is total fishing mortality, given by
 
 $$
-F_{t,p,s,a} = \sum_f^F\alpha_{a,s,f}q_{s,f}E_{t,p,f}
-$$
+u_{t,p,s,a} = \sum_{f=1}^F\alpha_{a,s,f}q_{s,f}E_{t,p,f}
+$$ {#eq-u}
 
@@
 
-```{=tex}
-\begin{equation}
+$$
 \alpha_{a,s,f}=\frac{1}{(1 + e^{-log(19)\times\frac{l_{a,s} - l^{sel}_{s,f}}{\delta^{sel}_{s,f}}})}
-  (\#eq:sel)
-\end{equation}
-```
+$$ {#eq-selectivity}
+
 where $l$ is length at age, $l^{sel}$ is the length at 50% selectivity,
@@
 
+or through a dome shaped form of
+
+XX add dome shaped equation XX
+
 *z* is total mortality given by
@@
 $$
- z_{t,p,s,a} = F_{t,p,s,a} + m_{s,a}
-$$
+ z_{t,p,s,a} = u_{t,p,s,a} + m_{s,a}
+$$ {#eq-z}
 
@@
 time step in a patch, individual fractions of which are portioned off as
-catches and subsequently revenues for each fleet. Given revenues, we
+catches and subsequently revenues for each fleet. Given revenues *R*, we
 then calculate profits $\phi$ per
@@
 $$
-\phi_{t,p,f} = R_{t,p,f} - c_f(E_{t,p,f} + 1)_{t,p,f}
-^{1.3}
+\phi_{t,p,f} = R_{t,p,f} - C_{t,p,f}
+$$ {#eq-profits}
+
+where costs *C* are calculated
+
+$$C_{t,p,f} = \gamma_f  \left( E_{t,p,f}^{\beta_f} + \eta_{f,p} E_{t,p,f} \right)$$
+
+where $$\gamma_f$$ the scales the cost per unit effort for fleet *f*, $\beta$ allows for the cost of effort to scale non-linearly, and $\eta$ is an additional cost per unit effort of fishing in each patch *p*, allowing for the model to account for travel costs for different patches.
+Travel costs are calculated based on the euclidean distance of each patch to the nearest port of a given fleet, users can specify any number from zero to the number of patches of port locations.
+When no ports are specified, travel costs are zero.
+
+#### Calculating Total Effort
+
+`marlin` allows for two general modes of effort. The simplest is "constant effort", in which the total effort of each fishing fleet remains constant over time.
+
+The more complex option is "open access".
+Under open access, the total amount of effort in a given time step for fleet *f* is a function of the profitability of that fleet in the proceeding time steps in which fishing was open.
+
 $$
+E_{t+1,f} = E_{t,f} \times e^{\theta_f log(R_{t,f} / C_{t,f})} 
+$$ {#eq-oa}
 
-where *c* is cost per unit effort and, the 1.3 exponent is in the manner
-of @costello2016a . The addition of one to the total effort is to
-constrain the value to be greater than 1, preventing undesired behavior
-where fishing effort actually gets less expensive if total effort is
-less than 1.
+There $\theta$ controls the responsiveness of fleet *f* to the log of the ratio of total revenues *R* to total costs *C*, and approximates the percentage amount effort changes in response to a one unit changes in the log revenue to cost ratio.
 
-The important feature here then is that the revenues and profits for a
-given fleet are a function of the prices and catchability of individual
-species, but that mortality itself is independent of price. So, a fleet
-that has selectivity and catchability greater than zero can exert
-fishing mortality on a species even if it obtains no value from that
-species.
+Three general exceptions exist for the effort dynamics.
+When MPAs are implemented, effort that used to operate inside the MPAs can either leave the fishery, or be redistributed outside the the MPA (the default behavior).
+The model can also accept total catch quotas for each species in each time step.
+When quotas are activated, if the total catch across all fleets for a given species under the model's effort dynamics would exceed the allowable quota for that species, the total effort for each fleet is decrease proportional to its contribution to the total catch until the quota is satisfied.
+`marlin` also allows users to specify closed fishing seasons for one or more species in the system.
+During closed seasons, no effort is allocated to patches that would result in capture of the covered species, potentially resulting in zero effort if the species cannot be avoided.
+Under these conditions, the amount of effort in the next open season is a function of the effort and profits in the last open fishing season.
 
-Each fishing fleet then decides how to allocate its effort in space
+#### Tuning Fishing Fleets
+
+The degree of fishing pressure exerted by a given set of fishing fleets on each species is a function of a range of parameters including the total amount of effort, the selectivity ogives, cost functions, the spatial distribution of the resource, the relative prices across species, and the catchability coefficient of each fleet for each species.
+`marlin` provides a tuning option to help users achieve desired biological outcomes from their fleets.
+
+Users can tune their fleets in one of two ways.
+First, they can specify a target exploitation rate *u* for each species in their model run.
+Taking all the other parameters of the model as given, `marlin` then adjusts the catchability coefficients *q_{f,s}* for each fleet *f* and species *s* such that the equilibrium exploitation rate for each species matches the desired level.
+Second, they can specify a target depletion, defined as total spawning stock biomass under fishing divided by total unfished spawning stock biomass, and the model will adjust the catchability coefficients *q_{f,s}* for each fleet *f* and species *s* such that the equilibrium depletion for each species matches the desired level.
+
+#### Spatial Allocation of Effort
+
+Each fishing fleet decides how to allocate its effort in space
 based on one of four possible spatial allocation strategies. The ideal
@@
 fleets in spaces (see @gillis2003 and references therein). However, the
-IFD presents a number of complications for our study. First, in order to
-simplify the assessment of the impacts of different fishing mortality
-rates we simulate scenarios with a fixed total pool of effort. As MPAs
-change the spatial distribution of profits though, the total amount of
-effort required to satisfy the IFD may not equal the total amount of
+IFD presents a number of complications for our model. First, under the constant effort model, the total amount of effort required to satisfy the IFD may not equal the total amount of
 effort required by the model. Second, even if we were to implement the
@@
 
-```{=tex}
-\begin{equation}
-n_{t,p,s,a}=\begin{cases}
+$$
+n_{t,p,s,a} = \begin{cases}
       = BH(ssb_{t-1,p,s,a}) & \text{if $a = 1$}\\
@@
   \end{cases}
-  (\#eq:pop)
-\end{equation}
-```
+$$ {#eq-pop}
+
 where *BH* is the Beverton-Holt recruitment function, *ssb* is
@@
 
-```{=tex}
-\begin{equation}
-  l_{a,s} = l\infty_s\left(1 - e^{-k_s(a - a0_{s})}\right)
-  (\#eq:len)
-\end{equation}
-```
+$$
+  l_{a,s} = l_{\infty,s}\left(1 - e^{-k_s(a - a0_{s})}\right)
+$$ {#eq-len}
+
 Biomass *b* at age is then given by
 
-```{=tex}
-\begin{equation}
+$$
   b_{a,s} = wa_s \times l_{a,s}^{wb_s}
-  (\#eq:weight)
-\end{equation}
-```
+$$ {#eq-weight}
+
 and maturity *mat* is calculated as
 
-```{=tex}
-\begin{equation}
+$$
   mat_{a,s} = \frac{1}{(1 + e^{-log(19)\times\frac{l_{a,s} - lmat_s}{{\delta}mat_s}}}
-  (\#eq:mat)
-\end{equation}
-```
+$$ {#eq-mat)}
+
 where $l_{mat}$ is the length at which on average 50% of individuals are
@@
 
-```{=tex}
-\begin{equation}
+$$
   ssb_{t,p,s} = \sum_{a=1}^Aw_{a,s,t}^{\gamma_s} mat_{t,a} n_{t,p,s,a}
-    (\#eq:ssb)
-\end{equation}
-```
-#### Recruitment {.unnumbered}
+$$ {#eq-ssb}
+
+#### Recruitment
 
@@
 
-```{=tex}
-\begin{equation}
+$$
 n_{t,p,s,a = 1} = \left(\frac{0.8{\times}\sum_{p=1}^P{r0_{p}}\times{h}\times\sum_{p=1}^P{ssb_{t-1,p}}}{0.2\times{\sum_{p=1}^Pssb0_p}\times(1 - h)+(h - 0.2)\times{\sum_{p=1}^Pssb_{t-1,p}}}\right)\times {hab_p}
-  (\#eq:dd2)
-\end{equation}
-```
+$$ {#eq-dd2}
+
 where *hab* is a vector of recruitment habitat quality by patch degined
@@
 
-$$hab_p = r0_p / \sum_p^P{r0_p}$$ 2. Local density dependence: Density
+$$hab_p = r0_p / \sum_p^P{r0_p}$$
+
+2. Local density dependence: Density
 dependence occurs independently in each patch and recruits are retained
@@
 
-```{=tex}
-\begin{equation}
+$$
 n_{t,p,s,a = 1,p} = \left(\frac{0.8{\times}r0_{s,p}\times{h_s}\times{ssb_{t-1,p,s}}}{0.2\times{ssb0_{p,s}}\times(1 - h_s)+(h_s - 0.2)\times{ssb_{t-1,p}}}\right)
-  (\#eq:dd1)
-\end{equation}
-```
+$$ {#eq-dd1}
+
 3.  Local density dependence then disperse: Density dependence occurs
@@
 
-```{=tex}
-\begin{equation}
+$$
   n_{t,p,s,a = 1,p} = \left(\frac{0.8{\times}r0_{s,p}\times{h_s}\times{ssb_{t-1,p,s}}}{0.2\times{ssb0_{p,s}}\times(1 - h_s)+(h_s - 0.2)\times{ssb_{t-1,p}}}\right)\times \boldsymbol{d^l} \times \epsilon_t
-  (\#eq:dd3)
-\end{equation}
-```
+$$ {#eq-dd3}
+
 where **d^l^** is the recruitment movement matrix
@@
 
-```{=tex}
-\begin{equation}
+$$
   larv_{t,p} = ssb_{t-1} \times\boldsymbol{d^l}
-  (\#eq:larvmove)
-\end{equation}
-```
-```{=tex}
-\begin{equation}
+$$ {#eq-larvmove}
+
+$$
 n_{t,a = 1,p} = \left(\frac{0.8{\times}r0_{p}\times{h}\times{larv_{t,p}}}{0.2\times{ssb0_p}\times(1 - h)+(h - 0.2)\times{larv_{t,p}}}\right) \times \epsilon_t
-  (\#eq:dd4)
-\end{equation}
-```
+$$ {#eq-dd4}
+
 #### Dispersal
@@
 `marlin`'s movement dynamics are based on a continuous-time Markov
-chain, as described in @thorson2021a. Within this framework, the model
+chain (CTMC), as described in @thorson2021a. Within this framework, the model
 allows for movement to be broken down into three components of
@@
 towards say a preferred habitat).
+For now, our model focuses just on the diffusion and taxis components of this model, though future extensions could easily incorporate advection vectors from say ROMs models (citations).
+In this way then, `marlin` allows users to run anything from a simple Gaussian dispersal kernel up to a system governed by species that
+passively diffuse around a dynamic thermal range.
 
-xx movement equations xx
+@thorson2021a described the exact steps of the CTMC approach in detail.
+We briefly the general CTMC method here, but see @thorson2021a for a detailed description of the method.
+Under this framework, movement from each patch to each other patch in the system in a given timestep *t* for life stage *g* of species *s* is defined by a movement matrix $\boldsymbol{M}_{t,s,g}$.
+$\boldsymbol{M}_{t,s,g}$ is calculated as the matrix exponent of the sum of the diffusion and taxis matrices scaled by the width of the time step that the model is running under.
 
-While this approach to movement simulation is not commonly seen in the
+$$
+\boldsymbol{M}_{t,s,g} = e^{\Delta t(\boldsymbol{D}_{t,s,g} + \boldsymbol{X}_{t,s,g})} 
+$$ {#eq-movement}
+
+Where $\boldsymbol{D}$ is the diffusion matrix and $\boldsymbol{X}$ is the taxis matrix.
+
+The diffusion matrix is a function of a user specified diffusion rate, see @thorson2021a. We slightly modify the taxis matrix in the following manner
+
+xx need to fix notation somewhere here, see this line of code for much clearer depiction of what's happening, since sometimes code is cleaner.
+https://github.com/DanOvando/marlin/blob/3563e7319ad0267f89789267a4c6ee58a9ad0d60/R/fish_class.R#L525 xx
+
+$$
+n_{t,p,s,a} = \begin{cases}
+      = (1 + \lambda d_{s,f})max(0, h(p2_{t,s,g}) - h(p1_{t,s,g})) & \text{if p2 and p1 are adjacent}\\
+     = -\sum_{g' \neq g1}\{ (1 + \lambda d_{s,f})max(0, h(p2_{t,s,g}) - h(p1_{t,s,g}))\} & \text{if p1 = p2}\\
+     = 0 & \text{otherwise.}
+  \end{cases}
+$$ {#eq-taxis}
+
+The $lambda d_{s,f}$ term is a scalar in which the magnitude of the habitat gradient between patches can be scaled relative to the diffusion parameter $d$.
+
+For both the diffusion and taxis matrix, we allow for the inclusion of physical barriers to movement (i.e.
+land).
+Pairs of patches that are adjacent but in which one or both patches are a barrier to movement are set as non-adjacent.
+The CTMC model then produces movement dynamics that move around barriers rather than over them.
+This is different than setting habitat in a patch to zero, which implies a patch that a given animal will not move towards for its own sake, but can move over towards another better patch.
+
+While this CTMC approach to movement simulation is to date not commonly seen in the
 marine modeling literature, it has numerous advantages that warrant its
 seeming complexity. First, when only diffusion is present, the model
-will generalize to the familiar dynamics of a Gaussian dispersal kernel
-xx check this xx. However, the time component *t* in equation xx allows
-the movement model to be run on arbitrary time steps while preserving
-its behavior.
-
-Second, the *advection* component allows users to input useful features
-such as the outputs of ROMs models.xxx
-
-Third, the *taxis* model allow for more active habitat choices by
+will generalize to the familiar dynamics of a Gaussian dispersal kernel.
+However, the time component *t* in @eq-movement allows the movement model to be run on arbitrarily fine time steps.
+Second, the *taxis* model allow for more active habitat choices by
 species, allowing us to simulate both preferences of species in space
-and time efficiently.
-
-In this way then, `marlin` allows users to run anything from a simple
-Gaussian dispersal kernel up to a system governed by species that
-passively diffuse around a dynamic thermal range
-
-Lastly, this form has the advantage that the parameters are directly
+and time efficiently.Lastly, this form has the advantage that the parameters are directly
 estimable from real data. So, if provided with say CPUE data and a
 tagging study, users can estimates the movement parameters in the same
-manner as @thorson2021, and then pass their estimated parameters to
+manner as @thorson2021a, and then pass their estimated parameters to
 `marlin` for simulation.
 
-### Rule-of-thumb MPA Design Strategies
+#### Reference Points
 
-MPA design is governed by two general parameters: the number of cells to
-place inside an MPAs, and the design strategy used to place the MPAs.
+Sigh. I need to work on this section, but I am *le tired*.
 
-We consider `r n_distinct(experiment_obj$placement_strategy)` different
-MPA placement strategies.
-
-1.  Depletion: MPA locations are prioritized in proportion to the
-    spawning stock biomass of the most depleted species pre-MPA. So,
-    patches with high spawning biomass of highly depleted species are
-    prioritized over patches with high spawning biomass of lightly
-    depleted species.
-
-2.  Rate: MPAs are placed in based on the pre-MPA depletion-weighted
-    catch relative to the total catch in a patch. So, patches with high
-    rates of catch of depleted species relative to total catch are
-    prioritized.
-
-3.  Avoid_fishing: MPAs locations are prioritized inversely to fishery
-    catches. So, patches with low catches are prioritized over patches
-    with high catches.
-
-4.  Target_fishing: MPAs locations are prioritized proportionally to
-    fishery catches. So, patches with high total catches are prioritized
-    over patches with low catches.
-
-5.  Area: For the "area" strategy, we simply assign patches in
-    sequential order until the desired total MPA size is reached.
-
-<!-- The information used to inform each MPA design strategy (except area) -->
-
-<!-- can include a multiplicative observation error term $\sigma_{obs}$ where -->
-
-<!-- $$information \sim lnorm(data,\sigma_{obs}) $$ -->
-
-<!-- For example, for the depletion scenario, $\sigma_{obs}$ controls the -->
-
-<!-- accuracy of the spatial depletion data used to prioritize MPA placement. -->
-
-## Reference Points
-
-A note here WHy no reference points? The problem: Fmsy and MSY are
+A note here Why no reference points? The problem: Fmsy and MSY are
 selectivity dependent, and once we allow for different fleets with
@@
 
-SSBmsy is slightly more conserved, as over the long term the key things
-is the flow of recruits, so we do provide an option for that
+SSBmsy is more conserved, as over the long term the key things
+is the flow of recruits, so we do provide an option for that, but I haven't messed with it for a while.
 
 There are many different strategies for estimating reference points
-conditional on selectivity and regimes, see references. We leave it to
+conditional on selectivity and regimes, see citations. We leave it to
 users to find relevant reference point as required by their methods on
@@
 
-## MPA Optimization
+### MPA Optimization
 
@@
 
-$$obj_p = \Phi \times scale(\sum_f^F \boldsymbol{\phi_{f,p}}) + (1 - \Phi)  scale(\boldsymbol{\sum_s^S\frac{ssb_s}{ssb0_{s,p}}})$$
+$$
+obj_p = \Phi \times scale(\sum_f^F \boldsymbol{\phi_{f,p}}) + (1 - \Phi)  scale(\boldsymbol{\sum_s^S\frac{ssb_s}{ssb0_{s,p}}})
+$$ {#eq-obj}
 
@@
 
-## Generalized Experiments
+### Use Cases
 
-Two features of marine ecosystems that are often omitted from
-theoretical and simulation-based considerations of marine spatial
-planning are *technical interactions* and *species on the move*. May
-canonical MPA models, informed by planning efforts centered around
-relatively static features such as coral reefs, have assumed that
-species inhabit a fixed and often uniform (e.g. @hastings2017). In
-reality, the habitat of many species are heterogeneous in space, and can
-experience both seasonal and directional changes, e.g. climate-change
-mediated range shifts.
+We include three examples demonstrating how `marlin` can be used to
+support marine spatial planning. In the first, we show how `marlin` can
+be used to compare the total and distributional impacts of marine
+protected areas design in a heavily fished coral reef ecosystem. In the
+second, we demonstrate how `marlin` can be used to assess components of
+climate resilience of alternative MPA design strategies. In the third,
+we show how `marlin` can be used to run large-scale simulation testing
+of management strategies.
 
-This reality creates a basic challenge to MPA design, in that given the
-need to protect an area less than the total range of a species, the more
-complex and dynamics the habitat of the species the more challenging it
-will be to identify an area that provides effective protection today and
-tomorrow. However, *technical interactions* interact with this spatial
-complexity to create another and potentially more difficult challenge.
-One of the key questions around spatial management strategies such as
-MPAs is to what extent do they *remove* vs *displace* threats.
+#### Rule-of-thumb MPA Design Strategies
 
-The potential benefits of MPAs as a bycatch management are complicated
-by two key issues: effort displacement and technical interactions. One
-of the primary questions around MPAs is to what extent do MPAs remove
-versus displace fishing effort. If MPAs remove fishing effort, then
-their direct effects on target and bycatch species can only be positive.
-If however MPAs displace fishing effort, evaluating their potential
-impacts becomes more complicated. The net conservation impact of MPAs
-becomes a balancing act between the effects of reduced fishing inside
-protected borders and increased fishing intensity outside MPAs. The
-outcomes of this displacement can be further complicated by technical
-interactions, i.e. the degree to which fishing effort for one species
-affects another, particularly when species have markedly different
-habitat distributions. For example, @abbott2012 demonstrated that an MPA
-designed to protect red king crab (*Paralithodes camtschaticus*) caused
-an unintended increase in bycatch of Pacific halibut (*Hippoglossus
-stenolepis*) by the displaced bottom-trawl fishery.
+Across all of our case studies we make use of four potential "rule of
+thumb" MPA design strategies:
 
-Given the complexity of the operating model used here, a complete
-sensitivity analysis across all parameters is not practical. Instead, we
-focus on the role of life history, fishing pressure, and MPA design.
+1.  Rate: MPAs are placed in based on the pre-MPA depletion-weighted
+    catch relative to the total catch in a patch. So, patches with high
+    rates of catch of depleted species relative to total catch are
+    prioritized.
 
-Each iteration uses one of three species: A stylized tuna modeled after
-skipjack tuna (), a stylized marlin modeled after (), and a stylized
-shark modeled after (). Each species has all life history parameters
+2.  Avoid_fishing: MPAs locations are prioritized inversely to fishery
+    catches. So, patches with low catches are prioritized over patches
+    with high catches.
+
+3.  Target_fishing: MPAs locations are prioritized proportionally to
+    fishery catches. So, patches with high total catches are prioritized
+    over patches with low catches.
+
+4.  Area: For the "area" strategy, we simply assign patches in
+    sequential order until the desired total MPA size is reached.
+
+MPAs never to our knowledge randomly placed; their locations are decided
+based on stakeholder negotiation based on available information and
+objectives. In theory the design MPA networks can be optimized through
+the use of a modeling framework, and indeed conditional on the validity
+of the model itself this process is likely to produce better outcomes
+than manually-designed strategies [@rassweiler2012; @rassweiler2014].
+However, in practice the ability of communities to conduct numerical
+optimization of MPA networks can be constrained by the number of
+objectives and dimensionality of the operating model, or by simple
+capacity of local stakeholders. As such, we show here how `marlin` can
+be used to examine the performance of potentially sub-optimal but more
+broadly usable design strategies xx.
+
+#### Coral Reefs
+
+In our coral reef example, we model the dynamics of four species: a
+grouper (XX), a shallow-reef snapper (xx), a deep-water snapper (XX),
+and a reef shark (XX). Groupers undergo a mass migration to a spawning
+aggregation once per year, followed by the sharks. Shallow-reef snappers
+stay in reefs closer to shower year-round, while deep-water snappers
+stay in the deeper reefs (@fig-coral).
+
+These species are targeted by two different fleets. Fleet one primarily
+targets the grouper and nearshore snapper populations, while obtaining
+some funds from sharks. Fleet one has a dome shaped selectivity pattern
+for both grouper and shallow-reef snapper due to demand for "plate
+sized" fish (citation), and logistic selectivity for sharks, as any
+shark caught is retained. This fleet is a subsistence fleet,
+where the local community takes every possible opportunity to fish that
+they can, and as such we model it as a "constant effort" fishery.
+
+The second fleet primarily targets the deep-water snapper population.
+While plate-sized snapper are still prized from this population, we
+model fleet two's selectivity as logistic due to high levels of discard
+mortality for deep-water snapper resulting from barotrauma (xx). This
+fleet catches some grouper, though less than fleet one, and receives no price for
+sharks due to the requirements of a certification program through which
+they sell their deep-water snapper, though they do accidentally catch and kill them on occasion. This fleet operates under open-access
+dynamics, as fishing is not the primary means of subsistence and so
+short-term effort expands and contracts in respond to profitability of
+the primarily grouper-driven fishery xx.
+
+```{r}
+#| label: fig-coral
+#| fig-cap: "Distribution of animals each season under unfished conditions. Numbers show the approximate location of each fleet's port."
+
+coral_fauna <- coral_fauna_and_fleets$fauna
+
+coral_fleets <- coral_fauna_and_fleets$fleets
+
+grid <- expand_grid(x = 1:resolution, y= 1:resolution) %>%
+  mutate(patch = 1:nrow(.))
+
+mpa_locations <- expand_grid(x = 1:resolution, y = 1:resolution) %>%
+  mutate(mpa = TRUE)
+
+coral_sim <- simmar(
+  fauna = coral_fauna,
+  fleets = coral_fleets,
+  manager = list(mpas = list(
+    locations = mpa_locations,
+    mpa_year = 1
+  )),
+  years = 2
+)
+
+patch_biomass <-
+  map_df(coral_sim, ~ map_df(.x, ~tibble(biomass = rowSums(.x$ssb_p_a), patch = 1:nrow(.x$ssb_p_a)), .id = "critter"), .id = "step") %>%
+  mutate(step = as.numeric(step)) %>%
+  left_join(grid, by = "patch")
+
+patch_biomass %>%
+  group_by(critter, step) %>%
+  mutate(sbiomass = biomass / max(biomass)) %>%
+  ungroup() %>%
+  filter(between(step,1,2)) %>% 
+  mutate(Season = step) %>% 
+  ggplot()+
+  geom_tile(aes(x,y, fill = sbiomass),show.legend = FALSE) +
+  geom_text_repel(data = coral_ports, aes(x, y, label = fleet), color = "red") +
+  facet_grid(Season~critter) +
+  scale_x_continuous(expand = c(0,0)) +
+  scale_y_continuous(expand = c(0,0)) +
+  scale_fill_viridis_c() + 
+  theme(axis.text = element_blank(),
+        axis.ticks = element_blank(),
+        axis.title = element_blank())
+
+```
+
+We then simulate the impacts of MPAs...
+
+#### Pelagic Systems
+
+We modeled our pelagic use case loosely on the characteristics of the
+Western and Central Pacific Ocean tuna fisheries. We simulate
+trajectories of XX species commonly caughts in the region, including
+both the highly abundance skipjack tuna (*Katsuwonus pelamis*) to the
+heavily depleted oceanic whitetip shark (*Carcharhinus longimanus*). We
+pulled publicly available data on catch-per-unit-effort of each of these
+species from the WCPO as a proxy for baseline habitat distributions
+(Fig.XX). However, for the pelagic case study we simulate a scenario
+where climate change is inducing a northerly migration of XX degress per
+year in XX species.
+
+These species are targeted by a longline fleet that targets large adult
+tunas such as bigeye and yellowfin primarily for high-grade consumption,
+and a purse-seine fishery that primarily targets skipjack tunas for bulk
+canning. XX. Both fleets are subject to a quota XX add this back in XX.
+
+```{r}
+#| label: fig-blue
+#| fig-cap: "Distribution of animals under unfished conditions for blue-water simulation."
+
+blue_fauna <- blue_water_fauna_and_fleets$fauna
+
+blue_fleets <- blue_water_fauna_and_fleets$fleets
+
+grid <- expand_grid(x = 1:resolution, y= 1:resolution) %>%
+  mutate(patch = 1:nrow(.))
+
+mpa_locations <- expand_grid(x = 1:resolution, y = 1:resolution) %>%
+  mutate(mpa = TRUE)
+
+blue_sim <- simmar(
+  fauna = blue_fauna,
+  fleets = blue_fleets,
+  manager = list(mpas = list(
+    locations = mpa_locations,
+    mpa_year = 1
+  )),
+  years = 2
+)
+
+patch_biomass <-
+  map_df(blue_sim, ~ map_df(.x, ~tibble(biomass = rowSums(.x$ssb_p_a), patch = 1:nrow(.x$ssb_p_a)), .id = "critter"), .id = "step") %>%
+  mutate(step = as.numeric(step)) %>%
+  left_join(grid, by = "patch")
+
+patch_biomass %>%
+  group_by(critter, step) %>%
+  mutate(sbiomass = biomass / max(biomass)) %>%
+  ungroup() %>%
+  filter(step == max(step)) %>% 
+  ggplot()+
+  geom_tile(aes(x,y, fill = sbiomass),show.legend = FALSE) +
+  facet_wrap(~critter) +
+  scale_x_continuous(expand = c(0,0)) +
+  scale_y_continuous(expand = c(0,0)) +
+  scale_fill_viridis_c() + 
+  theme(axis.text = element_blank(),
+        axis.ticks = element_blank(),
+        axis.title = element_blank())
+
+```
+
+#### Key Drivers
+
+The MPA literature points to the ability of a wide range of drivers to
+affect the performance of MPAs with regards to conservation and economic
+outcomes such as adult and larval dispersal [@gerber2005], the timing of
+density dependence [@babcock2011], correlations in space and time across
+species [@pons2022; @abbott2012] , fleet dynamics and economic
+incentives [@smith2003a; @cabral2019], relationships between size and
+fecundity (e.g. hyperallometry) [@marshall2021], heterogeneity in stock
+status [@hastings2017; @abbott2012], and many other traits [@gaines2010;
+@hilborn2004a].
+
+These studies have all helped guide the science of marine protected
+areas, but often focus on modeling the impact of subset of variables of
+interest at a time. We demonstrate here how `marlin` can be used to
+conduct Monte Carlo simulations across a wide range of drivers to assess
+the relative contributions each to the outcomes of protected areas under
+a particular set of circumstances.
+
+For this example we model the dynamics of a skipjack tuna *Katsuwonus
+pelamis*, a striped marlin *Kajikia audax*, and a oceanic whitetip shark
+*Carcharhinus longimanus*. Life history parameters for each species were
 pulled from FishLife (@thorson2020), with the exception of steepness,
@@
 high volume fishery, \$XX for marlin for a higher value low volume
-fishery, and \$0 for sharks to reflect at true bycatch species. We
-assign a selectivity of 60% of length at maturity for skipjack to allow
-for some overfishing when F/M is high enough, 70% for marlin, and 90%
-for shark.
+fishery, and \$0 for sharks to reflect a true bycatch species. We assign
+a selectivity of 60% of length at maturity for skipjack to allow for
+some overfishing when F/M is high enough, 70% for marlin, and 90% for
+shark xx.
 
@@
 | State                  | Distribution          | Description                                     |
-|-----------------------|-----------------------|-------------------------------------------------|
+|------------------|------------------|-----------------------------------|
 | $\sigma_{centroid}$    |                       | Standard deviation of habitat centroid          |
@@
 | F/M                    |                       | Fishing mortality relative to natural mortality |
-| Adult movement rate   |                       |                                                 |
-| Recruit movement rate |                       |                                                 |
-| Hyperallometry        |                       | Presence of hyperallometric fec                 |
+| Adult diffusion rate   |                       |                                                 |
+| Recruit diffusion rate |                       |                                                 |
+| Hyperallometry         | $U(1,2)$              | Presence of hyperallometric fec                 |
 | DD timing              |                       | Timing of density dependence                    |
 | Seasonal Movement      | $\subset[TRUE,FALSE]$ | Seasonal Movement Present                       |
-
-For each of these `r nrow(state_experiments)` state experiments, we then
-apply `r nrow(placement_experiments)` placement experiments defined by
-an MPA size, an MPA placement strategy, and ...XX
 
@@
 
-$$centroid_s \sim normal(center,\sigma_c)$$ where *s* denotes species,
-and *center* is the coordinate of the central cell. In this way, the
-larger $\sigma_c$ grows, the more variation there is between the
-centroid of the preferred habitat for each species.
+$$centroid_s \sim N(center,\sigma_c)$$ where *s* denotes species, and
+*center* is the coordinate of the central cell. In this way, the larger
+$\sigma_c$ grows, the more variation there is between the centroid of
+the preferred habitat for each species.
 
@@
 and $\sigma_h$ controls the dispersal of the habitat around that range.
+
+For each of these `r nrow(state_experiments)` state experiments, we then
+apply `r nrow(placement_experiments)` placement experiments defined by
+an MPA size, an MPA placement strategy, and the number of animals to be
+considered in the MPA planning, ranging from one to three, where one
+indicates that only one animal is considered in the MPA design process,
+while three means that all three species are considered.
+
+## Results
+
+### Coral Reefs
+
+```{r}
+#| label: fig-coral-results
+#| fig-cap: "Something something."
+#| fig-height: 8
+#| fig-width: 6
+
+coral_mpa_experiements$mpas <- map(coral_mpa_experiements$results, "mpa")
+
+coral_mpa_experiements$obj <- map(coral_mpa_experiements$results, "obj")
+
+coral_results <- coral_mpa_experiements %>%
+  unnest(cols = obj)
+
+coral_fleet_frontier <- coral_results %>%
+  pivot_longer(starts_with("fleet_"), names_to = "fleet", values_to = "fleet_yield") %>%
+  group_by(prop_mpa, fleet, placement_strategy) %>%
+  summarise(yield = sum(fleet_yield),biodiv = sum(unique(biodiv))) %>%
+  ggplot(aes(biodiv, yield, color = placement_strategy)) +
+  geom_line() +
+  scale_x_continuous(name = "Change in Total Depletion",limits = c(0, NA)) +
+    scale_y_continuous(name = "Total Yield",limits = c(0, NA)) +
+  facet_wrap(~ fleet, scales = "free_y")
+
+coral_yield <- coral_results %>% 
+  group_by(prop_mpa, placement_strategy) %>% 
+  summarise(biodiv = sum(biodiv), yield = sum(yield)) %>% 
+  group_by(placement_strategy) %>% 
+  mutate(delta_yield = yield / yield[prop_mpa == 0] - 1,
+         delta_biodiv = biodiv / biodiv[prop_mpa == 0] - 1) %>% 
+  ggplot(aes(delta_biodiv, delta_yield, color = placement_strategy)) + 
+  geom_vline(aes(xintercept = 0), linetype = 2) +
+  geom_hline(aes(yintercept = 0), linetype = 2) +
+  geom_line(size = 2) + 
+  scale_x_continuous(name = "Change in Conservation", labels = scales::label_percent(accuracy = 1)) +
+  scale_y_continuous(name = "Change in Yield", labels = scales::label_percent(accuracy = 1)) + 
+  theme(legend.position = "top")
+
+coral_fleet_yield <- coral_results %>%
+  pivot_longer(starts_with("fleet_"),
+               names_to = "fleet",
+               values_to = "fleet_yield") %>%
+  group_by(prop_mpa, fleet, placement_strategy) %>%
+  summarise(yield = sum(fleet_yield), biodiv = sum(unique(biodiv))) %>%
+  group_by(fleet, placement_strategy) %>%
+  mutate(delta_yield = yield / yield[prop_mpa == 0] - 1) %>%
+  ggplot(aes(prop_mpa, delta_yield, color = fleet)) +
+    geom_hline(aes(yintercept = 0), linetype = 2) +
+  geom_line() +
+  facet_wrap( ~ placement_strategy, scales = "free_y") +
+  scale_x_continuous(name = "MPA Size",
+                     labels = scales::label_percent(accuracy = 1)) +
+  scale_y_continuous(name = "Change in Yield",
+                     labels = scales::label_percent(accuracy = 1))
+
+coral_yield / coral_fleet_yield
+
+```
+
+### Pelagic Systems
+
+```{r}
+#| label: fig-pelagic
+#| fig-cap: "Placeholder for pelagic results is we stick with them"
+
+blue_water_climate_experiments$mpas <- map(blue_water_climate_experiments$results, "mpa")
+
+blue_water_climate_experiments$obj <- map(blue_water_climate_experiments$results, "obj")
+
+blue_results <- blue_water_climate_experiments %>%
+  unnest(cols = obj)
+
+blue_fleet_frontier <- blue_results %>%
+  pivot_longer(names(blue_fleets), names_to = "fleet", values_to = "fleet_yield") %>%
+  group_by(prop_mpa, fleet, placement_strategy) %>%
+  summarise(yield = sum(fleet_yield),biodiv = sum(unique(biodiv))) %>%
+  ggplot(aes(biodiv, yield, color = placement_strategy)) +
+  geom_line() +
+  scale_x_continuous(name = "Change in Total Depletion",limits = c(0, NA)) +
+    scale_y_continuous(name = "Total Yield",limits = c(0, NA)) +
+  facet_wrap(~ fleet, scales = "free_y")
+
+blue_frontier <- blue_results %>% 
+  group_by(prop_mpa, placement_strategy) %>% 
+  summarise(biodiv = sum(biodiv), yield = sum(yield)) %>% 
+  group_by(placement_strategy) %>% 
+  mutate(delta_yield = yield / yield[prop_mpa == 0] - 1,
+         delta_biodiv = biodiv / biodiv[prop_mpa == 0] - 1) %>% 
+  ggplot(aes(delta_biodiv, delta_yield, color = placement_strategy)) + 
+  geom_vline(aes(xintercept = 0), linetype = 2) +
+  geom_hline(aes(yintercept = 0), linetype = 2) +
+  geom_line(size = 2) + 
+  scale_x_continuous(name = "Change in Conservation", labels = scales::label_percent(accuracy = 1)) +
+  scale_y_continuous(name = "Change in Yield", labels = scales::label_percent(accuracy = 1)) + 
+  theme(legend.position = "top")
+
+blue_fleet_yield <- blue_results %>%
+  pivot_longer(names(blue_fleets),
+               names_to = "fleet",
+               values_to = "fleet_yield") %>%
+  group_by(prop_mpa, fleet, placement_strategy) %>%
+  summarise(yield = sum(fleet_yield), biodiv = sum(unique(biodiv))) %>%
+  group_by(fleet, placement_strategy) %>%
+  mutate(delta_yield = yield / yield[prop_mpa == 0] - 1) %>%
+  ggplot(aes(prop_mpa, delta_yield, color = fleet)) +
+    geom_hline(aes(yintercept = 0), linetype = 2) +
+  geom_line() +
+  facet_wrap( ~ placement_strategy, scales = "free_y") +
+  scale_x_continuous(name = "MPA Size",
+                     labels = scales::label_percent(accuracy = 1)) +
+  scale_y_continuous(name = "Change in Yield",
+                     labels = scales::label_percent(accuracy = 1))
+
+blue_critter_bio <- blue_results %>%
+  group_by(prop_mpa, critter, placement_strategy) %>%
+  summarise(biodiv = sum(unique(biodiv))) %>%
+  group_by(critter, placement_strategy) %>%
+  mutate(delta_bio = biodiv / biodiv[prop_mpa == 0] - 1) %>%
+  filter(prop_mpa < 0.4) %>% 
+  ggplot(aes(prop_mpa, pmin(1,delta_bio), color = critter)) +
+    geom_hline(aes(yintercept = 0), linetype = 2) +
+  geom_line() +
+  facet_wrap( ~ placement_strategy) +
+  scale_x_continuous(name = "MPA Size",
+                     labels = scales::label_percent(accuracy = 1)) +
+  scale_y_continuous(name = "Change in Spawning Biomass",
+                     labels = scales::label_percent(accuracy = 1))
+
+# blue_frontier / blue_critter_bio + plot_layout(heights = c(1,2))
+
+blue_critter_bio
+```
+
+### Key Drivers
+
+```{r}
+#| label: fig-drivers
+#| fig-cap: "Distribution of conservation and food security impacts of MPAs covering 30% of available habitat across a range of simulations."
+#| fig-height: 7
+#| fig-width: 7
+
+var_effects <- driver_results %>%
+  ungroup() %>%
+  mutate(biodiv_effect = biodiv / bau_biodiv - 1,
+         econ_effect = econ / bau_econ - 1) %>%
+  select(
+    biodiv_effect,
+    f_v_m,
+    econ_effect,
+    critter,
+    adult_diffusion,
+    recruit_diffusion,
+    density_dependence,
+    placement_strategy,
+    sigma_habitat,
+    hyper,
+    seasonal_movement,
+    fleet_model,
+    spawning_aggregation,
+    seasonal_movement,
+    prop_mpa
+  )
+
+continuous <- map_lgl(var_effects, is.numeric)
+
+bio_effects <- var_effects[,continuous] %>% 
+  filter(between(prop_mpa,0.29, .31)) %>%
+  select(-econ_effect, - prop_mpa) %>% 
+  pivot_longer(-c(contains("_effect")),names_to = "variable", values_to = "value")
+
+econ_effects <- var_effects[,continuous] %>% 
+  filter(between(prop_mpa,0.29, .31)) %>%
+  select(-biodiv_effect, - prop_mpa) %>% 
+  pivot_longer(-c(contains("_effect")),names_to = "variable", values_to = "value")
+
+biodiv_effects_plot <- bio_effects %>%
+  ggplot(aes(value, pmin(2.5,biodiv_effect))) +
+  geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
+  geom_bin2d(show.legend = FALSE) +
+  scale_fill_viridis_c() +
+  scale_y_continuous(name = "Conservation Impact", labels = scales::label_percent(accuracy = 1)) + 
+  scale_x_continuous(name = "") +
+  facet_wrap(~variable, scales = "free_x") +
+  theme(axis.text.x = element_blank(),
+        plot.margin = unit(c(0,0,0,0), units = "points"))
+
+econ_effects_plot <- econ_effects %>%
+  ggplot(aes(value, pmax(-1,econ_effect))) +
+  geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
+  geom_bin2d(show.legend = FALSE) +
+  scale_fill_viridis_c() +
+  scale_y_continuous(name = "Food Impact", labels = scales::label_percent(accuracy = 1)) + 
+  scale_x_continuous(name = "Metric Value") +
+  facet_wrap(~variable, scales = "free_x")
+
+biodiv_effects_plot / econ_effects_plot & theme(strip.text = element_text(size = 8))
+
+```
+
+## Discussion
+
+Waiting on a fleshed out discussion until we are all happy with the content, but in general
+
+* Growing need for more spatially explicit and ecosystem-based management and assessment
+
+* `marlin` fills an important gap in the modeling literature
+    * MPAs
+    * Policy interaction
+    * spatial-temporal MSE
+    * Range shifts
+    * make clear it's not just an MPA model
+
+* Examples demonstrate this
+    * total benefits and Distributional impacts in coral case
+    * Something in the pelagic case
+    * Complex interactions in the final case
+
+* `marlin` and the methods within as a platform for future practical and academic bio-economic modeling needs
+
+## References
+
+::: {#refs}
+:::
+
+## scraps

### 2022-10-12 coral_case_study.R

@@ -10,9 +10,12 @@ tune_type <- "explt"
 
 experiment_workers <- 8
 
-years <- 50
+run_casestudy <- TRUE
 
+years <- 75
+# generate state experiments. This is a somewhat tricky process where the actual generated values are created for many variables in create_experiment_critters
 # https://sustainablefish.org/roundtable/indonesian-snapper-and-grouper/
+
 
 # https://www.tandfonline.com/doi/full/10.1080/23308249.2018.1542420?casa_token=oyFyrvo42J4AAAAA%3A3OklvuxmJxB8TpTVomWzRawFsNRUX4WgwnMa_mcTECTLZn-OatxfLBgU2OES-KYynzhXMyfTC9smdQ
 #
@@ -35,7 +38,6 @@ long_spawning_ground <- expand.grid(x = 1:resolution, y = 1:resolution) %>%
   mutate(habitat = dnorm(x, reefs$x[1], reef_width / 2) * dnorm(y, reefs$y[1], reef_width / 2))
 
 
-
 for (i in 1:nrow(reefs)){
 
   long_reef_habitat$habitat <- long_reef_habitat$habitat + dnorm(long_reef_habitat$x, reefs$x[i], reef_width) * dnorm(long_reef_habitat$y, reefs$y[i], reef_width)
@@ -46,25 +48,7 @@ long_spawning_ground %>%
   ggplot(aes(x,y,fill = habitat)) +
   geom_tile()
 
-long_reef_habitat %>%
-  ggplot(aes(x,y,fill = habitat)) +
-  geom_tile()
-
 reef_habitat <- long_reef_habitat %>%
-  pivot_wider(names_from = y, values_from = habitat) %>%
-  select(-x) %>%
-  as.matrix()
-
-
-deep_reef_habitat <- long_reef_habitat %>%
-  mutate(habitat = habitat * 1 / (1 + exp(-(x - resolution / 2)))) %>%
-  pivot_wider(names_from = y, values_from = habitat) %>%
-  select(-x) %>%
-  as.matrix()
-
-
-shallow_reef_habitat <- long_reef_habitat %>%
-  mutate(habitat = habitat * (1 - 1 / (1 + exp(-(x - resolution / 2))))) %>%
   pivot_wider(names_from = y, values_from = habitat) %>%
   select(-x) %>%
   as.matrix()
@@ -76,43 +60,24 @@ spawning_ground <- long_spawning_ground %>%
[…]

### 2022-10-12 scripts/00_setup.R

@@ -27,6 +27,10 @@ library(rpart.plot)
 
 library(rstanarm)
 
+library(ggrepel)
+
+library(ggdist)
+
 options(dplyr.summarise.inform = FALSE)
 
 
@@ -44,11 +48,11 @@ if (!dir.exists(results_path)){
   dir.create(file.path(results_path,"sims"))
 }
 
-run_experiments <- TRUE
+run_experiments <- FALSE
 
 run_blue_water_example <- FALSE
 
-run_coral_example <- FALSE
+run_coral_example <- TRUE
 
 experiment_workers <- 8
 

### 2022-10-12 scripts/01_run_experiments.R

@@ -538,6 +538,8 @@ results <- results %>%
   mutate(bau_biodiv = biodiv[prop_mpa == 0],
          bau_econ = econ[prop_mpa == 0])
 
+write_rds(results, file.path(results_path,"driver_results.rds"))
+
 results %>%
   ggplot(aes(biodiv - bau_biodiv)) +
   geom_histogram()

### 2022-10-12 marlin-paper.qmd
rough draft

@@
 params:
-  run_name: ["v1.2"]
+  run_name: ["v0.1"]
 abstract: "Very rough cut.
@@
   - \usepackage{lineno}\linenumbers
-
 ---
@@
 Many models have been developed to support the theory and design of
-spatial management strategies (see @fulton2015 for a review of existing methods). However, they have tended
-to fall into one of either highly complex and tactical models designed
-for use in a specific location, or to be highly stylized representations
-intended to provide theoretical insights with less relevance for highly
-specific design questions. This complex landscape presents a challenge
-to stakeholders charged with marine spatial planning exercises lacking
-accessible models suited for more tactical applications, and a barrier
-to the scientific community, wherein comparisons across spatial
-management simulation studies are clouded by discrepancies in underlying
-model structures beyond differences in the core phenomenon in question.
+spatial management strategies (see @fulton2015 for a review of existing
+methods). However, they have tended to fall into one of either highly
+complex and tactical models designed for use in a specific location, or
+to be highly stylized representations intended to provide theoretical
+insights with less relevance for highly specific design questions. This
+complex landscape presents a challenge to stakeholders charged with
+marine spatial planning exercises lacking accessible models suited for
+more tactical applications, and a barrier to the scientific community,
+wherein comparisons across spatial management simulation studies are
+clouded by discrepancies in underlying model structures beyond
+differences in the core phenomenon in question.
 
@@
 
-`marlin` follows a "models of intermediate complexity" (MICE) philosophy [@plagányi2014]
-targeted specifically towards capturing heterogeneous habitats, movement dynamics, and technical interactions across species.
-In its general form, `marlin` simulates the behavior of populations of spatially explicit biologically independent age-structured animal populations affected by one or more fishing fleets in time steps specified by the user (generally annual or finer).
-While `marlin`
-does not simulate species interactions, fishing fleets in `marlin` are
-capable of targeting and affecting multiple species simultaneously.
-This allows the model to simulate processes such as fisheries bycatch or effort displacement in a way that accounts for the ways that changes in fleet behavior might affect multiple species in the system.
-The ability to track the impacts of fishing fleets across multiple species simultaneously is particularly important as very few fishing fleets are truly single species.
+`marlin` follows a "models of intermediate complexity" (MICE) philosophy
+[@plagányi2014] targeted specifically towards capturing heterogeneous
+habitats, movement dynamics, and technical interactions across species.
+In its general form, `marlin` simulates the behavior of populations of
+spatially explicit biologically independent age-structured animal
+populations affected by one or more fishing fleets in time steps
+specified by the user (generally annual or finer). While `marlin` does
+not simulate species interactions, fishing fleets in `marlin` are
+capable of targeting and affecting multiple species simultaneously. This
+allows the model to simulate processes such as fisheries bycatch or
+effort displacement in a way that accounts for the ways that changes in
+fleet behavior might affect multiple species in the system. The ability
+to track the impacts of fishing fleets across multiple species
+simultaneously is particularly important as very few fishing fleets are
+truly single species.
 
 Below we provide a summary of the `marlin` model, as well as details of
-the generalized simulation exercises and case study analyses.
-We note that `marlin` is a complex package with many options to control detailed behaviors.
-Given that these specific features will evolve over time, we do not describe in detail every option of the model here.
-Rather, we focus on the equations and methods describing the core foundations of our framework for efficiently simulating bio-economic dynamics in aquatic systems.
+the generalized simulation exercises and case study analyses. We note
+that `marlin` is a complex package with many options to control detailed
+behaviors. Given that these specific features will evolve over time, we
+do not describe in detail every option of the model here. Rather, we
+focus on the equations and methods describing the core foundations of
+our framework for efficiently simulating bio-economic dynamics in
+aquatic systems.
+
+While users can adapt the general equations and ideas presented here as
+they see fit, the software package accompanying this paper,
+[`marlin`](https://danovando.github.io/marlin/), is written in R
+[@rcoreteam2021] and C++, integrated through the `Rcpp` package
+@eddelbuettel2018 .
 
@@
 
-A fishing fleet is defined in `marlin` by a total pool of
-effort that can have different catchability, prices, and selectivity for
-individual species within a simulation (each of which we denote as a *metier*. For example, both a longline and
-purse seine fleet may capture bigeye (*Thunnus obesus*) and skipjack
-(*Katswamus pelamus*) tunas, but the longline fleet is much more likely to
-catch bigeye than skipjack, and *vice versa*. Each fishing fleet
-generates catches, revenues, and profits from individual species that it
-targets, and makes decisions based on its total objective function
-across all species.
+A fishing fleet is defined in `marlin` by a total pool of effort that
+can have different catchability, prices, and selectivity for individual
+species within a simulation (each of which we denote as a *metier*. For
+example, both a longline and purse seine fleet may capture bigeye
+(*Thunnus obesus*) and skipjack (*Katswamus pelamus*) tunas, but the
+longline fleet is much more likely to catch bigeye than skipjack, and
+*vice versa*. Each fishing fleet generates catches, revenues, and
+profits from individual species that it targets, and makes decisions
+based on its total objective function across all species.
 
@@
 
-where $$\gamma_f$$ the scales the cost per unit effort for fleet *f*, $\beta$ allows for the cost of effort to scale non-linearly, and $\eta$ is an additional cost per unit effort of fishing in each patch *p*, allowing for the model to account for travel costs for different patches.
-Travel costs are calculated based on the euclidean distance of each patch to the nearest port of a given fleet, users can specify any number from zero to the number of patches of port locations.
-When no ports are specified, travel costs are zero.
+where $$\gamma_f$$ the scales the cost per unit effort for fleet *f*,
+$\beta$ allows for the cost of effort to scale non-linearly, and $\eta$
+is an additional cost per unit effort of fishing in each patch *p*,
+allowing for the model to account for travel costs for different
+patches. Travel costs are calculated based on the euclidean distance of
+each patch to the nearest port of a given fleet, users can specify any
+number from zero to the number of patches of port locations. When no
+ports are specified, travel costs are zero.
 
@@
 
-`marlin` allows for two general modes of effort. The simplest is "constant effort", in which the total effort of each fishing fleet remains constant over time.
-
-The more complex option is "open access".
-Under open access, the total amount of effort in a given time step for fleet *f* is a function of the profitability of that fleet in the proceeding time steps in which fishing was open.
+`marlin` allows for two general modes of effort. The simplest is
+"constant effort", in which the total effort of each fishing fleet
+remains constant over time.
+
+The more complex option is "open access". Under open access, the total
+amount of effort in a given time step for fleet *f* is a function of the
+profitability of that fleet in the proceeding time steps in which
+fishing was open.
 
@@
 
-There $\theta$ controls the responsiveness of fleet *f* to the log of the ratio of total revenues *R* to total costs *C*, and approximates the percentage amount effort changes in response to a one unit changes in the log revenue to cost ratio.
-
-Three general exceptions exist for the effort dynamics.
-When MPAs are implemented, effort that used to operate inside the MPAs can either leave the fishery, or be redistributed outside the the MPA (the default behavior).
-The model can also accept total catch quotas for each species in each time step.
-When quotas are activated, if the total catch across all fleets for a given species under the model's effort dynamics would exceed the allowable quota for that species, the total effort for each fleet is decrease proportional to its contribution to the total catch until the quota is satisfied.
-`marlin` also allows users to specify closed fishing seasons for one or more species in the system.
-During closed seasons, no effort is allocated to patches that would result in capture of the covered species, potentially resulting in zero effort if the species cannot be avoided.
-Under these conditions, the amount of effort in the next open season is a function of the effort and profits in the last open fishing season.
+There $\theta$ controls the responsiveness of fleet *f* to the log of
+the ratio of total revenues *R* to total costs *C*, and approximates the
+percentage amount effort changes in response to a one unit changes in
+the log revenue to cost ratio.
+
+Three general exceptions exist for the effort dynamics. When MPAs are
+implemented, effort that used to operate inside the MPAs can either
+leave the fishery, or be redistributed outside the the MPA (the default
+behavior). The model can also accept total catch quotas for each species
+in each time step. When quotas are activated, if the total catch across
+all fleets for a given species under the model's effort dynamics would
+exceed the allowable quota for that species, the total effort for each
+fleet is decrease proportional to its contribution to the total catch
+until the quota is satisfied. `marlin` also allows users to specify
+closed fishing seasons for one or more species in the system. During
+closed seasons, no effort is allocated to patches that would result in
+capture of the covered species, potentially resulting in zero effort if
+the species cannot be avoided. Under these conditions, the amount of
+effort in the next open season is a function of the effort and profits
+in the last open fishing season.
 
@@
 
-The degree of fishing pressure exerted by a given set of fishing fleets on each species is a function of a range of parameters including the total amount of effort, the selectivity ogives, cost functions, the spatial distribution of the resource, the relative prices across species, and the catchability coefficient of each fleet for each species.
-`marlin` provides a tuning option to help users achieve desired biological outcomes from their fleets.
-
-Users can tune their fleets in one of two ways.
-First, they can specify a target exploitation rate *u* for each species in their model run.
-Taking all the other parameters of the model as given, `marlin` then adjusts the catchability coefficients *q_{f,s}* for each fleet *f* and species *s* such that the equilibrium exploitation rate for each species matches the desired level.
-Second, they can specify a target depletion, defined as total spawning stock biomass under fishing divided by total unfished spawning stock biomass, and the model will adjust the catchability coefficients *q_{f,s}* for each fleet *f* and species *s* such that the equilibrium depletion for each species matches the desired level.
+The degree of fishing pressure exerted by a given set of fishing fleets
+on each species is a function of a range of parameters including the
+total amount of effort, the selectivity ogives, cost functions, the
+spatial distribution of the resource, the relative prices across
+species, and the catchability coefficient of each fleet for each
+species. `marlin` provides a tuning option to help users achieve desired
+biological outcomes from their fleets.
+
+Users can tune their fleets in one of two ways. First, they can specify
+a target exploitation rate *u* for each species in their model run.
+Taking all the other parameters of the model as given, `marlin` then
+adjusts the catchability coefficients *q\_{f,s}* for each fleet *f* and
+species *s* such that the equilibrium exploitation rate for each species
+matches the desired level. Second, they can specify a target depletion,
+defined as total spawning stock biomass under fishing divided by total
+unfished spawning stock biomass, and the model will adjust the
+catchability coefficients *q\_{f,s}* for each fleet *f* and species *s*
+such that the equilibrium depletion for each species matches the desired
+level.
 
@@
 
-Each fishing fleet decides how to allocate its effort in space
-based on one of four possible spatial allocation strategies. The ideal
-free distribution (IFD) is the standard method for distributing fishing
+Each fishing fleet decides how to allocate its effort in space based on
+one of four possible spatial allocation strategies. The ideal free
+distribution (IFD) is the standard method for distributing fishing
 fleets in spaces (see @gillis2003 and references therein). However, the
-IFD presents a number of complications for our model. First, under the constant effort model, the total amount of effort required to satisfy the IFD may not equal the total amount of
-effort required by the model. Second, even if we were to implement the
-IFD for a single fleet, the model is capable of tracking multiple fleets
-at once. This means that the IFD for a single fleet would be a
-game-theoretic outcome of the actions of each of the individual fleet
-conditional on the actions of all other fleets. While possible to
-implement this would slow down our model runs to the point of
-practically preventing large-scale evaluation of the effects of MPAs on
-bycatch.
+IFD presents a number of complications for our model. First, under the
+constant effort model, the total amount of effort required to satisfy
+the IFD may not equal the total amount of effort required by the model.
+Second, even if we were to implement the IFD for a single fleet, the
+model is capable of tracking multiple fleets at once. This means that
+the IFD for a single fleet would be a game-theoretic outcome of the
+actions of each of the individual fleet conditional on the actions of
+all other fleets. While possible to implement this would slow down our
+model runs to the point of practically preventing large-scale evaluation
+of the effects of MPAs on bycatch.
 
@@
 
-2. Local density dependence: Density
-dependence occurs independently in each patch and recruits are retained
-in their home patch.
+2.  Local density dependence: Density dependence occurs independently in
+    each patch and recruits are retained in their home patch.
 
@@
 
-#### Dispersal
-
-`marlin`'s movement dynamics are based on a continuous-time Markov
-chain (CTMC), as described in @thorson2021a. Within this framework, the model
+#### Movement
+
+`marlin`'s movement dynamics are based on a continuous-time Markov chain
+(CTMC), as described in @thorson2021a. Within this framework, the model
 allows for movement to be broken down into three components of
 *diffusion*, *advection* (follow currents), and *taxis* (active movement
-towards say a preferred habitat).
-For now, our model focuses just on the diffusion and taxis components of this model, though future extensions could easily incorporate advection vectors from say ROMs models (citations).
-In this way then, `marlin` allows users to run anything from a simple Gaussian dispersal kernel up to a system governed by species that
-passively diffuse around a dynamic thermal range.
+towards say a preferred habitat). For now, our model focuses just on the
+diffusion and taxis components of this model, though future extensions
+could easily incorporate advection vectors from say ROMs models
+(citations). In this way then, `marlin` allows users to run anything
+from a simple Gaussian dispersal kernel up to a system governed by
+species that passively diffuse around a dynamic thermal range.
 
 @thorson2021a described the exact steps of the CTMC approach in detail.
-We briefly the general CTMC method here, but see @thorson2021a for a detailed description of the method.
-Under this framework, movement from each patch to each other patch in the system in a given timestep *t* for life stage *g* of species *s* is defined by a movement matrix $\boldsymbol{M}_{t,s,g}$.
-$\boldsymbol{M}_{t,s,g}$ is calculated as the matrix exponent of the sum of the diffusion and taxis matrices scaled by the width of the time step that the model is running under.
+We briefly the general CTMC method here, but see @thorson2021a for a
+detailed description of the method. Under this framework, movement from
+each patch to each other patch in the system in a given timestep *t* for
+life stage *g* of species *s* is defined by a movement matrix
+$\boldsymbol{M}_{t,s,g}$. $\boldsymbol{M}_{t,s,g}$ is calculated as the
+matrix exponent of the sum of the diffusion and taxis matrices scaled by
+the width of the time step that the model is running under.
 
@@
 
-Where $\boldsymbol{D}$ is the diffusion matrix and $\boldsymbol{X}$ is the taxis matrix.
-
-The diffusion matrix is a function of a user specified diffusion rate, see @thorson2021a. We slightly modify the taxis matrix in the following manner
-
-xx need to fix notation somewhere here, see this line of code for much clearer depiction of what's happening, since sometimes code is cleaner.
-https://github.com/DanOvando/marlin/blob/3563e7319ad0267f89789267a4c6ee58a9ad0d60/R/fish_class.R#L525 xx
+Where $\boldsymbol{D}$ is the diffusion matrix and $\boldsymbol{X}$ is
+the taxis matrix.
+
+The diffusion matrix is a function of a user specified diffusion rate,
+see @thorson2021a. We slightly modify the taxis matrix in the following
+manner
+
+xx need to fix notation somewhere here, see this line of code for much
+clearer depiction of what's happening, since sometimes code is cleaner.
+https://github.com/DanOvando/marlin/blob/3563e7319ad0267f89789267a4c6ee58a9ad0d60/R/fish_class.R#L525
+xx
 
@@
 
-The $lambda d_{s,f}$ term is a scalar in which the magnitude of the habitat gradient between patches can be scaled relative to the diffusion parameter $d$.
-
-For both the diffusion and taxis matrix, we allow for the inclusion of physical barriers to movement (i.e.
-land).
-Pairs of patches that are adjacent but in which one or both patches are a barrier to movement are set as non-adjacent.
-The CTMC model then produces movement dynamics that move around barriers rather than over them.
-This is different than setting habitat in a patch to zero, which implies a patch that a given animal will not move towards for its own sake, but can move over towards another better patch.
-
-While this CTMC approach to movement simulation is to date not commonly seen in the
-marine modeling literature, it has numerous advantages that warrant its
-seeming complexity. First, when only diffusion is present, the model
-will generalize to the familiar dynamics of a Gaussian dispersal kernel.
-However, the time component *t* in @eq-movement allows the movement model to be run on arbitrarily fine time steps.
-Second, the *taxis* model allow for more active habitat choices by
-species, allowing us to simulate both preferences of species in space
-and time efficiently.Lastly, this form has the advantage that the parameters are directly
-estimable from real data. So, if provided with say CPUE data and a
-tagging study, users can estimates the movement parameters in the same
-manner as @thorson2021a, and then pass their estimated parameters to
-`marlin` for simulation.
+The $lambda d_{s,f}$ term is a scalar in which the magnitude of the
+habitat gradient between patches can be scaled relative to the diffusion
+parameter $d$.
+
+For both the diffusion and taxis matrix, we allow for the inclusion of
+physical barriers to movement (i.e. land). Pairs of patches that are
+adjacent but in which one or both patches are a barrier to movement are
+set as non-adjacent. The CTMC model then produces movement dynamics that
+move around barriers rather than over them. This is different than
+setting habitat in a patch to zero, which implies a patch that a given
+animal will not move towards for its own sake, but can move over towards
+another better patch.
+
+While this CTMC approach to movement simulation is to date not commonly
+seen in the marine modeling literature, it has numerous advantages that
+warrant its seeming complexity. First, when only diffusion is present,
+the model will generalize to the familiar dynamics of a Gaussian
+dispersal kernel. However, the time component *t* in @eq-movement allows
+the movement model to be run on arbitrarily fine time steps. Second, the
+*taxis* model allow for more active habitat choices by species, allowing
+us to simulate both preferences of species in space and time
+efficiently.Lastly, this form has the advantage that the parameters are
+directly estimable from real data. So, if provided with say CPUE data
+and a tagging study, users can estimates the movement parameters in the
+same manner as @thorson2021a, and then pass their estimated parameters
+to `marlin` for simulation.
+
+One challenge is that the coefficients of the model matrices inside
+@eq-movement can be hard to interpret. To facilitate this, `marlin`
+allows users to specify a maximum number of cells that adults or
+recruits diffuse in an year. The model will then tune the diffusion
+parameter and habitat gradients to achieve this desired level of
+movement.
 
@@
 
-SSBmsy is more conserved, as over the long term the key things
-is the flow of recruits, so we do provide an option for that, but I haven't messed with it for a while.
+SSBmsy is more conserved, as over the long term the key things is the
+flow of recruits, so we do provide an option for that, but I haven't
+messed with it for a while.
 
@@
 sized" fish (citation), and logistic selectivity for sharks, as any
-shark caught is retained. This fleet is a subsistence fleet,
-where the local community takes every possible opportunity to fish that
-they can, and as such we model it as a "constant effort" fishery.
+shark caught is retained. This fleet is a subsistence fleet, where the
+local community takes every possible opportunity to fish that they can,
+and as such we model it as a "constant effort" fishery.
 
@@
 mortality for deep-water snapper resulting from barotrauma (xx). This
-fleet catches some grouper, though less than fleet one, and receives no price for
-sharks due to the requirements of a certification program through which
-they sell their deep-water snapper, though they do accidentally catch and kill them on occasion. This fleet operates under open-access
-dynamics, as fishing is not the primary means of subsistence and so
-short-term effort expands and contracts in respond to profitability of
-the primarily grouper-driven fishery xx.
+fleet catches some grouper, though less than fleet one, and receives no
+price for sharks due to the requirements of a certification program
+through which they sell their deep-water snapper, though they do
+accidentally catch and kill them on occasion. This fleet operates under
+open-access dynamics, as fishing is not the primary means of subsistence
+and so short-term effort expands and contracts in respond to
+profitability of the primarily grouper-driven fishery xx.
 
@@
 | State                  | Distribution          | Description                                     |
-|------------------|------------------|-----------------------------------|
+|-------------------|-------------------|-----------------------------------|
 | $\sigma_{centroid}$    |                       | Standard deviation of habitat centroid          |
@@
 #| label: fig-coral-results
-#| fig-cap: "Something something."
+#| fig-cap: "Frontier in conservation and yield outcomes under different MPA design strategies (top panel) for coral reef case study.
+Change in yield for each fleet as a function of MPA size and MPA design strategy (bottom panel) for coral reef case study."
 #| fig-height: 8
@@
 
-Waiting on a fleshed out discussion until we are all happy with the content, but in general
-
-* Growing need for more spatially explicit and ecosystem-based management and assessment
-
-* `marlin` fills an important gap in the modeling literature
-    * MPAs
-    * Policy interaction
-    * spatial-temporal MSE
-    * Range shifts
-    * make clear it's not just an MPA model
-
-* Examples demonstrate this
-    * total benefits and Distributional impacts in coral case
-    * Something in the pelagic case
-    * Complex interactions in the final case
-
-* `marlin` and the methods within as a platform for future practical and academic bio-economic modeling needs
+Waiting on a fleshed out discussion until we are all happy with the
+content, but in general
+
+-   Growing need for more spatially explicit and ecosystem-based
+    management and assessment
+
+-   `marlin` fills an important gap in the modeling literature
+
+    -   MPAs
+    -   Policy interaction
+    -   spatial-temporal MSE
+    -   Range shifts
+    -   make clear it's not just an MPA model
+    -   Draw contrast to virtual species and STEPS
+
+-   Examples demonstrate this
+
+    -   total benefits and Distributional impacts in coral case
+    -   Something in the pelagic case
+    -   Complex interactions in the final case
+
+-   `marlin` and the methods within as a platform for future practical
+    and academic bio-economic modeling needs
 

### 2022-10-12 scripts/00_setup.R

@@ -38,7 +38,7 @@ foos <- list.files(here("R"))
 
 walk(foos, ~ source(here("R", .x)))
 
-results_name <- "v1.2"
+results_name <- "v0.1"
 
 results_path <- here("results", results_name)
 

### 2022-11-17 03_marlin_paper.qmd
getting there. coral and pelagic case studies up and in paper

--- 2022-10-12 marlin-paper.qmd
@@
 ---
-title: "Efficient Simulation of Economic and Ecological Dynamics in Aquatic Systems"
+title: "Policy Implications of Coupled Economic and Ecological Dynamics in Aquatic Systems"
 author:
@@
 We present a framework for efficient simulation of the bio-economic dynamics of marine systems, designed to support both management and research.
-We demonstrate the model's capabilities by focusing on three case studies on the conservation and food security impacts of marine protected areas (MPAs)."
+We demonstrate the model's capabilities by focusing on two case studies on the conservation and food security impacts of marine protected areas (MPAs).
+We highlight how coupled bio-economic dynamics can result in complex trade-offs between conservation, fisheries yields, distributional outcomes in simulated coral reef and plegaic ecosystems.
+"
 header-includes:
@@
 
+XX explain hyperallometry option
+
+xx add in weird option 5 density dependence
+
 ```{r setup}
@@
 
-theme_set(marlin::theme_marlin())
+theme_set(theme_classic(base_size = 12, base_family = "Helvetica") + theme(strip.background = element_blank()))
 
@@
 
-rdatas <- results_files[str_detect(results_files,'.Rdata')]
-
-walk(rdatas, ~load(file.path(results_path,.x), envir = .GlobalEnv)) # load results into global environment
+# rdatas <- results_files[str_detect(results_files,'.Rdata')]
+# 
+# walk(rdatas, ~load(file.path(results_path,.x), envir = .GlobalEnv)) # load results into global environment
 
@@
 populations affected by one or more fishing fleets in time steps
-specified by the user (generally annual or finer). While `marlin` does
-not simulate species interactions, fishing fleets in `marlin` are
+specified by the user (generally annual or seasonal). While `marlin`
+does not simulate species interactions, fishing fleets in `marlin` are
 capable of targeting and affecting multiple species simultaneously. This
@@
 
-A fishing fleet is defined in `marlin` by a total pool of effort that
-can have different catchability, prices, and selectivity for individual
-species within a simulation (each of which we denote as a *metier*. For
+A fishing fleet is defined in `marlin` by a set of fishers that can have
+different catchability, prices, and selectivity for individual species
+within a simulation (each of which we denote as a *metier*). For
 example, both a longline and purse seine fleet may capture bigeye
@@
 longline fleet is much more likely to catch bigeye than skipjack, and
-*vice versa*. Each fishing fleet generates catches, revenues, and
-profits from individual species that it targets, and makes decisions
-based on its total objective function across all species.
-
-Revenues *R* for fleet *f* in time step *t* are given by
+*vice versa*. Each fishing fleet (*f*) generates catches or yield (*Y*),
+revenues (*R*), costs (*C*) from individual species (*s*) that it
+targets in a given time step (*t*) and patch (*p*). The fleet then makes
+decisions around how much and where to fish, subject to regulatory
+constraints, based on th total profitability across all species in space
+and time.
+
+Revenues *R* for fleet *f* in time step *t* are a function of the total
+amount caught of each species and its price ($\Pi$). The amount caught
+is a function of the selectivity at age of each species for each fleet
+($\alpha$), the fishing efficiency of the fleet for that that species
+(*q*), the amount of fishing effort of the fleet in question in that
+patch in a given time step (*E*), and the total fishing mortality at age
+(including all other fleets) for that particular species in that patch
+and time (*u*).
 
@@
 
-where *p* is patch, *s* is species, *a* is age, $\alpha$ is selectivity
-at age, *q* is catchability , *u* is total fishing mortality, $\Pi$ is
-price per unit weight, and *Y* is total catches.
+*Y* is total catches, which is calculated through the Baranov equation
+[@baranov1918], which appropriately accounts for the total instantaneous
+mortality (both fishing and natural) and divides the total amount of the
+population biomass (*b*) killed between natural (*m*) and fishing (*u*)
+sources, the amount of biomass killed through fishing being assigned to yield *Y*.
 
@@
 
-Selectivity $\alpha$ is modeled through a logistic form per
+Selectivity $\alpha$ is modeled through either a logistic form or a dome-shaped form.
+The logistic form is based on the lengths *l* at which 50% of individuals are selected by the fishing gear, $l^{sel}$, and $\delta$ which is the difference between the length at 50% selectivity and
+the length at 95% selectivity.
 
@@
 
-where $l$ is length at age, $l^{sel}$ is the length at 50% selectivity,
-and $\delta$ is the difference between the length at 50% selectivity and
-the length at 95% selectivity.
-
-or through a dome shaped form of
-
-XX add dome shaped equation XX
-
-*z* is total mortality given by
+We approximate the dome-shaped form as a simple normal distribution with mean $l^{sel}$ and standard deviation $\sigma$.
+The normal density function is re-scaled such that the selectivity is 1 at $l^{sel}$.
+
+$$
+\alpha_{a,s,f}= \frac{1}{\sigma_{s,f} \sqrt{2\pi}}e^{-0.5 \left( \frac{l_{a,s} - l^{sel}}{\sigma_{s,f}}  \right)^2}
+$$
+Lastly  *z* is total mortality, the sum of fishing mortality and natural mortality (*m*).
 
@@
 $$ {#eq-z}
-
-Where *m* is natural mortality.
 
@@
 catches and subsequently revenues for each fleet. Given revenues *R*, we
-then calculate profits $\phi$ per
+then calculate profits $\phi$ based on revenues and costs.
 
@@
 
-where costs *C* are calculated
-
-$$C_{t,p,f} = \gamma_f  \left( E_{t,p,f}^{\beta_f} + \eta_{f,p} E_{t,p,f} \right)$$
-
-where $$\gamma_f$$ the scales the cost per unit effort for fleet *f*,
-$\beta$ allows for the cost of effort to scale non-linearly, and $\eta$
+Costs *C* are calculated as a function of $\gamma_f$, that the scales the cost per unit effort for fleet *f*,$\beta$ allows for the cost of effort to scale non-linearly, and $\eta$
 is an additional cost per unit effort of fishing in each patch *p*,
@@
 
+$$C_{t,p,f} = \gamma_f  \left( E_{t,p,f}^{\beta_f} + \eta_{f,p} E_{t,p,f} \right)$$
+
 #### Calculating Total Effort
@@
 profitability of that fleet in the proceeding time steps in which
-fishing was open.
-
-$$
-E_{t+1,f} = E_{t,f} \times e^{\theta_f log(R_{t,f} / C_{t,f})} 
-$$ {#eq-oa}
-
-There $\theta$ controls the responsiveness of fleet *f* to the log of
+fishing was open, where $\theta$ controls the responsiveness of fleet *f* to the log of
 the ratio of total revenues *R* to total costs *C*, and approximates the
@@
 
-Three general exceptions exist for the effort dynamics. When MPAs are
-implemented, effort that used to operate inside the MPAs can either
+$$
+E_{t+1,f} = E_{t,f} \times e^{\theta_f log(R_{t,f} / C_{t,f})} 
+$$ {#eq-oa}
+
+When MPAs are implemented, effort that used to operate inside the MPAs can either
 leave the fishery, or be redistributed outside the the MPA (the default
@@
 fleet is decrease proportional to its contribution to the total catch
-until the quota is satisfied. `marlin` also allows users to specify
+until the quota is satisfied.
+As an alternative to catch quotas, users can input an effort cap per fleet, which prevents effort from exceeding a given amount under open-access dynamics, though effort may be reduced below this cap if required by the profit equation.
+`marlin` also allows users to specify
 closed fishing seasons for one or more species in the system. During
@@
 The underlying population model used is an age structured single-species
-model in the manner of @ovando2021a.
-
-For the population model, numbers *n* at time *t* for age *a* are given
-by
+model in the manner of @ovando2021a. The population model requires a
+large number of parameters. However, if the user supplies either a
+scientific or common name for the species in question (scientific
+preferred), the model will supply default values for that species based
+on the values reported in FishLife [@thorson2020].
+All of these values can be replaced with user-supplied values as needed though, the defaults simply ensure that reasonable and internally consistent values for the species in question are supplied as a starting point.
+
+For the population model, numbers *n* at time *t* for age *a* are a function of growth, death, and recruitment
 
@@
 where *BH* is the Beverton-Holt recruitment function, *ssb* is
-spawning-stock-biomass.
-
-where $l_a$ is the mean length at at age, $l_sel$ is the length at which
+spawning-stock-biomass. $l_a$ is the mean length at at age, $l_sel$ is the length at which
 on average 50% of individuals are selected by the fishery, and
@@
 
-*ssb* is calculated by converting age to mean length, calculating weight
+Spawning stock biomass *ssb* is calculated by converting age to mean length, calculating weight
 at age, maturity at age, and then calculating spawning stock biomass as
-the sum of spawning potential at age in a given time step.
-
-$$
-  l_{a,s} = l_{\infty,s}\left(1 - e^{-k_s(a - a0_{s})}\right)
+the sum of spawning potential at age in a given time step, taking into account the potential for hyperallometry.
+Age is converted to length through the von Bertalanffy  growth equation, assuming log-normally distributed variation in the length at age.
+
+$$
+  l_{a,s} = l_{\infty,s}\left(1 - e^{-k_s(a - a0_{s})}\right)e^{u_s}
 $$ {#eq-len}
 
-Biomass *b* at age is then given by
+Where *u* is a log-normally distributed term allowing for variation in
+the length at age relationship
+
+$$
+u_s \sim normal(0,\sigma_s)
+$$ {#eq-sigma}
+
+Biomass *b* at age is then given by the weight at length equation governed by a scaling coefficient $wa_s$ and an exponent $wb_s$ that controls the rate at which volume scales with length
 
@@
 
-and maturity *mat* is calculated as
-
-$$
-  mat_{a,s} = \frac{1}{(1 + e^{-log(19)\times\frac{l_{a,s} - lmat_s}{{\delta}mat_s}}}
+Maturity is then calculated as a logistic function where $l_{mat}$ is the length at which on average 50% of individuals are sexual maturity, and $\delta_{mat}$ is the units of length beyond
+$l_{mat}$ at which on average 95% of fish are sexually mature.
+
+$$
+  mat_{a,s} = \frac{1}{\left(1 + e^{-log(19)\times\frac{l_{a,s} - lmat_s}{{\delta}mat_s}}\right)}
 $$ {#eq-mat)}
 
-where $l_{mat}$ is the length at which on average 50% of individuals are
-sexual maturity, and $\delta_{mat}$ is the units of length beyond
-$l_{mat}$ at which on average 95% of fish are sexually mature.
-
-Spawning stock biomass at time *t* is then calculated as
+Spawning stock biomass at time *t* is then calculated as a function of the numbers at age, the maturity at age, and the weight at age scaled by a parameter $\gamma$.
+When $\gamma$ is greater than 1, the species is said to experience hyperallometric fecundity, i.e.
+fecundity increases faster than weight.
 
@@
 
-where **d^l^** is the recruitment movement matrix
+where **d^l^** is the recruitment movement matrix (xx make this notation
+consistent with the movement block)
 
@@
 $$ {#eq-dd4}
+
+5.  Global density dependence allocated by spawning biomass: Density dependence is a function of the
+    sum of spawning biomass across all patches, and recruits are then
+    distributed according to the distribution of spawning biomass
+
+$$
+n_{t,p,s,a = 1} = \left(\frac{0.8{\times}\sum_{p=1}^P{r0_{p}}\times{h}\times\sum_{p=1}^P{ssb_{t-1,p}}}{0.2\times{\sum_{p=1}^Pssb0_p}\times(1 - h)+(h - 0.2)\times{\sum_{p=1}^Pssb_{t-1,p}}}\right)\times \frac{ssb_{t-1,p}}{\sum_{p=1}^Pssb_{t-1,p}}
+$$ {#eq-dd5}
 
@@
 
-Sigh. I need to work on this section, but I am *le tired*.
-
-A note here Why no reference points? The problem: Fmsy and MSY are
-selectivity dependent, and once we allow for different fleets with
-different selectivities that critical can move in space, that movement
-affects the net selectivity and fishing mortality on the population.
-
-Bmsy is also affected, since if selectivity is very low, you might need
-more biomass of immature individuals relative to less biomass of more
-mature individuals; a biomass of X can be more or less reproductively
-productive depending on the size / age makeup of that biomass.
-
-SSBmsy is more conserved, as over the long term the key things is the
-flow of recruits, so we do provide an option for that, but I haven't
-messed with it for a while.
-
+Fisheries management is often concerned with measuring stock status relative to maximum sustainable yield (MSY) based reference points, though the exact level of stock status desired by societal objectives may vary widely.
+MS based reference points present a problem for a multi-fleet and spatial-temporal model such as `marlin`.
+MSY and the fishing mortality rate that would produce MSY, $F_{MSY}$ are a function of fishery selectivity.
+Fishery selectivity in this model can vary by fleet, and species can be distributed unevenly in both space and time.
+This means that the net fishing selectivity on a species can vary depending on the dynamics at a given moment, making the definition of an equilibrium concept such as MSY challenging.
+
+As such, we do not report MSY based reference points in the model by default, with the exception of the spawning stock biomass at MSY ($SSB_{MSY}$) which is more conserved across selectivity patterns as it is primarily a function of the mortality and flow of recruits relative to the maturity ogive.
 There are many different strategies for estimating reference points
-conditional on selectivity and regimes, see citations. We leave it to
+conditional on selectivity and regimes, see citations xx. We leave it to
 users to find relevant reference point as required by their methods on
@@
 
-### MPA Optimization
-
-Our MPA optimization algorithm seems to create an MPA network that
-maximizes an objective function, given by
-
-$$
-obj_p = \Phi \times scale(\sum_f^F \boldsymbol{\phi_{f,p}}) + (1 - \Phi)  scale(\boldsymbol{\sum_s^S\frac{ssb_s}{ssb0_{s,p}}})
-$$ {#eq-obj}
-
-Where *scale* rescales the values to be between 0 and 1 such that the
-weight parameter is comparable between vectors. The algorithm works by
-first assigning every patch in the system to an MPA. For the first
-round, the algorithm then calculates the equilibrium objective function
-value resulting from sequentially removing each individual patch from
-the MPA network. The cell whose removal from the MPA produces the
-highest objective function value is then removed from the MPA network,
-and the marginal value of the objective function each of the remaining
-cells from the last steps are then stored. For the next steps, we
-sampled 25% of the remaining MPA cells with probability proportional to
-their marginal contribution to the objective function in the prior
-iteration, selecting the cell whose removal produces the highest total
-objective function as the cell to remove from the network, and updating
-the marginal value of the sampled-but-not-selected cells. This in
-essence is a sampling-importance-resampling algorithm. The algorithm
-continues until no more cells are left in the MPA. This then provides
-for a given objective weighting value $\Phi$ the near-optimal MPA
-network at every MPA size from 0 to 100%. We say "near" as they is some
-path dependency here, in that for example the network that maximizes the
-objective function with 40% of the MPAs does not necessarily contain the
-optimal network at 30%. However, the reversed nature of the algorithm,
-starting at 100% MPA coverage and working better, produces a more stable
-network with regards to this problem than starting at 0% and growing to
-100%.
+<!-- ### MPA Optimization -->
+
+<!-- Our MPA optimization algorithm seems to create an MPA network that -->
+<!-- maximizes an objective function, given by -->
+
+<!-- $$ -->
+<!-- obj_p = \Phi \times scale(\sum_f^F \boldsymbol{\phi_{f,p}}) + (1 - \Phi)  scale(\boldsymbol{\sum_s^S\frac{ssb_s}{ssb0_{s,p}}}) -->
+<!-- $$ {#eq-obj} -->
+
+<!-- Where *scale* rescales the values to be between 0 and 1 such that the -->
+<!-- weight parameter is comparable between vectors. The algorithm works by -->
+<!-- first assigning every patch in the system to an MPA. For the first -->
+<!-- round, the algorithm then calculates the equilibrium objective function -->
+<!-- value resulting from sequentially removing each individual patch from -->
+<!-- the MPA network. The cell whose removal from the MPA produces the -->
+<!-- highest objective function value is then removed from the MPA network, -->
+<!-- and the marginal value of the objective function each of the remaining -->
+<!-- cells from the last steps are then stored. For the next steps, we -->
+<!-- sampled 25% of the remaining MPA cells with probability proportional to -->
+<!-- their marginal contribution to the objective function in the prior -->
+<!-- iteration, selecting the cell whose removal produces the highest total -->
+<!-- objective function as the cell to remove from the network, and updating -->
+<!-- the marginal value of the sampled-but-not-selected cells. This in -->
+<!-- essence is a sampling-importance-resampling algorithm. The algorithm -->
+<!-- continues until no more cells are left in the MPA. This then provides -->
+<!-- for a given objective weighting value $\Phi$ the near-optimal MPA -->
+<!-- network at every MPA size from 0 to 100%. We say "near" as they is some -->
+<!-- path dependency here, in that for example the network that maximizes the -->
+<!-- objective function with 40% of the MPAs does not necessarily contain the -->
+<!-- optimal network at 30%. However, the reversed nature of the algorithm, -->
+<!-- starting at 100% MPA coverage and working better, produces a more stable -->
+<!-- network with regards to this problem than starting at 0% and growing to -->
+<!-- 100%. -->
 
@@
 of management strategies.
+
+Each of the use cases contains too many parameters and options to be succinctly presented in the text here, and we are concerned that attempting to duplicate all of the parameters of the model here would simply introduce error.
+As such, readers should consult the accompanying code (provide clear instructions for which files) to view the precise details of each simulation.
+xx something like this to clarify why we don't have pages of tables with all the parameters and options).
+Targeted tactical applications though should be sure to consider and document all decisions made around model parameters.
 
@@
 In our coral reef example, we model the dynamics of four species: a
-grouper (XX), a shallow-reef snapper (xx), a deep-water snapper (XX),
-and a reef shark (XX). Groupers undergo a mass migration to a spawning
-aggregation once per year, followed by the sharks. Shallow-reef snappers
-stay in reefs closer to shower year-round, while deep-water snappers
-stay in the deeper reefs (@fig-coral).
+grouper (*Epinephelus fuscoguttatus*), a shallow-reef snapper (*Lutjanus malabaricus*), a deep-water snapper (*Pristipomoides filamentosus*), and
+a reef shark (*Carcharhinus amblyrhynchos*). The simulated groupers
+undergo a mass migration to a spawning aggregation once per year,
+followed by the sharks. Shallow-reef snappers stay in reefs closer to
+shower year-round, while deep-water snappers stay in the deeper reefs
+(@fig-coral).
 
@@
 and so short-term effort expands and contracts in respond to
-profitability of the primarily grouper-driven fishery xx.
+profitability of the primarily grouper-driven fishery.
+
+Each fleet has a different home-port location, and incurs an additional
+cost for fishing each patch in the system in proportion to the euclidean
+distance from each patch to each fleet's home port.
+
+We then simulated the outcomes for both food security and conservation
+for each of the species and both of the fleets as a function of both MPA
+size and MPA design strategy. For this exercise, MPAs are placed with
+perfect information and have no design constraints for continuity (i.e.
+the selected MPA design can be as patchwork as dictated by the data).
 
@@
 #| label: fig-coral
-#| fig-cap: "Distribution of animals each season under unfished conditions. Numbers show the approximate location of each fleet's port."
+#| fig-cap: "Distribution of fish spawning biomass (color) each season (row) under unfished conditions.
+The x-axis represents longitude, y-axis latitude.
+Numbers show the approximate location of each fleet's port.
+Rows indicate quarterly seasons that repeat each year."
 
@@
         axis.ticks = element_blank(),
-        axis.title = element_blank())
+        axis.title = element_blank(),
+        axis.line = element_blank())
 
@@
 
-We then simulate the impacts of MPAs...
-
 #### Pelagic Systems
+
+```{r}
+blue_fauna <- blue_water_fauna_and_fleets$fauna
+
+blue_fleets <- blue_water_fauna_and_fleets$fleets
+
+blue_resolution <- sqrt(blue_fauna[[1]]$patches)
+
+```
 
@@
 Western and Central Pacific Ocean tuna fisheries. We simulate
-trajectories of XX species commonly caughts in the region, including
-both the highly abundance skipjack tuna (*Katsuwonus pelamis*) to the
-heavily depleted oceanic whitetip shark (*Carcharhinus longimanus*). We
-pulled publicly available data on catch-per-unit-effort of each of these
-species from the WCPO as a proxy for baseline habitat distributions
-(Fig.XX). However, for the pelagic case study we simulate a scenario
-where climate change is inducing a northerly migration of XX degress per
-year in XX species.
-
-These species are targeted by a longline fleet that targets large adult
-tunas such as bigeye and yellowfin primarily for high-grade consumption,
-and a purse-seine fishery that primarily targets skipjack tunas for bulk
-canning. XX. Both fleets are subject to a quota XX add this back in XX.
+trajectories of `r n_distinct(names(blue_fauna))` species commonly
+caught in the region, including both the highly abundance skipjack tuna
+(*Katsuwonus pelamis*) to the heavily depleted oceanic whitetip shark
+(*Carcharhinus longimanus*) (@fig-blue-ssb). We pulled publicly
+available data on catch-per-unit-effort of each of these species from
+the WCPO as a proxy for baseline habitat distributions (@fig-blue).
+
+These pelagic species are caught by a longline fleet that primarily
+targets large adult tunas such as bigeye and yellowfin for high-grade
+consumption, and a purse-seine fishery that primarily targets skipjack
+tunas for bulk canning. Selectivities were modeled as logistic for the
+longline fleet, and dome-shaped for the purse-seine fleet, with
+selectivity values pulled from the literature where available (see SI
+xx). Both fleets operate under open-access dynamics with an effort cap.
+The effort cap was set at the level of effort that resulted in the
+desired levels of depletion for each species under open-access dynamics
+(@fig-blue), intended to simulate a scenario where managers step in to
+prevent further expansion of fishing effort in a fully developed
+fishery. For forward-simulation, open-access dynamics can result in
+effort decreasing in response to profitability, but cannot result in
+effort beyond the effort cap set for each fleet.
+
+For this exercise, we focused on using `marlin` to assess resilience of
+the selected "rule of thumb" MPA design strategies to a climate-driven
+range shift. Specifically, we simulate an extreme example where the
+centroid of each population shifts northward at a rate of 0.2 patches
+per year, over a `r experiment_years` time horizon. We designed MPA
+networks given the conditions in the starting year, and held that
+network constant over the years of the experiment, running one
+simulation with and another without the climate-drive range shift. We
+then compared effect of this range shift on food security and
+conservation outcomes from MPA networks designed based on the pre-range
+shift world.
 
 ```{r}
-#| label: fig-blue
-#| fig-cap: "Distribution of animals under unfished conditions for blue-water simulation."
-
-blue_fauna <- blue_water_fauna_and_fleets$fauna
-
-blue_fleets <- blue_water_fauna_and_fleets$fleets
-
-grid <- expand_grid(x = 1:resolution, y= 1:resolution) %>%
+#| label: fig-blue-ssb
+#| fig-cap: "Trajectory of depletion (spawning biomass divided by unfished spawning biomass) over time prior to implementation of MPAs."
+
+grid <- expand_grid(x = 1:blue_resolution, y= 1:blue_resolution) %>%
   mutate(patch = 1:nrow(.))
 
-mpa_locations <- expand_grid(x = 1:resolution, y = 1:resolution) %>%
+mpa_locations <- expand_grid(x = 1:blue_resolution, y = 1:blue_resolution) %>%
   mutate(mpa = TRUE)
@@
   fleets = blue_fleets,
+  years = 50
+)
+
+critter_biomass <-
+  map_df(blue_sim, ~ map_df(.x, ~tibble(ssb = rowSums(.x$ssb_p_a), patch = 1:nrow(.x$ssb_p_a)), .id = "critter"), .id = "step") %>%
+  mutate(step = as.numeric(step)) %>%
+  left_join(grid, by = "patch") %>% 
+  group_by(critter, step) %>% 
+  summarise(ssb = sum(ssb)) %>% 
+  group_by(critter) %>% 
+  mutate(depletion = ssb / ssb[step == 0])
+
+critter_biomass %>% 
+  ggplot(aes(step, depletion)) +
+  geom_line() + 
+  facet_wrap(~critter) +
+  scale_y_continuous(limits = c(0, 1))
+
+```
+
+```{r}
+#| label: fig-blue
+#| fig-cap: "Distribution of fish spawning biomass under status quo (blue) and range-shifted (red) conditions for blue-water simulation. The x-axis represents longitude, the y-axis latitude. "
+
+blue_climate_sim <- simmar(
+  fauna = blue_fauna,
+  fleets = blue_fleets,
+  habitat = future_habitat,
+  years = 20,
   manager = list(mpas = list(
@@
     mpa_year = 1
-  )),
-  years = 2
+  ))
 )
@@
 patch_biomass <-
-  map_df(blue_sim, ~ map_df(.x, ~tibble(biomass = rowSums(.x$ssb_p_a), patch = 1:nrow(.x$ssb_p_a)), .id = "critter"), .id = "step") %>%
+  map_df(blue_climate_sim, ~ map_df(.x, ~tibble(biomass = rowSums(.x$ssb_p_a), patch = 1:nrow(.x$ssb_p_a)), .id = "critter"), .id = "step") %>%
   mutate(step = as.numeric(step)) %>%
@@
 patch_biomass %>%
+  left_join(sci_to_com, by = "critter") %>% 
+    select(-critter) %>% 
+  rename(critter = common_name) %>% 
+  filter(step == min(step) | step == max(step)) %>% 
   group_by(critter, step) %>%
@@
   ungroup() %>%
-  filter(step == max(step)) %>% 
   ggplot()+
-  geom_tile(aes(x,y, fill = sbiomass),show.legend = FALSE) +
+  geom_contour(aes(x,y, z = sbiomass, color= factor(step)), linewidth = .1) +
   facet_wrap(~critter) +
@@
   scale_y_continuous(expand = c(0,0)) +
-  scale_fill_viridis_c() + 
   theme(axis.text = element_blank(),
         axis.ticks = element_blank(),
-        axis.title = element_blank())
+        axis.title = element_blank(),
+        legend.position = "top",
+        axis.line = element_blank()) + 
+  scale_color_manual(values = c("#63B8FF","#FF4500"), labels = c("Status Quo","Range Shift"), name = '')
 
@@
 
-#### Key Drivers
-
-The MPA literature points to the ability of a wide range of drivers to
-affect the performance of MPAs with regards to conservation and economic
-outcomes such as adult and larval dispersal [@gerber2005], the timing of
-density dependence [@babcock2011], correlations in space and time across
-species [@pons2022; @abbott2012] , fleet dynamics and economic
-incentives [@smith2003a; @cabral2019], relationships between size and
-fecundity (e.g. hyperallometry) [@marshall2021], heterogeneity in stock
-status [@hastings2017; @abbott2012], and many other traits [@gaines2010;
-@hilborn2004a].
-
-These studies have all helped guide the science of marine protected
-areas, but often focus on modeling the impact of subset of variables of
-interest at a time. We demonstrate here how `marlin` can be used to
-conduct Monte Carlo simulations across a wide range of drivers to assess
-the relative contributions each to the outcomes of protected areas under
-a particular set of circumstances.
-
-For this example we model the dynamics of a skipjack tuna *Katsuwonus
-pelamis*, a striped marlin *Kajikia audax*, and a oceanic whitetip shark
-*Carcharhinus longimanus*. Life history parameters for each species were
-pulled from FishLife (@thorson2020), with the exception of steepness,
-set to XX, and fecundity. For the tuna and marlin species, fecundity is
-assumed to scale with weight modified by the hyperallometric exponent.
-For the shark species, we assume a constant value of six pups per year
-at all ages greater than the age at maturity. We assume that tunas have
-an unfished biomass of XX, marlins XX, and sharks XX, to reflect a
-large, medium, and small population.
-
-For the generalized experiments we only include one fishing fleet. We
-assign a price per KG of \$XX for tuna, intended to reflect a low value
-high volume fishery, \$XX for marlin for a higher value low volume
-fishery, and \$0 for sharks to reflect a true bycatch species. We assign
-a selectivity of 60% of length at maturity for skipjack to allow for
-some overfishing when F/M is high enough, 70% for marlin, and 90% for
-shark xx.
-
-We ran `r nrow(state_experiments)` state experiments, varying the state
-of nature experienced by a simulation. The candidate states include
-
-| State                  | Distribution          | Description                                     |
-|-------------------|-------------------|-----------------------------------|
-| $\sigma_{centroid}$    |                       | Standard deviation of habitat centroid          |
-| $\sigma_{habitat}$     |                       | Standard deviation around centroid              |
-| Spatial catchability   |                       | Catchability varies in space                    |
-| Spatial allocation     |                       | Fleet allocation strategy                       |
-| F/M                    |                       | Fishing mortality relative to natural mortality |
-| Adult diffusion rate   |                       |                                                 |
-| Recruit diffusion rate |                       |                                                 |
-| Hyperallometry         | $U(1,2)$              | Presence of hyperallometric fec                 |
-| DD timing              |                       | Timing of density dependence                    |
-| Seasonal Movement      | $\subset[TRUE,FALSE]$ | Seasonal Movement Present                       |
-
-We simulate degrees of species overlap using two parameters, $\sigma_c$
-and $\sigma_h$. All species start with the assumption that the centroid
-of their preferred habitat occurs in the center-most patch of the two
-dimensional system simulated by marlin. For our simulations, we then
-generated a random deviation from that default centroid for each species
-using $\sigma_c$ in the manner
-
-$$centroid_s \sim N(center,\sigma_c)$$ where *s* denotes species, and
-*center* is the coordinate of the central cell. In this way, the larger
-$\sigma_c$ grows, the more variation there is between the centroid of
-the preferred habitat for each species.
-
-We then control the spread of a creatures preferred habitat around that
-centroid per a Gaussian dispersal kernel in the manner
-
-$$habitat_{p,s} = \frac{1}{\sqrt{2\pi\sigma_h^2}}e^{-\frac{d_p^2}{2\sigma_h^2}}$$
-
-Where $d$ is the euclidean distance from the $centroid_s$ to patch $p$
-and $\sigma_h$ controls the dispersal of the habitat around that range.
-
-For each of these `r nrow(state_experiments)` state experiments, we then
-apply `r nrow(placement_experiments)` placement experiments defined by
-an MPA size, an MPA placement strategy, and the number of animals to be
-considered in the MPA planning, ranging from one to three, where one
-indicates that only one animal is considered in the MPA design process,
-while three means that all three species are considered.
-
 ## Results
@@
 ### Coral Reefs
+
+For the coral reef example, both the area and target_fishing MPA design
+strategies were capable of producing a Pareto-improving outcome where
+the total spawning biomass divided by unfished spawning biomass across
+all species could be increased by up to 200% while maintaining at least
+the same levels of fisheries yields as would be caught in the absence of
+any MPAs. Conversely, achieving increases in total fish biomass through
+the rate or avoid fishing strategies resulted in immediate losses in
+total yields, and in losses to both yields and biomass in some cases
+(@fig-coral-results).
+
+However, these macro-level changes in total yield and conservation
+outcomes mask some internal tradeoffs. Nearly all of the food security
+benefits from this MPA network accrued to fleet one, with catches by
+fleet two being negatively impacted under nearly all scenarios,
+including those that resulted in the highest total gains to fleet one
+and subsequently to the system as a whole. For example, the MPA sizes
+and designs that result in the largest increase in yields for fleet one
+would produce in near total loss of yields for fleet two.
+
+Both the *rate* and *avoid fishing* MPA placement strategies produced
+lower total fishery gains but minimized the negatiev impacts on fleet
+two. However, these strategies produced much lower conservation gains
+than the *area* and *target_fishing* strategies, which produced minimal
+gains for any species besides deep snapper, and in fact resulting in
+slight losses to the nearshore snapper population for MPA sizes less
+than roughly 75%. Both the area and target_fishing strategies resulted
+in rapid conservation gains across all of the simulated species.
+
+MPAs were capable of producing positive outcomes in terms of total
+biomass and yields in this coral reef example, but none of the
+strategies on the table came without tradeoffs across species or fleets.
+The framework we present here could be further used then by a community
+such as the one simulated here to explore new design strategies, or to
+assess options for distributing fishery benefits across communities that
+may have very different outcomes from the same MPA network.
 
@@
 #| label: fig-coral-results
-#| fig-cap: "Frontier in conservation and yield outcomes under different MPA design strategies (top panel) for coral reef case study.
-Change in yield for each fleet as a function of MPA size and MPA design strategy (bottom panel) for coral reef case study."
-#| fig-height: 8
+#| fig-cap: "Frontier in conservation and yield outcomes under different MPA design strategies (A) for coral reef case study.
+Change in yield for each fleet (B) and depltion per species (C) as a function of MPA size and MPA design strategy () for coral reef case study."
+#| fig-height: 4
 #| fig-width: 6
@@
 
-coral_yield <- coral_results %>% 
+coral_frontier <- coral_results %>% 
   group_by(prop_mpa, placement_strategy) %>% 
@@
   geom_hline(aes(yintercept = 0), linetype = 2) +
-  geom_line(size = 2) + 
-  scale_x_continuous(name = "Change in Conservation", labels = scales::label_percent(accuracy = 1)) +
+  geom_line() + 
+  scale_x_continuous(name = "Change in Total Depletion", labels = scales::label_percent(accuracy = 1), guide = guide_axis(n.dodge = 2)) +
   scale_y_continuous(name = "Change in Yield", labels = scales::label_percent(accuracy = 1)) + 
-  theme(legend.position = "top")
+  scale_color_ordinal(name = '')+
+  guides(color = guide_legend(nrow = 4)) + 
+  labs(tag = "A") + 
+  theme(
+    legend.position = c(0.7, 1),
+    legend.background = element_rect(fill = "transparent"),
+    legend.text = element_text(size = 8)
+  )
 
@@
   geom_line() +
-  facet_wrap( ~ placement_strategy, scales = "free_y") +
+  facet_wrap( ~ placement_strategy) +
   scale_x_continuous(name = "MPA Size",
                      labels = scales::label_percent(accuracy = 1)) +
-  scale_y_continuous(name = "Change in Yield",
-                     labels = scales::label_percent(accuracy = 1))
-
-coral_yield / coral_fleet_yield
+  scale_y_continuous(name = "Yield",
+                     labels = scales::label_percent(accuracy = 1)) + 
+  theme(axis.text.x = element_blank(),
+        axis.title.x = element_blank(),
+        axis.text.y = element_text(size = 8))  + 
+    scale_color_discrete(name = '') + 
+    labs(tag = "B")
+
+coral_critter_bio <- coral_results %>%
+  group_by(prop_mpa, critter, placement_strategy) %>%
+  summarise(biodiv = sum(unique(biodiv))) %>%
+  group_by(critter, placement_strategy) %>%
+  mutate(delta_bio = biodiv) %>%
+  # filter(prop_mpa < 0.4) %>% 
+  ggplot(aes(prop_mpa, pmin(Inf,delta_bio), color = critter)) +
+  geom_line() +
+  facet_wrap( ~ placement_strategy) +
+  scale_x_continuous(name = "MPA Size",
+                     labels = scales::label_percent(accuracy = 1),
+                     guide = guide_axis(n.dodge = 2)) +
+  scale_y_continuous(name = "Depletion") + 
+  scale_color_viridis_d(name = '', option = "C") + 
+    labs(tag = "C") + 
+  theme(axis.text = element_text(size = 8))
+
+((coral_frontier + plot_layout(guides = "keep")) | (coral_fleet_yield / coral_critter_bio) + plot_layout(guides = "collect")) & theme(strip.text = element_text(size  = 8), plot.margin = margin(.1, .1, .1, .1, "cm"), legend.text = element_text(size = 7))
 
@@
 
+For the pelagic case study we focused on the climate resilience of two
+candidate MPA design strategies; target_fishing and rate, where both MPA
+networks are designed based on the species distributions at the start of
+the time series. We then compared the outcomes of these MPA networks for
+fisheries yield and spawning biomass under the two scenarios where the
+species distributions remain constant, and when the species undergo a
+range shift.
+
+The yield impacts of the proposed MPA strategies were much more
+sensitive to climate impacts for the purse seine fishery; under the
+*status quo* species distribution, the target fishing MPA design
+strategy produced increases in fisheries yields at sizes between 25-50%
+of the area; that same MPA network resulted in the near complete loss of
+yields from the purse seine fleet when applied to the climate change
+scenario. In contrast, the yield outcomes for the longline fleet were
+relatively insensitive to the simulated range shifts
+(@fig-blue-results).
+
+From a conservation perspective, species such as albacore and swordfish
+were relatively insensitive to the simulated range shifts, whereas
+species such as bigeye tuna and silky shark showed extreme variation in
+outcomes depending on the climate scenario. For bigeye tuna, under the
+*status quo* conditions the target fishing MPA design strategy would
+result in rapid increases in the population, but that same MPA network
+would produce much smaller and slower gains under climate change.
+However, for blue sharks this trend is flipped, with more rapid
+conservation gains under the range shift scenario than the *status quo*
+(@fig-blue-results).
+
 ```{r}
-#| label: fig-pelagic
-#| fig-cap: "Placeholder for pelagic results is we stick with them"
+#| include: false
+
+blue_water_experiments$mpas <- map(blue_water_experiments$results, "mpa")
+
+blue_water_experiments$obj <- map(blue_water_experiments$results, "obj")
 
@@
 
-blue_results <- blue_water_climate_experiments %>%
-  unnest(cols = obj)
-
-blue_fleet_frontier <- blue_results %>%
+static_results <- blue_water_experiments %>%
+  unnest(cols = obj) %>%
+  mutate(climate = FALSE)
+
+climate_results <- blue_water_climate_experiments %>%
+  unnest(cols = obj) %>%
+  mutate(climate = TRUE)
+
+blue_water_results <-static_results %>%
+  bind_rows(climate_results) %>% 
+  left_join(sci_to_com, by = "critter") %>% 
+  select(-critter) %>% 
+  rename(critter = common_name)
+
+fleet_climate_results <- blue_water_results %>%
+  select(placement_strategy,
+         fleet_model,
+         prop_mpa,
+         climate,
+         longline,
+         critter,
+         purseseine) %>%
+  pivot_longer(c(longline,purseseine), names_to = "fleet", values_to = "yield") %>%
+  pivot_wider(names_from = "climate", values_from =  "yield") %>%
+  mutate(delta = `TRUE` - `FALSE`) %>%
+  group_by(placement_strategy, prop_mpa, fleet) %>%
+  summarise(delta = sum(delta))
+
+fleet_climate_impacts <- fleet_climate_results %>%
+  ggplot(aes(prop_mpa, delta, color = fleet)) +
+  geom_hline(yintercept = 0, linetype = 2) +
+  geom_line() +
+  facet_wrap(~placement_strategy) +
+  scale_y_continuous(name = "Difference in Yield") +
+  theme(axis.text.x = element_blank(), axis.title.x = element_blank())
+
+critter_climate_results <- blue_water_results %>%
+  select(placement_strategy,
+         fleet_model,
+         prop_mpa,
+         climate,
+         critter,
+         biodiv) %>%
+  pivot_wider(names_from = "climate", values_from =  "biodiv") %>%
+  mutate(delta = `TRUE` - `FALSE`)
+
+critter_climate_impacts <- critter_climate_results %>%
+  ggplot(aes(prop_mpa, delta, color = critter)) +
+  geom_line() +
+  facet_wrap(~placement_strategy) +
+  scale_x_continuous(name = "MPA",labels = scales::label_percent()) +
+  scale_y_continuous(name = "Difference in Depletion")
+
+fleet_climate_impacts / critter_climate_impacts
+
+```
+
+```{r}
+#| label: fig-blue-results
+#| fig-cap: "Change in yield (A) and spawning biomass (B) as a function of MPA size and design strategy by fishing fleet and species.
+Blue lines indicate impacts of MPAs under status qup habitat, red impacts under climate-driven range shift."
+#| fig-height: 6
+#| fig-width: 6
+
+linesize <- 0.5
+
+# blue_water_results <- blue_water_results %>% 
+#   rename(`purse seine` = purseseine)
+
+blue_fleet_frontier <- blue_water_results %>%
   pivot_longer(names(blue_fleets), names_to = "fleet", values_to = "fleet_yield") %>%
-  group_by(prop_mpa, fleet, placement_strategy) %>%
+  group_by(prop_mpa, fleet, placement_strategy, climate) %>%
   summarise(yield = sum(fleet_yield),biodiv = sum(unique(biodiv))) %>%
-  ggplot(aes(biodiv, yield, color = placement_strategy)) +
+  ggplot(aes(biodiv, yield, color = placement_strategy, linetype = climate)) +
   geom_line() +
@@
 
-blue_frontier <- blue_results %>% 
-  group_by(prop_mpa, placement_strategy) %>% 
+blue_frontier <- blue_water_results %>% 
+  filter(climate == FALSE) %>% 
+  group_by(prop_mpa, placement_strategy, climate) %>% 
   summarise(biodiv = sum(biodiv), yield = sum(yield)) %>% 
@@
          delta_biodiv = biodiv / biodiv[prop_mpa == 0] - 1) %>% 
-  ggplot(aes(delta_biodiv, delta_yield, color = placement_strategy)) + 
+  ggplot(aes(delta_biodiv, delta_yield, color = placement_strategy, linetype = climate)) + 
   geom_vline(aes(xintercept = 0), linetype = 2) +
   geom_hline(aes(yintercept = 0), linetype = 2) +
-  geom_line(size = 2) + 
+  geom_line(size = linesize) + 
   scale_x_continuous(name = "Change in Conservation", labels = scales::label_percent(accuracy = 1)) +
@@
 
-blue_fleet_yield <- blue_results %>%
+blue_fleet_yield <- blue_water_results %>%
+  filter(placement_strategy %in% c("target_fishing", "rate")) %>%
   pivot_longer(names(blue_fleets),
@@
                values_to = "fleet_yield") %>%
-  group_by(prop_mpa, fleet, placement_strategy) %>%
+  group_by(prop_mpa, fleet, placement_strategy, climate) %>%
   summarise(yield = sum(fleet_yield), biodiv = sum(unique(biodiv))) %>%
@@
   mutate(delta_yield = yield / yield[prop_mpa == 0] - 1) %>%
-  ggplot(aes(prop_mpa, delta_yield, color = fleet)) +
+  ggplot(aes(prop_mpa, delta_yield, color = climate)) +
   geom_hline(aes(yintercept = 0), linetype = 2) +
-  geom_line() +
-  facet_wrap( ~ placement_strategy, scales = "free_y") +
+  geom_line(size = linesize, show.legend = FALSE) +
+  facet_grid(fleet ~ placement_strategy, scales = "free_y") +
   scale_x_continuous(name = "MPA Size",
@@
   scale_y_continuous(name = "Change in Yield",
-                     labels = scales::label_percent(accuracy = 1))
-
-blue_critter_bio <- blue_results %>%
-  group_by(prop_mpa, critter, placement_strategy) %>%
+                     labels = scales::label_percent(accuracy = 1)) +
+  scale_color_manual(
+    values = c("#63B8FF", "#FF4500"),
+    labels = c("Status Quo", "Range Shift"),
+    name = ''
+  ) + 
+  theme(strip.text = element_text(size = 8)) + 
+  labs(tag = "A")
+
+blue_critter_bio <- blue_water_results %>%
+    filter(placement_strategy %in% c("target_fishing", "rate")) %>% 
+  group_by(prop_mpa, critter, placement_strategy, climate) %>%
   summarise(biodiv = sum(unique(biodiv))) %>%
   group_by(critter, placement_strategy) %>%
-  mutate(delta_bio = biodiv / biodiv[prop_mpa == 0] - 1) %>%
-  filter(prop_mpa < 0.4) %>% 
-  ggplot(aes(prop_mpa, pmin(1,delta_bio), color = critter)) +
-    geom_hline(aes(yintercept = 0), linetype = 2) +
-  geom_line() +
-  facet_wrap( ~ placement_strategy) +
+  mutate(delta_bio = biodiv) %>%
+  ggplot(aes(prop_mpa, pmin(1,delta_bio), color = climate, linetype = placement_strategy)) +
+  geom_line(size = linesize) + 
+  facet_wrap(~critter) +
+  scale_color_manual(values = c("#63B8FF","#FF4500"), labels = c("Status Quo","Range Shift"), name = '') +
   scale_x_continuous(name = "MPA Size",
-                     labels = scales::label_percent(accuracy = 1)) +
-  scale_y_continuous(name = "Change in Spawning Biomass",
-                     labels = scales::label_percent(accuracy = 1))
+                     labels = scales::label_percent(accuracy = 1), guide = guide_axis(n.dodge = 2)) +
+  scale_y_continuous(name = "Change in Spawning Biomass",limits = c(0, NA)) +
+  scale_linetype(name = "MPA Strategy") + 
+    labs(tag = "B")+
+  theme(strip.text = element_text(size = 7),
+        axis.text.y = element_text(size = 6))
 
 # blue_frontier / blue_critter_bio + plot_layout(heights = c(1,2))
-
-blue_critter_bio
+(blue_fleet_yield / 
+blue_critter_bio) + plot_layout(heights = c(1.5,2), guides = "collect" )
 ```
 
-### Key Drivers
-
-```{r}
-#| label: fig-drivers
-#| fig-cap: "Distribution of conservation and food security impacts of MPAs covering 30% of available habitat across a range of simulations."
-#| fig-height: 7
-#| fig-width: 7
-
-var_effects <- driver_results %>%
-  ungroup() %>%
-  mutate(biodiv_effect = biodiv / bau_biodiv - 1,
-         econ_effect = econ / bau_econ - 1) %>%
-  select(
-    biodiv_effect,
-    f_v_m,
-    econ_effect,
-    critter,
-    adult_diffusion,
-    recruit_diffusion,
-    density_dependence,
-    placement_strategy,
-    sigma_habitat,
-    hyper,
-    seasonal_movement,
-    fleet_model,
-    spawning_aggregation,
-    seasonal_movement,
-    prop_mpa
-  )
-
-continuous <- map_lgl(var_effects, is.numeric)
-
-bio_effects <- var_effects[,continuous] %>% 
-  filter(between(prop_mpa,0.29, .31)) %>%
-  select(-econ_effect, - prop_mpa) %>% 
-  pivot_longer(-c(contains("_effect")),names_to = "variable", values_to = "value")
-
-econ_effects <- var_effects[,continuous] %>% 
-  filter(between(prop_mpa,0.29, .31)) %>%
-  select(-biodiv_effect, - prop_mpa) %>% 
-  pivot_longer(-c(contains("_effect")),names_to = "variable", values_to = "value")
-
-biodiv_effects_plot <- bio_effects %>%
-  ggplot(aes(value, pmin(2.5,biodiv_effect))) +
-  geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
-  geom_bin2d(show.legend = FALSE) +
-  scale_fill_viridis_c() +
-  scale_y_continuous(name = "Conservation Impact", labels = scales::label_percent(accuracy = 1)) + 
-  scale_x_continuous(name = "") +
-  facet_wrap(~variable, scales = "free_x") +
-  theme(axis.text.x = element_blank(),
-        plot.margin = unit(c(0,0,0,0), units = "points"))
-
-econ_effects_plot <- econ_effects %>%
-  ggplot(aes(value, pmax(-1,econ_effect))) +
-  geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
-  geom_bin2d(show.legend = FALSE) +
-  scale_fill_viridis_c() +
-  scale_y_continuous(name = "Food Impact", labels = scales::label_percent(accuracy = 1)) + 
-  scale_x_continuous(name = "Metric Value") +
-  facet_wrap(~variable, scales = "free_x")
-
-biodiv_effects_plot / econ_effects_plot & theme(strip.text = element_text(size = 8))
-
-```
-
 ## Discussion
 
-Waiting on a fleshed out discussion until we are all happy with the
-content, but in general
-
--   Growing need for more spatially explicit and ecosystem-based
-    management and assessment
-
--   `marlin` fills an important gap in the modeling literature
-
-    -   MPAs
-    -   Policy interaction
-    -   spatial-temporal MSE
-    -   Range shifts
-    -   make clear it's not just an MPA model
-    -   Draw contrast to virtual species and STEPS
-
--   Examples demonstrate this
-
-    -   total benefits and Distributional impacts in coral case
-    -   Something in the pelagic case
-    -   Complex interactions in the final case
-
--   `marlin` and the methods within as a platform for future practical
-    and academic bio-economic modeling needs
+Marine ecosystems are driven by complex social-ecological dynamics.
+Communities must often make decisions on how to manage these systems
+based on ... Modeling tools such as `marlin` can help users answer
+important scientific questions and design marine management policies
+informed by better understanding of sensitivities to key uncertainties.
+blah blah blah. `marlin` is to our knowledge the first bio-economic
+modeling tool that allows scientists and decision makers to efficiently
+simulate the outcomes of management strategies such as MPAs across
+multiple fleets and species while accounting for layered regulations,
+economic incentives, age structure, and realistic movement dynamics
+across a range of potential social-ecological conditions. xx this
+sentence annoys me but feels like we need to make something like it
+explicit to convince reviewers maybe less familiar with the modeling
+literature xx.
+
+There are a range of other useful modeling applications that can support
+spatio-temporal decision making in marine ecosystems. However, the
+`marlin` framework presented here fills an important gap in the modeling
+literature and in the policy support toolbox. More end-to-end models
+such as ATLANTIS [@audzijonyte2019] , OSMOSE [@shin2001], Ecopath with
+Ecosim [@christensen2004], POSEIDON [@bailey2019], or SEAPODYM
+[@lehodey2008] are capable of representing tremendous amounts of
+complexity, but can be extremely time consuming to design and run. Tools
+such as virtualspecies [@leroy2016] or STEPS [@visintin2020] are help
+model the distribution of species in space and time as a function of
+environmental covariates, but lack explicit process models for the xx
+Jim, can you expand on this section, I'm no too familiar with these
+methods xx. Simplified bio-economic models such as those used in
+@hastings2017, @sala2021, and @cabral2019 can be applied at scale and
+provide analytically tractable results, but as a result must abstract
+over many bio-economic dynamics that can be important for more tactical
+applications. `marlin` complements the existing spatial marine modeling
+literature by allowing scientists, decision makers, and other
+stakeholders to efficiently and productively examine the impacts of
+realistic bio-economic dynamics on academic and applied problems.
+
+In this paper we have documented the core system of equations underlying
+`marlin`, and demonstrated how this framework can be used to support
+efforts such as MPA planning. The variation in both the magnitude and
+distribution of costs and benefits stemming from alternative design
+choices and states of nature shown in our results demonstrates the value
+in our method. Beyond the current examples, we envision `marlin` being
+applicable to research on dynamic ocean management, range shifts,
+management strategy evaluation, policy interactions, and spatial stock
+assessment. Future iterations of the model may include life history
+modules for a broader range of species types such as marine mammals, sea
+birds, and sea turtles.
+
+Even the most complex marine model is a stylized cartoon of the true
+dynamics at play in these systems. However, for all their limitations
+models such as `marlin` can help users understand factors that drive the
+performance of potential marine management strategies. The goal of this
+tool is to empower people to design policies based on evaluation of key
+uncertainties and trade-offs, and in doing so support more effective and
+equitable marine resource management.
 

### 2022-11-17 R/create_critters.R

@@ -5,7 +5,9 @@ create_critters <-
            marlin_inputs,
            seasonal_movement = FALSE,
            ontogenetic_shift = FALSE,
-           random_rec = FALSE) {
+           random_rec = FALSE,
+           adult_diffusion = 10,
+           taxis_to_diff_ratio = 2) {
     # sciname <- marlin_inputs$scientific_name[[1]]
     #
     #
@@ -33,13 +35,11 @@ create_critters <-
     }
 
     if (random_rec){
-    rec_form <-  sample(c(0, 1,2,3), 1, replace = TRUE)
+      density_dependence <-  sample(c("global_habitat", "local_habitat","pre_dispersal","post_dispersal","global_ssb"), 1, replace = TRUE)
 
     } else {
-      rec_form <- 2 # local density dependence then disperse recruits per recruit movement
+      density_dependence <- "global_ssb" # local density dependence then disperse recruits per recruit movement
     }
-    # rec_form <- 1 # sample(c(0, 1, 2, 3), 1, replace = TRUE)
-
 
     if (ontogenetic_shift) {
       recruit_habitat <-
@@ -47,7 +47,7 @@ create_critters <-
 
       # set recruitment form to allow for recruit habitat
       #
-      rec_form <- 0
+      density_dependence <- "global_ssb"
     } else {
       recruit_habitat <- hab[[1]]
     }
@@ -56,9 +56,10 @@ create_critters <-
       scientific_name = sciname,
       base_habitat = hab,
       recruit_habitat = recruit_habitat,
-      adult_diffusion = 4,
-      recruit_diffusion = 2,
-      rec_form = rec_form,
+      adult_diffusion = adult_diffusion,
+      taxis_to_diff_ratio = taxis_to_diff_ratio,
+      recruit_diffusion = 0,
+      density_dependence = density_dependence,
       fec_form = ifelse(str_detect(sciname,"carcharhinus"),"pups","weight"),
       pups = 6,
       seasons = seasons,

### 2022-11-17 R/run_mpa_experiment.R

@@ -13,7 +13,8 @@ run_mpa_experiment <-
            max_delta = 1,
            resolution,
            years = 50,
-           future_habitat = list()) {
+           future_habitat = list(),
+           effort_cap) {
 
     options(dplyr.summarise.inform = FALSE)
 
@@ -178,12 +179,45 @@ run_mpa_experiment <-
       fleets = fleets,
       years = years,
       manager = list(mpas = list(locations = mpas,
-                  mpa_year = ceiling(starting_step))),
+                  mpa_year = ceiling(starting_step)),
+                  effort_cap = effort_cap),
       habitat = future_habitat,
       starting_step = starting_step,
       keep_starting_step = FALSE,
       initial_conditions = starting_conditions[[length(starting_conditions)]]
     )
+
+
+    # grid <- expand_grid(x = 1:resolution, y= 1:resolution) %>%
+    #   mutate(patch = 1:nrow(.))
+    #
+    # patch_biomass <-
+    #   map_df(mpa_sim, ~ map_df(.x, ~tibble(biomass = rowSums(.x$ssb_p_a), patch = 1:nrow(.x$ssb_p_a)), .id = "critter"), .id = "step") %>%
+    #   mutate(step = as.numeric(step)) %>%
+    #   left_join(grid, by = "patch")
+
+    # patch_biomass %>%
+    #   group_by(critter, step) %>%
+    #   mutate(sbiomass = biomass / sum(biomass)) %>%
+    #   ungroup() %>%
+    #   filter(step == max(step)) %>%
+    #   ggplot(aes(x,y, fill = sbiomass))+
+    #   geom_tile() +
+    #   facet_grid(critter~step) +
+    #   scale_fill_viridis_c()
+    #
+    #
+    # browser()
+    #
+    #
+    # effort <-
+    #   map_df(mpa_sim, ~ data.frame(effort = sum(.x$`katsuwonus pelamis`$e_p_fl$longline)), .id = "step") %>%
+    #   mutate(step = as.numeric(step))
+    #
+    # effort %>%
+    #   ungroup() %>%
+    #   ggplot(aes(step, effort)) +
+    #   geom_line()
 
 
     # process results

### 2022-11-17 cpue_and_yield.R

@@ -1,0 +1,202 @@
+mpa_locations <- expand_grid(x = 1:resolution, y = 1:resolution) %>%
+  mutate(mpa = x < 0.3 * resolution)
+
+# fleets$longline$fleet_model
+#
+# fleets$purseseine$fleet_model
+
+with_mpa <- simmar(
+  fauna = fauna,
+  fleets = fleets,
+  manager = list(mpas = list(
+    locations = mpa_locations,
+    mpa_year = round(years / 2)
+  )),
+  years = years
+)
+
+grid <- expand_grid(x = 1:resolution, y= 1:resolution) %>%
+  mutate(patch = 1:nrow(.))
+
+
+patch_effort <-   map_df(with_mpa, ~ map_df(.x, ~.x$e_p_fl %>% mutate(patch = 1:nrow(.)), .id = "critter"), .id = "step") %>%
+  mutate(step = as.numeric(step)) %>%
+  left_join(grid, by = "patch") %>%
+  filter(critter ==  names(fauna)[1]) %>%
+  pivot_longer(names(fleets), names_to = "fleet", values_to = "effort")
+
+patch_effort %>%
+  filter(step == max(step)) %>%
+  ggplot(aes(x,y, fill = effort))+
+  geom_tile() +
+  facet_wrap(fleet~step) +
+  scale_fill_viridis_c()
+
+
+patch_catch <-
+  map_df(with_mpa,
+         ~ map_df(.x, ~ as_tibble(.x$c_p_fl) %>% mutate(patch = 1:nrow(.)), .id = "critter"),
+         .id = "step") %>%
+  mutate(step = as.numeric(step)) %>%
+  left_join(grid, by = "patch") %>%
+  pivot_longer(names(fleets), names_to = "fleet", values_to = "catch")
+
+
+
+patch_cpue <- patch_catch %>%
+  left_join(patch_effort %>% select(step, patch, starts_with("fleet"), effort)) %>%
+  mutate(cpue = catch / effort) %>%
+  mutate(mpa = TRUE)
+
+patch_cpue %>%
+  group_by(critter, step, fleet) %>%
+  mutate(scpue = cpue / max(cpue, na.rm = TRUE)) %>%
+  ungroup() %>%
+  filter(step == max(step)) %>%
+  ggplot(aes(x,y, fill = scpue))+
+  geom_tile() +
+  facet_grid(critter~fleet) +
+  scale_fill_viridis_c()
+
[…]

### 2022-11-17 scripts/01_coral_case_study.R

@@ -1,6 +1,7 @@
 source(file.path("scripts", "00_setup.R"))
 
 library(gganimate)
+set.seed(42)
 
 resolution <- 20
 
@@ -8,7 +9,7 @@ seasons <- 4
 
 tune_type <- "explt"
 
-experiment_workers <- 8
+experiment_workers <- 6
 
 years <- 50
 
@@ -473,6 +474,10 @@ if (run_coral_example == TRUE){
       iter = 1
     )
 
+  effort_cap <- map(names(fleets), ~Inf, .id = "")
+
+  names(effort_cap) <- names(fleets)
+
   a <- Sys.time()
   coral_mpa_experiements <- case_study_experiments %>%
     ungroup() %>%
@@ -492,6 +497,7 @@ if (run_coral_example == TRUE){
         fleet_model = NA,
         fauna = fauna,
         fleets = fleets,
+        effort_cap = effort_cap,
         years = 20,
         .options = furrr_options(seed = 42),
         .progress = TRUE
@@ -504,22 +510,23 @@ if (run_coral_example == TRUE){
   future::plan(future::sequential)
 
   write_rds(coral_mpa_experiements, file = file.path(results_path, "coral_mpa_experiements.rds"))
-
-
-  optimized_networks <- tibble(alpha = seq(0, 1, by = .1)) %>%
-    ungroup() %>%
-    mutate(
-      opt_mpa = pmap(
-        list(alpha = alpha),
-        marlin::optimize_mpa,
-        fauna = fauna,
-        fleets = fleets,
-        resolution = resolution,
-        starting_conditions = reef_sim[[length(reef_sim)]],
-        max_prop_mpa = 0.3,
-        prop_sampled = 0.1
-      )
-    )
+# stop()
+# #
+#   optimized_networks <- tibble(alpha = seq(0, 1, by = .1)) %>%
+#     ungroup() %>%
+#     mutate(
+#       opt_mpa = pmap(
+#         list(alpha = alpha),
+#         marlin::optimize_mpa,
+#         fauna = fauna,
+#         fleets = fleets,
+#         resolution = resolution,
+#         starting_conditions = reef_sim[[length(reef_sim)]],
+#         max_prop_mpa = 0.3,
+#         prop_sampled = 0.05,
+#         patches_at_a_time = 10
+#       )
+#     )
 
 
 } else {

### 2022-11-17 scripts/02_blue_water_case_study.R

@@ -1,0 +1,575 @@
+source(file.path("scripts", "00_setup.R"))
+
+library(gganimate)
+
+resolution <- 12
+
+seasons <- 4
+
+tune_type <- "explt"
+
+experiment_workers <- 8
+
+years <- 25
+
+experiment_years <- 20
+
+
+# load bycatch risk layers ------------------------------------------------
+
+
+mats <- list.files(path = here("data", "species-matrices"))
+
+mats <- mats[str_detect(mats, "cpue_mean")]
+
+file_comps <- str_split(mats, "_", simplify = TRUE)
+
+species <- str_replace(file_comps[, 1], "-", " ")
+
+# go through and figure out the best matrix for each species in a really hacky manner
+
+best_mat <- rep(NA, n_distinct(species))
+for (s in seq_along(unique(species))){
+
+  fls <- mats[species == unique(species)[s]]
+
+  if (all(str_detect(fls,"ps_cpue"))){
+    # if the only thing available is purse seine CPUE
+
+    fls <- fls[str_detect(fls,"ps_cpue")]
+
+
+  } else {
+
+    fls <- fls[!str_detect(fls,"ps_cpue")]
+
+  }
+
+  if (any(str_detect(fls,"mtper"))){
+
+    fls <- fls[str_detect(fls,"mtper")]
+
+
+  } else {
+
+    fls <- fls[str_detect(fls,"countper")]
+  }
+
+
+  # preference: longline weight cpue
+
+  # then longline number cpue
+
+  # then purse seine
+
+
+  best_mat[s] <- fls
+
+}
+
+get_layer <- function(file) {
+
+  file_comps <- str_split(file, "_", simplify = TRUE)
+
+  species <- str_replace(file_comps[, 1], "-", " ")
+
+
+  hab <-
+    read_csv(here("data", "species-matrices", file),
+             col_names = FALSE,
+             skip = 1) %>%
+    rename(lat = X1) %>%
+    mutate(lat = 1:nrow(.)) %>%
+    pivot_longer(
+      -lat,
+      names_to = "lon",
+      values_to = "hab",
+      names_prefix = "X"
+    ) %>%
+    mutate(hab = tidyr::replace_na(hab, 0),
+           lon = as.numeric(lon) - 1)
+
+
+
+  x_binsize <- (max(hab$lon) - min(hab$lon) + 1) / resolution
+
+  y_binsize <- (max(hab$lat) - min(hab$lat) + 1) / resolution
+
+  if (x_binsize < 1 | y_binsize < 1) {
[…]

### 2022-11-17 03_marlin_paper.qmd
updating makefile instructions

@@
 population biomass (*b*) killed between natural (*m*) and fishing (*u*)
-sources, the amount of biomass killed through fishing being assigned to yield *Y*.
+sources, the amount of biomass killed through fishing being assigned to
+yield *Y*.
 
@@
 
-Selectivity $\alpha$ is modeled through either a logistic form or a dome-shaped form.
-The logistic form is based on the lengths *l* at which 50% of individuals are selected by the fishing gear, $l^{sel}$, and $\delta$ which is the difference between the length at 50% selectivity and
-the length at 95% selectivity.
+Selectivity $\alpha$ is modeled through either a logistic form or a
+dome-shaped form. The logistic form is based on the lengths *l* at which
+50% of individuals are selected by the fishing gear, $l^{sel}$, and
+$\delta$ which is the difference between the length at 50% selectivity
+and the length at 95% selectivity.
 
@@
 
-We approximate the dome-shaped form as a simple normal distribution with mean $l^{sel}$ and standard deviation $\sigma$.
-The normal density function is re-scaled such that the selectivity is 1 at $l^{sel}$.
+We approximate the dome-shaped form as a simple normal distribution with
+mean $l^{sel}$ and standard deviation $\sigma$. The normal density
+function is re-scaled such that the selectivity is 1 at $l^{sel}$.
 
@@
 \alpha_{a,s,f}= \frac{1}{\sigma_{s,f} \sqrt{2\pi}}e^{-0.5 \left( \frac{l_{a,s} - l^{sel}}{\sigma_{s,f}}  \right)^2}
-$$
-Lastly  *z* is total mortality, the sum of fishing mortality and natural mortality (*m*).
+$$ Lastly *z* is total mortality, the sum of fishing mortality and
+natural mortality (*m*).
 
@@
 
-Costs *C* are calculated as a function of $\gamma_f$, that the scales the cost per unit effort for fleet *f*,$\beta$ allows for the cost of effort to scale non-linearly, and $\eta$
-is an additional cost per unit effort of fishing in each patch *p*,
-allowing for the model to account for travel costs for different
-patches. Travel costs are calculated based on the euclidean distance of
-each patch to the nearest port of a given fleet, users can specify any
-number from zero to the number of patches of port locations. When no
-ports are specified, travel costs are zero.
+Costs *C* are calculated as a function of $\gamma_f$, that the scales
+the cost per unit effort for fleet *f*,$\beta$ allows for the cost of
+effort to scale non-linearly, and $\eta$ is an additional cost per unit
+effort of fishing in each patch *p*, allowing for the model to account
+for travel costs for different patches. Travel costs are calculated
+based on the euclidean distance of each patch to the nearest port of a
+given fleet, users can specify any number from zero to the number of
+patches of port locations. When no ports are specified, travel costs are
+zero.
 
@@
 profitability of that fleet in the proceeding time steps in which
-fishing was open, where $\theta$ controls the responsiveness of fleet *f* to the log of
-the ratio of total revenues *R* to total costs *C*, and approximates the
-percentage amount effort changes in response to a one unit changes in
-the log revenue to cost ratio.
+fishing was open, where $\theta$ controls the responsiveness of fleet
+*f* to the log of the ratio of total revenues *R* to total costs *C*,
+and approximates the percentage amount effort changes in response to a
+one unit changes in the log revenue to cost ratio.
 
@@
 
-When MPAs are implemented, effort that used to operate inside the MPAs can either
-leave the fishery, or be redistributed outside the the MPA (the default
-behavior). The model can also accept total catch quotas for each species
-in each time step. When quotas are activated, if the total catch across
-all fleets for a given species under the model's effort dynamics would
-exceed the allowable quota for that species, the total effort for each
-fleet is decrease proportional to its contribution to the total catch
-until the quota is satisfied.
-As an alternative to catch quotas, users can input an effort cap per fleet, which prevents effort from exceeding a given amount under open-access dynamics, though effort may be reduced below this cap if required by the profit equation.
-`marlin` also allows users to specify
-closed fishing seasons for one or more species in the system. During
-closed seasons, no effort is allocated to patches that would result in
-capture of the covered species, potentially resulting in zero effort if
-the species cannot be avoided. Under these conditions, the amount of
-effort in the next open season is a function of the effort and profits
-in the last open fishing season.
+When MPAs are implemented, effort that used to operate inside the MPAs
+can either leave the fishery, or be redistributed outside the the MPA
+(the default behavior). The model can also accept total catch quotas for
+each species in each time step. When quotas are activated, if the total
+catch across all fleets for a given species under the model's effort
+dynamics would exceed the allowable quota for that species, the total
+effort for each fleet is decrease proportional to its contribution to
+the total catch until the quota is satisfied. As an alternative to catch
+quotas, users can input an effort cap per fleet, which prevents effort
+from exceeding a given amount under open-access dynamics, though effort
+may be reduced below this cap if required by the profit equation.
+`marlin` also allows users to specify closed fishing seasons for one or
+more species in the system. During closed seasons, no effort is
+allocated to patches that would result in capture of the covered
+species, potentially resulting in zero effort if the species cannot be
+avoided. Under these conditions, the amount of effort in the next open
+season is a function of the effort and profits in the last open fishing
+season.
 
@@
 preferred), the model will supply default values for that species based
-on the values reported in FishLife [@thorson2020].
-All of these values can be replaced with user-supplied values as needed though, the defaults simply ensure that reasonable and internally consistent values for the species in question are supplied as a starting point.
-
-For the population model, numbers *n* at time *t* for age *a* are a function of growth, death, and recruitment
+on the values reported in FishLife [@thorson2020]. All of these values
+can be replaced with user-supplied values as needed though, the defaults
+simply ensure that reasonable and internally consistent values for the
+species in question are supplied as a starting point.
+
+For the population model, numbers *n* at time *t* for age *a* are a
+function of growth, death, and recruitment
 
@@
 where *BH* is the Beverton-Holt recruitment function, *ssb* is
-spawning-stock-biomass. $l_a$ is the mean length at at age, $l_sel$ is the length at which
-on average 50% of individuals are selected by the fishery, and
-$\delta_{sel}$ are the additional units of length at which on average
-95% of fish are selected by the fishery.
-
-Spawning stock biomass *ssb* is calculated by converting age to mean length, calculating weight
-at age, maturity at age, and then calculating spawning stock biomass as
-the sum of spawning potential at age in a given time step, taking into account the potential for hyperallometry.
-Age is converted to length through the von Bertalanffy  growth equation, assuming log-normally distributed variation in the length at age.
+spawning-stock-biomass. $l_a$ is the mean length at at age, $l_sel$ is
+the length at which on average 50% of individuals are selected by the
+fishery, and $\delta_{sel}$ are the additional units of length at which
+on average 95% of fish are selected by the fishery.
+
+Spawning stock biomass *ssb* is calculated by converting age to mean
+length, calculating weight at age, maturity at age, and then calculating
+spawning stock biomass as the sum of spawning potential at age in a
+given time step, taking into account the potential for hyperallometry.
+Age is converted to length through the von Bertalanffy growth equation,
+assuming log-normally distributed variation in the length at age.
 
@@
 
-Biomass *b* at age is then given by the weight at length equation governed by a scaling coefficient $wa_s$ and an exponent $wb_s$ that controls the rate at which volume scales with length
+Biomass *b* at age is then given by the weight at length equation
+governed by a scaling coefficient $wa_s$ and an exponent $wb_s$ that
+controls the rate at which volume scales with length
 
@@
 
-Maturity is then calculated as a logistic function where $l_{mat}$ is the length at which on average 50% of individuals are sexual maturity, and $\delta_{mat}$ is the units of length beyond
-$l_{mat}$ at which on average 95% of fish are sexually mature.
+Maturity is then calculated as a logistic function where $l_{mat}$ is
+the length at which on average 50% of individuals are sexual maturity,
+and $\delta_{mat}$ is the units of length beyond $l_{mat}$ at which on
+average 95% of fish are sexually mature.
 
@@
 
-Spawning stock biomass at time *t* is then calculated as a function of the numbers at age, the maturity at age, and the weight at age scaled by a parameter $\gamma$.
-When $\gamma$ is greater than 1, the species is said to experience hyperallometric fecundity, i.e.
-fecundity increases faster than weight.
+Spawning stock biomass at time *t* is then calculated as a function of
+the numbers at age, the maturity at age, and the weight at age scaled by
+a parameter $\gamma$. When $\gamma$ is greater than 1, the species is
+said to experience hyperallometric fecundity, i.e. fecundity increases
+faster than weight.
 
@@
 
-5.  Global density dependence allocated by spawning biomass: Density dependence is a function of the
-    sum of spawning biomass across all patches, and recruits are then
-    distributed according to the distribution of spawning biomass
+5.  Global density dependence allocated by spawning biomass: Density
+    dependence is a function of the sum of spawning biomass across all
+    patches, and recruits are then distributed according to the
+    distribution of spawning biomass
 
@@
 
-Fisheries management is often concerned with measuring stock status relative to maximum sustainable yield (MSY) based reference points, though the exact level of stock status desired by societal objectives may vary widely.
-MS based reference points present a problem for a multi-fleet and spatial-temporal model such as `marlin`.
-MSY and the fishing mortality rate that would produce MSY, $F_{MSY}$ are a function of fishery selectivity.
-Fishery selectivity in this model can vary by fleet, and species can be distributed unevenly in both space and time.
-This means that the net fishing selectivity on a species can vary depending on the dynamics at a given moment, making the definition of an equilibrium concept such as MSY challenging.
-
-As such, we do not report MSY based reference points in the model by default, with the exception of the spawning stock biomass at MSY ($SSB_{MSY}$) which is more conserved across selectivity patterns as it is primarily a function of the mortality and flow of recruits relative to the maturity ogive.
-There are many different strategies for estimating reference points
-conditional on selectivity and regimes, see citations xx. We leave it to
-users to find relevant reference point as required by their methods on
-their own, using the sort of MSE examples available in the package.
+Fisheries management is often concerned with measuring stock status
+relative to maximum sustainable yield (MSY) based reference points,
+though the exact level of stock status desired by societal objectives
+may vary widely. MS based reference points present a problem for a
+multi-fleet and spatial-temporal model such as `marlin`. MSY and the
+fishing mortality rate that would produce MSY, $F_{MSY}$ are a function
+of fishery selectivity. Fishery selectivity in this model can vary by
+fleet, and species can be distributed unevenly in both space and time.
+This means that the net fishing selectivity on a species can vary
+depending on the dynamics at a given moment, making the definition of an
+equilibrium concept such as MSY challenging.
+
+As such, we do not report MSY based reference points in the model by
+default, with the exception of the spawning stock biomass at MSY
+($SSB_{MSY}$) which is more conserved across selectivity patterns as it
+is primarily a function of the mortality and flow of recruits relative
+to the maturity ogive. There are many different strategies for
+estimating reference points conditional on selectivity and regimes, see
+citations xx. We leave it to users to find relevant reference point as
+required by their methods on their own, using the sort of MSE examples
+available in the package.
 
@@
 <!-- Our MPA optimization algorithm seems to create an MPA network that -->
+
 <!-- maximizes an objective function, given by -->
@@
 <!-- $$ -->
+
 <!-- obj_p = \Phi \times scale(\sum_f^F \boldsymbol{\phi_{f,p}}) + (1 - \Phi)  scale(\boldsymbol{\sum_s^S\frac{ssb_s}{ssb0_{s,p}}}) -->
+
 <!-- $$ {#eq-obj} -->
@@
 <!-- Where *scale* rescales the values to be between 0 and 1 such that the -->
+
 <!-- weight parameter is comparable between vectors. The algorithm works by -->
+
 <!-- first assigning every patch in the system to an MPA. For the first -->
+
 <!-- round, the algorithm then calculates the equilibrium objective function -->
+
 <!-- value resulting from sequentially removing each individual patch from -->
+
 <!-- the MPA network. The cell whose removal from the MPA produces the -->
+
 <!-- highest objective function value is then removed from the MPA network, -->
+
 <!-- and the marginal value of the objective function each of the remaining -->
+
 <!-- cells from the last steps are then stored. For the next steps, we -->
+
 <!-- sampled 25% of the remaining MPA cells with probability proportional to -->
+
 <!-- their marginal contribution to the objective function in the prior -->
+
 <!-- iteration, selecting the cell whose removal produces the highest total -->
+
 <!-- objective function as the cell to remove from the network, and updating -->
+
 <!-- the marginal value of the sampled-but-not-selected cells. This in -->
+
 <!-- essence is a sampling-importance-resampling algorithm. The algorithm -->
+
 <!-- continues until no more cells are left in the MPA. This then provides -->
+
 <!-- for a given objective weighting value $\Phi$ the near-optimal MPA -->
+
 <!-- network at every MPA size from 0 to 100%. We say "near" as they is some -->
+
 <!-- path dependency here, in that for example the network that maximizes the -->
+
 <!-- objective function with 40% of the MPAs does not necessarily contain the -->
+
 <!-- optimal network at 30%. However, the reversed nature of the algorithm, -->
+
 <!-- starting at 100% MPA coverage and working better, produces a more stable -->
+
 <!-- network with regards to this problem than starting at 0% and growing to -->
+
 <!-- 100%. -->
@@
 
-Each of the use cases contains too many parameters and options to be succinctly presented in the text here, and we are concerned that attempting to duplicate all of the parameters of the model here would simply introduce error.
-As such, readers should consult the accompanying code (provide clear instructions for which files) to view the precise details of each simulation.
-xx something like this to clarify why we don't have pages of tables with all the parameters and options).
-Targeted tactical applications though should be sure to consider and document all decisions made around model parameters.
+Each of the use cases contains too many parameters and options to be
+succinctly presented in the text here, and we are concerned that
+attempting to duplicate all of the parameters of the model here would
+simply introduce error. As such, readers should consult the accompanying
+code (provide clear instructions for which files) to view the precise
+details of each simulation. xx something like this to clarify why we
+don't have pages of tables with all the parameters and options).
+Targeted tactical applications though should be sure to consider and
+document all decisions made around model parameters.
 
@@
 In our coral reef example, we model the dynamics of four species: a
-grouper (*Epinephelus fuscoguttatus*), a shallow-reef snapper (*Lutjanus malabaricus*), a deep-water snapper (*Pristipomoides filamentosus*), and
+grouper (*Epinephelus fuscoguttatus*), a shallow-reef snapper (*Lutjanus
+malabaricus*), a deep-water snapper (*Pristipomoides filamentosus*), and
 a reef shark (*Carcharhinus amblyrhynchos*). The simulated groupers
@@
 #| fig-cap: "Frontier in conservation and yield outcomes under different MPA design strategies (A) for coral reef case study.
-Change in yield for each fleet (B) and depltion per species (C) as a function of MPA size and MPA design strategy () for coral reef case study."
+Change in yield for each fleet (B) and depltion per species (C) as a function of MPA size and MPA design strategy for the coral reef case study."
 #| fig-height: 4
@@
 
-coral_mpa_experiements$mpas <- map(coral_mpa_experiements$results, "mpa")
-
-coral_mpa_experiements$obj <- map(coral_mpa_experiements$results, "obj")
-
-coral_results <- coral_mpa_experiements %>%
+coral_mpa_experiments$mpas <- map(coral_mpa_experiments$results, "mpa")
+
+coral_mpa_experiments$obj <- map(coral_mpa_experiments$results, "obj")
+
+coral_results <- coral_mpa_experiments %>%
   unnest(cols = obj)
@@
 management strategy evaluation, policy interactions, and spatial stock
-assessment. Future iterations of the model may include life history
-modules for a broader range of species types such as marine mammals, sea
-birds, and sea turtles.
+assessment. In addition, the process-based movement model used in
+`marlin` can directly use empirical estimates of movement dynamics
+derived in the manner of @thorson2021a, providing a link between
+empirical and simulation based approaches to marine resource management.
+Future iterations of the model may include life history modules for a
+broader range of species types such as marine mammals, sea birds, and
+sea turtles.
 

### 2022-11-17 README.Rmd
updating makefile instructions

--- 2021-02-26 README.Rmd
@@
 
-# Something about MPAs and Bycatch
+# Policy Implications of Coupled Economic and Ecological Dynamics in Aquatic Systems
 
-Repo for manuscript exploring the ways in which habitat overlap, fleet spillover, and MPA design strategies can interact to affect conservation and fishery outcomes of MPAs, focused on WCPO-ish system.
+Paper Introducing the `marlin` package.
+
+## Reproducing results on UNIX machines
+
+This repo is equipped with a `make` file. To reproduce results this way
+
+1. Make sure you have make installed
+
+2. Open a terminal and make sure the working directory of the terminal is set to the location of this repo
+
+3. Make sure you have all dependencies installed by opening R and running `renv::restor()`
+
+4. In the terminal, run `RUN_NAME:MyRunName FORMAT:MyFormat make` where MyRunName is the name of the folder you want to create to store results (e.g. v0.1) and FORMAT is one of pdf, html, docx
+
+5. Wait for a bit! This command will run all the required scripts and knit the paper.

### 2022-11-17 README.Rmd
updating renv instructions

@@
 
+## renv
+
+This repository is equipped with [`renv`](https://rstudio.github.io/renv/) for package management.
+
+To install all dependencies for this project
+
+1. Open R
+
+2. Make sure you have `renv` installed (if not, `install.packages("renv")`)
+
+3. Inside R, making sure the working directory is set to the root directory of this repository, run `renv::restore()`
+
+4. Follow all prompts. This will install local versions of all the dependencies needed for this project
+
+5. If you have trouble installing `marlin`, see instructions [here](https://danovando.github.io/marlin/#installation-troubleshooting)
+
 ## Reproducing results on UNIX machines

### 2022-12-10 03_marlin_paper.qmd
circilng towards something

@@
   warning: false
+  cache: false
 bibliography: references.bib
 params:
-  run_name: ["v0.1"]
-abstract: "Very rough cut.
-Communities around the world are increasingly looking to spatial strategies for managing marine ecosystems.
+  run_name: ["v0.3"]
+abstract: "Communities around the world are increasingly looking to spatial strategies for managing marine ecosystems.
 However, the design of spatial management strategies faces numerous challenges.
-Marine ecosystems are highly dynamic and opaque, and extractive entities such as fishing fleets respond to changes in marine ecosystems in highly complex ways.
+Marine ecosystems are highly dynamic and opaque, and extractive entities such as fishing fleets respond endogenously to changes in marine ecosystems in ways that depend upon ecological and policy context.
 We present a framework for efficient simulation of the bio-economic dynamics of marine systems, designed to support both management and research.
 We demonstrate the model's capabilities by focusing on two case studies on the conservation and food security impacts of marine protected areas (MPAs).
-We highlight how coupled bio-economic dynamics can result in complex trade-offs between conservation, fisheries yields, distributional outcomes in simulated coral reef and plegaic ecosystems.
+We highlight how coupled bio-economic dynamics can result in complex trade-offs between conservation, fisheries yields, and distributional outcomes.
 "
@@
 
-XX explain hyperallometry option
-
-xx add in weird option 5 density dependence
-
 ```{r setup}
@@
 
-theme_set(theme_classic(base_size = 12, base_family = "Helvetica") + theme(strip.background = element_blank()))
+theme_set(theme_classic(base_size = 10, base_family = "Helvetica") + theme(strip.background = element_blank()))
 
@@
 movement calls for implementation of marine protected areas, a form of
-spatial management, across 30% of the world's oceans by 2030. While the
-objectives of individual spatial management strategies vary widely,
-spatial management strategies are often supported based on their ability
-to provide protection to broad swaths of marine ecosystems with one
-management action, with potentially fewer costs and complications than
-for example assessing, setting, and enforcing catch quotas for
-individual species.
+spatial management, across 30% of the world's oceans by the year 2030.
+While the objectives of individual spatial management strategies vary
+widely, these types of policies are often supported based on their
+ability to provide protection, along with potential broader benefits, to
+swaths of marine ecosystems with one management action, with potentially
+fewer costs and complications than for example assessing, setting, and
+enforcing catch quotas for individual species.
 
@@
 extractive entities such as fishing fleets respond to changes in marine
-ecosystems in highly complex ways. These dynamics post challenges even
+ecosystems in highly complex ways. These dynamics pose challenges even
 for single-species assessment and management strategies [@field2006a],
@@
 security must consider factors such as the optimal size and placement of
-a protected area network given a wide array of life history,
-exploitation levels, fleet dynamics, and preferred habitats, all of
-which may change over time, particularly given the impacts of climate
-change.
+a protected area network given a wide array of life history, species
+distributions, exploitation levels, fleet dynamics, and policy-dependent
+behavior [@reimer2017], all of which may change over time, particularly
+given the impacts of climate change.
 
 These challenges are amplified by the fact that our knowledge regarding
-the design and implementation of spatial management strategies such as
-MPAs is largely based on theoretical studies; the size of protected
-areas [@ovando2021a] and time required[@nickols2019] to see substantial
+the succesful design and implementation of spatial management strategies
+such as MPAs is largely based on theoretical studies; the size
+[@ovando2021a] and time required [@nickols2019] to see substantial
 effects from many MPAs means that we have relatively little empirical
-evidence as to their large-scale effects under a range of
+evidence as to their large-scale effects across a range of
 social-ecological settings [@mcclanahan2021].
@@
 to be highly stylized representations intended to provide theoretical
-insights with less relevance for highly specific design questions. This
-complex landscape presents a challenge to stakeholders charged with
-marine spatial planning exercises lacking accessible models suited for
-more tactical applications, and a barrier to the scientific community,
+insights with less relevance for specific applications. This lack of
+accessible models suited for more tactical application presents a
+challenge to stakeholders charged with marine spatial planning
+exercises. It also presents a barrier to the scientific community,
 wherein comparisons across spatial management simulation studies are
@@
 
-In part to help address this challenge, we present here a novel
-bio-economic modeling framework called `marlin` that allows for
-efficient simulation of spatio-temporal biological and economic
-dynamics. `marlin` allows users to simulation the impacts spatial
-policies such as MPAs across a range of species targeted by various
-fleets across heterogeneous and dynamic seascapes. Here we present the
-core methods and functionality of this framework, and demonstrate its
-use in three xx case-study applications. In the first, we show how
-`marlin` can be used to assess critical drivers for positive outcomes
-and unintended consequences of MPAs as a bycatch-mitigation strategy. In
-the second, we present a more tactical application, demonstrating how
-`marlin` can be used to compare and design MPA networks to achieve
-societal objectives.
+To help address this challenge, we present a novel bio-economic modeling
+framework called [`marlin`](https://danovando.github.io/marlin/) that
+allows for efficient simulation of spatio-temporal biological and
+economic dynamics. `marlin` allows users to simulate the impacts
+fisheries management policies across a range of species targeted by
+various fleets across heterogeneous and dynamic seascapes. This model
+can help users seek Parteo-optimal solutions that produce win-wins or
+minimize tradeoffs across multiple objectives, such as food security and
+conservation [@lester2013].
+
+Here we present the core methods and functionality of `marlin`, and
+demonstrate its use in two case-study applications; a coral reef system
+and a pelagic system. Our results show how interacting ecological,
+economic, and design attributes can affect the degree of benefits,
+costs, and tradeoffs between conservation and food security outcomes of
+MPAs. More broadly, this paper both demonstrates the functionality of
+our model and the critical need for considering sufficiently realistic
+coupled ecological and economic dynamics in policy evaluations.
 
@@
 [@plagányi2014] targeted specifically towards capturing heterogeneous
-habitats, movement dynamics, and technical interactions across species.
-In its general form, `marlin` simulates the behavior of populations of
-spatially explicit biologically independent age-structured animal
-populations affected by one or more fishing fleets in time steps
-specified by the user (generally annual or seasonal). While `marlin`
-does not simulate species interactions, fishing fleets in `marlin` are
-capable of targeting and affecting multiple species simultaneously. This
-allows the model to simulate processes such as fisheries bycatch or
-effort displacement in a way that accounts for the ways that changes in
+habitats, movement dynamics, and the simultaneous impacts of fishing
+fleets across multiple species (frequently called "technical
+interactions"). In its general form, `marlin` simulates the behavior of
+populations of spatially explicit biologically independent
+age-structured animal populations affected by one or more fishing fleets
+in time steps specified by the user (generally annual or seasonal).
+While `marlin` does not simulate species interactions, fishing fleets in
+`marlin` are capable of targeting and affecting multiple species
+simultaneously. This allows the model to simulate processes such as
+fisheries bycatch or effort displacement in a way that accounts for how
 fleet behavior might affect multiple species in the system. The ability
@@
 
+As a demonstration of the use of this modeling framework, we explore two
+general situations
+
+1.  Trade-offs and distributional outcomes for food security and
+    ecosystem biomass in a coral-reef fishery
+2.  Implications of climate-driven range shifts for MPA design in a
+    pelagic fishery
+
+We chose these two examples to illustrate the use of the model in
+contrasting systems in which MPAs are increasingly considered.
+
 Below we provide a summary of the `marlin` model, as well as details of
 the generalized simulation exercises and case study analyses. We note
-that `marlin` is a complex package with many options to control detailed
-behaviors. Given that these specific features will evolve over time, we
-do not describe in detail every option of the model here. Rather, we
-focus on the equations and methods describing the core foundations of
-our framework for efficiently simulating bio-economic dynamics in
-aquatic systems.
+that `marlin` is a complex software package with many options to control
+detailed behaviors. Given that these specific features will evolve over
+time, we do not describe in detail every option of the model here.
+Rather, we focus on the equations and methods describing the core
+foundations of our framework for efficiently simulating bio-economic
+dynamics in aquatic systems.
 
@@
 
+### Model Summary
+
+`marlin` simulates the dynamics of one or more species, currently best
+representing fish-like species, using age-structured population
+dynamics. Each time step, species move throughout the simulated seascape
+using both diffusion and "taxis" (active movement towards preferred
+habitat), experiences natural and potentially fishing mortality, and if
+in a spawning season spawn using one of the possible forms of density
+dependence implemented in the model, with potentially auto-correlated
+deviations from deterministic reproductive dynamics.
+
+These species are caught by fishing fleets. A fishing fleet is defined
+in `marlin` by a set of fishers that can have different fishing skill,
+prices, and selectivity for individual species (each of which we denote
+as a *metier*). For example, both a longline and purse seine fleet may
+capture bigeye (*Thunnus obesus*) and skipjack (*Katswamus pelamus*)
+tunas, but the longline fleet can be made much more likely to catch
+bigeye than skipjack, and *vice versa*. Each fishing fleet distributes
+its fishing effort in space according to a specified spatial allocation
+rule, for example in proportion to profit-per-unit-effort, conditional
+on the presence of any spatial restrictions such as MPAs.
+
+Unconstrained by management, the total amount of effort exerted by each
+fleet can follow one of two dynamics: open-access or constant-effort.
+Under constant effort, the total amount of effort of each fleet is fixed
+over time, with the possible exception of attrition due to MPA
+placement. Under open-access, the total amount of effort in a given time
+step is a function of the profitability of the fleet in the previous
+time step, evolving until a bionomic equilibrium of zero total profits
+is reached. Profitability is a function of the volume and price of each
+species caught, as well as the cost of the total amount of fishing
+effort per fleet and the distance from port that the fleet operates at.
+
+The dynamics of each fleet can be modified by management in a variety of
+ways. Managers can impose size limits for individual species within each
+fleet. They can also impose total catch quotas for each species in each
+time step. When quotas are activated, if the total catch across all
+fleets for a given species under the model's effort dynamics would
+exceed the allowable quota for that species, the total effort for each
+fleet is decrease proportional to its contribution to the total catch
+until the quota is satisfied. As an alternative to catch quotas,
+managers can set an effort cap per fleet, which prevents effort from
+exceeding a given amount under open-access dynamics, though effort may
+be reduced below this cap if required by the profit equation (i.e. the
+fleet can choose to fish less than the quota or effort cap). `marlin`
+also allows users to specify closed fishing seasons for one or more
+species in the system. During closed seasons, no effort is allocated to
+patches that would result in capture of the covered species, potentially
+resulting in zero effort if the species cannot be avoided. Under these
+conditions, the amount of effort in the next open season is a function
+of the effort and profits in the last open fishing season. Lastly,
+managers can specify locations of no-take MPAs, which could in theory
+change in size and location. When MPAs are implemented, fishing effort
+that used to operate inside the MPAs can either leave the fishery, or be
+redistributed outside the the MPA (the default behavior).
+
+Users can use the model to explore general properties of spatially
+explicit social-ecological systems, or can parameterize it to explore
+key uncertainties of more tactical applications. As an example of the
+later case, we provide two case studies here designed to reflect the
+kinds of questions stakeholders might face if designing an MPA network
+for a coral reef or pelagic system.
+
 ### Fleet Dynamics
 
-A fishing fleet is defined in `marlin` by a set of fishers that can have
-different catchability, prices, and selectivity for individual species
-within a simulation (each of which we denote as a *metier*). For
-example, both a longline and purse seine fleet may capture bigeye
-(*Thunnus obesus*) and skipjack (*Katswamus pelamus*) tunas, but the
-longline fleet is much more likely to catch bigeye than skipjack, and
-*vice versa*. Each fishing fleet (*f*) generates catches or yield (*Y*),
-revenues (*R*), costs (*C*) from individual species (*s*) that it
-targets in a given time step (*t*) and patch (*p*). The fleet then makes
-decisions around how much and where to fish, subject to regulatory
-constraints, based on th total profitability across all species in space
-and time.
+Each fishing fleet (*f*) generates catches or yield (*Y*), revenues
+(*R*), costs (*C*) from individual species (*s*) that it targets in a
+given time step (*t*) and patch (*p*). The fleet then makes decisions
+around how much and where to fish, subject to regulatory constraints,
+based on th total profitability across all species in space and time.
 
 Revenues *R* for fleet *f* in time step *t* are a function of the total
-amount caught of each species and its price ($\Pi$). The amount caught
-is a function of the selectivity at age of each species for each fleet
-($\alpha$), the fishing efficiency of the fleet for that that species
-(*q*), the amount of fishing effort of the fleet in question in that
-patch in a given time step (*E*), and the total fishing mortality at age
-(including all other fleets) for that particular species in that patch
-and time (*u*).
-
-$$
- R_{t,p,f} = \sum_s^S \sum_a^A \Pi_{s,f}\frac{\alpha_{a,s,f}q_{s,f}E_{t,p,f}}{u_{t,p,s,a}} \times Y_{t,p,s,a}
+amount caught of each species (*Y*) and its price ($\Pi$). The amount
+caught is a function of the selectivity at age of each species for each
+fleet ($\alpha$), the fishing efficiency of the fleet for that that
+species (also called "catchability", *q*), the amount of fishing effort
+of the fleet in question in that patch in a given time step (*E*), and
+the total fishing mortality at age (including all other fleets) for that
+particular species in that patch and time (*u*).
+
+$$
+ R_{t,p,f} = \sum_s^{N_s} \sum_a^{N_a} \Pi_{s,f}\frac{\alpha_{a,s,f}q_{s,f}E_{t,p,f}}{u_{t,p,s,a}} \times Y_{t,p,s,a}
 $$ {#eq-revenue}
 
-*Y* is total catches, which is calculated through the Baranov equation
+Total catches or yield *Y* are calculated through the Baranov equation
 [@baranov1918], which appropriately accounts for the total instantaneous
@@
 
-where *b* is biomass and *u* is total fishing mortality, given by
-
-$$
-u_{t,p,s,a} = \sum_{f=1}^F\alpha_{a,s,f}q_{s,f}E_{t,p,f}
+$$
+u_{t,p,s,a} = \sum_{f=1}^{N_f}\alpha_{a,s,f}q_{s,f}E_{t,p,f}
 $$ {#eq-u}
 
-Selectivity $\alpha$ is modeled through either a logistic form or a
-dome-shaped form. The logistic form is based on the lengths *l* at which
-50% of individuals are selected by the fishing gear, $l^{sel}$, and
-$\delta$ which is the difference between the length at 50% selectivity
-and the length at 95% selectivity.
+Selectivity at age $\alpha$ is modeled through either a logistic form or
+a dome-shaped form. The logistic form is based on the lengths *l* at
+which 50% of individuals are selected by the fishing gear, $l^{sel}$,
+and $\delta$ which is the difference between the length at 50%
+selectivity and the length at 95% selectivity.
 
@@
 \alpha_{a,s,f}= \frac{1}{\sigma_{s,f} \sqrt{2\pi}}e^{-0.5 \left( \frac{l_{a,s} - l^{sel}}{\sigma_{s,f}}  \right)^2}
-$$ Lastly *z* is total mortality, the sum of fishing mortality and
-natural mortality (*m*).
+$$ {#eq-dome}
+
+Lastly *z* is total mortality, the sum of fishing mortality and natural
+mortality (*m*).
 
@@
 for travel costs for different patches. Travel costs are calculated
-based on the euclidean distance of each patch to the nearest port of a
+based on the Euclidean distance of each patch to the nearest port of a
 given fleet, users can specify any number from zero to the number of
@@
 
-$$C_{t,p,f} = \gamma_f  \left( E_{t,p,f}^{\beta_f} + \eta_{f,p} E_{t,p,f} \right)$$
+$$
+C_{t,p,f} = \gamma_f  \left( E_{t,p,f}^{\beta_f} + \eta_{f,p} E_{t,p,f} \right)
+$$ {#eq-travel-costs}
 
@@
 
-When MPAs are implemented, effort that used to operate inside the MPAs
-can either leave the fishery, or be redistributed outside the the MPA
-(the default behavior). The model can also accept total catch quotas for
-each species in each time step. When quotas are activated, if the total
-catch across all fleets for a given species under the model's effort
-dynamics would exceed the allowable quota for that species, the total
-effort for each fleet is decrease proportional to its contribution to
-the total catch until the quota is satisfied. As an alternative to catch
-quotas, users can input an effort cap per fleet, which prevents effort
-from exceeding a given amount under open-access dynamics, though effort
-may be reduced below this cap if required by the profit equation.
-`marlin` also allows users to specify closed fishing seasons for one or
-more species in the system. During closed seasons, no effort is
-allocated to patches that would result in capture of the covered
-species, potentially resulting in zero effort if the species cannot be
-avoided. Under these conditions, the amount of effort in the next open
-season is a function of the effort and profits in the last open fishing
-season.
+This is essentially a Gompertz model for fishing effort, that has been
+used for other theoretical studies of fishing effort dynamics
+[@thorson2013].
 
@@
 Taking all the other parameters of the model as given, `marlin` then
-adjusts the catchability coefficients *q\_{f,s}* for each fleet *f* and
+adjusts the catchability coefficients $q_{f,s}$ for each fleet *f* and
 species *s* such that the equilibrium exploitation rate for each species
@@
 unfished spawning stock biomass, and the model will adjust the
-catchability coefficients *q\_{f,s}* for each fleet *f* and species *s*
+catchability coefficients $q_{f,s}$ for each fleet *f* and species *s*
 such that the equilibrium depletion for each species matches the desired
@@
 distribution (IFD) is the standard method for distributing fishing
-fleets in spaces (see @gillis2003 and references therein). However, the
-IFD presents a number of complications for our model. First, under the
-constant effort model, the total amount of effort required to satisfy
-the IFD may not equal the total amount of effort required by the model.
-Second, even if we were to implement the IFD for a single fleet, the
-model is capable of tracking multiple fleets at once. This means that
-the IFD for a single fleet would be a game-theoretic outcome of the
-actions of each of the individual fleet conditional on the actions of
-all other fleets. While possible to implement this would slow down our
-model runs to the point of practically preventing large-scale evaluation
-of the effects of MPAs on bycatch.
+fleets in spaces (see @gillis2003 and references therein). However,
+analytical solutions to the IFD present a number of complications for
+our model. Primarily, our model is capable of tracking multiple fleets
+at once. This means that the IFD for a single fleet would be a
+game-theoretic outcome of the actions of each of the individual fleet
+conditional on the actions of all other fleets. While possible to
+implement this would slow down our model runs to the point of
+practically preventing large-scale evaluation of the effects of MPAs on
+bycatch.
 
@@
 1.  Revenue per unit effort (RPUE): The fleet distributes itself in
-    space based on revenue per unit effort
-2.  Revenue: The fleet distributes itself in space based on revenue
+    space based on the realized revenue per unit effort in each patch in
+    the preceding time step
+2.  Revenue: The fleet distributes itself in space based on the realized
+    total revenue in each patch in the preceding time step
 3.  Profit per unit effort (PPUE): The fleet distributes itself in space
-    based on profit per unit effort
-4.  Profit: The fleet distributes itself in space based on profit
-
-Revenue based spatial distribution is not likely to be very realistic .
-However, due to the complexity of parameterizing cost functions fleet
-dynamics are often evaluated based on yield or revenue alone, and so we
-include those scenarios here to evaluate the potential implications of
-this choice. The decision on whether to allocate the fleet based on
-absolute or relative (per unit effort) is more complex. When effort
-represents the actions of separate and individual fishing actors (e.g.
-independent fishing vessels), a per-unit-effort strategy may be more
-realistic, in which fishers distribute themselves based on the expected
-catch of their individual efforts [@hilborn1987]. Conversely, a system
-defined by a sole owner seeking to maximize total profits might be
-better represented by a fleet model based on total profits
+    based on the realized profit per unit effort in each patch in the
+    preceding time step
+4.  Profit: The fleet distributes itself in space based on profit in
+    each patch in the preceding time step
+
+Revenue based spatial distribution is not likely to be very realistic;
+in general we would expect fishing fleets to respond to profits on some
+core level. However, due to the complexity of parameterizing cost
+functions fleet dynamics are often evaluated based on yield or revenue
+alone, and so we include those scenarios here to allow users to evalaute
+the potential implications of this choice. The decision on whether to
+allocate the fleet based on absolute or relative (per unit effort)
+metrics is more complex. When effort represents the actions of separate
+and individual fishing actors (e.g. independent fishing vessels), a
+per-unit-effort strategy may be more realistic, in which fishers
+distribute themselves based on the expected catch of their individual
+efforts [@hilborn1987]. Conversely, a system defined by a sole owner
+seeking to maximize total profits might be better represented by a fleet
+model based on total profits
 
@@
 The underlying population model used is an age structured single-species
-model in the manner of @ovando2021a. The population model requires a
-large number of parameters. However, if the user supplies either a
-scientific or common name for the species in question (scientific
-preferred), the model will supply default values for that species based
-on the values reported in FishLife [@thorson2020]. All of these values
-can be replaced with user-supplied values as needed though, the defaults
-simply ensure that reasonable and internally consistent values for the
-species in question are supplied as a starting point.
-
-For the population model, numbers *n* at time *t* for age *a* are a
-function of growth, death, and recruitment
-
-$$
-n_{t,p,s,a} = \begin{cases}
-      = BH(ssb_{t-1,p,s,a}) & \text{if $a = 1$}\\
-     = n_{t-1,p,s,a-1}e^{-(z_{t-1,p,s,a-1}}, & \text{if $1< a < max(age)$}\\
-     =  n_{t-1,p,s,a}e^{-(z_{t-1,p,s,a})} + n_{t-1,a-1}e^{-(z_{t-1,p,s,a-1})}, & \text{if $a = max(a)$}
-  \end{cases}
-$$ {#eq-pop}
-
-where *BH* is the Beverton-Holt recruitment function, *ssb* is
-spawning-stock-biomass. $l_a$ is the mean length at at age, $l_sel$ is
-the length at which on average 50% of individuals are selected by the
-fishery, and $\delta_{sel}$ are the additional units of length at which
-on average 95% of fish are selected by the fishery.
-
-Spawning stock biomass *ssb* is calculated by converting age to mean
-length, calculating weight at age, maturity at age, and then calculating
-spawning stock biomass as the sum of spawning potential at age in a
-given time step, taking into account the potential for hyperallometry.
-Age is converted to length through the von Bertalanffy growth equation,
-assuming log-normally distributed variation in the length at age.
-
-$$
-  l_{a,s} = l_{\infty,s}\left(1 - e^{-k_s(a - a0_{s})}\right)e^{u_s}
-$$ {#eq-len}
-
-Where *u* is a log-normally distributed term allowing for variation in
-the length at age relationship
-
-$$
-u_s \sim normal(0,\sigma_s)
-$$ {#eq-sigma}
-
-Biomass *b* at age is then given by the weight at length equation
-governed by a scaling coefficient $wa_s$ and an exponent $wb_s$ that
-controls the rate at which volume scales with length
-
-$$
-  b_{a,s} = wa_s \times l_{a,s}^{wb_s}
-$$ {#eq-weight}
-
-Maturity is then calculated as a logistic function where $l_{mat}$ is
-the length at which on average 50% of individuals are sexual maturity,
-and $\delta_{mat}$ is the units of length beyond $l_{mat}$ at which on
-average 95% of fish are sexually mature.
-
-$$
-  mat_{a,s} = \frac{1}{\left(1 + e^{-log(19)\times\frac{l_{a,s} - lmat_s}{{\delta}mat_s}}\right)}
-$$ {#eq-mat)}
-
-Spawning stock biomass at time *t* is then calculated as a function of
-the numbers at age, the maturity at age, and the weight at age scaled by
-a parameter $\gamma$. When $\gamma$ is greater than 1, the species is
-said to experience hyperallometric fecundity, i.e. fecundity increases
-faster than weight.
-
-$$
-  ssb_{t,p,s} = \sum_{a=1}^Aw_{a,s,t}^{\gamma_s} mat_{t,a} n_{t,p,s,a}
-$$ {#eq-ssb}
-
-#### Recruitment
-
-Recruitment follows Beverton-Holt dynamics. We do however allow for
-three variants in the timing of density dependence:
-
-1.  Global density dependence: Density dependence is a function of the
-    sum of spawning biomass across all patches, and recruits are then
-    distributed according to habitat quality
-
-$$
-n_{t,p,s,a = 1} = \left(\frac{0.8{\times}\sum_{p=1}^P{r0_{p}}\times{h}\times\sum_{p=1}^P{ssb_{t-1,p}}}{0.2\times{\sum_{p=1}^Pssb0_p}\times(1 - h)+(h - 0.2)\times{\sum_{p=1}^Pssb_{t-1,p}}}\right)\times {hab_p}
-$$ {#eq-dd2}
-
-where *hab* is a vector of recruitment habitat quality by patch degined
-by
-
-$$hab_p = r0_p / \sum_p^P{r0_p}$$
-
-2.  Local density dependence: Density dependence occurs independently in
-    each patch and recruits are retained in their home patch.
-
-$$
-n_{t,p,s,a = 1,p} = \left(\frac{0.8{\times}r0_{s,p}\times{h_s}\times{ssb_{t-1,p,s}}}{0.2\times{ssb0_{p,s}}\times(1 - h_s)+(h_s - 0.2)\times{ssb_{t-1,p}}}\right)
-$$ {#eq-dd1}
-
-3.  Local density dependence then disperse: Density dependence occurs
-    independently in each patch and recruits are then dispersed.
-
-$$
-  n_{t,p,s,a = 1,p} = \left(\frac{0.8{\times}r0_{s,p}\times{h_s}\times{ssb_{t-1,p,s}}}{0.2\times{ssb0_{p,s}}\times(1 - h_s)+(h_s - 0.2)\times{ssb_{t-1,p}}}\right)\times \boldsymbol{d^l} \times \epsilon_t
-$$ {#eq-dd3}
-
-where **d^l^** is the recruitment movement matrix (xx make this notation
-consistent with the movement block)
-
-4.  Post-dispersal density dependence: Larvae are distributed throughout
-    the system, and then density dependence occurs based on the density
-    of adult biomass at the destination patch.
-
-$$
-  larv_{t,p} = ssb_{t-1} \times\boldsymbol{d^l}
-$$ {#eq-larvmove}
-
-$$
-n_{t,a = 1,p} = \left(\frac{0.8{\times}r0_{p}\times{h}\times{larv_{t,p}}}{0.2\times{ssb0_p}\times(1 - h)+(h - 0.2)\times{larv_{t,p}}}\right) \times \epsilon_t
-$$ {#eq-dd4}
-
-5.  Global density dependence allocated by spawning biomass: Density
-    dependence is a function of the sum of spawning biomass across all
-    patches, and recruits are then distributed according to the
-    distribution of spawning biomass
-
-$$
-n_{t,p,s,a = 1} = \left(\frac{0.8{\times}\sum_{p=1}^P{r0_{p}}\times{h}\times\sum_{p=1}^P{ssb_{t-1,p}}}{0.2\times{\sum_{p=1}^Pssb0_p}\times(1 - h)+(h - 0.2)\times{\sum_{p=1}^Pssb_{t-1,p}}}\right)\times \frac{ssb_{t-1,p}}{\sum_{p=1}^Pssb_{t-1,p}}
-$$ {#eq-dd5}
+model in the manner of @ovando2021a. The population model requires many
+parameters. However, if the user supplies either a scientific or common
+name for the species in question (scientific preferred), the model will
+supply default values for that species based on the values reported in
+FishLife [@thorson2020]. All of the life history values can be replaced
+with user-supplied values as needed though, the defaults simply ensure
+that reasonable and internally consistent values for the species in
+question are supplied as a starting point.
 
@@
 diffusion and taxis components of this model, though future extensions
-could easily incorporate advection vectors from say ROMs models
-(citations). In this way then, `marlin` allows users to run anything
-from a simple Gaussian dispersal kernel up to a system governed by
-species that passively diffuse around a dynamic thermal range.
-
-@thorson2021a described the exact steps of the CTMC approach in detail.
+could easily incorporate advection vectors from oceanographic models. In
+this way then, `marlin` allows users to run anything from a simple
+Gaussian dispersal kernel up to a system governed by species that
+passively diffuse around a dynamic thermal range.
+
 We briefly the general CTMC method here, but see @thorson2021a for a
@@
 each patch to each other patch in the system in a given timestep *t* for
-life stage *g* of species *s* is defined by a movement matrix
-$\boldsymbol{M}_{t,s,g}$. $\boldsymbol{M}_{t,s,g}$ is calculated as the
+life stage *a* of species *s* is defined by a movement matrix
+$\boldsymbol{M}_{t,s,a}$. $\boldsymbol{M}_{t,s,a}$ is calculated as the
 matrix exponent of the sum of the diffusion and taxis matrices scaled by
@@
 $$
-\boldsymbol{M}_{t,s,g} = e^{\Delta t(\boldsymbol{D}_{t,s,g} + \boldsymbol{X}_{t,s,g})} 
+\boldsymbol{M}_{t,s,a} = e^{\Delta t(\boldsymbol{D}_{t,s,a} + \boldsymbol{X}_{t,s,a})} 
 $$ {#eq-movement}
@@
 Where $\boldsymbol{D}$ is the diffusion matrix and $\boldsymbol{X}$ is
-the taxis matrix.
-
-The diffusion matrix is a function of a user specified diffusion rate,
-see @thorson2021a. We slightly modify the taxis matrix in the following
+the taxis matrix.The diffusion matrix is a function of a user specified
+diffusion rate, see @thorson2021a.
+
+We slightly modify the taxis matrix, denoted as $\tau$ in the following
 manner
 
-xx need to fix notation somewhere here, see this line of code for much
-clearer depiction of what's happening, since sometimes code is cleaner.
-https://github.com/DanOvando/marlin/blob/3563e7319ad0267f89789267a4c6ee58a9ad0d60/R/fish_class.R#L525
-xx
-
-$$
-n_{t,p,s,a} = \begin{cases}
-      = (1 + \lambda d_{s,f})max(0, h(p2_{t,s,g}) - h(p1_{t,s,g})) & \text{if p2 and p1 are adjacent}\\
-     = -\sum_{g' \neq g1}\{ (1 + \lambda d_{s,f})max(0, h(p2_{t,s,g}) - h(p1_{t,s,g}))\} & \text{if p1 = p2}\\
+$$
+\tau_{p1,p2,t,s,a} = \begin{cases}
+      = (1 + \lambda d_{s,f})e^{h(p2,t,s,a) - h(p1,t,s,a)} & \text{if p2 and p1 are adjacent}\\
+     = -\sum_{p' \neq p1} (1 + \lambda d_{s,f})e^{h(p2,t,s,a) - h(p1,t,s,a)} & \text{if p1 = p2}\\
      = 0 & \text{otherwise.}
@@
 
-The $lambda d_{s,f}$ term is a scalar in which the magnitude of the
-habitat gradient between patches can be scaled relative to the diffusion
-parameter $d$.
+The $\lambda$ term is a scalar in which the magnitude of the habitat
+gradient between patches can be scaled relative to the diffusion
+parameter $d$. When the diffusion and taxis matrices are estimated
+jointly from actual data in the manner of @thorson2021a, $\lambda$
+should be set to zero. However, when users are manually creating both
+the diffusion and taxis matrices, the relative strength of diffusion to
+taxis must be defined by the user. Setting higher values of $\lambda$
+allows users to simulate a scenario where taxis is much more significant
+than diffusion.
 
@@
 move around barriers rather than over them. This is different than
-setting habitat in a patch to zero, which implies a patch that a given
-animal will not move towards for its own sake, but can move over towards
-another better patch.
+setting habitat in a patch to zero; an animal will preferentially move
+towards a patch with zero habitat all else being equal, but can move
+over a zero habitat patch towards another better patch.
 
@@
 us to simulate both preferences of species in space and time
-efficiently.Lastly, this form has the advantage that the parameters are
-directly estimable from real data. So, if provided with say CPUE data
-and a tagging study, users can estimates the movement parameters in the
-same manner as @thorson2021a, and then pass their estimated parameters
-to `marlin` for simulation.
+efficiently. Lastly, this form has the advantage that the parameters are
+directly estimable from real data. So, if provided with say spatial
+abundance data and a tagging study, users can estimate the movement
+parameters in the same manner as @thorson2021a, and then pass their
+estimated parameters to `marlin` for simulation.
 
@@
 allows users to specify a maximum number of cells that adults or
-recruits diffuse in an year. The model will then tune the diffusion
+recruits diffuse in a year. The model will then tune the diffusion
 parameter and habitat gradients to achieve this desired level of
 movement.
+
+#### Population Growth
+
+For the population model, numbers *n* at time *t* for age *a* are a
+function of growth, death, and recruitment
+
+$$
+n_{t,p,s,a} = \begin{cases}
+      = BH(ssb_{t-1,p,s,a}) & \text{if $a = 1$}\\
+     = n_{t-1,p,s,a-1}e^{-(z_{t-1,p,s,a-1}}, & \text{if $1< a < max(age)$}\\
+     =  n_{t-1,p,s,a}e^{-(z_{t-1,p,s,a})} + n_{t-1,a-1}e^{-(z_{t-1,p,s,a-1})}, & \text{if $a = max(a)$}
+  \end{cases}
+$$ {#eq-pop}
+
+where *BH* is the Beverton-Holt recruitment function, *ssb* is
+spawning-stock-biomass. $l_a$ is the mean length at at age, $l_{sel}$ is
+the length at which on average 50% of individuals are selected by the
+fishery, and $\delta_{sel}$ are the additional units of length at which
+on average 95% of fish are selected by the fishery.
+
+Spawning stock biomass *ssb* is calculated by converting age to mean
+length, calculating weight at age, maturity at age, and then calculating
+spawning stock biomass as the sum of spawning potential at age in a
+given time step, taking into account the potential for hyperallometry.
+Age is converted to length through the von Bertalanffy growth equation,
+assuming log-normally distributed variation in the length at age.
+
+$$
+  l_{a,s} = l_{\infty,s}\left(1 - e^{-k_s(a - a0_{s})}\right)e^{u_s}
+$$ {#eq-len}
+
+Where *u* is a log-normally distributed term allowing for a constant
+coefficient of variation in the length at age relationship
+
+$$
+u_s \sim N(0,\sigma_s)
+$$ {#eq-sigma}
+
+Biomass *b* at age is then given by the weight at length equation
+governed by a scaling coefficient $wa_s$ and an exponent $wb_s$ that
+controls the rate at which volume scales with length
+
+$$
+  b_{a,s} = wa_s \times l_{a,s}^{wb_s}
+$$ {#eq-weight}
+
+Maturity is then calculated as a logistic function where $l_{mat}$ is
+the length at which on average 50% of individuals are sexual maturity,
+and $\delta_{mat}$ is the units of length beyond $l_{mat}$ at which on
+average 95% of fish are sexually mature.
+
+$$
+  mat_{a,s} = \frac{1}{\left(1 + e^{-log(19)\times\frac{l_{a,s} - lmat_s}{{\delta}mat_s}}\right)}
+$$ {#eq-mat)}
+
+Spawning stock biomass at time *t* is then calculated as a function of
+the numbers at age, the maturity at age, and the weight at age scaled by
+a parameter $\gamma$. When $\gamma$ is greater than 1, the species is
+said to experience hyperallometric fecundity, i.e. fecundity increases
+faster than weight.
+
+$$
+  ssb_{t,p,s} = \sum_{a=1}^{N_a}w_{a,s,t}^{\gamma_s} mat_{t,a} n_{t,p,s,a}
+$$ {#eq-ssb}
+
+#### Recruitment
+
+Recruitment follows Beverton-Holt dynamics. We do however allow for five
+variants in the timing of density dependence:
+
+1.  Global density dependence: Density dependence is a function of the
+    sum of spawning biomass across all patches, and recruits are then
+    distributed according to habitat quality
+
+$$
+n_{t,p,s,a = 1} = \left(\frac{0.8{\times}\sum_{p=1}^P{r0_{p}}\times{h}\times\sum_{p=1}^P{ssb_{t-1,p}}}{0.2\times{\sum_{p=1}^Pssb0_p}\times(1 - h)+(h - 0.2)\times{\sum_{p=1}^Pssb_{t-1,p}}}\right)\times {hab_p}
+$$ {#eq-dd2}
+
+where *hab* is a vector of relative recruitment habitat quality in each
+patch.
+
+$$
+hab_p = r0_p / \sum_p^P{r0_p}
+$$ {#eq-hab}
+
+2.  Local density dependence: Density dependence occurs independently in
+    each patch and recruits are retained in their home patch.
+
+$$
+n_{t,p,s,a = 1,p} = \left(\frac{0.8{\times}r0_{s,p}\times{h_s}\times{ssb_{t-1,p,s}}}{0.2\times{ssb0_{p,s}}\times(1 - h_s)+(h_s - 0.2)\times{ssb_{t-1,p}}}\right)
+$$ {#eq-dd1}
+
+3.  Local density dependence then disperse: Density dependence occurs
+    independently in each patch and recruits are then dispersed.
+
+$$
+  n_{t,p,s,a = 1,p} = \left(\frac{0.8{\times}r0_{s,p}\times{h_s}\times{ssb_{t-1,p,s}}}{0.2\times{ssb0_{p,s}}\times(1 - h_s)+(h_s - 0.2)\times{ssb_{t-1,p}}}\right)\times \boldsymbol{d^l} \times \epsilon_t
+$$ {#eq-dd3}
+
+where **d^l^** is the recruitment movement matrix
+
+4.  Post-dispersal density dependence: Larvae are distributed throughout
+    the system, and then density dependence occurs based on the density
+    of adult biomass at the destination patch.
+
+$$
+  larv_{t,p} = ssb_{t-1} \times\boldsymbol{d^l}
+$$ {#eq-larvmove}
+
+$$
+n_{t,a = 1,p} = \left(\frac{0.8{\times}r0_{p}\times{h}\times{larv_{t,p}}}{0.2\times{ssb0_p}\times(1 - h)+(h - 0.2)\times{larv_{t,p}}}\right) \times \epsilon_t
+$$ {#eq-dd4}
+
+5.  Global density dependence allocated by spawning biomass: Density
+    dependence is a function of the sum of spawning biomass across all
+    patches, and recruits are then distributed according to the
+    distribution of spawning biomass
+
+$$
+n_{t,p,s,a = 1} = \left(\frac{0.8{\times}\sum_{p=1}^P{r0_{p}}\times{h}\times\sum_{p=1}^P{ssb_{t-1,p}}}{0.2\times{\sum_{p=1}^Pssb0_p}\times(1 - h)+(h - 0.2)\times{\sum_{p=1}^Pssb_{t-1,p}}}\right)\times \frac{ssb_{t-1,p}}{\sum_{p=1}^Pssb_{t-1,p}}
+$$ {#eq-dd5}
 
@@
 though the exact level of stock status desired by societal objectives
-may vary widely. MS based reference points present a problem for a
+may vary widely. MSY based reference points present a problem for a
 multi-fleet and spatial-temporal model such as `marlin`. MSY and the
@@
 fleet, and species can be distributed unevenly in both space and time.
-This means that the net fishing selectivity on a species can vary
-depending on the dynamics at a given moment, making the definition of an
-equilibrium concept such as MSY challenging.
+This means that the net effective fishing selectivity on a species can
+vary depending on the dynamics at a given moment, making the definition
+of an equilibrium concept such as MSY challenging.
 
@@
 estimating reference points conditional on selectivity and regimes, see
-citations xx. We leave it to users to find relevant reference point as
-required by their methods on their own, using the sort of MSE examples
-available in the package.
+citations xx. We leave it to users to define and find relevant reference
+point as required by their specific needs.
 
@@
 
-### Use Cases
-
-We include three examples demonstrating how `marlin` can be used to
+### Case Studies
+
+We include two examples demonstrating how `marlin` can be used to
 support marine spatial planning. In the first, we show how `marlin` can
@@
 second, we demonstrate how `marlin` can be used to assess components of
-climate resilience of alternative MPA design strategies. In the third,
-we show how `marlin` can be used to run large-scale simulation testing
-of management strategies.
-
-Each of the use cases contains too many parameters and options to be
+climate resilience of alternative MPA design strategies.
+
+Each of the case studies contains too many parameters and options to be
 succinctly presented in the text here, and we are concerned that
@@
 simply introduce error. As such, readers should consult the accompanying
-code (provide clear instructions for which files) to view the precise
-details of each simulation. xx something like this to clarify why we
-don't have pages of tables with all the parameters and options).
-Targeted tactical applications though should be sure to consider and
-document all decisions made around model parameters.
+code (xx provide clear instructions for which files xx) to view the
+precise details of each simulation. Targeted tactical applications
+though should be sure to consider and document all decisions made around
+model parameters.
 
@@
 be used to examine the performance of potentially sub-optimal but more
-broadly usable design strategies xx.
+broadly usable design strategies. We allow all MPAs to be designed in a
+"mosaic" fashion, but users can easily extend the analysis to for
+example compare outcomes between "contiguous" and "mosaic" MPA designs
+[@pons2022].
 
@@
 model fleet two's selectivity as logistic due to high levels of discard
-mortality for deep-water snapper resulting from barotrauma (xx). This
-fleet catches some grouper, though less than fleet one, and receives no
-price for sharks due to the requirements of a certification program
-through which they sell their deep-water snapper, though they do
-accidentally catch and kill them on occasion. This fleet operates under
-open-access dynamics, as fishing is not the primary means of subsistence
-and so short-term effort expands and contracts in respond to
-profitability of the primarily grouper-driven fishery.
+mortality for deep-water snapper resulting from barotrauma. Fleet two
+catches some grouper, though less than fleet one, and receives no price
+for sharks due to the requirements of a certification program through
+which they sell their deep-water snapper, though they do accidentally
+catch and kill them on occasion. This fleet operates under open-access
+dynamics, as fishing is not the primary means of subsistence and so
+short-term effort expands and contracts in respond to profitability of
+the primarily grouper-driven fishery.
 
@@
 
-We then simulated the outcomes for both food security and conservation
-for each of the species and both of the fleets as a function of both MPA
+We simulated the outcomes for both food security and conservation for
+each of the species and both of the fleets as a function of both MPA
 size and MPA design strategy. For this exercise, MPAs are placed with
@@
 longline fleet, and dome-shaped for the purse-seine fleet, with
-selectivity values pulled from the literature where available (see SI
-xx). Both fleets operate under open-access dynamics with an effort cap.
-The effort cap was set at the level of effort that resulted in the
-desired levels of depletion for each species under open-access dynamics
-(@fig-blue), intended to simulate a scenario where managers step in to
-prevent further expansion of fishing effort in a fully developed
-fishery. For forward-simulation, open-access dynamics can result in
-effort decreasing in response to profitability, but cannot result in
-effort beyond the effort cap set for each fleet.
+selectivity values pulled from the literature where available. Both
+fleets operate under open-access dynamics with an effort cap. The effort
+cap was set at the level of effort that resulted in the desired levels
+of depletion for each species under open-access dynamics (@fig-blue),
+intended to simulate a scenario where managers step in to prevent
+further expansion of fishing effort in a fully developed fishery. For
+forward-simulation, open-access dynamics can result in effort decreasing
+in response to profitability, but cannot result in effort beyond the
+effort cap set for each fleet.
 
@@
 centroid of each population shifts northward at a rate of 0.2 patches
-per year, over a `r experiment_years` time horizon. We designed MPA
-networks given the conditions in the starting year, and held that
-network constant over the years of the experiment, running one
-simulation with and another without the climate-drive range shift. We
-then compared effect of this range shift on food security and
+per year, over a `r experiment_years` year time horizon (@fig-blue-ssb).
+We designed MPA networks given the conditions in the starting year, and
+then held that network constant over the years of the experiment,
+running one simulation with and another without the climate-drive range
+shift. We then compared effect of this range shift on food security and
 conservation outcomes from MPA networks designed based on the pre-range
@@
 #| label: fig-blue-ssb
-#| fig-cap: "Trajectory of depletion (spawning biomass divided by unfished spawning biomass) over time prior to implementation of MPAs."
+#| fig-cap: "Trajectory of depletion (spawning biomass divided by unfished spawning biomass) over time prior to implementation of MPAs for the pelagic case study."
 
@@
 critter_biomass %>%
-  ggplot(aes(step, depletion)) +
-  geom_line() + 
-  facet_wrap(~critter) +
-  scale_y_continuous(limits = c(0, 1))
+  ungroup() %>% 
+  left_join(sci_to_com, by = "critter") %>%
+  select(-critter) %>%
+  rename(critter = common_name) %>%
+  ggplot(aes(step, depletion, color = critter)) +
+  geom_line(linewidth = 1) +
+  scale_y_continuous(limits = c(0, 1), name = "Depletion", expand = expansion(mult = c(0.0,0.05))) +
+  scale_x_continuous(name = "Time Step") +
+  scale_color_discrete(name = "")
 
@@
 #| label: fig-blue
-#| fig-cap: "Distribution of fish spawning biomass under status quo (blue) and range-shifted (red) conditions for blue-water simulation. The x-axis represents longitude, the y-axis latitude. "
+#| fig-cap: "Contour plot of fish spawning biomass in space under status quo (blue) and range-shifted (red) conditions for blue-water simulation.
+Packed together lines indicate higher biomass densities, farther spaced lines lower densities.
+The x-axis represents longitude, the y-axis latitude.
+"
 
@@
 
-## Results
+# Results
 
@@
 For the coral reef example, both the area and target_fishing MPA design
-strategies were capable of producing a Pareto-improving outcome where
-the total spawning biomass divided by unfished spawning biomass across
-all species could be increased by up to 200% while maintaining at least
-the same levels of fisheries yields as would be caught in the absence of
-any MPAs. Conversely, achieving increases in total fish biomass through
-the rate or avoid fishing strategies resulted in immediate losses in
-total yields, and in losses to both yields and biomass in some cases
-(@fig-coral-results).
+strategies were capable of producing a Pareto-improving, i.e.
+potentially win-win, outcome where the total spawning biomass divided by
+unfished spawning biomass across all species could be increased by up to
+200% while maintaining at least the same levels of fisheries yields as
+would be caught in the absence of any MPAs. Conversely, achieving
+increases in total fish biomass through the rate or avoid_fishing
+strategies resulted in immediate losses in total yields, and in losses
+to both yields and biomass in some cases (@fig-coral-results).
 
@@
 
-Both the *rate* and *avoid fishing* MPA placement strategies produced
-lower total fishery gains but minimized the negatiev impacts on fleet
-two. However, these strategies produced much lower conservation gains
-than the *area* and *target_fishing* strategies, which produced minimal
-gains for any species besides deep snapper, and in fact resulting in
-slight losses to the nearshore snapper population for MPA sizes less
-than roughly 75%. Both the area and target_fishing strategies resulted
-in rapid conservation gains across all of the simulated species.
+Both the rate and avoid_fishing MPA placement strategies produced lower
+total fishery gains but minimized the negative impacts on fleet two.
+However, these strategies produced much lower conservation gains than
+the area and target_fishing strategies, only benefiting deep_snapper
+meaningfully, and in fact resulting in slight losses to the nearshore
+snapper population for MPA sizes less than roughly 75%. Both the area
+and target_fishing strategies resulted in rapid conservation gains
+across all of the simulated species.
 
@@
 #| fig-cap: "Frontier in conservation and yield outcomes under different MPA design strategies (A) for coral reef case study.
-Change in yield for each fleet (B) and depltion per species (C) as a function of MPA size and MPA design strategy for the coral reef case study."
+Change in yield for each fleet (B) and depletion per species (C) as a function of MPA size and MPA design strategy for the coral reef case study."
 #| fig-height: 4
@@
 The yield impacts of the proposed MPA strategies were much more
-sensitive to climate impacts for the purse seine fishery; under the
-*status quo* species distribution, the target fishing MPA design
+sensitive to climate impacts for the purse seine fishery. Under the
+*status quo* species distribution, the target_fishing MPA design
 strategy produced increases in fisheries yields at sizes between 25-50%
@@
 
-## Discussion
+# Discussion
 
@@
 Communities must often make decisions on how to manage these systems
-based on ... Modeling tools such as `marlin` can help users answer
-important scientific questions and design marine management policies
-informed by better understanding of sensitivities to key uncertainties.
-blah blah blah. `marlin` is to our knowledge the first bio-economic
-modeling tool that allows scientists and decision makers to efficiently
-simulate the outcomes of management strategies such as MPAs across
-multiple fleets and species while accounting for layered regulations,
-economic incentives, age structure, and realistic movement dynamics
-across a range of potential social-ecological conditions. xx this
-sentence annoys me but feels like we need to make something like it
-explicit to convince reviewers maybe less familiar with the modeling
-literature xx.
+based on limited empirical evidence. Modeling tools such as `marlin` can
+help users answer important scientific questions and design marine
+management policies informed by better understanding of sensitivities to
+key uncertainties.
+
+<!-- `marlin` is to our knowledge the first bio-economic modeling tool that -->
+
+<!-- allows scientists and decision makers to efficiently simulate the -->
+
+<!-- outcomes of management strategies such as MPAs across multiple fleets -->
+
+<!-- and species while accounting for layered regulations, economic -->
+
+<!-- incentives, age structure, and realistic movement dynamics across a -->
+
+<!-- range of potential social-ecological conditions. xx this sentence annoys -->
+
+<!-- me but feels like we need to make something like it explicit to convince -->
+
+<!-- reviewers maybe less familiar with the modeling literature xx. -->
 
@@
 complexity, but can be extremely time consuming to design and run. Tools
-such as virtualspecies [@leroy2016] or STEPS [@visintin2020] are help
+such as virtualspecies [@leroy2016] or STEPS [@visintin2020] can help
 model the distribution of species in space and time as a function of
@@
 distribution of costs and benefits stemming from alternative design
-choices and states of nature shown in our results demonstrates the value
-in our method. Beyond the current examples, we envision `marlin` being
-applicable to research on dynamic ocean management, range shifts,
-management strategy evaluation, policy interactions, and spatial stock
-assessment. In addition, the process-based movement model used in
-`marlin` can directly use empirical estimates of movement dynamics
-derived in the manner of @thorson2021a, providing a link between
-empirical and simulation based approaches to marine resource management.
-Future iterations of the model may include life history modules for a
-broader range of species types such as marine mammals, sea birds, and
-sea turtles.
+choices and states of nature shown in our case study results
+demonstrates the value in our method. Beyond the current examples, we
+envision `marlin` being applicable to research on dynamic ocean
+management, range shifts, management strategy evaluation, policy
+interactions, and spatial stock assessment. In addition, the
+process-based movement model used in `marlin` can directly use empirical
+estimates of movement dynamics derived in the manner of @thorson2021a,
+providing a link between empirical and simulation based approaches to
+marine resource management. Future iterations of the model may include
+life history modules for a broader range of species types such as marine
+mammals, sea birds, and sea turtles.
 
@@
 models such as `marlin` can help users understand factors that drive the
-performance of potential marine management strategies. The goal of this
-tool is to empower people to design policies based on evaluation of key
+performance of marine management strategies. The goal of this tool is to
+empower people to design policies based on evaluation of key
 uncertainties and trade-offs, and in doing so support more effective and
@@
 
-## References
+# References
 
@@
 
-## scraps
+# Supporting Information
+
+# Scraps

### 2022-12-21 03_marlin_paper.qmd
out for internal review

@@
 ---
-title: "Policy Implications of Coupled Economic and Ecological Dynamics in Aquatic Systems"
+title: "Revealing Benefits, Costs, and Tradeoffs of Spatial Management in Marine Systems Arrising from Coupled Economic and Ecological Dynamics"
 author:
@@
     - James Thorson
-date: "`r Sys.Date()`"
+date: "Updated `r Sys.Date()`"
 format:
@@
   run_name: ["v0.3"]
-abstract: "Communities around the world are increasingly looking to spatial strategies for managing marine ecosystems.
-However, the design of spatial management strategies faces numerous challenges.
-Marine ecosystems are highly dynamic and opaque, and extractive entities such as fishing fleets respond endogenously to changes in marine ecosystems in ways that depend upon ecological and policy context.
-We present a framework for efficient simulation of the bio-economic dynamics of marine systems, designed to support both management and research.
-We demonstrate the model's capabilities by focusing on two case studies on the conservation and food security impacts of marine protected areas (MPAs).
-We highlight how coupled bio-economic dynamics can result in complex trade-offs between conservation, fisheries yields, and distributional outcomes.
-"
+abstract: "Abstract: Designing effective spatial management strategies is challenging because marine ecosystems are highly dynamic and opaque, and extractive entities such as fishing fleets respond endogenously to ecosystem changes in ways that depend upon ecological and policy context.
+We present a tool, `marlin`, that can be used to efficiently simulate the bio-economic dynamics of marine systems in support of both management and research.
+We demonstrate `marlin’s` capabilities by focusing on two case studies on the conservation and food production impacts of marine protected areas (MPAs): a coastal coral reef and a pelagic tuna fishery.
+In the coastal coral reef example, we explore how heterogeneity in species distributions and fleet preferences can affect distributional outcomes of MPAs.
+In the pelagic case study, we show how our model can be used to assess the climate resilience of different MPA design strategies, as well as the climate sensitivity of different fishing fleets.
+This paper demonstrates how intermediate complexity simulation of coupled bio-economic dynamics can help communities predict and potentially manage trade-offs between conservation, fisheries yields, and distributional outcomes of management policies affected by spatial bio-economic dynamics."
 header-includes:
@@
 
-sci_to_com <- taxize::sci2comm(unique(marlin_inputs$scientific_name)) %>% 
+sci_to_com <- taxize::sci2comm(unique(marlin_inputs$scientific_name),simplify= TRUE,db='ncbi') %>% 
   bind_rows(.id = "critter") %>% 
@@
 
+Abstract: Designing effective spatial management strategies is challenging because marine ecosystems are highly dynamic and opaque, and extractive entities such as fishing fleets respond endogenously to ecosystem changes in ways that depend upon ecological and policy context.
+We present a tool – marlin -- to efficiently simulate the bio-economic dynamics of marine systems, designed to support both management and research.
+We demonstrate `marlin’s` capabilities by focusing on two case studies on the conservation and food production impacts of marine protected areas (MPAs): a coastal coral reef and a pelagic tuna fishery.
+In the coastal coral reef example, we explore how heterogeneity in species distributions and fleet preferences can affect distributional outcomes of MPAs.
+In the pelagic case study, we demonstrate how our model can be used to assess the climate resilience of different MPA design strategies, as well as the climate sensitivity of different fishing fleets.
+This paper demonstrates how intermediate complexity simulation of coupled bio-economic dynamics can help communities predict and potentially manage trade-offs between conservation, fisheries yields, and distributional outcomes resulting from spatial management.
+
 ## Introduction
@@
 Communities around the world are increasingly looking to spatial
-strategies for managing marine ecosystems. For example, the "30x30"
-movement calls for implementation of marine protected areas, a form of
-spatial management, across 30% of the world's oceans by the year 2030.
-While the objectives of individual spatial management strategies vary
-widely, these types of policies are often supported based on their
-ability to provide protection, along with potential broader benefits, to
-swaths of marine ecosystems with one management action, with potentially
-fewer costs and complications than for example assessing, setting, and
-enforcing catch quotas for individual species.
-
-However, the design of spatial management strategies faces numerous
-challenges. Marine ecosystems are highly dynamic and opaque, and
-extractive entities such as fishing fleets respond to changes in marine
-ecosystems in highly complex ways. These dynamics pose challenges even
-for single-species assessment and management strategies [@field2006a],
-which are only amplified when we consider management policies
-specifically designed around spatial features in the oceans. For
-example, policies such as MPAs designed for both conservation and food
-security must consider factors such as the optimal size and placement of
-a protected area network given a wide array of life history, species
-distributions, exploitation levels, fleet dynamics, and policy-dependent
-behavior [@reimer2017], all of which may change over time, particularly
-given the impacts of climate change.
-
-These challenges are amplified by the fact that our knowledge regarding
-the succesful design and implementation of spatial management strategies
-such as MPAs is largely based on theoretical studies; the size
-[@ovando2021a] and time required [@nickols2019] to see substantial
-effects from many MPAs means that we have relatively little empirical
-evidence as to their large-scale effects across a range of
-social-ecological settings [@mcclanahan2021].
+strategies for managing marine ecosystems . For example, the "30x30"
+movement calls for implementation of marine protected areas (MPAs), a
+form of spatial management, across 30% of the world's oceans by the year
+2030 [@grorud-colvert2021 and references therein]. However , designing
+spatial management strategies to achieve desired objectives -- which may
+include recovery of overfished species, increased food production and
+economic well being, and equitable distribution of benefits - is not
+straightforward. Different species have different resilience to fishing,
+distributions in space and time, and value to fishing fleets; fishing
+fleets themselves can have varying ranges of incentives and reactions to
+policy structures. These dynamics pose challenges even for
+single-species assessment and management strategies [@field2006a], which
+are only amplified when we consider management policies specifically
+designed around spatial features in the oceans.
+
+For example, policies such as MPAs designed for both conservation and
+food production must consider factors such as the optimal size and
+placement of a protected area network given a wide array of life
+history, species distributions, exploitation levels, fleet dynamics, and
+policy-dependent behavior [@reimer2017], all of which may change over
+time, particularly given the impacts of climate change. Efforts to
+effectively design spatial management strategies such as MPAs are
+further constrained by a lack of empirical evidence describing the size
+[@ovando2021a] and time required [@nickols2019] for MPAs to produce
+substantial effects across a range of social-ecological settings
+[@mcclanahan2021].
 
@@
 To help address this challenge, we present a novel bio-economic modeling
-framework called [`marlin`](https://danovando.github.io/marlin/) that
-allows for efficient simulation of spatio-temporal biological and
-economic dynamics. `marlin` allows users to simulate the impacts
-fisheries management policies across a range of species targeted by
-various fleets across heterogeneous and dynamic seascapes. This model
-can help users seek Parteo-optimal solutions that produce win-wins or
-minimize tradeoffs across multiple objectives, such as food security and
-conservation [@lester2013].
+tool called [`marlin`](https://danovando.github.io/marlin/) that allows
+for efficient simulation of spatio-temporal biological and economic
+dynamics. `marlin` allows users to simulate the impacts of marine
+management policies across a range of species targeted by various fleets
+across heterogeneous and dynamic seascapes. This model can help users
+seek Parteo-optimal solutions that produce win-wins or minimize
+tradeoffs across multiple management objectives, such as food production
+and conservation [@lester2013].
 
@@
 economic, and design attributes can affect the degree of benefits,
-costs, and tradeoffs between conservation and food security outcomes of
-MPAs. More broadly, this paper both demonstrates the functionality of
+costs, and tradeoffs between conservation and food production outcomes
+of MPAs. More broadly, this paper both demonstrates the functionality of
 our model and the critical need for considering sufficiently realistic
@@
 
-As a demonstration of the use of this modeling framework, we explore two
+As a demonstration of the use of this modeling two, we explore two
 general situations
 
-1.  Trade-offs and distributional outcomes for food security and
-    ecosystem biomass in a coral-reef fishery
+1.  Trade-offs and distributional outcomes for food production and
+    multi-species biomass in a coral-reef fishery
 2.  Implications of climate-driven range shifts for MPA design in a
@@
 Rather, we focus on the equations and methods describing the core
-foundations of our framework for efficiently simulating bio-economic
-dynamics in aquatic systems.
+foundations of our tool for efficiently simulating bio-economic dynamics
+in aquatic systems.
 
@@
 in a spawning season spawn using one of the possible forms of density
-dependence implemented in the model, with potentially auto-correlated
-deviations from deterministic reproductive dynamics.
+dependence implemented in the model, with the potential to include
+auto-correlated stochastic deviations in the amount of offspring
+produced, generally called "recruits".
 
@@
 as a *metier*). For example, both a longline and purse seine fleet may
-capture bigeye (*Thunnus obesus*) and skipjack (*Katswamus pelamus*)
-tunas, but the longline fleet can be made much more likely to catch
-bigeye than skipjack, and *vice versa*. Each fishing fleet distributes
-its fishing effort in space according to a specified spatial allocation
-rule, for example in proportion to profit-per-unit-effort, conditional
-on the presence of any spatial restrictions such as MPAs.
+capture bigeye (*Thunnus obesus*) and skipjack tunas, but the longline
+fleet can be made much more likely to catch bigeye than skipjack, and
+*vice versa*. Each fishing fleet distributes its fishing effort in space
+according to a specified spatial allocation rule, for example in
+proportion to profit-per-unit-effort, conditional on the presence of any
+spatial restrictions such as MPAs.
 
@@
 exceed the allowable quota for that species, the total effort for each
-fleet is decrease proportional to its contribution to the total catch
+fleet is decreased proportional to its contribution to the total catch
 until the quota is satisfied. As an alternative to catch quotas,
@@
 that used to operate inside the MPAs can either leave the fishery, or be
-redistributed outside the the MPA (the default behavior).
+redistributed outside the MPA (the default behavior).
 
@@
 kinds of questions stakeholders might face if designing an MPA network
-for a coral reef or pelagic system.
+for a nearshore coral reef or a pelagic system.
 
@@
 around how much and where to fish, subject to regulatory constraints,
-based on th total profitability across all species in space and time.
+based on the total profitability across all species in space and time.
 
@@
 population biomass (*b*) killed between natural (*m*) and fishing (*u*)
-sources, the amount of biomass killed through fishing being assigned to
+sources, with the amount of biomass killed through fishing assigned to
 yield *Y*.
@@
 catches and subsequently revenues for each fleet. Given revenues *R*, we
-then calculate profits $\phi$ based on revenues and costs.
+then calculate profits $\phi$ based on revenues and costs. Costs *C* are
+calculated as a function of $\gamma_f$ that the scales the total cost
+per unit effort for fleet *f*. $\beta$ allows for the cost per unit
+effort effort to scale non-linearly, and $\eta$ is an additional cost
+per unit effort of fishing in each patch *p*, allowing for the model to
+account for travel costs for different patches. Travel costs are
+calculated based on the Euclidean distance of each patch to the nearest
+port of a given fleet, and users can specify any number from zero to the
+number of patches of port locations. When no ports are specified, travel
+costs are zero.
 
@@
 $$ {#eq-profits}
-
-Costs *C* are calculated as a function of $\gamma_f$, that the scales
-the cost per unit effort for fleet *f*,$\beta$ allows for the cost of
-effort to scale non-linearly, and $\eta$ is an additional cost per unit
-effort of fishing in each patch *p*, allowing for the model to account
-for travel costs for different patches. Travel costs are calculated
-based on the Euclidean distance of each patch to the nearest port of a
-given fleet, users can specify any number from zero to the number of
-patches of port locations. When no ports are specified, travel costs are
-zero.
 
@@
 *f* to the log of the ratio of total revenues *R* to total costs *C*,
-and approximates the percentage amount effort changes in response to a
-one unit changes in the log revenue to cost ratio.
+and approximates the proportional change in effort in response to a one
+unit changes in the log revenue to cost ratio.
 
@@
 functions fleet dynamics are often evaluated based on yield or revenue
-alone, and so we include those scenarios here to allow users to evalaute
+alone, and so we include those scenarios here to allow users to evaluate
 the potential implications of this choice. The decision on whether to
@@
 and individual fishing actors (e.g. independent fishing vessels), a
-per-unit-effort strategy may be more realistic, in which fishers
-distribute themselves based on the expected catch of their individual
-efforts [@hilborn1987]. Conversely, a system defined by a sole owner
-seeking to maximize total profits might be better represented by a fleet
-model based on total profits
+per-unit-effort strategy, in which fishers distribute themselves based
+on the expected catch of their individual efforts, may be more realistic
+[@hilborn1987]. Conversely, a system defined by a sole owner seeking to
+maximize total profits might be better represented by a fleet model
+based on total profits
 
@@
 supply default values for that species based on the values reported in
-FishLife [@thorson2020]. All of the life history values can be replaced
-with user-supplied values as needed though, the defaults simply ensure
-that reasonable and internally consistent values for the species in
-question are supplied as a starting point.
+FishLife [@thorson2020]. All of the life history values can
+alternatively be replaced with user-supplied values as needed; the
+default values simply ensure that reasonable and internally consistent
+values for the species in question are supplied as a starting point.
 
@@
 could easily incorporate advection vectors from oceanographic models. In
-this way then, `marlin` allows users to run anything from a simple
-Gaussian dispersal kernel up to a system governed by species that
-passively diffuse around a dynamic thermal range.
-
-We briefly the general CTMC method here, but see @thorson2021a for a
-detailed description of the method. Under this framework, movement from
-each patch to each other patch in the system in a given timestep *t* for
-life stage *a* of species *s* is defined by a movement matrix
+this way, `marlin` allows users to run anything from a simple Gaussian
+dispersal kernel up to a system governed by species that passively
+diffuse around a dynamic thermal range.
+
+We provide a brief overview of the the general CTMC method here, see
+@thorson2021a for a detailed description. Under this framework, movement
+from each patch to each other patch in the system in a given timestep
+*t* for life stage *a* of species *s* is defined by a movement matrix
 $\boldsymbol{M}_{t,s,a}$. $\boldsymbol{M}_{t,s,a}$ is calculated as the
@@
 Where $\boldsymbol{D}$ is the diffusion matrix and $\boldsymbol{X}$ is
-the taxis matrix.The diffusion matrix is a function of a user specified
-diffusion rate, see @thorson2021a.
+the taxis matrix. The diffusion matrix is a function of a user specified
+diffusion rate [@thorson2021a].
 
@@
 the movement model to be run on arbitrarily fine time steps. Second, the
-*taxis* model allow for more active habitat choices by species, allowing
+taxis model allow for more active habitat choices by species, allowing
 us to simulate both preferences of species in space and time
@@
 Maturity is then calculated as a logistic function where $l_{mat}$ is
-the length at which on average 50% of individuals are sexual maturity,
-and $\delta_{mat}$ is the units of length beyond $l_{mat}$ at which on
+the length at which on average 50% of individuals are sexually mature,
+and $\delta_{mat}$ is the unit of length beyond $l_{mat}$ at which on
 average 95% of fish are sexually mature.
@@
 relative to maximum sustainable yield (MSY) based reference points,
-though the exact level of stock status desired by societal objectives
-may vary widely. MSY based reference points present a problem for a
-multi-fleet and spatial-temporal model such as `marlin`. MSY and the
-fishing mortality rate that would produce MSY, $F_{MSY}$ are a function
-of fishery selectivity. Fishery selectivity in this model can vary by
-fleet, and species can be distributed unevenly in both space and time.
-This means that the net effective fishing selectivity on a species can
-vary depending on the dynamics at a given moment, making the definition
-of an equilibrium concept such as MSY challenging.
+though the exact level of stock status desired by management or societal
+objectives may vary widely. MSY based reference points present a problem
+for a multi-fleet and spatial-temporal model such as `marlin`. MSY and
+the fishing mortality rate that would produce MSY, $F_{MSY}$ are a
+function of fishery selectivity. Fishery selectivity in this model can
+vary by fleet, and species can be distributed unevenly in both space and
+time. This means that the net effective fishing selectivity on a species
+can vary depending on the dynamics at a given moment, making the
+definition of an equilibrium concept such as MSY challenging.
 
@@
 default, with the exception of the spawning stock biomass at MSY
-($SSB_{MSY}$) which is more conserved across selectivity patterns as it
-is primarily a function of the mortality and flow of recruits relative
-to the maturity ogive. There are many different strategies for
-estimating reference points conditional on selectivity and regimes, see
-citations xx. We leave it to users to define and find relevant reference
-point as required by their specific needs.
+($SSB_{MSY}$) which tends to be conserved across selectivity patterns as
+it is primarily a function of the mortality and flow of recruits
+relative to the maturity ogive. There are many different strategies for
+estimating reference points in spatially explicit systems [@kapur2021a].
+We leave it to users to define and find relevant reference point as
+required by their specific needs.
 
@@
 support marine spatial planning. In the first, we show how `marlin` can
-be used to compare the total and distributional impacts of marine
-protected areas design in a heavily fished coral reef ecosystem. In the
-second, we demonstrate how `marlin` can be used to assess components of
-climate resilience of alternative MPA design strategies.
+be used to compare the total and distributional impacts of MPAs designed
+in a heavily fished coral reef ecosystem. In the second, we demonstrate
+how `marlin` can be used to assess components of climate resilience of
+alternative MPA design strategies in a pelagic system.
 
@@
 code (xx provide clear instructions for which files xx) to view the
-precise details of each simulation. Targeted tactical applications
-though should be sure to consider and document all decisions made around
-model parameters.
+precise details of each simulation. Targeted tactical applications must
+carefully consider and document all decisions made around model
+parameters.
 
@@
 Across all of our case studies we make use of four potential "rule of
-thumb" MPA design strategies:
-
-1.  Rate: MPAs are placed in based on the pre-MPA depletion-weighted
-    catch relative to the total catch in a patch. So, patches with high
-    rates of catch of depleted species relative to total catch are
-    prioritized.
-
-2.  Avoid_fishing: MPAs locations are prioritized inversely to fishery
-    catches. So, patches with low catches are prioritized over patches
-    with high catches.
-
-3.  Target_fishing: MPAs locations are prioritized proportionally to
-    fishery catches. So, patches with high total catches are prioritized
-    over patches with low catches.
-
-4.  Area: For the "area" strategy, we simply assign patches in
-    sequential order until the desired total MPA size is reached.
-
-MPAs never to our knowledge randomly placed; their locations are decided
-based on stakeholder negotiation based on available information and
-objectives. In theory the design MPA networks can be optimized through
-the use of a modeling framework, and indeed conditional on the validity
-of the model itself this process is likely to produce better outcomes
-than manually-designed strategies [@rassweiler2012; @rassweiler2014].
-However, in practice the ability of communities to conduct numerical
-optimization of MPA networks can be constrained by the number of
-objectives and dimensionality of the operating model, or by simple
-capacity of local stakeholders. As such, we show here how `marlin` can
-be used to examine the performance of potentially sub-optimal but more
-broadly usable design strategies. We allow all MPAs to be designed in a
-"mosaic" fashion, but users can easily extend the analysis to for
-example compare outcomes between "contiguous" and "mosaic" MPA designs
-[@pons2022].
+thumb" MPA design strategies (xx add citation here) :
+
+1.  Depletion Rate (*rate*): MPAs are placed based on the pre-MPA
+    depletion-weighted catch relative to the total catch in a patch. So,
+    patches with high rates of catch of depleted species relative to
+    total catch are prioritized.
+
+2.  Avoid_fishing (*avoid_fishing*): MPAs locations are prioritized
+    inversely to fishery catches. So, patches with low catches are
+    prioritized over patches with high catches.
+
+3.  Target_fishing (*target_fishing*): MPAs locations are prioritized
+    proportionally to fishery catches. So, patches with high total
+    catches are prioritized over patches with low catches.
+
+4.  Area target (*area*): For the "area" strategy, we simply assign
+    patches in sequential order until the desired total MPA size is
+    reached.
+
+To our knowledge, MPAs are never randomly placed; their locations are
+decided based on stakeholder negotiation given available information and
+objectives. In theory, the design of MPA networks can be optimized
+through the use of a modeling framework, and indeed conditional on the
+validity of the model itself this process is likely to produce better
+outcomes than manually-designed strategies [@rassweiler2012;
+@rassweiler2014]. However, designing optimal MPA networks becomes
+increasingly difficult as the range of objectives and the complexity of
+the model increase. Therefore, while `marlin` provides some support for
+[optimization](https://danovando.github.io/marlin/articles/optimize-mpas.html),
+we focus here on the design and performance of these more "rule of
+thumb" design strategies that may be more accessible to a wider range of
+users. We allow all MPAs to be designed in a "mosaic" fashion, but users
+can easily extend the analysis to for example compare outcomes between
+"contiguous" and "mosaic" MPA designs [@pons2022].
 
@@
 These species are targeted by two different fleets. Fleet one primarily
-targets the grouper and nearshore snapper populations, while obtaining
-some funds from sharks. Fleet one has a dome shaped selectivity pattern
-for both grouper and shallow-reef snapper due to demand for "plate
-sized" fish (citation), and logistic selectivity for sharks, as any
-shark caught is retained. This fleet is a subsistence fleet, where the
-local community takes every possible opportunity to fish that they can,
-and as such we model it as a "constant effort" fishery.
+targets the grouper and nearshore snapper populations, but will land any
+incidentally captured sharks. Fleet one has a dome shaped selectivity
+pattern for both grouper and shallow-reef snapper due to demand for
+"plate sized" fish [@kindsvater2017], and logistic selectivity for
+sharks, as any shark caught is retained. This fleet is a subsistence
+fleet, where the local community takes every possible opportunity to
+fish that they can, and as such we model it as a "constant effort"
+fishery.
 
@@
 for sharks due to the requirements of a certification program through
-which they sell their deep-water snapper, though they do accidentally
-catch and kill them on occasion. This fleet operates under open-access
+which they sell their deep-water snapper, though accidental captures do
+occur, which results in mortality. This fleet operates under open-access
 dynamics, as fishing is not the primary means of subsistence and so
-short-term effort expands and contracts in respond to profitability of
+short-term effort expands and contracts in response to profitability of
 the primarily grouper-driven fishery.
@@
 Each fleet has a different home-port location, and incurs an additional
-cost for fishing each patch in the system in proportion to the euclidean
+cost for fishing each patch in the system in proportion to the Euclidean
 distance from each patch to each fleet's home port.
 
-We simulated the outcomes for both food security and conservation for
-each of the species and both of the fleets as a function of both MPA
-size and MPA design strategy. For this exercise, MPAs are placed with
-perfect information and have no design constraints for continuity (i.e.
-the selected MPA design can be as patchwork as dictated by the data).
+We use `marlin` to simulate the outcomes for both food production and
+conservation for each of the species and both of the fleets as a
+function of both MPA size and MPA design strategy. For this exercise,
+MPAs are placed with perfect information and have no design constraints
+for continuity (i.e. the selected MPA design can be as patchwork as
+dictated by the data).
 
@@
 
-We modeled our pelagic use case loosely on the characteristics of the
+We model our pelagic use case loosely on the characteristics of the
 Western and Central Pacific Ocean tuna fisheries. We simulate
 trajectories of `r n_distinct(names(blue_fauna))` species commonly
-caught in the region, including both the highly abundance skipjack tuna
-(*Katsuwonus pelamis*) to the heavily depleted oceanic whitetip shark
-(*Carcharhinus longimanus*) (@fig-blue-ssb). We pulled publicly
-available data on catch-per-unit-effort of each of these species from
-the WCPO as a proxy for baseline habitat distributions (@fig-blue).
+caught in the region, including both the highly abundant skipjack tuna
+to the heavily depleted oceanic whitetip shark (*Carcharhinus
+longimanus*) (@fig-blue-ssb). We use publicly available data on
+catch-per-unit-effort of each of these species from the WCPO as a proxy
+for baseline habitat distributions, noting that where possible
+fishery-independent abundance indices would be preferable (@fig-blue).
 
@@
 longline fleet, and dome-shaped for the purse-seine fleet, with
-selectivity values pulled from the literature where available. Both
-fleets operate under open-access dynamics with an effort cap. The effort
-cap was set at the level of effort that resulted in the desired levels
-of depletion for each species under open-access dynamics (@fig-blue),
-intended to simulate a scenario where managers step in to prevent
-further expansion of fishing effort in a fully developed fishery. For
-forward-simulation, open-access dynamics can result in effort decreasing
-in response to profitability, but cannot result in effort beyond the
-effort cap set for each fleet.
+selectivity values set to broadly match the patterns reported in
+relevant [stock
+assessments](https://oceanfish.spc.int/en/ofpsection/sam/sam/210-stock-assessment-input-and-result-files-)
+for each species. Both fleets operate under open-access dynamics with an
+effort cap. The effort cap was set at the level of effort that resulted
+in the desired levels of depletion for each species under open-access
+dynamics (@fig-blue), intended to simulate a scenario where managers
+step in to prevent further expansion of fishing effort in a fully
+developed fishery. For forward-simulation, open-access dynamics can
+result in effort decreasing in response to profitability, but cannot
+result in effort beyond the effort cap set for each fleet.
 
@@
 running one simulation with and another without the climate-drive range
-shift. We then compared effect of this range shift on food security and
-conservation outcomes from MPA networks designed based on the pre-range
-shift world.
+shift. We then compared the effects of this range shift on food
+production and conservation outcomes from MPA networks designed based on
+the pre-range shift world.
 
@@
 
-For the coral reef example, both the area and target_fishing MPA design
-strategies were capable of producing a Pareto-improving, i.e.
-potentially win-win, outcome where the total spawning biomass divided by
-unfished spawning biomass across all species could be increased by up to
-200% while maintaining at least the same levels of fisheries yields as
-would be caught in the absence of any MPAs. Conversely, achieving
-increases in total fish biomass through the rate or avoid_fishing
-strategies resulted in immediate losses in total yields, and in losses
-to both yields and biomass in some cases (@fig-coral-results).
-
-However, these macro-level changes in total yield and conservation
-outcomes mask some internal tradeoffs. Nearly all of the food security
-benefits from this MPA network accrued to fleet one, with catches by
-fleet two being negatively impacted under nearly all scenarios,
-including those that resulted in the highest total gains to fleet one
-and subsequently to the system as a whole. For example, the MPA sizes
-and designs that result in the largest increase in yields for fleet one
-would produce in near total loss of yields for fleet two.
-
-Both the rate and avoid_fishing MPA placement strategies produced lower
-total fishery gains but minimized the negative impacts on fleet two.
-However, these strategies produced much lower conservation gains than
-the area and target_fishing strategies, only benefiting deep_snapper
-meaningfully, and in fact resulting in slight losses to the nearshore
-snapper population for MPA sizes less than roughly 75%. Both the area
-and target_fishing strategies resulted in rapid conservation gains
-across all of the simulated species.
+For the nearshore coral reef example, both the *area* and
+*target_fishing* MPA design strategies were capable of producing a
+Pareto-improving, i.e. potentially win-win, outcome where the total
+spawning biomass divided by unfished spawning biomass across all species
+increase by up to 200% while maintaining at least the same levels of
+fisheries yields as would be caught in the absence of any MPAs.
+Conversely, achieving increases in total fish biomass through the *rate*
+or *avoid_fishing* strategies resulted in immediate losses in total
+yields, and in losses to both yields and biomass in some cases
+(@fig-coral-results).
+
+These macro-level changes in total yield and conservation outcomes mask
+some internal tradeoffs. Nearly all of the food production benefits from
+this MPA network accrued to fleet one, while catches by fleet two were
+negatively impacted under nearly all scenarios, including those that
+resulted in the highest total gains to fleet one and subsequently to the
+system as a whole. For example, the MPA sizes and designs that result in
+the largest increase in yields for fleet one would produce a near total
+loss of yields for fleet two.
+
+Both the *rate* and *avoid_fishing* MPA placement strategies produced
+lower total fishery gains but minimized the negative impacts on fleet
+two. However, these strategies produced much lower conservation gains
+than the *area* and *target_fishing* strategies, only benefiting
+deep_snapper meaningfully, and in fact resulting in slight losses to the
+nearshore snapper population for MPA sizes less than roughly 75% of the
+simulated space. Both the *area* and *target_fishing* strategies
+resulted in rapid conservation gains across all of the simulated
+species.
 
 MPAs were capable of producing positive outcomes in terms of total
-biomass and yields in this coral reef example, but none of the
-strategies on the table came without tradeoffs across species or fleets.
-The framework we present here could be further used then by a community
-such as the one simulated here to explore new design strategies, or to
-assess options for distributing fishery benefits across communities that
-may have very different outcomes from the same MPA network.
+biomass and yields in the nearshore coral reef example, but all the
+tested strategies resulted in tradeoffs. `marlin` could be further used
+by a community such as the one simulated here to explore new design
+strategies, or to assess options for distributing fishery benefits
+across communities that may experience very different outcomes from the
+same MPA network.
 
@@
 #| label: fig-coral-results
-#| fig-cap: "Frontier in conservation and yield outcomes under different MPA design strategies (A) for coral reef case study.
+#| fig-cap: "Frontier in conservation and yield outcomes under different MPA design strategies (A) for nearshore coral reef case study.
 Change in yield for each fleet (B) and depletion per species (C) as a function of MPA size and MPA design strategy for the coral reef case study."
@@
 
-For the pelagic case study we focused on the climate resilience of two
-candidate MPA design strategies; target_fishing and rate, where both MPA
-networks are designed based on the species distributions at the start of
-the time series. We then compared the outcomes of these MPA networks for
-fisheries yield and spawning biomass under the two scenarios where the
-species distributions remain constant, and when the species undergo a
-range shift.
-
 The yield impacts of the proposed MPA strategies were much more
@@
 strategy produced increases in fisheries yields at sizes between 25-50%
-of the area; that same MPA network resulted in the near complete loss of
-yields from the purse seine fleet when applied to the climate change
-scenario. In contrast, the yield outcomes for the longline fleet were
-relatively insensitive to the simulated range shifts
+of the simulated area; that same MPA network resulted in the near
+complete loss of yields from the purse seine fleet when applied to the
+climate change scenario. In contrast, the yield outcomes for the
+longline fleet were relatively insensitive to the simulated range shifts
 (@fig-blue-results).
@@
 would produce much smaller and slower gains under climate change.
-However, for blue sharks this trend is flipped, with more rapid
+However, for blue sharks this trend is inverted, with more rapid
 conservation gains under the range shift scenario than the *status quo*
@@
 Marine ecosystems are driven by complex social-ecological dynamics.
-Communities must often make decisions on how to manage these systems
-based on limited empirical evidence. Modeling tools such as `marlin` can
-help users answer important scientific questions and design marine
-management policies informed by better understanding of sensitivities to
-key uncertainties.
-
-<!-- `marlin` is to our knowledge the first bio-economic modeling tool that -->
-
-<!-- allows scientists and decision makers to efficiently simulate the -->
-
-<!-- outcomes of management strategies such as MPAs across multiple fleets -->
-
-<!-- and species while accounting for layered regulations, economic -->
-
-<!-- incentives, age structure, and realistic movement dynamics across a -->
-
-<!-- range of potential social-ecological conditions. xx this sentence annoys -->
-
-<!-- me but feels like we need to make something like it explicit to convince -->
-
-<!-- reviewers maybe less familiar with the modeling literature xx. -->
-
-There are a range of other useful modeling applications that can support
-spatio-temporal decision making in marine ecosystems. However, the
-`marlin` framework presented here fills an important gap in the modeling
-literature and in the policy support toolbox. More end-to-end models
-such as ATLANTIS [@audzijonyte2019] , OSMOSE [@shin2001], Ecopath with
-Ecosim [@christensen2004], POSEIDON [@bailey2019], or SEAPODYM
-[@lehodey2008] are capable of representing tremendous amounts of
-complexity, but can be extremely time consuming to design and run. Tools
-such as virtualspecies [@leroy2016] or STEPS [@visintin2020] can help
-model the distribution of species in space and time as a function of
-environmental covariates, but lack explicit process models for the xx
-Jim, can you expand on this section, I'm no too familiar with these
-methods xx. Simplified bio-economic models such as those used in
-@hastings2017, @sala2021, and @cabral2019 can be applied at scale and
-provide analytically tractable results, but as a result must abstract
-over many bio-economic dynamics that can be important for more tactical
+Communities must make decisions on how to manage these systems based on
+limited empirical evidence. Modeling tools such as `marlin` can help
+users answer important scientific questions and design marine management
+policies informed by a better understanding of sensitivities to key
+uncertainties.
+
+The role of this paper is not to conduct a review of the many
+tools for spatial-temporal modeling available in the literature.
+However, we highlight here the specific gaps that we feel `marlin` fills
+in the modeling literature and in the policy support toolbox using some
+selected publications. More end-to-end models such as ATLANTIS
+[@audzijonyte2019] , OSMOSE [@shin2001], Ecopath with Ecosim
+[@christensen2004], POSEIDON [@bailey2019], or SEAPODYM [@lehodey2008]
+are capable of representing tremendous amounts of complexity, but can be
+extremely time consuming to design and run.
+
+Tools such as virtualspecies
+[@leroy2016], STEPS [@visintin2020], RangeShifter [@bocedi2014], and
+[SMS](https://www.aquaveo.com/software/sms-surface-water-modeling-system-introduction)
+can be efficiently constructed to model the dispersal and distribution
+of species in space and time as a function of environmental covariates.
+However, these typically require specifying a covariate-response curve
+without explicitly acknowledging how this arises from habitat-specific
+movement and demography (virtualspecies), or model movement using a
+dispersal kernel (RangeShifter, STEPS) or a least-cost path algorithm
+(SMS, STEPS). By contrast, we specify a continuous-time Markov chain
+movement model, which integrates multiple paths (including their
+path-dependent probability based on intervening habitat types and
+species preferences) while using scale-free parameters that can be
+measured experimentally in laboratory or tagging studies.
+
+Simplified bio-economic models such as those used in @hastings2017,
+@sala2021, and @cabral2019 can be applied at scale and provide
+analytically tractable results, but as a result must abstract over many
+bio-economic dynamics that can be important for more tactical
 applications. `marlin` complements the existing spatial marine modeling
@@
 In this paper we have documented the core system of equations underlying
-`marlin`, and demonstrated how this framework can be used to support
-efforts such as MPA planning. The variation in both the magnitude and
+`marlin`, and demonstrated how `marlin` can be used to support efforts
+such as MPA planning. The variation in both the magnitude and
 distribution of costs and benefits stemming from alternative design
 choices and states of nature shown in our case study results
-demonstrates the value in our method. Beyond the current examples, we
-envision `marlin` being applicable to research on dynamic ocean
-management, range shifts, management strategy evaluation, policy
-interactions, and spatial stock assessment. In addition, the
-process-based movement model used in `marlin` can directly use empirical
-estimates of movement dynamics derived in the manner of @thorson2021a,
-providing a link between empirical and simulation based approaches to
-marine resource management. Future iterations of the model may include
-life history modules for a broader range of species types such as marine
-mammals, sea birds, and sea turtles.
+demonstrates the value in our method. For example, our coastal coral
+reef example illustrates both the potential for MPAs to benefit
+conservation and food production in these systems, and the potential for
+the same MPA to benefit some fleets while harming others when fishing
+fleets affected by an MPA do not share the same objectives. Our pelagic
+case study illustrates how `marlin` can be used to assess the climate
+resiliency of spatial management strategies, and identify strategies
+that best meet the needs of both current and future conditions. Results
+from this type of work could be used to help prioritize communities and
+species at particular risk to climate change impacts.
+
+Beyond the current examples, we envision `marlin` being applicable to
+research on dynamic ocean management, range shifts, management strategy
+evaluation, policy interactions, and spatial stock assessment. In
+addition, the process-based movement model used in `marlin` can directly
+use empirical estimates of movement dynamics derived in the manner of
+@thorson2021a, providing a link between empirical and simulation based
+approaches to marine resource management. Future iterations of the model
+may include life history modules for a broader range of species types
+such as marine mammals, sea birds, and sea turtles.
 

### 2022-12-23 03_marlin_paper.qmd
now with scale-free CTMC and some edits (#7)

@@
 
+if (!file.exists(here("data","sci_to_com.csv"))){
   sci_to_com <- taxize::sci2comm(unique(marlin_inputs$scientific_name),simplify= TRUE,db='ncbi') %>% 
@@
 
+  write_csv(sci_to_com, file = here("data","sci_to_com.csv"))
+
+} else{
+ sci_to_com <-  read_csv(file = here("data","sci_to_com.csv"))
+
+}
+
+print(results_path)
 results_files <- list.files(file.path(results_path))
@@
 
-`marlin` follows a "models of intermediate complexity" (MICE) philosophy
+`marlin` follows a "models of intermediate complexity" philosophy
 [@plagányi2014] targeted specifically towards capturing heterogeneous
@@
 age-structured animal populations affected by one or more fishing fleets
-in time steps specified by the user (generally annual or seasonal).
+in time steps and a spatial resolution specified by the user.
+
 While `marlin` does not simulate species interactions, fishing fleets in
@@
 
-As a demonstration of the use of this modeling two, we explore two
+As a demonstration of the use of this modeling tool, we explore two
 general situations
@@
 1.  Trade-offs and distributional outcomes for food production and
-    multi-species biomass in a coral-reef fishery
+    multi-species biomass in a coastal coral-reef fishery
 2.  Implications of climate-driven range shifts for MPA design in a
@@
 
-We chose these two examples to illustrate the use of the model in
-contrasting systems in which MPAs are increasingly considered.
+We chose these two examples to illustrate the use of `marlin` in
+contrasting systems in which spatial management strategies such as MPAs are increasingly considered.
 
 Below we provide a summary of the `marlin` model, as well as details of
-the generalized simulation exercises and case study analyses. We note
+the case studies. We note
 that `marlin` is a complex software package with many options to control
@@
 [@rcoreteam2021] and C++, integrated through the `Rcpp` package
-@eddelbuettel2018 .
+[@eddelbuettel2018].
 
@@
 using both diffusion and "taxis" (active movement towards preferred
-habitat), experiences natural and potentially fishing mortality, and if
+habitat), experience natural and potentially fishing mortality, and if
 in a spawning season spawn using one of the possible forms of density
@@
 auto-correlated stochastic deviations in the amount of offspring
-produced, generally called "recruits".
-
-These species are caught by fishing fleets. A fishing fleet is defined
+produced, generally called "recruitment deviates".
+
+These species can be caught by fishing fleets. A fishing fleet is defined
 in `marlin` by a set of fishers that can have different fishing skill,
@@
 according to a specified spatial allocation rule, for example in
-proportion to profit-per-unit-effort, conditional on the presence of any
-spatial restrictions such as MPAs.
+proportion to profit-per-unit-effort, conditional on management policies in places such as quotas and/or the presence of any spatial restrictions such as MPAs.
 
 Unconstrained by management, the total amount of effort exerted by each
-fleet can follow one of two dynamics: open-access or constant-effort.
+fleet can follow one of two dynamics: open access or constant effort.
 Under constant effort, the total amount of effort of each fleet is fixed
 over time, with the possible exception of attrition due to MPA
-placement. Under open-access, the total amount of effort in a given time
+placement. Under open access, the total amount of effort in a given time
 step is a function of the profitability of the fleet in the previous
@@
 species caught, as well as the cost of the total amount of fishing
-effort per fleet and the distance from port that the fleet operates at.
+effort per fleet and the travel costs as a function of distance from a port.
 
@@
 managers can set an effort cap per fleet, which prevents effort from
-exceeding a given amount under open-access dynamics, though effort may
+exceeding a given amount under open access dynamics, though effort may
 be reduced below this cap if required by the profit equation (i.e. the
-fleet can choose to fish less than the quota or effort cap). `marlin`
-also allows users to specify closed fishing seasons for one or more
+fleet can choose to fish less than the quota or effort cap).
+
+`marlin` also allows users to specify closed fishing seasons for one or more
 species in the system. During closed seasons, no effort is allocated to
@@
 kinds of questions stakeholders might face if designing an MPA network
-for a nearshore coral reef or a pelagic system.
+for a coastal coral reef or a pelagic system.
 
@@
 
-Revenues *R* for fleet *f* in time step *t* are a function of the total
-amount caught of each species (*Y*) and its price ($\Pi$). The amount
+Revenues for each fleet in a time step are a function of the total
+amount caught of each species and its price ($\Pi$). The amount
 caught is a function of the selectivity at age of each species for each
@@
 
-We approximate the dome-shaped form as a simple normal distribution with
+We approximate the dome-shaped form as a normal distribution with
 mean $l^{sel}$ and standard deviation $\sigma$. The normal density
@@
 Users can tune their fleets in one of two ways. First, they can specify
-a target exploitation rate *u* for each species in their model run.
+a target exploitation rate *u* for each species in their simulation.
 Taking all the other parameters of the model as given, `marlin` then
@@
 our model. Primarily, our model is capable of tracking multiple fleets
-at once. This means that the IFD for a single fleet would be a
-game-theoretic outcome of the actions of each of the individual fleet
+at once. This means that the IFD for a single fleet would commonly be modeled as a Nash-equilibrium based on the actions of each of the individual fleet
 conditional on the actions of all other fleets. While possible to
-implement this would slow down our model runs to the point of
-practically preventing large-scale evaluation of the effects of MPAs on
-bycatch.
+implement, this would slow down our model runs to the point of
+practically preventing large-scale evaluation of the spatial management policies.
 
@@
 core level. However, due to the complexity of parameterizing cost
-functions fleet dynamics are often evaluated based on yield or revenue
+functions, fleet dynamics are often evaluated based on yield or revenue
 alone, and so we include those scenarios here to allow users to evaluate
@@
 supply default values for that species based on the values reported in
-FishLife [@thorson2020]. All of the life history values can
-alternatively be replaced with user-supplied values as needed; the
-default values simply ensure that reasonable and internally consistent
+FishLife [internet connection required, @thorson2020]. All of the life history values can
+alternatively be replaced with user-supplied values as needed and should be for more tactical applications; the default values simply ensure that reasonable and internally consistent
 values for the species in question are supplied as a starting point.
@@
 (CTMC), as described in @thorson2021a. Within this framework, the model
-allows for movement to be broken down into three components of
-*diffusion*, *advection* (follow currents), and *taxis* (active movement
-towards say a preferred habitat). For now, our model focuses just on the
+allows for movement to be broken down into three components of *advection* (follow currents), and *taxis* (active movement towards preferred habitat) and *diffusion* (essentially remaining variation in movement not explained by diffusion or taxis).
+For now, `marlin` focuses just on the
 diffusion and taxis components of this model, though future extensions
@@
 $\boldsymbol{M}_{t,s,a}$. $\boldsymbol{M}_{t,s,a}$ is calculated as the
-matrix exponent of the sum of the diffusion and taxis matrices scaled by
-the width of the time step that the model is running under.
-
-$$
-\boldsymbol{M}_{t,s,a} = e^{\Delta t(\boldsymbol{D}_{t,s,a} + \boldsymbol{X}_{t,s,a})} 
+matrix exponent of the sum of the diffusion $\boldsymbol{D}$ and taxis $\boldsymbol{\tau}$ matrices scaled by the width of the time step $\Delta_{t}$ and the length of the edge of each cell $\Delta_d$ that the model is running under.
+So for example, if we are running a model on a monthly timestep, and we define the units of time in years, then $\Delta_{t} = 1/12$.
+Similarly, if we are running a model where one cell equals 100km^2^, then $\Delta_d = 100km$.
+This parameterization allows users to set the effective area of the spatial domain through two avenues; the number of cells, which effectively scales the resolution of the model, and the area of each cell, which scales the spatial extent of the simulation.
+
+$$
+\boldsymbol{M}_{t,s,a} = e^{\Delta_{t}(\boldsymbol{D}_{t,s,a} + \boldsymbol{\tau}_{t,s,a})} 
 $$ {#eq-movement}
 
-Where $\boldsymbol{D}$ is the diffusion matrix and $\boldsymbol{X}$ is
-the taxis matrix. The diffusion matrix is a function of a user specified
-diffusion rate [@thorson2021a].
-
-We slightly modify the taxis matrix, denoted as $\tau$ in the following
-manner
+The individual components of the diffusion matrix are filled based on an adjacency matrix, which defines whether two cells are both adjacent and water (as opposed to land or another physical barrier), and a diffusion rate $D$ definend in units of length of the side of a cell per unit of time.
+For example, if we are defining the time units as years and the distance units as kilometers, for a tuna $D$ might be 1,000 $\frac{KM}{Year}$.
+The $\Delta_{t}$ and $\Delta_{p}$ parameters then translate the movement rate $D$ to match the time step and cell size used in a simulation.
+This "scale free" parameterization means that appropriate value of $D$ can be identified for a species and then set, regardless of the time step or cell size used in the simulation model itself.
+
+$$ 
+d_{p1,p2,t,s,a} = \begin{cases}
+      = \frac{\Delta_{t}}{\Delta_{d}^2}D & \text{if p2 and p1 are adjacent}\\
+     = -\sum_{p' \neq p1} d_{p1,p2,t,s,a} & \text{if p1 = p2}\\
+     = 0 & \text{otherwise.}
+\end{cases}
+$$ {#eq-diffusion}
+
+The taxis matrix $\boldsymbol{\tau}$ is built off of the difference in log habitat quality between each patch, the adjacency matrix, and temporal and spatial resolution of the model ($\Delta_{t}$ and $\Delta_{p}$), as well as a scalar term $\lambda$.
+$\lambda$ scales the magnitude of the habitat
+gradient between patches relative to the diffusion
+parameter $D$. When the diffusion and taxis matrices are estimated
+jointly from actual data in the manner of @thorson2021a, $\lambda$
+should be set to zero. However, when users are manually creating both
+the diffusion and taxis matrices for simulation, the relative strength of diffusion to
+taxis must be defined by the user. Setting higher values of $\lambda$
+allows users to simulate a scenario where taxis is much more significant
+than diffusion.
 
@@
 \tau_{p1,p2,t,s,a} = \begin{cases}
-      = (1 + \lambda d_{s,f})e^{h(p2,t,s,a) - h(p1,t,s,a)} & \text{if p2 and p1 are adjacent}\\
-     = -\sum_{p' \neq p1} (1 + \lambda d_{s,f})e^{h(p2,t,s,a) - h(p1,t,s,a)} & \text{if p1 = p2}\\
+      = \frac{\Delta_{t}}{\Delta_{d}}(1 + \lambda d_{s,f})e^{h(p2,t,s,a) - h(p1,t,s,a)} & \text{if p2 and p1 are adjacent}\\
+     = -\sum_{p' \neq p1} \tau_{p1,p2,t,s,a} & \text{if p1 = p2}\\
      = 0 & \text{otherwise.}
@@
 $$ {#eq-taxis}
-
-The $\lambda$ term is a scalar in which the magnitude of the habitat
-gradient between patches can be scaled relative to the diffusion
-parameter $d$. When the diffusion and taxis matrices are estimated
-jointly from actual data in the manner of @thorson2021a, $\lambda$
-should be set to zero. However, when users are manually creating both
-the diffusion and taxis matrices, the relative strength of diffusion to
-taxis must be defined by the user. Setting higher values of $\lambda$
-allows users to simulate a scenario where taxis is much more significant
-than diffusion.
 
@@
 move around barriers rather than over them. This is different than
-setting habitat in a patch to zero; an animal will preferentially move
+setting habitat in a patch to zero; an animal will not preferentially move
 towards a patch with zero habitat all else being equal, but can move
-over a zero habitat patch towards another better patch.
+over a zero habitat patch towards another better patch. They most move around a physical barrier.
 
@@
 
-One challenge is that the coefficients of the model matrices inside
-@eq-movement can be hard to interpret. To facilitate this, `marlin`
-allows users to specify a maximum number of cells that adults or
-recruits diffuse in a year. The model will then tune the diffusion
-parameter and habitat gradients to achieve this desired level of
-movement.
-
 #### Population Growth
@@
 be used to compare the total and distributional impacts of MPAs designed
-in a heavily fished coral reef ecosystem. In the second, we demonstrate
+in a heavily fished coastal coral reef ecosystem. In the second, we demonstrate
 how `marlin` can be used to assess components of climate resilience of
@@
 
-Across all of our case studies we make use of four potential "rule of
-thumb" MPA design strategies (xx add citation here) :
+Across all of our case studies we make use of two potential "rule of
+thumb" MPA design strategies:
 
@@
 
-2.  Avoid_fishing (*avoid_fishing*): MPAs locations are prioritized
-    inversely to fishery catches. So, patches with low catches are
-    prioritized over patches with high catches.
-
-3.  Target_fishing (*target_fishing*): MPAs locations are prioritized
-    proportionally to fishery catches. So, patches with high total
+2.  Target_fishing (*target_fishing*): MPAs locations are prioritized
+    proportional to fishery catches. So, patches with high total
     catches are prioritized over patches with low catches.
-
-4.  Area target (*area*): For the "area" strategy, we simply assign
-    patches in sequential order until the desired total MPA size is
-    reached.
 
@@
 for continuity (i.e. the selected MPA design can be as patchwork as
-dictated by the data).
+dictated by the data). We ran the simulation on a quarterly level, meaning that $\Delta_t = 1/ 4$ and set the area of each cell to an arbitrary value of 1 (meaning $\Delta_p = 1$ ).
 
@@
 production and conservation outcomes from MPA networks designed based on
-the pre-range shift world.
+the pre-range shift world. We ran the simulation on a quarterly level, meaning that $\Delta_t = 1/4$ and set the area of each cell to an arbitrary value of 1 (meaning $\Delta_p = 1$ ).
 
@@
 
-For the nearshore coral reef example, both the *area* and
-*target_fishing* MPA design strategies were capable of producing a
-Pareto-improving, i.e. potentially win-win, outcome where the total
+For the coastal coral reef example the *target_fishing* MPA design strategies was capable of producing a Pareto-improving, i.e. potentially win-win, outcome where the total
 spawning biomass divided by unfished spawning biomass across all species
@@
 fisheries yields as would be caught in the absence of any MPAs.
-Conversely, achieving increases in total fish biomass through the *rate*
-or *avoid_fishing* strategies resulted in immediate losses in total
-yields, and in losses to both yields and biomass in some cases
+Conversely, achieving increases in total fish biomass through the *rate* strategy resulted in immediate losses in total yields, and in losses to both yields and total biomass in some cases
 (@fig-coral-results).
@@
 some internal tradeoffs. Nearly all of the food production benefits from
-this MPA network accrued to fleet one, while catches by fleet two were
-negatively impacted under nearly all scenarios, including those that
-resulted in the highest total gains to fleet one and subsequently to the
-system as a whole. For example, the MPA sizes and designs that result in
-the largest increase in yields for fleet one would produce a near total
-loss of yields for fleet two.
-
-Both the *rate* and *avoid_fishing* MPA placement strategies produced
+this MPA network accrued to fleet one, while catches by fleet two were generally
+negatively impacted under both scenarios. For example, the MPA sizes and designs that result in the largest increase in yields for fleet one would produce a near total
+loss of fish catch for fleet two.
+
+The *rate* MPA placement strategy produced
 lower total fishery gains but minimized the negative impacts on fleet
-two. However, these strategies produced much lower conservation gains
-than the *area* and *target_fishing* strategies, only benefiting
-deep_snapper meaningfully, and in fact resulting in slight losses to the
-nearshore snapper population for MPA sizes less than roughly 75% of the
-simulated space. Both the *area* and *target_fishing* strategies
+two. However, this strategy produced much lower conservation gains
+than the *target_fishing* strategy, only benefiting
+deep snapper meaningfully, and in fact resulting in slight losses to the
+nearshore snapper and grouper populations for MPA sizes less than roughly 75% of the
+simulated space. The *target_fishing* strategy
 resulted in rapid conservation gains across all of the simulated
@@
 MPAs were capable of producing positive outcomes in terms of total
-biomass and yields in the nearshore coral reef example, but all the
-tested strategies resulted in tradeoffs. `marlin` could be further used
-by a community such as the one simulated here to explore new design
+biomass and yields in the coastal coral reef example, but all the
+tested strategies resulted in tradeoffs across at least one dimension. `marlin` could be further used
+by a community such as the one simulated here to explore alternative design
 strategies, or to assess options for distributing fishery benefits
-across communities that may experience very different outcomes from the
+across communities that may experience very different food provision outcomes from the
 same MPA network.
@@
 #| label: fig-coral-results
-#| fig-cap: "Frontier in conservation and yield outcomes under different MPA design strategies (A) for nearshore coral reef case study.
+#| fig-cap: "Frontier in conservation and yield outcomes under different MPA design strategies (A) for coastal coral reef case study.
 Change in yield for each fleet (B) and depletion per species (C) as a function of MPA size and MPA design strategy for the coral reef case study."
@@
 sensitive to climate impacts for the purse seine fishery. Under the
-*status quo* species distribution, the target_fishing MPA design
+*status quo* species distribution, the *target_fishing* MPA design
 strategy produced increases in fisheries yields at sizes between 25-50%
@@
 outcomes depending on the climate scenario. For bigeye tuna, under the
-*status quo* conditions the target fishing MPA design strategy would
-result in rapid increases in the population, but that same MPA network
-would produce much smaller and slower gains under climate change.
+*status quo* conditions the *target_fishing* MPA design strategy
+resulted in rapid increases in the population, but that same MPA network produced much smaller and slower gains under climate change.
 However, for blue sharks this trend is inverted, with more rapid
@@
 Marine ecosystems are driven by complex social-ecological dynamics.
-Communities must make decisions on how to manage these systems based on
+Communities must often make decisions on how to manage these systems based on
 limited empirical evidence. Modeling tools such as `marlin` can help
@@
 conservation and food production in these systems, and the potential for
-the same MPA to benefit some fleets while harming others when fishing
-fleets affected by an MPA do not share the same objectives. Our pelagic
+the same MPA to benefit some fleets and species while harming others when fishing
+fleets affected by an MPA do not share the same objectives and species are not uniformly distributed. Our pelagic
 case study illustrates how `marlin` can be used to assess the climate
@@
 Even the most complex marine model is a stylized cartoon of the true
-dynamics at play in these systems. However, for all their limitations
+dynamics at play in the ocean. However, for all their limitations
 models such as `marlin` can help users understand factors that drive the

### 2022-12-23 scripts/01_coral_case_study.R

@@ -462,11 +462,11 @@ mpa_sim <- simmar(
 
 if (run_coral_example == TRUE){
 
-  plan(multisession, workers = experiment_workers)
+  future::plan(future::multisession, workers = experiment_workers)
 
   case_study_experiments <-
     expand_grid(
-      placement_strategy = c("rate", "avoid_fishing", "target_fishing", "area"),
+      placement_strategy = c("rate", "target_fishing"),
       prop_mpa = seq(0, 1, by = 0.05),
       critters_considered = length(fauna),
       placement_error = c(0),

### 2023-01-05 03_marlin_paper.qmd
echelle edits

@@
 `marlin` also allows users to specify closed fishing seasons for one or more
-species in the system. During closed seasons, no effort is allocated to
-patches that would result in capture of the covered species, potentially
-resulting in zero effort if the species cannot be avoided. Under these
-conditions, the amount of effort in the next open season is a function
-of the effort and profits in the last open fishing season. Lastly,
+species in the system.
+When all species in a simulation share the same closed seasons, no fishing occurs in those seasons.
+When not all species share the same closed seasons, fishing still occurs in the closed seasons, but the fleet avoids catching the species for which fishing is closed in a given season, assuming no discard mortality.
+When closed seasons are in effect, the amount of effort in the next open season is a function of the effort and profits in the last open fishing season.
+Lastly,
 managers can specify locations of no-take MPAs, which could in theory
@@
 Each fishing fleet (*f*) generates catches or yield (*Y*), revenues
-(*R*), costs (*C*) from individual species (*s*) that it targets in a
+(*R*), and costs (*C*) from fishing individual species (*s*) that it targets in a
 given time step (*t*) and patch (*p*). The fleet then makes decisions
-around how much and where to fish, subject to regulatory constraints,
+around fishing locations and intensity, subject to regulatory constraints,
 based on the total profitability across all species in space and time.
@@
 Revenues for each fleet in a time step are a function of the total
-amount caught of each species and its price ($\Pi$). The amount
+amount of each species caught and its price ($\Pi$). The amount
 caught is a function of the selectivity at age of each species for each
@@
 (CTMC), as described in @thorson2021a. Within this framework, the model
-allows for movement to be broken down into three components of *advection* (follow currents), and *taxis* (active movement towards preferred habitat) and *diffusion* (essentially remaining variation in movement not explained by diffusion or taxis).
+allows for movement to be broken down into three components of *advection* (follow currents), and *taxis* (active movement towards preferred habitat) and *diffusion* (essentially remaining variation in movement not explained by advection or taxis).
 For now, `marlin` focuses just on the
-diffusion and taxis components of this model, though future extensions
+diffusion and taxis components of this model, assuming that advection is zero, though future extensions
 could easily incorporate advection vectors from oceanographic models. In
@@
 
-The individual components of the diffusion matrix are filled based on an adjacency matrix, which defines whether two cells are both adjacent and water (as opposed to land or another physical barrier), and a diffusion rate $D$ definend in units of length of the side of a cell per unit of time.
+The individual components of the diffusion matrix are filled based on an adjacency matrix, which defines whether two cells are both adjacent and water (as opposed to land or another physical barrier), and a diffusion rate $D$ defined in units of length of the side of a cell per unit of time.
 For example, if we are defining the time units as years and the distance units as kilometers, for a tuna $D$ might be 1,000 $\frac{KM}{Year}$.
@@
 where *BH* is the Beverton-Holt recruitment function, *ssb* is
-spawning-stock-biomass. $l_a$ is the mean length at at age, $l_{sel}$ is
+spawning-stock-biomass, $l_a$ is the mean length at at age, $l_{sel}$ is
 the length at which on average 50% of individuals are selected by the
@@
 for a multi-fleet and spatial-temporal model such as `marlin`. MSY and
-the fishing mortality rate that would produce MSY, $F_{MSY}$ are a
+the fishing mortality rate that would produce MSY, $F_{MSY}$, are a
 function of fishery selectivity. Fishery selectivity in this model can
@@
 
-2.  Target_fishing (*target_fishing*): MPAs locations are prioritized
-    proportional to fishery catches. So, patches with high total
+2.  Target Fishing (*target_fishing*): MPAs locations are prioritized
+    proportional to fishery catches. Patches with high total
     catches are prioritized over patches with low catches.
@@
 objectives. In theory, the design of MPA networks can be optimized
-through the use of a modeling framework, and indeed conditional on the
-validity of the model itself this process is likely to produce better
+through the use of a modeling framework, and depending on the
+validity of the model, this process is likely to produce better
 outcomes than manually-designed strategies [@rassweiler2012;
@@
 users. We allow all MPAs to be designed in a "mosaic" fashion, but users
-can easily extend the analysis to for example compare outcomes between
-"contiguous" and "mosaic" MPA designs [@pons2022].
+can easily extend the analysis to compare outcomes between
+"contiguous" (MPA is made up of one continuously connected block) and "mosaic" (MPAs can be separated in space) MPA designs [@pons2022].
 
@@
 followed by the sharks. Shallow-reef snappers stay in reefs closer to
-shower year-round, while deep-water snappers stay in the deeper reefs
+shore year-round, while deep-water snappers stay in the deeper reefs
 (@fig-coral).
@@
 sharks, as any shark caught is retained. This fleet is a subsistence
-fleet, where the local community takes every possible opportunity to
-fish that they can, and as such we model it as a "constant effort"
+fleet, where the local community takes advantage of every possible opportunity to
+fish, and as such we model it as a "constant effort"
 fishery.
@@
 for sharks due to the requirements of a certification program through
-which they sell their deep-water snapper, though accidental captures do
+which they sell their deep-water snapper. Accidental captures (bycatch) do
 occur, which results in mortality. This fleet operates under open-access
-dynamics, as fishing is not the primary means of subsistence and so
+dynamics, as fishing is not the primary means of subsistence;
 short-term effort expands and contracts in response to profitability of
@@
 for continuity (i.e. the selected MPA design can be as patchwork as
-dictated by the data). We ran the simulation on a quarterly level, meaning that $\Delta_t = 1/ 4$ and set the area of each cell to an arbitrary value of 1 (meaning $\Delta_p = 1$ ).
+dictated by the data). We ran the simulation on a quarterly level ($\Delta_t = 1/4$) and set the area of each cell to an arbitrary value of 1 ($\Delta_p = 1$).
 
@@
 We model our pelagic use case loosely on the characteristics of the
-Western and Central Pacific Ocean tuna fisheries. We simulate
+Western and Central Pacific Ocean (WCPO) tuna fisheries. We simulate
 trajectories of `r n_distinct(names(blue_fauna))` species commonly
@@
 catch-per-unit-effort of each of these species from the WCPO as a proxy
-for baseline habitat distributions, noting that where possible
+for baseline habitat distributions, noting that where possible,
 fishery-independent abundance indices would be preferable (@fig-blue).
@@
 production and conservation outcomes from MPA networks designed based on
-the pre-range shift world. We ran the simulation on a quarterly level, meaning that $\Delta_t = 1/4$ and set the area of each cell to an arbitrary value of 1 (meaning $\Delta_p = 1$ ).
+the pre-range shift world. We ran the simulation on a quarterly level ($\Delta_t = 1/4$) and set the area of each cell to an arbitrary value of 1 ($\Delta_p = 1$).
 
@@
 These macro-level changes in total yield and conservation outcomes mask
-some internal tradeoffs. Nearly all of the food production benefits from
-this MPA network accrued to fleet one, while catches by fleet two were generally
+some internal tradeoffs. Nearly all of the food production benefits from the simulated MPA networks accrued to fleet one, while catches by fleet two were generally
 negatively impacted under both scenarios. For example, the MPA sizes and designs that result in the largest increase in yields for fleet one would produce a near total
@@
 The role of this paper is not to conduct a review of the many
-tools for spatial-temporal modeling available in the literature.
+tools for spatial-temporal modeling available in the literature, each of which provide useful functionality for specific applications.
 However, we highlight here the specific gaps that we feel `marlin` fills
@@
 are capable of representing tremendous amounts of complexity, but can be
-extremely time consuming to design and run.
+extremely time consuming to design and run. `marlin` allows some of the realism of these end-to-end models while being simpler and faster to parameterize and run.
 
@@
 dispersal kernel (RangeShifter, STEPS) or a least-cost path algorithm
-(SMS, STEPS). By contrast, we specify a continuous-time Markov chain
+(SMS, STEPS). By contrast, `marlin` uses a continuous-time Markov chain
 movement model, which integrates multiple paths (including their
@@
 bio-economic dynamics that can be important for more tactical
-applications. `marlin` complements the existing spatial marine modeling
+applications. `marlin` allows for more realistic representations of spatial social-ecological systems while maintaining processing speed.
+
+In sum, `marlin` complements the existing spatial marine modeling
 literature by allowing scientists, decision makers, and other
@@
 realistic bio-economic dynamics on academic and applied problems.
+For example, scientists could use `marlin` to consider the relative importance of different biological variables in affecting policy outcomes.
+Managers could use `marlin` to assess the potential of proposed MPA networks to achieve stated objectives, such as maximizing conservation gains for a threatened species while minimizing or removing negative impacts to affected fishing fleets.
 

### 2023-01-09 03_marlin_paper.qmd
prepping repo for submission

@@
 params:
-  run_name: ["v0.3"]
+  run_name: ["v0.5"]
 abstract: "Abstract: Designing effective spatial management strategies is challenging because marine ecosystems are highly dynamic and opaque, and extractive entities such as fishing fleets respond endogenously to ecosystem changes in ways that depend upon ecological and policy context.
@@
 
-source(here::here("scripts", "00_setup.R"))
+source(here::here("00_setup.R"))
 
@@
 walk(rdses,~assign(str_remove(.x,".rds"), read_rds(file.path(results_path,.x)), envir = .GlobalEnv)) #load and assign all .rds objects in the appropriate results folder
+
+# simple function to make prettier labels
+titler <- function(x){
+
+  stringr::str_to_title(stringr::str_replace_all(x,"_"," "))
+
+}
 
@@
 
+We define a few commonly used terms here.
+*Yields* refers to the volume of catch from fishing activity.
+*Spawning stock biomass* refers to the total biomass of reproductively mature fish in the water, a function of numbers, weight, and sexual maturity at age.
+*Depletion* is a common term in the fisheries literature, and reflects here the *spawning stock biomass* in a given time step relative to *unfished spawning stock biomass*.
+So, a depletion value of 1 means that the population is unfished, a depletion value of 0 means the population is extinct.
+
 As a demonstration of the use of this modeling tool, we explore two
@@
 simply introduce error. As such, readers should consult the accompanying
-code (xx provide clear instructions for which files xx) to view the
+code to view the
 precise details of each simulation. Targeted tactical applications must
@@
 
-#### Rule-of-thumb MPA Design Strategies
+#### MPA Design Strategies
 
@@
 
-1.  Depletion Rate (*rate*): MPAs are placed based on the pre-MPA
+1.  *Rate*: MPAs are placed based on the pre-MPA
     depletion-weighted catch relative to the total catch in a patch. So,
@@
 
-2.  Target Fishing (*target_fishing*): MPAs locations are prioritized
+2.  *Target Fishing*: MPAs locations are prioritized
     proportional to fishery catches. Patches with high total
@@
 #| label: fig-coral
-#| fig-cap: "Distribution of fish spawning biomass (color) each season (row) under unfished conditions.
+#| fig-cap: "Distribution of fish spawning biomass (cell color, yellow = higher, blue = lower) each season (columns) under unfished conditions.
 The x-axis represents longitude, y-axis latitude.
-Numbers show the approximate location of each fleet's port.
-Rows indicate quarterly seasons that repeat each year."
+Numbers show the approximate location of each fleet's (one or two) port.
+Columns indicate quarterly seasons that repeat each year."
 
@@
   filter(between(step,1,2)) %>%
+  mutate(step = fct_recode(as.factor(step), "2" = "1.25", "3" = "1.5", "4" = "1.75")) %>% 
   mutate(Season = step) %>% 
@@
   geom_text_repel(data = coral_ports, aes(x, y, label = fleet), color = "red") +
-  facet_grid(Season~critter) +
+  facet_grid(critter~Season, labeller = labeller(critter = titler, Season = label_both)) +
   scale_x_continuous(expand = c(0,0)) +
@@
 
+# critter_biomass %>%
+#   ungroup() %>% 
+#   left_join(sci_to_com, by = "critter") %>%
+#   select(-critter) %>%
+#   rename(critter = common_name) %>%
+#   ggplot(aes(step, depletion, color = critter)) +
+#   geom_line(linewidth = 1) +
+#   scale_y_continuous(limits = c(0, 1), name = "Depletion", expand = expansion(mult = c(0.0,0.05))) +
+#   scale_x_continuous(name = "Time Step") +
+#   scale_color_discrete(name = "")
+
 critter_biomass %>%
@@
   rename(critter = common_name) %>%
-  ggplot(aes(step, depletion, color = critter)) +
-  geom_line(linewidth = 1) +
+  filter(step == max(step)) %>% 
+  ggplot(aes(reorder(critter, depletion), depletion)) +
+  geom_col() +
   scale_y_continuous(limits = c(0, 1), name = "Depletion", expand = expansion(mult = c(0.0,0.05))) +
-  scale_x_continuous(name = "Time Step") +
-  scale_color_discrete(name = "")
+  labs(x = "") + 
+  coord_flip()
 
@@
 
-For the coastal coral reef example the *target_fishing* MPA design strategies was capable of producing a Pareto-improving, i.e. potentially win-win, outcome where the total
+For the coastal coral reef example the *Target fishing* MPA design strategies was capable of producing a Pareto-improving, i.e. potentially win-win, outcome where the total
 spawning biomass divided by unfished spawning biomass across all species
@@
 fisheries yields as would be caught in the absence of any MPAs.
-Conversely, achieving increases in total fish biomass through the *rate* strategy resulted in immediate losses in total yields, and in losses to both yields and total biomass in some cases
+Conversely, achieving increases in total fish biomass through the *Rate* strategy resulted in immediate losses in total yields, and in losses to both yields and total biomass in some cases
 (@fig-coral-results).
@@
 
-The *rate* MPA placement strategy produced
+The *Rate* MPA placement strategy produced
 lower total fishery gains but minimized the negative impacts on fleet
 two. However, this strategy produced much lower conservation gains
-than the *target_fishing* strategy, only benefiting
+than the *Target Fishing* strategy, only benefiting
 deep snapper meaningfully, and in fact resulting in slight losses to the
 nearshore snapper and grouper populations for MPA sizes less than roughly 75% of the
-simulated space. The *target_fishing* strategy
+simulated space. The *Target Fishing* strategy
 resulted in rapid conservation gains across all of the simulated
@@
 #| label: fig-coral-results
-#| fig-cap: "Frontier in conservation and yield outcomes under different MPA design strategies (A) for coastal coral reef case study.
-Change in yield for each fleet (B) and depletion per species (C) as a function of MPA size and MPA design strategy for the coral reef case study."
+#| fig-cap: "Percent change in yield per fleet (A) and depletion per species (B) as a function of MPA size and MPA design strategy for the coastal coral reef case study.
+Percent change in total yield across both fleets (y-axis) and total depletion across all species (x-axis) by MPA size (point size) and placement strategy (point color) (C)."
 #| fig-height: 4
@@
          delta_biodiv = biodiv / biodiv[prop_mpa == 0] - 1) %>% 
-  ggplot(aes(delta_biodiv, delta_yield, color = placement_strategy)) + 
-  geom_vline(aes(xintercept = 0), linetype = 2) +
+  ungroup() %>% 
+    mutate(placement_strategy= fct_relabel(placement_strategy, titler)) %>% 
+  ggplot(aes(delta_biodiv, delta_yield, fill = placement_strategy, size = prop_mpa)) + 
   geom_hline(aes(yintercept = 0), linetype = 2) +
-  geom_line() + 
+  geom_point(alpha = 0.8, shape = 21) + 
   scale_x_continuous(name = "Change in Total Depletion", labels = scales::label_percent(accuracy = 1), guide = guide_axis(n.dodge = 2)) +
   scale_y_continuous(name = "Change in Yield", labels = scales::label_percent(accuracy = 1)) + 
-  scale_color_ordinal(name = '')+
-  guides(color = guide_legend(nrow = 4)) + 
-  labs(tag = "A") + 
+  scale_fill_viridis_d(name = '')+
+  scale_size(name = "MPA Size", labels = scales::label_percent(accuracy = 1), breaks = seq(0.25,1, by = 0.25)) +
+  guides(fill = guide_legend(order = 1, override.aes = list(size = 4)),
+         size = guide_legend(order = 2, ncol = 2, byrow = TRUE)) + 
+  labs(tag = "C")  + 
   theme(
-    legend.position = c(0.7, 1),
-    legend.background = element_rect(fill = "transparent"),
+    legend.position = c(0.725, .85),
+    legend.background = element_rect(fill = "transparent", color = "black"),
     legend.text = element_text(size = 8)
@@
   mutate(delta_yield = yield / yield[prop_mpa == 0] - 1) %>%
+  ungroup() %>% 
+  mutate(fleet= fct_relabel(fleet, titler)) %>% 
   ggplot(aes(prop_mpa, delta_yield, color = fleet)) +
@@
   geom_line() +
-  facet_wrap( ~ placement_strategy) +
+  facet_wrap( ~ placement_strategy, labeller = labeller(placement_strategy = titler)) +
   scale_x_continuous(name = "MPA Size",
@@
     scale_color_discrete(name = '') + 
-    labs(tag = "B")
+    labs(tag = "A")
 
@@
   mutate(delta_bio = biodiv) %>%
+  mutate(critter = fct_relabel(critter, titler)) %>% 
   # filter(prop_mpa < 0.4) %>% 
@@
   scale_y_continuous(name = "Depletion") + 
-  scale_color_viridis_d(name = '', option = "C") + 
-    labs(tag = "C") + 
-  theme(axis.text = element_text(size = 8))
-
-((coral_frontier + plot_layout(guides = "keep")) | (coral_fleet_yield / coral_critter_bio) + plot_layout(guides = "collect")) & theme(strip.text = element_text(size  = 8), plot.margin = margin(.1, .1, .1, .1, "cm"), legend.text = element_text(size = 7))
+  scale_color_brewer(name = '', type = "qual", palette = "Dark2") + 
+    labs(tag = "B") + 
+  theme(axis.text = element_text(size = 8), axis.text.x = element_text(size = 8))
+
+((coral_fleet_yield / coral_critter_bio) | (coral_frontier + plot_layout(guides = "keep")) ) & theme(strip.text = element_text(size  = 8), plot.margin = margin(.1, .1, .1, .1, "cm"), panel.spacing = unit(1, "lines"), legend.text = element_text(size = 7))
 
@@
 sensitive to climate impacts for the purse seine fishery. Under the
-*status quo* species distribution, the *target_fishing* MPA design
+*status quo* species distribution, the *Target Fishing* MPA design
 strategy produced increases in fisheries yields at sizes between 25-50%
@@
 outcomes depending on the climate scenario. For bigeye tuna, under the
-*status quo* conditions the *target_fishing* MPA design strategy
+*status quo* conditions the *Target Fishing* MPA design strategy
 resulted in rapid increases in the population, but that same MPA network produced much smaller and slower gains under climate change.
@@
   geom_line(size = linesize, show.legend = FALSE) +
-  facet_grid(fleet ~ placement_strategy, scales = "free_y") +
+  facet_grid(fleet ~ placement_strategy, scales = "free_y", labeller = labeller(fleet = titler, placement_strategy = titler)) +
   scale_x_continuous(name = "MPA Size",
@@
   mutate(delta_bio = biodiv) %>%
+  ungroup() %>% 
+  mutate(placement_strategy = fct_relabel(as.factor(placement_strategy), titler)) %>% 
   ggplot(aes(prop_mpa, pmin(1,delta_bio), color = climate, linetype = placement_strategy)) +
@@
                      labels = scales::label_percent(accuracy = 1), guide = guide_axis(n.dodge = 2)) +
-  scale_y_continuous(name = "Change in Spawning Biomass",limits = c(0, NA)) +
+  scale_y_continuous(name = "Depletion",limits = c(0, NA)) +
   scale_linetype(name = "MPA Strategy") + 
@@
 (blue_fleet_yield / 
-blue_critter_bio) + plot_layout(heights = c(1.5,2), guides = "collect" )
+blue_critter_bio) + plot_layout(heights = c(1.5,2))
 ```
@@
 
+# Data Availability Statement
+
+All code and materials needed to reproduce this paper is available at (https://github.com/DanOvando/marlin-paper)[https://github.com/DanOvando/marlin-paper].
+
 # References
@@
 # Supporting Information
-
-# Scraps

### 2023-01-09 README.Rmd
circilng towards something
prepping repo for submission

@@
 
+## Navigating Repo
+
+THe main scripts needed to reproduce results are
+
+1. 00_setup.R
+2. 01_coral_case_study.R
+3. 02_blue_water_cast_study.R
+4. 03_marlin_paper.qmd
+
+The best way to reproduce the results is through the `make` commands described below.
+To reproduce results manually, first open 00_setup.R and set parameters as desired.
+Then, run 01_..
+02_...
+and 03_.
+
 ## Reproducing results on UNIX machines
@@
 
-3. Make sure you have all dependencies installed by opening R and running `renv::restor()`
+3. Make sure you have all dependencies installed by opening R and running `renv::restore()`
 
-4. In the terminal, run `RUN_NAME:MyRunName FORMAT:MyFormat make` where MyRunName is the name of the folder you want to create to store results (e.g. v0.1) and FORMAT is one of pdf, html, docx
+4. In the terminal, run `RUN_NAME:MyRunName FORMAT:MyFormat make` where MyRunName is the name of the folder you want to create to store results (e.g. v0.1) and FORMAT is one of pdf, html, or docx
 
-5. Wait for a bit! This command will run all the required scripts and knit the paper.
+5. Wait for a bit! This command will run all the required scripts and knit the paper. Should take no more than a few hours.

### 2023-01-09 R/compile_fleet.R

@@ -1,224 +1,0 @@
-compile_fleet <- function(fauna, tune_type = "explt",  effort_cost_exponent = 1,
-                          spatial_allocation = "ppue", fleet_model = "open access") {
-
-
-  fleets <- list(
-    "longline" = create_fleet(
-      list(
-        "thunnus obesus" = Metier$new(
-          critter = fauna$`thunnus obesus`,
-          price = 10,
-          sel_form = "logistic",
-          sel_start = 100,
-          sel_delta = .01,
-          catchability = .1,
-          p_explt = .5,
-          sel_unit = "length"
-        ),
-        "thunnus alalunga" = Metier$new(
-          critter = fauna$`thunnus alalunga`,
-          price = 4.57,
-          sel_form = "logistic",
-          sel_start = 50,
-          sel_delta = .01,
-          catchability = .1,
-          p_explt = .98,
-          sel_unit = "length"
-        ),
-        "katsuwonus pelamis" = Metier$new(
-          critter = fauna$`katsuwonus pelamis`,
-          price = 1.36,
-          sel_form = "logistic",
-          sel_start = 65,
-          sel_delta = 0.01,
-          catchability = .1,
-          p_explt = .01,
-          sel_unit = "length"
-
-        ),
-        "thunnus albacares" = Metier$new(
-          critter = fauna$`thunnus albacares`,
-          price = 7.52,
-          sel_form = "logistic",
-          sel_start = 90,
-          sel_delta = 0.01,
-          catchability = .1,
-          p_explt = 0.28,
-          sel_unit = "length"
-
-        ),
-        "carcharhinus longimanus" = Metier$new(
-          critter = fauna$`carcharhinus longimanus`,
-          price = 1.89,
-          sel_form = "logistic",
-          sel_start = 50,
-          sel_delta = 0.01,
-          catchability = .1,
-          p_explt = .98,
-          sel_unit = "length"
-
-        ),
-        "carcharhinus falciformis" = Metier$new(
-          critter = fauna$`carcharhinus falciformis`,
-          price = 2.16,
-          sel_form = "logistic",
-          sel_start = 110,
-          sel_delta = 0.01,
-          catchability = .1,
-          p_explt = .79,
-          sel_unit = "length"
-
-        ),
-        "prionace glauca" = Metier$new(
-          critter = fauna$`prionace glauca`,
-          price = 10,
-          sel_form = "logistic",
-          sel_start = 100,
[…]

### 2023-01-09 R/compile_fleet_highf.R

@@ -1,222 +1,0 @@
-compile_fleet_highf <- function(fauna, tune_type = "explt",  effort_cost_exponent = 1.3,
-                          spatial_allocation = "rpue") {
-
-
-  fleets <- list(
-    "longline" = create_fleet(
-      list(
-        "thunnus obesus" = Metier$new(
-          critter = fauna$`thunnus obesus`,
-          price = 10,
-          sel_form = "logistic",
-          sel_start = 1,
-          sel_delta = .01,
-          catchability = .1,
-          p_explt = .5,
-          sel_unit = "length"
-        ),
-        "thunnus alalunga" = Metier$new(
-          critter = fauna$`thunnus alalunga`,
-          price = 4.57,
-          sel_form = "logistic",
-          sel_start = 1,
-          sel_delta = .01,
-          catchability = .1,
-          p_explt = .98,
-          sel_unit = "length"
-        ),
-        "katsuwonus pelamis" = Metier$new(
-          critter = fauna$`katsuwonus pelamis`,
-          price = 1.36,
-          sel_form = "logistic",
-          sel_start = 1,
-          sel_delta = 0.01,
-          catchability = .1,
-          p_explt = .01,
-          sel_unit = "length"
-
-        ),
-        "thunnus albacares" = Metier$new(
-          critter = fauna$`thunnus albacares`,
-          price = 7.52,
-          sel_form = "logistic",
-          sel_start = 1,
-          sel_delta = 0.01,
-          catchability = .1,
-          p_explt = 0.28,
-          sel_unit = "length"
-
-        ),
-        "carcharhinus longimanus" = Metier$new(
-          critter = fauna$`carcharhinus longimanus`,
-          price = 1.89,
-          sel_form = "logistic",
-          sel_start = 175,
-          sel_delta = 0.01,
-          catchability = .1,
-          p_explt = .98,
-          sel_unit = "length"
-
-        ),
-        "carcharhinus falciformis" = Metier$new(
-          critter = fauna$`carcharhinus falciformis`,
-          price = 2.16,
-          sel_form = "logistic",
-          sel_start = 200,
-          sel_delta = 0.01,
-          catchability = .1,
-          p_explt = .79,
-          sel_unit = "length"
-
-        ),
-        "prionace glauca" = Metier$new(
-          critter = fauna$`prionace glauca`,
-          price = 10,
-          sel_form = "logistic",
-          sel_start = 200,
[…]

### 2023-01-09 R/optimize_mpa.R

@@ -1,303 +1,0 @@
-#' Optimize MPA network
-#'
-#' @param fauna fauna object
-#' @param fleets fleet object
-#' @param alpha the weight given to conservation relative to economics (range 0 to 1)
-#' @param max_prop_mpa max proportion of cells to place in MPA
-#' @param resolution the resolution of the simulated system
-#' @param prop_sampled the proportion of cells to sample in each SIR iteration
-#' @param workers number of workers for parallel process
-#' @param starting_conditions starting conditions of the simmar object
-#' @param objective one of max_ssb to maximize spawning stock biomass or min_loss to prioritize not having any losses
-#'
-#' @return a list with results of MPA optimization
-#' @export
-#'
-optimize_mpa <-
-  function(fauna,
-           fleets,
-           starting_conditions = NA,
-           alpha = 0.33,
-           max_prop_mpa = 1,
-           resolution,
-           prop_sampled = .2,
-           max_delta = 2,
-           workers = 6,
-           objective = "max_ssb") {
-    # workers <- 6
-    future::plan(future::multisession, workers = workers)
-
-    on.exit(future::plan(future::sequential))
-
-    # fauna <- casestudy$fauna[[1]]
-    #
-    # fleets <- casestudy$fleet[[1]]
-    #
-    # alpha <- 0
-    #
-    # percent_mpa <- 0.3
-    #
-    # prop_sampled <- 0.2
-    #
-    # max_prop_mpa <-  1
-
-    patches <- resolution ^ 2
-
-    samps <- round(prop_sampled * patches)
-
-    mpas <- expand_grid(x = 1:resolution, y = 1:resolution) %>%
-      mutate(patch = 1:nrow(.)) %>%
-      mutate(mpa = TRUE)
-
-    max_patches_protected <- round(patches * max_prop_mpa)
-
-    candidate_patches <- 1:patches
-
-    patch_value <- tibble(patch = 1:patches, obj_value = 0.5)
-
-    results <- vector(mode = "list", length = max_patches_protected)
-
-    mpa_network <-
-      vector(mode = "list", length = max_patches_protected)
-
-    # set up open access
-
-    calc_objective_function <-
-      function(candidate_patch, fauna, fleets, mpas,starting_conditions) {
-        tmp_mpas <- mpas
-
-        tmp_mpas$mpa[tmp_mpas$patch %in% candidate_patch] <- FALSE
-
[…]

### 2023-01-09 cpue_and_yield.R

@@ -1,202 +1,0 @@
-mpa_locations <- expand_grid(x = 1:resolution, y = 1:resolution) %>%
-  mutate(mpa = x < 0.3 * resolution)
-
-# fleets$longline$fleet_model
-#
-# fleets$purseseine$fleet_model
-
-with_mpa <- simmar(
-  fauna = fauna,
-  fleets = fleets,
-  manager = list(mpas = list(
-    locations = mpa_locations,
-    mpa_year = round(years / 2)
-  )),
-  years = years
-)
-
-grid <- expand_grid(x = 1:resolution, y= 1:resolution) %>%
-  mutate(patch = 1:nrow(.))
-
-
-patch_effort <-   map_df(with_mpa, ~ map_df(.x, ~.x$e_p_fl %>% mutate(patch = 1:nrow(.)), .id = "critter"), .id = "step") %>%
-  mutate(step = as.numeric(step)) %>%
-  left_join(grid, by = "patch") %>%
-  filter(critter ==  names(fauna)[1]) %>%
-  pivot_longer(names(fleets), names_to = "fleet", values_to = "effort")
-
-patch_effort %>%
-  filter(step == max(step)) %>%
-  ggplot(aes(x,y, fill = effort))+
-  geom_tile() +
-  facet_wrap(fleet~step) +
-  scale_fill_viridis_c()
-
-
-patch_catch <-
-  map_df(with_mpa,
-         ~ map_df(.x, ~ as_tibble(.x$c_p_fl) %>% mutate(patch = 1:nrow(.)), .id = "critter"),
-         .id = "step") %>%
-  mutate(step = as.numeric(step)) %>%
-  left_join(grid, by = "patch") %>%
-  pivot_longer(names(fleets), names_to = "fleet", values_to = "catch")
-
-
-
-patch_cpue <- patch_catch %>%
-  left_join(patch_effort %>% select(step, patch, starts_with("fleet"), effort)) %>%
-  mutate(cpue = catch / effort) %>%
-  mutate(mpa = TRUE)
-
-patch_cpue %>%
-  group_by(critter, step, fleet) %>%
-  mutate(scpue = cpue / max(cpue, na.rm = TRUE)) %>%
-  ungroup() %>%
-  filter(step == max(step)) %>%
-  ggplot(aes(x,y, fill = scpue))+
-  geom_tile() +
-  facet_grid(critter~fleet) +
-  scale_fill_viridis_c()
-
[…]

### 2023-01-09 functions/capitalize.R

@@ -1,0 +1,4 @@
+capitalize <- function(string) {
+  substr(string, 1, 1) <- toupper(substr(string, 1, 1))
+  string
+}

### 2023-01-09 functions/compile_fleet.R

@@ -1,0 +1,224 @@
+compile_fleet <- function(fauna, tune_type = "explt",  effort_cost_exponent = 1,
+                          spatial_allocation = "ppue", fleet_model = "open access") {
+
+
+  fleets <- list(
+    "longline" = create_fleet(
+      list(
+        "thunnus obesus" = Metier$new(
+          critter = fauna$`thunnus obesus`,
+          price = 10,
+          sel_form = "logistic",
+          sel_start = 100,
+          sel_delta = .01,
+          catchability = .1,
+          p_explt = .5,
+          sel_unit = "length"
+        ),
+        "thunnus alalunga" = Metier$new(
+          critter = fauna$`thunnus alalunga`,
+          price = 4.57,
+          sel_form = "logistic",
+          sel_start = 50,
+          sel_delta = .01,
+          catchability = .1,
+          p_explt = .98,
+          sel_unit = "length"
+        ),
+        "katsuwonus pelamis" = Metier$new(
+          critter = fauna$`katsuwonus pelamis`,
+          price = 1.36,
+          sel_form = "logistic",
+          sel_start = 65,
+          sel_delta = 0.01,
+          catchability = .1,
+          p_explt = .01,
+          sel_unit = "length"
+
+        ),
+        "thunnus albacares" = Metier$new(
+          critter = fauna$`thunnus albacares`,
+          price = 7.52,
+          sel_form = "logistic",
+          sel_start = 90,
+          sel_delta = 0.01,
+          catchability = .1,
+          p_explt = 0.28,
+          sel_unit = "length"
+
+        ),
+        "carcharhinus longimanus" = Metier$new(
+          critter = fauna$`carcharhinus longimanus`,
+          price = 1.89,
+          sel_form = "logistic",
+          sel_start = 50,
+          sel_delta = 0.01,
+          catchability = .1,
+          p_explt = .98,
+          sel_unit = "length"
+
+        ),
+        "carcharhinus falciformis" = Metier$new(
+          critter = fauna$`carcharhinus falciformis`,
+          price = 2.16,
+          sel_form = "logistic",
+          sel_start = 110,
+          sel_delta = 0.01,
+          catchability = .1,
+          p_explt = .79,
+          sel_unit = "length"
+
+        ),
+        "prionace glauca" = Metier$new(
+          critter = fauna$`prionace glauca`,
+          price = 10,
+          sel_form = "logistic",
+          sel_start = 100,
[…]

### 2023-01-09 functions/create_critters.R

@@ -1,0 +1,78 @@
+create_critters <-
+  function(sciname,
+           habitat,
+           seasons = 1,
+           marlin_inputs,
+           seasonal_movement = FALSE,
+           ontogenetic_shift = FALSE,
+           random_rec = FALSE,
+           adult_diffusion = 10,
+           taxis_to_diff_ratio = 2) {
+    # sciname <- marlin_inputs$scientific_name[[1]]
+    #
+    #
+
+    # habitat <- experiments$habitat[[1]]
+    tmp_inputs <- marlin_inputs %>%
+      filter(scientific_name == sciname)
+
+    hab <- habitat %>%
+      pivot_wider(names_from = y, values_from = habitat) %>%
+      select(-x) %>%
+      as.matrix()
+
+    # seasonal_movement <- sample(c(1, 0), 1, replace = TRUE)
+
+    # uniform_rec_hab <- sample(c(1, 0), 1, replace = TRUE)
+
+
+    # seasonal_movement <- 0
+    if (seasonal_movement) {
+      hab <- list(hab,-hab - (min(-hab)))
+
+    } else {
+      hab <- list(hab, hab)
+    }
+
+    if (random_rec){
+      density_dependence <-  sample(c("global_habitat", "local_habitat","pre_dispersal","post_dispersal","global_ssb"), 1, replace = TRUE)
+
+    } else {
+      density_dependence <- "global_ssb" # local density dependence then disperse recruits per recruit movement
+    }
+
+    if (ontogenetic_shift) {
+      recruit_habitat <-
+        -hab[[1]] - min(-hab[[1]]) # place recruits in different places than adults
+
+      # set recruitment form to allow for recruit habitat
+      #
+      density_dependence <- "global_ssb"
+    } else {
+      recruit_habitat <- hab[[1]]
+    }
+
+    critter <- marlin::create_critter(
+      scientific_name = sciname,
+      base_habitat = hab,
+      recruit_habitat = recruit_habitat,
+      adult_diffusion = adult_diffusion,
+      taxis_to_diff_ratio = taxis_to_diff_ratio,
+      recruit_diffusion = 0,
+      density_dependence = density_dependence,
+      fec_form = ifelse(str_detect(sciname,"carcharhinus"),"pups","weight"),
+      pups = 6,
+      seasons = seasons,
[…]

### 2023-01-09 functions/create_fauna.R

@@ -1,0 +1,25 @@
+create_fauna <- function(scientific_name,
+                         seasonal_habitat,
+                         season_blocks,
+                         adult_diffusion,
+                         rec_form,
+                         seasons,
+                         init_explt,
+                         explt_type = "f",
+                         recruit_habitat = NA){
+
+  tmp <-  create_critter(
+    scientific_name = scientific_name,
+    seasonal_habitat = seasonal_habitat,
+    season_blocks = season_blocks,
+    adult_diffusion = adult_diffusion,
+    rec_form = rec_form,
+    seasons = seasons,
+    init_explt = init_explt,
+    explt_type = "explt_type",
+    recruit_habitat = recruit_habitat
+  )
+
+
+
+}

### 2023-01-09 functions/run_mpa_experiment.R

@@ -1,0 +1,261 @@
+run_mpa_experiment <-
+  function(placement_strategy = "depletion",
+           fleet_model = "constant effort",
+           starting_conditions,
+           proc_starting_conditions,
+           prop_mpa = 0.3,
+           fauna,
+           fleets,
+           placement_error = 0,
+           mpa_response = "stay",
+           critters_considered = NA,
+           random_mpas = FALSE,
+           max_delta = 1,
+           resolution,
+           years = 50,
+           future_habitat = list(),
+           effort_cap) {
+
+    options(dplyr.summarise.inform = FALSE)
+
+
+    fleets <-  purrr::modify_in(fleets, list(1, "mpa_response"), ~ mpa_response)
+
+    if (!is.na(fleet_model)) {
+      fleets <-
+        purrr::modify_in(fleets, list(1, "fleet_model"), ~ fleet_model)
+
+
+    }
+
+    n_mpa <- round(prop_mpa * resolution^2)
+
+    # set up open access
+
+    #
+    if (all(is.na(critters_considered))){
+      critters_considered <-  length(fauna)
+    }
+
+    critters_considered <- sample(names(fauna), critters_considered, replace = FALSE)
+
+    if (placement_strategy == "depletion"){
+
+      # place MPAs in proportion to depletion-weighted spawning stock biomass
+      depletion <-
+        map_df(starting_conditions[1], ~ (1 - (map_df(.x, ~ sum(.x$ssb_p_a) / .x$ssb0))) %>%  # depletion of each species
+        pivot_longer(everything(), names_to = "critter", values_to = "weight"), .id = "step") %>%
+        select(-step)
+
+      priorities <- proc_starting_conditions$fauna %>%
+        left_join(depletion, by = "critter") %>%
+        filter(critter %in% critters_considered) %>%
+        group_by(critter, step) %>%
+        mutate(patch_weight = ssb / sum(ssb) * weight) %>%
+        group_by(patch, step) %>%
+        summarise(patch_weight = sum(patch_weight)) %>%
+        ungroup() %>%
+        mutate(patch_weight = patch_weight * rlnorm(nrow(.), 0, placement_error)) %>%
+        arrange(desc(patch_weight))
+
[…]

### 2023-01-09 scripts/00_setup.R

@@ -1,107 +1,0 @@
-set.seed(42)
-library(tidyverse)
-
-library(marlin)
-
-library(here)
-
-library(furrr)
-
-library(ggridges)
-
-library(googledrive)
-
-library(gamm4)
-
-library(rgdal)
-
-library(sf)
-
-library(mvtnorm)
-
-library(patchwork)
-
-library(progress)
-
-library(rpart.plot)
-
-library(rstanarm)
-
-library(ggrepel)
-
-library(ggdist)
-
-library(gganimate)
-
-library(gifski)
-
-options(dplyr.summarise.inform = FALSE)
-
-
-foos <- list.files(here("R"))
-
-walk(foos, ~ source(here("R", .x)))
-
-print(Sys.getenv("RUN_NAME"))
-if (Sys.getenv("RUN_NAME") == ''){
-  run_name <- "v0.2"
-
-} else {
-  run_name <- Sys.getenv("RUN_NAME")
-}
-
-
-results_path <- here("results", run_name)
-
-if (!dir.exists(results_path)){
-  dir.create(results_path, recursive = TRUE)
-
-  dir.create(file.path(results_path,"sims"))
-}
-
-run_coral_example <- TRUE
-
-run_blue_water_example <- TRUE
-
-experiment_workers <- 8
-
-experiment_years <- 20
-
-safety_stop <- FALSE
-
-draws <- 20
-
-resolution <- 20 # resolution is in squared patches, so 20 implies a 20X20 system, i.e. 400 patches
-
-years <- 80
-
-seasons <- 2
-
-time_step <- 1 / seasons
-
-max_delta <- 1
-
-workers <- round(parallel::detectCores() - 8)
-
-steps <- years * seasons
-
-time_steps <- seq(0,years - 1, by = time_step)
-
-theme_set(marlin::theme_marlin())
-
-
-# load data ---------------------------------------------------------------
-
-if (!file.exists(here("data","marlin-inputs.xlsx"))){
-
-  dir.create("data")
-
-  marlin_inputs_path <- googledrive::drive_get("https://docs.google.com/spreadsheets/d/1eiJIxiDYZLQlBuSLy-yIofhbc-KfWBR9v4FpThgXrQ0/edit#gid=0&fvid=1954704204")
-
-
-  googledrive::drive_download(marlin_inputs_path, path = here("data","marlin-inputs.xlsx"), overwrite = TRUE)
-
-}
-
[…]

### 2023-01-09 scripts/old_scripts/02_run_case_study.R

@@ -1,938 +1,0 @@
-source(file.path("scripts", "00_setup.R"))
-
-
-# load bycatch risk layers ------------------------------------------------
-
-
-mats <- list.files(path = here("data", "species-matrices"))
-
-mats <- mats[str_detect(mats, "cpue_mean")]
-
-file_comps <- str_split(mats, "_", simplify = TRUE)
-
-species <- str_replace(file_comps[, 1], "-", " ")
-
-# go through and figure out the best matrix for each species in a really hacky manner
-
-best_mat <- rep(NA, n_distinct(species))
-for (s in seq_along(unique(species))){
-
-  fls <- mats[species == unique(species)[s]]
-
-  if (all(str_detect(fls,"ps_cpue"))){
-    # if the only thing available is purse seine CPUE
-
-    fls <- fls[str_detect(fls,"ps_cpue")]
-
-
-  } else {
-
-    fls <- fls[!str_detect(fls,"ps_cpue")]
-
-  }
-
-  if (any(str_detect(fls,"mtper"))){
-
-    fls <- fls[str_detect(fls,"mtper")]
-
-
-  } else {
-
-    fls <- fls[str_detect(fls,"countper")]
-  }
-
-
-  # preference: longline weight cpue
-
-  # then longline number cpue
-
-  # then purse seine
-
-
-  best_mat[s] <- fls
-
-}
-
-get_layer <- function(file) {
-
-  file_comps <- str_split(file, "_", simplify = TRUE)
-
-  species <- str_replace(file_comps[, 1], "-", " ")
-
-
-  hab <-
-    read_csv(here("data", "species-matrices", file),
-             col_names = FALSE,
-             skip = 1) %>%
-    rename(lat = X1) %>%
-    mutate(lat = 1:nrow(.)) %>%
-    pivot_longer(
-      -lat,
-      names_to = "lon",
-      values_to = "hab",
-      names_prefix = "X"
-    ) %>%
-    mutate(hab = tidyr::replace_na(hab, 0),
-           lon = as.numeric(lon) - 1)
-
-  # hab %>%
-  #   ggplot(aes(lon, lat, fill = hab)) +
-  #   geom_tile()
-
-
-  x_binsize <- (max(hab$lon) - min(hab$lon) + 1) / resolution
-
-  y_binsize <- (max(hab$lat) - min(hab$lat) + 1) / resolution
-
-  if (x_binsize < 1 | y_binsize < 1) {
-    stop("resolution is too high for input data")
-  }
-
-
-  hab <- hab %>%
[…]

### 2023-01-10 03_marlin_paper.qmd
out for internal

@@
 params:
-  run_name: ["v0.5"]
+  run_name: ["v1.0"]
 abstract: "Abstract: Designing effective spatial management strategies is challenging because marine ecosystems are highly dynamic and opaque, and extractive entities such as fishing fleets respond endogenously to ecosystem changes in ways that depend upon ecological and policy context.
@@
 
+The front end of the software package accompanying this paper,
+[`marlin`](https://danovando.github.io/marlin/), is written in R
+[@rcoreteam2021] to facilitate use, while the underlying population model is written in C++, integrated through the `Rcpp` package
+[@eddelbuettel2018], for speed. Users on standard computers should be able to simulate one age-structured population distributed across a 100 patch system for 20 years in fractions of a second.
+
 While `marlin` does not simulate species interactions, fishing fleets in
@@
 in aquatic systems.
-
-While users can adapt the general equations and ideas presented here as
-they see fit, the software package accompanying this paper,
-[`marlin`](https://danovando.github.io/marlin/), is written in R
-[@rcoreteam2021] and C++, integrated through the `Rcpp` package
-[@eddelbuettel2018].
 
@@
 #| label: fig-coral-results
-#| fig-cap: "Percent change in yield per fleet (A) and depletion per species (B) as a function of MPA size and MPA design strategy for the coastal coral reef case study.
-Percent change in total yield across both fleets (y-axis) and total depletion across all species (x-axis) by MPA size (point size) and placement strategy (point color) (C)."
+#| fig-cap: "Percent change in yield per fleet (A) and depletion (spawning biomass divided by unfished spawning biomass) per species (B) as a function of MPA size and MPA design strategy for the coastal coral reef case study.
+Percent change in total yield across both fleets (y-axis) and total depletion across all species (x-axis) and placement strategies (line color) (C).
+Text labels in panel C indicate the percent of area in an MPA at selected points along curves."
 #| fig-height: 4
@@
 
-coral_frontier <- coral_results %>% 
+tmp <- coral_results %>% 
   group_by(prop_mpa, placement_strategy) %>% 
@@
   ungroup() %>% 
-    mutate(placement_strategy= fct_relabel(placement_strategy, titler)) %>% 
-  ggplot(aes(delta_biodiv, delta_yield, fill = placement_strategy, size = prop_mpa)) + 
+    mutate(placement_strategy= fct_relabel(placement_strategy, titler))
+
+frontier_labels <- tmp %>% 
+  mutate(delta_50 = (delta_yield - -0.5)^2,
+         delta_0 = delta_yield^2) %>% 
+  group_by(placement_strategy) %>% 
+  filter(prop_mpa > 0) %>% 
+  filter(yield == max(yield) | delta_0 == min(delta_0) | delta_50 == min(delta_50)) %>% 
+  ungroup() %>% 
+  mutate(pmpa = scales::percent(prop_mpa))
+
+coral_frontier <- tmp %>% 
+  ggplot(aes(delta_biodiv, delta_yield)) + 
   geom_hline(aes(yintercept = 0), linetype = 2) +
-  geom_point(alpha = 0.8, shape = 21) + 
+  geom_line(alpha = 0.8, aes(color = placement_strategy),size = 1.1) + 
+  geom_text_repel(data = frontier_labels, aes(label = pmpa), min.segment.length = 0, box.padding = 0.5) +
   scale_x_continuous(name = "Change in Total Depletion", labels = scales::label_percent(accuracy = 1), guide = guide_axis(n.dodge = 2)) +
   scale_y_continuous(name = "Change in Yield", labels = scales::label_percent(accuracy = 1)) + 
-  scale_fill_viridis_d(name = '')+
-  scale_size(name = "MPA Size", labels = scales::label_percent(accuracy = 1), breaks = seq(0.25,1, by = 0.25)) +
-  guides(fill = guide_legend(order = 1, override.aes = list(size = 4)),
-         size = guide_legend(order = 2, ncol = 2, byrow = TRUE)) + 
+  scale_color_manual(values = c("tomato", "steelblue"),name = '') +
+  guides(fill = guide_legend(order = 1, override.aes = list(size = 4))) + 
   labs(tag = "C")  + 
   theme(
-    legend.position = c(0.725, .85),
-    legend.background = element_rect(fill = "transparent", color = "black"),
+    legend.position = c(0.725, .95),
+    legend.background = element_rect(fill = "transparent"),
     legend.text = element_text(size = 8)
@@
 #| label: fig-blue-results
-#| fig-cap: "Change in yield (A) and spawning biomass (B) as a function of MPA size and design strategy by fishing fleet and species.
+#| fig-cap: "Change in yield (A) and depletion (spawning biomass divided by unfished spawning biomass) (B) as a function of MPA size and design strategy by fishing fleet and species.
 Blue lines indicate impacts of MPAs under status qup habitat, red impacts under climate-driven range shift."

### 2023-02-16 03_marlin_paper.qmd
prepping for edits on squid

@@
 
+resolution <- sqrt((coral_fauna[[1]]$patches))
+
 grid <- expand_grid(x = 1:resolution, y= 1:resolution) %>%
@@
 centroid of each population shifts northward at a rate of 0.2 patches
-per year, over a `r experiment_years` year time horizon (@fig-blue-ssb).
+per year, over a 20 year time horizon (@fig-blue-ssb).
 We designed MPA networks given the conditions in the starting year, and

### 2023-02-16 00_setup.R

@@ -1,5 +1,19 @@
 set.seed(42)
-library(tidyverse)
+library(purrr)
+
+library(dplyr)
+
+library(tidyr)
+
+library(readr)
+
+library(forcats)
+
+library(stringr)
+
+library(ggplot2)
+
+library(viridis)
 
 library(marlin)
 
@@ -7,33 +21,17 @@ library(here)
 
 library(furrr)
 
-library(ggridges)
-
-library(googledrive)
-
 library(gamm4)
 
-library(rgdal)
-
-library(sf)
-
-library(mvtnorm)
-
 library(patchwork)
-
-library(progress)
-
-library(rpart.plot)
-
-library(rstanarm)
 
 library(ggrepel)
 
 library(ggdist)
 
-library(gganimate)
+library(rmarkdown)
 
-library(gifski)
+library(quarto)
 
 options(dplyr.summarise.inform = FALSE)
 
@@ -62,30 +60,6 @@ if (!dir.exists(results_path)){
 run_coral_example <- TRUE
 
 run_blue_water_example <- TRUE
-
-experiment_workers <- 8
-
-experiment_years <- 20
-
-safety_stop <- FALSE
-
-draws <- 20
-
-resolution <- 20 # resolution is in squared patches, so 20 implies a 20X20 system, i.e. 400 patches
-
-years <- 80
-
-seasons <- 2
-
-time_step <- 1 / seasons
-
-max_delta <- 1
-
-workers <- round(parallel::detectCores() - 8)
-?unzip
-steps <- years * seasons
-
-time_steps <- seq(0,years - 1, by = time_step)
 
 theme_set(marlin::theme_marlin())
 

### 2023-02-16 01_coral_case_study.R

@@ -1,6 +1,5 @@
 source(file.path("00_setup.R"))
 
-library(gganimate)
 set.seed(42)
 
 resolution <- 20
@@ -544,15 +543,15 @@ coral_mpa_experiments$obj <- map(coral_mpa_experiments$results, "obj")
 examine_mpas <- coral_mpa_experiments %>%
   unnest(cols = mpas)
 
-mpas <- examine_mpas %>%
-  filter(mpa_response == "stay") %>%
-  ggplot(aes(x,y,fill = mpa)) +
-  geom_tile() +
-  transition_time(prop_mpa) +
-  scale_x_continuous(name = "longitude") +
-  scale_y_continuous(name = "latitude") +
-  labs(title = 'Step: {frame_time}') +
-  facet_wrap(~placement_strategy)
+# mpas <- examine_mpas %>%
+#   filter(mpa_response == "stay") %>%
+#   ggplot(aes(x,y,fill = mpa)) +
+#   geom_tile() +
+#   transition_time(prop_mpa) +
+#   scale_x_continuous(name = "longitude") +
+#   scale_y_continuous(name = "latitude") +
+#   labs(title = 'Step: {frame_time}') +
+#   facet_wrap(~placement_strategy)
 
 
 examine_results <- coral_mpa_experiments %>%

### 2023-02-23 03_marlin_paper.qmd
off to abalone

@@
 Abstract: Designing effective spatial management strategies is challenging because marine ecosystems are highly dynamic and opaque, and extractive entities such as fishing fleets respond endogenously to ecosystem changes in ways that depend upon ecological and policy context.
-We present a tool – marlin -- to efficiently simulate the bio-economic dynamics of marine systems, designed to support both management and research.
+We present a tool -- marlin -- to efficiently simulate the bio-economic dynamics of marine systems, designed to support both management and research.
 We demonstrate `marlin’s` capabilities by focusing on two case studies on the conservation and food production impacts of marine protected areas (MPAs): a coastal coral reef and a pelagic tuna fishery.
@@
 
-Communities around the world are increasingly looking to spatial
-strategies for managing marine ecosystems . For example, the "30x30"
-movement calls for implementation of marine protected areas (MPAs), a
-form of spatial management, across 30% of the world's oceans by the year
-2030 [@grorud-colvert2021 and references therein]. However , designing
-spatial management strategies to achieve desired objectives -- which may
-include recovery of overfished species, increased food production and
-economic well being, and equitable distribution of benefits - is not
-straightforward. Different species have different resilience to fishing,
-distributions in space and time, and value to fishing fleets; fishing
-fleets themselves can have varying ranges of incentives and reactions to
-policy structures. These dynamics pose challenges even for
-single-species assessment and management strategies [@field2006a], which
-are only amplified when we consider management policies specifically
-designed around spatial features in the oceans.
-
-For example, policies such as MPAs designed for both conservation and
-food production must consider factors such as the optimal size and
-placement of a protected area network given a wide array of life
-history, species distributions, exploitation levels, fleet dynamics, and
-policy-dependent behavior [@reimer2017], all of which may change over
-time, particularly given the impacts of climate change. Efforts to
-effectively design spatial management strategies such as MPAs are
-further constrained by a lack of empirical evidence describing the size
-[@ovando2021a] and time required [@nickols2019] for MPAs to produce
-substantial effects across a range of social-ecological settings
-[@mcclanahan2021].
-
-Many models have been developed to support the theory and design of
-spatial management strategies (see @fulton2015 for a review of existing
-methods). However, they have tended to fall into one of either highly
-complex and tactical models designed for use in a specific location, or
-to be highly stylized representations intended to provide theoretical
-insights with less relevance for specific applications. This lack of
-accessible models suited for more tactical application presents a
-challenge to stakeholders charged with marine spatial planning
-exercises. It also presents a barrier to the scientific community,
-wherein comparisons across spatial management simulation studies are
-clouded by discrepancies in underlying model structures beyond
-differences in the core phenomenon in question.
-
-To help address this challenge, we present a novel bio-economic modeling
-tool called [`marlin`](https://danovando.github.io/marlin/) that allows
-for efficient simulation of spatio-temporal biological and economic
-dynamics. `marlin` allows users to simulate the impacts of marine
-management policies across a range of species targeted by various fleets
-across heterogeneous and dynamic seascapes. This model can help users
-seek Parteo-optimal solutions that produce win-wins or minimize
-tradeoffs across multiple management objectives, such as food production
-and conservation [@lester2013].
-
-Here we present the core methods and functionality of `marlin`, and
-demonstrate its use in two case-study applications; a coral reef system
-and a pelagic system. Our results show how interacting ecological,
-economic, and design attributes can affect the degree of benefits,
-costs, and tradeoffs between conservation and food production outcomes
-of MPAs. More broadly, this paper both demonstrates the functionality of
-our model and the critical need for considering sufficiently realistic
-coupled ecological and economic dynamics in policy evaluations.
+Communities around the world are increasingly looking to spatial strategies for managing marine ecosystems .
+For example, the "30x30" movement calls for implementation of marine protected areas (MPAs), a form of spatial management, across 30% of the world's oceans by the year 2030 [@grorud-colvert2021 and references therein].
+However , designing spatial management strategies to achieve desired objectives -- which may include recovery of overfished species, increased food production and economic well being, and equitable distribution of benefits - is not straightforward.
+Different species have different resilience to fishing, distributions in space and time, and value to fishing fleets; fishing fleets themselves can have varying ranges of incentives and reactions to policy structures.
+These dynamics pose challenges even for single-species assessment and management strategies [@field2006a], which are only amplified when we consider management policies specifically designed around spatial features in the oceans.
+
+For example, policies such as MPAs designed for both conservation and food production must consider factors such as the optimal size and placement of a protected area network given a wide array of life history, species distributions, exploitation levels, fleet dynamics, and policy-dependent behavior [@reimer2017], all of which may change over time, particularly given the impacts of climate change.
+Efforts to effectively design spatial management strategies such as MPAs are further constrained by a lack of empirical evidence describing the size [@ovando2021a] and time required [@nickols2019] for MPAs to produce substantial effects across a range of social-ecological settings [@mcclanahan2021].
+
+Many models have been developed to support the theory and design of spatial management strategies (see @fulton2015 for a review of existing methods).
+However, they have tended to fall into one of either highly complex and tactical models designed for use in a specific location, or to be highly stylized representations intended to provide theoretical insights with less relevance for specific applications.
+This lack of accessible models suited for more tactical application presents a challenge to stakeholders charged with marine spatial planning exercises.
+It also presents a barrier to the scientific community, wherein comparisons across spatial management simulation studies are clouded by discrepancies in underlying model structures beyond differences in the core phenomenon in question.
+
+To help address this challenge, we present a novel bio-economic modeling tool called [`marlin`](https://danovando.github.io/marlin/) that allows for efficient simulation of spatio-temporal biological and economic dynamics.
+`marlin` allows users to simulate the impacts of marine management policies across a range of species targeted by various fleets across heterogeneous and dynamic seascapes.
+This model can help users seek Parteo-optimal solutions that produce win-wins or minimize tradeoffs across multiple management objectives, such as food production and conservation [@lester2013].
+
+Here we present the core methods and functionality of `marlin`, and demonstrate its use in two case-study applications; a coral reef system and a pelagic system.
+Our results show how interacting ecological, economic, and design attributes can affect the degree of benefits, costs, and tradeoffs between conservation and food production outcomes of MPAs.
+More broadly, this paper both demonstrates the functionality of our model and the critical need for considering sufficiently realistic coupled ecological and economic dynamics in policy evaluations.
 
@@
 
-`marlin` follows a "models of intermediate complexity" philosophy
-[@plagányi2014] targeted specifically towards capturing heterogeneous
-habitats, movement dynamics, and the simultaneous impacts of fishing
-fleets across multiple species (frequently called "technical
-interactions"). In its general form, `marlin` simulates the behavior of
-populations of spatially explicit biologically independent
-age-structured animal populations affected by one or more fishing fleets
-in time steps and a spatial resolution specified by the user.
-
-The front end of the software package accompanying this paper,
-[`marlin`](https://danovando.github.io/marlin/), is written in R
-[@rcoreteam2021] to facilitate use, while the underlying population model is written in C++, integrated through the `Rcpp` package
-[@eddelbuettel2018], for speed. Users on standard computers should be able to simulate one age-structured population distributed across a 100 patch system for 20 years in fractions of a second.
-
-While `marlin` does not simulate species interactions, fishing fleets in
-`marlin` are capable of targeting and affecting multiple species
-simultaneously. This allows the model to simulate processes such as
-fisheries bycatch or effort displacement in a way that accounts for how
-fleet behavior might affect multiple species in the system. The ability
-to track the impacts of fishing fleets across multiple species
-simultaneously is particularly important as very few fishing fleets are
-truly single species.
+`marlin` follows a "models of intermediate complexity" philosophy [@plagányi2014] targeted specifically towards capturing heterogeneous habitats, movement dynamics, and the simultaneous impacts of fishing fleets across multiple species (frequently called "technical interactions").
+In its general form, `marlin` simulates the behavior of populations of spatially explicit biologically independent age-structured animal populations affected by one or more fishing fleets in time steps and a spatial resolution specified by the user.
+
+The front end of the software package accompanying this paper, [`marlin`](https://danovando.github.io/marlin/), is written in R [@rcoreteam2021] to facilitate use, while the underlying population model is written in C++, integrated through the `Rcpp` package [@eddelbuettel2018], for speed.
+Users on standard computers should be able to simulate one age-structured population distributed across a 100 patch system for 20 years in fractions of a second.
+
+While `marlin` does not simulate species interactions, fishing fleets in `marlin` are capable of targeting and affecting multiple species simultaneously.
+This allows the model to simulate processes such as fisheries bycatch or effort displacement in a way that accounts for how fleet behavior might affect multiple species in the system.
+The ability to track the impacts of fishing fleets across multiple species simultaneously is particularly important as very few fishing fleets are truly single species.
 
@@
 *Spawning stock biomass* refers to the total biomass of reproductively mature fish in the water, a function of numbers, weight, and sexual maturity at age.
-*Depletion* is a common term in the fisheries literature, and reflects here the *spawning stock biomass* in a given time step relative to *unfished spawning stock biomass*.
-So, a depletion value of 1 means that the population is unfished, a depletion value of 0 means the population is extinct.
-
-As a demonstration of the use of this modeling tool, we explore two
-general situations
-
-1.  Trade-offs and distributional outcomes for food production and
-    multi-species biomass in a coastal coral-reef fishery
-2.  Implications of climate-driven range shifts for MPA design in a
-    pelagic fishery
-
-We chose these two examples to illustrate the use of `marlin` in
-contrasting systems in which spatial management strategies such as MPAs are increasingly considered.
-
-Below we provide a summary of the `marlin` model, as well as details of
-the case studies. We note
-that `marlin` is a complex software package with many options to control
-detailed behaviors. Given that these specific features will evolve over
-time, we do not describe in detail every option of the model here.
-Rather, we focus on the equations and methods describing the core
-foundations of our tool for efficiently simulating bio-economic dynamics
-in aquatic systems.
+We measure the size of the population by th ratio of the *spawning stock biomass* (SSB) in a given time step relative to *unfished spawning stock biomass* (SSB0), SSB/SSB0.
+So, a SSB/SSB0 value of 1 means that the population is unfished, a SSB/SSB0 value of 0 means the population is extinct.
+
+As a demonstration of the use of this modeling tool, we explore two general situations
+
+1.  Trade-offs and distributional outcomes for food production and multi-species biomass in a coastal coral-reef fishery
+2.  Implications of climate-driven range shifts for MPA design in a pelagic fishery
+
+We chose these two examples to illustrate the use of `marlin` in contrasting systems in which spatial management strategies such as MPAs are increasingly considered.
+
+Below we provide a summary of the `marlin` model, as well as details of the case studies.
+We note that `marlin` is a complex software package with many options to control detailed behaviors.
+Given that these specific features will evolve over time, we do not describe in detail every option of the model here.
+Rather, we focus on the equations and methods describing the core foundations of our tool for efficiently simulating bio-economic dynamics in aquatic systems.
 
@@
 
-`marlin` simulates the dynamics of one or more species, currently best
-representing fish-like species, using age-structured population
-dynamics. Each time step, species move throughout the simulated seascape
-using both diffusion and "taxis" (active movement towards preferred
-habitat), experience natural and potentially fishing mortality, and if
-in a spawning season spawn using one of the possible forms of density
-dependence implemented in the model, with the potential to include
-auto-correlated stochastic deviations in the amount of offspring
-produced, generally called "recruitment deviates".
-
-These species can be caught by fishing fleets. A fishing fleet is defined
-in `marlin` by a set of fishers that can have different fishing skill,
-prices, and selectivity for individual species (each of which we denote
-as a *metier*). For example, both a longline and purse seine fleet may
-capture bigeye (*Thunnus obesus*) and skipjack tunas, but the longline
-fleet can be made much more likely to catch bigeye than skipjack, and
-*vice versa*. Each fishing fleet distributes its fishing effort in space
-according to a specified spatial allocation rule, for example in
-proportion to profit-per-unit-effort, conditional on management policies in places such as quotas and/or the presence of any spatial restrictions such as MPAs.
-
-Unconstrained by management, the total amount of effort exerted by each
-fleet can follow one of two dynamics: open access or constant effort.
-Under constant effort, the total amount of effort of each fleet is fixed
-over time, with the possible exception of attrition due to MPA
-placement. Under open access, the total amount of effort in a given time
-step is a function of the profitability of the fleet in the previous
-time step, evolving until a bionomic equilibrium of zero total profits
-is reached. Profitability is a function of the volume and price of each
-species caught, as well as the cost of the total amount of fishing
-effort per fleet and the travel costs as a function of distance from a port.
-
-The dynamics of each fleet can be modified by management in a variety of
-ways. Managers can impose size limits for individual species within each
-fleet. They can also impose total catch quotas for each species in each
-time step. When quotas are activated, if the total catch across all
-fleets for a given species under the model's effort dynamics would
-exceed the allowable quota for that species, the total effort for each
-fleet is decreased proportional to its contribution to the total catch
-until the quota is satisfied. As an alternative to catch quotas,
-managers can set an effort cap per fleet, which prevents effort from
-exceeding a given amount under open access dynamics, though effort may
-be reduced below this cap if required by the profit equation (i.e. the
-fleet can choose to fish less than the quota or effort cap).
-
-`marlin` also allows users to specify closed fishing seasons for one or more
-species in the system.
+`marlin` simulates the dynamics of one or more species, currently best representing fish-like species, using age-structured population dynamics.
+Each time step, species move throughout the simulated seascape using both diffusion and "taxis" (active movement towards preferred habitat), experience natural and potentially fishing mortality, and if in a spawning season spawn using one of the possible forms of density dependence implemented in the model, with the potential to include auto-correlated stochastic deviations in the amount of offspring produced, generally called "recruitment deviates".
+
+These species can be caught by fishing fleets.
+A fishing fleet is defined in `marlin` by a set of fishers that can have different fishing skill, prices, and selectivity for individual species (each of which we denote as a *metier*).
+For example, both a longline and purse seine fleet may capture bigeye (*Thunnus obesus*) and skipjack tunas, but the longline fleet can be made much more likely to catch bigeye than skipjack, and *vice versa*.
+Each fishing fleet distributes its fishing effort in space according to a specified spatial allocation rule, for example in proportion to profit-per-unit-effort, conditional on management policies in places such as quotas and/or the presence of any spatial restrictions such as MPAs.
+
+Unconstrained by management, the total amount of effort exerted by each fleet can follow one of two dynamics: open access or constant effort.
+Under constant effort, the total amount of effort of each fleet is fixed over time, with the possible exception of attrition due to MPA placement.
+Under open access, the total amount of effort in a given time step is a function of the profitability of the fleet in the previous time step, evolving until a bionomic equilibrium of zero total profits is reached.
+Profitability is a function of the volume and price of each species caught, as well as the cost of the total amount of fishing effort per fleet and the travel costs as a function of distance from a port.
+
+The dynamics of each fleet can be modified by management in a variety of ways.
+Managers can impose size limits for individual species within each fleet.
+They can also impose total catch quotas for each species in each time step.
+When quotas are activated, if the total catch across all fleets for a given species under the model's effort dynamics would exceed the allowable quota for that species, the total effort for each fleet is decreased proportional to its contribution to the total catch until the quota is satisfied.
+As an alternative to catch quotas, managers can set an effort cap per fleet, which prevents effort from exceeding a given amount under open access dynamics, though effort may be reduced below this cap if required by the profit equation (i.e.
+the fleet can choose to fish less than the quota or effort cap).
+
+`marlin` also allows users to specify closed fishing seasons for one or more species in the system.
 When all species in a simulation share the same closed seasons, no fishing occurs in those seasons.
@@
 When closed seasons are in effect, the amount of effort in the next open season is a function of the effort and profits in the last open fishing season.
-Lastly,
-managers can specify locations of no-take MPAs, which could in theory
-change in size and location. When MPAs are implemented, fishing effort
-that used to operate inside the MPAs can either leave the fishery, or be
-redistributed outside the MPA (the default behavior).
-
-Users can use the model to explore general properties of spatially
-explicit social-ecological systems, or can parameterize it to explore
-key uncertainties of more tactical applications. As an example of the
-later case, we provide two case studies here designed to reflect the
-kinds of questions stakeholders might face if designing an MPA network
-for a coastal coral reef or a pelagic system.
+Lastly, managers can specify locations of no-take MPAs, which could in theory change in size and location.
+When MPAs are implemented, fishing effort that used to operate inside the MPAs can either leave the fishery, or be redistributed outside the MPA (the default behavior).
+
+Users can use the model to explore general properties of spatially explicit social-ecological systems, or can parameterize it to explore key uncertainties of more tactical applications.
+As an example of the later case, we provide two case studies here designed to reflect the kinds of questions stakeholders might face if designing an MPA network for a coastal coral reef or a pelagic system.
 
@@
 
-Each fishing fleet (*f*) generates catches or yield (*Y*), revenues
-(*R*), and costs (*C*) from fishing individual species (*s*) that it targets in a
-given time step (*t*) and patch (*p*). The fleet then makes decisions
-around fishing locations and intensity, subject to regulatory constraints,
-based on the total profitability across all species in space and time.
-
-Revenues for each fleet in a time step are a function of the total
-amount of each species caught and its price ($\Pi$). The amount
-caught is a function of the selectivity at age of each species for each
-fleet ($\alpha$), the fishing efficiency of the fleet for that that
-species (also called "catchability", *q*), the amount of fishing effort
-of the fleet in question in that patch in a given time step (*E*), and
-the total fishing mortality at age (including all other fleets) for that
-particular species in that patch and time (*u*).
+Each fishing fleet (*f*) generates catches or yield (*Y*), revenues (*R*), and costs (*C*) from fishing individual species (*s*) that it targets in a given time step (*t*) and patch (*p*).
+The fleet then makes decisions around fishing locations and intensity, subject to regulatory constraints, based on the total profitability across all species in space and time.
+
+Revenues for each fleet in a time step are a function of the total amount of each species caught and its price ($\Pi$).
+The amount caught is a function of the selectivity at age of each species for each fleet ($\alpha$), the fishing efficiency of the fleet for that that species (also called "catchability", *q*), the amount of fishing effort of the fleet in question in that patch in a given time step (*E*), and the total instantaneous fishing mortality (*u*) at age (*a*) (including all other fleets) for that particular species in that patch and time step.
 
@@
 
-Total catches or yield *Y* are calculated through the Baranov equation
-[@baranov1918], which appropriately accounts for the total instantaneous
-mortality (both fishing and natural) and divides the total amount of the
-population biomass (*b*) killed between natural (*m*) and fishing (*u*)
-sources, with the amount of biomass killed through fishing assigned to
-yield *Y*.
+Total catches or yield *Y* are calculated through the Baranov equation [@baranov1918], which appropriately accounts for the total instantaneous mortality (both fishing and natural.
+*z*) and divides the total amount of the population biomass (*b*) killed between natural (*m*) and fishing (*u*) sources, with the amount of biomass killed through fishing called "yield".
 
@@
 
-Selectivity at age $\alpha$ is modeled through either a logistic form or
-a dome-shaped form. The logistic form is based on the lengths *l* at
-which 50% of individuals are selected by the fishing gear, $l^{sel}$,
-and $\delta$ which is the difference between the length at 50%
-selectivity and the length at 95% selectivity.
+Selectivity at age $\alpha$ is modeled through either a logistic form or a dome-shaped form.
+The logistic form is based on the lengths *l* at which 50% of individuals are selected by the fishing gear, $l^{sel}$, and $\delta$ which is the difference between the length at 50% selectivity and the length at 95% selectivity.
 
@@
 
-We approximate the dome-shaped form as a normal distribution with
-mean $l^{sel}$ and standard deviation $\sigma$. The normal density
-function is re-scaled such that the selectivity is 1 at $l^{sel}$.
+We approximate the dome-shaped form as a normal distribution with mean $l^{sel}$ and standard deviation $\sigma$. The normal density function is re-scaled such that the selectivity is 1 at $l^{sel}$.
 
@@
 
-Lastly *z* is total mortality, the sum of fishing mortality and natural
-mortality (*m*).
+Lastly *z* is total mortality, the sum of fishing mortality and natural mortality (*m*).
 
@@
 
-This means that each species experiences a total mortality at age in a
-time step in a patch, individual fractions of which are portioned off as
-catches and subsequently revenues for each fleet. Given revenues *R*, we
-then calculate profits $\phi$ based on revenues and costs. Costs *C* are
-calculated as a function of $\gamma_f$ that the scales the total cost
-per unit effort for fleet *f*. $\beta$ allows for the cost per unit
-effort effort to scale non-linearly, and $\eta$ is an additional cost
-per unit effort of fishing in each patch *p*, allowing for the model to
-account for travel costs for different patches. Travel costs are
-calculated based on the Euclidean distance of each patch to the nearest
-port of a given fleet, and users can specify any number from zero to the
-number of patches of port locations. When no ports are specified, travel
-costs are zero.
+This means that each species experiences a total mortality at age in a time step in a patch, individual fractions of which are portioned off as catches and subsequently revenues for each fleet.
+Given revenues *R*, we then calculate profits $\phi$ based on revenues and costs.
+Costs *C* are calculated as a function of base costs per unit effort ($\gamma$) as well as potential additional costs per unit effort of fishing in particular patches ($\eta$) for fleet *f*.
+$\beta$ allows for the cost per unit effort effort to scale non-linearly.
+Travel costs ($\eta$) are calculated based on the Euclidean distance of each patch to the nearest port of a given fleet, and users can specify any number from zero to the number of patches of port locations.
+When no ports are specified, travel costs are zero.
 
@@
 
-`marlin` allows for two general modes of effort. The simplest is
-"constant effort", in which the total effort of each fishing fleet
-remains constant over time.
-
-The more complex option is "open access". Under open access, the total
-amount of effort in a given time step for fleet *f* is a function of the
-profitability of that fleet in the proceeding time steps in which
-fishing was open, where $\theta$ controls the responsiveness of fleet
-*f* to the log of the ratio of total revenues *R* to total costs *C*,
-and approximates the proportional change in effort in response to a one
-unit changes in the log revenue to cost ratio.
+`marlin` allows for two general modes of effort. The simplest is "constant effort", in which the total effort of each fishing fleet remains constant over time.
+
+The more complex option is "open access".
+Under open access, the total amount of effort in a given time step for fleet *f* is a function of the profitability of that fleet in the proceeding time steps in which fishing was open, where $\theta$ controls the responsiveness of fleet *f* to the log of the ratio of total revenues *R* to total costs *C*, and approximates the proportional change in effort in response to a one unit changes in the log revenue to cost ratio.
 
@@
 
-This is essentially a Gompertz model for fishing effort, that has been
-used for other theoretical studies of fishing effort dynamics
-[@thorson2013].
-
-#### Tuning Fishing Fleets
-
-The degree of fishing pressure exerted by a given set of fishing fleets
-on each species is a function of a range of parameters including the
-total amount of effort, the selectivity ogives, cost functions, the
-spatial distribution of the resource, the relative prices across
-species, and the catchability coefficient of each fleet for each
-species. `marlin` provides a tuning option to help users achieve desired
-biological outcomes from their fleets.
-
-Users can tune their fleets in one of two ways. First, they can specify
-a target exploitation rate *u* for each species in their simulation.
-Taking all the other parameters of the model as given, `marlin` then
-adjusts the catchability coefficients $q_{f,s}$ for each fleet *f* and
-species *s* such that the equilibrium exploitation rate for each species
-matches the desired level. Second, they can specify a target depletion,
-defined as total spawning stock biomass under fishing divided by total
-unfished spawning stock biomass, and the model will adjust the
-catchability coefficients $q_{f,s}$ for each fleet *f* and species *s*
-such that the equilibrium depletion for each species matches the desired
-level.
+This is essentially a Gompertz model for fishing effort, that has been used for other theoretical studies of fishing effort dynamics [@thorson2013].
+
+#### Tuning Fishing Fleet Parameters
+
+The degree of fishing pressure exerted by a given set of fishing fleets on each species is a function of a range of parameters including the total amount of effort (*E*), the selectivity ogives ($\alpha$,the curve specifying the proportion of individuals at a given age or size that will be caught by a unit of fishing mortality), fishing cost (*C*), the spatial distribution of the species affected by the fleet, the relative prices across species ($\Pi$), and the catchability coefficient of each fleet for each species ($q$).
+`marlin` provides a tuning option to help users achieve desired biological outcomes from their fleets.
+
+Users can tune their fleets in one of two ways.
+First, they can specify a target exploitation rate *u* for each species in their simulation.
+Taking all the other parameters of the model as given, `marlin` then adjusts the catchability coefficients $q_{f,s}$ for each fleet *f* and species *s* such that the equilibrium exploitation rate for each species matches the desired level.
+Second, they can specify a target total spawning stock biomass under fishing divided by total unfished spawning stock biomass, and the model will adjust the catchability coefficients $q_{f,s}$ for each fleet *f* and species *s* such that the equilibrium ratio of fished to unfished spawning biomass for each species matches the desired level.
+
+Users can also tune the fleet dynamics by specifying the ratio of costs to revenues.
+Price data for use in the model can be obtained relatively simply through literature reviews, market surveys, or local experts, so long as they are converted to the same units of biomass used in the model.
+However, cost parameters are more complicated, as translating say cost per day of fishing into the same representation of fishing effort used in the model is not straight forward to accomplish.
+As an alternative, users can specify an equilibrium cost-to-revenue ratio for the fleet, essentially the profit margins of the fleet in question, and `marlin` will tune the cost parameter to achieve this desired cost to revenue ratio given the other parameters in the model (e.g.
+price and biomass of the target species).
+So for example, while users may not have access to exact cost parameters, they may have a general sense that when operating under ideal conditions a given fleet has a profit margin of 10%, i.e.
+the cost to revenue ratio is 90%.
+They can input the cost to revenue ratio as 0.9 then and allow `marlin` to tune the cost parameters accordingly.
 
@@
 
-Each fishing fleet decides how to allocate its effort in space based on
-one of four possible spatial allocation strategies. The ideal free
-distribution (IFD) is the standard method for distributing fishing
-fleets in spaces (see @gillis2003 and references therein). However,
-analytical solutions to the IFD present a number of complications for
-our model. Primarily, our model is capable of tracking multiple fleets
-at once. This means that the IFD for a single fleet would commonly be modeled as a Nash-equilibrium based on the actions of each of the individual fleet
-conditional on the actions of all other fleets. While possible to
-implement, this would slow down our model runs to the point of
-practically preventing large-scale evaluation of the spatial management policies.
-
-As such we explored a series of "next best" fleet distribution
-algorithms. While not the IFD in any individual time step, over time
-they start to approximate the IFD, as each assumes that fleets base
-their decisions for the current time step on the outcomes in the prior
-timestep, meaning that the impacts of the actions of other fishing
-fleets are somewhat internalized. The four possible fleet distribution
-algorithms are
-
-1.  Revenue per unit effort (RPUE): The fleet distributes itself in
-    space based on the realized revenue per unit effort in each patch in
-    the preceding time step
-2.  Revenue: The fleet distributes itself in space based on the realized
-    total revenue in each patch in the preceding time step
-3.  Profit per unit effort (PPUE): The fleet distributes itself in space
-    based on the realized profit per unit effort in each patch in the
-    preceding time step
-4.  Profit: The fleet distributes itself in space based on profit in
-    each patch in the preceding time step
-
-Revenue based spatial distribution is not likely to be very realistic;
-in general we would expect fishing fleets to respond to profits on some
-core level. However, due to the complexity of parameterizing cost
-functions, fleet dynamics are often evaluated based on yield or revenue
-alone, and so we include those scenarios here to allow users to evaluate
-the potential implications of this choice. The decision on whether to
-allocate the fleet based on absolute or relative (per unit effort)
-metrics is more complex. When effort represents the actions of separate
-and individual fishing actors (e.g. independent fishing vessels), a
-per-unit-effort strategy, in which fishers distribute themselves based
-on the expected catch of their individual efforts, may be more realistic
-[@hilborn1987]. Conversely, a system defined by a sole owner seeking to
-maximize total profits might be better represented by a fleet model
-based on total profits
+Each fishing fleet decides how to allocate its effort in space based on one of four possible spatial allocation strategies.
+The ideal free distribution (IFD) is the standard method for distributing fishing fleets in spaces (see @gillis2003 and references therein).
+However, analytical solutions to the IFD present a number of complications for our model.
+Primarily, our model is capable of tracking multiple fleets at once.
+This means that the IFD for a single fleet would commonly be modeled as a Nash-equilibrium based on the actions of each of the individual fleet conditional on the actions of all other fleets.
+While possible to implement, this would slow down our model runs to the point of practically preventing large-scale evaluation of the spatial management policies.
+
+As such we explored a series of "next best" fleet distribution algorithms.
+While not the IFD in any individual time step, over time they start to approximate the IFD, as each assumes that fleets base their decisions for the current time step on the outcomes in the prior timestep, meaning that the impacts of the actions of other fishing fleets are somewhat internalized.
+The four possible fleet distribution algorithms are
+
+1.  Revenue per unit effort (RPUE): The fleet distributes itself in space based on the realized revenue per unit effort in each patch in the preceding time step
+2.  Revenue: The fleet distributes itself in space based on the realized total revenue in each patch in the preceding time step
+3.  Profit per unit effort (PPUE): The fleet distributes itself in space based on the realized profit per unit effort in each patch in the preceding time step
+4.  Profit: The fleet distributes itself in space based on profit in each patch in the preceding time step
+
+Revenue based spatial distribution is not likely to be very realistic; in general we would expect fishing fleets to respond to profits on some core level.
+However, due to the complexity of parameterizing cost functions, fleet dynamics are often evaluated based on yield or revenue alone, and so we include those scenarios here to allow users to evaluate the potential implications of this choice.
+The decision on whether to allocate the fleet based on absolute or relative (per unit effort) metrics is more complex.
+When effort represents the actions of separate and individual fishing actors (e.g.
+independent fishing vessels), a per-unit-effort strategy, in which fishers distribute themselves based on the expected catch of their individual efforts, may be more realistic [@hilborn1987].
+Conversely, a system defined by a sole owner seeking to maximize total profits might be better represented by a fleet model based on total profits
 
@@
 
-The underlying population model used is an age structured single-species
-model in the manner of @ovando2021a. The population model requires many
-parameters. However, if the user supplies either a scientific or common
-name for the species in question (scientific preferred), the model will
-supply default values for that species based on the values reported in
-FishLife [internet connection required, @thorson2020]. All of the life history values can
-alternatively be replaced with user-supplied values as needed and should be for more tactical applications; the default values simply ensure that reasonable and internally consistent
-values for the species in question are supplied as a starting point.
+The underlying population model used is an age structured single-species model in the manner of @ovando2021a.
+The population model requires many parameters.
+However, if the user supplies either a scientific or common name for the species in question (scientific preferred), the model will supply default values for that species based on the values reported in FishLife [internet connection required, @thorson2020].
+FishLife provides estimates of core life history parameters for fish species based on a model integrating published values and evolutionary connections.
+FishLife provides reasonable default values for given species, but these default values should be not be taken as definitive and users should input best available parameters specific to the stock in question if they wish to best represent the dynamics of their specific system, and should check the plausibility of the default values prior to use.
+All of the life history values can be replaced with user-supplied values as needed and should be for more tactical applications; the default values simply ensure that reasonable and internally consistent values for the species in question are supplied as a starting point.
 
@@
 
-`marlin`'s movement dynamics are based on a continuous-time Markov chain
-(CTMC), as described in @thorson2021a. Within this framework, the model
-allows for movement to be broken down into three components of *advection* (follow currents), and *taxis* (active movement towards preferred habitat) and *diffusion* (essentially remaining variation in movement not explained by advection or taxis).
-For now, `marlin` focuses just on the
-diffusion and taxis components of this model, assuming that advection is zero, though future extensions
-could easily incorporate advection vectors from oceanographic models. In
-this way, `marlin` allows users to run anything from a simple Gaussian
-dispersal kernel up to a system governed by species that passively
-diffuse around a dynamic thermal range.
-
-We provide a brief overview of the the general CTMC method here, see
-@thorson2021a for a detailed description. Under this framework, movement
-from each patch to each other patch in the system in a given timestep
-*t* for life stage *a* of species *s* is defined by a movement matrix
-$\boldsymbol{M}_{t,s,a}$. $\boldsymbol{M}_{t,s,a}$ is calculated as the
-matrix exponent of the sum of the diffusion $\boldsymbol{D}$ and taxis $\boldsymbol{\tau}$ matrices scaled by the width of the time step $\Delta_{t}$ and the length of the edge of each cell $\Delta_d$ that the model is running under.
+`marlin`'s movement dynamics are based on a continuous-time Markov chain (CTMC), as described in @thorson2021a.
+Within this framework, the model allows for movement to be broken down into three components of *advection* (follow currents), and *taxis* (active movement towards preferred habitat) and *diffusion* (essentially remaining variation in movement not explained by advection or taxis).
+For now, `marlin` focuses just on the diffusion and taxis components of this model, assuming that advection is zero, though future extensions could easily incorporate advection vectors from oceanographic models.
+In this way, `marlin` allows users to run anything from a simple Gaussian dispersal kernel up to a system governed by species that passively diffuse out from a core habitat defined by a dynamic thermal range.
+
+We provide a brief overview of the the general CTMC method here, see @thorson2021a for a detailed description.
+Under this framework, movement from each patch to each other patch in the system in a given timestep *t* for life stage *a* of species *s* is defined by a movement matrix $\boldsymbol{M}_{t,s,a}$.
+$\boldsymbol{M}_{t,s,a}$ is calculated as the matrix exponent of the sum of the diffusion $\boldsymbol{D}$ and taxis $\boldsymbol{\tau}$ matrices scaled by the width of the time step $\Delta_{t}$ and the length of the edge of each patch $\Delta_d$ that the model is running under.
 So for example, if we are running a model on a monthly timestep, and we define the units of time in years, then $\Delta_{t} = 1/12$.
-Similarly, if we are running a model where one cell equals 100km^2^, then $\Delta_d = 100km$.
-This parameterization allows users to set the effective area of the spatial domain through two avenues; the number of cells, which effectively scales the resolution of the model, and the area of each cell, which scales the spatial extent of the simulation.
+Similarly, if we are running a model where one patch equals 100km^2^, then $\Delta_d = 100km$.
+This parameterization allows users to set the effective area of the spatial domain through two avenues; the number of patches, which effectively scales the resolution of the model, and the area of each patch, which scales the spatial extent of the simulation.
 
@@
 
-The individual components of the diffusion matrix are filled based on an adjacency matrix, which defines whether two cells are both adjacent and water (as opposed to land or another physical barrier), and a diffusion rate $D$ defined in units of length of the side of a cell per unit of time.
+The individual components of the diffusion matrix are filled based on an adjacency matrix, which defines whether two patches are both adjacent and water (as opposed to land or another physical barrier), and a diffusion rate $D$ defined in units of length of the side of a patch per unit of time.
 For example, if we are defining the time units as years and the distance units as kilometers, for a tuna $D$ might be 1,000 $\frac{KM}{Year}$.
-The $\Delta_{t}$ and $\Delta_{p}$ parameters then translate the movement rate $D$ to match the time step and cell size used in a simulation.
-This "scale free" parameterization means that appropriate value of $D$ can be identified for a species and then set, regardless of the time step or cell size used in the simulation model itself.
+The $\Delta_{t}$ and $\Delta_{d}$ parameters then translate the movement rate $D$ to match the time step and patch size used in a simulation.
+This "scale free" parameterization means that appropriate value of $D$ can be identified for a species and then set, regardless of the time step or patch size used in the simulation model itself.
+These parameters are then used to calculate the value of a particular entry the diffusion matrix (*d*).
 
@@
 
-The taxis matrix $\boldsymbol{\tau}$ is built off of the difference in log habitat quality between each patch, the adjacency matrix, and temporal and spatial resolution of the model ($\Delta_{t}$ and $\Delta_{p}$), as well as a scalar term $\lambda$.
-$\lambda$ scales the magnitude of the habitat
-gradient between patches relative to the diffusion
-parameter $D$. When the diffusion and taxis matrices are estimated
-jointly from actual data in the manner of @thorson2021a, $\lambda$
-should be set to zero. However, when users are manually creating both
-the diffusion and taxis matrices for simulation, the relative strength of diffusion to
-taxis must be defined by the user. Setting higher values of $\lambda$
-allows users to simulate a scenario where taxis is much more significant
-than diffusion.
+The taxis matrix $\boldsymbol{\tau}$ is built off of the difference in log habitat quality between each patch (the habitat gradient), the adjacency matrix, and temporal and spatial resolution of the model ($\Delta_{t}$ and $\Delta_{p}$), as well as a scalar term $\lambda$.
+$\lambda$ scales the magnitude of the habitat gradient between patches relative to the diffusion parameter $D$.
+When the diffusion and taxis matrices are estimated jointly from actual data in the manner of @thorson2021a, $\lambda$ should be set to zero.
+However, when users are manually creating both the diffusion and taxis matrices for simulation, the relative strength of diffusion to taxis must be defined by the user.
+Setting higher values of $\lambda$ allows users to simulate a scenario where taxis is much more significant than diffusion.
 
@@
 \tau_{p1,p2,t,s,a} = \begin{cases}
-      = \frac{\Delta_{t}}{\Delta_{d}}(1 + \lambda d_{s,f})e^{h(p2,t,s,a) - h(p1,t,s,a)} & \text{if p2 and p1 are adjacent}\\
+      = \frac{\Delta_{t}}{\Delta_{d}}(1 + \lambda d_{s,f})e^{log(h(p2,t,s,a) - log(h(p1,t,s,a))} & \text{if p2 and p1 are adjacent}\\
      = -\sum_{p' \neq p1} \tau_{p1,p2,t,s,a} & \text{if p1 = p2}\\
@@
 
-For both the diffusion and taxis matrix, we allow for the inclusion of
-physical barriers to movement (i.e. land). Pairs of patches that are
-adjacent but in which one or both patches are a barrier to movement are
-set as non-adjacent. The CTMC model then produces movement dynamics that
-move around barriers rather than over them. This is different than
-setting habitat in a patch to zero; an animal will not preferentially move
-towards a patch with zero habitat all else being equal, but can move
-over a zero habitat patch towards another better patch. They most move around a physical barrier.
-
-While this CTMC approach to movement simulation is to date not commonly
-seen in the marine modeling literature, it has numerous advantages that
-warrant its seeming complexity. First, when only diffusion is present,
-the model will generalize to the familiar dynamics of a Gaussian
-dispersal kernel. However, the time component *t* in @eq-movement allows
-the movement model to be run on arbitrarily fine time steps. Second, the
-taxis model allow for more active habitat choices by species, allowing
-us to simulate both preferences of species in space and time
-efficiently. Lastly, this form has the advantage that the parameters are
-directly estimable from real data. So, if provided with say spatial
-abundance data and a tagging study, users can estimate the movement
-parameters in the same manner as @thorson2021a, and then pass their
-estimated parameters to `marlin` for simulation.
+For both the diffusion and taxis matrix, we allow for the inclusion of physical barriers to movement (i.e.
+land).
+Pairs of patches that are adjacent but in which one or both patches are a barrier to movement are set as non-adjacent.
+The CTMC model then produces movement dynamics that move around barriers rather than over them.
+This is different than setting habitat in a patch to zero; an animal will not preferentially move towards a patch with zero habitat all else being equal, but can move over a patch with zero habitat towards another better patch.
+An example might be a species that lives and feeds on coastal habitats transiting through a patch of open ocean to reach a new feeding ground.
+In contrast, a physical barrier represents something like a peninsula that an animal must move around to transit from one side to another.
+
+While this CTMC approach to movement simulation is to date not commonly seen in the marine modeling literature, it has numerous advantages that warrant its seeming complexity.
+First, when only diffusion is present, the model will generalize to the familiar dynamics of a Gaussian dispersal kernel.
+However, the time component *t* in @eq-movement allows the movement model to be run on arbitrarily fine time steps.
+Second, the taxis model allow for more active habitat choices by species, allowing us to simulate both preferences of species in space and time efficiently.
+Lastly, this form has the advantage that the parameters are directly estimable from real data.
+So, if provided with say spatial abundance data and a tagging study, users can estimate the diffusion and taxis movement parameters in the same manner as @thorson2021a, and then pass their estimated parameters to `marlin` for simulation.
 
@@
 
-For the population model, numbers *n* at time *t* for age *a* are a
-function of growth, death, and recruitment
+For the population model, numbers *n* at time *t* for age *a* are a function of growth, death, and recruitment
 
@@
 
-where *BH* is the Beverton-Holt recruitment function, *ssb* is
-spawning-stock-biomass, $l_a$ is the mean length at at age, $l_{sel}$ is
-the length at which on average 50% of individuals are selected by the
-fishery, and $\delta_{sel}$ are the additional units of length at which
-on average 95% of fish are selected by the fishery.
-
-Spawning stock biomass *ssb* is calculated by converting age to mean
-length, calculating weight at age, maturity at age, and then calculating
-spawning stock biomass as the sum of spawning potential at age in a
-given time step, taking into account the potential for hyperallometry.
-Age is converted to length through the von Bertalanffy growth equation,
-assuming log-normally distributed variation in the length at age.
+where *BH* is the Beverton-Holt recruitment function [@beverton1957] and *ssb* is spawning-stock-biomass.
+Per convention, the model allows for a "plus group", wherein rather than tracking numbers of every possible age, individuals greater than or equal to a given maximum age are grouped together.
+
+Spawning stock biomass *ssb* is calculated by converting age to mean length, calculating weight at age, maturity at age, and then calculating spawning stock biomass as the sum of spawning potential at age in a given time step, taking into account the potential for hyperallometry.
+Age is converted to length through the von Bertalanffy growth equation given parameters asymptotic length ($l_\infty$), growth (*k*) and theoretical age at length zero ($a0$) assuming log-normally distributed variation *u* in the length at age with CV $\sigma_s$.
 
@@
 
-Where *u* is a log-normally distributed term allowing for a constant
-coefficient of variation in the length at age relationship
-
 $$
@@
 
-Biomass *b* at age is then given by the weight at length equation
-governed by a scaling coefficient $wa_s$ and an exponent $wb_s$ that
-controls the rate at which volume scales with length
-
-$$
-  b_{a,s} = wa_s \times l_{a,s}^{wb_s}
+Biomass *b* at age is then given by the weight at length equation governed by a scaling coefficient $\Omega_s$ and an exponent $\Phi_s$ that controls the rate at which volume scales with length
+
+$$
+  b_{a,s} = _s \times l_{a,s}^{wb_s}
 $$ {#eq-weight}
 
-Maturity is then calculated as a logistic function where $l_{mat}$ is
-the length at which on average 50% of individuals are sexually mature,
-and $\delta_{mat}$ is the unit of length beyond $l_{mat}$ at which on
-average 95% of fish are sexually mature.
+The proportion of sexually mature individuals (*mat*) at a given age is then calculated as a logistic function where $l_{mat}$ is the length at which on average 50% of individuals are sexually mature, and $\delta_{mat}$ is the unit of length beyond $l_{mat}$ at which on average 95% of fish are sexually mature.
 
@@
 
-Spawning stock biomass at time *t* is then calculated as a function of
-the numbers at age, the maturity at age, and the weight at age scaled by
-a parameter $\gamma$. When $\gamma$ is greater than 1, the species is
-said to experience hyperallometric fecundity, i.e. fecundity increases
-faster than weight.
+Spawning stock biomass at time *t* is then calculated as a function of the numbers at age, the maturity at age, and the weight at age raised by a parameter $\gamma$.
+When $\gamma$ is greater than 1, the species is said to experience hyperallometric fecundity, i.e.
+fecundity increases faster than weight.
 
@@
 
-Recruitment follows Beverton-Holt dynamics. We do however allow for five
-variants in the timing of density dependence:
-
-1.  Global density dependence: Density dependence is a function of the
-    sum of spawning biomass across all patches, and recruits are then
-    distributed according to habitat quality
-
-$$
-n_{t,p,s,a = 1} = \left(\frac{0.8{\times}\sum_{p=1}^P{r0_{p}}\times{h}\times\sum_{p=1}^P{ssb_{t-1,p}}}{0.2\times{\sum_{p=1}^Pssb0_p}\times(1 - h)+(h - 0.2)\times{\sum_{p=1}^Pssb_{t-1,p}}}\right)\times {hab_p}
+Recruitment (i.e.
+the number of age 1 individuals entering the population) follows Beverton-Holt dynamics parameterized around steepness (*h*) with log-normally distributed recruitment deviates $\epsilon$.
+When steepness is one recruitment is independent of spawning biomass.
+As steepness approaches 0.2 recruitment becomes a linear function of spawning biomass.
+We allow for five variants in the timing of density dependent recruitment:
+
+1.  Global density dependence: Density dependent recruitment is a function of the sum of spawning biomass across all patches, and recruits are then distributed according to habitat quality
+
+$$
+n_{t,p,s,a = 1} = \left(\frac{0.8{\times}\sum_{p=1}^P{r0_{p,s}}\times{h_s}\times\sum_{p=1}^P{ssb_{t-1,p,s}}}{0.2\times{\sum_{p=1}^Pssb0_{p,s}}\times(1 - h_s)+(h_s - 0.2)\times{\sum_{p=1}^Pssb_{t-1,p,s}}}\right)\times {r0_{p,s} / \sum_p^P{r0_{p,s}}\times \epsilon_{t,s}
+}
 $$ {#eq-dd2}
 
-where *hab* is a vector of relative recruitment habitat quality in each
-patch.
-
-$$
-hab_p = r0_p / \sum_p^P{r0_p}
-$$ {#eq-hab}
-
-2.  Local density dependence: Density dependence occurs independently in
-    each patch and recruits are retained in their home patch.
-
-$$
-n_{t,p,s,a = 1,p} = \left(\frac{0.8{\times}r0_{s,p}\times{h_s}\times{ssb_{t-1,p,s}}}{0.2\times{ssb0_{p,s}}\times(1 - h_s)+(h_s - 0.2)\times{ssb_{t-1,p}}}\right)
+where $r0$ is is a vector of recruits under unfished conditions in a given patch.
+
+2.  Local density dependence: Density dependent recruitment occurs independently in each patch and recruits are retained in their home patch.
+
+$$
+n_{t,p,s,a = 1} = \left(\frac{0.8{\times}r0_{p,s}\times{h_s}\times{ssb_{t-1,p,s}}}{0.2\times{ssb0_{p,s}}\times(1 - h_s)+(h_s - 0.2)\times{ssb_{t-1,p,s}}}\right) \times \epsilon_{t,s}
 $$ {#eq-dd1}
 
-3.  Local density dependence then disperse: Density dependence occurs
-    independently in each patch and recruits are then dispersed.
-
-$$
-  n_{t,p,s,a = 1,p} = \left(\frac{0.8{\times}r0_{s,p}\times{h_s}\times{ssb_{t-1,p,s}}}{0.2\times{ssb0_{p,s}}\times(1 - h_s)+(h_s - 0.2)\times{ssb_{t-1,p}}}\right)\times \boldsymbol{d^l} \times \epsilon_t
+3.  Local density dependence then disperse: Density dependent recruitment occurs independently in each patch and recruits are then dispersed.
+
+$$
+  n_{t,p,s,a = 1} = \left(\frac{0.8{\times}r0_{p,s}\times{h_s}\times{ssb_{t-1,p,s}}}{0.2\times{ssb0_{p,s}}\times(1 - h_s)+(h_s - 0.2)\times{ssb_{t-1,p,s}}}\right)\times \boldsymbol{d^l_s} \times \epsilon_{t,s}
 $$ {#eq-dd3}
@@
 
-4.  Post-dispersal density dependence: Larvae are distributed throughout
-    the system, and then density dependence occurs based on the density
-    of adult biomass at the destination patch.
-
-$$
-  larv_{t,p} = ssb_{t-1} \times\boldsymbol{d^l}
+4.  Post-dispersal density dependence: Larvae are distributed throughout the system, and then density dependent recruitment occurs based on the density of adult biomass at the destination patch.
+
+$$
+  larv_{t,p,s} = ssb_{t-1,p,s} \times\boldsymbol{d^l_s}
 $$ {#eq-larvmove}
@@
 $$
-n_{t,a = 1,p} = \left(\frac{0.8{\times}r0_{p}\times{h}\times{larv_{t,p}}}{0.2\times{ssb0_p}\times(1 - h)+(h - 0.2)\times{larv_{t,p}}}\right) \times \epsilon_t
+n_{t,a = 1,p,s} = \left(\frac{0.8{\times}r0_{p,s}\times{h_s}\times{larv_{t,p,s}}}{0.2\times{ssb0_{p,s}}\times(1 - h_s)+(h_s - 0.2)\times{larv_{t,p,s}}}\right) \times \epsilon_{t,s}
 $$ {#eq-dd4}
 
-5.  Global density dependence allocated by spawning biomass: Density
-    dependence is a function of the sum of spawning biomass across all
-    patches, and recruits are then distributed according to the
-    distribution of spawning biomass
-
-$$
-n_{t,p,s,a = 1} = \left(\frac{0.8{\times}\sum_{p=1}^P{r0_{p}}\times{h}\times\sum_{p=1}^P{ssb_{t-1,p}}}{0.2\times{\sum_{p=1}^Pssb0_p}\times(1 - h)+(h - 0.2)\times{\sum_{p=1}^Pssb_{t-1,p}}}\right)\times \frac{ssb_{t-1,p}}{\sum_{p=1}^Pssb_{t-1,p}}
+5.
+Global density dependence allocated by spawning biomass: Density dependence is a function of the sum of spawning biomass across all patches, and recruits are then distributed according to the distribution of spawning biomass
+
+$$
+n_{t,p,s,a = 1} = \left(\frac{0.8{\times}\sum_{p=1}^P{r0_{p,s}}\times{h_s}\times\sum_{p=1}^P{ssb_{t-1,p,s}}}{0.2\times{\sum_{p=1}^Pssb0_{p,s}}\times(1 - h_s)+(h_s - 0.2)\times{\sum_{p=1}^Pssb_{t-1,p,s}}}\right)\times \frac{ssb_{t-1,p,s}}{\sum_{p=1}^Pssb_{t-1,p,s}} \times \epsilon_{t,s}
 $$ {#eq-dd5}
 
+Log-normal recruitment deviates are calculated with the potential for autocorrelation defined with strength $\rho$
+
+$$ 
+log(\epsilon_{t,s}) \sim \begin{cases}
+       N(0,\sigma_{r,s}), & \text{if $t = 1$}\\
+       \rho_s log(\epsilon_{t-1,s}) + \sqrt{1 - \rho_s^2} N(0,\sigma_{r,s}), & \text{if $t > 1$}
+  \end{cases}
+$$ {#eq-recdev}
+
+And log recruitment deviates are converted to raw units using the bias correction factor
+
+$$
+\epsilon_{t,s} = e^{log(\epsilon_{t,s}) - \sigma_{r,s}^2/2}
+$$ {#eq-bias}
+
 #### Reference Points
 
-Fisheries management is often concerned with measuring stock status
-relative to maximum sustainable yield (MSY) based reference points,
-though the exact level of stock status desired by management or societal
-objectives may vary widely. MSY based reference points present a problem
-for a multi-fleet and spatial-temporal model such as `marlin`. MSY and
-the fishing mortality rate that would produce MSY, $F_{MSY}$, are a
-function of fishery selectivity. Fishery selectivity in this model can
-vary by fleet, and species can be distributed unevenly in both space and
-time. This means that the net effective fishing selectivity on a species
-can vary depending on the dynamics at a given moment, making the
-definition of an equilibrium concept such as MSY challenging.
-
-As such, we do not report MSY based reference points in the model by
-default, with the exception of the spawning stock biomass at MSY
-($SSB_{MSY}$) which tends to be conserved across selectivity patterns as
-it is primarily a function of the mortality and flow of recruits
-relative to the maturity ogive. There are many different strategies for
-estimating reference points in spatially explicit systems [@kapur2021a].
-We leave it to users to define and find relevant reference point as
-required by their specific needs.
+Fisheries management is often concerned with measuring stock status relative to maximum sustainable yield (MSY) based reference points, though the exact level of stock status desired by management or societal objectives may vary widely.
+MSY based reference points present a problem for a multi-fleet and spatial-temporal model such as `marlin`.
+MSY and the fishing mortality rate that would produce MSY, $F_{MSY}$, are a function of fishery selectivity.
+Fishery selectivity in this model can vary by fleet, and species can be distributed unevenly in both space and time.
+This means that the net effective fishing selectivity on a species can vary depending on the dynamics at a given moment, making the definition of an equilibrium concept such as MSY challenging.
+
+As such, we do not report MSY based reference points in the model by default, with the exception of the spawning stock biomass at MSY ($SSB_{MSY}$) which tends to be conserved across selectivity patterns as it is primarily a function of the mortality and flow of recruits relative to the maturity ogive.
+There are many different strategies for estimating reference points in spatially explicit systems [@kapur2021a].
+We leave it to users to define and find relevant reference point as required by their specific needs.
 
@@
 
-We include two examples demonstrating how `marlin` can be used to
-support marine spatial planning. In the first, we show how `marlin` can
-be used to compare the total and distributional impacts of MPAs designed
-in a heavily fished coastal coral reef ecosystem. In the second, we demonstrate
-how `marlin` can be used to assess components of climate resilience of
-alternative MPA design strategies in a pelagic system.
-
-Each of the case studies contains too many parameters and options to be
-succinctly presented in the text here, and we are concerned that
-attempting to duplicate all of the parameters of the model here would
-simply introduce error. As such, readers should consult the accompanying
-code to view the
-precise details of each simulation. Targeted tactical applications must
-carefully consider and document all decisions made around model
-parameters.
+We include two examples demonstrating how `marlin` can be used to support marine spatial planning.
+In the first, we show how `marlin` can be used to compare the total and distributional impacts of MPAs designed in a heavily fished coastal coral reef ecosystem.
+In the second, we demonstrate how `marlin` can be used to assess components of climate resilience of alternative MPA design strategies in a pelagic system.
+
+Each of the case studies contains too many parameters and options to be succinctly presented in the text here, and we are concerned that attempting to duplicate all of the parameters of the model here would simply introduce error.
+As such, readers should consult the accompanying code to view the precise details of each simulation.
+Targeted tactical applications must carefully consider and document all decisions made around model parameters.
 
@@
 
-Across all of our case studies we make use of two potential "rule of
-thumb" MPA design strategies:
-
-1.  *Rate*: MPAs are placed based on the pre-MPA
-    depletion-weighted catch relative to the total catch in a patch. So,
-    patches with high rates of catch of depleted species relative to
-    total catch are prioritized.
-
-2.  *Target Fishing*: MPAs locations are prioritized
-    proportional to fishery catches. Patches with high total
-    catches are prioritized over patches with low catches.
-
-To our knowledge, MPAs are never randomly placed; their locations are
-decided based on stakeholder negotiation given available information and
-objectives. In theory, the design of MPA networks can be optimized
-through the use of a modeling framework, and depending on the
-validity of the model, this process is likely to produce better
-outcomes than manually-designed strategies [@rassweiler2012;
-@rassweiler2014]. However, designing optimal MPA networks becomes
-increasingly difficult as the range of objectives and the complexity of
-the model increase. Therefore, while `marlin` provides some support for
-[optimization](https://danovando.github.io/marlin/articles/optimize-mpas.html),
-we focus here on the design and performance of these more "rule of
-thumb" design strategies that may be more accessible to a wider range of
-users. We allow all MPAs to be designed in a "mosaic" fashion, but users
-can easily extend the analysis to compare outcomes between
-"contiguous" (MPA is made up of one continuously connected block) and "mosaic" (MPAs can be separated in space) MPA designs [@pons2022].
+Across all of our case studies we make use of two potential "rule of thumb" MPA design strategies:
+
+1.
+*Rate*: MPAs are placed based on the pre-MPA SSB/SSB0 weighted catch relative to the total catch in a patch.
+So, patches with high rates of catch of depleted species relative to total catch are prioritized.
+
+2.  *Target Fishing*: MPAs locations are prioritized proportional to fishery catches. Patches with high total catches are prioritized over patches with low catches.
+
+To our knowledge, MPAs are never randomly placed; their locations are decided based on stakeholder negotiation given available information and objectives.
+In theory, the design of MPA networks can be optimized through the use of a modeling framework, and depending on the validity of the model, this process is likely to produce better outcomes than manually-designed strategies [@rassweiler2012; @rassweiler2014].
+However, designing optimal MPA networks becomes increasingly difficult as the range of objectives and the complexity of the model increase.
+Therefore, while `marlin` provides some support for [optimization](https://danovando.github.io/marlin/articles/optimize-mpas.html), we focus here on the design and performance of these more "rule of thumb" design strategies that may be more accessible to a wider range of users.
+We allow all MPAs to be designed in a "mosaic" fashion, but users can easily extend the analysis to compare outcomes between "contiguous" (MPA is made up of one continuously connected block) and "mosaic" (MPAs can be separated in space) MPA designs [@pons2022] by developing and passing their own MPA design algorithms to the model.
 
@@
 
-In our coral reef example, we model the dynamics of four species: a
-grouper (*Epinephelus fuscoguttatus*), a shallow-reef snapper (*Lutjanus
-malabaricus*), a deep-water snapper (*Pristipomoides filamentosus*), and
-a reef shark (*Carcharhinus amblyrhynchos*). The simulated groupers
-undergo a mass migration to a spawning aggregation once per year,
-followed by the sharks. Shallow-reef snappers stay in reefs closer to
-shore year-round, while deep-water snappers stay in the deeper reefs
-(@fig-coral).
-
-These species are targeted by two different fleets. Fleet one primarily
-targets the grouper and nearshore snapper populations, but will land any
-incidentally captured sharks. Fleet one has a dome shaped selectivity
-pattern for both grouper and shallow-reef snapper due to demand for
-"plate sized" fish [@kindsvater2017], and logistic selectivity for
-sharks, as any shark caught is retained. This fleet is a subsistence
-fleet, where the local community takes advantage of every possible opportunity to
-fish, and as such we model it as a "constant effort"
-fishery.
-
-The second fleet primarily targets the deep-water snapper population.
-While plate-sized snapper are still prized from this population, we
-model fleet two's selectivity as logistic due to high levels of discard
-mortality for deep-water snapper resulting from barotrauma. Fleet two
-catches some grouper, though less than fleet one, and receives no price
-for sharks due to the requirements of a certification program through
-which they sell their deep-water snapper. Accidental captures (bycatch) do
-occur, which results in mortality. This fleet operates under open-access
-dynamics, as fishing is not the primary means of subsistence;
-short-term effort expands and contracts in response to profitability of
-the primarily grouper-driven fishery.
-
-Each fleet has a different home-port location, and incurs an additional
-cost for fishing each patch in the system in proportion to the Euclidean
-distance from each patch to each fleet's home port.
-
-We use `marlin` to simulate the outcomes for both food production and
-conservation for each of the species and both of the fleets as a
-function of both MPA size and MPA design strategy. For this exercise,
-MPAs are placed with perfect information and have no design constraints
-for continuity (i.e. the selected MPA design can be as patchwork as
-dictated by the data). We ran the simulation on a quarterly level ($\Delta_t = 1/4$) and set the area of each cell to an arbitrary value of 1 ($\Delta_p = 1$).
+In our coral reef example, we model the dynamics of four tropical Pacific species: a grouper (*Epinephelus fuscoguttatus*), a shallow-reef snapper (*Lutjanus malabaricus*), a deep-water snapper (*Pristipomoides filamentosus*), and a reef shark (*Carcharhinus amblyrhynchos*).
+The simulated groupers undergo a mass migration to a spawning aggregation once per year, followed by the sharks.
+Shallow-reef snappers stay in reefs closer to shore year-round, while deep-water snappers stay in the deeper reefs (@fig-coral).
+
+These species are targeted by two different fleets.
+Fleet One primarily targets the grouper and nearshore snapper populations, but will land any incidentally captured sharks.
+Fleet One has a dome shaped selectivity pattern for both grouper and shallow-reef snapper due to demand for "plate sized" fish [@kindsvater2017], and logistic selectivity for sharks, as any shark caught is retained.
+Fleet One is totally dependent on fishing for their livelihood, meaning the local community takes advantage of every possible opportunity to fish, and as such we model it as a "constant effort" fishery.
+
+Fleet Two primarily targets the deep-water snapper population.
+While plate-sized snapper are still prized from this population, we model Fleet Two's selectivity as logistic due to high levels of discard mortality for deep-water snapper resulting from barotrauma.
+Fleet Two catches some grouper, though less than Fleet One, and receives no price for sharks due to the requirements of a certification program through which they sell their deep-water snapper.
+Accidental captures (bycatch) do occur, which results in mortality.
+Fleet Two operates under open-access dynamics, as fishing is not the primary means of subsistence; short-term effort expands and contracts in response to profitability of the primarily grouper-driven fishery.
+
+Each fleet has a different home-port location, and incurs an additional cost for fishing each patch in the system in proportion to the Euclidean distance from each patch to each fleet's home port.
+
+We use `marlin` to simulate the outcomes for both food production and conservation for each of the species and both of the fleets as a function of both MPA size and MPA design strategy.
+For this exercise, MPAs are placed with perfect information and have no design constraints for continuity (i.e.
+the selected MPA design can be as patchwork as dictated by the data).
+We ran the simulation on a quarterly level ($\Delta_t = 1/4$) and set the area of each patch to an arbitrary value of 1 ($\Delta_p = 1$).
 
@@
 
-We model our pelagic use case loosely on the characteristics of the
-Western and Central Pacific Ocean (WCPO) tuna fisheries. We simulate
-trajectories of `r n_distinct(names(blue_fauna))` species commonly
-caught in the region, including both the highly abundant skipjack tuna
-to the heavily depleted oceanic whitetip shark (*Carcharhinus
-longimanus*) (@fig-blue-ssb). We use publicly available data on
-catch-per-unit-effort of each of these species from the WCPO as a proxy
-for baseline habitat distributions, noting that where possible,
-fishery-independent abundance indices would be preferable (@fig-blue).
-
-These pelagic species are caught by a longline fleet that primarily
-targets large adult tunas such as bigeye and yellowfin for high-grade
-consumption, and a purse-seine fishery that primarily targets skipjack
-tunas for bulk canning. Selectivities were modeled as logistic for the
-longline fleet, and dome-shaped for the purse-seine fleet, with
-selectivity values set to broadly match the patterns reported in
-relevant [stock
-assessments](https://oceanfish.spc.int/en/ofpsection/sam/sam/210-stock-assessment-input-and-result-files-)
-for each species. Both fleets operate under open-access dynamics with an
-effort cap. The effort cap was set at the level of effort that resulted
-in the desired levels of depletion for each species under open-access
-dynamics (@fig-blue), intended to simulate a scenario where managers
-step in to prevent further expansion of fishing effort in a fully
-developed fishery. For forward-simulation, open-access dynamics can
-result in effort decreasing in response to profitability, but cannot
-result in effort beyond the effort cap set for each fleet.
-
-For this exercise, we focused on using `marlin` to assess resilience of
-the selected "rule of thumb" MPA design strategies to a climate-driven
-range shift. Specifically, we simulate an extreme example where the
-centroid of each population shifts northward at a rate of 0.2 patches
-per year, over a 20 year time horizon (@fig-blue-ssb).
-We designed MPA networks given the conditions in the starting year, and
-then held that network constant over the years of the experiment,
-running one simulation with and another without the climate-drive range
-shift. We then compared the effects of this range shift on food
-production and conservation outcomes from MPA networks designed based on
-the pre-range shift world. We ran the simulation on a quarterly level ($\Delta_t = 1/4$) and set the area of each cell to an arbitrary value of 1 ($\Delta_p = 1$).
+We model our pelagic use case loosely on the characteristics of the Western and Central Pacific Ocean (WCPO) tuna fisheries.
+We simulate trajectories of `r n_distinct(names(blue_fauna))` species commonly caught in the region, including both the highly abundant skipjack tuna to the heavily depleted oceanic whitetip shark (*Carcharhinus longimanus*) (@fig-blue-ssb).
+We use publicly available data on catch-per-unit-effort of each of these species from the WCPO as a proxy for baseline habitat distributions, noting that where possible, fishery-independent abundance indices would be preferable (@fig-blue).
+
+These pelagic species are caught by a longline fleet that primarily targets large adult tunas such as bigeye and yellowfin for high-grade consumption, and a purse-seine fishery that primarily targets skipjack tunas for bulk canning.
+Selectivities were modeled as logistic for the longline fleet, and dome-shaped for the purse-seine fleet, with selectivity values set to broadly match the patterns reported in relevant [stock assessments](https://oceanfish.spc.int/en/ofpsection/sam/sam/210-stock-assessment-input-and-result-files-) for each species.
+Both fleets operate under open-access dynamics with an effort cap.
+The effort cap was set at the level of effort that resulted in the desired levels of SSB/SSB0 for each species under open-access dynamics (@fig-blue), intended to simulate a scenario where managers step in to prevent further expansion of fishing effort in a fully developed fishery.
+For forward-simulation, open-access dynamics can result in effort decreasing in response to profitability, but cannot result in effort beyond the effort cap set for each fleet.
+
+For this exercise, we focused on using `marlin` to assess resilience of the selected "rule of thumb" MPA design strategies to a climate-driven range shift.
+Specifically, we simulate an extreme example where the centroid of each population shifts northward at a rate of 0.2 patches per year, over a 20 year time horizon (@fig-blue-ssb).
+We designed MPA networks given the conditions in the starting year, and then held that network constant over the years of the experiment, running one simulation with and another without the climate-drive range shift.
+We then compared the effects of this range shift on food production and conservation outcomes from MPA networks designed based on the pre-range shift world.
+We ran the simulation on a quarterly level ($\Delta_t = 1/4$) and set the area of each patch to an arbitrary value of 1 ($\Delta_p = 1$).
 
@@
 #| label: fig-blue-ssb
-#| fig-cap: "Trajectory of depletion (spawning biomass divided by unfished spawning biomass) over time prior to implementation of MPAs for the pelagic case study."
+#| fig-cap: Spawning biomass divided by unfished spawning biomass in the time period prior to implementation of MPAs for the pelagic case study."
 
@@
   geom_col() +
-  scale_y_continuous(limits = c(0, 1), name = "Depletion", expand = expansion(mult = c(0.0,0.05))) +
+  scale_y_continuous(limits = c(0, 1), name = "SSB/SSB0", expand = expansion(mult = c(0.0,0.05))) +
   labs(x = "") + 
@@
 
-For the coastal coral reef example the *Target fishing* MPA design strategies was capable of producing a Pareto-improving, i.e. potentially win-win, outcome where the total
-spawning biomass divided by unfished spawning biomass across all species
-increase by up to 200% while maintaining at least the same levels of
-fisheries yields as would be caught in the absence of any MPAs.
-Conversely, achieving increases in total fish biomass through the *Rate* strategy resulted in immediate losses in total yields, and in losses to both yields and total biomass in some cases
-(@fig-coral-results).
-
-These macro-level changes in total yield and conservation outcomes mask
-some internal tradeoffs. Nearly all of the food production benefits from the simulated MPA networks accrued to fleet one, while catches by fleet two were generally
-negatively impacted under both scenarios. For example, the MPA sizes and designs that result in the largest increase in yields for fleet one would produce a near total
-loss of fish catch for fleet two.
-
-The *Rate* MPA placement strategy produced
-lower total fishery gains but minimized the negative impacts on fleet
-two. However, this strategy produced much lower conservation gains
-than the *Target Fishing* strategy, only benefiting
-deep snapper meaningfully, and in fact resulting in slight losses to the
-nearshore snapper and grouper populations for MPA sizes less than roughly 75% of the
-simulated space. The *Target Fishing* strategy
-resulted in rapid conservation gains across all of the simulated
-species.
-
-MPAs were capable of producing positive outcomes in terms of total
-biomass and yields in the coastal coral reef example, but all the
-tested strategies resulted in tradeoffs across at least one dimension. `marlin` could be further used
-by a community such as the one simulated here to explore alternative design
-strategies, or to assess options for distributing fishery benefits
-across communities that may experience very different food provision outcomes from the
-same MPA network.
+For the coastal coral reef example the *Target fishing* MPA design strategies was capable of producing a Pareto-improving, i.e.
+potentially win-win, outcome where the total spawning biomass divided by unfished spawning biomass across all species increase by up to 200% while maintaining at least the same levels of fisheries yields as would be caught in the absence of any MPAs.
+Conversely, achieving increases in total fish biomass through the *Rate* strategy resulted in immediate losses in total yields, and in losses to both yields and total biomass in some cases (@fig-coral-results).
+
+These macro-level changes in total yield and conservation outcomes mask some internal tradeoffs.
+Nearly all of the food production benefits from the simulated MPA networks accrued to fleet one, while catches by fleet two were generally negatively impacted under both scenarios.
+For example, the MPA sizes and designs that result in the largest increase in yields for fleet one would produce a near total loss of fish catch for fleet two.
+
+The *Rate* MPA placement strategy produced lower total fishery gains but minimized the negative impacts on fleet two.
+However, this strategy produced much lower conservation gains than the *Target Fishing* strategy, only benefiting deep snapper meaningfully, and in fact resulting in slight losses to the nearshore snapper and grouper populations for MPA sizes less than roughly 75% of the simulated space.
+The *Target Fishing* strategy resulted in rapid conservation gains across all of the simulated species.
+
+MPAs were capable of producing positive outcomes in terms of total biomass and yields in the coastal coral reef example, but all the tested strategies resulted in tradeoffs across at least one dimension.
+`marlin` could be further used by a community such as the one simulated here to explore alternative design strategies, or to assess options for distributing fishery benefits across communities that may experience very different food provision outcomes from the same MPA network.
 
@@
 #| label: fig-coral-results
-#| fig-cap: "Percent change in yield per fleet (A) and depletion (spawning biomass divided by unfished spawning biomass) per species (B) as a function of MPA size and MPA design strategy for the coastal coral reef case study.
-Percent change in total yield across both fleets (y-axis) and total depletion across all species (x-axis) and placement strategies (line color) (C).
+#| fig-cap: "Percent change in yield per fleet (A) and SSb/SSB0 (spawning biomass divided by unfished spawning biomass) per species (B) as a function of MPA size and MPA design strategy for the coastal coral reef case study.
+Percent change in total yield across both fleets (y-axis) and total SSB/SSB0 across all species (x-axis) and placement strategies (line color) (C).
 Text labels in panel C indicate the percent of area in an MPA at selected points along curves."
@@
   geom_line() +
-  scale_x_continuous(name = "Change in Total Depletion",limits = c(0, NA)) +
+  scale_x_continuous(name = "Change in SSB/SSB0",limits = c(0, NA)) +
     scale_y_continuous(name = "Total Yield",limits = c(0, NA)) +
@@
   geom_text_repel(data = frontier_labels, aes(label = pmpa), min.segment.length = 0, box.padding = 0.5) +
-  scale_x_continuous(name = "Change in Total Depletion", labels = scales::label_percent(accuracy = 1), guide = guide_axis(n.dodge = 2)) +
+  scale_x_continuous(name = "Change in Total SSB/SSB0", labels = scales::label_percent(accuracy = 1), guide = guide_axis(n.dodge = 2)) +
   scale_y_continuous(name = "Change in Yield", labels = scales::label_percent(accuracy = 1)) + 
@@
                      guide = guide_axis(n.dodge = 2)) +
-  scale_y_continuous(name = "Depletion") + 
+  scale_y_continuous(name = "SSB/SSB0") + 
   scale_color_brewer(name = '', type = "qual", palette = "Dark2") + 
@@
 
-The yield impacts of the proposed MPA strategies were much more
-sensitive to climate impacts for the purse seine fishery. Under the
-*status quo* species distribution, the *Target Fishing* MPA design
-strategy produced increases in fisheries yields at sizes between 25-50%
-of the simulated area; that same MPA network resulted in the near
-complete loss of yields from the purse seine fleet when applied to the
-climate change scenario. In contrast, the yield outcomes for the
-longline fleet were relatively insensitive to the simulated range shifts
-(@fig-blue-results).
-
-From a conservation perspective, species such as albacore and swordfish
-were relatively insensitive to the simulated range shifts, whereas
-species such as bigeye tuna and silky shark showed extreme variation in
-outcomes depending on the climate scenario. For bigeye tuna, under the
-*status quo* conditions the *Target Fishing* MPA design strategy
-resulted in rapid increases in the population, but that same MPA network produced much smaller and slower gains under climate change.
-However, for blue sharks this trend is inverted, with more rapid
-conservation gains under the range shift scenario than the *status quo*
-(@fig-blue-results).
+The yield impacts of the proposed MPA strategies were much more sensitive to climate impacts for the purse seine fishery.
+Under the *status quo* species distribution, the *Target Fishing* MPA design strategy produced increases in fisheries yields when the MPA covered 25-50% of the simulated area; that same MPA network resulted in the near complete loss of yields from the purse seine fleet when applied to the climate change scenario.
+In contrast, the yield outcomes for the longline fleet were relatively insensitive to the simulated range shifts (@fig-blue-results).
+
+From a conservation perspective, spawning biomass of species such as albacore and swordfish was relatively insensitive to the simulated range shifts, whereas species such as bigeye tuna and silky shark showed extreme variation in outcomes depending on the climate scenario.
+For bigeye tuna, under the *status quo* conditions the *Target Fishing* MPA design strategy resulted in rapid increases in the population, but that same MPA network produced much smaller and slower gains under climate change.
+However, for blue sharks this trend is inverted, with more rapid conservation gains under the range shift scenario than the *status quo* (@fig-blue-results).
 
@@
   scale_x_continuous(name = "MPA",labels = scales::label_percent()) +
-  scale_y_continuous(name = "Difference in Depletion")
+  scale_y_continuous(name = "Difference in SSB/SSB0")
 
@@
 #| label: fig-blue-results
-#| fig-cap: "Change in yield (A) and depletion (spawning biomass divided by unfished spawning biomass) (B) as a function of MPA size and design strategy by fishing fleet and species.
+#| fig-cap: "Change in yield (A) and SSB/SSB0 (spawning biomass divided by unfished spawning biomass) (B) as a function of MPA size and design strategy by fishing fleet and species.
 Blue lines indicate impacts of MPAs under status qup habitat, red impacts under climate-driven range shift."
@@
   geom_line() +
-  scale_x_continuous(name = "Change in Total Depletion",limits = c(0, NA)) +
+  scale_x_continuous(name = "Change in Total SSB/SSB0",limits = c(0, NA)) +
     scale_y_continuous(name = "Total Yield",limits = c(0, NA)) +
@@
                      labels = scales::label_percent(accuracy = 1), guide = guide_axis(n.dodge = 2)) +
-  scale_y_continuous(name = "Depletion",limits = c(0, NA)) +
+  scale_y_continuous(name = "SSB/SSB0",limits = c(0, NA)) +
   scale_linetype(name = "MPA Strategy") + 
@@
 Marine ecosystems are driven by complex social-ecological dynamics.
-Communities must often make decisions on how to manage these systems based on
-limited empirical evidence. Modeling tools such as `marlin` can help
-users answer important scientific questions and design marine management
-policies informed by a better understanding of sensitivities to key
-uncertainties.
-
-The role of this paper is not to conduct a review of the many
-tools for spatial-temporal modeling available in the literature, each of which provide useful functionality for specific applications.
-However, we highlight here the specific gaps that we feel `marlin` fills
-in the modeling literature and in the policy support toolbox using some
-selected publications. More end-to-end models such as ATLANTIS
-[@audzijonyte2019] , OSMOSE [@shin2001], Ecopath with Ecosim
-[@christensen2004], POSEIDON [@bailey2019], or SEAPODYM [@lehodey2008]
-are capable of representing tremendous amounts of complexity, but can be
-extremely time consuming to design and run. `marlin` allows some of the realism of these end-to-end models while being simpler and faster to parameterize and run.
-
-Tools such as virtualspecies
-[@leroy2016], STEPS [@visintin2020], RangeShifter [@bocedi2014], and
-[SMS](https://www.aquaveo.com/software/sms-surface-water-modeling-system-introduction)
-can be efficiently constructed to model the dispersal and distribution
-of species in space and time as a function of environmental covariates.
-However, these typically require specifying a covariate-response curve
-without explicitly acknowledging how this arises from habitat-specific
-movement and demography (virtualspecies), or model movement using a
-dispersal kernel (RangeShifter, STEPS) or a least-cost path algorithm
-(SMS, STEPS). By contrast, `marlin` uses a continuous-time Markov chain
-movement model, which integrates multiple paths (including their
-path-dependent probability based on intervening habitat types and
-species preferences) while using scale-free parameters that can be
-measured experimentally in laboratory or tagging studies.
-
-Simplified bio-economic models such as those used in @hastings2017,
-@sala2021, and @cabral2019 can be applied at scale and provide
-analytically tractable results, but as a result must abstract over many
-bio-economic dynamics that can be important for more tactical
-applications. `marlin` allows for more realistic representations of spatial social-ecological systems while maintaining processing speed.
-
-In sum, `marlin` complements the existing spatial marine modeling
-literature by allowing scientists, decision makers, and other
-stakeholders to efficiently and productively examine the impacts of
-realistic bio-economic dynamics on academic and applied problems.
+Communities must often make decisions on how to manage these systems based on limited empirical evidence.
+Modeling tools such as `marlin` can help users answer important scientific questions and design marine management policies informed by a better understanding of sensitivities to key uncertainties.
+In this paper we have documented the core system of equations underlying `marlin`, and demonstrated how `marlin` can be used to support efforts such as MPA planning.
+The variation in both the magnitude and distribution of costs and benefits stemming from alternative design choices and states of nature shown in our case study results demonstrates the value in our method.
+For example, our coastal coral reef example illustrates both the potential for MPAs to benefit conservation and food production in these systems, and the potential for the same MPA to benefit some fleets and species while harming others when fishing fleets affected by an MPA do not share the same objectives and species are not uniformly distributed.
+Our pelagic case study illustrates how `marlin` can be used to assess the climate resiliency of spatial management strategies, and identify strategies that best meet the needs of both current and future conditions.
+Results from this type of work could be used to help prioritize communities and species at particular risk to climate change impacts.
+
+The role of this paper is not to conduct a review of the many tools for spatial-temporal modeling available in the literature, each of which provide useful functionality for specific applications.
+However, we highlight here the specific gaps that we feel `marlin` fills in the modeling literature and in the policy support toolbox using some selected publications.
+More end-to-end models such as ATLANTIS [@audzijonyte2019] , OSMOSE [@shin2001], Ecopath with Ecosim [@christensen2004], POSEIDON [@bailey2019], DISPLACE [@bastardie2013], or SEAPODYM [@lehodey2008] are capable of representing tremendous amounts of complexity, but can be extremely time consuming to design and run.
+`marlin` allows some of the realism of these end-to-end models while being simpler and faster to parameterize and run.
+
+Tools such as virtualspecies [@leroy2016], STEPS [@visintin2020], RangeShifter [@bocedi2014], and [SMS](https://www.aquaveo.com/software/sms-surface-water-modeling-system-introduction) can be efficiently constructed to model the dispersal and distribution of species in space and time as a function of environmental covariates.
+However, these typically require specifying a covariate-response curve without explicitly acknowledging how this arises from habitat-specific movement and demography (virtualspecies), or model movement using a dispersal kernel (RangeShifter, STEPS) or a least-cost path algorithm (SMS, STEPS).
+By contrast, `marlin` uses a continuous-time Markov chain movement model, which integrates multiple paths (including their path-dependent probability based on intervening habitat types and species preferences) while using scale-free parameters that can be measured experimentally in laboratory or tagging studies.
+
+Simplified bio-economic models such as those used in @hastings2017, @sala2021, and @cabral2019 can be applied at scale and provide analytically tractable results, but as a result must abstract over many bio-economic dynamics that can be important for more tactical applications.
+`marlin` allows for more realistic representations of spatial social-ecological systems while maintaining processing speed.
+
+In sum, `marlin` complements the existing spatial marine modeling literature by allowing scientists, decision makers, and other stakeholders to efficiently and productively examine the impacts of realistic bio-economic dynamics on academic and applied problems.
 For example, scientists could use `marlin` to consider the relative importance of different biological variables in affecting policy outcomes.
@@
 
-In this paper we have documented the core system of equations underlying
-`marlin`, and demonstrated how `marlin` can be used to support efforts
-such as MPA planning. The variation in both the magnitude and
-distribution of costs and benefits stemming from alternative design
-choices and states of nature shown in our case study results
-demonstrates the value in our method. For example, our coastal coral
-reef example illustrates both the potential for MPAs to benefit
-conservation and food production in these systems, and the potential for
-the same MPA to benefit some fleets and species while harming others when fishing
-fleets affected by an MPA do not share the same objectives and species are not uniformly distributed. Our pelagic
-case study illustrates how `marlin` can be used to assess the climate
-resiliency of spatial management strategies, and identify strategies
-that best meet the needs of both current and future conditions. Results
-from this type of work could be used to help prioritize communities and
-species at particular risk to climate change impacts.
-
-Beyond the current examples, we envision `marlin` being applicable to
-research on dynamic ocean management, range shifts, management strategy
-evaluation, policy interactions, and spatial stock assessment. In
-addition, the process-based movement model used in `marlin` can directly
-use empirical estimates of movement dynamics derived in the manner of
-@thorson2021a, providing a link between empirical and simulation based
-approaches to marine resource management. Future iterations of the model
-may include life history modules for a broader range of species types
-such as marine mammals, sea birds, and sea turtles.
-
-Even the most complex marine model is a stylized cartoon of the true
-dynamics at play in the ocean. However, for all their limitations
-models such as `marlin` can help users understand factors that drive the
-performance of marine management strategies. The goal of this tool is to
-empower people to design policies based on evaluation of key
-uncertainties and trade-offs, and in doing so support more effective and
-equitable marine resource management.
+Beyond the current examples, we envision `marlin` being applicable to research on dynamic ocean management, range shifts, management strategy evaluation, policy interactions, and spatial stock assessment.
+In addition, the process-based movement model used in `marlin` can directly use empirical estimates of movement dynamics derived in the manner of @thorson2021a, providing a link between empirical and simulation based approaches to marine resource management.
+Future iterations of the model may include life history modules for a broader range of species types such as marine mammals, sea birds, and sea turtles.
+
+Even the most complex marine model is a stylized cartoon of the true dynamics at play in the ocean.
+However, for all their limitations models such as `marlin` can help users understand factors that drive the performance of marine management strategies.
+The goal of this tool is to empower people to design policies based on evaluation of key uncertainties and trade-offs, and in doing so support more effective and equitable marine resource management.
 
@@
 
-All code and materials needed to reproduce this paper is available at (https://github.com/DanOvando/marlin-paper)[https://github.com/DanOvando/marlin-paper].
+All code and materials needed to reproduce this paper is available at (https://github.com/DanOvando/marlin-paper)\[https://github.com/DanOvando/marlin-paper\].
 

### 2023-03-22 03_marlin_paper.qmd
tuna time

@@
 
-Abstract: Designing effective spatial management strategies is challenging because marine ecosystems are highly dynamic and opaque, and extractive entities such as fishing fleets respond endogenously to ecosystem changes in ways that depend upon ecological and policy context.
-We present a tool -- marlin -- to efficiently simulate the bio-economic dynamics of marine systems, designed to support both management and research.
-We demonstrate `marlin’s` capabilities by focusing on two case studies on the conservation and food production impacts of marine protected areas (MPAs): a coastal coral reef and a pelagic tuna fishery.
-In the coastal coral reef example, we explore how heterogeneity in species distributions and fleet preferences can affect distributional outcomes of MPAs.
-In the pelagic case study, we demonstrate how our model can be used to assess the climate resilience of different MPA design strategies, as well as the climate sensitivity of different fishing fleets.
-This paper demonstrates how intermediate complexity simulation of coupled bio-economic dynamics can help communities predict and potentially manage trade-offs between conservation, fisheries yields, and distributional outcomes resulting from spatial management.
+Abstract: Designing effective spatial management strategies is
+challenging because marine ecosystems are highly dynamic and opaque, and
+extractive entities such as fishing fleets respond endogenously to
+ecosystem changes in ways that depend upon ecological and policy
+context. We present a tool -- marlin -- to efficiently simulate the
+bio-economic dynamics of marine systems, designed to support both
+management and research. We demonstrate `marlin’s` capabilities by
+focusing on two case studies on the conservation and food production
+impacts of marine protected areas (MPAs): a coastal coral reef and a
+pelagic tuna fishery. In the coastal coral reef example, we explore how
+heterogeneity in species distributions and fleet preferences can affect
+distributional outcomes of MPAs. In the pelagic case study, we
+demonstrate how our model can be used to assess the climate resilience
+of different MPA design strategies, as well as the climate sensitivity
+of different fishing fleets. This paper demonstrates how intermediate
+complexity simulation of coupled bio-economic dynamics can help
+communities predict and potentially manage trade-offs between
+conservation, fisheries yields, and distributional outcomes resulting
+from spatial management.
 
@@
 
-Communities around the world are increasingly looking to spatial strategies for managing marine ecosystems .
-For example, the "30x30" movement calls for implementation of marine protected areas (MPAs), a form of spatial management, across 30% of the world's oceans by the year 2030 [@grorud-colvert2021 and references therein].
-However , designing spatial management strategies to achieve desired objectives -- which may include recovery of overfished species, increased food production and economic well being, and equitable distribution of benefits - is not straightforward.
-Different species have different resilience to fishing, distributions in space and time, and value to fishing fleets; fishing fleets themselves can have varying ranges of incentives and reactions to policy structures.
-These dynamics pose challenges even for single-species assessment and management strategies [@field2006a], which are only amplified when we consider management policies specifically designed around spatial features in the oceans.
-
-For example, policies such as MPAs designed for both conservation and food production must consider factors such as the optimal size and placement of a protected area network given a wide array of life history, species distributions, exploitation levels, fleet dynamics, and policy-dependent behavior [@reimer2017], all of which may change over time, particularly given the impacts of climate change.
-Efforts to effectively design spatial management strategies such as MPAs are further constrained by a lack of empirical evidence describing the size [@ovando2021a] and time required [@nickols2019] for MPAs to produce substantial effects across a range of social-ecological settings [@mcclanahan2021].
-
-Many models have been developed to support the theory and design of spatial management strategies (see @fulton2015 for a review of existing methods).
-However, they have tended to fall into one of either highly complex and tactical models designed for use in a specific location, or to be highly stylized representations intended to provide theoretical insights with less relevance for specific applications.
-This lack of accessible models suited for more tactical application presents a challenge to stakeholders charged with marine spatial planning exercises.
-It also presents a barrier to the scientific community, wherein comparisons across spatial management simulation studies are clouded by discrepancies in underlying model structures beyond differences in the core phenomenon in question.
-
-To help address this challenge, we present a novel bio-economic modeling tool called [`marlin`](https://danovando.github.io/marlin/) that allows for efficient simulation of spatio-temporal biological and economic dynamics.
-`marlin` allows users to simulate the impacts of marine management policies across a range of species targeted by various fleets across heterogeneous and dynamic seascapes.
-This model can help users seek Parteo-optimal solutions that produce win-wins or minimize tradeoffs across multiple management objectives, such as food production and conservation [@lester2013].
-
-Here we present the core methods and functionality of `marlin`, and demonstrate its use in two case-study applications; a coral reef system and a pelagic system.
-Our results show how interacting ecological, economic, and design attributes can affect the degree of benefits, costs, and tradeoffs between conservation and food production outcomes of MPAs.
-More broadly, this paper both demonstrates the functionality of our model and the critical need for considering sufficiently realistic coupled ecological and economic dynamics in policy evaluations.
+Communities around the world are increasingly looking to spatial
+strategies for managing marine ecosystems . For example, the "30x30"
+movement calls for implementation of marine protected areas (MPAs), a
+form of spatial management, across 30% of the world's oceans by the year
+2030 [@grorud-colvert2021 and references therein]. However , designing
+spatial management strategies to achieve desired objectives -- which may
+include recovery of overfished species, increased food production and
+economic well being, and equitable distribution of benefits - is not
+straightforward. Different species have different resilience to fishing,
+distributions in space and time, and value to fishing fleets; fishing
+fleets themselves can have varying ranges of incentives and reactions to
+policy structures. These dynamics pose challenges even for
+single-species assessment and management strategies [@field2006a], which
+are only amplified when we consider management policies specifically
+designed around spatial features in the oceans.
+
+For example, policies such as MPAs designed for both conservation and
+food production must consider factors such as the optimal size and
+placement of a protected area network given a wide array of life
+history, species distributions, exploitation levels, fleet dynamics, and
+policy-dependent behavior [@reimer2017], all of which may change over
+time, particularly given the impacts of climate change. Efforts to
+effectively design spatial management strategies such as MPAs are
+further constrained by a lack of empirical evidence describing the size
+[@ovando2021a] and time required [@nickols2019] for MPAs to produce
+substantial effects across a range of social-ecological settings
+[@mcclanahan2021].
+
+Many models have been developed to support the theory and design of
+spatial management strategies (see @fulton2015 for a review of existing
+methods). However, they have tended to be either highly complex and
+tactical models designed for use in a specific location, or to be
+extremely stylized representations intended to provide theoretical
+insights with less relevance for specific applications. This lack of
+accessible models suited for more tactical application presents a
+challenge to stakeholders charged with marine spatial planning
+exercises. It also presents a barrier to the scientific community,
+wherein comparisons across spatial management simulation studies are
+clouded by discrepancies in underlying model structures beyond
+differences in the core phenomenon in question.
+
+To help address this challenge, we present a novel bio-economic modeling
+tool called [`marlin`](https://danovando.github.io/marlin/) that allows
+for efficient simulation of spatio-temporal biological and economic
+dynamics. `marlin` allows users to simulate the impacts of marine
+management policies across a range of species targeted by various fleets
+across heterogeneous and dynamic seascapes. This model can help users
+seek Pareto-optimal solutions that produce win-wins or minimize
+tradeoffs across multiple management objectives, such as food production
+and conservation [@lester2013].
+
+Here we present the core methods and functionality of `marlin`, and
+demonstrate its use in two case-study applications; a coral reef system
+and a pelagic system. Our results show how interacting ecological,
+economic, and design attributes can affect the degree of benefits,
+costs, and tradeoffs between conservation and food production outcomes
+of MPAs. More broadly, this paper both demonstrates the functionality of
+our model and the critical need for considering sufficiently realistic
+coupled ecological and economic dynamics in policy evaluations.
 
@@
 
-`marlin` follows a "models of intermediate complexity" philosophy [@plagányi2014] targeted specifically towards capturing heterogeneous habitats, movement dynamics, and the simultaneous impacts of fishing fleets across multiple species (frequently called "technical interactions").
-In its general form, `marlin` simulates the behavior of populations of spatially explicit biologically independent age-structured animal populations affected by one or more fishing fleets in time steps and a spatial resolution specified by the user.
-
-The front end of the software package accompanying this paper, [`marlin`](https://danovando.github.io/marlin/), is written in R [@rcoreteam2021] to facilitate use, while the underlying population model is written in C++, integrated through the `Rcpp` package [@eddelbuettel2018], for speed.
-Users on standard computers should be able to simulate one age-structured population distributed across a 100 patch system for 20 years in fractions of a second.
-
-While `marlin` does not simulate species interactions, fishing fleets in `marlin` are capable of targeting and affecting multiple species simultaneously.
-This allows the model to simulate processes such as fisheries bycatch or effort displacement in a way that accounts for how fleet behavior might affect multiple species in the system.
-The ability to track the impacts of fishing fleets across multiple species simultaneously is particularly important as very few fishing fleets are truly single species.
-
-We define a few commonly used terms here.
-*Yields* refers to the volume of catch from fishing activity.
-*Spawning stock biomass* refers to the total biomass of reproductively mature fish in the water, a function of numbers, weight, and sexual maturity at age.
-We measure the size of the population by th ratio of the *spawning stock biomass* (SSB) in a given time step relative to *unfished spawning stock biomass* (SSB0), SSB/SSB0.
-So, a SSB/SSB0 value of 1 means that the population is unfished, a SSB/SSB0 value of 0 means the population is extinct.
-
-As a demonstration of the use of this modeling tool, we explore two general situations
-
-1.  Trade-offs and distributional outcomes for food production and multi-species biomass in a coastal coral-reef fishery
-2.  Implications of climate-driven range shifts for MPA design in a pelagic fishery
-
-We chose these two examples to illustrate the use of `marlin` in contrasting systems in which spatial management strategies such as MPAs are increasingly considered.
-
-Below we provide a summary of the `marlin` model, as well as details of the case studies.
-We note that `marlin` is a complex software package with many options to control detailed behaviors.
-Given that these specific features will evolve over time, we do not describe in detail every option of the model here.
-Rather, we focus on the equations and methods describing the core foundations of our tool for efficiently simulating bio-economic dynamics in aquatic systems.
+`marlin` follows a "models of intermediate complexity" philosophy
+[@plagányi2014] targeted specifically towards capturing heterogeneous
+habitats, movement dynamics, and the simultaneous impacts of fishing
+fleets across multiple species (frequently called "technical
+interactions"). In its general form, `marlin` simulates the behavior of
+populations of spatially explicit biologically independent
+age-structured animal populations affected by one or more fishing fleets
+in time steps and a spatial resolution specified by the user.
+
+The front end of the software package accompanying this paper,
+[`marlin`](https://danovando.github.io/marlin/), is written in R
+[@rcoreteam2021] to facilitate use, while the underlying population
+model is written in C++, integrated through the `Rcpp` package
+[@eddelbuettel2018], for speed. Users on standard computers should be
+able to simulate one age-structured population distributed across a 100
+patch system for 20 years in fractions of a second.
+
+While `marlin` does not simulate species interactions, fishing fleets in
+`marlin` are capable of targeting and affecting multiple species
+simultaneously. This allows the model to simulate processes such as
+fisheries bycatch or effort displacement in a way that accounts for how
+fleet behavior might affect multiple species in the system. The ability
+to track the impacts of fishing fleets across multiple species
+simultaneously is particularly important as very few fishing fleets are
+truly single species.
+
+We define a few commonly used terms here. *Yields* refers to the volume
+of catch from fishing activity. *Spawning stock biomass* refers to the
+total biomass of reproductively mature fish in the water, a function of
+numbers, weight, and sexual maturity at age. We measure the size of the
+population by th ratio of the *spawning stock biomass* (SSB) in a given
+time step relative to *unfished spawning stock biomass* (SSB0),
+SSB/SSB0. So, a SSB/SSB0 value of 1 means that the population is
+unfished, a SSB/SSB0 value of 0 means the population is extinct.
+
+As a demonstration of the use of this modeling tool, we explore two
+general situations
+
+1.  Trade-offs and distributional outcomes for food production and
+    multi-species biomass in a coastal coral-reef fishery
+2.  Implications of climate-driven range shifts for MPA design in a
+    pelagic fishery
+
+We chose these two examples to illustrate the use of `marlin` in
+contrasting systems in which spatial management strategies such as MPAs
+are increasingly considered.
+
+Below we provide a summary of the `marlin` model, as well as details of
+the case studies. We note that `marlin` is a complex software package
+with many options to control detailed behaviors. Given that these
+specific features will evolve over time, we do not describe in detail
+every option of the model here. Rather, we focus on the equations and
+methods describing the core foundations of our tool for efficiently
+simulating bio-economic dynamics in aquatic systems.
 
@@
 
-`marlin` simulates the dynamics of one or more species, currently best representing fish-like species, using age-structured population dynamics.
-Each time step, species move throughout the simulated seascape using both diffusion and "taxis" (active movement towards preferred habitat), experience natural and potentially fishing mortality, and if in a spawning season spawn using one of the possible forms of density dependence implemented in the model, with the potential to include auto-correlated stochastic deviations in the amount of offspring produced, generally called "recruitment deviates".
-
-These species can be caught by fishing fleets.
-A fishing fleet is defined in `marlin` by a set of fishers that can have different fishing skill, prices, and selectivity for individual species (each of which we denote as a *metier*).
-For example, both a longline and purse seine fleet may capture bigeye (*Thunnus obesus*) and skipjack tunas, but the longline fleet can be made much more likely to catch bigeye than skipjack, and *vice versa*.
-Each fishing fleet distributes its fishing effort in space according to a specified spatial allocation rule, for example in proportion to profit-per-unit-effort, conditional on management policies in places such as quotas and/or the presence of any spatial restrictions such as MPAs.
-
-Unconstrained by management, the total amount of effort exerted by each fleet can follow one of two dynamics: open access or constant effort.
-Under constant effort, the total amount of effort of each fleet is fixed over time, with the possible exception of attrition due to MPA placement.
-Under open access, the total amount of effort in a given time step is a function of the profitability of the fleet in the previous time step, evolving until a bionomic equilibrium of zero total profits is reached.
-Profitability is a function of the volume and price of each species caught, as well as the cost of the total amount of fishing effort per fleet and the travel costs as a function of distance from a port.
-
-The dynamics of each fleet can be modified by management in a variety of ways.
-Managers can impose size limits for individual species within each fleet.
-They can also impose total catch quotas for each species in each time step.
-When quotas are activated, if the total catch across all fleets for a given species under the model's effort dynamics would exceed the allowable quota for that species, the total effort for each fleet is decreased proportional to its contribution to the total catch until the quota is satisfied.
-As an alternative to catch quotas, managers can set an effort cap per fleet, which prevents effort from exceeding a given amount under open access dynamics, though effort may be reduced below this cap if required by the profit equation (i.e.
-the fleet can choose to fish less than the quota or effort cap).
-
-`marlin` also allows users to specify closed fishing seasons for one or more species in the system.
-When all species in a simulation share the same closed seasons, no fishing occurs in those seasons.
-When not all species share the same closed seasons, fishing still occurs in the closed seasons, but the fleet avoids catching the species for which fishing is closed in a given season, assuming no discard mortality.
-When closed seasons are in effect, the amount of effort in the next open season is a function of the effort and profits in the last open fishing season.
-Lastly, managers can specify locations of no-take MPAs, which could in theory change in size and location.
-When MPAs are implemented, fishing effort that used to operate inside the MPAs can either leave the fishery, or be redistributed outside the MPA (the default behavior).
-
-Users can use the model to explore general properties of spatially explicit social-ecological systems, or can parameterize it to explore key uncertainties of more tactical applications.
-As an example of the later case, we provide two case studies here designed to reflect the kinds of questions stakeholders might face if designing an MPA network for a coastal coral reef or a pelagic system.
+`marlin` simulates the dynamics of one or more species, currently best
+representing fish-like species, using age-structured population
+dynamics. Each time step, species move throughout the simulated seascape
+using both diffusion and "taxis" (active movement towards preferred
+habitat), experience natural and potentially fishing mortality, and if
+in a spawning season spawn using one of the possible forms of density
+dependence implemented in the model, with the potential to include
+auto-correlated stochastic deviations in the amount of offspring
+produced, generally called "recruitment deviates".
+
+These species can be caught by fishing fleets. A fishing fleet is
+defined in `marlin` by a set of fishers that have the same fishing
+skill, prices, and selectivity for individual species (each of which we
+denote as a *metier*). For example, both a longline and purse seine
+fleet may capture bigeye (*Thunnus obesus*) and skipjack tunas, but the
+longline fleet can be made much more likely to catch bigeye than
+skipjack, and *vice versa*. Each fishing fleet distributes its fishing
+effort in space according to a specified spatial allocation rule (see
+*Spatial Allocation of Effort* ), for example in proportion to
+profit-per-unit-effort, conditional on management policies in places
+such as quotas and/or the presence of any spatial restrictions such as
+MPAs.
+
+Unconstrained by management, the total amount of effort exerted by each
+fleet can follow one of two dynamics: open access or constant effort
+(see *Calculating Total Effort*). Under constant effort, the total
+amount of effort of each fleet is fixed over time, with the possible
+exception of attrition due to MPA placement. Under open access, the
+total amount of effort in a given time step is a function of the
+profitability of the fleet in the previous time step, evolving until a
+bionomic equilibrium of zero total profits is reached. Profitability is
+a function of the volume and price of each species caught, as well as
+the cost of the total amount of fishing effort per fleet and the travel
+costs as a function of distance from a port (see *Fleet Dynamics*).
+
+The dynamics of each fleet can be modified by management in a variety of
+ways. Managers can impose size limits for individual species within each
+fleet. They can also impose total catch quotas for each species in each
+time step. When quotas are activated, if the total catch across all
+fleets for a given species under the model's effort dynamics would
+exceed the allowable quota for that species, the total effort for each
+fleet is decreased proportional to its contribution to the total catch
+until the quota is satisfied. As an alternative to catch quotas,
+managers can set an effort cap per fleet, which prevents effort from
+exceeding a given amount under open access dynamics, though effort may
+be reduced below this cap if required by the profit equation (i.e. the
+fleet can choose to fish less than the quota or effort cap).
+
+`marlin` also allows users to specify closed fishing seasons for one or
+more species in the system. When all species in a simulation share the
+same closed seasons, no fishing occurs in those seasons. When not all
+species share the same closed seasons, fishing still occurs in the
+closed seasons, but the fleet avoids catching the species for which
+fishing is closed in a given season, assuming no discard mortality. When
+closed seasons are in effect, the amount of effort in the next open
+season is a function of the effort and profits in the last open fishing
+season. Lastly, managers can specify locations of no-take MPAs, which
+could in theory change in size and location. When MPAs are implemented,
+fishing effort that used to operate inside the MPAs can either leave the
+fishery, or be redistributed outside the MPA (the default behavior).
+
+Users can use the model to explore general properties of spatially
+explicit social-ecological systems, or can parameterize it to explore
+key uncertainties of more tactical applications. As an example of the
+later case, we provide two case studies here designed to reflect the
+kinds of questions stakeholders might face if designing an MPA network
+for a coastal coral reef or a pelagic system.
 
@@
 
-Each fishing fleet (*f*) generates catches or yield (*Y*), revenues (*R*), and costs (*C*) from fishing individual species (*s*) that it targets in a given time step (*t*) and patch (*p*).
-The fleet then makes decisions around fishing locations and intensity, subject to regulatory constraints, based on the total profitability across all species in space and time.
-
-Revenues for each fleet in a time step are a function of the total amount of each species caught and its price ($\Pi$).
-The amount caught is a function of the selectivity at age of each species for each fleet ($\alpha$), the fishing efficiency of the fleet for that that species (also called "catchability", *q*), the amount of fishing effort of the fleet in question in that patch in a given time step (*E*), and the total instantaneous fishing mortality (*u*) at age (*a*) (including all other fleets) for that particular species in that patch and time step.
+Each fishing fleet (*f*) generates catches or yield (*Y*), revenues
+(*R*), and costs (*C*) from fishing individual species (*s*) that it
+targets in a given time step (*t*) and patch (*p*). The fleet then makes
+decisions around fishing locations and intensity, subject to regulatory
+constraints, based on the total profitability across all species in
+space and time.
+
+Revenues for each fleet in a time step are a function of the total
+amount of each species caught and its price ($\Pi$). The amount caught
+is a function of the selectivity at age of each species for each fleet
+($\alpha$), the fishing efficiency of the fleet for that that species
+(also called "catchability", *q*), the amount of fishing effort of the
+fleet in question in that patch in a given time step (*E*), and the
+total instantaneous fishing mortality (*u*) at age (*a*) (including all
+other fleets) for that particular species in that patch and time step.
 
@@
 
-Total catches or yield *Y* are calculated through the Baranov equation [@baranov1918], which appropriately accounts for the total instantaneous mortality (both fishing and natural.
-*z*) and divides the total amount of the population biomass (*b*) killed between natural (*m*) and fishing (*u*) sources, with the amount of biomass killed through fishing called "yield".
+Total catches or yield *Y* are calculated through the Baranov equation
+[@baranov1918], which appropriately accounts for the total instantaneous
+mortality (both fishing and natural. *z*) and divides the total amount
+of the population biomass (*b*) killed between natural (*m*) and fishing
+(*u*) sources, with the amount of biomass killed through fishing called
+"yield".
 
@@
 
-Selectivity at age $\alpha$ is modeled through either a logistic form or a dome-shaped form.
-The logistic form is based on the lengths *l* at which 50% of individuals are selected by the fishing gear, $l^{sel}$, and $\delta$ which is the difference between the length at 50% selectivity and the length at 95% selectivity.
+Selectivity at age $\alpha$ is modeled through either a logistic form or
+a dome-shaped form. The logistic form is based on the lengths *l* at
+which 50% of individuals are selected by the fishing gear, $l^{sel}$,
+and $\delta$ which is the difference between the length at 50%
+selectivity and the length at 95% selectivity.
 
@@
 
-We approximate the dome-shaped form as a normal distribution with mean $l^{sel}$ and standard deviation $\sigma$. The normal density function is re-scaled such that the selectivity is 1 at $l^{sel}$.
+We approximate the dome-shaped form as a normal distribution with mean
+$l^{sel}$ and standard deviation $\sigma$. The normal density function
+is re-scaled such that the selectivity is 1 at $l^{sel}$.
 
@@
 
-Lastly *z* is total mortality, the sum of fishing mortality and natural mortality (*m*).
+Lastly *z* is total mortality, the sum of fishing mortality and natural
+mortality (*m*).
 
@@
 
-This means that each species experiences a total mortality at age in a time step in a patch, individual fractions of which are portioned off as catches and subsequently revenues for each fleet.
-Given revenues *R*, we then calculate profits $\phi$ based on revenues and costs.
-Costs *C* are calculated as a function of base costs per unit effort ($\gamma$) as well as potential additional costs per unit effort of fishing in particular patches ($\eta$) for fleet *f*.
-$\beta$ allows for the cost per unit effort effort to scale non-linearly.
-Travel costs ($\eta$) are calculated based on the Euclidean distance of each patch to the nearest port of a given fleet, and users can specify any number from zero to the number of patches of port locations.
-When no ports are specified, travel costs are zero.
+This means that each species experiences a total mortality at age in a
+time step in a patch, individual fractions of which are portioned off as
+catches and subsequently revenues for each fleet. Given revenues *R*, we
+then calculate profits $\phi$ based on revenues and costs. Costs *C* are
+calculated as a function of base costs per unit effort ($\gamma$) as
+well as potential additional costs per unit effort of fishing in
+particular patches ($\eta$) for fleet *f*. $\beta$ allows for the cost
+per unit effort effort to scale non-linearly. Travel costs ($\eta$) are
+calculated based on the Euclidean distance of each patch to the nearest
+port of a given fleet, and users can specify any number from zero to the
+number of patches of port locations. When no ports are specified, travel
+costs are zero.
 
@@
 
-`marlin` allows for two general modes of effort. The simplest is "constant effort", in which the total effort of each fishing fleet remains constant over time.
-
-The more complex option is "open access".
-Under open access, the total amount of effort in a given time step for fleet *f* is a function of the profitability of that fleet in the proceeding time steps in which fishing was open, where $\theta$ controls the responsiveness of fleet *f* to the log of the ratio of total revenues *R* to total costs *C*, and approximates the proportional change in effort in response to a one unit changes in the log revenue to cost ratio.
+`marlin` allows for two general modes of effort. The simplest is
+"constant effort", in which the total effort of each fishing fleet
+remains constant over time.
+
+The more complex option is "open access". Under open access, the total
+amount of effort in a given time step for fleet *f* is a function of the
+profitability of that fleet in the proceeding time steps in which
+fishing was open, where $\theta$ controls the responsiveness of fleet
+*f* to the log of the ratio of total revenues *R* to total costs *C*,
+and approximates the proportional change in effort in response to a one
+unit changes in the log revenue to cost ratio.
 
@@
 
-This is essentially a Gompertz model for fishing effort, that has been used for other theoretical studies of fishing effort dynamics [@thorson2013].
+This is essentially a Gompertz model for fishing effort, that has been
+used for other theoretical studies of fishing effort dynamics
+[@thorson2013].
 
@@
 
-The degree of fishing pressure exerted by a given set of fishing fleets on each species is a function of a range of parameters including the total amount of effort (*E*), the selectivity ogives ($\alpha$,the curve specifying the proportion of individuals at a given age or size that will be caught by a unit of fishing mortality), fishing cost (*C*), the spatial distribution of the species affected by the fleet, the relative prices across species ($\Pi$), and the catchability coefficient of each fleet for each species ($q$).
-`marlin` provides a tuning option to help users achieve desired biological outcomes from their fleets.
-
-Users can tune their fleets in one of two ways.
-First, they can specify a target exploitation rate *u* for each species in their simulation.
-Taking all the other parameters of the model as given, `marlin` then adjusts the catchability coefficients $q_{f,s}$ for each fleet *f* and species *s* such that the equilibrium exploitation rate for each species matches the desired level.
-Second, they can specify a target total spawning stock biomass under fishing divided by total unfished spawning stock biomass, and the model will adjust the catchability coefficients $q_{f,s}$ for each fleet *f* and species *s* such that the equilibrium ratio of fished to unfished spawning biomass for each species matches the desired level.
-
-Users can also tune the fleet dynamics by specifying the ratio of costs to revenues.
-Price data for use in the model can be obtained relatively simply through literature reviews, market surveys, or local experts, so long as they are converted to the same units of biomass used in the model.
-However, cost parameters are more complicated, as translating say cost per day of fishing into the same representation of fishing effort used in the model is not straight forward to accomplish.
-As an alternative, users can specify an equilibrium cost-to-revenue ratio for the fleet, essentially the profit margins of the fleet in question, and `marlin` will tune the cost parameter to achieve this desired cost to revenue ratio given the other parameters in the model (e.g.
-price and biomass of the target species).
-So for example, while users may not have access to exact cost parameters, they may have a general sense that when operating under ideal conditions a given fleet has a profit margin of 10%, i.e.
-the cost to revenue ratio is 90%.
-They can input the cost to revenue ratio as 0.9 then and allow `marlin` to tune the cost parameters accordingly.
+The degree of fishing pressure exerted by a given set of fishing fleets
+on each species is a function of a range of parameters including the
+total amount of effort (*E*), the selectivity ogives ($\alpha$,the curve
+specifying the proportion of individuals at a given age or size that
+will be caught by a unit of fishing mortality), fishing cost (*C*), the
+spatial distribution of the species affected by the fleet, the relative
+prices across species ($\Pi$), and the catchability coefficient of each
+fleet for each species ($q$). `marlin` provides a tuning option to help
+users achieve desired biological outcomes from their fleets.
+
+Users can tune their fleets in one of two ways. First, they can specify
+a target exploitation rate *u* for each species in their simulation.
+Taking all the other parameters of the model as given, `marlin` then
+adjusts the catchability coefficients $q_{f,s}$ for each fleet *f* and
+species *s* such that the equilibrium exploitation rate for each species
+matches the desired level. Second, they can specify a target total
+spawning stock biomass under fishing divided by total unfished spawning
+stock biomass, and the model will adjust the catchability coefficients
+$q_{f,s}$ for each fleet *f* and species *s* such that the equilibrium
+ratio of fished to unfished spawning biomass for each species matches
+the desired level.
+
+Users can also tune the fleet dynamics by specifying the ratio of costs
+to revenues. Price data for use in the model can be obtained relatively
+simply through literature reviews, market surveys, or local experts, so
+long as they are converted to the same units of biomass used in the
+model. However, cost parameters are more complicated, as translating say
+cost per day of fishing into the same representation of fishing effort
+used in the model is not straight forward to accomplish. As an
+alternative, users can specify an equilibrium cost-to-revenue ratio for
+the fleet, essentially the profit margins of the fleet in question, and
+`marlin` will tune the cost parameter to achieve this desired cost to
+revenue ratio given the other parameters in the model (e.g. price and
+biomass of the target species). So for example, while users may not have
+access to exact cost parameters, they may have a general sense that when
+operating under ideal conditions a given fleet has a profit margin of
+10%, i.e. the cost to revenue ratio is 90%. They can input the cost to
+revenue ratio as 0.9 then and allow `marlin` to tune the cost parameters
+accordingly.
 
@@
 
-Each fishing fleet decides how to allocate its effort in space based on one of four possible spatial allocation strategies.
-The ideal free distribution (IFD) is the standard method for distributing fishing fleets in spaces (see @gillis2003 and references therein).
-However, analytical solutions to the IFD present a number of complications for our model.
-Primarily, our model is capable of tracking multiple fleets at once.
-This means that the IFD for a single fleet would commonly be modeled as a Nash-equilibrium based on the actions of each of the individual fleet conditional on the actions of all other fleets.
-While possible to implement, this would slow down our model runs to the point of practically preventing large-scale evaluation of the spatial management policies.
-
-As such we explored a series of "next best" fleet distribution algorithms.
-While not the IFD in any individual time step, over time they start to approximate the IFD, as each assumes that fleets base their decisions for the current time step on the outcomes in the prior timestep, meaning that the impacts of the actions of other fishing fleets are somewhat internalized.
-The four possible fleet distribution algorithms are
-
-1.  Revenue per unit effort (RPUE): The fleet distributes itself in space based on the realized revenue per unit effort in each patch in the preceding time step
-2.  Revenue: The fleet distributes itself in space based on the realized total revenue in each patch in the preceding time step
-3.  Profit per unit effort (PPUE): The fleet distributes itself in space based on the realized profit per unit effort in each patch in the preceding time step
-4.  Profit: The fleet distributes itself in space based on profit in each patch in the preceding time step
-
-Revenue based spatial distribution is not likely to be very realistic; in general we would expect fishing fleets to respond to profits on some core level.
-However, due to the complexity of parameterizing cost functions, fleet dynamics are often evaluated based on yield or revenue alone, and so we include those scenarios here to allow users to evaluate the potential implications of this choice.
-The decision on whether to allocate the fleet based on absolute or relative (per unit effort) metrics is more complex.
-When effort represents the actions of separate and individual fishing actors (e.g.
-independent fishing vessels), a per-unit-effort strategy, in which fishers distribute themselves based on the expected catch of their individual efforts, may be more realistic [@hilborn1987].
-Conversely, a system defined by a sole owner seeking to maximize total profits might be better represented by a fleet model based on total profits
+Each fishing fleet decides how to allocate its effort in space based on
+one of four possible spatial allocation strategies. The ideal free
+distribution (IFD) is the standard method for distributing fishing
+fleets in spaces (see @gillis2003 and references therein). However,
+analytical solutions to the IFD present a number of complications for
+our model. Primarily, our model is capable of tracking multiple fleets
+at once. This means that the IFD for a single fleet would commonly be
+modeled as a Nash-equilibrium based on the actions of each of the
+individual fleet conditional on the actions of all other fleets. While
+possible to implement, this would slow down our model runs to the point
+of practically preventing large-scale evaluation of the spatial
+management policies.
+
+As such we explored a series of "next best" fleet distribution
+algorithms. While not the IFD in any individual time step, over time
+they start to approximate the IFD, as each assumes that fleets base
+their decisions for the current time step on the outcomes in the prior
+timestep, meaning that the impacts of the actions of other fishing
+fleets are somewhat internalized. The four possible fleet distribution
+algorithms are
+
+1.  Revenue per unit effort (RPUE): The fleet distributes itself in
+    space based on the realized revenue per unit effort in each patch in
+    the preceding time step
+2.  Revenue: The fleet distributes itself in space based on the realized
+    total revenue in each patch in the preceding time step
+3.  Profit per unit effort (PPUE): The fleet distributes itself in space
+    based on the realized profit per unit effort in each patch in the
+    preceding time step
+4.  Profit: The fleet distributes itself in space based on profit in
+    each patch in the preceding time step
+
+Revenue based spatial distribution is not likely to be very realistic;
+in general we would expect fishing fleets to respond to profits on some
+core level. However, due to the complexity of parameterizing cost
+functions, fleet dynamics are often evaluated based on yield or revenue
+alone, and so we include those scenarios here to allow users to evaluate
+the potential implications of this choice. The decision on whether to
+allocate the fleet based on absolute or relative (per unit effort)
+metrics is more complex. When effort represents the actions of separate
+and individual fishing actors (e.g. independent fishing vessels), a
+per-unit-effort strategy, in which fishers distribute themselves based
+on the expected catch of their individual efforts, may be more realistic
+[@hilborn1987]. Conversely, a system defined by a sole owner seeking to
+maximize total profits might be better represented by a fleet model
+based on total profits
 
@@
 
-The underlying population model used is an age structured single-species model in the manner of @ovando2021a.
-The population model requires many parameters.
-However, if the user supplies either a scientific or common name for the species in question (scientific preferred), the model will supply default values for that species based on the values reported in FishLife [internet connection required, @thorson2020].
-FishLife provides estimates of core life history parameters for fish species based on a model integrating published values and evolutionary connections.
-FishLife provides reasonable default values for given species, but these default values should be not be taken as definitive and users should input best available parameters specific to the stock in question if they wish to best represent the dynamics of their specific system, and should check the plausibility of the default values prior to use.
-All of the life history values can be replaced with user-supplied values as needed and should be for more tactical applications; the default values simply ensure that reasonable and internally consistent values for the species in question are supplied as a starting point.
+The underlying population model used is an age structured single-species
+model in the manner of @ovando2021a. The population model requires many
+parameters. However, if the user supplies either a scientific or common
+name for the species in question (scientific preferred), the model will
+supply default values for that species based on the values reported in
+FishLife [internet connection required, @thorson2020]. FishLife provides
+estimates of core life history parameters for fish species based on a
+model integrating published values and evolutionary connections.
+FishLife provides reasonable default values for given species, but these
+default values should be not be taken as definitive and users should
+input best available parameters specific to the stock in question if
+they wish to best represent the dynamics of their specific system, and
+should check the plausibility of the default values prior to use. All of
+the life history values can be replaced with user-supplied values as
+needed and should be for more tactical applications; the default values
+simply ensure that reasonable and internally consistent values for the
+species in question are supplied as a starting point.
 
@@
 
-`marlin`'s movement dynamics are based on a continuous-time Markov chain (CTMC), as described in @thorson2021a.
-Within this framework, the model allows for movement to be broken down into three components of *advection* (follow currents), and *taxis* (active movement towards preferred habitat) and *diffusion* (essentially remaining variation in movement not explained by advection or taxis).
-For now, `marlin` focuses just on the diffusion and taxis components of this model, assuming that advection is zero, though future extensions could easily incorporate advection vectors from oceanographic models.
-In this way, `marlin` allows users to run anything from a simple Gaussian dispersal kernel up to a system governed by species that passively diffuse out from a core habitat defined by a dynamic thermal range.
-
-We provide a brief overview of the the general CTMC method here, see @thorson2021a for a detailed description.
-Under this framework, movement from each patch to each other patch in the system in a given timestep *t* for life stage *a* of species *s* is defined by a movement matrix $\boldsymbol{M}_{t,s,a}$.
-$\boldsymbol{M}_{t,s,a}$ is calculated as the matrix exponent of the sum of the diffusion $\boldsymbol{D}$ and taxis $\boldsymbol{\tau}$ matrices scaled by the width of the time step $\Delta_{t}$ and the length of the edge of each patch $\Delta_d$ that the model is running under.
-So for example, if we are running a model on a monthly timestep, and we define the units of time in years, then $\Delta_{t} = 1/12$.
-Similarly, if we are running a model where one patch equals 100km^2^, then $\Delta_d = 100km$.
-This parameterization allows users to set the effective area of the spatial domain through two avenues; the number of patches, which effectively scales the resolution of the model, and the area of each patch, which scales the spatial extent of the simulation.
+`marlin`'s movement dynamics are based on a continuous-time Markov chain
+(CTMC), as described in @thorson2021a. Within this framework, the model
+allows for movement to be broken down into three components of
+*advection* (follow currents), and *taxis* (active movement towards
+preferred habitat) and *diffusion* (essentially remaining variation in
+movement not explained by advection or taxis). For now, `marlin` focuses
+just on the diffusion and taxis components of this model, assuming that
+advection is zero, though future extensions could easily incorporate
+advection vectors from oceanographic models. In this way, `marlin`
+allows users to run anything from a simple Gaussian dispersal kernel up
+to a system governed by species that passively diffuse out from a core
+habitat defined by a dynamic thermal range.
+
+We provide a brief overview of the the general CTMC method here, see
+@thorson2021a for a detailed description. Under this framework, movement
+from each patch to each other patch in the system in a given timestep
+*t* for life stage *a* of species *s* is defined by a movement matrix
+$\boldsymbol{M}_{t,s,a}$. $\boldsymbol{M}_{t,s,a}$ is calculated as the
+matrix exponent of the sum of the diffusion $\boldsymbol{D}$ and taxis
+$\boldsymbol{\tau}$ matrices scaled by the width of the time step
+$\Delta_{t}$ and the length of the edge of each patch $\Delta_d$ that
+the model is running under. So for example, if we are running a model on
+a monthly timestep, and we define the units of time in years, then
+$\Delta_{t} = 1/12$. Similarly, if we are running a model where one
+patch equals 100km^2^, then $\Delta_d = 100km$. This parameterization
+allows users to set the effective area of the spatial domain through two
+avenues; the number of patches, which effectively scales the resolution
+of the model, and the area of each patch, which scales the spatial
+extent of the simulation.
 
@@
 
-The individual components of the diffusion matrix are filled based on an adjacency matrix, which defines whether two patches are both adjacent and water (as opposed to land or another physical barrier), and a diffusion rate $D$ defined in units of length of the side of a patch per unit of time.
-For example, if we are defining the time units as years and the distance units as kilometers, for a tuna $D$ might be 1,000 $\frac{KM}{Year}$.
-The $\Delta_{t}$ and $\Delta_{d}$ parameters then translate the movement rate $D$ to match the time step and patch size used in a simulation.
-This "scale free" parameterization means that appropriate value of $D$ can be identified for a species and then set, regardless of the time step or patch size used in the simulation model itself.
-These parameters are then used to calculate the value of a particular entry the diffusion matrix (*d*).
+The individual components of the diffusion matrix are filled based on an
+adjacency matrix, which defines whether two patches are both adjacent
+and water (as opposed to land or another physical barrier), and a
+diffusion rate $D$ defined in units of length of the side of a patch per
+unit of time. For example, if we are defining the time units as years
+and the distance units as kilometers, for a tuna $D$ might be 1,000
+$\frac{KM}{Year}$. The $\Delta_{t}$ and $\Delta_{d}$ parameters then
+translate the movement rate $D$ to match the time step and patch size
+used in a simulation. This "scale free" parameterization means that
+appropriate value of $D$ can be identified for a species and then set,
+regardless of the time step or patch size used in the simulation model
+itself. These parameters are then used to calculate the value of a
+particular entry the diffusion matrix (*d*).
 
@@
 
-The taxis matrix $\boldsymbol{\tau}$ is built off of the difference in log habitat quality between each patch (the habitat gradient), the adjacency matrix, and temporal and spatial resolution of the model ($\Delta_{t}$ and $\Delta_{p}$), as well as a scalar term $\lambda$.
-$\lambda$ scales the magnitude of the habitat gradient between patches relative to the diffusion parameter $D$.
-When the diffusion and taxis matrices are estimated jointly from actual data in the manner of @thorson2021a, $\lambda$ should be set to zero.
-However, when users are manually creating both the diffusion and taxis matrices for simulation, the relative strength of diffusion to taxis must be defined by the user.
-Setting higher values of $\lambda$ allows users to simulate a scenario where taxis is much more significant than diffusion.
+The taxis matrix $\boldsymbol{\tau}$ is built off of the difference in
+log habitat quality between each patch (the habitat gradient), the
+adjacency matrix, and temporal and spatial resolution of the model
+($\Delta_{t}$ and $\Delta_{p}$), as well as a scalar term $\lambda$.
+$\lambda$ scales the magnitude of the habitat gradient between patches
+relative to the diffusion parameter $D$. When the diffusion and taxis
+matrices are estimated jointly from actual data in the manner of
+@thorson2021a, $\lambda$ should be set to zero. However, when users are
+manually creating both the diffusion and taxis matrices for simulation,
+the relative strength of diffusion to taxis must be defined by the user.
+Setting higher values of $\lambda$ allows users to simulate a scenario
+where taxis is much more significant than diffusion.
 
@@
 
-For both the diffusion and taxis matrix, we allow for the inclusion of physical barriers to movement (i.e.
-land).
-Pairs of patches that are adjacent but in which one or both patches are a barrier to movement are set as non-adjacent.
-The CTMC model then produces movement dynamics that move around barriers rather than over them.
-This is different than setting habitat in a patch to zero; an animal will not preferentially move towards a patch with zero habitat all else being equal, but can move over a patch with zero habitat towards another better patch.
-An example might be a species that lives and feeds on coastal habitats transiting through a patch of open ocean to reach a new feeding ground.
-In contrast, a physical barrier represents something like a peninsula that an animal must move around to transit from one side to another.
-
-While this CTMC approach to movement simulation is to date not commonly seen in the marine modeling literature, it has numerous advantages that warrant its seeming complexity.
-First, when only diffusion is present, the model will generalize to the familiar dynamics of a Gaussian dispersal kernel.
-However, the time component *t* in @eq-movement allows the movement model to be run on arbitrarily fine time steps.
-Second, the taxis model allow for more active habitat choices by species, allowing us to simulate both preferences of species in space and time efficiently.
-Lastly, this form has the advantage that the parameters are directly estimable from real data.
-So, if provided with say spatial abundance data and a tagging study, users can estimate the diffusion and taxis movement parameters in the same manner as @thorson2021a, and then pass their estimated parameters to `marlin` for simulation.
+For both the diffusion and taxis matrices, we allow for the inclusion of
+physical barriers to movement (i.e. land). Pairs of patches that are
+adjacent but in which one or both patches are a barrier to movement are
+set as non-adjacent. The CTMC model then produces movement dynamics that
+move around barriers rather than over them. This is different than
+setting habitat in a patch to zero; an animal will not preferentially
+move towards a patch with zero habitat all else being equal, but can
+move over a patch with zero habitat towards another better patch. An
+example might be a species that lives and feeds on coastal habitats
+transiting through a patch of open ocean to reach a new feeding ground.
+In contrast, a physical barrier represents something like a peninsula
+that an animal must move around to transit from one side to another.
+
+While this CTMC approach to movement simulation is to date not commonly
+seen in the marine modeling literature, it has numerous advantages that
+warrant its seeming complexity. First, when only diffusion is present,
+the model will generalize to the familiar dynamics of a Gaussian
+dispersal kernel. However, the time component *t* in @eq-movement allows
+the movement model to be run on arbitrarily fine time steps. Second, the
+taxis model allow for more active habitat choices by species, allowing
+us to simulate preferences of species in space and time efficiently.
+Lastly, this form has the advantage that the parameters are directly
+estimable from real data. So, if provided with for example spatial
+abundance data and a tagging study, users can estimate the diffusion and
+taxis movement parameters in the same manner as @thorson2021a, and then
+pass their estimated parameters to `marlin` for simulation.
 
@@
 
-For the population model, numbers *n* at time *t* for age *a* are a function of growth, death, and recruitment
+For the population model, numbers *n* at time *t* for age *a* are a
+function of growth, death, and recruitment
 
@@
 
-where *BH* is the Beverton-Holt recruitment function [@beverton1957] and *ssb* is spawning-stock-biomass.
-Per convention, the model allows for a "plus group", wherein rather than tracking numbers of every possible age, individuals greater than or equal to a given maximum age are grouped together.
-
-Spawning stock biomass *ssb* is calculated by converting age to mean length, calculating weight at age, maturity at age, and then calculating spawning stock biomass as the sum of spawning potential at age in a given time step, taking into account the potential for hyperallometry.
-Age is converted to length through the von Bertalanffy growth equation given parameters asymptotic length ($l_\infty$), growth (*k*) and theoretical age at length zero ($a0$) assuming log-normally distributed variation *u* in the length at age with CV $\sigma_s$.
+where *BH* is the Beverton-Holt recruitment function [@beverton1957] and
+*ssb* is spawning-stock-biomass. Per convention, the model allows for a
+"plus group", wherein rather than tracking numbers of every possible
+age, individuals greater than or equal to a given maximum age are
+grouped together.
+
+Spawning stock biomass *ssb* is calculated by converting age to mean
+length, calculating weight at age, maturity at age, and then calculating
+spawning stock biomass as the sum of spawning potential at age in a
+given time step, taking into account the potential for hyperallometry in
+the manner of @marshall2021. Age is converted to length through the von
+Bertalanffy growth equation given parameters asymptotic length
+($l_\infty$), growth (*k*) and theoretical age at length zero ($a0$)
+assuming log-normally distributed variation *u* in the length at age
+with CV $\sigma_s$.
 
@@
 
-Biomass *b* at age is then given by the weight at length equation governed by a scaling coefficient $\Omega_s$ and an exponent $\Phi_s$ that controls the rate at which volume scales with length
+Biomass *b* at age is then given by the weight at length equation
+governed by a scaling coefficient $\Omega_s$ and an exponent $\Phi_s$
+that controls the rate at which volume scales with length
 
@@
 
-The proportion of sexually mature individuals (*mat*) at a given age is then calculated as a logistic function where $l_{mat}$ is the length at which on average 50% of individuals are sexually mature, and $\delta_{mat}$ is the unit of length beyond $l_{mat}$ at which on average 95% of fish are sexually mature.
+The proportion of sexually mature individuals (*mat*) at a given age is
+then calculated as a logistic function where $l_{mat}$ is the length at
+which on average 50% of individuals are sexually mature, and
+$\delta_{mat}$ is the unit of length beyond $l_{mat}$ at which on
+average 95% of fish are sexually mature.
 
@@
 
-Spawning stock biomass at time *t* is then calculated as a function of the numbers at age, the maturity at age, and the weight at age raised by a parameter $\gamma$.
-When $\gamma$ is greater than 1, the species is said to experience hyperallometric fecundity, i.e.
-fecundity increases faster than weight.
+Spawning stock biomass at time *t* is then calculated as a function of
+the numbers at age, the maturity at age, and the weight at age raised by
+a parameter $\gamma$. When $\gamma$ is greater than 1, the species is
+said to experience hyperallometric fecundity, i.e. fecundity increases
+faster than weight.
 
@@
 
-Recruitment (i.e.
-the number of age 1 individuals entering the population) follows Beverton-Holt dynamics parameterized around steepness (*h*) with log-normally distributed recruitment deviates $\epsilon$.
-When steepness is one recruitment is independent of spawning biomass.
-As steepness approaches 0.2 recruitment becomes a linear function of spawning biomass.
-We allow for five variants in the timing of density dependent recruitment:
-
-1.  Global density dependence: Density dependent recruitment is a function of the sum of spawning biomass across all patches, and recruits are then distributed according to habitat quality
+Recruitment (i.e. the number of age 1 individuals entering the
+population) follows Beverton-Holt dynamics parameterized around
+steepness (*h*) with log-normally distributed recruitment deviates
+$\epsilon$. When steepness is one recruitment is independent of spawning
+biomass. As steepness approaches 0.2 recruitment becomes a linear
+function of spawning biomass. We allow for five variants in the timing
+of density dependent recruitment:
+
+1.  Global density dependence: Density dependent recruitment is a
+    function of the sum of spawning biomass across all patches, and
+    recruits are then distributed according to habitat quality
 
@@
 
-where $r0$ is is a vector of recruits under unfished conditions in a given patch.
-
-2.  Local density dependence: Density dependent recruitment occurs independently in each patch and recruits are retained in their home patch.
+where $r0$ is is a vector of recruits under unfished conditions in a
+given patch.
+
+2.  Local density dependence: Density dependent recruitment occurs
+    independently in each patch and recruits are retained in their home
+    patch.
 
@@
 
-3.  Local density dependence then disperse: Density dependent recruitment occurs independently in each patch and recruits are then dispersed.
+3.  Local density dependence then disperse: Density dependent
+    recruitment occurs independently in each patch and recruits are then
+    dispersed.
 
@@
 
-4.  Post-dispersal density dependence: Larvae are distributed throughout the system, and then density dependent recruitment occurs based on the density of adult biomass at the destination patch.
+4.  Post-dispersal density dependence: Larvae are distributed throughout
+    the system, and then density dependent recruitment occurs based on
+    the density of adult biomass at the destination patch.
 
@@
 
-5.
-Global density dependence allocated by spawning biomass: Density dependence is a function of the sum of spawning biomass across all patches, and recruits are then distributed according to the distribution of spawning biomass
+5.  Global density dependence allocated by spawning biomass: Density
+    dependence is a function of the sum of spawning biomass across all
+    patches, and recruits are then distributed according to the
+    distribution of spawning biomass
 
@@
 
-Log-normal recruitment deviates are calculated with the potential for autocorrelation defined with strength $\rho$
+Log-normal recruitment deviates are calculated with the potential for
+autocorrelation defined with strength $\rho$
 
@@
 
-And log recruitment deviates are converted to raw units using the bias correction factor
+And log recruitment deviates are converted to raw units using the bias
+correction factor
 
@@
 
-Fisheries management is often concerned with measuring stock status relative to maximum sustainable yield (MSY) based reference points, though the exact level of stock status desired by management or societal objectives may vary widely.
-MSY based reference points present a problem for a multi-fleet and spatial-temporal model such as `marlin`.
-MSY and the fishing mortality rate that would produce MSY, $F_{MSY}$, are a function of fishery selectivity.
-Fishery selectivity in this model can vary by fleet, and species can be distributed unevenly in both space and time.
-This means that the net effective fishing selectivity on a species can vary depending on the dynamics at a given moment, making the definition of an equilibrium concept such as MSY challenging.
-
-As such, we do not report MSY based reference points in the model by default, with the exception of the spawning stock biomass at MSY ($SSB_{MSY}$) which tends to be conserved across selectivity patterns as it is primarily a function of the mortality and flow of recruits relative to the maturity ogive.
-There are many different strategies for estimating reference points in spatially explicit systems [@kapur2021a].
-We leave it to users to define and find relevant reference point as required by their specific needs.
+Fisheries management is often concerned with measuring stock status
+relative to maximum sustainable yield (MSY) based reference points,
+though the exact level of stock status desired by management or societal
+objectives may vary widely. MSY based reference points present a problem
+for a multi-fleet and spatial-temporal model such as `marlin`. MSY and
+the fishing mortality rate that would produce MSY, $F_{MSY}$, are a
+function of fishery selectivity. Fishery selectivity in this model can
+vary by fleet, and species can be distributed unevenly in both space and
+time. This means that the net effective fishing selectivity on a species
+can vary depending on the dynamics at a given moment, making the
+definition of an equilibrium concept such as MSY challenging.
+
+As such, we do not report MSY based reference points in the model by
+default, with the exception of the spawning stock biomass at MSY
+($SSB_{MSY}$) which tends to be conserved across selectivity patterns as
+it is primarily a function of the mortality and flow of recruits
+relative to the maturity ogive. There are many different strategies for
+estimating reference points in spatially explicit systems [@kapur2021a].
+We leave it to users to define and find relevant reference point as
+required by their specific needs.
 
@@
 
-We include two examples demonstrating how `marlin` can be used to support marine spatial planning.
-In the first, we show how `marlin` can be used to compare the total and distributional impacts of MPAs designed in a heavily fished coastal coral reef ecosystem.
-In the second, we demonstrate how `marlin` can be used to assess components of climate resilience of alternative MPA design strategies in a pelagic system.
-
-Each of the case studies contains too many parameters and options to be succinctly presented in the text here, and we are concerned that attempting to duplicate all of the parameters of the model here would simply introduce error.
-As such, readers should consult the accompanying code to view the precise details of each simulation.
-Targeted tactical applications must carefully consider and document all decisions made around model parameters.
+We include two examples demonstrating how `marlin` can be used to
+support marine spatial planning. In the first, we show how `marlin` can
+be used to compare the total and distributional impacts of MPAs designed
+in a heavily fished coastal coral reef ecosystem. In the second, we
+demonstrate how `marlin` can be used to assess components of climate
+resilience of alternative MPA design strategies in a pelagic system.
+
+Each of the case studies contains too many parameters and options to be
+succinctly presented in the text here, and we are concerned that
+attempting to duplicate all of the parameters of the model here would
+simply introduce error. As such, readers should consult the accompanying
+code to view the precise details of each simulation. Targeted tactical
+applications must carefully consider and document all decisions made
+around model parameters.
 
@@
 
-Across all of our case studies we make use of two potential "rule of thumb" MPA design strategies:
-
-1.
-*Rate*: MPAs are placed based on the pre-MPA SSB/SSB0 weighted catch relative to the total catch in a patch.
-So, patches with high rates of catch of depleted species relative to total catch are prioritized.
-
-2.  *Target Fishing*: MPAs locations are prioritized proportional to fishery catches. Patches with high total catches are prioritized over patches with low catches.
-
-To our knowledge, MPAs are never randomly placed; their locations are decided based on stakeholder negotiation given available information and objectives.
-In theory, the design of MPA networks can be optimized through the use of a modeling framework, and depending on the validity of the model, this process is likely to produce better outcomes than manually-designed strategies [@rassweiler2012; @rassweiler2014].
-However, designing optimal MPA networks becomes increasingly difficult as the range of objectives and the complexity of the model increase.
-Therefore, while `marlin` provides some support for [optimization](https://danovando.github.io/marlin/articles/optimize-mpas.html), we focus here on the design and performance of these more "rule of thumb" design strategies that may be more accessible to a wider range of users.
-We allow all MPAs to be designed in a "mosaic" fashion, but users can easily extend the analysis to compare outcomes between "contiguous" (MPA is made up of one continuously connected block) and "mosaic" (MPAs can be separated in space) MPA designs [@pons2022] by developing and passing their own MPA design algorithms to the model.
+Across all of our case studies we make use of two potential "rule of
+thumb" MPA design strategies:
+
+1.  *Rate*: MPAs are placed based on the pre-MPA SSB/SSB0 weighted catch
+    relative to the total catch in a patch. So, patches with high rates
+    of catch of depleted species relative to total catch are
+    prioritized.
+
+2.  *Target Fishing*: MPAs locations are prioritized proportional to
+    fishery catches. Patches with high total catches are prioritized
+    over patches with low catches.
+
+To our knowledge, MPAs are never randomly placed; their locations are
+decided based on stakeholder negotiation given available information and
+objectives. In theory, the design of MPA networks can be optimized
+through the use of a modeling framework, and depending on the validity
+of the model, this process is likely to produce better outcomes than
+manually-designed strategies [@rassweiler2012; @rassweiler2014].
+However, designing optimal MPA networks becomes increasingly difficult
+as the range of objectives and the complexity of the model increase.
+Therefore, while `marlin` provides some support for
+[optimization](https://danovando.github.io/marlin/articles/optimize-mpas.html),
+we focus here on the design and performance of these more "rule of
+thumb" design strategies that may be more accessible to a wider range of
+users. We allow all MPAs to be designed in a "mosaic" fashion, but users
+can easily extend the analysis to compare outcomes between "contiguous"
+(MPA is made up of one continuously connected block) and "mosaic" (MPAs
+can be separated in space) MPA designs [@pons2022] by developing and
+passing their own MPA design algorithms to the model.
 
@@
 
-In our coral reef example, we model the dynamics of four tropical Pacific species: a grouper (*Epinephelus fuscoguttatus*), a shallow-reef snapper (*Lutjanus malabaricus*), a deep-water snapper (*Pristipomoides filamentosus*), and a reef shark (*Carcharhinus amblyrhynchos*).
-The simulated groupers undergo a mass migration to a spawning aggregation once per year, followed by the sharks.
-Shallow-reef snappers stay in reefs closer to shore year-round, while deep-water snappers stay in the deeper reefs (@fig-coral).
-
-These species are targeted by two different fleets.
-Fleet One primarily targets the grouper and nearshore snapper populations, but will land any incidentally captured sharks.
-Fleet One has a dome shaped selectivity pattern for both grouper and shallow-reef snapper due to demand for "plate sized" fish [@kindsvater2017], and logistic selectivity for sharks, as any shark caught is retained.
-Fleet One is totally dependent on fishing for their livelihood, meaning the local community takes advantage of every possible opportunity to fish, and as such we model it as a "constant effort" fishery.
-
-Fleet Two primarily targets the deep-water snapper population.
-While plate-sized snapper are still prized from this population, we model Fleet Two's selectivity as logistic due to high levels of discard mortality for deep-water snapper resulting from barotrauma.
-Fleet Two catches some grouper, though less than Fleet One, and receives no price for sharks due to the requirements of a certification program through which they sell their deep-water snapper.
-Accidental captures (bycatch) do occur, which results in mortality.
-Fleet Two operates under open-access dynamics, as fishing is not the primary means of subsistence; short-term effort expands and contracts in response to profitability of the primarily grouper-driven fishery.
-
-Each fleet has a different home-port location, and incurs an additional cost for fishing each patch in the system in proportion to the Euclidean distance from each patch to each fleet's home port.
-
-We use `marlin` to simulate the outcomes for both food production and conservation for each of the species and both of the fleets as a function of both MPA size and MPA design strategy.
-For this exercise, MPAs are placed with perfect information and have no design constraints for continuity (i.e.
-the selected MPA design can be as patchwork as dictated by the data).
-We ran the simulation on a quarterly level ($\Delta_t = 1/4$) and set the area of each patch to an arbitrary value of 1 ($\Delta_p = 1$).
+In our coral reef example, we model the dynamics of four tropical
+Pacific species: a grouper (*Epinephelus fuscoguttatus*), a shallow-reef
+snapper (*Lutjanus malabaricus*), a deep-water snapper (*Pristipomoides
+filamentosus*), and a reef shark (*Carcharhinus amblyrhynchos*). The
+simulated groupers undergo a mass migration to a spawning aggregation
+once per year, followed by the sharks. Shallow-reef snappers stay in
+reefs closer to shore year-round, while deep-water snappers stay in the
+deeper reefs (@fig-coral).
+
+These species are targeted by two different fleets. Fleet One primarily
+targets the grouper and nearshore snapper populations, but will land any
+incidentally captured sharks. Fleet One has a dome shaped selectivity
+pattern for both grouper and shallow-reef snapper due to demand for
+"plate sized" fish [@kindsvater2017], and logistic selectivity for
+sharks, as any shark caught is retained. Fleet One is totally dependent
+on fishing for their livelihood, meaning the local community takes
+advantage of every possible opportunity to fish, and as such we model it
+as a "constant effort" fishery.
+
+Fleet Two primarily targets the deep-water snapper population. While
+plate-sized snapper are still prized from this population, we model
+Fleet Two's selectivity as logistic due to high levels of discard
+mortality for deep-water snapper resulting from barotrauma. Fleet Two
+catches some grouper, though less than Fleet One, and receives no price
+for sharks due to the requirements of a certification program through
+which they sell their deep-water snapper. Accidental captures (bycatch)
+do occur, which results in mortality. Fleet Two operates under
+open-access dynamics, as fishing is not the primary means of
+subsistence; short-term effort expands and contracts in response to
+profitability of the primarily grouper-driven fishery.
+
+Each fleet has a different home-port location, and incurs an additional
+cost for fishing each patch in the system in proportion to the Euclidean
+distance from each patch to each fleet's home port.
+
+We use `marlin` to simulate the outcomes for both food production and
+conservation for each of the species and both of the fleets as a
+function of both MPA size and MPA design strategy. For this exercise,
+MPAs are placed with perfect information and have no design constraints
+for continuity (i.e. the selected MPA design can be as patchwork as
+dictated by the data). We ran the simulation on a quarterly level
+($\Delta_t = 1/4$) and set the area of each patch to an arbitrary value
+of 1 ($\Delta_p = 1$).
 
@@
 
-We model our pelagic use case loosely on the characteristics of the Western and Central Pacific Ocean (WCPO) tuna fisheries.
-We simulate trajectories of `r n_distinct(names(blue_fauna))` species commonly caught in the region, including both the highly abundant skipjack tuna to the heavily depleted oceanic whitetip shark (*Carcharhinus longimanus*) (@fig-blue-ssb).
-We use publicly available data on catch-per-unit-effort of each of these species from the WCPO as a proxy for baseline habitat distributions, noting that where possible, fishery-independent abundance indices would be preferable (@fig-blue).
-
-These pelagic species are caught by a longline fleet that primarily targets large adult tunas such as bigeye and yellowfin for high-grade consumption, and a purse-seine fishery that primarily targets skipjack tunas for bulk canning.
-Selectivities were modeled as logistic for the longline fleet, and dome-shaped for the purse-seine fleet, with selectivity values set to broadly match the patterns reported in relevant [stock assessments](https://oceanfish.spc.int/en/ofpsection/sam/sam/210-stock-assessment-input-and-result-files-) for each species.
-Both fleets operate under open-access dynamics with an effort cap.
-The effort cap was set at the level of effort that resulted in the desired levels of SSB/SSB0 for each species under open-access dynamics (@fig-blue), intended to simulate a scenario where managers step in to prevent further expansion of fishing effort in a fully developed fishery.
-For forward-simulation, open-access dynamics can result in effort decreasing in response to profitability, but cannot result in effort beyond the effort cap set for each fleet.
-
-For this exercise, we focused on using `marlin` to assess resilience of the selected "rule of thumb" MPA design strategies to a climate-driven range shift.
-Specifically, we simulate an extreme example where the centroid of each population shifts northward at a rate of 0.2 patches per year, over a 20 year time horizon (@fig-blue-ssb).
-We designed MPA networks given the conditions in the starting year, and then held that network constant over the years of the experiment, running one simulation with and another without the climate-drive range shift.
-We then compared the effects of this range shift on food production and conservation outcomes from MPA networks designed based on the pre-range shift world.
-We ran the simulation on a quarterly level ($\Delta_t = 1/4$) and set the area of each patch to an arbitrary value of 1 ($\Delta_p = 1$).
+We model our pelagic use case loosely on the characteristics of the
+Western and Central Pacific Ocean (WCPO) tuna fisheries. We simulate
+trajectories of `r n_distinct(names(blue_fauna))` species commonly
+caught in the region, including both the highly abundant skipjack tuna
+and the heavily depleted oceanic whitetip shark (*Carcharhinus
+longimanus*) (@fig-blue-ssb). We use publicly available data on
+catch-per-unit-effort of each of these species from the WCPO as a proxy
+for baseline habitat distributions, noting that where possible,
+fishery-independent abundance indices would be preferable (@fig-blue).
+
+These pelagic species are caught by a longline fleet that primarily
+targets large adult tunas such as bigeye and yellowfin for high-grade
+consumption, and a purse-seine fishery that primarily targets skipjack
+tunas for bulk canning. Selectivities were modeled as logistic for the
+longline fleet, and dome-shaped for the purse-seine fleet, with
+selectivity values set to broadly match the patterns reported in
+relevant [stock
+assessments](https://oceanfish.spc.int/en/ofpsection/sam/sam/210-stock-assessment-input-and-result-files-)
+for each species. Both fleets operate under open-access dynamics with an
+effort cap. The effort cap was set at the level of effort that resulted
+in the desired levels of SSB/SSB0 for each species under open-access
+dynamics (@fig-blue), intended to simulate a scenario where managers
+step in to prevent further expansion of fishing effort in a fully
+developed fishery. For forward-simulation, open-access dynamics can
+result in effort decreasing in response to profitability, but cannot
+result in effort beyond the effort cap set for each fleet.
+
+For this exercise, we focused on using `marlin` to assess resilience of
+the selected "rule of thumb" MPA design strategies to a climate-driven
+range shift. Specifically, we simulate an extreme example where the
+centroid of each population shifts northward at a rate of 0.2 patches
+per year, over a 20 year time horizon (@fig-blue-ssb). We designed MPA
+networks given the conditions in the starting year, and then held that
+network constant over the years of the experiment, running one
+simulation with and another without the climate-drive range shift. We
+then compared the effects of this range shift on food production and
+conservation outcomes from MPA networks designed based on the pre-range
+shift world. We ran the simulation on a quarterly level
+($\Delta_t = 1/4$) and set the area of each patch to an arbitrary value
+of 1 ($\Delta_p = 1$).
 
@@
 
-For the coastal coral reef example the *Target fishing* MPA design strategies was capable of producing a Pareto-improving, i.e.
-potentially win-win, outcome where the total spawning biomass divided by unfished spawning biomass across all species increase by up to 200% while maintaining at least the same levels of fisheries yields as would be caught in the absence of any MPAs.
-Conversely, achieving increases in total fish biomass through the *Rate* strategy resulted in immediate losses in total yields, and in losses to both yields and total biomass in some cases (@fig-coral-results).
-
-These macro-level changes in total yield and conservation outcomes mask some internal tradeoffs.
-Nearly all of the food production benefits from the simulated MPA networks accrued to fleet one, while catches by fleet two were generally negatively impacted under both scenarios.
-For example, the MPA sizes and designs that result in the largest increase in yields for fleet one would produce a near total loss of fish catch for fleet two.
-
-The *Rate* MPA placement strategy produced lower total fishery gains but minimized the negative impacts on fleet two.
-However, this strategy produced much lower conservation gains than the *Target Fishing* strategy, only benefiting deep snapper meaningfully, and in fact resulting in slight losses to the nearshore snapper and grouper populations for MPA sizes less than roughly 75% of the simulated space.
-The *Target Fishing* strategy resulted in rapid conservation gains across all of the simulated species.
-
-MPAs were capable of producing positive outcomes in terms of total biomass and yields in the coastal coral reef example, but all the tested strategies resulted in tradeoffs across at least one dimension.
-`marlin` could be further used by a community such as the one simulated here to explore alternative design strategies, or to assess options for distributing fishery benefits across communities that may experience very different food provision outcomes from the same MPA network.
+For the coastal coral reef example the *Target fishing* MPA design
+strategies was capable of producing a Pareto-improving, i.e. potentially
+win-win, outcome where the total spawning biomass divided by unfished
+spawning biomass across all species increase by up to 200% while
+maintaining at least the same levels of fisheries yields as would be
+caught in the absence of any MPAs. Conversely, achieving increases in
+total fish biomass through the *Rate* strategy resulted in immediate
+losses in total yields, and in losses to both yields and total biomass
+in some cases (@fig-coral-results).
+
+These macro-level changes in total yield and conservation outcomes mask
+some internal tradeoffs. Nearly all of the food production benefits from
+the simulated MPA networks accrued to fleet one, while catches by fleet
+two were generally negatively impacted under both scenarios. For
+example, the MPA sizes and designs that result in the largest increase
+in yields for fleet one would produce a near total loss of fish catch
+for fleet two.
+
+The *Rate* MPA placement strategy produced lower total fishery gains but
+minimized the negative impacts on fleet two. However, this strategy
+produced much lower conservation gains than the *Target Fishing*
+strategy, only benefiting deep snapper meaningfully, and in fact
+resulting in slight losses to the nearshore snapper and grouper
+populations for MPA sizes less than roughly 75% of the simulated space.
+The *Target Fishing* strategy resulted in rapid conservation gains
+across all of the simulated species.
+
+MPAs were capable of producing positive outcomes in terms of total
+biomass and yields in the coastal coral reef example, but all the tested
+strategies resulted in tradeoffs across at least one dimension. `marlin`
+could be further used by a community such as the one simulated here to
+explore alternative design strategies, or to assess options for
+distributing fishery benefits across communities that may experience
+very different food provision outcomes from the same MPA network.
 
@@
 
-The yield impacts of the proposed MPA strategies were much more sensitive to climate impacts for the purse seine fishery.
-Under the *status quo* species distribution, the *Target Fishing* MPA design strategy produced increases in fisheries yields when the MPA covered 25-50% of the simulated area; that same MPA network resulted in the near complete loss of yields from the purse seine fleet when applied to the climate change scenario.
-In contrast, the yield outcomes for the longline fleet were relatively insensitive to the simulated range shifts (@fig-blue-results).
-
-From a conservation perspective, spawning biomass of species such as albacore and swordfish was relatively insensitive to the simulated range shifts, whereas species such as bigeye tuna and silky shark showed extreme variation in outcomes depending on the climate scenario.
-For bigeye tuna, under the *status quo* conditions the *Target Fishing* MPA design strategy resulted in rapid increases in the population, but that same MPA network produced much smaller and slower gains under climate change.
-However, for blue sharks this trend is inverted, with more rapid conservation gains under the range shift scenario than the *status quo* (@fig-blue-results).
+The yield impacts of the proposed MPA strategies were much more
+sensitive to climate impacts for the purse seine fishery than for the longline fishery. Under the
+*status quo* species distribution, the *Target Fishing* MPA design
+strategy produced increases in fisheries yields when the MPA covered
+25-50% of the simulated area; that same MPA network resulted in the near
+complete loss of yields from the purse seine fleet when applied to the
+climate change scenario. In contrast, the yield outcomes for the
+longline fleet were relatively insensitive to the simulated range shifts
+(@fig-blue-results).
+
+From a conservation perspective, spawning biomass of species such as
+albacore and swordfish was relatively insensitive to the simulated range
+shifts, whereas species such as bigeye tuna and silky shark showed
+extreme variation in outcomes depending on the climate scenario. For
+bigeye tuna, under the *status quo* conditions the *Target Fishing* MPA
+design strategy resulted in rapid increases in the population, but that
+same MPA network produced much smaller and slower gains under climate
+change. However, for blue sharks this trend is inverted, with more rapid
+conservation gains under the range shift scenario than the *status quo*
+(@fig-blue-results).
 
@@
 Marine ecosystems are driven by complex social-ecological dynamics.
-Communities must often make decisions on how to manage these systems based on limited empirical evidence.
-Modeling tools such as `marlin` can help users answer important scientific questions and design marine management policies informed by a better understanding of sensitivities to key uncertainties.
-In this paper we have documented the core system of equations underlying `marlin`, and demonstrated how `marlin` can be used to support efforts such as MPA planning.
-The variation in both the magnitude and distribution of costs and benefits stemming from alternative design choices and states of nature shown in our case study results demonstrates the value in our method.
-For example, our coastal coral reef example illustrates both the potential for MPAs to benefit conservation and food production in these systems, and the potential for the same MPA to benefit some fleets and species while harming others when fishing fleets affected by an MPA do not share the same objectives and species are not uniformly distributed.
-Our pelagic case study illustrates how `marlin` can be used to assess the climate resiliency of spatial management strategies, and identify strategies that best meet the needs of both current and future conditions.
-Results from this type of work could be used to help prioritize communities and species at particular risk to climate change impacts.
-
-The role of this paper is not to conduct a review of the many tools for spatial-temporal modeling available in the literature, each of which provide useful functionality for specific applications.
-However, we highlight here the specific gaps that we feel `marlin` fills in the modeling literature and in the policy support toolbox using some selected publications.
-More end-to-end models such as ATLANTIS [@audzijonyte2019] , OSMOSE [@shin2001], Ecopath with Ecosim [@christensen2004], POSEIDON [@bailey2019], DISPLACE [@bastardie2013], or SEAPODYM [@lehodey2008] are capable of representing tremendous amounts of complexity, but can be extremely time consuming to design and run.
-`marlin` allows some of the realism of these end-to-end models while being simpler and faster to parameterize and run.
-
-Tools such as virtualspecies [@leroy2016], STEPS [@visintin2020], RangeShifter [@bocedi2014], and [SMS](https://www.aquaveo.com/software/sms-surface-water-modeling-system-introduction) can be efficiently constructed to model the dispersal and distribution of species in space and time as a function of environmental covariates.
-However, these typically require specifying a covariate-response curve without explicitly acknowledging how this arises from habitat-specific movement and demography (virtualspecies), or model movement using a dispersal kernel (RangeShifter, STEPS) or a least-cost path algorithm (SMS, STEPS).
-By contrast, `marlin` uses a continuous-time Markov chain movement model, which integrates multiple paths (including their path-dependent probability based on intervening habitat types and species preferences) while using scale-free parameters that can be measured experimentally in laboratory or tagging studies.
-
-Simplified bio-economic models such as those used in @hastings2017, @sala2021, and @cabral2019 can be applied at scale and provide analytically tractable results, but as a result must abstract over many bio-economic dynamics that can be important for more tactical applications.
-`marlin` allows for more realistic representations of spatial social-ecological systems while maintaining processing speed.
-
-In sum, `marlin` complements the existing spatial marine modeling literature by allowing scientists, decision makers, and other stakeholders to efficiently and productively examine the impacts of realistic bio-economic dynamics on academic and applied problems.
-For example, scientists could use `marlin` to consider the relative importance of different biological variables in affecting policy outcomes.
-Managers could use `marlin` to assess the potential of proposed MPA networks to achieve stated objectives, such as maximizing conservation gains for a threatened species while minimizing or removing negative impacts to affected fishing fleets.
-
-Beyond the current examples, we envision `marlin` being applicable to research on dynamic ocean management, range shifts, management strategy evaluation, policy interactions, and spatial stock assessment.
-In addition, the process-based movement model used in `marlin` can directly use empirical estimates of movement dynamics derived in the manner of @thorson2021a, providing a link between empirical and simulation based approaches to marine resource management.
-Future iterations of the model may include life history modules for a broader range of species types such as marine mammals, sea birds, and sea turtles.
-
-Even the most complex marine model is a stylized cartoon of the true dynamics at play in the ocean.
-However, for all their limitations models such as `marlin` can help users understand factors that drive the performance of marine management strategies.
-The goal of this tool is to empower people to design policies based on evaluation of key uncertainties and trade-offs, and in doing so support more effective and equitable marine resource management.
+Communities must often make decisions on how to manage these systems
+based on limited empirical evidence. Modeling tools such as `marlin` can
+help users answer important scientific questions and design marine
+management policies informed by a better understanding of sensitivities
+to key uncertainties. In this paper we have documented the core system
+of equations underlying `marlin`, and demonstrated how `marlin` can be
+used to support efforts such as MPA planning. The variation in both the
+magnitude and distribution of costs and benefits stemming from
+alternative design choices and states of nature shown in our case study
+results demonstrates the value in our method. For example, our coastal
+coral reef example illustrates both the potential for MPAs to benefit
+conservation and food production in these systems, and the potential for
+the same MPA to benefit some fleets and species while harming others
+when fishing fleets affected by an MPA do not share the same objectives
+and species are not uniformly distributed. Our pelagic case study
+illustrates how `marlin` can be used to assess the climate resiliency of
+spatial management strategies, and identify strategies that best meet
+the needs of both current and future conditions. Results from this type
+of work could be used to help prioritize communities and species at
+particular risk to climate change impacts.
+
+The role of this paper is not to conduct a review of the many tools for
+spatial-temporal modeling available in the literature, each of which
+provide useful functionality for specific applications. However, we
+highlight here the specific gaps that we feel `marlin` fills in the
+modeling literature and in the policy support toolbox using some
+selected publications. More end-to-end models such as ATLANTIS
+[@audzijonyte2019] , OSMOSE [@shin2001], Ecopath with Ecosim
+[@christensen2004], POSEIDON [@bailey2019], DISPLACE [@bastardie2013],
+or SEAPODYM [@lehodey2008] are capable of representing tremendous
+amounts of complexity, but can be extremely time consuming to design and
+run. `marlin` allows some of the realism of these end-to-end models
+while being simpler and faster to parameterize and run.
+
+Tools such as virtualspecies [@leroy2016], STEPS [@visintin2020],
+RangeShifter [@bocedi2014], and
+[SMS](https://www.aquaveo.com/software/sms-surface-water-modeling-system-introduction)
+can be efficiently constructed to model the dispersal and distribution
+of species in space and time as a function of environmental covariates.
+However, these typically require specifying a covariate-response curve
+without explicitly acknowledging how this arises from habitat-specific
+movement and demography (virtualspecies), or model movement using a
+dispersal kernel (RangeShifter, STEPS) or a least-cost path algorithm
+(SMS, STEPS). By contrast, `marlin` uses a continuous-time Markov chain
+movement model, which integrates multiple paths (including their
+path-dependent probability based on intervening habitat types and
+species preferences) while using scale-free parameters that can be
+measured experimentally in laboratory or tagging studies.
+
+Simplified bio-economic models such as those used in @hastings2017,
+@sala2021, and @cabral2019 can be applied at scale and provide
+analytically tractable results, but as a result must abstract over many
+bio-economic dynamics that can be important for more tactical
+applications. `marlin` allows for more realistic representations of
+spatial social-ecological systems while maintaining processing speed.
+
+In sum, `marlin` complements the existing spatial marine modeling
+literature by allowing scientists, decision makers, and other
+stakeholders to efficiently and productively examine the impacts of
+realistic bio-economic dynamics on academic and applied problems. For
+example, scientists could use `marlin` to consider the relative
+importance of different biological variables in affecting policy
+outcomes. Managers could use `marlin` to assess the potential of
+proposed MPA networks to achieve stated objectives, such as maximizing
+conservation gains for a threatened species while minimizing or removing
+negative impacts to affected fishing fleets.
+
+Beyond the current examples, we envision `marlin` being applicable to
+research on dynamic ocean management, range shifts, management strategy
+evaluation, policy interactions, and spatial stock assessment. In
+addition, the process-based movement model used in `marlin` can directly
+use empirical estimates of movement dynamics derived in the manner of
+@thorson2021a, providing a link between empirical and simulation based
+approaches to marine resource management. Future iterations of the model
+may include life history modules for a broader range of species types
+such as marine mammals, sea birds, and sea turtles.
+
+Even the most complex marine model is a stylized cartoon of the true
+dynamics at play in the ocean. However, for all their limitations models
+such as `marlin` can help users understand factors that drive the
+performance of marine management strategies. The goal of this tool is to
+empower people to design policies based on evaluation of key
+uncertainties and trade-offs, and in doing so support more effective and
+equitable marine resource management.
 
@@
 
-All code and materials needed to reproduce this paper is available at (https://github.com/DanOvando/marlin-paper)\[https://github.com/DanOvando/marlin-paper\].
+All code and materials needed to reproduce this paper is available at
+(https://github.com/DanOvando/marlin-paper)\[https://github.com/DanOvando/marlin-paper\].
 

### 2023-03-22 functions/create_critters.R

@@ -6,8 +6,7 @@ create_critters <-
            seasonal_movement = FALSE,
            ontogenetic_shift = FALSE,
            random_rec = FALSE,
-           adult_diffusion = 10,
-           taxis_to_diff_ratio = 2) {
+           adult_diffusion = 10) {
     # sciname <- marlin_inputs$scientific_name[[1]]
     #
     #
@@ -54,10 +53,9 @@ create_critters <-
 
     critter <- marlin::create_critter(
       scientific_name = sciname,
-      base_habitat = hab,
+      habitat = hab,
       recruit_habitat = recruit_habitat,
       adult_diffusion = adult_diffusion,
-      taxis_to_diff_ratio = taxis_to_diff_ratio,
       recruit_diffusion = 0,
       density_dependence = density_dependence,
       fec_form = ifelse(str_detect(sciname,"carcharhinus"),"pups","weight"),

### 2023-03-28 02_blue_water_case_study.R
passing back to tuna for edits

@@ -1,8 +1,22 @@
 source(file.path("00_setup.R"))
 
-resolution <- 12
+resolution <- 12 # number of patches per side of simulation grid
+
+patches <- resolution ^ 2
+
+# target simulation area: 14,000,000 km2 = 3741 side length
+
+patch_area <- 14e6 / patches # km2
+
+patch_side <- sqrt(patch_area) # km
+
+simulation_area <- patch_area * patches #km2
+
+message(glue::glue("Blue water simulation area is {scales::comma(patch_area)} km2, spread over {patches} patches each with an area of {scales::comma(patch_area)} km2 or {scales::comma(patch_side)} km per side"))
 
 seasons <- 4
+
+time_step <- 1 / seasons # in years
 
 tune_type <- "explt"
 
@@ -10,7 +24,9 @@ experiment_workers <- 8
 
 experiment_years <- 20
 
-adult_diffusion <- 10
+adult_diffusion <- 0.8 * simulation_area  # km2, which means that each year a fish can roam an area 4km
+
+max_hab_mult = 20
 
 # load bycatch risk layers ------------------------------------------------
 
@@ -115,8 +131,8 @@ get_layer <- function(file) {
 
   hab$interp_cpue <- as.numeric(predict(mod$gam))
 
-  hab$habitat <- scales::rescale(hab$habitat, c(0,log(3)))
-  
+  hab$habitat <- scales::rescale(hab$interp_cpue, c(0,log(20)))
+
   # hab$habitat <- pmax(0, hab$interp_cpue / max(hab$interp_cpue) * adult_diffusion)
 
 
@@ -169,8 +185,10 @@ casestudy <- casestudy %>%
     create_critters,
     marlin_inputs = marlin_inputs,
     seasons = seasons,
-    adult_diffusion = adult_diffusion
-  )) %>%
+    adult_diffusion = adult_diffusion,
+    resolution = resolution,
+    max_hab_mult = max_hab_mult,
+    patch_area = patch_area)) %>%
   mutate(xid = 1)
 
 casestudy <- casestudy %>%
@@ -271,7 +289,7 @@ if (run_blue_water_example == TRUE){
 
       # range shift the preferred habitat, i.e. keep the shape of the habitat the same, but move it northward
       future_hab <- tmp_hab %>%
-        mutate(y = y + .2 * year) %>%
+        mutate(y = y + .33 * year) %>%
         mutate(xy = x * y)
 
[…]

### 2023-03-28 functions/run_mpa_experiment.R

@@ -14,6 +14,7 @@ run_mpa_experiment <-
            resolution,
            years = 50,
            future_habitat = list(),
+           spawning_ground = spawning_ground,
            effort_cap) {
 
     options(dplyr.summarise.inform = FALSE)
@@ -51,7 +52,7 @@ run_mpa_experiment <-
         left_join(depletion, by = "critter") %>%
         filter(critter %in% critters_considered) %>%
         group_by(critter, step) %>%
-        mutate(patch_weight = ssb / sum(ssb) * weight) %>%
+        mutate(patch_weight = (ssb / sum(ssb)) * weight) %>%
         group_by(patch, step) %>%
         summarise(patch_weight = sum(patch_weight)) %>%
         ungroup() %>%
@@ -67,7 +68,6 @@ run_mpa_experiment <-
       priorities <-  priorities %>%
         filter(step == step_priority$step[1]) %>%
         arrange(desc(patch_weight))
-
 
     } else if (placement_strategy == "rate"){
 
@@ -153,6 +153,13 @@ run_mpa_experiment <-
         group_by(patch) %>%
         summarise(patch_weight = unique(patch)) %>%
         arrange((patch_weight))
+
+
+    } else if (placement_strategy == "spawning_ground"){
+
+      priorities <- spawning_ground |>
+        mutate(patch = 1:length(x)) %>%
+        arrange(desc(habitat))
 
 
     } else {

### 2023-03-28 03_marlin_paper.qmd
passing back to tuna for edits
passing back to abalone for a final top to bottom run

@@
 
-# simple function to make prettier labels
-titler <- function(x){
-
-  stringr::str_to_title(stringr::str_replace_all(x,"_"," "))
-
-}
-
 # rdatas <- results_files[str_detect(results_files,'.Rdata')]
@@
 
-Abstract: Designing effective spatial management strategies is
-challenging because marine ecosystems are highly dynamic and opaque, and
-extractive entities such as fishing fleets respond endogenously to
-ecosystem changes in ways that depend upon ecological and policy
-context. We present a tool -- marlin -- to efficiently simulate the
-bio-economic dynamics of marine systems, designed to support both
-management and research. We demonstrate `marlin’s` capabilities by
-focusing on two case studies on the conservation and food production
-impacts of marine protected areas (MPAs): a coastal coral reef and a
-pelagic tuna fishery. In the coastal coral reef example, we explore how
-heterogeneity in species distributions and fleet preferences can affect
-distributional outcomes of MPAs. In the pelagic case study, we
-demonstrate how our model can be used to assess the climate resilience
-of different MPA design strategies, as well as the climate sensitivity
-of different fishing fleets. This paper demonstrates how intermediate
-complexity simulation of coupled bio-economic dynamics can help
-communities predict and potentially manage trade-offs between
-conservation, fisheries yields, and distributional outcomes resulting
-from spatial management.
-
 ## Introduction
 
-Communities around the world are increasingly looking to spatial
-strategies for managing marine ecosystems . For example, the "30x30"
-movement calls for implementation of marine protected areas (MPAs), a
-form of spatial management, across 30% of the world's oceans by the year
-2030 [@grorud-colvert2021 and references therein]. However , designing
-spatial management strategies to achieve desired objectives -- which may
-include recovery of overfished species, increased food production and
-economic well being, and equitable distribution of benefits - is not
-straightforward. Different species have different resilience to fishing,
-distributions in space and time, and value to fishing fleets; fishing
-fleets themselves can have varying ranges of incentives and reactions to
-policy structures. These dynamics pose challenges even for
-single-species assessment and management strategies [@field2006a], which
-are only amplified when we consider management policies specifically
-designed around spatial features in the oceans.
-
-For example, policies such as MPAs designed for both conservation and
-food production must consider factors such as the optimal size and
-placement of a protected area network given a wide array of life
-history, species distributions, exploitation levels, fleet dynamics, and
-policy-dependent behavior [@reimer2017], all of which may change over
-time, particularly given the impacts of climate change. Efforts to
-effectively design spatial management strategies such as MPAs are
-further constrained by a lack of empirical evidence describing the size
-[@ovando2021a] and time required [@nickols2019] for MPAs to produce
-substantial effects across a range of social-ecological settings
-[@mcclanahan2021].
-
-Many models have been developed to support the theory and design of
-spatial management strategies (see @fulton2015 for a review of existing
-methods). However, they have tended to be either highly complex and
-tactical models designed for use in a specific location, or to be
-extremely stylized representations intended to provide theoretical
-insights with less relevance for specific applications. This lack of
-accessible models suited for more tactical application presents a
-challenge to stakeholders charged with marine spatial planning
-exercises. It also presents a barrier to the scientific community,
-wherein comparisons across spatial management simulation studies are
-clouded by discrepancies in underlying model structures beyond
-differences in the core phenomenon in question.
-
-To help address this challenge, we present a novel bio-economic modeling
-tool called [`marlin`](https://danovando.github.io/marlin/) that allows
-for efficient simulation of spatio-temporal biological and economic
-dynamics. `marlin` allows users to simulate the impacts of marine
-management policies across a range of species targeted by various fleets
-across heterogeneous and dynamic seascapes. This model can help users
-seek Pareto-optimal solutions that produce win-wins or minimize
-tradeoffs across multiple management objectives, such as food production
-and conservation [@lester2013].
-
-Here we present the core methods and functionality of `marlin`, and
-demonstrate its use in two case-study applications; a coral reef system
-and a pelagic system. Our results show how interacting ecological,
-economic, and design attributes can affect the degree of benefits,
-costs, and tradeoffs between conservation and food production outcomes
-of MPAs. More broadly, this paper both demonstrates the functionality of
-our model and the critical need for considering sufficiently realistic
-coupled ecological and economic dynamics in policy evaluations.
+Communities around the world are increasingly looking to spatial strategies for managing marine ecosystems .
+For example, the "30x30" movement calls for implementation of marine protected areas (MPAs), a form of spatial management, across 30% of the world's oceans by the year 2030 [@grorud-colvert2021 and references therein].
+However , designing spatial management strategies to achieve desired objectives -- which may include recovery and resilience of overfished species, increased food production and economic well being, and equitable distribution of benefits - is not straightforward.
+Different species have different resilience to fishing, distributions in space and time, and value to fishing fleets; fishing fleets themselves can have varying ranges of incentives and reactions to policy structures.
+These dynamics pose challenges even for single-species assessment and management strategies , which are only amplified when we consider management policies specifically designed around multiple species, fleets, and spatial features in the oceans [@field2006a].
+
+To illustrate, policies such as MPAs designed for both conservation and food production must consider factors such as the optimal size and placement of a protected area network given a wide array of life history, species distributions, exploitation levels, fleet dynamics, and policy-dependent behavior [@reimer2017], all of which may change over time, particularly given the impacts of climate change.
+Efforts to effectively design spatial management strategies such as MPAs are further constrained by a lack of empirical evidence describing the size [@ovando2021a] and time required [@nickols2019] for MPAs to produce substantial effects across a range of social-ecological settings [@mcclanahan2021].
+
+Many models have been developed to support the theory and design of spatial management strategies (see @fulton2015 for a review of existing methods).
+However, these models have tended to be either highly complex and tactical models designed for use in a specific location, or to be extremely stylized representations intended to provide theoretical insights with less relevance for specific applications.
+This lack of accessible models suited for more tactical application presents a challenge to stakeholders charged with marine spatial planning exercises.
+It also presents a barrier to the scientific community, wherein comparisons across spatial management simulation studies are clouded by discrepancies in underlying model structures beyond differences in the core phenomenon in question.
+
+To help address this challenge, we present a novel bio-economic modeling tool called [`marlin`](https://danovando.github.io/marlin/) that allows for efficient simulation of spatio-temporal biological and economic dynamics.
+`marlin` allows users to simulate the impacts of marine management policies across a range of species targeted by various fleets across heterogeneous and dynamic seascapes.
+This model can help users seek Pareto-optimal solutions that produce win-wins or minimize trade-offs across multiple management objectives, such as food production and conservation [@lester2013].
+
+Here we present the core methods and functionality of `marlin`, and demonstrate its use in two case-study applications; a coral reef system and a pelagic system.
+Our results show how interacting ecological, economic, and design attributes can affect the degree of benefits, costs, and trade offs between conservation and food production outcomes of MPAs.
+More broadly, this paper both demonstrates the functionality of our model and the critical need for considering sufficiently realistic coupled ecological and economic dynamics in policy evaluations.
 
@@
 
-`marlin` follows a "models of intermediate complexity" philosophy
-[@plagányi2014] targeted specifically towards capturing heterogeneous
-habitats, movement dynamics, and the simultaneous impacts of fishing
-fleets across multiple species (frequently called "technical
-interactions"). In its general form, `marlin` simulates the behavior of
-populations of spatially explicit biologically independent
-age-structured animal populations affected by one or more fishing fleets
-in time steps and a spatial resolution specified by the user.
-
-The front end of the software package accompanying this paper,
-[`marlin`](https://danovando.github.io/marlin/), is written in R
-[@rcoreteam2021] to facilitate use, while the underlying population
-model is written in C++, integrated through the `Rcpp` package
-[@eddelbuettel2018], for speed. Users on standard computers should be
-able to simulate one age-structured population distributed across a 100
-patch system for 20 years in fractions of a second.
-
-While `marlin` does not simulate species interactions, fishing fleets in
-`marlin` are capable of targeting and affecting multiple species
-simultaneously. This allows the model to simulate processes such as
-fisheries bycatch or effort displacement in a way that accounts for how
-fleet behavior might affect multiple species in the system. The ability
-to track the impacts of fishing fleets across multiple species
-simultaneously is particularly important as very few fishing fleets are
-truly single species.
-
-We define a few commonly used terms here. *Yields* refers to the volume
-of catch from fishing activity. *Spawning stock biomass* refers to the
-total biomass of reproductively mature fish in the water, a function of
-numbers, weight, and sexual maturity at age. We measure the size of the
-population by th ratio of the *spawning stock biomass* (SSB) in a given
-time step relative to *unfished spawning stock biomass* (SSB0),
-SSB/SSB0. So, a SSB/SSB0 value of 1 means that the population is
-unfished, a SSB/SSB0 value of 0 means the population is extinct.
-
-As a demonstration of the use of this modeling tool, we explore two
-general situations
-
-1.  Trade-offs and distributional outcomes for food production and
-    multi-species biomass in a coastal coral-reef fishery
-2.  Implications of climate-driven range shifts for MPA design in a
-    pelagic fishery
-
-We chose these two examples to illustrate the use of `marlin` in
-contrasting systems in which spatial management strategies such as MPAs
-are increasingly considered.
-
-Below we provide a summary of the `marlin` model, as well as details of
-the case studies. We note that `marlin` is a complex software package
-with many options to control detailed behaviors. Given that these
-specific features will evolve over time, we do not describe in detail
-every option of the model here. Rather, we focus on the equations and
-methods describing the core foundations of our tool for efficiently
-simulating bio-economic dynamics in aquatic systems.
+`marlin` follows a "models of intermediate complexity" philosophy [@plagányi2014] targeted specifically towards capturing heterogeneous habitats, movement dynamics, and the simultaneous impacts of fishing fleets across multiple species (frequently called "technical interactions").
+In its general form, `marlin` simulates the behavior of populations of spatially explicit biologically independent age-structured animal populations affected by one or more fishing fleets in time steps and a spatial resolution specified by the user.
+
+The front end of the software package accompanying this paper, [`marlin`](https://danovando.github.io/marlin/), is written in R [@rcoreteam2021] to facilitate use, while the underlying population model is written in C++, integrated through the `Rcpp` package [@eddelbuettel2018], for speed.
+Users on standard computers should be able to simulate one age-structured population distributed across a 10 by 10 patch system for 20 years in fractions of a second.
+
+While `marlin` does not simulate species interactions, fishing fleets in `marlin` are capable of targeting and affecting multiple species simultaneously.
+This allows the model to simulate processes such as fisheries bycatch or effort displacement in a way that accounts for how fleet behavior might affect multiple species in the system.
+The ability to track the impacts of fishing fleets across multiple species simultaneously is particularly important as very few fishing fleets are truly single species.
+
+We define a few commonly used terms here.
+*Yields* refers to the volume of catch from fishing activity.
+*Spawning stock biomass* refers to the total biomass of reproductively mature fish in the water, a function of numbers, weight, fecundity, and sexual maturity at age.
+We measure the size of the population by th ratio of the *spawning stock biomass* (SSB) in a given time step relative to *unfished spawning stock biomass* (SSB0), SSB/SSB0.
+So, a SSB/SSB0 value of 1 means that the population is unfished, a SSB/SSB0 value of 0 means the population is extinct.
+
+As a demonstration of the use of this modeling tool, we explore two general situations
+
+1.  Trade-offs and distributional outcomes for food production and multi-species biomass in a coastal coral-reef fishery
+2.  Implications of climate-driven range shifts for MPA design in a pelagic fishery
+
+We chose these two examples to illustrate the use of `marlin` in contrasting systems in which spatial management strategies such as MPAs are increasingly considered.
+
+Below we provide a summary of the `marlin` model, as well as details of the case studies.
+We note that `marlin` is a complex software package with many options to control detailed behaviors.
+Given that these specific features will evolve over time, we do not describe in detail every option of the model here.
+Rather, we focus on the equations and methods describing the core foundations of our tool for efficiently simulating bio-economic dynamics in aquatic systems.
 
@@
 
-`marlin` simulates the dynamics of one or more species, currently best
-representing fish-like species, using age-structured population
-dynamics. Each time step, species move throughout the simulated seascape
-using both diffusion and "taxis" (active movement towards preferred
-habitat), experience natural and potentially fishing mortality, and if
-in a spawning season spawn using one of the possible forms of density
-dependence implemented in the model, with the potential to include
-auto-correlated stochastic deviations in the amount of offspring
-produced, generally called "recruitment deviates".
-
-These species can be caught by fishing fleets. A fishing fleet is
-defined in `marlin` by a set of fishers that have the same fishing
-skill, prices, and selectivity for individual species (each of which we
-denote as a *metier*). For example, both a longline and purse seine
-fleet may capture bigeye (*Thunnus obesus*) and skipjack tunas, but the
-longline fleet can be made much more likely to catch bigeye than
-skipjack, and *vice versa*. Each fishing fleet distributes its fishing
-effort in space according to a specified spatial allocation rule (see
-*Spatial Allocation of Effort* ), for example in proportion to
-profit-per-unit-effort, conditional on management policies in places
-such as quotas and/or the presence of any spatial restrictions such as
-MPAs.
-
-Unconstrained by management, the total amount of effort exerted by each
-fleet can follow one of two dynamics: open access or constant effort
-(see *Calculating Total Effort*). Under constant effort, the total
-amount of effort of each fleet is fixed over time, with the possible
-exception of attrition due to MPA placement. Under open access, the
-total amount of effort in a given time step is a function of the
-profitability of the fleet in the previous time step, evolving until a
-bionomic equilibrium of zero total profits is reached. Profitability is
-a function of the volume and price of each species caught, as well as
-the cost of the total amount of fishing effort per fleet and the travel
-costs as a function of distance from a port (see *Fleet Dynamics*).
-
-The dynamics of each fleet can be modified by management in a variety of
-ways. Managers can impose size limits for individual species within each
-fleet. They can also impose total catch quotas for each species in each
-time step. When quotas are activated, if the total catch across all
-fleets for a given species under the model's effort dynamics would
-exceed the allowable quota for that species, the total effort for each
-fleet is decreased proportional to its contribution to the total catch
-until the quota is satisfied. As an alternative to catch quotas,
-managers can set an effort cap per fleet, which prevents effort from
-exceeding a given amount under open access dynamics, though effort may
-be reduced below this cap if required by the profit equation (i.e. the
-fleet can choose to fish less than the quota or effort cap).
-
-`marlin` also allows users to specify closed fishing seasons for one or
-more species in the system. When all species in a simulation share the
-same closed seasons, no fishing occurs in those seasons. When not all
-species share the same closed seasons, fishing still occurs in the
-closed seasons, but the fleet avoids catching the species for which
-fishing is closed in a given season, assuming no discard mortality. When
-closed seasons are in effect, the amount of effort in the next open
-season is a function of the effort and profits in the last open fishing
-season. Lastly, managers can specify locations of no-take MPAs, which
-could in theory change in size and location. When MPAs are implemented,
-fishing effort that used to operate inside the MPAs can either leave the
-fishery, or be redistributed outside the MPA (the default behavior).
-
-Users can use the model to explore general properties of spatially
-explicit social-ecological systems, or can parameterize it to explore
-key uncertainties of more tactical applications. As an example of the
-later case, we provide two case studies here designed to reflect the
-kinds of questions stakeholders might face if designing an MPA network
-for a coastal coral reef or a pelagic system.
+`marlin` simulates the dynamics of one or more species, currently best representing fish-like species, using age-structured population dynamics.
+Each time step, species move throughout the simulated seascape using both diffusion and "taxis" (active movement towards preferred habitat), experience natural and potentially fishing mortality, and if in a spawning season spawn using one of the possible forms of density dependence implemented in the model, with the potential to include auto-correlated stochastic deviations in the amount of offspring produced, generally called "recruitment deviates".
+
+These species can be caught by fishing fleets.
+A fishing fleet is defined in `marlin` by a set of fishers that have the same fishing skill, prices, and selectivity for individual species (each of which we denote as a *metier*).
+For example, both a longline and purse seine fleet may capture bigeye (*Thunnus obesus*) and skipjack tunas, but the longline fleet can be made much more likely to catch larger bigeye than skipjack, and *vice versa*.
+Each fishing fleet distributes its fishing effort in space according to a specified spatial allocation rule (see *Spatial Allocation of Effort* ), for example in proportion to profit-per-unit-effort, conditional on management policies in places such as quotas and/or the presence of any spatial restrictions such as MPAs.
+
+Unconstrained by management, the total amount of effort exerted by each fleet can follow one of two dynamics: open access or constant effort (see *Calculating Total Effort*).
+Under constant effort, the total amount of effort of each fleet is fixed over time, with the possible exception of attrition due to MPA placement.
+Under open access, the total amount of effort in a given time step is a function of the profitability of the fleet in the previous time step, evolving until a bionomic equilibrium of zero total profits is reached.
+Profitability is a function of the volume and price of each species caught, as well as the cost of the total amount of fishing effort per fleet and the travel costs as a function of distance from a port (see *Fleet Dynamics*).
+
+The dynamics of each fleet can be modified by management in a variety of ways.
+Managers can impose size limits for individual species within each fleet.
+They can also impose total catch quotas for each species in each time step.
+When quotas are activated, if the total catch across all fleets for a given species under the model's effort dynamics would exceed the allowable quota for that species, the total effort for each fleet is decreased proportional to its contribution to the total catch until the quota is satisfied.
+As an alternative to catch quotas, managers can set an effort cap per fleet, which prevents effort from exceeding a given amount under open access dynamics, though effort may be reduced below this cap if required by the profit equation (i.e.
+the fleet can choose to fish less than the quota or effort cap).
+
+`marlin` also allows users to specify closed fishing seasons for one or more species in the system.
+When all species in a simulation share the same closed seasons, no fishing occurs in those seasons.
+When not all species share the same closed seasons, fishing still occurs in the closed seasons, but the fleet avoids catching the species for which fishing is closed in a given season, assuming no discard mortality.
+When closed seasons are in effect, the amount of effort in the next open season is a function of the effort and profits in the last open fishing season.
+Lastly, managers can specify locations of no-take MPAs, which can change in size and location if needed.
+When MPAs are implemented, fishing effort that used to operate inside the MPAs can either leave the fishery, or be redistributed outside the MPA (the default behavior).
+
+Users can use the model to explore general properties of spatially explicit social-ecological systems, or can parameterize it to explore key uncertainties of more tactical applications.
+As an example of the later case, we provide two case studies here designed to reflect the kinds of questions stakeholders might face if designing an MPA network for a coastal coral reef or a pelagic system.
 
@@
 
-Each fishing fleet (*f*) generates catches or yield (*Y*), revenues
-(*R*), and costs (*C*) from fishing individual species (*s*) that it
-targets in a given time step (*t*) and patch (*p*). The fleet then makes
-decisions around fishing locations and intensity, subject to regulatory
-constraints, based on the total profitability across all species in
-space and time.
-
-Revenues for each fleet in a time step are a function of the total
-amount of each species caught and its price ($\Pi$). The amount caught
-is a function of the selectivity at age of each species for each fleet
-($\alpha$), the fishing efficiency of the fleet for that that species
-(also called "catchability", *q*), the amount of fishing effort of the
-fleet in question in that patch in a given time step (*E*), and the
-total instantaneous fishing mortality (*u*) at age (*a*) (including all
-other fleets) for that particular species in that patch and time step.
+Each fishing fleet (*f*) generates catches or yield (*Y*), revenues (*R*), and costs (*C*) from fishing individual species (*s*) that it targets in a given time step (*t*) and patch (*p*).
+The fleet then makes decisions around fishing locations and intensity, subject to regulatory constraints, based on the total profitability across all species in space and time.
+
+Revenues for each fleet in a time step are a function of the total amount of each species caught and its price ($\Pi$).
+The amount caught is a function of the selectivity at age of each species for each fleet ($\alpha$), the fishing efficiency of the fleet for that that species (also called "catchability", *q*), the amount of fishing effort of the fleet in question in that patch in a given time step (*E*), and the total instantaneous fishing mortality (*u*) at age (*a*) (including all other fleets) for that particular species in that patch and time step.
 
@@
 
-Total catches or yield *Y* are calculated through the Baranov equation
-[@baranov1918], which appropriately accounts for the total instantaneous
-mortality (both fishing and natural. *z*) and divides the total amount
-of the population biomass (*b*) killed between natural (*m*) and fishing
-(*u*) sources, with the amount of biomass killed through fishing called
-"yield".
+Total catches or yield *Y* are calculated through the Baranov equation [@baranov1918], which appropriately accounts for the total instantaneous mortality (both fishing and natural.
+*z*) and divides the total amount of the population biomass (*b*) killed between natural (*m*) and fishing (*u*) sources, with the amount of biomass killed through fishing called "yield".
 
@@
 
-We approximate the dome-shaped form as a normal distribution with mean
-$l^{sel}$ and standard deviation $\sigma$. The normal density function
-is re-scaled such that the selectivity is 1 at $l^{sel}$.
+We approximate the dome-shaped form as a normal distribution with mean $l^{sel}$ and standard deviation $\sigma$.
+The normal density function is re-scaled such that the selectivity is 1 at $l^{sel}$.
 
@@
 
-This means that each species experiences a total mortality at age in a
-time step in a patch, individual fractions of which are portioned off as
-catches and subsequently revenues for each fleet. Given revenues *R*, we
-then calculate profits $\phi$ based on revenues and costs. Costs *C* are
-calculated as a function of base costs per unit effort ($\gamma$) as
-well as potential additional costs per unit effort of fishing in
-particular patches ($\eta$) for fleet *f*. $\beta$ allows for the cost
-per unit effort effort to scale non-linearly. Travel costs ($\eta$) are
-calculated based on the Euclidean distance of each patch to the nearest
-port of a given fleet, and users can specify any number from zero to the
-number of patches of port locations. When no ports are specified, travel
-costs are zero.
+This means that each species experiences a total mortality at age in a time step in a patch, individual fractions of which are portioned off as catches and subsequently revenues for each fleet.
+Given revenues *R*, we then calculate profits $\phi$ based on revenues and costs.
+Costs *C* are calculated as a function of base costs per unit effort ($\gamma$) as well as potential additional costs per unit effort of fishing in particular patches ($\eta$) for fleet *f*.
+$\beta$ allows for the cost per unit effort effort to scale non-linearly.
+Travel costs ($\eta$) are calculated based on the Euclidean distance of each patch to the nearest port of a given fleet, and users can specify any number from zero to the number of patches of port locations.
+When no ports are specified, travel costs are zero.
 
@@
 
-`marlin` allows for two general modes of effort. The simplest is
-"constant effort", in which the total effort of each fishing fleet
-remains constant over time.
-
-The more complex option is "open access". Under open access, the total
-amount of effort in a given time step for fleet *f* is a function of the
-profitability of that fleet in the proceeding time steps in which
-fishing was open, where $\theta$ controls the responsiveness of fleet
-*f* to the log of the ratio of total revenues *R* to total costs *C*,
-and approximates the proportional change in effort in response to a one
-unit changes in the log revenue to cost ratio.
+`marlin` allows for two general modes of effort.
+The simplest is "constant effort", in which the total effort of each fishing fleet remains constant over time.
+
+The more complex option is "open access".
+Under open access, the total amount of effort in a given time step for fleet *f* is a function of the profitability of that fleet in the proceeding time steps in which fishing was open, where $\theta$ controls the responsiveness of fleet *f* to the log of the ratio of total revenues *R* to total costs *C*, and approximates the proportional change in effort in response to a one unit changes in the log revenue to cost ratio.
 
@@
 
-This is essentially a Gompertz model for fishing effort, that has been
-used for other theoretical studies of fishing effort dynamics
-[@thorson2013].
+This is essentially a Gompertz model for fishing effort, that has been used for other theoretical studies of fishing effort dynamics [@thorson2013].
 
@@
 
-The degree of fishing pressure exerted by a given set of fishing fleets
-on each species is a function of a range of parameters including the
-total amount of effort (*E*), the selectivity ogives ($\alpha$,the curve
-specifying the proportion of individuals at a given age or size that
-will be caught by a unit of fishing mortality), fishing cost (*C*), the
-spatial distribution of the species affected by the fleet, the relative
-prices across species ($\Pi$), and the catchability coefficient of each
-fleet for each species ($q$). `marlin` provides a tuning option to help
-users achieve desired biological outcomes from their fleets.
-
-Users can tune their fleets in one of two ways. First, they can specify
-a target exploitation rate *u* for each species in their simulation.
-Taking all the other parameters of the model as given, `marlin` then
-adjusts the catchability coefficients $q_{f,s}$ for each fleet *f* and
-species *s* such that the equilibrium exploitation rate for each species
-matches the desired level. Second, they can specify a target total
-spawning stock biomass under fishing divided by total unfished spawning
-stock biomass, and the model will adjust the catchability coefficients
-$q_{f,s}$ for each fleet *f* and species *s* such that the equilibrium
-ratio of fished to unfished spawning biomass for each species matches
-the desired level.
-
-Users can also tune the fleet dynamics by specifying the ratio of costs
-to revenues. Price data for use in the model can be obtained relatively
-simply through literature reviews, market surveys, or local experts, so
-long as they are converted to the same units of biomass used in the
-model. However, cost parameters are more complicated, as translating say
-cost per day of fishing into the same representation of fishing effort
-used in the model is not straight forward to accomplish. As an
-alternative, users can specify an equilibrium cost-to-revenue ratio for
-the fleet, essentially the profit margins of the fleet in question, and
-`marlin` will tune the cost parameter to achieve this desired cost to
-revenue ratio given the other parameters in the model (e.g. price and
-biomass of the target species). So for example, while users may not have
-access to exact cost parameters, they may have a general sense that when
-operating under ideal conditions a given fleet has a profit margin of
-10%, i.e. the cost to revenue ratio is 90%. They can input the cost to
-revenue ratio as 0.9 then and allow `marlin` to tune the cost parameters
-accordingly.
+The degree of fishing pressure exerted by a given set of fishing fleets on each species is a function of a range of parameters including the total amount of effort (*E*), the selectivity ogives ($\alpha$,the curve specifying the proportion of individuals at a given age or size that will be caught by a unit of fishing mortality), fishing cost (*C*), the spatial distribution of the species affected by the fleet, the relative prices across species ($\Pi$), and the catchability coefficient of each fleet for each species ($q$).
+`marlin` provides a tuning option to help users achieve desired biological outcomes from their fleets.
+
+Users can tune their fleets in one of two ways.
+First, they can specify a target exploitation rate *u* for each species in their simulation.
+Taking all the other parameters of the model as given, `marlin` then adjusts the catchability coefficients $q_{f,s}$ for each fleet *f* and species *s* such that the equilibrium exploitation rate for each species matches the desired level.
+Second, they can specify a target total spawning stock biomass under fishing divided by total unfished spawning stock biomass, and the model will adjust the catchability coefficients $q_{f,s}$ for each fleet *f* and species *s* such that the equilibrium ratio of fished to unfished spawning biomass for each species matches the desired level.
+
+Users can also tune the fleet dynamics by specifying the ratio of costs to revenues.
+Price data for use in the model can be obtained relatively simply through literature reviews, market surveys, or local experts, so long as they are converted to the same units of biomass used in the model.
+However, cost parameters are more complicated, as translating say cost per day of fishing into the same representation of fishing effort used in the model is not straight forward to accomplish.
+As an alternative, users can specify an equilibrium cost-to-revenue ratio for the fleet, essentially the profit margins of the fleet in question, and `marlin` will tune the cost parameter to achieve this desired cost to revenue ratio given the other parameters in the model (e.g.
+price and biomass of the target species).
+So for example, while users may not have access to exact cost parameters, they may have a general sense that when operating under ideal conditions a given fleet has a profit margin of 10%, i.e.
+the cost to revenue ratio is 90%.
+They can input the cost to revenue ratio as 0.9 then and allow `marlin` to tune the cost parameters accordingly.
 
@@
 
-Each fishing fleet decides how to allocate its effort in space based on
-one of four possible spatial allocation strategies. The ideal free
-distribution (IFD) is the standard method for distributing fishing
-fleets in spaces (see @gillis2003 and references therein). However,
-analytical solutions to the IFD present a number of complications for
-our model. Primarily, our model is capable of tracking multiple fleets
-at once. This means that the IFD for a single fleet would commonly be
-modeled as a Nash-equilibrium based on the actions of each of the
-individual fleet conditional on the actions of all other fleets. While
-possible to implement, this would slow down our model runs to the point
-of practically preventing large-scale evaluation of the spatial
-management policies.
-
-As such we explored a series of "next best" fleet distribution
-algorithms. While not the IFD in any individual time step, over time
-they start to approximate the IFD, as each assumes that fleets base
-their decisions for the current time step on the outcomes in the prior
-timestep, meaning that the impacts of the actions of other fishing
-fleets are somewhat internalized. The four possible fleet distribution
-algorithms are
-
-1.  Revenue per unit effort (RPUE): The fleet distributes itself in
-    space based on the realized revenue per unit effort in each patch in
-    the preceding time step
-2.  Revenue: The fleet distributes itself in space based on the realized
-    total revenue in each patch in the preceding time step
-3.  Profit per unit effort (PPUE): The fleet distributes itself in space
-    based on the realized profit per unit effort in each patch in the
-    preceding time step
-4.  Profit: The fleet distributes itself in space based on profit in
-    each patch in the preceding time step
-
-Revenue based spatial distribution is not likely to be very realistic;
-in general we would expect fishing fleets to respond to profits on some
-core level. However, due to the complexity of parameterizing cost
-functions, fleet dynamics are often evaluated based on yield or revenue
-alone, and so we include those scenarios here to allow users to evaluate
-the potential implications of this choice. The decision on whether to
-allocate the fleet based on absolute or relative (per unit effort)
-metrics is more complex. When effort represents the actions of separate
-and individual fishing actors (e.g. independent fishing vessels), a
-per-unit-effort strategy, in which fishers distribute themselves based
-on the expected catch of their individual efforts, may be more realistic
-[@hilborn1987]. Conversely, a system defined by a sole owner seeking to
-maximize total profits might be better represented by a fleet model
-based on total profits
+Each fishing fleet decides how to allocate its effort in space based on one of four possible spatial allocation strategies.
+The ideal free distribution (IFD) is the standard method for distributing fishing fleets in spaces (see @gillis2003 and references therein).
+However, analytical solutions to the IFD present a number of complications for our model.
+Our model is capable of tracking multiple fleets at once.
+This means that the IFD for a single fleet would commonly be modeled as a Nash-equilibrium based on the actions of each of the individual fleet conditional on the actions of all other fleets.
+While possible to implement, this would slow down our model runs to the point of practically preventing large-scale evaluation of the spatial management policies.
+
+As such we explored a series of "next best" fleet distribution algorithms.
+While not the IFD in any individual time step, over time they start to approximate the IFD, as each assumes that fleets base their decisions for the current time step on the outcomes in the prior timestep, meaning that the impacts of the actions of other fishing fleets are somewhat internalized.
+The four possible fleet distribution algorithms are
+
+1.  Revenue per unit effort (RPUE): The fleet distributes itself in space based on the realized revenue per unit effort in each patch in the preceding time step
+2.  Revenue: The fleet distributes itself in space based on the realized total revenue in each patch in the preceding time step
+3.  Profit per unit effort (PPUE): The fleet distributes itself in space based on the realized profit per unit effort in each patch in the preceding time step
+4.  Profit: The fleet distributes itself in space based on profit in each patch in the preceding time step
+
+Revenue based spatial distribution is not likely to be very realistic; in general we would expect fishing fleets to respond to profits on some core level.
+However, due to the complexity of parameterizing cost functions, fleet dynamics are often evaluated based on yield or revenue alone, and so we include those scenarios here to allow users to evaluate the potential implications of this choice.
+The decision on whether to allocate the fleet based on absolute or relative (per unit effort) metrics is more complex.
+When effort represents the actions of separate and individual fishing actors (e.g.
+independent fishing vessels), a per-unit-effort strategy, in which fishers distribute themselves based on the expected catch of their individual efforts, may be more realistic [@hilborn1987].
+Conversely, a system defined by a sole owner seeking to maximize total profits might be better represented by a fleet model based on total profits
 
@@
 
-The underlying population model used is an age structured single-species
-model in the manner of @ovando2021a. The population model requires many
-parameters. However, if the user supplies either a scientific or common
-name for the species in question (scientific preferred), the model will
-supply default values for that species based on the values reported in
-FishLife [internet connection required, @thorson2020]. FishLife provides
-estimates of core life history parameters for fish species based on a
-model integrating published values and evolutionary connections.
-FishLife provides reasonable default values for given species, but these
-default values should be not be taken as definitive and users should
-input best available parameters specific to the stock in question if
-they wish to best represent the dynamics of their specific system, and
-should check the plausibility of the default values prior to use. All of
-the life history values can be replaced with user-supplied values as
-needed and should be for more tactical applications; the default values
-simply ensure that reasonable and internally consistent values for the
-species in question are supplied as a starting point.
+The underlying population model used is an age structured single-species model in the manner of @ovando2021a.
+The population model requires many parameters.
+However, if the user supplies either a scientific or common name for the species in question (scientific preferred), the model will supply default values for that species based on the values reported in FishLife [internet connection required, @thorson2020].
+FishLife provides estimates of core life history parameters for fish species based on a model integrating published values and evolutionary connections.
+FishLife provides reasonable default values for given species, but these default values should be not be taken as definitive and users should input best available parameters specific to the stock in question if they wish to best represent the dynamics of their specific system, and should check the plausibility of the default values prior to use.
+All of the life history values can be replaced with user-supplied values as needed and should be for more tactical applications; the default values simply ensure that reasonable and internally consistent values for the species in question are supplied as a starting point.
 
@@
 
-`marlin`'s movement dynamics are based on a continuous-time Markov chain
-(CTMC), as described in @thorson2021a. Within this framework, the model
-allows for movement to be broken down into three components of
-*advection* (follow currents), and *taxis* (active movement towards
-preferred habitat) and *diffusion* (essentially remaining variation in
-movement not explained by advection or taxis). For now, `marlin` focuses
-just on the diffusion and taxis components of this model, assuming that
-advection is zero, though future extensions could easily incorporate
-advection vectors from oceanographic models. In this way, `marlin`
-allows users to run anything from a simple Gaussian dispersal kernel up
-to a system governed by species that passively diffuse out from a core
-habitat defined by a dynamic thermal range.
-
-We provide a brief overview of the the general CTMC method here, see
-@thorson2021a for a detailed description. Under this framework, movement
-from each patch to each other patch in the system in a given timestep
-*t* for life stage *a* of species *s* is defined by a movement matrix
-$\boldsymbol{M}_{t,s,a}$. $\boldsymbol{M}_{t,s,a}$ is calculated as the
-matrix exponent of the sum of the diffusion $\boldsymbol{D}$ and taxis
-$\boldsymbol{\tau}$ matrices scaled by the width of the time step
-$\Delta_{t}$ and the length of the edge of each patch $\Delta_d$ that
-the model is running under. So for example, if we are running a model on
-a monthly timestep, and we define the units of time in years, then
-$\Delta_{t} = 1/12$. Similarly, if we are running a model where one
-patch equals 100km^2^, then $\Delta_d = 100km$. This parameterization
-allows users to set the effective area of the spatial domain through two
-avenues; the number of patches, which effectively scales the resolution
-of the model, and the area of each patch, which scales the spatial
-extent of the simulation.
+`marlin`'s movement dynamics are based on a continuous-time Markov chain (CTMC), as described in @thorson2021a.
+Within this framework, the model allows for movement to be broken down into three components of *advection* (drifting with currents), *taxis* (active movement towards preferred habitat), and *diffusion* (essentially remaining variation in movement not explained by advection or taxis).
+For now, `marlin` focuses just on the diffusion and taxis components of this model, assuming that advection is zero, though future extensions could easily incorporate advection vectors from oceanographic models.
+In this way, `marlin` allows users to run anything from a simple Gaussian dispersal kernel up to a system governed by species that passively diffuse out from a core habitat defined by a dynamic thermal range.
+
+We provide a brief overview of the the general CTMC method here, see @thorson2021a for a detailed description.
+Under this framework, movement from each patch to each other patch in the system in a given timestep *t* for life stage *a* of species *s* is defined by a movement matrix $\boldsymbol{M}_{t,s,a}$.
+$\boldsymbol{M}_{t,s,a}$ is calculated as the matrix exponent of a function of the diffusion $\boldsymbol{D}$ and taxis $\boldsymbol{\tau}$ matrices scaled by the width of the time step $\Delta_{t}$ and the length of the edge of each patch $\Delta_d$ that the model is running under.
+So for example, if we are running a model on a monthly timestep, and we define the units of time in years, then $\Delta_{t} = 1/12$.
+Similarly, if we are running a model where one patch equals 100km^2^, then $\Delta_d = 10km$.
+This parameterization allows users to set the effective area of the spatial domain through two avenues; the number of patches, which effectively scales the resolution of the model, and the area of each patch, which scales the spatial extent of the simulation.
 
@@
 
-The individual components of the diffusion matrix are filled based on an
-adjacency matrix, which defines whether two patches are both adjacent
-and water (as opposed to land or another physical barrier), and a
-diffusion rate $D$ defined in units of length of the side of a patch per
-unit of time. For example, if we are defining the time units as years
-and the distance units as kilometers, for a tuna $D$ might be 1,000
-$\frac{KM}{Year}$. The $\Delta_{t}$ and $\Delta_{d}$ parameters then
-translate the movement rate $D$ to match the time step and patch size
-used in a simulation. This "scale free" parameterization means that
-appropriate value of $D$ can be identified for a species and then set,
-regardless of the time step or patch size used in the simulation model
-itself. These parameters are then used to calculate the value of a
-particular entry the diffusion matrix (*d*).
-
-$$ 
-d_{p1,p2,t,s,a} = \begin{cases}
-      = \frac{\Delta_{t}}{\Delta_{d}^2}D & \text{if p2 and p1 are adjacent}\\
-     = -\sum_{p' \neq p1} d_{p1,p2,t,s,a} & \text{if p1 = p2}\\
+The individual components (*m*) of the movement matrix are filled based on an adjacency matrix, which defines whether two patches are both adjacent and water (as opposed to land or another physical barrier), a diffusion rate $D$ defined in units of area of a patch per unit of time, and a habitat preference function *h* in units of length of a side of a patch per unit time.
+For example, if we are defining the time units as years and the distance units as kilometers, for a tuna $D$ might be 1,000 $\frac{KM^2}{Year}$.
+We then use parameters $\Delta_{t}$ and $\Delta_{d}$ parameters to translate the diffusion rate $D$ to match the time step and patch size used in a simulation.
+This "scale free" parameterization means that appropriate value of $D$ can be identified for a species and then set, regardless of the time step or patch size used in the simulation model itself.
+The taxis component of the movement process is a function of the difference in habitat quality *h*.
+The habitat preference function itself can take any form the user wishes.
+Exponentiating the difference in the habitat preference function between patches turns the taxis matrix into a multiplier of the diffusion rate *D*.
+As such, when creating habitat layers for simulation, users can tune the scale of the habitat gradient function to result in realistic multipliers of the diffusion rate.
+This parameterization ensure that the off-diagonal elements of the movement matrix $\boldsymbol{M}_{t,s,a}$ are all non-negative, a requirement of the CTMC method.
+
+$$ 
+m_{p1,p2,t,s,a} = \begin{cases}
+      = \frac{\Delta_{t}}{\Delta_{d}^2}De^{\frac{\Delta_t(h(p2,t,s,a) - h(p1,t,s,a))}{\Delta_d}} & \text{if p2 and p1 are adjacent}\\
+     = -\sum_{p' \neq p1} m_{p1,p2,t,s,a} & \text{if p1 = p2}\\
      = 0 & \text{otherwise.}
@@
 
-The taxis matrix $\boldsymbol{\tau}$ is built off of the difference in
-log habitat quality between each patch (the habitat gradient), the
-adjacency matrix, and temporal and spatial resolution of the model
-($\Delta_{t}$ and $\Delta_{p}$), as well as a scalar term $\lambda$.
-$\lambda$ scales the magnitude of the habitat gradient between patches
-relative to the diffusion parameter $D$. When the diffusion and taxis
-matrices are estimated jointly from actual data in the manner of
-@thorson2021a, $\lambda$ should be set to zero. However, when users are
-manually creating both the diffusion and taxis matrices for simulation,
-the relative strength of diffusion to taxis must be defined by the user.
-Setting higher values of $\lambda$ allows users to simulate a scenario
-where taxis is much more significant than diffusion.
-
-$$
-\tau_{p1,p2,t,s,a} = \begin{cases}
-      = \frac{\Delta_{t}}{\Delta_{d}}(1 + \lambda d_{s,f})e^{log(h(p2,t,s,a) - log(h(p1,t,s,a))} & \text{if p2 and p1 are adjacent}\\
-     = -\sum_{p' \neq p1} \tau_{p1,p2,t,s,a} & \text{if p1 = p2}\\
-     = 0 & \text{otherwise.}
-  \end{cases}
-$$ {#eq-taxis}
-
-For both the diffusion and taxis matrices, we allow for the inclusion of
-physical barriers to movement (i.e. land). Pairs of patches that are
-adjacent but in which one or both patches are a barrier to movement are
-set as non-adjacent. The CTMC model then produces movement dynamics that
-move around barriers rather than over them. This is different than
-setting habitat in a patch to zero; an animal will not preferentially
-move towards a patch with zero habitat all else being equal, but can
-move over a patch with zero habitat towards another better patch. An
-example might be a species that lives and feeds on coastal habitats
-transiting through a patch of open ocean to reach a new feeding ground.
-In contrast, a physical barrier represents something like a peninsula
-that an animal must move around to transit from one side to another.
-
-While this CTMC approach to movement simulation is to date not commonly
-seen in the marine modeling literature, it has numerous advantages that
-warrant its seeming complexity. First, when only diffusion is present,
-the model will generalize to the familiar dynamics of a Gaussian
-dispersal kernel. However, the time component *t* in @eq-movement allows
-the movement model to be run on arbitrarily fine time steps. Second, the
-taxis model allow for more active habitat choices by species, allowing
-us to simulate preferences of species in space and time efficiently.
-Lastly, this form has the advantage that the parameters are directly
-estimable from real data. So, if provided with for example spatial
-abundance data and a tagging study, users can estimate the diffusion and
-taxis movement parameters in the same manner as @thorson2021a, and then
-pass their estimated parameters to `marlin` for simulation.
+For both the diffusion and taxis matrices, we allow for the inclusion of physical barriers to movement (i.e. land).
+Pairs of patches that are adjacent but in which one or both patches are a barrier to movement are set as non-adjacent.
+The CTMC model then produces movement dynamics that move around barriers rather than over them.
+This is different than setting habitat in a patch to zero; an animal will not preferentially move towards a patch with zero habitat all else being equal, but can move over a patch with zero habitat towards another better patch.
+An example might be a species that lives and feeds on coastal habitats transiting through a patch of open ocean to reach a new feeding ground.
+In contrast, a physical barrier represents something like a peninsula that an animal must move around to transit from one side to another.
+
+While this CTMC approach to movement simulation is to date not commonly seen in the marine modeling literature, it has numerous advantages that warrant its seeming complexity.
+First, when only diffusion is present, the model will generalize to the familiar dynamics of a Gaussian dispersal kernel.
+However, the time component *t* in @eq-movement allows the movement model to be run on arbitrarily fine time steps.
+Second, the taxis model allow for more active habitat choices by species, allowing us to simulate preferences of species in space and time efficiently.
+Lastly, this form has the advantage that the parameters are directly estimable from real data.
+So, if provided with for example spatial abundance data and a tagging study, users can estimate the diffusion and taxis movement parameters in the same manner as @thorson2021a, and then pass their estimated parameters to `marlin` for simulation (so long as the estimating method uses the same functional form as the movement model in `marlin`).
 
@@
 
-where *BH* is the Beverton-Holt recruitment function [@beverton1957] and
-*ssb* is spawning-stock-biomass. Per convention, the model allows for a
-"plus group", wherein rather than tracking numbers of every possible
-age, individuals greater than or equal to a given maximum age are
-grouped together.
-
-Spawning stock biomass *ssb* is calculated by converting age to mean
-length, calculating weight at age, maturity at age, and then calculating
-spawning stock biomass as the sum of spawning potential at age in a
-given time step, taking into account the potential for hyperallometry in
-the manner of @marshall2021. Age is converted to length through the von
-Bertalanffy growth equation given parameters asymptotic length
-($l_\infty$), growth (*k*) and theoretical age at length zero ($a0$)
-assuming log-normally distributed variation *u* in the length at age
-with CV $\sigma_s$.
+where *BH* is the Beverton-Holt recruitment function [@beverton1957] and *ssb* is spawning-stock-biomass.
+Per convention, the model allows for a "plus group", wherein rather than tracking numbers of every possible age, individuals greater than or equal to a given maximum age are grouped together.
+
+Spawning stock biomass *ssb* is calculated by converting age to mean length, calculating weight at age, maturity at age, and then calculating spawning stock biomass as the sum of spawning potential at age in a given time step, taking into account the potential for hyperallometry in the manner of @marshall2021.
+Age is converted to length through the von Bertalanffy growth equation given parameters asymptotic length ($l_\infty$), growth (*k*) and theoretical age at length zero ($a0$) assuming log-normally distributed variation *u* in the length at age with CV $\sigma_s$.
 
@@
 
-Biomass *b* at age is then given by the weight at length equation
-governed by a scaling coefficient $\Omega_s$ and an exponent $\Phi_s$
-that controls the rate at which volume scales with length
+Biomass *b* at age is then given by the weight at length equation governed by a scaling coefficient $\Omega_s$ and an exponent $\Phi_s$ that controls the rate at which volume scales with length
 
@@
 
-The proportion of sexually mature individuals (*mat*) at a given age is
-then calculated as a logistic function where $l_{mat}$ is the length at
-which on average 50% of individuals are sexually mature, and
-$\delta_{mat}$ is the unit of length beyond $l_{mat}$ at which on
-average 95% of fish are sexually mature.
+The proportion of sexually mature individuals (*mat*) at a given age is then calculated as a logistic function where $l_{mat}$ is the length at which on average 50% of individuals are sexually mature, and $\delta_{mat}$ is the unit of length beyond $l_{mat}$ at which on average 95% of fish are sexually mature.
 
@@
 
-Spawning stock biomass at time *t* is then calculated as a function of
-the numbers at age, the maturity at age, and the weight at age raised by
-a parameter $\gamma$. When $\gamma$ is greater than 1, the species is
-said to experience hyperallometric fecundity, i.e. fecundity increases
-faster than weight.
+Spawning stock biomass at time *t* is then calculated as a function of the numbers at age, the maturity at age, and the weight at age raised by a parameter $\gamma$.
+When $\gamma$ is greater than 1, the species is said to experience hyperallometric fecundity, i.e. fecundity increases faster than weight.
 
@@
 
-Recruitment (i.e. the number of age 1 individuals entering the
-population) follows Beverton-Holt dynamics parameterized around
-steepness (*h*) with log-normally distributed recruitment deviates
-$\epsilon$. When steepness is one recruitment is independent of spawning
-biomass. As steepness approaches 0.2 recruitment becomes a linear
-function of spawning biomass. We allow for five variants in the timing
-of density dependent recruitment:
-
-1.  Global density dependence: Density dependent recruitment is a
-    function of the sum of spawning biomass across all patches, and
-    recruits are then distributed according to habitat quality
+Recruitment (i.e.
+the number of age 1 individuals entering the population) follows Beverton-Holt dynamics parameterized around steepness (*h*) with log-normally distributed recruitment deviates $\epsilon$.
+When steepness is one recruitment is independent of spawning biomass.
+As steepness approaches 0.2 recruitment becomes a linear function of spawning biomass.
+We allow for five variants in the timing of density dependent recruitment, building off of @babcock2011 :
+
+1.  Global density dependence: Density dependent recruitment is a function of the sum of spawning biomass across all patches, and recruits are then distributed according to habitat quality
 
@@
 
-where $r0$ is is a vector of recruits under unfished conditions in a
-given patch.
-
-2.  Local density dependence: Density dependent recruitment occurs
-    independently in each patch and recruits are retained in their home
-    patch.
+where $r0$ is is a vector of recruits under unfished conditions in a given patch.
+
+2.  Local density dependence: Density dependent recruitment occurs independently in each patch and recruits are retained in their home patch.
 
@@
 
-3.  Local density dependence then disperse: Density dependent
-    recruitment occurs independently in each patch and recruits are then
-    dispersed.
+3.  Local density dependence then disperse: Density dependent recruitment occurs independently in each patch and recruits are then dispersed.
 
@@
 
-4.  Post-dispersal density dependence: Larvae are distributed throughout
-    the system, and then density dependent recruitment occurs based on
-    the density of adult biomass at the destination patch.
+4.  Post-dispersal density dependence: Larvae are distributed throughout the system, and then density dependent recruitment occurs based on the density of spawning biomass at the destination patch.
 
@@
 
-5.  Global density dependence allocated by spawning biomass: Density
-    dependence is a function of the sum of spawning biomass across all
-    patches, and recruits are then distributed according to the
-    distribution of spawning biomass
+5.
+Global density dependence allocated by spawning biomass: Density dependence is a function of the sum of spawning biomass across all patches, and recruits are then distributed according to the distribution of spawning biomass
 
@@
 
-Log-normal recruitment deviates are calculated with the potential for
-autocorrelation defined with strength $\rho$
+Log-normal recruitment deviates are calculated with the potential for autocorrelation defined with strength $\rho$
 
@@
 
-And log recruitment deviates are converted to raw units using the bias
-correction factor
+And log recruitment deviates are converted to raw units using the bias correction factor
 
@@
 
-Fisheries management is often concerned with measuring stock status
-relative to maximum sustainable yield (MSY) based reference points,
-though the exact level of stock status desired by management or societal
-objectives may vary widely. MSY based reference points present a problem
-for a multi-fleet and spatial-temporal model such as `marlin`. MSY and
-the fishing mortality rate that would produce MSY, $F_{MSY}$, are a
-function of fishery selectivity. Fishery selectivity in this model can
-vary by fleet, and species can be distributed unevenly in both space and
-time. This means that the net effective fishing selectivity on a species
-can vary depending on the dynamics at a given moment, making the
-definition of an equilibrium concept such as MSY challenging.
-
-As such, we do not report MSY based reference points in the model by
-default, with the exception of the spawning stock biomass at MSY
-($SSB_{MSY}$) which tends to be conserved across selectivity patterns as
-it is primarily a function of the mortality and flow of recruits
-relative to the maturity ogive. There are many different strategies for
-estimating reference points in spatially explicit systems [@kapur2021a].
-We leave it to users to define and find relevant reference point as
-required by their specific needs.
+Fisheries management is often concerned with measuring stock status relative to maximum sustainable yield (MSY) based reference points, though the exact level of stock status relative to MSY reference points desired by management or societal objectives may vary widely.
+MSY based reference points present a problem for a multi-fleet and spatial-temporal model such as `marlin`.
+MSY and the fishing mortality rate that would produce MSY, $F_{MSY}$, are a function of fishery selectivity.
+Fishery selectivity in this model can vary by fleet, and species can be distributed unevenly in both space and time.
+This means that the net effective fishing selectivity on a species can vary depending on the dynamics at a given moment, making the definition of an equilibrium concept such as MSY challenging.
+
+As such, we do not report MSY based reference points in the model by default, with the exception of the spawning stock biomass at MSY ($SSB_{MSY}$) which tends to be conserved across selectivity patterns as it is primarily a function of the mortality and flow of recruits relative to the maturity ogive.
+There are many different strategies for estimating reference points in spatially explicit systems [@kapur2021a].
+We leave it to users to define and find relevant reference point as required by their specific needs.
 
@@
 
-We include two examples demonstrating how `marlin` can be used to
-support marine spatial planning. In the first, we show how `marlin` can
-be used to compare the total and distributional impacts of MPAs designed
-in a heavily fished coastal coral reef ecosystem. In the second, we
-demonstrate how `marlin` can be used to assess components of climate
-resilience of alternative MPA design strategies in a pelagic system.
-
-Each of the case studies contains too many parameters and options to be
-succinctly presented in the text here, and we are concerned that
-attempting to duplicate all of the parameters of the model here would
-simply introduce error. As such, readers should consult the accompanying
-code to view the precise details of each simulation. Targeted tactical
-applications must carefully consider and document all decisions made
-around model parameters.
+We include two examples demonstrating how `marlin` can be used to support marine spatial planning.
+In the first, we show how `marlin` can be used to compare the total and distributional impacts of MPAs designed in a heavily fished coastal coral reef ecosystem.
+In the second, we demonstrate how `marlin` can be used to assess components of climate resilience of alternative MPA design strategies in a pelagic system.
+
+Each of the case studies contains too many parameters and options to be succinctly presented in the text here, and we are concerned that attempting to duplicate all of the parameters of the model within the text would simply introduce errors and confusion.
+As such, and since these results are for illustrative purposes only, readers should consult the accompanying code to view the precise details of each simulation.
+Targeted tactical applications must carefully consider and document all decisions made around model parameters.
 
@@
 
-Across all of our case studies we make use of two potential "rule of
-thumb" MPA design strategies:
-
-1.  *Rate*: MPAs are placed based on the pre-MPA SSB/SSB0 weighted catch
-    relative to the total catch in a patch. So, patches with high rates
-    of catch of depleted species relative to total catch are
-    prioritized.
-
-2.  *Target Fishing*: MPAs locations are prioritized proportional to
-    fishery catches. Patches with high total catches are prioritized
-    over patches with low catches.
-
-To our knowledge, MPAs are never randomly placed; their locations are
-decided based on stakeholder negotiation given available information and
-objectives. In theory, the design of MPA networks can be optimized
-through the use of a modeling framework, and depending on the validity
-of the model, this process is likely to produce better outcomes than
-manually-designed strategies [@rassweiler2012; @rassweiler2014].
-However, designing optimal MPA networks becomes increasingly difficult
-as the range of objectives and the complexity of the model increase.
-Therefore, while `marlin` provides some support for
-[optimization](https://danovando.github.io/marlin/articles/optimize-mpas.html),
-we focus here on the design and performance of these more "rule of
-thumb" design strategies that may be more accessible to a wider range of
-users. We allow all MPAs to be designed in a "mosaic" fashion, but users
-can easily extend the analysis to compare outcomes between "contiguous"
-(MPA is made up of one continuously connected block) and "mosaic" (MPAs
-can be separated in space) MPA designs [@pons2022] by developing and
-passing their own MPA design algorithms to the model.
-
-#### Coral Reefs
-
-In our coral reef example, we model the dynamics of four tropical
-Pacific species: a grouper (*Epinephelus fuscoguttatus*), a shallow-reef
-snapper (*Lutjanus malabaricus*), a deep-water snapper (*Pristipomoides
-filamentosus*), and a reef shark (*Carcharhinus amblyrhynchos*). The
-simulated groupers undergo a mass migration to a spawning aggregation
-once per year, followed by the sharks. Shallow-reef snappers stay in
-reefs closer to shore year-round, while deep-water snappers stay in the
-deeper reefs (@fig-coral).
-
-These species are targeted by two different fleets. Fleet One primarily
-targets the grouper and nearshore snapper populations, but will land any
-incidentally captured sharks. Fleet One has a dome shaped selectivity
-pattern for both grouper and shallow-reef snapper due to demand for
-"plate sized" fish [@kindsvater2017], and logistic selectivity for
-sharks, as any shark caught is retained. Fleet One is totally dependent
-on fishing for their livelihood, meaning the local community takes
-advantage of every possible opportunity to fish, and as such we model it
-as a "constant effort" fishery.
-
-Fleet Two primarily targets the deep-water snapper population. While
-plate-sized snapper are still prized from this population, we model
-Fleet Two's selectivity as logistic due to high levels of discard
-mortality for deep-water snapper resulting from barotrauma. Fleet Two
-catches some grouper, though less than Fleet One, and receives no price
-for sharks due to the requirements of a certification program through
-which they sell their deep-water snapper. Accidental captures (bycatch)
-do occur, which results in mortality. Fleet Two operates under
-open-access dynamics, as fishing is not the primary means of
-subsistence; short-term effort expands and contracts in response to
-profitability of the primarily grouper-driven fishery.
-
-Each fleet has a different home-port location, and incurs an additional
-cost for fishing each patch in the system in proportion to the Euclidean
-distance from each patch to each fleet's home port.
-
-We use `marlin` to simulate the outcomes for both food production and
-conservation for each of the species and both of the fleets as a
-function of both MPA size and MPA design strategy. For this exercise,
-MPAs are placed with perfect information and have no design constraints
-for continuity (i.e. the selected MPA design can be as patchwork as
-dictated by the data). We ran the simulation on a quarterly level
-($\Delta_t = 1/4$) and set the area of each patch to an arbitrary value
-of 1 ($\Delta_p = 1$).
+We make use of three potential "rule of thumb" MPA design strategies:
+
+1.  *Rate*: MPAs are placed based on the pre-MPA SSB/SSB0 weighted catch relative to the total catch in a patch.
+    So, patches with high rates of catch of depleted species relative to total catch are prioritized.
+
+2.  *Target Fishing*: MPAs locations are prioritized proportional to fishery catches.
+    Patches with high total catches are prioritized over patches with low catches.
+
+3.  *Spawning Aggregation*: MPAs are centered on the location of a known spawning aggregation
+
+To our knowledge, MPAs are never randomly placed; their locations are decided based on stakeholder negotiation given available information and objectives.
+In theory, the design of MPA networks can be optimized through the use of a modeling framework, and depending on the validity of the model, this process is likely to produce better outcomes than manually-designed strategies [@rassweiler2012; @rassweiler2014].
+However, designing optimal MPA networks becomes increasingly difficult as the range of objectives and the complexity of the model increase.
+Therefore, while `marlin` provides some support for [optimization](https://danovando.github.io/marlin/articles/optimize-mpas.html), we focus here on the design and performance of these more "rule of thumb" design strategies that may be more accessible to a wider range of users.
+We allow all MPAs to be designed in a "mosaic" fashion in these examples, but users can easily extend the analysis to compare outcomes between "contiguous" (MPA is made up of one continuously connected block) and "mosaic" (MPAs can be separated in space) MPA designs [@pons2022] by developing and passing their own MPA design algorithms to the model.
+
+#### Coastal Coral Reef Fishery
+
+In our coastal coral reef example, we model the dynamics of four tropical Pacific species: a grouper (*Epinephelus fuscoguttatus*), a shallow-reef snapper (*Lutjanus malabaricus*), a deep-water snapper (*Pristipomoides filamentosus*), and a reef shark (*Carcharhinus amblyrhynchos*).
+The simulated groupers undergo a mass migration to a spawning aggregation once per year, followed by the sharks.
+Shallow-reef snappers stay in reefs closer to shore above a steep drop-off year-round, while deep-water snappers stay in the deeper reefs past the drop-off (@fig-coral).
+
+These species are targeted by two different fleets.
+Fleet One primarily targets the grouper and near-shore snapper populations, but will land any incidentally captured sharks.
+Fleet One has a logistic selectivity pattern for all species, as they retain any fish caught for consumption or sale.
+both grouper and shallow-reef snapper due to demand for "plate sized" fish , and logistic selectivity for sharks, as any shark caught is retained.
+Fleet One is totally dependent on fishing for their livelihood, meaning the local community takes advantage of every possible opportunity to fish, and as such we model it as a "constant effort" fishery.
+Do to having less efficient boats, Fleet One has a higher cost per distance coefficient than Fleet Two.
+Fleet One's home port is located on the site of the grouper spawning aggregation.
+
+Fleet Two is a more commercial fleet that primarily targets the deep-water snapper population.
+This fleet primarily sells their catch to local restaurants and distributors where plate-sized fish are prized, and so for both snapper and grouper Fleet Two has a dome-shaped selectivity pattern [@kindsvater2017].
+While plate-sized deep snapper are the primary target of Fleet Two, we model Fleet Two's selectivity for deep snapper as logistic due to high levels of discard mortality for deep-water snapper resulting from barotrauma.
+Fleet Two catches groupers, though less than Fleet One, and receives no price for sharks due to the requirements of a certification program through which they sell their deep-water snapper.
+Accidental captures (bycatch) of sharks do occur, which results in mortality.
+Fleet Two operates under open-access dynamics, as fishing is not the only means of subsistence for this community; short-term effort expands and contracts in response to profitability of the primarily grouper-driven fishery.
+Fleet two's home port is located further north from Fleet One.
+
+We use `marlin` to simulate the outcomes for both food production and conservation for each of the species and both of the fleets as a function of both MPA size and MPA design strategy.
+For this exercise, MPAs are placed with perfect information and have no design constraints for continuity (i.e. the selected MPA design can be as patchwork as dictated by the data).
+We ran the simulation on a quarterly level ($\Delta_t = 1/4$) and set the area of each patch to be 5km^2^ ($\Delta_p = \sqrt{5}$), using 400 patches for a total area of 2,000 KM^2^.
 
@@
 
-We model our pelagic use case loosely on the characteristics of the
-Western and Central Pacific Ocean (WCPO) tuna fisheries. We simulate
-trajectories of `r n_distinct(names(blue_fauna))` species commonly
-caught in the region, including both the highly abundant skipjack tuna
-and the heavily depleted oceanic whitetip shark (*Carcharhinus
-longimanus*) (@fig-blue-ssb). We use publicly available data on
-catch-per-unit-effort of each of these species from the WCPO as a proxy
-for baseline habitat distributions, noting that where possible,
-fishery-independent abundance indices would be preferable (@fig-blue).
-
-These pelagic species are caught by a longline fleet that primarily
-targets large adult tunas such as bigeye and yellowfin for high-grade
-consumption, and a purse-seine fishery that primarily targets skipjack
-tunas for bulk canning. Selectivities were modeled as logistic for the
-longline fleet, and dome-shaped for the purse-seine fleet, with
-selectivity values set to broadly match the patterns reported in
-relevant [stock
-assessments](https://oceanfish.spc.int/en/ofpsection/sam/sam/210-stock-assessment-input-and-result-files-)
-for each species. Both fleets operate under open-access dynamics with an
-effort cap. The effort cap was set at the level of effort that resulted
-in the desired levels of SSB/SSB0 for each species under open-access
-dynamics (@fig-blue), intended to simulate a scenario where managers
-step in to prevent further expansion of fishing effort in a fully
-developed fishery. For forward-simulation, open-access dynamics can
-result in effort decreasing in response to profitability, but cannot
-result in effort beyond the effort cap set for each fleet.
-
-For this exercise, we focused on using `marlin` to assess resilience of
-the selected "rule of thumb" MPA design strategies to a climate-driven
-range shift. Specifically, we simulate an extreme example where the
-centroid of each population shifts northward at a rate of 0.2 patches
-per year, over a 20 year time horizon (@fig-blue-ssb). We designed MPA
-networks given the conditions in the starting year, and then held that
-network constant over the years of the experiment, running one
-simulation with and another without the climate-drive range shift. We
-then compared the effects of this range shift on food production and
-conservation outcomes from MPA networks designed based on the pre-range
-shift world. We ran the simulation on a quarterly level
-($\Delta_t = 1/4$) and set the area of each patch to an arbitrary value
-of 1 ($\Delta_p = 1$).
+We model our pelagic use case loosely on the characteristics of the Western and Central Pacific Ocean (WCPO) tuna fisheries.
+We simulate trajectories of `r n_distinct(names(blue_fauna))` species commonly caught in the region, including both the highly abundant skipjack tuna and the heavily depleted oceanic whitetip shark (*Carcharhinus longimanus*) (@fig-blue-ssb).
+We use publicly available data on catch-per-unit-effort of each of these species from the WCPO as a very rough proxy for baseline habitat distributions, noting that where possible, fishery-independent abundance indices would be preferable (@fig-blue).
+
+These pelagic species are caught by a longline fleet that primarily targets large adult tunas such as bigeye and yellowfin for high-grade consumption, and a purse seine fishery that primarily targets skipjack tunas for bulk canning.
+Selectivities were modeled as logistic for the longline fleet, and dome-shaped for the purse seine fleet, with selectivity values set to broadly match the patterns reported in relevant [stock assessments](https://oceanfish.spc.int/en/ofpsection/sam/sam/210-stock-assessment-input-and-result-files-) for each species.
+Both fleets operate under open-access dynamics with an effort cap.
+The effort cap was set at the level of effort that resulted in the desired levels of SSB/SSB0 for each species under open-access dynamics (@fig-blue), intended to simulate a scenario where managers step in to prevent further expansion of fishing effort in a fully developed fishery.
+For forward-simulation, open-access dynamics can result in effort decreasing in response to profitability, but cannot result in effort beyond the effort cap set for each fleet.
+
+For this exercise, we focused on using `marlin` to assess resilience of the selected Target Fishing and Rate "rule of thumb" MPA design strategies to a climate-driven range shift.
+Specifically, we simulate an extreme example where the centroid of each population shifts northward at a rate of 0.33 patches per year, over a 20 year time horizon (@fig-blue-ssb).
+We designed MPA networks given the conditions in the starting year, and then held that network constant over the years of the experiment, running one simulation with and another without the climate-drive range shift.
+We then compared the effects of this range shift on food production and conservation outcomes from MPA networks designed based on the pre-range shift world.
+We ran the pelagic simulation at a quarterly level ($\Delta_t$ = 1/4 year), and set the area of each cell to be roughly 97,000 KM^2^ across 144 patches each with a side length of roughly 311 KM, for a total area of 14,000,000 KM^2^, broadly commensurate with the area of the WCPO.
 
@@
   mutate(depletion = ssb / ssb[step == 0])
-
-# critter_biomass %>%
-#   ungroup() %>% 
-#   left_join(sci_to_com, by = "critter") %>%
-#   select(-critter) %>%
-#   rename(critter = common_name) %>%
-#   ggplot(aes(step, depletion, color = critter)) +
-#   geom_line(linewidth = 1) +
-#   scale_y_continuous(limits = c(0, 1), name = "Depletion", expand = expansion(mult = c(0.0,0.05))) +
-#   scale_x_continuous(name = "Time Step") +
-#   scale_color_discrete(name = "")
 
@@
 
-For the coastal coral reef example the *Target fishing* MPA design
-strategies was capable of producing a Pareto-improving, i.e. potentially
-win-win, outcome where the total spawning biomass divided by unfished
-spawning biomass across all species increase by up to 200% while
-maintaining at least the same levels of fisheries yields as would be
-caught in the absence of any MPAs. Conversely, achieving increases in
-total fish biomass through the *Rate* strategy resulted in immediate
-losses in total yields, and in losses to both yields and total biomass
-in some cases (@fig-coral-results).
-
-These macro-level changes in total yield and conservation outcomes mask
-some internal tradeoffs. Nearly all of the food production benefits from
-the simulated MPA networks accrued to fleet one, while catches by fleet
-two were generally negatively impacted under both scenarios. For
-example, the MPA sizes and designs that result in the largest increase
-in yields for fleet one would produce a near total loss of fish catch
-for fleet two.
-
-The *Rate* MPA placement strategy produced lower total fishery gains but
-minimized the negative impacts on fleet two. However, this strategy
-produced much lower conservation gains than the *Target Fishing*
-strategy, only benefiting deep snapper meaningfully, and in fact
-resulting in slight losses to the nearshore snapper and grouper
-populations for MPA sizes less than roughly 75% of the simulated space.
-The *Target Fishing* strategy resulted in rapid conservation gains
-across all of the simulated species.
-
-MPAs were capable of producing positive outcomes in terms of total
-biomass and yields in the coastal coral reef example, but all the tested
-strategies resulted in tradeoffs across at least one dimension. `marlin`
-could be further used by a community such as the one simulated here to
-explore alternative design strategies, or to assess options for
-distributing fishery benefits across communities that may experience
-very different food provision outcomes from the same MPA network.
+For the coastal coral reef example both the *Spawning Ground* and *Target fishing* MPA design strategies were capable of producing a Pareto-optimal or improving, i.e.
+potentially win-win, outcome where change in total spawning biomass across all species and yield across all fleets was greater than or equal to zero for some MPA size.
+However, producing a total increase in SSB/SSB0 greater than 50% came at a net cost to fisheries yield across both design strategies.
+The *Spawning Aggregation* strategy produced more immediate trade-offs between food security and conservation (@fig-coral-results).
+
+These macro-level changes in total yield and conservation outcomes mask interesting internal dynamics.
+The *Spawning Ground* strategy increased yield to Fleet One when MPAs were relatively small, but was able to maintain yields near *status quo* levels up to a very large MPA size to the dependence of Fleet One on only a relatively small portion of the total simulation area.
+However, *Spawning Ground* MPAs were unable to provide increases in yield to Fleet Two.
+The *Spawning Ground* strategy produced the most rapid conservation gains for the snapper population.
+Interestingly, due to effort displacement from the spawning grounds onto the fishing areas inhabited by the groupers the rest of the year, closure of the spawning aggregation had limited impact on the grouper population until the MPA size exceeded 50% of the simulation area.
+Effort displacement from the spawning grounds also resulted in a small decrease in the deep snapper population for MPA sizes less than 50%.
+The *Target Fishing* design strategy has similar yield outcomes to the *Spawning Ground* strategy, but less dramatic negative impacts on Fleet Two, at the expense of slightly fewer benefits for Fleet One.
+However, the *Target Fishing* strategy provided more uniform conservation benefits to all species (@fig-coral-results).
+
+MPAs were capable of producing positive outcomes in terms of total biomass and yields in the coastal coral reef example, but all the tested strategies resulted in trade-offs across at least one dimension.
+`marlin` could be further used by a community such as the one simulated here to explore alternative design strategies, or to assess options for distributing fishery benefits across communities that may experience very different food provision outcomes from the same MPA network.
 
@@
 #| label: fig-coral-results
-#| fig-cap: "Percent change in yield per fleet (A) and SSb/SSB0 (spawning biomass divided by unfished spawning biomass) per species (B) as a function of MPA size and MPA design strategy for the coastal coral reef case study.
+#| fig-cap: "Percent change in yield per fleet (A) and SSB/SSB0 (spawning biomass divided by unfished spawning biomass) per species (B) as a function of MPA size and MPA design strategy for the coastal coral reef case study.
 Percent change in total yield across both fleets (y-axis) and total SSB/SSB0 across all species (x-axis) and placement strategies (line color) (C).
-Text labels in panel C indicate the percent of area in an MPA at selected points along curves."
+Text labels in panel C indicate the percent of the simulation area in an MPA when change in yields is maximized, is 0, and is -50%.
+All results reflect outcomes after 20 years of simulated MPA protection."
 #| fig-height: 4
@@
   geom_line(alpha = 0.8, aes(color = placement_strategy),size = 1.1) + 
-  geom_text_repel(data = frontier_labels, aes(label = pmpa), min.segment.length = 0, box.padding = 0.5) +
+  geom_text_repel(data = frontier_labels, aes(label = pmpa, color = placement_strategy), min.segment.length = 0, box.padding = 0.2, show.legend = FALSE) +
   scale_x_continuous(name = "Change in Total SSB/SSB0", labels = scales::label_percent(accuracy = 1), guide = guide_axis(n.dodge = 2)) +
@@
   theme(
-    legend.position = c(0.725, .95),
+    legend.position = c(0.7, .85),
     legend.background = element_rect(fill = "transparent"),
@@
 
-((coral_fleet_yield / coral_critter_bio) | (coral_frontier + plot_layout(guides = "keep")) ) & theme(strip.text = element_text(size  = 8), plot.margin = margin(.1, .1, .1, .1, "cm"), panel.spacing = unit(1, "lines"), legend.text = element_text(size = 7))
+((coral_fleet_yield / coral_critter_bio) | (coral_frontier + plot_layout(guides = "keep")) ) & theme(strip.text = element_text(size  = 6), plot.margin = margin(.1, .1, .1, .1, "cm"), panel.spacing = unit(0.5, "lines"), legend.text = element_text(size = 6))
 
@@
 
-The yield impacts of the proposed MPA strategies were much more
-sensitive to climate impacts for the purse seine fishery than for the longline fishery. Under the
-*status quo* species distribution, the *Target Fishing* MPA design
-strategy produced increases in fisheries yields when the MPA covered
-25-50% of the simulated area; that same MPA network resulted in the near
-complete loss of yields from the purse seine fleet when applied to the
-climate change scenario. In contrast, the yield outcomes for the
-longline fleet were relatively insensitive to the simulated range shifts
-(@fig-blue-results).
-
-From a conservation perspective, spawning biomass of species such as
-albacore and swordfish was relatively insensitive to the simulated range
-shifts, whereas species such as bigeye tuna and silky shark showed
-extreme variation in outcomes depending on the climate scenario. For
-bigeye tuna, under the *status quo* conditions the *Target Fishing* MPA
-design strategy resulted in rapid increases in the population, but that
-same MPA network produced much smaller and slower gains under climate
-change. However, for blue sharks this trend is inverted, with more rapid
-conservation gains under the range shift scenario than the *status quo*
-(@fig-blue-results).
+The yield impacts of the proposed MPA strategies varied across each of our simulations.
+Yields of the purse seine fleet were more resilient to the range shift under the *Rate* design strategy, while the longline fleet had the exact opposite result.
+Under the *Rate* strategy, longline yields were relatively stable across a large MPA size under the *status quo* conditions, but declined rapidly under the *range shift* conditions.
+For the purse seine fleet, yields were more stable under the *range shift* conditions under the *Target Fishing* design scenario, but declined quickly as a function of MPA size under the *status quo* conditions (@fig-blue-results).
+In terms of fisheries yield, the *Target Fishing* strategy would produce better yield outcomes for both fishing fleets under the *range shift* conditions, but the *Rate* strategy would perform best under the *status quo*.
+
+From a conservation perspective, the primary tuna species (bigeye, skipjack, and yellowfin) were most sensitive to design strategy and climate scenario, with the *Target Fishing* strategies producing for example rapid conservation gains for bigeye under the *status quo*, while the *Rate* strategy resulted in net conservation losses until the MPA became extremely large.
+However, under the *Range Shift* scenario the MPAs had little impact on many of the tuna populations until the MPA size became extremely large, do to movement of the primary fishing grounds outside of the current hotpots.
+The 20 year time horizon simulated here was not enough to produce substantial gains for any of the shark species even with 100% closures except for the more rapidly growing blue shark (@fig-blue-results).
 
@@
 #| fig-cap: "Change in yield (A) and SSB/SSB0 (spawning biomass divided by unfished spawning biomass) (B) as a function of MPA size and design strategy by fishing fleet and species.
-Blue lines indicate impacts of MPAs under status qup habitat, red impacts under climate-driven range shift."
+Blue lines indicate impacts of MPAs under status quo habitat, red impacts under climate-driven range shift.
+Results reflect the outcome of 20 years of simulated MPA protection."
 #| fig-height: 6
@@
 Marine ecosystems are driven by complex social-ecological dynamics.
-Communities must often make decisions on how to manage these systems
-based on limited empirical evidence. Modeling tools such as `marlin` can
-help users answer important scientific questions and design marine
-management policies informed by a better understanding of sensitivities
-to key uncertainties. In this paper we have documented the core system
-of equations underlying `marlin`, and demonstrated how `marlin` can be
-used to support efforts such as MPA planning. The variation in both the
-magnitude and distribution of costs and benefits stemming from
-alternative design choices and states of nature shown in our case study
-results demonstrates the value in our method. For example, our coastal
-coral reef example illustrates both the potential for MPAs to benefit
-conservation and food production in these systems, and the potential for
-the same MPA to benefit some fleets and species while harming others
-when fishing fleets affected by an MPA do not share the same objectives
-and species are not uniformly distributed. Our pelagic case study
-illustrates how `marlin` can be used to assess the climate resiliency of
-spatial management strategies, and identify strategies that best meet
-the needs of both current and future conditions. Results from this type
-of work could be used to help prioritize communities and species at
-particular risk to climate change impacts.
-
-The role of this paper is not to conduct a review of the many tools for
-spatial-temporal modeling available in the literature, each of which
-provide useful functionality for specific applications. However, we
-highlight here the specific gaps that we feel `marlin` fills in the
-modeling literature and in the policy support toolbox using some
-selected publications. More end-to-end models such as ATLANTIS
-[@audzijonyte2019] , OSMOSE [@shin2001], Ecopath with Ecosim
-[@christensen2004], POSEIDON [@bailey2019], DISPLACE [@bastardie2013],
-or SEAPODYM [@lehodey2008] are capable of representing tremendous
-amounts of complexity, but can be extremely time consuming to design and
-run. `marlin` allows some of the realism of these end-to-end models
-while being simpler and faster to parameterize and run.
-
-Tools such as virtualspecies [@leroy2016], STEPS [@visintin2020],
-RangeShifter [@bocedi2014], and
-[SMS](https://www.aquaveo.com/software/sms-surface-water-modeling-system-introduction)
-can be efficiently constructed to model the dispersal and distribution
-of species in space and time as a function of environmental covariates.
-However, these typically require specifying a covariate-response curve
-without explicitly acknowledging how this arises from habitat-specific
-movement and demography (virtualspecies), or model movement using a
-dispersal kernel (RangeShifter, STEPS) or a least-cost path algorithm
-(SMS, STEPS). By contrast, `marlin` uses a continuous-time Markov chain
-movement model, which integrates multiple paths (including their
-path-dependent probability based on intervening habitat types and
-species preferences) while using scale-free parameters that can be
-measured experimentally in laboratory or tagging studies.
-
-Simplified bio-economic models such as those used in @hastings2017,
-@sala2021, and @cabral2019 can be applied at scale and provide
-analytically tractable results, but as a result must abstract over many
-bio-economic dynamics that can be important for more tactical
-applications. `marlin` allows for more realistic representations of
-spatial social-ecological systems while maintaining processing speed.
-
-In sum, `marlin` complements the existing spatial marine modeling
-literature by allowing scientists, decision makers, and other
-stakeholders to efficiently and productively examine the impacts of
-realistic bio-economic dynamics on academic and applied problems. For
-example, scientists could use `marlin` to consider the relative
-importance of different biological variables in affecting policy
-outcomes. Managers could use `marlin` to assess the potential of
-proposed MPA networks to achieve stated objectives, such as maximizing
-conservation gains for a threatened species while minimizing or removing
-negative impacts to affected fishing fleets.
-
-Beyond the current examples, we envision `marlin` being applicable to
-research on dynamic ocean management, range shifts, management strategy
-evaluation, policy interactions, and spatial stock assessment. In
-addition, the process-based movement model used in `marlin` can directly
-use empirical estimates of movement dynamics derived in the manner of
-@thorson2021a, providing a link between empirical and simulation based
-approaches to marine resource management. Future iterations of the model
-may include life history modules for a broader range of species types
-such as marine mammals, sea birds, and sea turtles.
-
-Even the most complex marine model is a stylized cartoon of the true
-dynamics at play in the ocean. However, for all their limitations models
-such as `marlin` can help users understand factors that drive the
-performance of marine management strategies. The goal of this tool is to
-empower people to design policies based on evaluation of key
-uncertainties and trade-offs, and in doing so support more effective and
-equitable marine resource management.
+Communities must often make decisions on how to manage these systems based on limited empirical evidence.
+Modeling tools such as `marlin` can help users answer important scientific questions and design marine management policies informed by a better understanding of sensitivities to key uncertainties.
+In this paper we have documented the core system of equations underlying `marlin`, and demonstrated how `marlin` can be used to support efforts such as MPA planning.
+The variation in both the magnitude and distribution of costs and benefits stemming from alternative design choices and states of nature shown in our case study results demonstrates the value in our method.
+For example, our coastal coral reef example illustrates both the potential for MPAs to benefit conservation and food production in these systems, and the potential for the same MPA to benefit some fleets and species while harming others when fishing fleets affected by an MPA do not share the same objectives and species are not uniformly distributed.
+Our pelagic case study illustrates how `marlin` can be used to assess the climate resiliency of spatial management strategies, and identify strategies that best meet the needs of both current and future conditions.
+Results from this type of work could be used to help prioritize communities and species at particular risk to climate change impacts.
+
+The role of this paper is not to conduct a review of the many tools for spatial-temporal modeling available in the literature, each of which provide useful functionality for specific applications.
+However, we highlight here the specific gaps that we feel `marlin` fills in the modeling literature and in the policy support toolbox using some selected publications.
+More end-to-end models such as ATLANTIS [@audzijonyte2019] , OSMOSE [@shin2001], Ecopath with Ecosim [@christensen2004], POSEIDON [@bailey2019], DISPLACE [@bastardie2013], or SEAPODYM [@lehodey2008] are capable of representing tremendous amounts of complexity, but can be extremely time consuming to design and run.
+`marlin` allows some of the realism of these end-to-end models while being simpler and faster to parameterize and run.
+
+Tools such as virtualspecies [@leroy2016], STEPS [@visintin2020], RangeShifter [@bocedi2014], and [SMS](https://www.aquaveo.com/software/sms-surface-water-modeling-system-introduction) can be efficiently constructed to model the dispersal and distribution of species in space and time as a function of environmental covariates.
+However, these typically require specifying a covariate-response curve without explicitly acknowledging how this arises from habitat-specific movement and demography (virtualspecies), or model movement using a dispersal kernel (RangeShifter, STEPS) or a least-cost path algorithm (SMS, STEPS).
+By contrast, `marlin` uses a continuous-time Markov chain movement model, which integrates multiple paths (including their path-dependent probability based on intervening habitat types and species preferences) while using scale-free parameters that can be measured experimentally in laboratory or tagging studies.
+
+Simplified bio-economic models such as those used in @hastings2017, @sala2021, and @cabral2019 can be applied at scale and provide analytically tractable results, but as a result must abstract over many bio-economic dynamics that can be important for more tactical applications.
+Our results demonstrate how nuances in fleet dynamics and species distributions can dramatically impact MPA outcomes.
+`marlin` allows for more realistic representations of spatial social-ecological systems while maintaining processing speed.
+
+In sum, `marlin` complements the existing spatial marine modeling literature by allowing scientists, decision makers, and other stakeholders to efficiently examine the impacts of realistic bio-economic dynamics on academic and applied problems.
+For example, scientists could use `marlin` to consider the relative importance of different biological variables in affecting policy outcomes.
+Managers could use `marlin` to assess the potential of proposed MPA networks to achieve stated objectives, such as maximizing conservation gains for a threatened species while minimizing or removing negative impacts to affected fishing fleets.
+
+Beyond the current examples, we envision `marlin` being applicable to research on dynamic ocean management, range shifts, management strategy evaluation, policy interactions, and spatial stock assessment.
+In addition, the process-based movement model used in `marlin` can directly use empirical estimates of movement dynamics derived in the manner of @thorson2021a, providing a link between empirical and simulation based approaches to marine resource management.
+Future iterations of the model may include life history modules for a broader range of species types such as marine mammals, sea birds, and sea turtles.
+
+Even the most complex marine model is a stylized cartoon of the true dynamics at play in the ocean.
+However, for all their limitations, models such as `marlin` can help users understand factors that drive the performance of marine management strategies.
+The goal of this tool is to empower people to design policies based on evaluation of key uncertainties and trade-offs, and in doing so support more effective and equitable marine resource management.
 
@@
 
-All code and materials needed to reproduce this paper is available at
-(https://github.com/DanOvando/marlin-paper)\[https://github.com/DanOvando/marlin-paper\].
+All code and materials needed to reproduce this paper is available at (https://github.com/DanOvando/marlin-paper)\[https://github.com/DanOvando/marlin-paper\].
 

### 2023-03-30 03_marlin_paper.qmd
merging back to main prior to submission

@@
 ---
-title: "Revealing Benefits, Costs, and Tradeoffs of Spatial Management in Marine Systems Arrising from Coupled Economic and Ecological Dynamics"
-author:
-    - Daniel Ovando
-    - Darcy Bradley
-    - Echelle Burns
-    - Lennon Thomas
-    - James Thorson
-date: "Updated `r Sys.Date()`"
 format:
   pdf:
-    toc: true
+    toc: false
     fig-pos: 'H'
+    mainfont: Times New Roman
+    sansfont: Times New Roman
+    fontsize: 12pt
   html: default
@@
   cache: false
+csl: chicago-author-date-16th-edition.csl
 bibliography: references.bib
@@
   run_name: ["v1.0"]
-abstract: "Abstract: Designing effective spatial management strategies is challenging because marine ecosystems are highly dynamic and opaque, and extractive entities such as fishing fleets respond endogenously to ecosystem changes in ways that depend upon ecological and policy context.
-We present a tool, `marlin`, that can be used to efficiently simulate the bio-economic dynamics of marine systems in support of both management and research.
-We demonstrate `marlin’s` capabilities by focusing on two case studies on the conservation and food production impacts of marine protected areas (MPAs): a coastal coral reef and a pelagic tuna fishery.
-In the coastal coral reef example, we explore how heterogeneity in species distributions and fleet preferences can affect distributional outcomes of MPAs.
-In the pelagic case study, we show how our model can be used to assess the climate resilience of different MPA design strategies, as well as the climate sensitivity of different fishing fleets.
-This paper demonstrates how intermediate complexity simulation of coupled bio-economic dynamics can help communities predict and potentially manage trade-offs between conservation, fisheries yields, and distributional outcomes of management policies affected by spatial bio-economic dynamics."
 header-includes:
@@
 
+## Title Page
+
+Journal Name: Ecological Applications
+
+Manuscript Type: Article
+
+Manuscript Title: Simulating Benefits, Costs, and Tradeoffs of Management Policies in Marine Social-Ecological Systems
+
+Author Names: Daniel Ovando^1,2^, Darcy Bradley^3,4,5^, Echelle Burns^3,4,5^,Lennon Thomas^3,4,5^, James Thorson^6^
+
+Affiliations:
+^1^ School of Aquatic and Fishery Sciences, University of Washington,1122 NE Boat St Box 355020,Seattle, WA, USA
+^2^ Current Address: Inter-American Tropical Tuna Commission,8901 La Jolla Shores Drive, CA, USA
+La Jolla, CA 92037, USA
+^3^ Marine Science Institute, University of California, Santa Barbara, CA, USA
+^4^ Bren School of Environmental Science & Management, University of California, Santa Barbara, CA, USA
+^5^ Environmental Markets Lab, University of California, Santa Barbara, CA, USA
+^6^ Habitat and Ecological Processes Research Program, Alaska Fisheries Science Center, NOAA, NMFS, NOAA, Seattle, WA, USA
+
+Corresponding Author: Daniel Ovando <dovando@iattc.org>
+
+Open Research Statement: All code and materials needed to reproduce this paper is available at [https://github.com/DanOvando/marlin-paper](https://github.com/DanOvando/marlin-paper).
+
+Key Words: 
+Bio-economic modeling, spatio-temporal modeling, marine protected areas, fisheries management, marine spatial planning, conservation planning
+
+\newpage
+
+## Abstract
+
+Designing effective spatial management strategies is challenging because marine ecosystems are highly dynamic and opaque, and extractive entities such as fishing fleets respond endogenously to ecosystem changes in ways that depend upon ecological and policy context.
+We present a tool, `marlin`, that can be used to efficiently simulate the bio-economic dynamics of marine systems in support of both management and research.
+We demonstrate `marlin’s` capabilities by focusing on two case studies on the conservation and food production impacts of marine protected areas (MPAs): a coastal coral reef and a pelagic tuna fishery.
+In the coastal coral reef example, we explore how heterogeneity in species distributions and fleet preferences can affect distributional outcomes of MPAs.
+In the pelagic case study, we show how our model can be used to assess the climate resilience of different MPA design strategies, as well as the climate sensitivity of different fishing fleets.
+This paper demonstrates how intermediate complexity simulation of coupled bio-economic dynamics can help communities predict and potentially manage trade-offs between conservation, fisheries yields, and distributional outcomes of management policies affected by spatial bio-economic dynamics.
+
 ## Introduction
 
-Communities around the world are increasingly looking to spatial strategies for managing marine ecosystems .
-For example, the "30x30" movement calls for implementation of marine protected areas (MPAs), a form of spatial management, across 30% of the world's oceans by the year 2030 [@grorud-colvert2021 and references therein].
-However , designing spatial management strategies to achieve desired objectives -- which may include recovery and resilience of overfished species, increased food production and economic well being, and equitable distribution of benefits - is not straightforward.
-Different species have different resilience to fishing, distributions in space and time, and value to fishing fleets; fishing fleets themselves can have varying ranges of incentives and reactions to policy structures.
-These dynamics pose challenges even for single-species assessment and management strategies , which are only amplified when we consider management policies specifically designed around multiple species, fleets, and spatial features in the oceans [@field2006a].
-
-To illustrate, policies such as MPAs designed for both conservation and food production must consider factors such as the optimal size and placement of a protected area network given a wide array of life history, species distributions, exploitation levels, fleet dynamics, and policy-dependent behavior [@reimer2017], all of which may change over time, particularly given the impacts of climate change.
-Efforts to effectively design spatial management strategies such as MPAs are further constrained by a lack of empirical evidence describing the size [@ovando2021a] and time required [@nickols2019] for MPAs to produce substantial effects across a range of social-ecological settings [@mcclanahan2021].
-
-Many models have been developed to support the theory and design of spatial management strategies (see @fulton2015 for a review of existing methods).
-However, these models have tended to be either highly complex and tactical models designed for use in a specific location, or to be extremely stylized representations intended to provide theoretical insights with less relevance for specific applications.
-This lack of accessible models suited for more tactical application presents a challenge to stakeholders charged with marine spatial planning exercises.
-It also presents a barrier to the scientific community, wherein comparisons across spatial management simulation studies are clouded by discrepancies in underlying model structures beyond differences in the core phenomenon in question.
-
-To help address this challenge, we present a novel bio-economic modeling tool called [`marlin`](https://danovando.github.io/marlin/) that allows for efficient simulation of spatio-temporal biological and economic dynamics.
-`marlin` allows users to simulate the impacts of marine management policies across a range of species targeted by various fleets across heterogeneous and dynamic seascapes.
-This model can help users seek Pareto-optimal solutions that produce win-wins or minimize trade-offs across multiple management objectives, such as food production and conservation [@lester2013].
-
-Here we present the core methods and functionality of `marlin`, and demonstrate its use in two case-study applications; a coral reef system and a pelagic system.
-Our results show how interacting ecological, economic, and design attributes can affect the degree of benefits, costs, and trade offs between conservation and food production outcomes of MPAs.
-More broadly, this paper both demonstrates the functionality of our model and the critical need for considering sufficiently realistic coupled ecological and economic dynamics in policy evaluations.
+Communities around the world are increasingly looking to spatial
+strategies for managing marine ecosystems . For example, the "30x30"
+movement calls for implementation of marine protected areas (MPAs), a
+form of spatial management, across 30% of the world's oceans by the year
+2030 [@grorud-colvert2021 and references therein]. However , designing
+spatial management strategies to achieve desired objectives -- which may
+include recovery and resilience of overfished species, increased food
+production and economic well being, and equitable distribution of
+benefits - is not straightforward. Different species have different
+resilience to fishing, distributions in space and time, and value to
+fishing fleets; fishing fleets themselves can have varying ranges of
+incentives and reactions to policy structures. These dynamics pose
+challenges even for single-species assessment and management strategies
+, which are only amplified when we consider management policies
+specifically designed around multiple species, fleets, and spatial
+features in the oceans [@field2006a].
+
+To illustrate, policies such as MPAs designed for both conservation and
+food production must consider factors such as the optimal size and
+placement of a protected area network given a wide array of life
+history, species distributions, exploitation levels, fleet dynamics, and
+policy-dependent behavior [@reimer2017], all of which may change over
+time, particularly given the impacts of climate change. Efforts to
+effectively design spatial management strategies such as MPAs are
+further constrained by a lack of empirical evidence describing the size
+[@ovando2021a] and time required [@nickols2019] for MPAs to produce
+substantial effects across a range of social-ecological settings
+[@mcclanahan2021].
+
+Many models have been developed to support the theory and design of
+spatial management strategies (see @fulton2015 for a review of existing
+methods). However, these models have tended to be either highly complex
+and tactical models designed for use in a specific location, or to be
+extremely stylized representations intended to provide theoretical
+insights with less relevance for specific applications. This lack of
+accessible models suited for more tactical application presents a
+challenge to stakeholders charged with marine spatial planning
+exercises. It also presents a barrier to the scientific community,
+wherein comparisons across spatial management simulation studies are
+clouded by discrepancies in underlying model structures beyond
+differences in the core phenomenon in question.
+
+To help address this challenge, we present a novel bio-economic modeling
+tool called [`marlin`](https://danovando.github.io/marlin/) that allows
+for efficient simulation of spatio-temporal biological and economic
+dynamics. `marlin` allows users to simulate the impacts of marine
+management policies across a range of species targeted by various fleets
+across heterogeneous and dynamic seascapes. This model can help users
+seek Pareto-optimal solutions that produce win-wins or minimize
+trade-offs across multiple management objectives, such as food
+production and conservation [@lester2013].
+
+Here we present the core methods and functionality of `marlin`, and
+demonstrate its use in two case-study applications; a coral reef system
+and a pelagic system. Our results show how interacting ecological,
+economic, and design attributes can affect the degree of benefits,
+costs, and trade offs between conservation and food production outcomes
+of MPAs. More broadly, this paper both demonstrates the functionality of
+our model and the critical need for considering sufficiently realistic
+coupled ecological and economic dynamics in policy evaluations.
 
@@
 
-`marlin` follows a "models of intermediate complexity" philosophy [@plagányi2014] targeted specifically towards capturing heterogeneous habitats, movement dynamics, and the simultaneous impacts of fishing fleets across multiple species (frequently called "technical interactions").
-In its general form, `marlin` simulates the behavior of populations of spatially explicit biologically independent age-structured animal populations affected by one or more fishing fleets in time steps and a spatial resolution specified by the user.
-
-The front end of the software package accompanying this paper, [`marlin`](https://danovando.github.io/marlin/), is written in R [@rcoreteam2021] to facilitate use, while the underlying population model is written in C++, integrated through the `Rcpp` package [@eddelbuettel2018], for speed.
-Users on standard computers should be able to simulate one age-structured population distributed across a 10 by 10 patch system for 20 years in fractions of a second.
-
-While `marlin` does not simulate species interactions, fishing fleets in `marlin` are capable of targeting and affecting multiple species simultaneously.
-This allows the model to simulate processes such as fisheries bycatch or effort displacement in a way that accounts for how fleet behavior might affect multiple species in the system.
-The ability to track the impacts of fishing fleets across multiple species simultaneously is particularly important as very few fishing fleets are truly single species.
-
-We define a few commonly used terms here.
-*Yields* refers to the volume of catch from fishing activity.
-*Spawning stock biomass* refers to the total biomass of reproductively mature fish in the water, a function of numbers, weight, fecundity, and sexual maturity at age.
-We measure the size of the population by th ratio of the *spawning stock biomass* (SSB) in a given time step relative to *unfished spawning stock biomass* (SSB0), SSB/SSB0.
-So, a SSB/SSB0 value of 1 means that the population is unfished, a SSB/SSB0 value of 0 means the population is extinct.
-
-As a demonstration of the use of this modeling tool, we explore two general situations
-
-1.  Trade-offs and distributional outcomes for food production and multi-species biomass in a coastal coral-reef fishery
-2.  Implications of climate-driven range shifts for MPA design in a pelagic fishery
-
-We chose these two examples to illustrate the use of `marlin` in contrasting systems in which spatial management strategies such as MPAs are increasingly considered.
-
-Below we provide a summary of the `marlin` model, as well as details of the case studies.
-We note that `marlin` is a complex software package with many options to control detailed behaviors.
-Given that these specific features will evolve over time, we do not describe in detail every option of the model here.
-Rather, we focus on the equations and methods describing the core foundations of our tool for efficiently simulating bio-economic dynamics in aquatic systems.
+`marlin` follows a "models of intermediate complexity" philosophy
+[@plagányi2014] targeted specifically towards capturing heterogeneous
+habitats, movement dynamics, and the simultaneous impacts of fishing
+fleets across multiple species (frequently called "technical
+interactions"). In its general form, `marlin` simulates the behavior of
+populations of spatially explicit biologically independent
+age-structured animal populations affected by one or more fishing fleets
+in time steps and a spatial resolution specified by the user.
+
+The front end of the software package accompanying this paper,
+[`marlin`](https://danovando.github.io/marlin/), is written in R
+[@rcoreteam2021] to facilitate use, while the underlying population
+model is written in C++, integrated through the `Rcpp` package
+[@eddelbuettel2018], for speed. Users on standard computers should be
+able to simulate one age-structured population distributed across a 10
+by 10 patch system for 20 years in fractions of a second.
+
+While `marlin` does not simulate species interactions, fishing fleets in
+`marlin` are capable of targeting and affecting multiple species
+simultaneously. This allows the model to simulate processes such as
+fisheries bycatch or effort displacement in a way that accounts for how
+fleet behavior might affect multiple species in the system. The ability
+to track the impacts of fishing fleets across multiple species
+simultaneously is particularly important as very few fishing fleets are
+truly single species.
+
+We define a few commonly used terms here. *Yields* refers to the volume
+of catch from fishing activity. *Spawning stock biomass* refers to the
+total biomass of reproductively mature fish in the water, a function of
+numbers, weight, fecundity, and sexual maturity at age. We measure the
+size of the population by th ratio of the *spawning stock biomass* (SSB)
+in a given time step relative to *unfished spawning stock biomass*
+(SSB0), SSB/SSB0. So, a SSB/SSB0 value of 1 means that the population is
+unfished, a SSB/SSB0 value of 0 means the population is extinct.
+
+As a demonstration of the use of this modeling tool, we explore two
+general situations
+
+1.  Trade-offs and distributional outcomes for food production and
+    multi-species biomass in a coastal coral-reef fishery
+2.  Implications of climate-driven range shifts for MPA design in a
+    pelagic fishery
+
+We chose these two examples to illustrate the use of `marlin` in
+contrasting systems in which spatial management strategies such as MPAs
+are increasingly considered.
+
+Below we provide a summary of the `marlin` model, as well as details of
+the case studies. We note that `marlin` is a complex software package
+with many options to control detailed behaviors. Given that these
+specific features will evolve over time, we do not describe in detail
+every option of the model here. Rather, we focus on the equations and
+methods describing the core foundations of our tool for efficiently
+simulating bio-economic dynamics in aquatic systems.
 
@@
 
-`marlin` simulates the dynamics of one or more species, currently best representing fish-like species, using age-structured population dynamics.
-Each time step, species move throughout the simulated seascape using both diffusion and "taxis" (active movement towards preferred habitat), experience natural and potentially fishing mortality, and if in a spawning season spawn using one of the possible forms of density dependence implemented in the model, with the potential to include auto-correlated stochastic deviations in the amount of offspring produced, generally called "recruitment deviates".
-
-These species can be caught by fishing fleets.
-A fishing fleet is defined in `marlin` by a set of fishers that have the same fishing skill, prices, and selectivity for individual species (each of which we denote as a *metier*).
-For example, both a longline and purse seine fleet may capture bigeye (*Thunnus obesus*) and skipjack tunas, but the longline fleet can be made much more likely to catch larger bigeye than skipjack, and *vice versa*.
-Each fishing fleet distributes its fishing effort in space according to a specified spatial allocation rule (see *Spatial Allocation of Effort* ), for example in proportion to profit-per-unit-effort, conditional on management policies in places such as quotas and/or the presence of any spatial restrictions such as MPAs.
-
-Unconstrained by management, the total amount of effort exerted by each fleet can follow one of two dynamics: open access or constant effort (see *Calculating Total Effort*).
-Under constant effort, the total amount of effort of each fleet is fixed over time, with the possible exception of attrition due to MPA placement.
-Under open access, the total amount of effort in a given time step is a function of the profitability of the fleet in the previous time step, evolving until a bionomic equilibrium of zero total profits is reached.
-Profitability is a function of the volume and price of each species caught, as well as the cost of the total amount of fishing effort per fleet and the travel costs as a function of distance from a port (see *Fleet Dynamics*).
-
-The dynamics of each fleet can be modified by management in a variety of ways.
-Managers can impose size limits for individual species within each fleet.
-They can also impose total catch quotas for each species in each time step.
-When quotas are activated, if the total catch across all fleets for a given species under the model's effort dynamics would exceed the allowable quota for that species, the total effort for each fleet is decreased proportional to its contribution to the total catch until the quota is satisfied.
-As an alternative to catch quotas, managers can set an effort cap per fleet, which prevents effort from exceeding a given amount under open access dynamics, though effort may be reduced below this cap if required by the profit equation (i.e.
-the fleet can choose to fish less than the quota or effort cap).
-
-`marlin` also allows users to specify closed fishing seasons for one or more species in the system.
-When all species in a simulation share the same closed seasons, no fishing occurs in those seasons.
-When not all species share the same closed seasons, fishing still occurs in the closed seasons, but the fleet avoids catching the species for which fishing is closed in a given season, assuming no discard mortality.
-When closed seasons are in effect, the amount of effort in the next open season is a function of the effort and profits in the last open fishing season.
-Lastly, managers can specify locations of no-take MPAs, which can change in size and location if needed.
-When MPAs are implemented, fishing effort that used to operate inside the MPAs can either leave the fishery, or be redistributed outside the MPA (the default behavior).
-
-Users can use the model to explore general properties of spatially explicit social-ecological systems, or can parameterize it to explore key uncertainties of more tactical applications.
-As an example of the later case, we provide two case studies here designed to reflect the kinds of questions stakeholders might face if designing an MPA network for a coastal coral reef or a pelagic system.
+`marlin` simulates the dynamics of one or more species, currently best
+representing fish-like species, using age-structured population
+dynamics. Each time step, species move throughout the simulated seascape
+using both diffusion and "taxis" (active movement towards preferred
+habitat), experience natural and potentially fishing mortality, and if
+in a spawning season spawn using one of the possible forms of density
+dependence implemented in the model, with the potential to include
+auto-correlated stochastic deviations in the amount of offspring
+produced, generally called "recruitment deviates".
+
+These species can be caught by fishing fleets. A fishing fleet is
+defined in `marlin` by a set of fishers that have the same fishing
+skill, prices, and selectivity for individual species (each of which we
+denote as a *metier*). For example, both a longline and purse seine
+fleet may capture bigeye (*Thunnus obesus*) and skipjack tunas, but the
+longline fleet can be made much more likely to catch larger bigeye than
+skipjack, and *vice versa*. Each fishing fleet distributes its fishing
+effort in space according to a specified spatial allocation rule (see
+*Spatial Allocation of Effort* ), for example in proportion to
+profit-per-unit-effort, conditional on management policies in places
+such as quotas and/or the presence of any spatial restrictions such as
+MPAs.
+
+Unconstrained by management, the total amount of effort exerted by each
+fleet can follow one of two dynamics: open access or constant effort
+(see *Calculating Total Effort*). Under constant effort, the total
+amount of effort of each fleet is fixed over time, with the possible
+exception of attrition due to MPA placement. Under open access, the
+total amount of effort in a given time step is a function of the
+profitability of the fleet in the previous time step, evolving until a
+bionomic equilibrium of zero total profits is reached. Profitability is
+a function of the volume and price of each species caught, as well as
+the cost of the total amount of fishing effort per fleet and the travel
+costs as a function of distance from a port (see *Fleet Dynamics*).
+
+The dynamics of each fleet can be modified by management in a variety of
+ways. Managers can impose size limits for individual species within each
+fleet. They can also impose total catch quotas for each species in each
+time step. When quotas are activated, if the total catch across all
+fleets for a given species under the model's effort dynamics would
+exceed the allowable quota for that species, the total effort for each
+fleet is decreased proportional to its contribution to the total catch
+until the quota is satisfied. As an alternative to catch quotas,
+managers can set an effort cap per fleet, which prevents effort from
+exceeding a given amount under open access dynamics, though effort may
+be reduced below this cap if required by the profit equation (i.e. the
+fleet can choose to fish less than the quota or effort cap).
+
+`marlin` also allows users to specify closed fishing seasons for one or
+more species in the system. When all species in a simulation share the
+same closed seasons, no fishing occurs in those seasons. When not all
+species share the same closed seasons, fishing still occurs in the
+closed seasons, but the fleet avoids catching the species for which
+fishing is closed in a given season, assuming no discard mortality. When
+closed seasons are in effect, the amount of effort in the next open
+season is a function of the effort and profits in the last open fishing
+season. Lastly, managers can specify locations of no-take MPAs, which
+can change in size and location if needed. When MPAs are implemented,
+fishing effort that used to operate inside the MPAs can either leave the
+fishery, or be redistributed outside the MPA (the default behavior).
+
+Users can use the model to explore general properties of spatially
+explicit social-ecological systems, or can parameterize it to explore
+key uncertainties of more tactical applications. As an example of the
+later case, we provide two case studies here designed to reflect the
+kinds of questions stakeholders might face if designing an MPA network
+for a coastal coral reef or a pelagic system.
 
@@
 
-We approximate the dome-shaped form as a normal distribution with mean $l^{sel}$ and standard deviation $\sigma$.
-The normal density function is re-scaled such that the selectivity is 1 at $l^{sel}$.
+We approximate the dome-shaped form as a normal distribution with mean
+$l^{sel}$ and standard deviation $\sigma$. The normal density function
+is re-scaled such that the selectivity is 1 at $l^{sel}$.
 
@@
 
-`marlin` allows for two general modes of effort.
-The simplest is "constant effort", in which the total effort of each fishing fleet remains constant over time.
-
-The more complex option is "open access".
-Under open access, the total amount of effort in a given time step for fleet *f* is a function of the profitability of that fleet in the proceeding time steps in which fishing was open, where $\theta$ controls the responsiveness of fleet *f* to the log of the ratio of total revenues *R* to total costs *C*, and approximates the proportional change in effort in response to a one unit changes in the log revenue to cost ratio.
+`marlin` allows for two general modes of effort. The simplest is
+"constant effort", in which the total effort of each fishing fleet
+remains constant over time.
+
+The more complex option is "open access". Under open access, the total
+amount of effort in a given time step for fleet *f* is a function of the
+profitability of that fleet in the proceeding time steps in which
+fishing was open, where $\theta$ controls the responsiveness of fleet
+*f* to the log of the ratio of total revenues *R* to total costs *C*,
+and approximates the proportional change in effort in response to a one
+unit changes in the log revenue to cost ratio.
 
@@
 
-Each fishing fleet decides how to allocate its effort in space based on one of four possible spatial allocation strategies.
-The ideal free distribution (IFD) is the standard method for distributing fishing fleets in spaces (see @gillis2003 and references therein).
-However, analytical solutions to the IFD present a number of complications for our model.
-Our model is capable of tracking multiple fleets at once.
-This means that the IFD for a single fleet would commonly be modeled as a Nash-equilibrium based on the actions of each of the individual fleet conditional on the actions of all other fleets.
-While possible to implement, this would slow down our model runs to the point of practically preventing large-scale evaluation of the spatial management policies.
-
-As such we explored a series of "next best" fleet distribution algorithms.
-While not the IFD in any individual time step, over time they start to approximate the IFD, as each assumes that fleets base their decisions for the current time step on the outcomes in the prior timestep, meaning that the impacts of the actions of other fishing fleets are somewhat internalized.
-The four possible fleet distribution algorithms are
-
-1.  Revenue per unit effort (RPUE): The fleet distributes itself in space based on the realized revenue per unit effort in each patch in the preceding time step
-2.  Revenue: The fleet distributes itself in space based on the realized total revenue in each patch in the preceding time step
-3.  Profit per unit effort (PPUE): The fleet distributes itself in space based on the realized profit per unit effort in each patch in the preceding time step
-4.  Profit: The fleet distributes itself in space based on profit in each patch in the preceding time step
-
-Revenue based spatial distribution is not likely to be very realistic; in general we would expect fishing fleets to respond to profits on some core level.
-However, due to the complexity of parameterizing cost functions, fleet dynamics are often evaluated based on yield or revenue alone, and so we include those scenarios here to allow users to evaluate the potential implications of this choice.
-The decision on whether to allocate the fleet based on absolute or relative (per unit effort) metrics is more complex.
-When effort represents the actions of separate and individual fishing actors (e.g.
-independent fishing vessels), a per-unit-effort strategy, in which fishers distribute themselves based on the expected catch of their individual efforts, may be more realistic [@hilborn1987].
-Conversely, a system defined by a sole owner seeking to maximize total profits might be better represented by a fleet model based on total profits
+Each fishing fleet decides how to allocate its effort in space based on
+one of four possible spatial allocation strategies. The ideal free
+distribution (IFD) is the standard method for distributing fishing
+fleets in spaces (see @gillis2003 and references therein). However,
+analytical solutions to the IFD present a number of complications for
+our model. Our model is capable of tracking multiple fleets at once.
+This means that the IFD for a single fleet would commonly be modeled as
+a Nash-equilibrium based on the actions of each of the individual fleet
+conditional on the actions of all other fleets. While possible to
+implement, this would slow down our model runs to the point of
+practically preventing large-scale evaluation of the spatial management
+policies.
+
+As such we explored a series of "next best" fleet distribution
+algorithms. While not the IFD in any individual time step, over time
+they start to approximate the IFD, as each assumes that fleets base
+their decisions for the current time step on the outcomes in the prior
+timestep, meaning that the impacts of the actions of other fishing
+fleets are somewhat internalized. The four possible fleet distribution
+algorithms are
+
+1.  Revenue per unit effort (RPUE): The fleet distributes itself in
+    space based on the realized revenue per unit effort in each patch in
+    the preceding time step
+2.  Revenue: The fleet distributes itself in space based on the realized
+    total revenue in each patch in the preceding time step
+3.  Profit per unit effort (PPUE): The fleet distributes itself in space
+    based on the realized profit per unit effort in each patch in the
+    preceding time step
+4.  Profit: The fleet distributes itself in space based on profit in
+    each patch in the preceding time step
+
+Revenue based spatial distribution is not likely to be very realistic;
+in general we would expect fishing fleets to respond to profits on some
+core level. However, due to the complexity of parameterizing cost
+functions, fleet dynamics are often evaluated based on yield or revenue
+alone, and so we include those scenarios here to allow users to evaluate
+the potential implications of this choice. The decision on whether to
+allocate the fleet based on absolute or relative (per unit effort)
+metrics is more complex. When effort represents the actions of separate
+and individual fishing actors (e.g. independent fishing vessels), a
+per-unit-effort strategy, in which fishers distribute themselves based
+on the expected catch of their individual efforts, may be more realistic
+[@hilborn1987]. Conversely, a system defined by a sole owner seeking to
+maximize total profits might be better represented by a fleet model
+based on total profits
 
@@
 
-`marlin`'s movement dynamics are based on a continuous-time Markov chain (CTMC), as described in @thorson2021a.
-Within this framework, the model allows for movement to be broken down into three components of *advection* (drifting with currents), *taxis* (active movement towards preferred habitat), and *diffusion* (essentially remaining variation in movement not explained by advection or taxis).
-For now, `marlin` focuses just on the diffusion and taxis components of this model, assuming that advection is zero, though future extensions could easily incorporate advection vectors from oceanographic models.
-In this way, `marlin` allows users to run anything from a simple Gaussian dispersal kernel up to a system governed by species that passively diffuse out from a core habitat defined by a dynamic thermal range.
-
-We provide a brief overview of the the general CTMC method here, see @thorson2021a for a detailed description.
-Under this framework, movement from each patch to each other patch in the system in a given timestep *t* for life stage *a* of species *s* is defined by a movement matrix $\boldsymbol{M}_{t,s,a}$.
-$\boldsymbol{M}_{t,s,a}$ is calculated as the matrix exponent of a function of the diffusion $\boldsymbol{D}$ and taxis $\boldsymbol{\tau}$ matrices scaled by the width of the time step $\Delta_{t}$ and the length of the edge of each patch $\Delta_d$ that the model is running under.
-So for example, if we are running a model on a monthly timestep, and we define the units of time in years, then $\Delta_{t} = 1/12$.
-Similarly, if we are running a model where one patch equals 100km^2^, then $\Delta_d = 10km$.
-This parameterization allows users to set the effective area of the spatial domain through two avenues; the number of patches, which effectively scales the resolution of the model, and the area of each patch, which scales the spatial extent of the simulation.
+`marlin`'s movement dynamics are based on a continuous-time Markov chain
+(CTMC), as described in @thorson2021a. Within this framework, the model
+allows for movement to be broken down into three components of
+*advection* (drifting with currents), *taxis* (active movement towards
+preferred habitat), and *diffusion* (essentially remaining variation in
+movement not explained by advection or taxis). For now, `marlin` focuses
+just on the diffusion and taxis components of this model, assuming that
+advection is zero, though future extensions could easily incorporate
+advection vectors from oceanographic models. In this way, `marlin`
+allows users to run anything from a simple Gaussian dispersal kernel up
+to a system governed by species that passively diffuse out from a core
+habitat defined by a dynamic thermal range.
+
+We provide a brief overview of the the general CTMC method here, see
+@thorson2021a for a detailed description. Under this framework, movement
+from each patch to each other patch in the system in a given timestep
+*t* for life stage *a* of species *s* is defined by a movement matrix
+$\boldsymbol{M}_{t,s,a}$. $\boldsymbol{M}_{t,s,a}$ is calculated as the
+matrix exponent of a function of the diffusion $\boldsymbol{D}$ and
+taxis $\boldsymbol{\tau}$ matrices scaled by the width of the time step
+$\Delta_{t}$ and the length of the edge of each patch $\Delta_d$ that
+the model is running under. So for example, if we are running a model on
+a monthly timestep, and we define the units of time in years, then
+$\Delta_{t} = 1/12$. Similarly, if we are running a model where one
+patch equals 100km^2^, then $\Delta_d = 10km$. This parameterization
+allows users to set the effective area of the spatial domain through two
+avenues; the number of patches, which effectively scales the resolution
+of the model, and the area of each patch, which scales the spatial
+extent of the simulation.
 
@@
 
-The individual components (*m*) of the movement matrix are filled based on an adjacency matrix, which defines whether two patches are both adjacent and water (as opposed to land or another physical barrier), a diffusion rate $D$ defined in units of area of a patch per unit of time, and a habitat preference function *h* in units of length of a side of a patch per unit time.
-For example, if we are defining the time units as years and the distance units as kilometers, for a tuna $D$ might be 1,000 $\frac{KM^2}{Year}$.
-We then use parameters $\Delta_{t}$ and $\Delta_{d}$ parameters to translate the diffusion rate $D$ to match the time step and patch size used in a simulation.
-This "scale free" parameterization means that appropriate value of $D$ can be identified for a species and then set, regardless of the time step or patch size used in the simulation model itself.
-The taxis component of the movement process is a function of the difference in habitat quality *h*.
-The habitat preference function itself can take any form the user wishes.
-Exponentiating the difference in the habitat preference function between patches turns the taxis matrix into a multiplier of the diffusion rate *D*.
-As such, when creating habitat layers for simulation, users can tune the scale of the habitat gradient function to result in realistic multipliers of the diffusion rate.
-This parameterization ensure that the off-diagonal elements of the movement matrix $\boldsymbol{M}_{t,s,a}$ are all non-negative, a requirement of the CTMC method.
+The individual components (*m*) of the movement matrix are filled based
+on an adjacency matrix, which defines whether two patches are both
+adjacent and water (as opposed to land or another physical barrier), a
+diffusion rate $D$ defined in units of area of a patch per unit of time,
+and a habitat preference function *h* in units of length of a side of a
+patch per unit time. For example, if we are defining the time units as
+years and the distance units as kilometers, for a tuna $D$ might be
+1,000 $\frac{KM^2}{Year}$. We then use parameters $\Delta_{t}$ and
+$\Delta_{d}$ parameters to translate the diffusion rate $D$ to match the
+time step and patch size used in a simulation. This "scale free"
+parameterization means that appropriate value of $D$ can be identified
+for a species and then set, regardless of the time step or patch size
+used in the simulation model itself. The taxis component of the movement
+process is a function of the difference in habitat quality *h*. The
+habitat preference function itself can take any form the user wishes.
+Exponentiating the difference in the habitat preference function between
+patches turns the taxis matrix into a multiplier of the diffusion rate
+*D*. As such, when creating habitat layers for simulation, users can
+tune the scale of the habitat gradient function to result in realistic
+multipliers of the diffusion rate. This parameterization ensure that the
+off-diagonal elements of the movement matrix $\boldsymbol{M}_{t,s,a}$
+are all non-negative, a requirement of the CTMC method.
 
@@
 
-For both the diffusion and taxis matrices, we allow for the inclusion of physical barriers to movement (i.e. land).
-Pairs of patches that are adjacent but in which one or both patches are a barrier to movement are set as non-adjacent.
-The CTMC model then produces movement dynamics that move around barriers rather than over them.
-This is different than setting habitat in a patch to zero; an animal will not preferentially move towards a patch with zero habitat all else being equal, but can move over a patch with zero habitat towards another better patch.
-An example might be a species that lives and feeds on coastal habitats transiting through a patch of open ocean to reach a new feeding ground.
-In contrast, a physical barrier represents something like a peninsula that an animal must move around to transit from one side to another.
-
-While this CTMC approach to movement simulation is to date not commonly seen in the marine modeling literature, it has numerous advantages that warrant its seeming complexity.
-First, when only diffusion is present, the model will generalize to the familiar dynamics of a Gaussian dispersal kernel.
-However, the time component *t* in @eq-movement allows the movement model to be run on arbitrarily fine time steps.
-Second, the taxis model allow for more active habitat choices by species, allowing us to simulate preferences of species in space and time efficiently.
-Lastly, this form has the advantage that the parameters are directly estimable from real data.
-So, if provided with for example spatial abundance data and a tagging study, users can estimate the diffusion and taxis movement parameters in the same manner as @thorson2021a, and then pass their estimated parameters to `marlin` for simulation (so long as the estimating method uses the same functional form as the movement model in `marlin`).
+For both the diffusion and taxis matrices, we allow for the inclusion of
+physical barriers to movement (i.e. land). Pairs of patches that are
+adjacent but in which one or both patches are a barrier to movement are
+set as non-adjacent. The CTMC model then produces movement dynamics that
+move around barriers rather than over them. This is different than
+setting habitat in a patch to zero; an animal will not preferentially
+move towards a patch with zero habitat all else being equal, but can
+move over a patch with zero habitat towards another better patch. An
+example might be a species that lives and feeds on coastal habitats
+transiting through a patch of open ocean to reach a new feeding ground.
+In contrast, a physical barrier represents something like a peninsula
+that an animal must move around to transit from one side to another.
+
+While this CTMC approach to movement simulation is to date not commonly
+seen in the marine modeling literature, it has numerous advantages that
+warrant its seeming complexity. First, when only diffusion is present,
+the model will generalize to the familiar dynamics of a Gaussian
+dispersal kernel. However, the time component *t* in @eq-movement allows
+the movement model to be run on arbitrarily fine time steps. Second, the
+taxis model allow for more active habitat choices by species, allowing
+us to simulate preferences of species in space and time efficiently.
+Lastly, this form has the advantage that the parameters are directly
+estimable from real data. So, if provided with for example spatial
+abundance data and a tagging study, users can estimate the diffusion and
+taxis movement parameters in the same manner as @thorson2021a, and then
+pass their estimated parameters to `marlin` for simulation (so long as
+the estimating method uses the same functional form as the movement
+model in `marlin`).
 
@@
 
-where *BH* is the Beverton-Holt recruitment function [@beverton1957] and *ssb* is spawning-stock-biomass.
-Per convention, the model allows for a "plus group", wherein rather than tracking numbers of every possible age, individuals greater than or equal to a given maximum age are grouped together.
-
-Spawning stock biomass *ssb* is calculated by converting age to mean length, calculating weight at age, maturity at age, and then calculating spawning stock biomass as the sum of spawning potential at age in a given time step, taking into account the potential for hyperallometry in the manner of @marshall2021.
-Age is converted to length through the von Bertalanffy growth equation given parameters asymptotic length ($l_\infty$), growth (*k*) and theoretical age at length zero ($a0$) assuming log-normally distributed variation *u* in the length at age with CV $\sigma_s$.
+where *BH* is the Beverton-Holt recruitment function [@beverton1957] and
+*ssb* is spawning-stock-biomass. Per convention, the model allows for a
+"plus group", wherein rather than tracking numbers of every possible
+age, individuals greater than or equal to a given maximum age are
+grouped together.
+
+Spawning stock biomass *ssb* is calculated by converting age to mean
+length, calculating weight at age, maturity at age, and then calculating
+spawning stock biomass as the sum of spawning potential at age in a
+given time step, taking into account the potential for hyperallometry in
+the manner of @marshall2021. Age is converted to length through the von
+Bertalanffy growth equation given parameters asymptotic length
+($l_\infty$), growth (*k*) and theoretical age at length zero ($a0$)
+assuming log-normally distributed variation *u* in the length at age
+with CV $\sigma_s$.
 
@@
 
-Biomass *b* at age is then given by the weight at length equation governed by a scaling coefficient $\Omega_s$ and an exponent $\Phi_s$ that controls the rate at which volume scales with length
+Users can manually supply a vector of of natural mortality at age (*m*).
+Or, they can supply one value of natural mortality which is then
+converted into mortality at age through one of two means. Under the
+default behavior, natural mortality at age given a target mean mortality
+across all ages $m_s$ is calculated using a length-inverse mortality
+function [@lorenzen2022].
+
+$$
+minv_{s_a} = (\frac{l_{s,a}}{l_{\infty,s}})^{-1}
+$$ {#eq-minv}
+
+$$
+m_{s,a} = \frac{minv_{s,a}}{mean( minv_{s,a})}m_s
+$$ {#eq-mata}
+
+Alternatively, users can set mortality at age to be constant
+
+$$
+m_{s,a} = m_s
+$$ {#eq-mata2}
+
+Biomass *b* at age is then given by the weight at length equation
+governed by a scaling coefficient $\Omega_s$ and an exponent $\Phi_s$
+that controls the rate at which volume scales with length
 
@@
 
-Spawning stock biomass at time *t* is then calculated as a function of the numbers at age, the maturity at age, and the weight at age raised by a parameter $\gamma$.
-When $\gamma$ is greater than 1, the species is said to experience hyperallometric fecundity, i.e. fecundity increases faster than weight.
+Spawning stock biomass at time *t* is then calculated as a function of
+the numbers at age, the maturity at age, and the weight at age raised by
+a parameter $\gamma$. When $\gamma$ is greater than 1, the species is
+said to experience hyperallometric fecundity, i.e. fecundity increases
+faster than weight.
 
@@
 
-Recruitment (i.e.
-the number of age 1 individuals entering the population) follows Beverton-Holt dynamics parameterized around steepness (*h*) with log-normally distributed recruitment deviates $\epsilon$.
-When steepness is one recruitment is independent of spawning biomass.
-As steepness approaches 0.2 recruitment becomes a linear function of spawning biomass.
-We allow for five variants in the timing of density dependent recruitment, building off of @babcock2011 :
-
-1.  Global density dependence: Density dependent recruitment is a function of the sum of spawning biomass across all patches, and recruits are then distributed according to habitat quality
+Recruitment (i.e. the number of age 1 individuals entering the
+population) follows Beverton-Holt dynamics parameterized around
+steepness (*h*) with log-normally distributed recruitment deviates
+$\epsilon$. When steepness is one recruitment is independent of spawning
+biomass. As steepness approaches 0.2 recruitment becomes a linear
+function of spawning biomass. We allow for five variants in the timing
+of density dependent recruitment, building off of @babcock2011 :
+
+1.  Global density dependence: Density dependent recruitment is a
+    function of the sum of spawning biomass across all patches, and
+    recruits are then distributed according to habitat quality
 
@@
 
-4.  Post-dispersal density dependence: Larvae are distributed throughout the system, and then density dependent recruitment occurs based on the density of spawning biomass at the destination patch.
+4.  Post-dispersal density dependence: Larvae are distributed throughout
+    the system, and then density dependent recruitment occurs based on
+    the density of spawning biomass at the destination patch.
 
@@
 
-Fisheries management is often concerned with measuring stock status relative to maximum sustainable yield (MSY) based reference points, though the exact level of stock status relative to MSY reference points desired by management or societal objectives may vary widely.
-MSY based reference points present a problem for a multi-fleet and spatial-temporal model such as `marlin`.
-MSY and the fishing mortality rate that would produce MSY, $F_{MSY}$, are a function of fishery selectivity.
-Fishery selectivity in this model can vary by fleet, and species can be distributed unevenly in both space and time.
-This means that the net effective fishing selectivity on a species can vary depending on the dynamics at a given moment, making the definition of an equilibrium concept such as MSY challenging.
-
-As such, we do not report MSY based reference points in the model by default, with the exception of the spawning stock biomass at MSY ($SSB_{MSY}$) which tends to be conserved across selectivity patterns as it is primarily a function of the mortality and flow of recruits relative to the maturity ogive.
-There are many different strategies for estimating reference points in spatially explicit systems [@kapur2021a].
-We leave it to users to define and find relevant reference point as required by their specific needs.
+Fisheries management is often concerned with measuring stock status
+relative to maximum sustainable yield (MSY) based reference points,
+though the exact level of stock status relative to MSY reference points
+desired by management or societal objectives may vary widely. MSY based
+reference points present a problem for a multi-fleet and
+spatial-temporal model such as `marlin`. MSY and the fishing mortality
+rate that would produce MSY, $F_{MSY}$, are a function of fishery
+selectivity. Fishery selectivity in this model can vary by fleet, and
+species can be distributed unevenly in both space and time. This means
+that the net effective fishing selectivity on a species can vary
+depending on the dynamics at a given moment, making the definition of an
+equilibrium concept such as MSY challenging.
+
+As such, we do not report MSY based reference points in the model by
+default, with the exception of the spawning stock biomass at MSY
+($SSB_{MSY}$) which tends to be conserved across selectivity patterns as
+it is primarily a function of the mortality and flow of recruits
+relative to the maturity ogive. There are many different strategies for
+estimating reference points in spatially explicit systems [@kapur2021a].
+We leave it to users to define and find relevant reference point as
+required by their specific needs.
 
@@
 
-We include two examples demonstrating how `marlin` can be used to support marine spatial planning.
-In the first, we show how `marlin` can be used to compare the total and distributional impacts of MPAs designed in a heavily fished coastal coral reef ecosystem.
-In the second, we demonstrate how `marlin` can be used to assess components of climate resilience of alternative MPA design strategies in a pelagic system.
-
-Each of the case studies contains too many parameters and options to be succinctly presented in the text here, and we are concerned that attempting to duplicate all of the parameters of the model within the text would simply introduce errors and confusion.
-As such, and since these results are for illustrative purposes only, readers should consult the accompanying code to view the precise details of each simulation.
-Targeted tactical applications must carefully consider and document all decisions made around model parameters.
+We include two examples demonstrating how `marlin` can be used to
+support marine spatial planning. In the first, we show how `marlin` can
+be used to compare the total and distributional impacts of MPAs designed
+in a heavily fished coastal coral reef ecosystem. In the second, we
+demonstrate how `marlin` can be used to assess components of climate
+resilience of alternative MPA design strategies in a pelagic system.
+
+Each of the case studies contains too many parameters and options to be
+succinctly presented in the text here, and we are concerned that
+attempting to duplicate all of the parameters of the model within the
+text would simply introduce errors and confusion. As such, and since
+these results are for illustrative purposes only, readers should consult
+the accompanying code to view the precise details of each simulation.
+Targeted tactical applications must carefully consider and document all
+decisions made around model parameters.
 
@@
 
-1.  *Rate*: MPAs are placed based on the pre-MPA SSB/SSB0 weighted catch relative to the total catch in a patch.
-    So, patches with high rates of catch of depleted species relative to total catch are prioritized.
-
-2.  *Target Fishing*: MPAs locations are prioritized proportional to fishery catches.
-    Patches with high total catches are prioritized over patches with low catches.
-
-3.  *Spawning Aggregation*: MPAs are centered on the location of a known spawning aggregation
-
-To our knowledge, MPAs are never randomly placed; their locations are decided based on stakeholder negotiation given available information and objectives.
-In theory, the design of MPA networks can be optimized through the use of a modeling framework, and depending on the validity of the model, this process is likely to produce better outcomes than manually-designed strategies [@rassweiler2012; @rassweiler2014].
-However, designing optimal MPA networks becomes increasingly difficult as the range of objectives and the complexity of the model increase.
-Therefore, while `marlin` provides some support for [optimization](https://danovando.github.io/marlin/articles/optimize-mpas.html), we focus here on the design and performance of these more "rule of thumb" design strategies that may be more accessible to a wider range of users.
-We allow all MPAs to be designed in a "mosaic" fashion in these examples, but users can easily extend the analysis to compare outcomes between "contiguous" (MPA is made up of one continuously connected block) and "mosaic" (MPAs can be separated in space) MPA designs [@pons2022] by developing and passing their own MPA design algorithms to the model.
+1.  *Rate*: MPAs are placed based on the pre-MPA SSB/SSB0 weighted catch
+    relative to the total catch in a patch. So, patches with high rates
+    of catch of depleted species relative to total catch are
+    prioritized.
+
+2.  *Target Fishing*: MPAs locations are prioritized proportional to
+    fishery catches. Patches with high total catches are prioritized
+    over patches with low catches.
+
+3.  *Spawning Aggregation*: MPAs are centered on the location of a known
+    spawning aggregation
+
+To our knowledge, MPAs are never randomly placed; their locations are
+decided based on stakeholder negotiation given available information and
+objectives. In theory, the design of MPA networks can be optimized
+through the use of a modeling framework, and depending on the validity
+of the model, this process is likely to produce better outcomes than
+manually-designed strategies [@rassweiler2012; @rassweiler2014].
+However, designing optimal MPA networks becomes increasingly difficult
+as the range of objectives and the complexity of the model increase.
+Therefore, while `marlin` provides some support for
+[optimization](https://danovando.github.io/marlin/articles/optimize-mpas.html),
+we focus here on the design and performance of these more "rule of
+thumb" design strategies that may be more accessible to a wider range of
+users. We allow all MPAs to be designed in a "mosaic" fashion in these
+examples, but users can easily extend the analysis to compare outcomes
+between "contiguous" (MPA is made up of one continuously connected
+block) and "mosaic" (MPAs can be separated in space) MPA designs
+[@pons2022] by developing and passing their own MPA design algorithms to
+the model.
 
@@
 
-In our coastal coral reef example, we model the dynamics of four tropical Pacific species: a grouper (*Epinephelus fuscoguttatus*), a shallow-reef snapper (*Lutjanus malabaricus*), a deep-water snapper (*Pristipomoides filamentosus*), and a reef shark (*Carcharhinus amblyrhynchos*).
-The simulated groupers undergo a mass migration to a spawning aggregation once per year, followed by the sharks.
-Shallow-reef snappers stay in reefs closer to shore above a steep drop-off year-round, while deep-water snappers stay in the deeper reefs past the drop-off (@fig-coral).
-
-These species are targeted by two different fleets.
-Fleet One primarily targets the grouper and near-shore snapper populations, but will land any incidentally captured sharks.
-Fleet One has a logistic selectivity pattern for all species, as they retain any fish caught for consumption or sale.
-both grouper and shallow-reef snapper due to demand for "plate sized" fish , and logistic selectivity for sharks, as any shark caught is retained.
-Fleet One is totally dependent on fishing for their livelihood, meaning the local community takes advantage of every possible opportunity to fish, and as such we model it as a "constant effort" fishery.
-Do to having less efficient boats, Fleet One has a higher cost per distance coefficient than Fleet Two.
-Fleet One's home port is located on the site of the grouper spawning aggregation.
-
-Fleet Two is a more commercial fleet that primarily targets the deep-water snapper population.
-This fleet primarily sells their catch to local restaurants and distributors where plate-sized fish are prized, and so for both snapper and grouper Fleet Two has a dome-shaped selectivity pattern [@kindsvater2017].
-While plate-sized deep snapper are the primary target of Fleet Two, we model Fleet Two's selectivity for deep snapper as logistic due to high levels of discard mortality for deep-water snapper resulting from barotrauma.
-Fleet Two catches groupers, though less than Fleet One, and receives no price for sharks due to the requirements of a certification program through which they sell their deep-water snapper.
-Accidental captures (bycatch) of sharks do occur, which results in mortality.
-Fleet Two operates under open-access dynamics, as fishing is not the only means of subsistence for this community; short-term effort expands and contracts in response to profitability of the primarily grouper-driven fishery.
-Fleet two's home port is located further north from Fleet One.
-
-We use `marlin` to simulate the outcomes for both food production and conservation for each of the species and both of the fleets as a function of both MPA size and MPA design strategy.
-For this exercise, MPAs are placed with perfect information and have no design constraints for continuity (i.e. the selected MPA design can be as patchwork as dictated by the data).
-We ran the simulation on a quarterly level ($\Delta_t = 1/4$) and set the area of each patch to be 5km^2^ ($\Delta_p = \sqrt{5}$), using 400 patches for a total area of 2,000 KM^2^.
+In our coastal coral reef example, we model the dynamics of four
+tropical Pacific species: a grouper (*Epinephelus fuscoguttatus*), a
+shallow-reef snapper (*Lutjanus malabaricus*), a deep-water snapper
+(*Pristipomoides filamentosus*), and a reef shark (*Carcharhinus
+amblyrhynchos*). The simulated groupers undergo a mass migration to a
+spawning aggregation once per year, followed by the sharks. Shallow-reef
+snappers stay in reefs closer to shore above a steep drop-off
+year-round, while deep-water snappers stay in the deeper reefs past the
+drop-off (@fig-coral).
+
+These species are targeted by two different fleets. Fleet One primarily
+targets the grouper and near-shore snapper populations, but will land
+any incidentally captured sharks. Fleet One has a logistic selectivity
+pattern for all species, as they retain any fish caught for consumption
+or sale. both grouper and shallow-reef snapper due to demand for "plate
+sized" fish , and logistic selectivity for sharks, as any shark caught
+is retained. Fleet One is totally dependent on fishing for their
+livelihood, meaning the local community takes advantage of every
+possible opportunity to fish, and as such we model it as a "constant
+effort" fishery. Do to having less efficient boats, Fleet One has a
+higher cost per distance coefficient than Fleet Two. Fleet One's home
+port is located on the site of the grouper spawning aggregation.
+
+Fleet Two is a more commercial fleet that primarily targets the
+deep-water snapper population. This fleet primarily sells their catch to
+local restaurants and distributors where plate-sized fish are prized,
+and so for both snapper and grouper Fleet Two has a dome-shaped
+selectivity pattern [@kindsvater2017]. While plate-sized deep snapper
+are the primary target of Fleet Two, we model Fleet Two's selectivity
+for deep snapper as logistic due to high levels of discard mortality for
+deep-water snapper resulting from barotrauma. Fleet Two catches
+groupers, though less than Fleet One, and receives no price for sharks
+due to the requirements of a certification program through which they
+sell their deep-water snapper. Accidental captures (bycatch) of sharks
+do occur, which results in mortality. Fleet Two operates under
+open-access dynamics, as fishing is not the only means of subsistence
+for this community; short-term effort expands and contracts in response
+to profitability of the primarily grouper-driven fishery. Fleet two's
+home port is located further north from Fleet One.
+
+We use `marlin` to simulate the outcomes for both food production and
+conservation for each of the species and both of the fleets as a
+function of both MPA size and MPA design strategy. For this exercise,
+MPAs are placed with perfect information and have no design constraints
+for continuity (i.e. the selected MPA design can be as patchwork as
+dictated by the data). We ran the simulation on a quarterly level
+($\Delta_t = 1/4$) and set the area of each patch to be 5km^2^
+($\Delta_p = \sqrt{5}$), using 400 patches for a total area of 2,000
+KM^2^.
 
@@
 
-We model our pelagic use case loosely on the characteristics of the Western and Central Pacific Ocean (WCPO) tuna fisheries.
-We simulate trajectories of `r n_distinct(names(blue_fauna))` species commonly caught in the region, including both the highly abundant skipjack tuna and the heavily depleted oceanic whitetip shark (*Carcharhinus longimanus*) (@fig-blue-ssb).
-We use publicly available data on catch-per-unit-effort of each of these species from the WCPO as a very rough proxy for baseline habitat distributions, noting that where possible, fishery-independent abundance indices would be preferable (@fig-blue).
-
-These pelagic species are caught by a longline fleet that primarily targets large adult tunas such as bigeye and yellowfin for high-grade consumption, and a purse seine fishery that primarily targets skipjack tunas for bulk canning.
-Selectivities were modeled as logistic for the longline fleet, and dome-shaped for the purse seine fleet, with selectivity values set to broadly match the patterns reported in relevant [stock assessments](https://oceanfish.spc.int/en/ofpsection/sam/sam/210-stock-assessment-input-and-result-files-) for each species.
-Both fleets operate under open-access dynamics with an effort cap.
-The effort cap was set at the level of effort that resulted in the desired levels of SSB/SSB0 for each species under open-access dynamics (@fig-blue), intended to simulate a scenario where managers step in to prevent further expansion of fishing effort in a fully developed fishery.
-For forward-simulation, open-access dynamics can result in effort decreasing in response to profitability, but cannot result in effort beyond the effort cap set for each fleet.
-
-For this exercise, we focused on using `marlin` to assess resilience of the selected Target Fishing and Rate "rule of thumb" MPA design strategies to a climate-driven range shift.
-Specifically, we simulate an extreme example where the centroid of each population shifts northward at a rate of 0.33 patches per year, over a 20 year time horizon (@fig-blue-ssb).
-We designed MPA networks given the conditions in the starting year, and then held that network constant over the years of the experiment, running one simulation with and another without the climate-drive range shift.
-We then compared the effects of this range shift on food production and conservation outcomes from MPA networks designed based on the pre-range shift world.
-We ran the pelagic simulation at a quarterly level ($\Delta_t$ = 1/4 year), and set the area of each cell to be roughly 97,000 KM^2^ across 144 patches each with a side length of roughly 311 KM, for a total area of 14,000,000 KM^2^, broadly commensurate with the area of the WCPO.
+We model our pelagic use case loosely on the characteristics of the
+Western and Central Pacific Ocean (WCPO) tuna fisheries. We simulate
+trajectories of `r n_distinct(names(blue_fauna))` species commonly
+caught in the region, including both the highly abundant skipjack tuna
+and the heavily depleted oceanic whitetip shark (*Carcharhinus
+longimanus*) (@fig-blue-ssb). We use publicly available data on
+catch-per-unit-effort of each of these species from the WCPO as a very
+rough proxy for baseline habitat distributions, noting that where
+possible, fishery-independent abundance indices would be preferable
+(@fig-blue).
+
+These pelagic species are caught by a longline fleet that primarily
+targets large adult tunas such as bigeye and yellowfin for high-grade
+consumption, and a purse seine fishery that primarily targets skipjack
+tunas for bulk canning. Selectivities were modeled as logistic for the
+longline fleet, and dome-shaped for the purse seine fleet, with
+selectivity values set to broadly match the patterns reported in
+relevant [stock
+assessments](https://oceanfish.spc.int/en/ofpsection/sam/sam/210-stock-assessment-input-and-result-files-)
+for each species. Both fleets operate under open-access dynamics with an
+effort cap. The effort cap was set at the level of effort that resulted
+in the desired levels of SSB/SSB0 for each species under open-access
+dynamics (@fig-blue), intended to simulate a scenario where managers
+step in to prevent further expansion of fishing effort in a fully
+developed fishery. For forward-simulation, open-access dynamics can
+result in effort decreasing in response to profitability, but cannot
+result in effort beyond the effort cap set for each fleet.
+
+For this exercise, we focused on using `marlin` to assess resilience of
+the selected Target Fishing and Rate "rule of thumb" MPA design
+strategies to a climate-driven range shift. Specifically, we simulate an
+extreme example where the centroid of each population shifts northward
+at a rate of 0.33 patches per year, over a 20 year time horizon
+(@fig-blue-ssb). We designed MPA networks given the conditions in the
+starting year, and then held that network constant over the years of the
+experiment, running one simulation with and another without the
+climate-drive range shift. We then compared the effects of this range
+shift on food production and conservation outcomes from MPA networks
+designed based on the pre-range shift world. We ran the pelagic
+simulation at a quarterly level ($\Delta_t$ = 1/4 year), and set the
+area of each cell to be roughly 97,000 KM^2^ across 144 patches each
+with a side length of roughly 311 KM, for a total area of 14,000,000
+KM^2^, broadly commensurate with the area of the WCPO.
 
@@
 
-For the coastal coral reef example both the *Spawning Ground* and *Target fishing* MPA design strategies were capable of producing a Pareto-optimal or improving, i.e.
-potentially win-win, outcome where change in total spawning biomass across all species and yield across all fleets was greater than or equal to zero for some MPA size.
-However, producing a total increase in SSB/SSB0 greater than 50% came at a net cost to fisheries yield across both design strategies.
-The *Spawning Aggregation* strategy produced more immediate trade-offs between food security and conservation (@fig-coral-results).
-
-These macro-level changes in total yield and conservation outcomes mask interesting internal dynamics.
-The *Spawning Ground* strategy increased yield to Fleet One when MPAs were relatively small, but was able to maintain yields near *status quo* levels up to a very large MPA size to the dependence of Fleet One on only a relatively small portion of the total simulation area.
-However, *Spawning Ground* MPAs were unable to provide increases in yield to Fleet Two.
-The *Spawning Ground* strategy produced the most rapid conservation gains for the snapper population.
-Interestingly, due to effort displacement from the spawning grounds onto the fishing areas inhabited by the groupers the rest of the year, closure of the spawning aggregation had limited impact on the grouper population until the MPA size exceeded 50% of the simulation area.
-Effort displacement from the spawning grounds also resulted in a small decrease in the deep snapper population for MPA sizes less than 50%.
-The *Target Fishing* design strategy has similar yield outcomes to the *Spawning Ground* strategy, but less dramatic negative impacts on Fleet Two, at the expense of slightly fewer benefits for Fleet One.
-However, the *Target Fishing* strategy provided more uniform conservation benefits to all species (@fig-coral-results).
-
-MPAs were capable of producing positive outcomes in terms of total biomass and yields in the coastal coral reef example, but all the tested strategies resulted in trade-offs across at least one dimension.
-`marlin` could be further used by a community such as the one simulated here to explore alternative design strategies, or to assess options for distributing fishery benefits across communities that may experience very different food provision outcomes from the same MPA network.
+For the coastal coral reef example both the *Spawning Ground* and
+*Target fishing* MPA design strategies were capable of producing a
+Pareto-optimal or improving, i.e. potentially win-win, outcome where
+change in total spawning biomass across all species and yield across all
+fleets was greater than or equal to zero for some MPA size. However,
+producing a total increase in SSB/SSB0 greater than 50% came at a net
+cost to fisheries yield across both design strategies. The *Spawning
+Aggregation* strategy produced more immediate trade-offs between food
+security and conservation (@fig-coral-results).
+
+These macro-level changes in total yield and conservation outcomes mask
+interesting internal dynamics. The *Spawning Ground* strategy increased
+yield to Fleet One when MPAs were relatively small, but was able to
+maintain yields near *status quo* levels up to a very large MPA size to
+the dependence of Fleet One on only a relatively small portion of the
+total simulation area. However, *Spawning Ground* MPAs were unable to
+provide increases in yield to Fleet Two. The *Spawning Ground* strategy
+produced the most rapid conservation gains for the snapper population.
+Interestingly, due to effort displacement from the spawning grounds onto
+the fishing areas inhabited by the groupers the rest of the year,
+closure of the spawning aggregation had limited impact on the grouper
+population until the MPA size exceeded 50% of the simulation area.
+Effort displacement from the spawning grounds also resulted in a small
+decrease in the deep snapper population for MPA sizes less than 50%. The
+*Target Fishing* design strategy has similar yield outcomes to the
+*Spawning Ground* strategy, but less dramatic negative impacts on Fleet
+Two, at the expense of slightly fewer benefits for Fleet One. However,
+the *Target Fishing* strategy provided more uniform conservation
+benefits to all species (@fig-coral-results).
+
+MPAs were capable of producing positive outcomes in terms of total
+biomass and yields in the coastal coral reef example, but all the tested
+strategies resulted in trade-offs across at least one dimension.
+`marlin` could be further used by a community such as the one simulated
+here to explore alternative design strategies, or to assess options for
+distributing fishery benefits across communities that may experience
+very different food provision outcomes from the same MPA network.
 
@@
 
-The yield impacts of the proposed MPA strategies varied across each of our simulations.
-Yields of the purse seine fleet were more resilient to the range shift under the *Rate* design strategy, while the longline fleet had the exact opposite result.
-Under the *Rate* strategy, longline yields were relatively stable across a large MPA size under the *status quo* conditions, but declined rapidly under the *range shift* conditions.
-For the purse seine fleet, yields were more stable under the *range shift* conditions under the *Target Fishing* design scenario, but declined quickly as a function of MPA size under the *status quo* conditions (@fig-blue-results).
-In terms of fisheries yield, the *Target Fishing* strategy would produce better yield outcomes for both fishing fleets under the *range shift* conditions, but the *Rate* strategy would perform best under the *status quo*.
-
-From a conservation perspective, the primary tuna species (bigeye, skipjack, and yellowfin) were most sensitive to design strategy and climate scenario, with the *Target Fishing* strategies producing for example rapid conservation gains for bigeye under the *status quo*, while the *Rate* strategy resulted in net conservation losses until the MPA became extremely large.
-However, under the *Range Shift* scenario the MPAs had little impact on many of the tuna populations until the MPA size became extremely large, do to movement of the primary fishing grounds outside of the current hotpots.
-The 20 year time horizon simulated here was not enough to produce substantial gains for any of the shark species even with 100% closures except for the more rapidly growing blue shark (@fig-blue-results).
+The yield impacts of the proposed MPA strategies varied across each of
+our simulations. Yields of the purse seine fleet were more resilient to
+the range shift under the *Rate* design strategy, while the longline
+fleet had the exact opposite result. Under the *Rate* strategy, longline
+yields were relatively stable across a large MPA size under the *status
+quo* conditions, but declined rapidly under the *range shift*
+conditions. For the purse seine fleet, yields were more stable under the
+*range shift* conditions under the *Target Fishing* design scenario, but
+declined quickly as a function of MPA size under the *status quo*
+conditions (@fig-blue-results). In terms of fisheries yield, the *Target
+Fishing* strategy would produce better yield outcomes for both fishing
+fleets under the *range shift* conditions, but the *Rate* strategy would
+perform best under the *status quo*.
+
+From a conservation perspective, the primary tuna species (bigeye,
+skipjack, and yellowfin) were most sensitive to design strategy and
+climate scenario, with the *Target Fishing* strategies producing for
+example rapid conservation gains for bigeye under the *status quo*,
+while the *Rate* strategy resulted in net conservation losses until the
+MPA became extremely large. However, under the *Range Shift* scenario
+the MPAs had little impact on many of the tuna populations until the MPA
+size became extremely large, do to movement of the primary fishing
+grounds outside of the current hotpots. The 20 year time horizon
+simulated here was not enough to produce substantial gains for any of
+the shark species even with 100% closures except for the more rapidly
+growing blue shark (@fig-blue-results).
 
@@
 Marine ecosystems are driven by complex social-ecological dynamics.
-Communities must often make decisions on how to manage these systems based on limited empirical evidence.
-Modeling tools such as `marlin` can help users answer important scientific questions and design marine management policies informed by a better understanding of sensitivities to key uncertainties.
-In this paper we have documented the core system of equations underlying `marlin`, and demonstrated how `marlin` can be used to support efforts such as MPA planning.
-The variation in both the magnitude and distribution of costs and benefits stemming from alternative design choices and states of nature shown in our case study results demonstrates the value in our method.
-For example, our coastal coral reef example illustrates both the potential for MPAs to benefit conservation and food production in these systems, and the potential for the same MPA to benefit some fleets and species while harming others when fishing fleets affected by an MPA do not share the same objectives and species are not uniformly distributed.
-Our pelagic case study illustrates how `marlin` can be used to assess the climate resiliency of spatial management strategies, and identify strategies that best meet the needs of both current and future conditions.
-Results from this type of work could be used to help prioritize communities and species at particular risk to climate change impacts.
-
-The role of this paper is not to conduct a review of the many tools for spatial-temporal modeling available in the literature, each of which provide useful functionality for specific applications.
-However, we highlight here the specific gaps that we feel `marlin` fills in the modeling literature and in the policy support toolbox using some selected publications.
-More end-to-end models such as ATLANTIS [@audzijonyte2019] , OSMOSE [@shin2001], Ecopath with Ecosim [@christensen2004], POSEIDON [@bailey2019], DISPLACE [@bastardie2013], or SEAPODYM [@lehodey2008] are capable of representing tremendous amounts of complexity, but can be extremely time consuming to design and run.
-`marlin` allows some of the realism of these end-to-end models while being simpler and faster to parameterize and run.
-
-Tools such as virtualspecies [@leroy2016], STEPS [@visintin2020], RangeShifter [@bocedi2014], and [SMS](https://www.aquaveo.com/software/sms-surface-water-modeling-system-introduction) can be efficiently constructed to model the dispersal and distribution of species in space and time as a function of environmental covariates.
-However, these typically require specifying a covariate-response curve without explicitly acknowledging how this arises from habitat-specific movement and demography (virtualspecies), or model movement using a dispersal kernel (RangeShifter, STEPS) or a least-cost path algorithm (SMS, STEPS).
-By contrast, `marlin` uses a continuous-time Markov chain movement model, which integrates multiple paths (including their path-dependent probability based on intervening habitat types and species preferences) while using scale-free parameters that can be measured experimentally in laboratory or tagging studies.
-
-Simplified bio-economic models such as those used in @hastings2017, @sala2021, and @cabral2019 can be applied at scale and provide analytically tractable results, but as a result must abstract over many bio-economic dynamics that can be important for more tactical applications.
-Our results demonstrate how nuances in fleet dynamics and species distributions can dramatically impact MPA outcomes.
-`marlin` allows for more realistic representations of spatial social-ecological systems while maintaining processing speed.
-
-In sum, `marlin` complements the existing spatial marine modeling literature by allowing scientists, decision makers, and other stakeholders to efficiently examine the impacts of realistic bio-economic dynamics on academic and applied problems.
-For example, scientists could use `marlin` to consider the relative importance of different biological variables in affecting policy outcomes.
-Managers could use `marlin` to assess the potential of proposed MPA networks to achieve stated objectives, such as maximizing conservation gains for a threatened species while minimizing or removing negative impacts to affected fishing fleets.
-
-Beyond the current examples, we envision `marlin` being applicable to research on dynamic ocean management, range shifts, management strategy evaluation, policy interactions, and spatial stock assessment.
-In addition, the process-based movement model used in `marlin` can directly use empirical estimates of movement dynamics derived in the manner of @thorson2021a, providing a link between empirical and simulation based approaches to marine resource management.
-Future iterations of the model may include life history modules for a broader range of species types such as marine mammals, sea birds, and sea turtles.
-
-Even the most complex marine model is a stylized cartoon of the true dynamics at play in the ocean.
-However, for all their limitations, models such as `marlin` can help users understand factors that drive the performance of marine management strategies.
-The goal of this tool is to empower people to design policies based on evaluation of key uncertainties and trade-offs, and in doing so support more effective and equitable marine resource management.
-
-# Data Availability Statement
-
-All code and materials needed to reproduce this paper is available at (https://github.com/DanOvando/marlin-paper)\[https://github.com/DanOvando/marlin-paper\].
-
-# References
+Communities must often make decisions on how to manage these systems
+based on limited empirical evidence. Modeling tools such as `marlin` can
+help users answer important scientific questions and design marine
+management policies informed by a better understanding of sensitivities
+to key uncertainties. In this paper we have documented the core system
+of equations underlying `marlin`, and demonstrated how `marlin` can be
+used to support efforts such as MPA planning. The variation in both the
+magnitude and distribution of costs and benefits stemming from
+alternative design choices and states of nature shown in our case study
+results demonstrates the value in our method. For example, our coastal
+coral reef example illustrates both the potential for MPAs to benefit
+conservation and food production in these systems, and the potential for
+the same MPA to benefit some fleets and species while harming others
+when fishing fleets affected by an MPA do not share the same objectives
+and species are not uniformly distributed. Our pelagic case study
+illustrates how `marlin` can be used to assess the climate resiliency of
+spatial management strategies, and identify strategies that best meet
+the needs of both current and future conditions. Results from this type
+of work could be used to help prioritize communities and species at
+particular risk to climate change impacts.
+
+The role of this paper is not to conduct a review of the many tools for
+spatial-temporal modeling available in the literature, each of which
+provide useful functionality for specific applications. However, we
+highlight here the specific gaps that we feel `marlin` fills in the
+modeling literature and in the policy support toolbox using some
+selected publications. More end-to-end models such as ATLANTIS
+[@audzijonyte2019] , OSMOSE [@shin2001], Ecopath with Ecosim
+[@christensen2004], POSEIDON [@bailey2019], DISPLACE [@bastardie2013],
+or SEAPODYM [@lehodey2008] are capable of representing tremendous
+amounts of complexity, but can be extremely time consuming to design and
+run. `marlin` allows some of the realism of these end-to-end models
+while being simpler and faster to parameterize and run.
+
+Tools such as virtualspecies [@leroy2016], STEPS [@visintin2020],
+RangeShifter [@bocedi2014], and
+[SMS](https://www.aquaveo.com/software/sms-surface-water-modeling-system-introduction)
+can be efficiently constructed to model the dispersal and distribution
+of species in space and time as a function of environmental covariates.
+However, these typically require specifying a covariate-response curve
+without explicitly acknowledging how this arises from habitat-specific
+movement and demography (virtualspecies), or model movement using a
+dispersal kernel (RangeShifter, STEPS) or a least-cost path algorithm
+(SMS, STEPS). By contrast, `marlin` uses a continuous-time Markov chain
+movement model, which integrates multiple paths (including their
+path-dependent probability based on intervening habitat types and
+species preferences) while using scale-free parameters that can be
+measured experimentally in laboratory or tagging studies.
+
+Simplified bio-economic models such as those used in @hastings2017,
+@sala2021, and @cabral2019 can be applied at scale and provide
+analytically tractable results, but as a result must abstract over many
+bio-economic dynamics that can be important for more tactical
+applications. Our results demonstrate how nuances in fleet dynamics and
+species distributions can dramatically impact MPA outcomes. `marlin`
+allows for more realistic representations of spatial social-ecological
+systems while maintaining processing speed.
+
+In sum, `marlin` complements the existing spatial marine modeling
+literature by allowing scientists, decision makers, and other
+stakeholders to efficiently examine the impacts of realistic
+bio-economic dynamics on academic and applied problems. For example,
+scientists could use `marlin` to consider the relative importance of
+different biological variables in affecting policy outcomes. Managers
+could use `marlin` to assess the potential of proposed MPA networks to
+achieve stated objectives, such as maximizing conservation gains for a
+threatened species while minimizing or removing negative impacts to
+affected fishing fleets.
+
+Beyond the current examples, we envision `marlin` being applicable to
+research on dynamic ocean management, range shifts, management strategy
+evaluation, policy interactions, and spatial stock assessment. In
+addition, the process-based movement model used in `marlin` can directly
+use empirical estimates of movement dynamics derived in the manner of
+@thorson2021a, providing a link between empirical and simulation based
+approaches to marine resource management. Future iterations of the model
+may include life history modules for a broader range of species types
+such as marine mammals, sea birds, and sea turtles.
+
+Even the most complex marine model is a stylized cartoon of the true
+dynamics at play in the ocean. However, for all their limitations,
+models such as `marlin` can help users understand factors that drive the
+performance of marine management strategies. The goal of this tool is to
+empower people to design policies based on evaluation of key
+uncertainties and trade-offs, and in doing so support more effective and
+equitable marine resource management.
+
+## Acknowledgements
+
+This project was funded by the Waitt Foundation. We thank Dr. Alexa Fredston for helpful revisions of this manuscript.
+
+## Author Contribitions
+
+Methods and code written by DO and JT. Results produced by DO. Data collected by EB and LT. All authors contributed to the conceptualization and writing of the manuscript.
+
+## Conflit of Interest Statement
+
+The authors declare no conflicts of interest.
+
+## References
 
@@
 :::
-
-# Supporting Information

### 2023-03-30 README.Rmd
prepping for edits on squid
Final steps (#8)
* final model runs with updated CTMC parameterization and Lorenzen mortality
---------

Co-authored-by: dovando_iattc <danovan@iattc.org>

@@
 
-4. In the terminal, run `RUN_NAME:MyRunName FORMAT:MyFormat make` where MyRunName is the name of the folder you want to create to store results (e.g. v0.1) and FORMAT is one of pdf, html, or docx
+4. In the terminal, run `RUN_NAME="MyRunName" FORMAT="MyFormat" make` where MyRunName is the name of the folder you want to create to store results (e.g. v0.1) and FORMAT is one of pdf, html, or docx
 
-5. Wait for a bit! This command will run all the required scripts and knit the paper. Should take no more than a few hours.
+5. Wait for a bit! This command will run all the required scripts and knit the paper. Shouldn't take more than an hour on a reasonable computer with parallel computation capabilities.

### 2023-03-30 marlin_paper_coverletter.qmd
Final steps (#8)
* final model runs with updated CTMC parameterization and Lorenzen mortality
---------

Co-authored-by: dovando_iattc <danovan@iattc.org>

---
format: pdf
editor: visual
---

Dear Editors,

We are writing to submit our manuscript titled "Simulating Benefits, Costs, and Tradeoffs of Management Policies in Marine Social-Ecological Systems" for consideration at Ecological Applications. Spatial management policies such as marine protected areas are increasingly being looked to to help manage and conserve marine ecosystems. However , designing
spatial management strategies to achieve desired objectives -- which may
include recovery and resilience of overfished species, increased food
production and economic well being, and equitable distribution of
benefits - is not straightforward due to the complex, dynamic, and opaque nature of marine social-ecological systems.

In this paper we present a bio-economic model intended to facilitate the study and design of marine management policies called [`marlin`](https://github.com/DanOvando/marlin/). `marlin` allows users to efficiently simulate the dynamics of multiple species and fishing fleets across space and time, using realistic representations of life history, economic decision making, and movement dynamics.

We show in this paper how `marlin` can be used to help inform policy decisions related to the design and placement of marine protected areas through two case studies, and discuss how this tool can help answer a wider array of research and management questions in marine social-ecological systems.

Thank you for your consideration,

Dan Ovando & Co-Authors

### 2023-03-30 ovando_tunacamp2023_abstract.qmd
Final steps (#8)
* final model runs with updated CTMC parameterization and Lorenzen mortality
---------

Co-authored-by: dovando_iattc <danovan@iattc.org>

---
title: "Revealing Benefits, Costs, and Tradeoffs of Spatial Management Arrising from Coupled Economic and Ecological Dynamics in Marine Systems"
author:
    - Daniel Ovando
address: "8901 La Jolla Shores Drive,La Jolla, CA 92037, USA"
affiliation: "Inter-American Tropical Tuna Comission"
format:
  docx: 
    reference-doc: tuna_template.docx
---

There is a growing movement to expand the use of spatial management measures such as Marine Protected Areas (MPAs) in both coastal and pelagic ecosystems. However, designing effective spatial management strategies is challenging because marine ecosystems are highly dynamic and opaque, and extractive entities such as fishing fleets respond endogenously to ecosystem changes in ways that depend upon ecological and policy context. We present a tool, [marlin](https://danovando.github.io/marlin/), that can be used to efficiently simulate the bio-economic dynamics of coastal and pelagic marine systems in support of both management and research. A key feature of `marlin` is that it allows for realistic simulation of fish movement dynamics in response to changing environmental conditions or foraging opportunities through use of a continuous-time Markov chain process. We demonstrate `marlin's` capabilities in two case studies on the conservation and food production impacts of MPAs: a coastal coral reef and a pelagic tuna fishery. In the coastal coral reef example, we show how heterogeneity in species distributions and fleet preferences can affect distributional outcomes of MPAs. In the pelagic tuna fishery case study, we show how `marlin` can simulate the effects of alternative MPA design strategies on the biomass of pelagic species and the economics of fishing fleets under climate-driven range shifts. Our work demonstrates how intermediate complexity simulation of coupled bio-economic dynamics can help communities predict and potentially manage trade-offs between conservation, fisheries yields, and distributional outcomes of management policies affected by spatial bio-economic dynamics in coastal and pelagic ecosystems.

### 2023-03-30 00_setup.R

@@ -33,6 +33,8 @@ library(rmarkdown)
 
 library(quarto)
 
+library(extrafont)
+
 options(dplyr.summarise.inform = FALSE)
 
 
@@ -61,7 +63,7 @@ run_coral_example <- TRUE
 
 run_blue_water_example <- TRUE
 
-theme_set(marlin::theme_marlin())
+theme_set(marlin::theme_marlin(base_size = 12))
 
 
 # load data ---------------------------------------------------------------
@@ -79,3 +81,15 @@ if (!file.exists(here("data","marlin-inputs.xlsx"))){
 
 marlin_inputs <- readxl::read_xlsx(here("data","marlin-inputs.xlsx"), sheet = "inputs",na = c("NA",""))
 
+if (!file.exists(here("data","sci_to_com.csv"))){
+  sci_to_com <- taxize::sci2comm(unique(marlin_inputs$scientific_name),simplify= TRUE,db='ncbi') %>%
+    bind_rows(.id = "critter") %>%
+    pivot_longer(everything(),names_to = "critter",values_to = "common_name") %>%
+    mutate(common_name = stringr::str_to_title(common_name))
+
+  write_csv(sci_to_com, file = here("data","sci_to_com.csv"))
+
+} else{
+  sci_to_com <-  read_csv(file = here("data","sci_to_com.csv"))
+
+}

### 2023-03-30 02_blue_water_case_study.R

@@ -1,8 +1,22 @@
 source(file.path("00_setup.R"))
 
-resolution <- 12
+resolution <- 12 # number of patches per side of simulation grid
+
+patches <- resolution ^ 2
+
+# target simulation area: 14,000,000 km2 = 3741 side length
+
+patch_area <- 14e6 / patches # km2
+
+patch_side <- sqrt(patch_area) # km
+
+simulation_area <- patch_area * patches #km2
+
+message(glue::glue("Blue water simulation area is {scales::comma(patch_area)} km2, spread over {patches} patches each with an area of {scales::comma(patch_area)} km2 or {scales::comma(patch_side)} km per side"))
 
 seasons <- 4
+
+time_step <- 1 / seasons # in years
 
 tune_type <- "explt"
 
@@ -10,6 +24,9 @@ experiment_workers <- 8
 
 experiment_years <- 20
 
+adult_diffusion <- 0.8 * simulation_area  # km2, which means that each year a fish can roam an area 4km
+
+max_hab_mult = 20
 
 # load bycatch risk layers ------------------------------------------------
 
@@ -114,7 +131,9 @@ get_layer <- function(file) {
 
   hab$interp_cpue <- as.numeric(predict(mod$gam))
 
-  hab$habitat <- pmax(0, hab$interp_cpue / max(hab$interp_cpue))
+  hab$habitat <- scales::rescale(hab$interp_cpue, c(0,log(20)))
+
+  # hab$habitat <- pmax(0, hab$interp_cpue / max(hab$interp_cpue) * adult_diffusion)
 
 
   if (sqrt(nrow(hab)) != resolution) {
@@ -166,9 +185,10 @@ casestudy <- casestudy %>%
     create_critters,
     marlin_inputs = marlin_inputs,
     seasons = seasons,
-    taxis_to_diff_ratio = 2,
-    adult_diffusion = 10
-  )) %>%
+    adult_diffusion = adult_diffusion,
+    resolution = resolution,
+    max_hab_mult = max_hab_mult,
+    patch_area = patch_area)) %>%
   mutate(xid = 1)
 
 casestudy <- casestudy %>%
@@ -183,7 +203,6 @@ check_pop_sizes <- map_dbl(casestudy$fauna[[1]], "ssb0")
 #   geom_col() +
 #   coord_flip()
 
-image(casestudy$fauna[[1]]$`prionace glauca`$base_movement[[1]])
 
 
 # tune fleet --------------------------------------------------------------
[…]

### 2023-03-30 functions/create_critters.R

@@ -7,32 +7,25 @@ create_critters <-
            ontogenetic_shift = FALSE,
            random_rec = FALSE,
            adult_diffusion = 10,
-           taxis_to_diff_ratio = 2) {
-    # sciname <- marlin_inputs$scientific_name[[1]]
-    #
-    #
+           resolution = resolution,
+           max_hab_mult = 5,
+           patch_area = 1) {
 
-    # habitat <- experiments$habitat[[1]]
     tmp_inputs <- marlin_inputs %>%
-      filter(scientific_name == sciname)
+      filter(scientific_name == sciname & !is.na(scientific_name))
 
     hab <- habitat %>%
       pivot_wider(names_from = y, values_from = habitat) %>%
       select(-x) %>%
       as.matrix()
 
-    # seasonal_movement <- sample(c(1, 0), 1, replace = TRUE)
 
-    # uniform_rec_hab <- sample(c(1, 0), 1, replace = TRUE)
-
-
-    # seasonal_movement <- 0
-    if (seasonal_movement) {
-      hab <- list(hab,-hab - (min(-hab)))
-
-    } else {
-      hab <- list(hab, hab)
-    }
+    # if (seasonal_movement) {
+    #   hab <- list(hab,-hab - (min(-hab)))
+    #
+    # } else {
+    #   hab <- list(hab, hab)
+    # }
 
     if (random_rec){
       density_dependence <-  sample(c("global_habitat", "local_habitat","pre_dispersal","post_dispersal","global_ssb"), 1, replace = TRUE)
@@ -41,24 +34,12 @@ create_critters <-
       density_dependence <- "global_ssb" # local density dependence then disperse recruits per recruit movement
     }
 
-    if (ontogenetic_shift) {
-      recruit_habitat <-
-        -hab[[1]] - min(-hab[[1]]) # place recruits in different places than adults
-
-      # set recruitment form to allow for recruit habitat
-      #
-      density_dependence <- "global_ssb"
-    } else {
-      recruit_habitat <- hab[[1]]
-    }
-
     critter <- marlin::create_critter(
       scientific_name = sciname,
-      base_habitat = hab,
-      recruit_habitat = recruit_habitat,
+      habitat = hab,
+      recruit_habitat = hab,
       adult_diffusion = adult_diffusion,
[…]

### 2023-03-30 functions/titler.R

@@ -1,0 +1,6 @@
+# simple function to make prettier labels
+titler <- function(x){
+
+  stringr::str_to_title(stringr::str_replace_all(x,"_"," "))
+
+}

### 2023-04-03 03_marlin_paper.qmd
added option for coastal MPA design ands fixed some equation alignments

@@
 $$
-n_{t,p,s,a = 1} = \left(\frac{0.8{\times}\sum_{p=1}^P{r0_{p,s}}\times{h_s}\times\sum_{p=1}^P{ssb_{t-1,p,s}}}{0.2\times{\sum_{p=1}^Pssb0_{p,s}}\times(1 - h_s)+(h_s - 0.2)\times{\sum_{p=1}^Pssb_{t-1,p,s}}}\right)\times {r0_{p,s} / \sum_p^P{r0_{p,s}}\times \epsilon_{t,s}
+\begin{aligned}
+n_{t,p,s,a = 1} = \left(\frac{0.8{\times}\sum_{p=1}^P{r0_{p,s}}\times{h_s}\times\sum_{p=1}^P{ssb_{t-1,p,s}}}{0.2\times{\sum_{p=1}^Pssb0_{p,s}} \times(1 - h_s)+(h_s - 0.2) \times{\sum_{p=1}^Pssb_{t-1,p,s}}}\right) \times \\ {r0_{p,s} / \sum_p^P{r0_{p,s}}\times \epsilon_{t,s}
 }
+\end{aligned}
 $$ {#eq-dd2}
@@
 $$
-n_{t,p,s,a = 1} = \left(\frac{0.8{\times}\sum_{p=1}^P{r0_{p,s}}\times{h_s}\times\sum_{p=1}^P{ssb_{t-1,p,s}}}{0.2\times{\sum_{p=1}^Pssb0_{p,s}}\times(1 - h_s)+(h_s - 0.2)\times{\sum_{p=1}^Pssb_{t-1,p,s}}}\right)\times \frac{ssb_{t-1,p,s}}{\sum_{p=1}^Pssb_{t-1,p,s}} \times \epsilon_{t,s}
+\begin{aligned}
+n_{t,p,s,a = 1} = \left(\frac{0.8{\times}\sum_{p=1}^P{r0_{p,s}}\times{h_s}\times\sum_{p=1}^P{ssb_{t-1,p,s}}}{0.2\times{\sum_{p=1}^Pssb0_{p,s}}\times(1 - h_s)+(h_s - 0.2)\times{\sum_{p=1}^Pssb_{t-1,p,s}}}\right) \times \\ \frac{ssb_{t-1,p,s}}{\sum_{p=1}^Pssb_{t-1,p,s}} \times \epsilon_{t,s}
+\end{aligned}
 $$ {#eq-dd5}

### 2023-04-03 03_marlin_paper.qmd
slight edits to movement notation

@@
 
-We provide a brief overview of the the general CTMC method here, see
-@thorson2021a for a detailed description. Under this framework, movement
+We provide a brief overview of the the general CTMC method here (see
+@thorson2021a for a detailed description). Under this framework, movement of indivuduals
 from each patch to each other patch in the system in a given timestep
@@
 
-$$
-\boldsymbol{M}_{t,s,a} = e^{\Delta_{t}(\boldsymbol{D}_{t,s,a} + \boldsymbol{\tau}_{t,s,a})} 
-$$ {#eq-movement}
-
-The individual components (*m*) of the movement matrix are filled based
+The individual components (*M*) of the movement matrix ($\boldsymbol{M}$) are filled based
 on an adjacency matrix, which defines whether two patches are both
@@
 $$ 
-m_{p1,p2,t,s,a} = \begin{cases}
+M_{p1,p2,t,s,a} = \begin{cases}
       = \frac{\Delta_{t}}{\Delta_{d}^2}De^{\frac{\Delta_t(h(p2,t,s,a) - h(p1,t,s,a))}{\Delta_d}} & \text{if p2 and p1 are adjacent}\\
-     = -\sum_{p' \neq p1} m_{p1,p2,t,s,a} & \text{if p1 = p2}\\
+     = -\sum_{p' \neq p1} M_{p1,p2,t,s,a} & \text{if p1 = p2}\\
      = 0 & \text{otherwise.}
@@
 that an animal must move around to transit from one side to another.
+
+The movement of individuals across patches is then calculated by matrix multiplication of the pre-movement vector of the number of individuals ($\boldsymbol{n}$) of species *s* at age *a* in time step *t* across all patches *p* times the matrix exponential of the movement matrix $\boldsymbol{M}$
+
+$$
+\boldsymbol{n}_{t+1,s,a} = \boldsymbol{n}_{t,s,a}e^{\boldsymbol{M}_{t,s,a}}
+$$ {#eq-movement}
 

### 2023-04-03 03_marlin_paper.qmd
fun with math notation latex issues

--- 2023-04-03 03_marlin_paper.qmd
@@
   - \usepackage{lineno}\linenumbers
+  - \usepackage{bm}
 ---
@@
 
-We provide a brief overview of the the general CTMC method here, see
-@thorson2021a for a detailed description. Under this framework, movement
+We provide a brief overview of the the general CTMC method here (see
+@thorson2021a for a detailed description). Under this framework, movement of individuals
 from each patch to each other patch in the system in a given timestep
 *t* for life stage *a* of species *s* is defined by a movement matrix
-$\boldsymbol{M}_{t,s,a}$. $\boldsymbol{M}_{t,s,a}$ is calculated as the
-matrix exponent of a function of the diffusion $\boldsymbol{D}$ and
-taxis $\boldsymbol{\tau}$ matrices scaled by the width of the time step
+$\pmb{M}_{t,s,a}$. $\pmb{M}_{t,s,a}$ is calculated as a function of diffusion $\pmb{D}$ and
+taxis $\pmb{\tau}$ matrices scaled by the width of the time step
 $\Delta_{t}$ and the length of the edge of each patch $\Delta_d$ that
-the model is running under. So for example, if we are running a model on
-a monthly timestep, and we define the units of time in years, then
-$\Delta_{t} = 1/12$. Similarly, if we are running a model where one
-patch equals 100km^2^, then $\Delta_d = 10km$. This parameterization
+the model is running under. This parameterization
 allows users to set the effective area of the spatial domain through two
@@
 
-$$
-\boldsymbol{M}_{t,s,a} = e^{\Delta_{t}(\boldsymbol{D}_{t,s,a} + \boldsymbol{\tau}_{t,s,a})} 
-$$ {#eq-movement}
-
-The individual components (*m*) of the movement matrix are filled based
+The individual components (*M*) of the movement matrix ($\pmb{M}$) are filled based
 on an adjacency matrix, which defines whether two patches are both
@@
 diffusion rate $D$ defined in units of area of a patch per unit of time,
-and a habitat preference function *h* in units of length of a side of a
+and a habitat preference function *H* in units of length of a side of a
 patch per unit time. For example, if we are defining the time units as
@@
 $\Delta_{d}$ parameters to translate the diffusion rate $D$ to match the
-time step and patch size used in a simulation. This "scale free"
+time step and patch size used in a simulation. For example, if we were to run a model on
+a monthly timestep given time units of years, then
+$\Delta_{t} = 1/12 years$. If one square patch in the simulation has an area of 100km^2^, then $\Delta_d = 10KM$. This "scale free"
 parameterization means that appropriate value of $D$ can be identified
@@
 used in the simulation model itself. The taxis component of the movement
-process is a function of the difference in habitat quality *h*. The
+process is a function of the difference in habitat quality *H*. The
 habitat preference function itself can take any form the user wishes.
@@
 multipliers of the diffusion rate. This parameterization ensure that the
-off-diagonal elements of the movement matrix $\boldsymbol{M}_{t,s,a}$
+off-diagonal elements of the movement matrix $\pmb{M}_{t,s,a}$
 are all non-negative, a requirement of the CTMC method.
@@
 $$ 
-m_{p1,p2,t,s,a} = \begin{cases}
-      = \frac{\Delta_{t}}{\Delta_{d}^2}De^{\frac{\Delta_t(h(p2,t,s,a) - h(p1,t,s,a))}{\Delta_d}} & \text{if p2 and p1 are adjacent}\\
-     = -\sum_{p' \neq p1} m_{p1,p2,t,s,a} & \text{if p1 = p2}\\
+M_{p1,p2,t,s,a} = \begin{cases}
+      = \frac{\Delta_{t}}{\Delta_{d}^2}De^{\frac{\Delta_t(H(p2,t,s,a) - H(p1,t,s,a))}{\Delta_d}} & \text{if p2 and p1 are adjacent}\\
+     = -\sum_{p' \neq p1} M_{p1,p2,t,s,a} & \text{if p1 = p2}\\
      = 0 & \text{otherwise.}
@@
 
+The movement of individuals across patches is then calculated by matrix multiplication of the pre-movement vector of the number of individuals ($\pmb{n}$) of species *s* at age *a* in time step *t* across all patches *p* times the matrix exponential of the movement matrix $\pmb{M}$
+
+$$
+\pmb{n}_{t+1,s,a} = \pmb{n}_{t,s,a}e^{\pmb{M}_{t,s,a}}
+$$ {#eq-movement}
+
 While this CTMC approach to movement simulation is to date not commonly
 seen in the marine modeling literature, it has numerous advantages that
-warrant its seeming complexity. First, when only diffusion is present,
+warrant its seeming complexity. First, the parameters of the model have interpretable biological meaning (e.g. the diffusion rate *D*). Second, when only diffusion is present,
 the model will generalize to the familiar dynamics of a Gaussian
-dispersal kernel. However, the time component *t* in @eq-movement allows
-the movement model to be run on arbitrarily fine time steps. Second, the
-taxis model allow for more active habitat choices by species, allowing
+dispersal kernel at whatever spatial and temporal resolution the simulation is set to. Third, the
+taxis model allow for clearly parameterized active habitat choices by species, allowing
 us to simulate preferences of species in space and time efficiently.
-Lastly, this form has the advantage that the parameters are directly
+Lastly, the CTMC form has the advantage that its parameters are directly
 estimable from real data. So, if provided with for example spatial
@@
 $$
-  n_{t,p,s,a = 1} = \left(\frac{0.8{\times}r0_{p,s}\times{h_s}\times{ssb_{t-1,p,s}}}{0.2\times{ssb0_{p,s}}\times(1 - h_s)+(h_s - 0.2)\times{ssb_{t-1,p,s}}}\right)\times \boldsymbol{d^l_s} \times \epsilon_{t,s}
+  n_{t,p,s,a = 1} = \left(\frac{0.8{\times}r0_{p,s}\times{h_s}\times{ssb_{t-1,p,s}}}{0.2\times{ssb0_{p,s}}\times(1 - h_s)+(h_s - 0.2)\times{ssb_{t-1,p,s}}}\right)\times \pmb{d^l_s} \times \epsilon_{t,s}
 $$ {#eq-dd3}
@@
 $$
-  larv_{t,p,s} = ssb_{t-1,p,s} \times\boldsymbol{d^l_s}
+  larv_{t,p,s} = ssb_{t-1,p,s} \times\pmb{d^l_s}
 $$ {#eq-larvmove}
@@
 
-<!-- obj_p = \Phi \times scale(\sum_f^F \boldsymbol{\phi_{f,p}}) + (1 - \Phi)  scale(\boldsymbol{\sum_s^S\frac{ssb_s}{ssb0_{s,p}}}) -->
+<!-- obj_p = \Phi \times scale(\sum_f^F \pmb{\phi_{f,p}}) + (1 - \Phi)  scale(\pmb{\sum_s^S\frac{ssb_s}{ssb0_{s,p}}}) -->
 

### 2023-04-27 03_marlin_paper.qmd
initial fandf submission

@@
   cache: false
-csl: chicago-author-date-16th-edition.csl
+csl: apa-6th-edition.csl
 bibliography: references.bib
@@
 
+<!-- referees -->
+
+<!-- Maya Tolstoy -->
+
+<!-- Mikaela Provost -->
+
+<!-- Leah Gerber -->
+
+<!-- David M. Kaplan -->
+
 ```{r setup}
@@
 
-<!-- Single column, double line spaced -->
-
-<!-- Within the word count (no more than 7000-8000 words for Standard Articles, 3000-4000 words for Applications and Practical Tools) -->
-
-<!-- Continuous line and page numbering throughout -->
-
-<!-- Clearly defined manuscript structure as standard: Author details, Abstract (must be numbered 1-4 according to Manuscript Specifications), Keywords, Introduction, Materials and Methods, Results, Discussion, Figures and Tables with captions -->
-
-<!-- Figures and Tables can be embedded within the text where referenced to facilitate reviewing -->
-
-<!--         Author Contributions (if >1 author – see Authorship Policy). -->
-
-<!--         Statement of where you intend to archive your data -->
-
 ## Title Page
 
-Journal Name: Ecological Applications
-
-Manuscript Type: Article
-
-Manuscript Title: Simulating Benefits, Costs, and Tradeoffs of Management Policies in Marine Social-Ecological Systems
-
-Author Names: Daniel Ovando^1,2^, Darcy Bradley^3,4,5^, Echelle Burns^3,4,5^,Lennon Thomas^3,4,5^, James Thorson^6^
-
-Affiliations:
-^1^ School of Aquatic and Fishery Sciences, University of Washington,1122 NE Boat St Box 355020,Seattle, WA, USA
-^2^ Current Address: Inter-American Tropical Tuna Commission,8901 La Jolla Shores Drive, CA, USA
-La Jolla, CA 92037, USA
-^3^ Marine Science Institute, University of California, Santa Barbara, CA, USA
-^4^ Bren School of Environmental Science & Management, University of California, Santa Barbara, CA, USA
-^5^ Environmental Markets Lab, University of California, Santa Barbara, CA, USA
-^6^ Habitat and Ecological Processes Research Program, Alaska Fisheries Science Center, NOAA, NMFS, NOAA, Seattle, WA, USA
-
-Corresponding Author: Daniel Ovando <dovando@iattc.org>
-
-Open Research Statement: All code and materials needed to reproduce this paper is available at [https://github.com/DanOvando/marlin-paper](https://github.com/DanOvando/marlin-paper).
-
-Key Words: 
-Bio-economic modeling, spatio-temporal modeling, marine protected areas, fisheries management, marine spatial planning, conservation planning
+Title: Simulating Benefits, Costs, and Tradeoffs of Spatial Management
+in Marine Social-Ecological Systems
+
+Running Title: Spatial Social-Ecological Simulation
+
+Author Names: Daniel Ovando^1,2^, Darcy Bradley^3,4,5^, Echelle
+Burns^3,4,5^,Lennon Thomas^3,4,5^, James Thorson^6^
+
+Affiliations: ^1^ School of Aquatic and Fishery Sciences, University of
+Washington,1122 NE Boat St Box 355020,Seattle, WA, USA ^3^ Marine
+Science Institute, University of California, Santa Barbara, CA, USA ^4^
+Bren School of Environmental Science & Management, University of
+California, Santa Barbara, CA, USA ^5^ Environmental Markets Lab,
+University of California, Santa Barbara, CA, USA ^6^ Habitat and
+Ecological Processes Research Program, Alaska Fisheries Science Center,
+NOAA, NMFS, NOAA, Seattle, WA, USA
+
+Corresponding Author: Daniel Ovando
+[dovando\@iattc.org](mailto:dovando@iattc.org){.email}
+
+Present Address: \^2 Inter-American Tropical Tuna Commission,8901 La
+Jolla Shores Drive, La Jolla, CA 92037, USA
+
+Author Contributions: Methods and code written by DO and JT. Results
+produced by DO. Data collected by EB and LT. All authors contributed to
+the conceptualization and writing of the manuscript.
+
+Conflict of Interest Statement: The authors declare no conflicts of
+interest.
 
@@
 
-Designing effective spatial management strategies is challenging because marine ecosystems are highly dynamic and opaque, and extractive entities such as fishing fleets respond endogenously to ecosystem changes in ways that depend upon ecological and policy context.
-We present a tool, `marlin`, that can be used to efficiently simulate the bio-economic dynamics of marine systems in support of both management and research.
-We demonstrate `marlin’s` capabilities by focusing on two case studies on the conservation and food production impacts of marine protected areas (MPAs): a coastal coral reef and a pelagic tuna fishery.
-In the coastal coral reef example, we explore how heterogeneity in species distributions and fleet preferences can affect distributional outcomes of MPAs.
-In the pelagic case study, we show how our model can be used to assess the climate resilience of different MPA design strategies, as well as the climate sensitivity of different fishing fleets.
-This paper demonstrates how intermediate complexity simulation of coupled bio-economic dynamics can help communities predict and potentially manage trade-offs between conservation, fisheries yields, and distributional outcomes of management policies affected by spatial bio-economic dynamics.
+Designing effective spatial management strategies is challenging because
+marine ecosystems are highly dynamic and opaque, and extractive entities
+such as fishing fleets respond endogenously to ecosystem changes in ways
+that depend upon ecological and policy context. We present a tool,
+`marlin`, that can be used to efficiently simulate the bio-economic
+dynamics of marine systems in support of both management and research.
+We demonstrate `marlin’s` capabilities by focusing on two case studies
+on the conservation and food production impacts of marine protected
+areas (MPAs): a coastal coral reef and a pelagic tuna fishery. In the
+coastal coral reef example, we explore how heterogeneity in species
+distributions and fleet preferences can affect distributional outcomes
+of MPAs. In the pelagic case study, we show how our model can be used to
+assess the climate resilience of different MPA design strategies, as
+well as the climate sensitivity of different fishing fleets. This paper
+demonstrates how intermediate complexity simulation of coupled
+bio-economic dynamics can help communities predict and potentially
+manage trade-offs between conservation, fisheries yields, and
+distributional outcomes of management policies affected by spatial
+bio-economic dynamics.
+
+## Keywords
+
+Bio-economic modeling, spatio-temporal modeling, marine protected areas,
+fisheries management, marine spatial planning, conservation planning
+
+## Table of Contents
+
+1.  Introduction
+2.  Methods
+    1.  Model Summary
+    2.  Fleet Dynamics
+    3.  Population Model
+    4.  Case Studies
+3.  Results
+    1.  Coral Reefs
+    2.  Pelagic Systems
+4.  Discussion
+5.  Acknowledgements
+6.  Data Availability Statement
+7.  References
 
@@
 We provide a brief overview of the the general CTMC method here (see
-@thorson2021a for a detailed description). Under this framework, movement of individuals
-from each patch to each other patch in the system in a given timestep
-*t* for life stage *a* of species *s* is defined by a movement matrix
-$\pmb{M}_{t,s,a}$. $\pmb{M}_{t,s,a}$ is calculated as a function of diffusion $\pmb{D}$ and
-taxis $\pmb{\tau}$ matrices scaled by the width of the time step
-$\Delta_{t}$ and the length of the edge of each patch $\Delta_d$ that
-the model is running under. This parameterization
-allows users to set the effective area of the spatial domain through two
-avenues; the number of patches, which effectively scales the resolution
-of the model, and the area of each patch, which scales the spatial
-extent of the simulation.
-
-The individual components (*M*) of the movement matrix ($\pmb{M}$) are filled based
-on an adjacency matrix, which defines whether two patches are both
-adjacent and water (as opposed to land or another physical barrier), a
-diffusion rate $D$ defined in units of area of a patch per unit of time,
-and a habitat preference function *H* in units of length of a side of a
-patch per unit time. For example, if we are defining the time units as
-years and the distance units as kilometers, for a tuna $D$ might be
-1,000 $\frac{KM^2}{Year}$. We then use parameters $\Delta_{t}$ and
-$\Delta_{d}$ parameters to translate the diffusion rate $D$ to match the
-time step and patch size used in a simulation. For example, if we were to run a model on
-a monthly timestep given time units of years, then
-$\Delta_{t} = 1/12 years$. If one square patch in the simulation has an area of 100km^2^, then $\Delta_d = 10KM$. This "scale free"
+@thorson2021a for a detailed description). Under this framework,
+movement of individuals from each patch to each other patch in the
+system in a given timestep *t* for life stage *a* of species *s* is
+defined by a movement matrix $\pmb{M}_{t,s,a}$. $\pmb{M}_{t,s,a}$ is
+calculated as a function of diffusion $\pmb{D}$ and taxis $\pmb{\tau}$
+matrices scaled by the width of the time step $\Delta_{t}$ and the
+length of the edge of each patch $\Delta_d$ that the model is running
+under. This parameterization allows users to set the effective area of
+the spatial domain through two avenues; the number of patches, which
+effectively scales the resolution of the model, and the area of each
+patch, which scales the spatial extent of the simulation.
+
+The individual components (*M*) of the movement matrix ($\pmb{M}$) are
+filled based on an adjacency matrix, which defines whether two patches
+are both adjacent and water (as opposed to land or another physical
+barrier), a diffusion rate $D$ defined in units of area of a patch per
+unit of time, and a habitat preference function *H* in units of length
+of a side of a patch per unit time. For example, if we are defining the
+time units as years and the distance units as kilometers, for a tuna $D$
+might be 1,000 $\frac{KM^2}{Year}$. We then use parameters $\Delta_{t}$
+and $\Delta_{d}$ parameters to translate the diffusion rate $D$ to match
+the time step and patch size used in a simulation. For example, if we
+were to run a model on a monthly timestep given time units of years,
+then $\Delta_{t} = 1/12 years$. If one square patch in the simulation
+has an area of 100km^2^, then $\Delta_d = 10KM$. This "scale free"
 parameterization means that appropriate value of $D$ can be identified
@@
 multipliers of the diffusion rate. This parameterization ensure that the
-off-diagonal elements of the movement matrix $\pmb{M}_{t,s,a}$
-are all non-negative, a requirement of the CTMC method.
+off-diagonal elements of the movement matrix $\pmb{M}_{t,s,a}$ are all
+non-negative, a requirement of the CTMC method.
 
@@
 
-The movement of individuals across patches is then calculated by matrix multiplication of the pre-movement vector of the number of individuals ($\pmb{n}$) of species *s* at age *a* in time step *t* across all patches *p* times the matrix exponential of the movement matrix $\pmb{M}$
+The movement of individuals across patches is then calculated by matrix
+multiplication of the pre-movement vector of the number of individuals
+($\pmb{n}$) of species *s* at age *a* in time step *t* across all
+patches *p* times the matrix exponential of the movement matrix
+$\pmb{M}$
 
@@
 seen in the marine modeling literature, it has numerous advantages that
-warrant its seeming complexity. First, the parameters of the model have interpretable biological meaning (e.g. the diffusion rate *D*). Second, when only diffusion is present,
-the model will generalize to the familiar dynamics of a Gaussian
-dispersal kernel at whatever spatial and temporal resolution the simulation is set to. Third, the
-taxis model allow for clearly parameterized active habitat choices by species, allowing
-us to simulate preferences of species in space and time efficiently.
-Lastly, the CTMC form has the advantage that its parameters are directly
-estimable from real data. So, if provided with for example spatial
-abundance data and a tagging study, users can estimate the diffusion and
-taxis movement parameters in the same manner as @thorson2021a, and then
-pass their estimated parameters to `marlin` for simulation (so long as
-the estimating method uses the same functional form as the movement
-model in `marlin`).
+warrant its seeming complexity. First, the parameters of the model have
+interpretable biological meaning (e.g. the diffusion rate *D*). Second,
+when only diffusion is present, the model will generalize to the
+familiar dynamics of a Gaussian dispersal kernel at whatever spatial and
+temporal resolution the simulation is set to. Third, the taxis model
+allow for clearly parameterized active habitat choices by species,
+allowing us to simulate preferences of species in space and time
+efficiently. Lastly, the CTMC form has the advantage that its parameters
+are directly estimable from real data. So, if provided with for example
+spatial abundance data and a tagging study, users can estimate the
+diffusion and taxis movement parameters in the same manner as
+@thorson2021a, and then pass their estimated parameters to `marlin` for
+simulation (so long as the estimating method uses the same functional
+form as the movement model in `marlin`).
 
@@
 #| label: fig-blue-ssb
-#| fig-cap: Spawning biomass divided by unfished spawning biomass in the time period prior to implementation of MPAs for the pelagic case study."
+#| fig-cap: "Spawning biomass divided by unfished spawning biomass in the time period prior to implementation of MPAs for the pelagic case study."
 
@@
 
-# Results
+## Results
 
@@
 
-# Discussion
+## Discussion
 
@@
 
-This project was funded by the Waitt Foundation. We thank Dr. Alexa Fredston for helpful revisions of this manuscript.
-
-## Author Contribitions
-
-Methods and code written by DO and JT. Results produced by DO. Data collected by EB and LT. All authors contributed to the conceptualization and writing of the manuscript.
-
-## Conflit of Interest Statement
-
-The authors declare no conflicts of interest.
+This project was funded by the Waitt Foundation. We thank Dr. Alexa
+Fredston for helpful revisions of this manuscript.
+
+## Data Availability Statement
+
+The code, novel code, data, and materials needed to reproduce the
+results and all aspects of this manuscript are publicly available at
+https://github.com/DanOvando/marlin-paper, with supporting novel code
+available at https://github.com/DanOvando/marlin/. Upon acceptance, all
+data, materials, code, and novel code needed to reproduce the results
+and all aspects of this manuscript will be publicly available via GitHub
+at https://github.com/DanOvando/marlin-paper and
+https://github.com/DanOvando/marlin/ and through figshare at
+https://figshare.com/articles/preprint/marlin-paper/21843582
 

### 2023-05-23 presentations/impac_slides.qmd
testing slides deploy

---
title: "Assessing the Population-level Impacts of Marine Protected Areas"
subtitle: "IMPAC 2023"
author: "Dan Ovando"
date: "2023/02/04"
format:
  revealjs:
    fig-align: center
    auto-stretch: true
  html: default
  pptx:
    reference-doc: "template.pptx"
execute: 
  eval: true
  message: false
  warning: false
---

```{r}
#| label: setup
#| include: false
library(knitr)

library(tufte)

library(tidyverse)

library(here)

```

## Population-Effects of MPAs

:::: {.columns}
::: {.column width="60%"}
- Thanks to co-authors and PISCO divers
- MPAs a growing part of marine management toolkit
- *Predicting* and *estimating* effects critical for planning and adaptation

**What effects do protected areas have on populations inside and outside of borders?**
:::
::: {.column width="40%"}
![](imgs/paper_cover.png)
:::
::::

## MPAs have Complex Effects {.smaller}

:::: {.columns}
::: {.column width="50%"}

Fishing closures have complex impacts beyond borders

- "spillover" of adult or larval organisms

- Displacement of human activities

- Across multiple species and fleets!

**Creates a challenge for both predicting and estimating effects of spatial policies**

:::

::: {.column width="50%"}
![](imgs/IMG_1679.JPG)
:::
::::

## MPAs - The Evidence {.smaller}

:::: {.columns}
::: {.column width="50%}

:::{.incremental}
- MPAs seem simple: 
  - *Less fishing, more fish*

- Ample evidence of **more everything** inside MPAs than outside (Lester *et al.* 2009)

- Problem solved?

  - Fish and Fishers move

  - What do we do when the **treatment** affects the **control**

  How do we estimate **net effect** inside and out?
:::
:::
::: {.column width="50%"}
![](imgs/lester.jpg)
:::
::::

## Population-Level Effects of MPAs

![](imgs/bio_anim.gif)

::: notes
Relatively sedentary adults, widely dispersed larvae with post-settlement density dependence
:::

## Population-Level Effects of MPAs

![](imgs/fleet_anim.gif)

::: notes
Relatively sedentary adults, widely dispersed larvae with post-settlement density dependence
:::

## The Challenge of Spatial MPA evaluation

::: {.incremental}

- The goal of a control site is to account for **differences and changes caused by things other than the MPA that you do not directly observe**

- The **closer** your control site, the more risk of spillover and fleet movement affecting your "control"

- The **further** your control site, the less likely it will experience common environmental shocks

- Is there an alternative to spatial controls?
:::
## Some useful tips here

![](imgs/hilborn.png)

## Case Study: Channel Islands MPAs

![](imgs/ci.png)

## What Happend in the Channel Islands?

![](imgs/raw_targ_trend.png){fig.align="center"}

## What Happend in the Channel Islands?

![](imgs/raw_trend.png)

## Estimating Population-Level Effects

![](imgs/did.001.jpeg)

## Estimating Population-Level Effects

![](imgs/did.002.jpeg)

---

## Estimating Population-Level Effects

![](imgs/did.003.jpeg)

---

## Estimating Population-Level Effects

### General form

$$(log(D_{MPA=1, T = 1})-log(D_{MPA=0, T = 1}))-(log(D_{MPA = 1,T = 0})-log(D_{MPA =0, T = 0}))$$

### Bayesian Gamma GLM

$$d_{i} \sim Gamma(e^{\beta_0 + \beta_1T_{i} +  \beta_2MPA_{i} + \beta_{3}T_iMPA_i + \mathbf{B^cX_i} + \mathbf{B^sS_i}},shape)$$

### Hierarchical clustering of sites by island
$$\mathbf{B^s} \sim N(\beta_r,\sigma_r)$$

## Population-Level Effects of MPAs

![](imgs/a_priori_plot.png)

## Population-Level Effects of MPAs {.nostretch}

![](imgs/mpa_effect_plot.png)

## Effects of Protected Areas {.smaller}

::: columns
::: {.column width="50%"}

- Critical to consider effects **inside** and **outside**

-  **Population** effects may be smaller and harder to find than conventionally thought

- Uncertainty does not preclude action

- Communities should have a clear understanding of full range of possible effects of protected areas

:::
::: {.column width="50%"}
![](imgs/IMG_2947.JPG)
:::
:::

## The Next Generation of MPA Models

:::: {.columns}
::: {.column width="60%"}
Empirical strategies should...

- Acknowledge uncertainty
- Have a clear and defensible strategy for causal inference
- Be grounded in theory

:::
::: {.column width="40%"}

Predictive models should...

- Assess where they fall on the "tactical" to "strategic" spectrum
- Acknowledge, explore, and communicate key uncertainties
- Remember that all models are wrong, some are useful
:::
::::

## Reliable Estimation Matters

![](imgs/complexity.png)

## `marlin`

Be on the lookout for a new MPA modeling tool called [`marlin`](https://danovando.github.io/marlin/) in the coming year.

- Full age structure
- 2D spatial structure
- Multi-species & Multi-fleet
- Realistic movement dynamics

**Facilitate simulation testing of MPA design outcomes**

## Questions?

:::: {.columns}
::: {.column width="60%"}

**email:** dovando@iattc.org

**website:** [danovando.com](https://www.weirdfishes.blog/)

**twitter:** [@danovand0](https://twitter.com/DanOvand0)

**Funding:** NOAA Sea Grant Population and Ecosystem Dynamics Fellowship

:::
::: {.column width="40%"}

![](imgs/squats.jpg)
:::
::::

# Extras

## Inside vs. Outside = Effect?

![](imgs/rr1.jpg)

## Inside vs. Outside = Effect?

![](https://raw.githubusercontent.com/DanOvando/weird-fishes/master/static/blog/2021-07-06-pop-effects-of-mpas_files/biased_dr_plot.jpg)

## Details Matter

![](imgs/complexity)

## Details Matter

![](imgs/need.png)

## Effect Size Matters

![](imgs/ERRORS.PNG)

## MPAs - The Evidence

<br>
<br>
<br>

> The jury is in on marine reserves: They work.
> Research has repeatedly shown that fish numbers quickly climb following well-enforced fishing bans, 
> creating tangible benefits for fishers who work the surrounding waters.

[NPR 2018](https://www.npr.org/sections/thesalt/2018/09/14/647441547/could-a-ban-on-fishing-in-international-waters-become-a-reality)

### 2023-05-23 presentations/tunacamp_slides.qmd
testing slides deploy

---
title: "Revealing Benefits, Costs, and Tradeoffs of Spatial Management Arrising from Coupled Economic and Ecological Dynamics in Marine Systems"
author: "Dan Ovando"
format:
  revealjs:
    embed-resources: true
editor: visual
---

## Quarto

Quarto enables you to weave together content and executable code into a
finished presentation. To learn more about Quarto presentations see
<https://quarto.org/docs/presentations/>.

## Bullets

When you click the **Render** button a document will be generated that
includes:

-   Content authored with markdown
-   Output from executable code

## Code

When you click the **Render** button a presentation will be generated
that includes both content and the output of embedded code. You can
embed code like this:

```{r}
1 + 1
```

### 2023-05-23 presentations/tunacamp_slides.qmd
slides testing for tunacamp

@@
   revealjs:
-    embed-resources: true
-editor: visual
+    embed-resources: false
 ---
 
-## Quarto
+## Complexity Matters!
 
-Quarto enables you to weave together content and executable code into a
-finished presentation. To learn more about Quarto presentations see
-<https://quarto.org/docs/presentations/>.
-
-## Bullets
-
-When you click the **Render** button a document will be generated that
-includes:
-
--   Content authored with markdown
--   Output from executable code
-
-## Code
-
-When you click the **Render** button a presentation will be generated
-that includes both content and the output of embedded code. You can
-embed code like this:
-
-```{r}
-1 + 1
-```
+![](imgs/complexity.png)

### 2023-05-23 presentations/tunacamp_slides.qmd
embed-resources: true

@@
   revealjs:
-    embed-resources: false
+    embed-resources: true
 ---
+
+## Testing
+
+Hello
 

### 2023-05-23 presentations/tunacamp_slides.qmd
testing equations
pushing with less cmbedding

--- 2023-05-23 presentations/tunacamp_slides.qmd
@@
     embed-resources: false
+    fig-align: center
+    auto-stretch: true
+    theme: moon
 ---
+
+## Acknowledgements
+
+Daniel Ovando^1,2^, Darcy Bradley^3,4,5^, Echelle
+Burns^3,4,5^,Lennon Thomas^3,4,5^, James Thorson^6^
+
+Affiliations: ^1^ School of Aquatic and Fishery Sciences, University of
+Washington,1122 NE Boat St Box 355020,Seattle, WA, USA ^3^ Marine
+Science Institute, University of California, Santa Barbara, CA, USA ^4^
+Bren School of Environmental Science & Management, University of
+California, Santa Barbara, CA, USA ^5^ Environmental Markets Lab,
+University of California, Santa Barbara, CA, USA ^6^ Habitat and
+Ecological Processes Research Program, Alaska Fisheries Science Center,
+NOAA, NMFS, NOAA, Seattle, WA, USA
+
+## All Models are Wrong
+
+:::: {.columns}
+
+::: {.column width="60%"}
+
+- But some are hopefully useful for something?
+
+-
+
+:::
+
+::: {.column width="40%"}
+Right column
+:::
+
+::::
+
+## Lots of Spatio-Temporal Models Out There
+
+![](imgs/fulton-mpas.jpg)
+
+::: aside
+[Fulton *et al.* 2015](https://royalsocietypublishing.org/doi/10.1098/rstb.2014.0278)
+:::
+
+## The marlin package
+
+[`marlin`](https://danovando.github.io/marlin/) is an R package currently available on GitHub
+
+```{r}
+#| eval: false
+
+remotes::install_github("danovando/marlin")
+
+```
+
+Core Features
+
+- Full age structure of multiple fish(y) critters (automated life history with [FishLife](https://github.com/James-Thorson-NOAA/FishLife))
+
+- Multiple fleets and *métiers* enabling **technical interactions**
+  - Input controls, output controls, bio-economic dynamics
+
+- 2-D spatial structure with scale-free and estimable movement
+
+- Fast! (fractions of a second for 10x10 system over 20 years with quarterly steps)
+
+## The marlin Package
+
+Best suited for *what if* as opposed to *what will* questions
+
+  -
+
+## Fish Model
+
+## Movement Model
+
+## Fleet Model
+
+## Manager Model
+
+## Coral Reef Case Study
+
+![](../03_marlin_paper_files/figure-pdf/fig-coral-1.png)
+
+## Equations
+
+$$
+\begin{aligned}
+n_{t,p,s,a = 1} = \left(\frac{0.8{\times}\sum_{p=1}^P{r0_{p,s}}\times{h_s}\times\sum_{p=1}^P{ssb_{t-1,p,s}}}{0.2\times{\sum_{p=1}^Pssb0_{p,s}} \times(1 - h_s)+(h_s - 0.2) \times{\sum_{p=1}^Pssb_{t-1,p,s}}}\right) \times \\ {r0_{p,s} / \sum_p^P{r0_{p,s}}\times \epsilon_{t,s}
+}
+\end{aligned}
+$$ {#eq-dd2}
 
@@
 ![](imgs/complexity.png)
+
+## Linking to Team Tuna
+
+- Still stuck in the dietary dark ages 
+  - Diet data could be used to inform taxis matrix?
+
+- Humans need to be part of the "diet Renaissance"!
+
+- Provides a direct and mechanistic link between tagging studies and simulation work
+  - Can be linked to advection and environmental covariates as a bonus!
+
+- Future work adding in more species modules
+  - Seabirds?
+  - Marine mammals?
+  - Turtles?

### 2023-05-23 presentations/tunacamp_slides.qmd
stop tracking slides, not worth it

--- 2023-05-23 presentations/tunacamp_slides.qmd
@@
 ---
-title: "Revealing Benefits, Costs, and Tradeoffs of Spatial Management Arrising from Coupled Economic and Ecological Dynamics in Marine Systems"
+title: "Simulating Benefits, Costs, and Tradeoffs of Spatial Management in Marine Social-Ecological Systems"
+subtitle: "73rd Annual Tuna Conference"
 author: "Dan Ovando"
@@
   revealjs:
-    embed-resources: false
+    embed-resources: true
+    fig-align: center
+    auto-stretch: true
+    theme: [moon, custom.scss]
 ---
+
+## Acknowledgements
+
+**Daniel Ovando^1,2^, Darcy Bradley^3,4,5^, Echelle Burns^3,4,5^,Lennon
+Thomas^3,4,5^, James Thorson^6^**
+
+**Funding**: Waitt Foundation
+
+::: aside
+Affiliations: ^1^ School of Aquatic and Fishery Sciences, University of
+Washington,1122 NE Boat St Box 355020,Seattle, WA, USA ^3^ Marine
+Science Institute, University of California, Santa Barbara, CA, USA ^4^
+Bren School of Environmental Science & Management, University of
+California, Santa Barbara, CA, USA ^5^ Environmental Markets Lab,
+University of California, Santa Barbara, CA, USA ^6^ Habitat and
+Ecological Processes Research Program, Alaska Fisheries Science Center,
+NOAA, NMFS, NOAA, Seattle, WA, USA
+:::
+
+## All Models are Wrong
+
+But some are hopefully useful for something?
+
+- At the *strategic* level can help provide general insights around *what kinds of things matter*
+
+- At the *tactical* level can help provide *specific answers* when experiments are off the table
+
+. . .
+
+What is the best model of an elephant?
+
+. . .
+
+Depends on objectives
+
+## Lots of Spatio-Temporal Models Out There
+
+![](imgs/fulton-mpas.jpg)
+
+::: aside
+[Fulton *et al.*
+2015](https://royalsocietypublishing.org/doi/10.1098/rstb.2014.0278)
+:::
+
+## The marlin package {.smaller}
+
+[`marlin`](https://danovando.github.io/marlin/) is an R package
+currently available on GitHub
+
+```{r}
+#| eval: false
+
+remotes::install_github("danovando/marlin")
+
+```
+
+#### Core Features
+
+-   Full age structure of multiple fish(y) critters (automated life
+    history with
+    [FishLife](https://github.com/James-Thorson-NOAA/FishLife))
+
+-   Multiple fleets and *métiers* enabling **technical interactions**
+
+    -   Input controls, output controls, bio-economic dynamics
+
+    - Bycatch impacts
+
+    - Distributional impacts of policies on fish and people
+
+-   2-D spatial structure with scale-free and estimable movement
+
+-   Fast! (fractions of a second for 10x10 system over 20 years with
+    quarterly steps)
+
+# Help users test the impacts of bio-economic uncertainties on social-ecological outcomes
+
+. . .
+
+Best suited for *what if* as opposed to *what will* questions
+
+## Fish Model {.smaller}
+
+- Flexible time steps
+
+- Simultaneously simulate *N* age-structured fish(ish) populations
+
+- von Bertalanffy growth
+
+- Flexible fecundity at age
+
+- [Lorenzen natural mortality](https://linkinghub.elsevier.com/retrieve/pii/S0165783622002314) by default
+
+- 6 different timing of Beverton-Holt recruitment
+  - e.g. pre vs. post settlement
+  - With or without potentially auto-correlated recruitment deviates
+
+## Movement Model
+
+Movement simulated using a *continuous-time Markov chain (CTMC)^[[Thorson et al. 2021](https://onlinelibrary.wiley.com/doi/10.1111/faf.12592)]*
+
+- Movement governed by *diffusion* and *taxis* matrices
+
+  - Can be separate for larvae vs. adults
+
+  - See later releases for age-specific and *advection*
+
+  - Can be empirically estimated!
+
+  - Can dodge barriers (e.g. land)
+
+## Movement Model {.smaller}
+
+*Instantaneous* movement matrix is a function of diffusion rate *D*, habitat matrix *H*, and the spatial ($\Delta_d$) and temporal ($\Delta_t$) resolution
+
+$$ 
+M_{p1,p2,t,s,a} = \begin{cases}
+      = \frac{\Delta_{t}}{\Delta_{d}^2}De^{\frac{\Delta_t(H(p2,t,s,a) - H(p1,t,s,a))}{\Delta_d}} & \text{if p2 and p1 are adjacent}\\
+     = -\sum_{p' \neq p1} M_{p1,p2,t,s,a} & \text{if p1 = p2}\\
+     = 0 & \text{otherwise.}
+\end{cases}
+$$ {#eq-diffusion}
+
+. . .
+
+Numbers at age by patch then distributed each time step via matrix exponential
+
+$$
+\pmb{n}_{t+1,s,a} = \pmb{n}_{t,s,a}e^{\pmb{M}_{t,s,a}}
+$$ {#eq-movement}
+
+. . .
+
+Scale-free parameterization means parameters can be preserved and compared across models with different spatio-temporal resolution
+
+## {background-image=imgs/ctmc.gif background-size=contain}
+
+## Fleet Model
+
+- *N* fleets defined by *M* *metiers*
+
+- Fleets have
+  - Cost per unit effort
+  - Travel costs from ports
+
+- *metier* defined by
+  - Species
+  - Price
+  - Selectivity at age
+  - Catchability in space
+
+## Fleet Model {.smaller}
+
+Catch per species calculated by Baranov equation
+
+$$
+Y_{t,p,s,a} = \frac{u_{t,p,s,a}}{z_{t,p,s,a}} \times b_{t,p,s,a} \times (1 - e^{-z_{t,p,s,a}})
+$$ {#eq-yield}
+
+. . . 
+And then revenues divided to each fleet proportionally
+
+$$
+ R_{t,p,f} = \sum_s^{N_s} \sum_a^{N_a} \Pi_{s,f}\frac{\alpha_{a,s,f}q_{s,f}E_{t,p,f}}{u_{t,p,s,a}} \times Y_{t,p,s,a}
+$$ {#eq-revenue}
+
+. . .
+
+Fishers then allocate effort in space based on total profits in space in last time step (conditional on management...)^[Allows for economic MPAs!]
+
+## Fleet Model
+
+Absent management fleet can either be *constant effort*^[with or without MPA attrition] or *open access*
+
+When *open access*, fishing effort expands and contracts in response to total profits
+
+. . .
+
+$$
+E_{t+1,f} = E_{t,f} \times e^{\theta_f log(R_{t,f} / C_{t,f})} 
+$$ {#eq-oa}
+
+## Manager Model
+
+Managers can set arbitrary combinations of
+
+- Size / slot limits
+
+- Fishing seasons by species
+
+- Effort limits by species
+
+- Catch quotas by species
+
+- MPAs
+
+- "Bycatch" penalties
+
+## Coral Reef Case Study
+
+![](imgs/03_marlin_paper_files/figure-pdf/fig-coral-1.png){.absolute left=50 height="600"}
+
+## Coral Reef Case Study
+
+![](imgs/03_marlin_paper_files/figure-pdf/fig-coral-results-1.png){.absolute left=50 height="600"}
+
+## Pelagic Case Study
+
+![](imgs/03_marlin_paper_files/figure-pdf/fig-blue-1.png){.absolute left=50 height="600"}
+
+## Pelagic Case Study
+
+![](imgs/03_marlin_paper_files/figure-pdf/fig-blue-results-1.png){.absolute top=100 left=200 width="650" height="550"}
 
@@
 
+. . .
+
 ![](imgs/complexity.png)
+
+## Linking to Team Tuna {.smaller}
+
+-   Many models still stuck in the dietary dark ages
+
+    -   Diet data could be used to inform taxis matrix?
+
+-   Humans need to be part of the "diet Renaissance"!
+
+-   `marlin` provides a mechanistic link between tagging studies and
+    simulation work
+
+    -   Can be linked to advection and environmental covariates as a
+        bonus!
+
+-   Future work adding in more species modules
+
+    -   Seabirds?
+    -   Marine mammals?
+    -   Turtles?
+
+## Of MICE and `marlin`
+
+::: columns
+::: {.column width="60%"}
+
+- Spatial stock assessment and management increasingly needed
+
+- BBNJ and 30x30 means that "tunas" are now firmly in the great MPA debate
+
+- Lack of consistent / transparent assumptions can produce inconsistent outcomes
+
+- MICE models like `marlin` can help illuminate drivers of social-ecological outcomes in fisheries
+
+:::
+::: {.column width="40%"}
+
+![](imgs/bbnj.png)
+
+![](imgs/medoff.png)
+
+![](imgs/hampton.png)
+
+:::
+
+:::
+## Questions?
+
+::: columns
+::: {.column width="65%"}
+-   Be on the lookout for `marlin` paper sometime in 2023 (fingers
+    crossed)
+
+-   Check out the model at <br> [https://danovando.github.io/marlin/](https://danovando.github.io/marlin/)
+
+-   Open source! Contributions welcome
+
+-   email: dovando@iattc.org
+
+-   github: danovando
+
+-   twitter: @danovand0
+:::
+
+::: {.column width="35%"}
+![](imgs/IMG_1679.JPG)
+:::
+:::

### 2023-05-23 03_marlin_paper.qmd
Fandf (#9)
* initial fandf submission
Slides (#10)
* stopping slides branch 
* use     embed-resources: true to generate shareeable html files
* first real draft of tunacamp presentation

@@
     fontsize: 12pt
+    keep-tex: true
+    fig-format: png
   html: default
@@
 ---
-
-<!-- referees -->
-
-<!-- Maya Tolstoy -->
-
-<!-- Mikaela Provost -->
-
-<!-- Leah Gerber -->
-
-<!-- David M. Kaplan -->
 

### 2023-06-05 presentations/tunacamp_slides.qmd
updating lockfile

@@
 ---
+
+## Time for Models! {background-image="https://c.tenor.com/FtuByrhdiwIAAAAC/monty-python.gif" background-size=contain}
 
@@
     quarterly steps)
+      - R in the front, C++ in the back
+
+- Recursive (you can pass it to itself) for customized applications (e.g. MSE)
 
@@
 ## Fish Model {.smaller}
+
+::: columns
+::: {.column width="60%"}
 
@@
 
+- Flexible spawning seasons
+
+- We'll get to movement...
+:::
+::: {.column width="40%"}
+
+![](imgs/fishy.jpeg)
+
+:::
+:::
+
 ## Movement Model
@@
 
-  - See later releases for age-specific and *advection*
+  - See later releases for age-specific and *advection* (aka oceanography!)
 
@@
 . . .
+
+This is the important part
+
+$$ 
+\frac{\Delta_{t}}{\Delta_{d}^2}De^{\frac{\Delta_t(H(p2,t,s,a) - H(p1,t,s,a))}{\Delta_d}} 
+$$ {#eq-diffusion2}
+
+## Movement Model
 
@@
 
-Absent management fleet can either be *constant effort*^[with or without MPA attrition] or *open access*
+Conditional on management management fleet can either be *constant effort*^[with or without MPA attrition] or *open access*
 
@@
 
+::: {.notes}
+
+Fleet One primarily targets the grouper and near-shore snapper populations, but will land
+any incidentally captured sharks. Fleet One has a logistic selectivity
+pattern for all species, as they retain any fish caught for consumption
+or sale. both grouper and shallow-reef snapper due to demand for "plate
+sized" fish , and logistic selectivity for sharks, as any shark caught
+is retained. Fleet One is totally dependent on fishing for their
+livelihood, meaning the local community takes advantage of every
+possible opportunity to fish, and as such we model it as a "constant
+effort" fishery. Do to having less efficient boats, Fleet One has a
+higher cost per distance coefficient than Fleet Two. Fleet One's home
+port is located on the site of the grouper spawning aggregation.
+
+Fleet Two is a more commercial fleet that primarily targets the
+deep-water snapper population. This fleet primarily sells their catch to
+local restaurants and distributors where plate-sized fish are prized,
+and so for both snapper and grouper Fleet Two has a dome-shaped
+selectivity pattern [@kindsvater2017]. While plate-sized deep snapper
+are the primary target of Fleet Two, we model Fleet Two's selectivity
+for deep snapper as logistic due to high levels of discard mortality for
+deep-water snapper resulting from barotrauma. Fleet Two catches
+groupers, though less than Fleet One, and receives no price for sharks
+due to the requirements of a certification program through which they
+sell their deep-water snapper. Accidental captures (bycatch) of sharks
+do occur, which results in mortality. Fleet Two operates under
+open-access dynamics, as fishing is not the only means of subsistence
+for this community; short-term effort expands and contracts in response
+to profitability of the primarily grouper-driven fishery. Fleet two's
+home port is located further north from Fleet One.
+
+:::
+
 ## Coral Reef Case Study
@@
 
-![](imgs/03_marlin_paper_files/figure-pdf/fig-blue-results-1.png){.absolute top=100 left=200 width="650" height="550"}
+![](imgs/03_marlin_paper_files/figure-pdf/fig-blue-results-1.png){.absolute top=100 left=200 width="750" height="600"}
 
-## Complexity Matters!
+## Complexity Matters! (but it's costly)
 
@@
 
-## Of MICE and `marlin`
+## Of MICE and `marlin` {.smaller}
 
@@
 
-- MICE models like `marlin` can help illuminate drivers of social-ecological outcomes in fisheries
+- [MICE(ish)](https://research.csiro.au/mice/) models like `marlin` can help illuminate drivers of social-ecological outcomes in fisheries
 
@@
 -   twitter: @danovand0
+
+- website: [danovando.com](https://www.weirdfishes.blog/)
 :::

### 2023-07-25 03_marlin_paper.qmd
updating a few things to reflect some changes to marlin

@@
   map_df(coral_sim, ~ map_df(.x, ~tibble(biomass = rowSums(.x$ssb_p_a), patch = 1:nrow(.x$ssb_p_a)), .id = "critter"), .id = "step") %>%
-  mutate(step = as.numeric(step)) %>%
+  mutate(step = marlin::clean_steps(step)) %>%
   left_join(grid, by = "patch")
@@
   map_df(blue_sim, ~ map_df(.x, ~tibble(ssb = rowSums(.x$ssb_p_a), patch = 1:nrow(.x$ssb_p_a)), .id = "critter"), .id = "step") %>%
-  mutate(step = as.numeric(step)) %>%
+  mutate(step = marlin::clean_steps(step)) %>%
   left_join(grid, by = "patch") %>% 
@@
   map_df(blue_climate_sim, ~ map_df(.x, ~tibble(biomass = rowSums(.x$ssb_p_a), patch = 1:nrow(.x$ssb_p_a)), .id = "critter"), .id = "step") %>%
-  mutate(step = as.numeric(step)) %>%
+  mutate(step = marlin::clean_steps(step)) %>%
   left_join(grid, by = "patch")
@@
   geom_line(alpha = 0.8, aes(color = placement_strategy),size = 1.1) + 
-  geom_text_repel(data = frontier_labels, aes(label = pmpa, color = placement_strategy), min.segment.length = 0, box.padding = 0.2, show.legend = FALSE) +
+  geom_text_repel(data = frontier_labels, aes(label = pmpa, color = placement_strategy), min.segment.length = 0, box.padding = 0, show.legend = FALSE, nudge_x = .25, size = 5) +
   scale_x_continuous(name = "Change in Total SSB/SSB0", labels = scales::label_percent(accuracy = 1), guide = guide_axis(n.dodge = 2)) +

### 2023-07-25 01_coral_case_study.R

@@ -219,7 +219,7 @@ reef_shark <- create_critter(
   max_hab_mult = max_hab_mult,
   patch_area = patch_area,
   lorenzen_m = FALSE,
-  ssb0 = 10
+  ssb0 = 100
 )
 
 # critters
@@ -372,7 +372,7 @@ patch_biomass <-
     biomass = rowSums(.x$b_p_a),
     patch = 1:nrow(.x$b_p_a)
   ), .id = "critter"), .id = "step") %>%
-  mutate(step = as.numeric(step)) %>%
+  mutate(step = marlin::clean_steps(step)) |>
   left_join(grid, by = "patch")
 
 patch_biomass %>%
@@ -388,7 +388,7 @@ patch_biomass %>%
 
 biomass <-
   map_df(reef_sim, ~ map_df(.x, ~ tibble(biomass = sum(.x$b_p_a)), .id = "critter"), .id = "step") %>%
-  mutate(step = as.numeric(step))
+  mutate(step = marlin::clean_steps(step))
 
 
 biomass %>%
@@ -411,7 +411,7 @@ biomass %>%
 
 fleet_catch <-
   map_df(reef_sim, ~ map_df(.x, ~ colSums(.x$c_p_fl), .id = "critter"), .id = "step") %>%
-  mutate(step = as.numeric(step)) |>
+  mutate(step = marlin::clean_steps(step)) |>
   pivot_longer(starts_with("fleet_"),
                names_to = "fleet",
                values_to = "catch")
@@ -438,7 +438,7 @@ patch_effort %>%
 
 effort <-
   map_df(reef_sim, ~ data.frame(effort = sum(.x$grouper$e_p_fl$fleet_two)), .id = "step") %>%
-  mutate(step = as.numeric(step))
+  mutate(step = marlin::clean_steps(step))
 
 effort %>%
   ungroup() %>%
@@ -448,7 +448,7 @@ effort %>%
 
 profits <-
   map_df(reef_sim, ~ map_df(.x, ~ tibble(profit = colSums(.x$prof_p_fl)), .id = "critter"), .id = "step") %>%
-  mutate(step = as.numeric(step))
+  mutate(step = marlin::clean_steps(step))
 
 profits %>%
   ungroup() %>%
@@ -469,7 +469,7 @@ patch_biomass <-
     biomass = rowSums(.x$ssb_p_a),
     patch = 1:nrow(.x$ssb_p_a)
   ), .id = "critter"), .id = "step") %>%
[…]

### 2023-09-28 referee-replies.qmd
getting somewhere

---
title: "Referee Replies"
format: pdf
---
## Notes
- emphasize that this is one model that doesn't capture a lot of stuff

"Spatial Heterogeneous" as definition for the kind of stock structure modeled here, goes from "panmictic" to "Spatial Heterogeneous", per the definitions in bosley et el. 2021. Acutally it can range all the way to metapopulation (set density dependence timing to local, shut down movement at one or more life stages)

## R1

Comments to the Author
This paper presents interesting, well-documented, and useful modeling software for spatial management of fisheries targeting multiple non-interacting species. I could see it becoming a go-to for both fisheries scientists and advanced fisheries courses. It should be published following minor revisions.

**We thank the reviewer for their helpful comments. **

My main suggestion for revision is to explain and emphasize the case-study results a bit more. Ultimately, the insights from the case studies will be one of the main things that convince readers to use the model. For example, what is the intuition behind the results in Figure 5A, showing target-fishing design being much more robust to range shifts than rate design? More generally, what are the most important things we should take away from the case studies?

**We have added a specific "Insights from Case Studies" section to the discussion to provide clearer insights from our results.**

Second, the authors should clarify that their model has both strategic and tactical uses, not just tactical. There are a lot of basic strategic questions that could be asked with this model. The case studies are arguably strategic as much or more so than they are tactical.

**We have added a "General Recommendations for using `marlin`" section to the discussion, in which we have emphasized this point amount others.**

Minor comments:
Line 116: typo ‘th’.
**Corrected.**

Line 169: This seems like a potentially strong assumption that could artificially eliminate weak-stock-strong-stock tradeoffs. This should be discussed and justified.

**This is a very good point, particularly as 0% discard mortality is clearly the exception rather than the rule. We agree that this is too strong of assumption to use in many cases, and are working to implement both a discard rate and a discard mortality function into the model. However, as seasonal management is not a key feature of the model at this time, and is not used in our case study, we have elected to remove the mention of seasonal closures for now from the manuscript, though we are pursuing the addition of this feature in the package itself.**

XX Probably not going to do this yet since it is a pain in the ass XX. 
**To address this concern, we have modified the model such that "closed" seasons are now implemented by a multiplier on fishing mortality that can range from 0 to 1. When the multiplier is set to 0, the closed season is 100% effective (i.e. no discard mortality). Setting the multiplier to 0.5 instead reduces F by 50%, reflecting a 50% discard mortality rate**

Line 280: missing period.

**Corrected.**

Line 283: that is a very nice model feature.

**We are glad the reviewer finds this feature helpful, many thanks to the folks at FishBase and Thorson's prior work with FishLife for making it possible.**

Line 301: Is omitting advection only a significant limitation for drifting species like plankton or jellyfish, or is it also a limitation for large pelagics that concentrate at SST fronts?

**Good question. A strength of the CTMC approach is that, given sufficient data, one can separate out the components of movement into advection, taxis, and diffusion. Having this knowledge is certainly interesting, but from a simulation perspective is less critical. One could simulate the dynamics of a large pelagic concentrating at an SST front using simply the taxis matrix in `marlin`; while the SST fronts themselves are clearly a partial product of advective forces, so long as we know where they are, we can just pass their location to marlin and the fish will go there. What would be missing there might be some timing dynamics whereby a fish might take longer to get to ideal habitat when fighting against a current than with it. The biggest limitation of omitting the advection matrix at this point is the potential to overlook cases where a current is strong enough to essentially serve as a barrier preventing an organism from reaching preferred habitat that it otherwise would go to under the taxis + diffusion model. We have clarified this in the text of the discussion, and pointed out the value of research into the importance of advective forces as well. **

Line 473: Typo: “Do” -> “Due”.
**Corrected.**

##R2

FaF-23-Apr-OA-149 Review 02 Aug 2023
Summary
This study presents a bio-economic modeling tool aimed to provide a tractable, efficient, tactical management simulation framework for marine fisheries. It seeks to complement earlier, similar, models which are highly complex, bespoke, or overly-generalized. Case study applications to a coral reef and pelagic ecosystem is presented. The study conclusions does not evaluate spatial management paradigms per se, but rather illustrate the value of “sufficiently realistic” biodynamic models for policy evaluation.

This model and paper provide the following novel contributions:
1) the modeling framework allows users to test either eliminating or displacing fishing effort given the placement of an MPA; ignoring these alternative possibilities was the cause of much scandal and scientific misinformation in recent MPA analyses,

**It certainly has been, we are glad the reviewer finds this feature useful**

2) a wise avoidance of the calculation of equilibrium-based quantities for spatio-temporal models, precluding scientists’ conclusion that certain policies do or do not meet management targets (as these are not definable under such conditions), and

**Though dynamic and spatially explicit reference points are certainly an area of interesting research going forward.**

3) an explicit manner of investigating MPA effectiveness in the context of range shifts, which are important for climate-enhanced fishery management.

I recommend publication with minor revision.

Major Comments
This is a well-written paper, and I appreciated that the methods are clear and high-level, with minimal equations. There are two central things that need to be addressed in the document, which would bolster and clarify how marlin fits into the spatial modelling world, and how we can interpret the results. I believe these can be readdressed through a fairly simple simulation exercise and the addition of a few paragraphs to the methods & discussion.

Most importantly, the reader wants to know whether marlin is providing insights consistent with our understanding of how population and fishery dynamics behave in space, and whether the reader has enough information to make it worth testing the model on their fishery. If the aim is for this model to be used by global communities with little data or information about their fishery, please discuss if/how the model can be simplified to provide insights useful in the absence of key processes (i.e., selectivity).

**We have added a "General Recommendations for using `marlin`" section to the discussion, in which we have emphasized this point amount others.**

How dependent is the simulation accuracy on full confidence in the inputs? I am not requesting a full misspecification analysis. Rather, please illustrate or discuss how the results of the coral reef case study, for example, might change if we are incorrect that Fleet Two has dome-shaped selectivity. We know this to be problematic in spatial models. How do your findings compare to earlier examinations of mis- specified selectivity in spatial context, e.g. works by Aaron Berger and Lynn Waterhouse?

**We have added a "General Recommendations for using `marlin`" section to the discussion, in which we have emphasized this point amount others.**

XX I still need to integrate some more of these XX s

Bosley, Katelyn M., Amy M. Schueller, Daniel R. Goethel, Dana H. Hanselman, Kari H. Fenske, Aaron M. Berger, Jonathan J. Deroba, and Brian J. Langseth. “Finding the Perfect Mismatch: Evaluating Misspecification of Population Structure within Spatially Explicit Integrated Population Models.” Fish and Fisheries 23, no. 2 (2022): 294–315. https://doi.org/10.1111/faf.12616.

Berger, Aaron M., Daniel R. Goethel, Patrick D. Lynch, Terrance Quinn, Sophie Mormede, Jeremy McKenzie, and Alistair Dunn. “Space Oddity: The Mission for Spatial Integration.” Canadian Journal of Fisheries and Aquatic Sciences 74, no. 11 (November 2017): 1698–1716. https://doi.org/10.1139/cjfas-2017-0150.

Berger, Aaron M., Michael L. Jones, Yingming Zhao, and James R. Bence. “Accounting for Spatial Population Structure at Scales Relevant to Life History Improves Stock Assessment: The Case for Lake Erie Walleye Sander Vitreus.” Fisheries Research 115–116 (March 1, 2012): 44–59. https://doi.org/10.1016/j.fishres.2011.11.006.

Berger, Aaron M, Jonathan J Deroba, Katelyn M Bosley, Daniel R Goethel, Brian J Langseth, Amy M Schueller, and Dana H Hanselman. “Incoherent Dimensionality in Fisheries Management: Consequences of Misaligned Stock Assessment and Population Boundaries.” ICES Journal of Marine Science 78, no. 1 (February 1, 2021): 155–71. https://doi.org/10.1093/icesjms/fsaa203.

Waterhouse, Lynn, David B. Sampson, Mark Maunder, and Brice X. Semmens. “Using Areas-as-Fleets Selectivity to Model Spatial Fishing: Asymptotic Curves Are Unlikely under Equilibrium Conditions.” Fisheries Research, SI: Selectivity, 158 (October 1, 2014): 15–25. https://doi.org/10.1016/j.fishres.2014.01.009.

- Add discussion to the effect of "marlin has a lot of options. A recommended use case would be to try and narrow down things that are "known", and then focus sensitivity runs on outcomes that are uncertain and likely to be important. For example, testing the sensitivity of outcomes to misspecified selectivity curves

**We have added a "General Recommendations for using `marlin`" section to the discussion, in which we have emphasized this point amount others.**

The value of this contribution would be stronger if you directly compared the findings from your case studies with similar works. For example, there have been several studies (empirical and simulation- based) examining MPAs and coral reef ecosystems. Several have found confounding effects of MPA closures, often caused by inter-species interactions, which I recognize are not modeled here. Does marlin provide similar insights on less information (i.e. without predator dynamics)?

**case study section literature time here. Really struggling with this one. Thoughts on how to address?**

Of secondary importance, I am not sure I agree that marlin follows the MICE paradigm. Much of the model specification (I’m thinking movement especially) is highly granular even for well-studied fisheries.

**We have removed the reference to MICE models per say..**

It is common for tactical applications to model irregular polygons of distinct biological and/or political attributes, connected by a user-input matrix of movement values. Confirm whether the current setup requires individual patches to share biological information and the harvest strategies among them. Discuss how spatial heterogeneity in population demography can impact model results.

**This is a very good point; spatio-temporal variation in population demography (and regulation) can be extremely important and is not captured here. `marlin` assumes that while habitat can vary in space and time, population processes are constant within a species. That being said,  users could treat different morphs of a population as separate species with different habitats. For example, if an invertebrate species grows at a different rate in warmer coastal waters than cooler deeper waters, these could simply be modeled as two separate species with different growth parameters. However, the model would not then allow for interbreeding of those morphs..xx** Cite @waterhouse2014 here, need to find where to cite and add to paper XX

see "finding the perfect mismatch bosley paper" for discussions around frequency of spatial complexity paired with frequency of ignoring it

Minor Comments
I recognize the authors didn’t want to detail every aspect of the model specifications, but these are things the reader might want to know from the outset:

L106 “10 x 10 patch system” – units? what is a “patch” here? (related to point above about what info is shared within a patch)

**Users on standard computers should be able to simulate one age-structured population distributed across spatial surface represented by a ten by ten grid of cells over 20 years in fractions of a second**

L157 you mention size limits as a potential harvest control – but don’t say that size is also represented in the pop dynamics until much later

**We have now clarified in the opening description that the model also tracks lengths at age**.

Model Summary – somewhere early in here, confirm that there is no estimation/conditioning capacity within marlin (to discourage folks from thinking they might be able to run assessments within it, for example). Add a sentence or two near the beginning describing what the user must provide to the model, and state that meta-analyses like FishLife will be consulted in their absence.

**We have added a few sentences to the model summary section describing the minimum inputs needed, and making the important point that there is no estimation feature built into marlin**

L232 Related to above: I see the description of the tuning settings. When total SSB is specified in the context of MPA, how is Q adjusted? Does this step interact with the option to either displace or turn off effort outside of an MPA?

**We have clarified in the text that "`marlin` allows users to specify a target unfinished spawning stock biomass ($SSB0$), which will be achieved by tuning the total unfished recruitment (R0), given the remaining life history parameters and independent of any characteristics of the fishing fleets.". We have also clarified in the "tuning fishing fleet parameters" section that all tuning is done conditional on the underlying population dynamics.**

L259 what sort of time scale is needed for this approximation to be true?

**We have clarified in the text that "The timeline required for the fleets to reach an equilibrium distribution will vary and depend on factors such as the population dynamics of the species in question (e.g. movement, habitat heterogeneity, growth rates, density dependence) and the number and degree of competition across fleets. Users should explore different simulation times to ensure that any results they wish to use are not simply a reflection of the fleet dynamics fluctuating on their way to an equilibrium condition. "**

L310 “that the model is running under” – rephrase

**Clarified**.

L363 be consistent about capitalization of SSB

**Fixed**

L447 too many quotes in this section

**Removed**

L493 quarterly timestep? Only one year (four seasons total)?

**We have clarified that the simulation runs for 100 years in time steps of quarter years (i.e. four seasons per year)**

Figure 3 I’d prefer this to have the same colorscale as Figure 1; do a panel for each condition, don’t overlay them.

**We have changed Figure 3 to reflect the same format as Figure 1, and agree it is much improved, thank you!**

L534 Do they actually mask them, or do interesting dynamics underlie these findings? Add a sentence here what users should do with this info (i.e. be sure to look at the tradeoffs one by one – is there functionality in marlin to do so, such as automated plots or summary tables?)

**Good point, we clarify that "These macro-level changes in total yield and conservation are the result of interesting internal dynamics that `marlin` allows users to explore through output tables at the level of individual species and fleets." We do not specify specific functions within the marlin package here as the details of those may evolve over time.**

FIG 4. The results state three MPA strategies were used: spawning ground, spawning aggregation, and target fishing, but I only see two here.

**We have fixed this typo (we had accidentally refered to the same policy by two different names).**

FIG 4C This figure is confusing; I’m not clear what the skinny short lines are. The caption doesn’t clarify which color of label corresponds to which % change in yield; instead just use a colored point for these values.

**We have modified Figure 4C and we hope the reviewer finds it clearer now. Much of the visual clutter of Fig 4C was due to the long tail to the right of high SSB increases and catch decreases that occured once MPA size approaches 100%. To help illustrate the more interesting points of the results, we have limited the figure to MPA sizes <= 65%, and removed the text labels for MPA sie and replaced them with colored points to indicate MPA size along each of the MPA design strategy curves.**

### 2023-09-28 03_marlin_paper.qmd
getting somewhere

@@
 params:
-  run_name: ["v1.0"]
+  run_name: ["v1.2"]
 header-includes:
@@
   - \usepackage{bm}
+editor: 
+  markdown: 
+    wrap: 80
 ---
@@
 
-Title: Simulating Benefits, Costs, and Tradeoffs of Spatial Management
-in Marine Social-Ecological Systems
+Title: Simulating Benefits, Costs, and Tradeoffs of Spatial Management in Marine
+Social-Ecological Systems
 
@@
 Affiliations: ^1^ School of Aquatic and Fishery Sciences, University of
-Washington,1122 NE Boat St Box 355020,Seattle, WA, USA ^3^ Marine
-Science Institute, University of California, Santa Barbara, CA, USA ^4^
-Bren School of Environmental Science & Management, University of
-California, Santa Barbara, CA, USA ^5^ Environmental Markets Lab,
-University of California, Santa Barbara, CA, USA ^6^ Habitat and
-Ecological Processes Research Program, Alaska Fisheries Science Center,
-NOAA, NMFS, NOAA, Seattle, WA, USA
+Washington,1122 NE Boat St Box 355020,Seattle, WA, USA ^3^ Marine Science
+Institute, University of California, Santa Barbara, CA, USA ^4^ Bren School of
+Environmental Science & Management, University of California, Santa Barbara, CA,
+USA ^5^ Environmental Markets Lab, University of California, Santa Barbara, CA,
+USA ^6^ Habitat and Ecological Processes Research Program, Alaska Fisheries
+Science Center, NOAA, NMFS, NOAA, Seattle, WA, USA
 
@@
 
-Present Address: \^2 Inter-American Tropical Tuna Commission,8901 La
-Jolla Shores Drive, La Jolla, CA 92037, USA
-
-Author Contributions: Methods and code written by DO and JT. Results
-produced by DO. Data collected by EB and LT. All authors contributed to
-the conceptualization and writing of the manuscript.
-
-Conflict of Interest Statement: The authors declare no conflicts of
-interest.
+Present Address: \^2 Inter-American Tropical Tuna Commission,8901 La Jolla
+Shores Drive, La Jolla, CA 92037, USA
+
+Author Contributions: Methods and code written by DO and JT. Results produced by
+DO. Data collected by EB and LT. All authors contributed to the
+conceptualization and writing of the manuscript.
+
+Conflict of Interest Statement: The authors declare no conflicts of interest.
 
@@
 
-Designing effective spatial management strategies is challenging because
-marine ecosystems are highly dynamic and opaque, and extractive entities
-such as fishing fleets respond endogenously to ecosystem changes in ways
-that depend upon ecological and policy context. We present a tool,
-`marlin`, that can be used to efficiently simulate the bio-economic
-dynamics of marine systems in support of both management and research.
-We demonstrate `marlin’s` capabilities by focusing on two case studies
-on the conservation and food production impacts of marine protected
-areas (MPAs): a coastal coral reef and a pelagic tuna fishery. In the
-coastal coral reef example, we explore how heterogeneity in species
-distributions and fleet preferences can affect distributional outcomes
-of MPAs. In the pelagic case study, we show how our model can be used to
-assess the climate resilience of different MPA design strategies, as
-well as the climate sensitivity of different fishing fleets. This paper
-demonstrates how intermediate complexity simulation of coupled
-bio-economic dynamics can help communities predict and potentially
-manage trade-offs between conservation, fisheries yields, and
-distributional outcomes of management policies affected by spatial
-bio-economic dynamics.
+Designing effective spatial management strategies is challenging because marine
+ecosystems are highly dynamic and opaque, and extractive entities such as
+fishing fleets respond endogenously to ecosystem changes in ways that depend
+upon ecological and policy context. We present a tool, `marlin`, that can be
+used to efficiently simulate the bio-economic dynamics of marine systems in
+support of both management and research. We demonstrate `marlin’s` capabilities
+by focusing on two case studies on the conservation and food production impacts
+of marine protected areas (MPAs): a coastal coral reef and a pelagic tuna
+fishery. In the coastal coral reef example, we explore how heterogeneity in
+species distributions and fleet preferences can affect distributional outcomes
+of MPAs. In the pelagic case study, we show how our model can be used to assess
+the climate resilience of different MPA design strategies, as well as the
+climate sensitivity of different fishing fleets. This paper demonstrates how
+intermediate complexity simulation of coupled bio-economic dynamics can help
+communities predict and potentially manage trade-offs between conservation,
+fisheries yields, and distributional outcomes of management policies affected by
+spatial bio-economic dynamics.
 
@@
 
+## Need to cite..
+
+@bosley2022
+
+@berger2017
+
 ## Introduction
 
-Communities around the world are increasingly looking to spatial
-strategies for managing marine ecosystems . For example, the "30x30"
-movement calls for implementation of marine protected areas (MPAs), a
-form of spatial management, across 30% of the world's oceans by the year
-2030 [@grorud-colvert2021 and references therein]. However , designing
-spatial management strategies to achieve desired objectives -- which may
-include recovery and resilience of overfished species, increased food
-production and economic well being, and equitable distribution of
-benefits - is not straightforward. Different species have different
-resilience to fishing, distributions in space and time, and value to
-fishing fleets; fishing fleets themselves can have varying ranges of
-incentives and reactions to policy structures. These dynamics pose
-challenges even for single-species assessment and management strategies
-, which are only amplified when we consider management policies
-specifically designed around multiple species, fleets, and spatial
-features in the oceans [@field2006a].
-
-To illustrate, policies such as MPAs designed for both conservation and
-food production must consider factors such as the optimal size and
-placement of a protected area network given a wide array of life
-history, species distributions, exploitation levels, fleet dynamics, and
-policy-dependent behavior [@reimer2017], all of which may change over
-time, particularly given the impacts of climate change. Efforts to
-effectively design spatial management strategies such as MPAs are
-further constrained by a lack of empirical evidence describing the size
-[@ovando2021a] and time required [@nickols2019] for MPAs to produce
-substantial effects across a range of social-ecological settings
+Communities around the world are increasingly looking to spatial strategies for
+managing marine ecosystems . For example, the "30x30" movement calls for
+implementation of marine protected areas (MPAs), a form of spatial management,
+across 30% of the world's oceans by the year 2030 [@grorud-colvert2021 and
+references therein]. However , designing spatial management strategies to
+achieve desired objectives -- which may include recovery and resilience of
+overfished species, increased food production and economic well being, and
+equitable distribution of benefits - is not straightforward. Different species
+have different resilience to fishing, distributions in space and time, and value
+to fishing fleets; fishing fleets themselves can have varying ranges of
+incentives and reactions to policy structures. These dynamics pose challenges
+even for single-species assessment and management strategies , which are only
+amplified when we consider management policies specifically designed around
+multiple species, fleets, and spatial features in the oceans [@field2006a].
+
+To illustrate, policies such as MPAs designed for both conservation and food
+production must consider factors such as the optimal size and placement of a
+protected area network given a wide array of life history, species
+distributions, exploitation levels, fleet dynamics, and policy-dependent
+behavior [@reimer2017], all of which may change over time, particularly given
+the impacts of climate change. Efforts to effectively design spatial management
+strategies such as MPAs are further constrained by a lack of empirical evidence
+describing the size [@ovando2021a] and time required [@nickols2019] for MPAs to
+produce substantial effects across a range of social-ecological settings
 [@mcclanahan2021].
 
-Many models have been developed to support the theory and design of
-spatial management strategies (see @fulton2015 for a review of existing
-methods). However, these models have tended to be either highly complex
-and tactical models designed for use in a specific location, or to be
-extremely stylized representations intended to provide theoretical
-insights with less relevance for specific applications. This lack of
-accessible models suited for more tactical application presents a
-challenge to stakeholders charged with marine spatial planning
-exercises. It also presents a barrier to the scientific community,
-wherein comparisons across spatial management simulation studies are
-clouded by discrepancies in underlying model structures beyond
-differences in the core phenomenon in question.
-
-To help address this challenge, we present a novel bio-economic modeling
-tool called [`marlin`](https://danovando.github.io/marlin/) that allows
-for efficient simulation of spatio-temporal biological and economic
-dynamics. `marlin` allows users to simulate the impacts of marine
-management policies across a range of species targeted by various fleets
-across heterogeneous and dynamic seascapes. This model can help users
-seek Pareto-optimal solutions that produce win-wins or minimize
-trade-offs across multiple management objectives, such as food
+Many models have been developed to support the theory and design of spatial
+management strategies (see @fulton2015 for a review of existing methods).
+However, these models have tended to be either highly complex and tactical
+models designed for use in a specific location, or to be extremely stylized
+representations intended to provide theoretical insights with less relevance for
+specific applications. This lack of accessible models suited for more tactical
+application presents a challenge to stakeholders charged with marine spatial
+planning exercises. It also presents a barrier to the scientific community,
+wherein comparisons across spatial management simulation studies are clouded by
+discrepancies in underlying model structures beyond differences in the core
+phenomenon in question.
+
+To help address this challenge, we present a novel bio-economic modeling tool
+called [`marlin`](https://danovando.github.io/marlin/) that allows for efficient
+simulation of spatio-temporal biological and economic dynamics. `marlin` allows
+users to simulate the impacts of marine management policies across a range of
+species targeted by various fleets across heterogeneous and dynamic seascapes.
+This model can help users seek Pareto-optimal solutions that produce win-wins or
+minimize trade-offs across multiple management objectives, such as food
 production and conservation [@lester2013].
 
-Here we present the core methods and functionality of `marlin`, and
-demonstrate its use in two case-study applications; a coral reef system
-and a pelagic system. Our results show how interacting ecological,
-economic, and design attributes can affect the degree of benefits,
-costs, and trade offs between conservation and food production outcomes
-of MPAs. More broadly, this paper both demonstrates the functionality of
-our model and the critical need for considering sufficiently realistic
-coupled ecological and economic dynamics in policy evaluations.
+Here we present the core methods and functionality of `marlin`, and demonstrate
+its use in two case-study applications; a coral reef system and a pelagic
+system. Our results show how interacting ecological, economic, and design
+attributes can affect the degree of benefits, costs, and trade offs between
+conservation and food production outcomes of MPAs. More broadly, this paper both
+demonstrates the functionality of our model and the critical need for
+considering sufficiently realistic coupled ecological and economic dynamics in
+policy evaluations.
 
@@
 `marlin` follows a "models of intermediate complexity" philosophy
-[@plagányi2014] targeted specifically towards capturing heterogeneous
-habitats, movement dynamics, and the simultaneous impacts of fishing
-fleets across multiple species (frequently called "technical
-interactions"). In its general form, `marlin` simulates the behavior of
-populations of spatially explicit biologically independent
-age-structured animal populations affected by one or more fishing fleets
-in time steps and a spatial resolution specified by the user.
+[@plagányi2014] targeted specifically towards capturing heterogeneous habitats,
+movement dynamics, and the simultaneous impacts of fishing fleets across
+multiple species (frequently called "technical interactions"). In its general
+form, `marlin` simulates the behavior of populations of spatially explicit
+biologically independent age-structured animal populations affected by one or
+more fishing fleets in time steps and a spatial resolution specified by the
+user.
 
@@
 [`marlin`](https://danovando.github.io/marlin/), is written in R
-[@rcoreteam2021] to facilitate use, while the underlying population
-model is written in C++, integrated through the `Rcpp` package
-[@eddelbuettel2018], for speed. Users on standard computers should be
-able to simulate one age-structured population distributed across a 10
-by 10 patch system for 20 years in fractions of a second.
+[@rcoreteam2021] to facilitate use, while the underlying population model is
+written in C++, integrated through the `Rcpp` package [@eddelbuettel2018], for
+speed. Users on standard computers should be able to simulate one age-structured
+population distributed across two-dimensional spatial surface represented by a
+ten by ten grid of cells over 20 years in fractions of a second.
 
 While `marlin` does not simulate species interactions, fishing fleets in
-`marlin` are capable of targeting and affecting multiple species
-simultaneously. This allows the model to simulate processes such as
-fisheries bycatch or effort displacement in a way that accounts for how
-fleet behavior might affect multiple species in the system. The ability
-to track the impacts of fishing fleets across multiple species
-simultaneously is particularly important as very few fishing fleets are
-truly single species.
-
-We define a few commonly used terms here. *Yields* refers to the volume
-of catch from fishing activity. *Spawning stock biomass* refers to the
-total biomass of reproductively mature fish in the water, a function of
-numbers, weight, fecundity, and sexual maturity at age. We measure the
-size of the population by th ratio of the *spawning stock biomass* (SSB)
-in a given time step relative to *unfished spawning stock biomass*
-(SSB0), SSB/SSB0. So, a SSB/SSB0 value of 1 means that the population is
-unfished, a SSB/SSB0 value of 0 means the population is extinct.
-
-As a demonstration of the use of this modeling tool, we explore two
-general situations
-
-1.  Trade-offs and distributional outcomes for food production and
-    multi-species biomass in a coastal coral-reef fishery
-2.  Implications of climate-driven range shifts for MPA design in a
-    pelagic fishery
-
-We chose these two examples to illustrate the use of `marlin` in
-contrasting systems in which spatial management strategies such as MPAs
-are increasingly considered.
-
-Below we provide a summary of the `marlin` model, as well as details of
-the case studies. We note that `marlin` is a complex software package
-with many options to control detailed behaviors. Given that these
-specific features will evolve over time, we do not describe in detail
-every option of the model here. Rather, we focus on the equations and
-methods describing the core foundations of our tool for efficiently
-simulating bio-economic dynamics in aquatic systems.
+`marlin` are capable of targeting and affecting multiple species simultaneously.
+This allows the model to simulate processes such as fisheries bycatch or effort
+displacement in a way that accounts for how fleet behavior might affect multiple
+species in the system. The ability to track the impacts of fishing fleets across
+multiple species simultaneously is particularly important as very few fishing
+fleets are truly single species.
+
+We define a few commonly used terms here. *Yields* refers to the volume of catch
+from fishing activity. *Spawning stock biomass* refers to the total biomass of
+reproductively mature fish in the water, a function of numbers, weight,
+fecundity, and sexual maturity at age. We measure the size of the population by
+the ratio of the *spawning stock biomass* (SSB) in a given time step relative to
+*unfished spawning stock biomass* (SSB0), SSB/SSB0. So, a SSB/SSB0 value of 1
+means that the population is unfished, a SSB/SSB0 value of 0 means the
+population is extinct.
+
+As a demonstration of the use of this modeling tool, we explore two general
+situations
+
+1.  Trade-offs and distributional outcomes for food production and multi-species
+    biomass in a coastal coral-reef fishery
+2.  Implications of climate-driven range shifts for MPA design in a pelagic
+    fishery
+
+We chose these two examples to illustrate the use of `marlin` in contrasting
+systems in which spatial management strategies such as MPAs are increasingly
+considered.
+
+Below we provide a summary of the `marlin` model, as well as details of the case
+studies. We note that `marlin` is a complex software package with many options
+to control detailed behaviors. Given that these specific features will evolve
+over time, we do not describe in detail every option of the model here. Rather,
+we focus on the equations and methods describing the core foundations of our
+tool for efficiently simulating bio-economic dynamics in aquatic systems.
 
@@
 `marlin` simulates the dynamics of one or more species, currently best
-representing fish-like species, using age-structured population
-dynamics. Each time step, species move throughout the simulated seascape
-using both diffusion and "taxis" (active movement towards preferred
-habitat), experience natural and potentially fishing mortality, and if
-in a spawning season spawn using one of the possible forms of density
-dependence implemented in the model, with the potential to include
-auto-correlated stochastic deviations in the amount of offspring
-produced, generally called "recruitment deviates".
-
-These species can be caught by fishing fleets. A fishing fleet is
-defined in `marlin` by a set of fishers that have the same fishing
-skill, prices, and selectivity for individual species (each of which we
-denote as a *metier*). For example, both a longline and purse seine
-fleet may capture bigeye (*Thunnus obesus*) and skipjack tunas, but the
-longline fleet can be made much more likely to catch larger bigeye than
-skipjack, and *vice versa*. Each fishing fleet distributes its fishing
-effort in space according to a specified spatial allocation rule (see
-*Spatial Allocation of Effort* ), for example in proportion to
-profit-per-unit-effort, conditional on management policies in places
-such as quotas and/or the presence of any spatial restrictions such as
-MPAs.
-
-Unconstrained by management, the total amount of effort exerted by each
-fleet can follow one of two dynamics: open access or constant effort
-(see *Calculating Total Effort*). Under constant effort, the total
-amount of effort of each fleet is fixed over time, with the possible
-exception of attrition due to MPA placement. Under open access, the
-total amount of effort in a given time step is a function of the
-profitability of the fleet in the previous time step, evolving until a
-bionomic equilibrium of zero total profits is reached. Profitability is
-a function of the volume and price of each species caught, as well as
-the cost of the total amount of fishing effort per fleet and the travel
-costs as a function of distance from a port (see *Fleet Dynamics*).
-
-The dynamics of each fleet can be modified by management in a variety of
-ways. Managers can impose size limits for individual species within each
-fleet. They can also impose total catch quotas for each species in each
-time step. When quotas are activated, if the total catch across all
-fleets for a given species under the model's effort dynamics would
-exceed the allowable quota for that species, the total effort for each
-fleet is decreased proportional to its contribution to the total catch
-until the quota is satisfied. As an alternative to catch quotas,
-managers can set an effort cap per fleet, which prevents effort from
-exceeding a given amount under open access dynamics, though effort may
-be reduced below this cap if required by the profit equation (i.e. the
-fleet can choose to fish less than the quota or effort cap).
-
-`marlin` also allows users to specify closed fishing seasons for one or
-more species in the system. When all species in a simulation share the
-same closed seasons, no fishing occurs in those seasons. When not all
-species share the same closed seasons, fishing still occurs in the
-closed seasons, but the fleet avoids catching the species for which
-fishing is closed in a given season, assuming no discard mortality. When
-closed seasons are in effect, the amount of effort in the next open
-season is a function of the effort and profits in the last open fishing
-season. Lastly, managers can specify locations of no-take MPAs, which
-can change in size and location if needed. When MPAs are implemented,
-fishing effort that used to operate inside the MPAs can either leave the
-fishery, or be redistributed outside the MPA (the default behavior).
-
-Users can use the model to explore general properties of spatially
-explicit social-ecological systems, or can parameterize it to explore
-key uncertainties of more tactical applications. As an example of the
-later case, we provide two case studies here designed to reflect the
-kinds of questions stakeholders might face if designing an MPA network
-for a coastal coral reef or a pelagic system.
+representing fish-like species, using age-structured population dynamics. Ages
+are then converted to lengths using the von Bertalanffy growth equation with
+log-normally distributed variation in the length at age. Each time step, species
+move throughout the simulated seascape using both diffusion and "taxis" (active
+movement towards preferred habitat), experience natural and potentially fishing
+mortality, and if in a spawning season spawn using one of the possible forms of
+density dependence implemented in the model, with the potential to include
+auto-correlated stochastic deviations in the amount of offspring produced,
+generally called "recruitment deviates".
+
+These species can be caught by fishing fleets. A fishing fleet is defined in
+`marlin` by a set of fishers that have the same fishing skill, prices, and
+selectivity for individual species (each of which we denote as a *metier*). For
+example, both a longline and purse seine fleet may capture bigeye (*Thunnus
+obesus*) and skipjack tunas, but the longline fleet can be made much more likely
+to catch larger bigeye than skipjack, and *vice versa*. Each fishing fleet
+distributes its fishing effort in space according to a specified spatial
+allocation rule (see *Spatial Allocation of Effort* ), for example in proportion
+to profit-per-unit-effort, conditional on management policies in places such as
+quotas and/or the presence of any spatial restrictions such as MPAs.
+
+Unconstrained by management, the total amount of effort exerted by each fleet
+can follow one of two dynamics: open access or constant effort (see *Calculating
+Total Effort*). Under constant effort, the total amount of effort of each fleet
+is fixed over time, with the possible exception of attrition due to MPA
+placement. Under open access, the total amount of effort in a given time step is
+a function of the profitability of the fleet in the previous time step, evolving
+until a bionomic equilibrium of zero total profits is reached. Profitability is
+a function of the volume and price of each species caught, as well as the cost
+of the total amount of fishing effort per fleet and the travel costs as a
+function of distance from a port (see *Fleet Dynamics*).
+
+The dynamics of each fleet can be modified by management in a variety of ways.
+Managers can impose size limits for individual species within each fleet. They
+can also impose total catch quotas for each species in each time step. When
+quotas are activated, if the total catch across all fleets for a given species
+under the model's effort dynamics would exceed the allowable quota for that
+species, the total effort for each fleet is decreased proportional to its
+contribution to the total catch until the quota is satisfied. As an alternative
+to catch quotas, managers can set an effort cap per fleet, which prevents effort
+from exceeding a given amount under open access dynamics, though effort may be
+reduced below this cap if required by the profit equation (i.e. the fleet can
+choose to fish less than the quota or effort cap).
+
+`marlin` also allows users to specify closed fishing seasons for one or more
+species in the system. When all species in a simulation share the same closed
+seasons, no fishing occurs in those seasons. When not all species share the same
+closed seasons, fishing still occurs in the closed seasons, but the fleet avoids
+catching the species for which fishing is closed in a given season, assuming no
+discard mortality. When closed seasons are in effect, the amount of effort in
+the next open season is a function of the effort and profits in the last open
+fishing season. Lastly, managers can specify locations of no-take MPAs, which
+can change in size and location if needed. When MPAs are implemented, fishing
+effort that used to operate inside the MPAs can either leave the fishery, or be
+redistributed outside the MPA (the default behavior).
+
+Users can use the model to explore general properties of spatially explicit
+social-ecological systems, or can parameterize it to explore key uncertainties
+of more tactical applications. It is very important note that the parameters of
+this model cannot be fit to data directly within the `marlin` package; users
+must set model parameters themselves based on externally available data and
+their best judgement. On the biological side, at a minimum users must for each
+species they wish to simulate supply the common or scientific name of the
+species in question, a measure of the level of fishing intensity the species is
+experiencing at the start of the simulation, and the diffusion rates for adult
+and larvae. Given a common or scientific name, the model will then input default
+life history parameters based on FishLife [internet connection required,
+@thorson2020], though users are encouraged to check these values and supply
+their own life history parameter values when possible. On the fleet side, users
+must for each fleet they wish to simulate at a minimum specify the selectivity
+curves for each species caught by the fleet, as well relative price per unit
+weights greater than 0 for any target species. Species caught by but not
+targeted by the fleet can be represented by prices of 0 (not targeted) or below
+0 (actively avoided). We provide two case studies here designed to show users
+how the `marlin` model can be used to address the kinds of questions
+stakeholders might face if designing an MPA network for a coastal coral reef or
+a pelagic system.
 
@@
 
-Each fishing fleet (*f*) generates catches or yield (*Y*), revenues
-(*R*), and costs (*C*) from fishing individual species (*s*) that it
-targets in a given time step (*t*) and patch (*p*). The fleet then makes
-decisions around fishing locations and intensity, subject to regulatory
-constraints, based on the total profitability across all species in
-space and time.
-
-Revenues for each fleet in a time step are a function of the total
-amount of each species caught and its price ($\Pi$). The amount caught
-is a function of the selectivity at age of each species for each fleet
-($\alpha$), the fishing efficiency of the fleet for that that species
-(also called "catchability", *q*), the amount of fishing effort of the
-fleet in question in that patch in a given time step (*E*), and the
-total instantaneous fishing mortality (*u*) at age (*a*) (including all
-other fleets) for that particular species in that patch and time step.
+Each fishing fleet (*f*) generates catches or yield (*Y*), revenues (*R*), and
+costs (*C*) from fishing individual species (*s*) that it targets in a given
+time step (*t*) and patch (*p*). The fleet then makes decisions around fishing
+locations and intensity, subject to regulatory constraints, based on the total
+profitability across all species in space and time.
+
+Revenues for each fleet in a time step are a function of the total amount of
+each species caught and its price ($\Pi$). The amount caught is a function of
+the selectivity at age of each species for each fleet ($\alpha$), the fishing
+efficiency of the fleet for that that species (also called "catchability", *q*),
+the amount of fishing effort of the fleet in question in that patch in a given
+time step (*E*), and the total instantaneous fishing mortality (*u*) at age
+(*a*) (including all other fleets) for that particular species in that patch and
+time step.
 
@@
 [@baranov1918], which appropriately accounts for the total instantaneous
-mortality (both fishing and natural. *z*) and divides the total amount
-of the population biomass (*b*) killed between natural (*m*) and fishing
-(*u*) sources, with the amount of biomass killed through fishing called
-"yield".
+mortality (both fishing and natural. *z*) and divides the total amount of the
+population biomass (*b*) killed between natural (*m*) and fishing (*u*) sources,
+with the amount of biomass killed through fishing called "yield".
 
@@
 
-Selectivity at age $\alpha$ is modeled through either a logistic form or
-a dome-shaped form. The logistic form is based on the lengths *l* at
-which 50% of individuals are selected by the fishing gear, $l^{sel}$,
-and $\delta$ which is the difference between the length at 50%
-selectivity and the length at 95% selectivity.
+Selectivity at age $\alpha$ is modeled through either a logistic form or a
+dome-shaped form. The logistic form is based on the lengths *l* at which 50% of
+individuals are selected by the fishing gear, $l^{sel}$, and $\delta$ which is
+the difference between the length at 50% selectivity and the length at 95%
+selectivity.
 
@@
 
-We approximate the dome-shaped form as a normal distribution with mean
-$l^{sel}$ and standard deviation $\sigma$. The normal density function
-is re-scaled such that the selectivity is 1 at $l^{sel}$.
+We approximate the dome-shaped form as a normal distribution with mean $l^{sel}$
+and standard deviation $\sigma$. The normal density function is re-scaled such
+that the selectivity is 1 at $l^{sel}$.
 
@@
 
-This means that each species experiences a total mortality at age in a
-time step in a patch, individual fractions of which are portioned off as
-catches and subsequently revenues for each fleet. Given revenues *R*, we
-then calculate profits $\phi$ based on revenues and costs. Costs *C* are
-calculated as a function of base costs per unit effort ($\gamma$) as
-well as potential additional costs per unit effort of fishing in
-particular patches ($\eta$) for fleet *f*. $\beta$ allows for the cost
-per unit effort effort to scale non-linearly. Travel costs ($\eta$) are
-calculated based on the Euclidean distance of each patch to the nearest
-port of a given fleet, and users can specify any number from zero to the
-number of patches of port locations. When no ports are specified, travel
-costs are zero.
+This means that each species experiences a total mortality at age in a time step
+in a patch, individual fractions of which are portioned off as catches and
+subsequently revenues for each fleet. Given revenues *R*, we then calculate
+profits $\phi$ based on revenues and costs. Costs *C* are calculated as a
+function of base costs per unit effort ($\gamma$) as well as potential
+additional costs per unit effort of fishing in particular patches ($\eta$) for
+fleet *f*. $\beta$ allows for the cost per unit effort effort to scale
+non-linearly. Travel costs ($\eta$) are calculated based on the Euclidean
+distance of each patch to the nearest port of a given fleet, and users can
+specify any number from zero to the number of patches of port locations. When no
+ports are specified, travel costs are zero.
 
@@
 
-`marlin` allows for two general modes of effort. The simplest is
-"constant effort", in which the total effort of each fishing fleet
-remains constant over time.
-
-The more complex option is "open access". Under open access, the total
-amount of effort in a given time step for fleet *f* is a function of the
-profitability of that fleet in the proceeding time steps in which
-fishing was open, where $\theta$ controls the responsiveness of fleet
-*f* to the log of the ratio of total revenues *R* to total costs *C*,
-and approximates the proportional change in effort in response to a one
-unit changes in the log revenue to cost ratio.
+`marlin` allows for two general modes of effort. The simplest is "constant
+effort", in which the total effort of each fishing fleet remains constant over
+time.
+
+The more complex option is "open access". Under open access, the total amount of
+effort in a given time step for fleet *f* is a function of the profitability of
+that fleet in the proceeding time steps in which fishing was open, where
+$\theta$ controls the responsiveness of fleet *f* to the log of the ratio of
+total revenues *R* to total costs *C*, and approximates the proportional change
+in effort in response to a one unit changes in the log revenue to cost ratio.
 
@@
 
-This is essentially a Gompertz model for fishing effort, that has been
-used for other theoretical studies of fishing effort dynamics
-[@thorson2013].
+This is essentially a Gompertz model for fishing effort, that has been used for
+other theoretical studies of fishing effort dynamics [@thorson2013].
 
@@
 
-The degree of fishing pressure exerted by a given set of fishing fleets
-on each species is a function of a range of parameters including the
-total amount of effort (*E*), the selectivity ogives ($\alpha$,the curve
-specifying the proportion of individuals at a given age or size that
-will be caught by a unit of fishing mortality), fishing cost (*C*), the
-spatial distribution of the species affected by the fleet, the relative
-prices across species ($\Pi$), and the catchability coefficient of each
-fleet for each species ($q$). `marlin` provides a tuning option to help
-users achieve desired biological outcomes from their fleets.
-
-Users can tune their fleets in one of two ways. First, they can specify
-a target exploitation rate *u* for each species in their simulation.
-Taking all the other parameters of the model as given, `marlin` then
-adjusts the catchability coefficients $q_{f,s}$ for each fleet *f* and
-species *s* such that the equilibrium exploitation rate for each species
-matches the desired level. Second, they can specify a target total
-spawning stock biomass under fishing divided by total unfished spawning
-stock biomass, and the model will adjust the catchability coefficients
-$q_{f,s}$ for each fleet *f* and species *s* such that the equilibrium
-ratio of fished to unfished spawning biomass for each species matches
-the desired level.
-
-Users can also tune the fleet dynamics by specifying the ratio of costs
-to revenues. Price data for use in the model can be obtained relatively
-simply through literature reviews, market surveys, or local experts, so
-long as they are converted to the same units of biomass used in the
-model. However, cost parameters are more complicated, as translating say
-cost per day of fishing into the same representation of fishing effort
-used in the model is not straight forward to accomplish. As an
-alternative, users can specify an equilibrium cost-to-revenue ratio for
-the fleet, essentially the profit margins of the fleet in question, and
-`marlin` will tune the cost parameter to achieve this desired cost to
-revenue ratio given the other parameters in the model (e.g. price and
-biomass of the target species). So for example, while users may not have
-access to exact cost parameters, they may have a general sense that when
-operating under ideal conditions a given fleet has a profit margin of
-10%, i.e. the cost to revenue ratio is 90%. They can input the cost to
-revenue ratio as 0.9 then and allow `marlin` to tune the cost parameters
-accordingly.
+The degree of fishing pressure exerted by a given set of fishing fleets on each
+species is a function of a range of parameters including the total amount of
+effort (*E*), the selectivity ogives ($\alpha$,the curve specifying the
+proportion of individuals at a given age or size that will be caught by a unit
+of fishing mortality), fishing cost (*C*), the spatial distribution of the
+species affected by the fleet, the relative prices across species ($\Pi$), and
+the catchability coefficient of each fleet for each species ($q$). `marlin`
+provides a tuning option to help users achieve desired biological outcomes from
+their fleets.
+
+Users can tune their fleets in one of two ways, conditional on the underlying
+population dynamics of the species in question. First, they can specify a target
+exploitation rate *u* for each species in their simulation. Taking all the other
+parameters of the model as given, `marlin` then adjusts the catchability
+coefficients $q_{f,s}$ for each fleet *f* and species *s* such that the
+equilibrium exploitation rate for each species matches the desired level.
+Second, they can specify a target total spawning stock biomass under fishing
+divided by total unfished spawning stock biomass, and the model will adjust the
+catchability coefficients $q_{f,s}$ for each fleet *f* and species *s* such that
+the equilibrium ratio of fished to unfished spawning biomass for each species
+matches the desired level.
+
+Users can also tune the fleet dynamics by specifying the ratio of costs to
+revenues. Price data for use in the model can be obtained relatively simply
+through literature reviews, market surveys, or local experts, so long as they
+are converted to the same units of biomass used in the model. However, cost
+parameters are more complicated, as translating say cost per day of fishing into
+the same representation of fishing effort used in the model is not straight
+forward to accomplish. As an alternative, users can specify an equilibrium
+cost-to-revenue ratio for the fleet, essentially the profit margins of the fleet
+in question, and `marlin` will tune the cost parameter to achieve this desired
+cost to revenue ratio given the other parameters in the model (e.g. price and
+biomass of the target species). So for example, while users may not have access
+to exact cost parameters, they may have a general sense that when operating
+under ideal conditions a given fleet has a profit margin of 10%, i.e. the cost
+to revenue ratio is 90%. They can input the cost to revenue ratio as 0.9 then
+and allow `marlin` to tune the cost parameters accordingly.
 
@@
 
-Each fishing fleet decides how to allocate its effort in space based on
-one of four possible spatial allocation strategies. The ideal free
-distribution (IFD) is the standard method for distributing fishing
-fleets in spaces (see @gillis2003 and references therein). However,
-analytical solutions to the IFD present a number of complications for
-our model. Our model is capable of tracking multiple fleets at once.
-This means that the IFD for a single fleet would commonly be modeled as
-a Nash-equilibrium based on the actions of each of the individual fleet
-conditional on the actions of all other fleets. While possible to
-implement, this would slow down our model runs to the point of
+Each fishing fleet decides how to allocate its effort in space based on one of
+four possible spatial allocation strategies. The ideal free distribution (IFD)
+is the standard method for distributing fishing fleets in spaces (see
+@gillis2003 and references therein). However, analytical solutions to the IFD
+present a number of complications for our model. Our model is capable of
+tracking multiple fleets at once. This means that the IFD for a single fleet
+would commonly be modeled as a Nash-equilibrium based on the actions of each of
+the individual fleet conditional on the actions of all other fleets. While
+possible to implement, this would slow down our model runs to the point of
 practically preventing large-scale evaluation of the spatial management
@@
 
-As such we explored a series of "next best" fleet distribution
-algorithms. While not the IFD in any individual time step, over time
-they start to approximate the IFD, as each assumes that fleets base
-their decisions for the current time step on the outcomes in the prior
-timestep, meaning that the impacts of the actions of other fishing
-fleets are somewhat internalized. The four possible fleet distribution
-algorithms are
-
-1.  Revenue per unit effort (RPUE): The fleet distributes itself in
-    space based on the realized revenue per unit effort in each patch in
-    the preceding time step
-2.  Revenue: The fleet distributes itself in space based on the realized
-    total revenue in each patch in the preceding time step
-3.  Profit per unit effort (PPUE): The fleet distributes itself in space
-    based on the realized profit per unit effort in each patch in the
-    preceding time step
-4.  Profit: The fleet distributes itself in space based on profit in
-    each patch in the preceding time step
-
-Revenue based spatial distribution is not likely to be very realistic;
-in general we would expect fishing fleets to respond to profits on some
-core level. However, due to the complexity of parameterizing cost
-functions, fleet dynamics are often evaluated based on yield or revenue
-alone, and so we include those scenarios here to allow users to evaluate
-the potential implications of this choice. The decision on whether to
-allocate the fleet based on absolute or relative (per unit effort)
-metrics is more complex. When effort represents the actions of separate
-and individual fishing actors (e.g. independent fishing vessels), a
-per-unit-effort strategy, in which fishers distribute themselves based
-on the expected catch of their individual efforts, may be more realistic
-[@hilborn1987]. Conversely, a system defined by a sole owner seeking to
-maximize total profits might be better represented by a fleet model
-based on total profits
+As such we explored a series of "next best" fleet distribution algorithms. While
+not the IFD in any individual time step, over time they start to approximate the
+IFD, as each assumes that fleets base their decisions for the current time step
+on the outcomes in the prior timestep, meaning that the impacts of the actions
+of other fishing fleets are somewhat internalized. The timeline required for the
+fleets to reach an equilibrium distribution will vary and depend on factors such
+as the population dynamics of the species in question (e.g. movement, habitat
+heterogeneity, growth rates, density dependence) and the number and degree of
+competition across fleets. Users should explore different simulation times to
+ensure that any results they wish to use are not simply a reflection of the
+fleet dynamics fluctuating on their way to an equilibrium condition.
+
+The four possible fleet distribution algorithms are
+
+1.  Revenue per unit effort (RPUE): The fleet distributes itself in space based
+    on the realized revenue per unit effort in each patch in the preceding time
+    step
+2.  Revenue: The fleet distributes itself in space based on the realized total
+    revenue in each patch in the preceding time step
+3.  Profit per unit effort (PPUE): The fleet distributes itself in space based
+    on the realized profit per unit effort in each patch in the preceding time
+    step
+4.  Profit: The fleet distributes itself in space based on profit in each patch
+    in the preceding time step
+
+Revenue based spatial distribution is not likely to be very realistic; in
+general we would expect fishing fleets to respond to profits on some core level.
+However, due to the complexity of parameterizing cost functions, fleet dynamics
+are often evaluated based on yield or revenue alone, and so we include those
+scenarios here to allow users to evaluate the potential implications of this
+choice. The decision on whether to allocate the fleet based on absolute or
+relative (per unit effort) metrics is more complex. When effort represents the
+actions of separate and individual fishing actors (e.g. independent fishing
+vessels), a per-unit-effort strategy, in which fishers distribute themselves
+based on the expected catch of their individual efforts, may be more realistic
+[@hilborn1987]. Conversely, a system defined by a sole owner seeking to maximize
+total profits might be better represented by a fleet model based on total
+profits.
 
@@
 
-The underlying population model used is an age structured single-species
-model in the manner of @ovando2021a. The population model requires many
-parameters. However, if the user supplies either a scientific or common
-name for the species in question (scientific preferred), the model will
-supply default values for that species based on the values reported in
-FishLife [internet connection required, @thorson2020]. FishLife provides
-estimates of core life history parameters for fish species based on a
-model integrating published values and evolutionary connections.
-FishLife provides reasonable default values for given species, but these
-default values should be not be taken as definitive and users should
-input best available parameters specific to the stock in question if
-they wish to best represent the dynamics of their specific system, and
-should check the plausibility of the default values prior to use. All of
-the life history values can be replaced with user-supplied values as
-needed and should be for more tactical applications; the default values
-simply ensure that reasonable and internally consistent values for the
-species in question are supplied as a starting point.
+The underlying population model used is an age structured single-species model
+in the manner of @ovando2021a. The population model requires many parameters.
+However, if the user supplies either a scientific or common name for the species
+in question (scientific preferred), the model will supply default values for
+that species based on the values reported in FishLife [internet connection
+required, @thorson2020]. FishLife provides estimates of core life history
+parameters for fish species based on a model integrating published values and
+evolutionary connections. FishLife provides reasonable default values for given
+species, but these default values should be not be taken as definitive and users
+should input best available parameters specific to the stock in question if they
+wish to best represent the dynamics of their specific system, and should check
+the plausibility of the default values prior to use. All of the life history
+values can be replaced with user-supplied values as needed and should be for
+more tactical applications; the default values simply ensure that reasonable and
+internally consistent values for the species in question are supplied as a
+starting point.
 
@@
 
-`marlin`'s movement dynamics are based on a continuous-time Markov chain
-(CTMC), as described in @thorson2021a. Within this framework, the model
-allows for movement to be broken down into three components of
-*advection* (drifting with currents), *taxis* (active movement towards
-preferred habitat), and *diffusion* (essentially remaining variation in
-movement not explained by advection or taxis). For now, `marlin` focuses
-just on the diffusion and taxis components of this model, assuming that
-advection is zero, though future extensions could easily incorporate
-advection vectors from oceanographic models. In this way, `marlin`
-allows users to run anything from a simple Gaussian dispersal kernel up
-to a system governed by species that passively diffuse out from a core
-habitat defined by a dynamic thermal range.
+`marlin`'s movement dynamics are based on a continuous-time Markov chain (CTMC),
+as described in @thorson2021a. Within this framework, the model allows for
+movement to be broken down into three components of *advection* (drifting with
+currents), *taxis* (active movement towards preferred habitat), and *diffusion*
+(essentially remaining variation in movement not explained by advection or
+taxis). For now, `marlin` focuses just on the diffusion and taxis components of
+this model, assuming that advection is zero, though future extensions could
+easily incorporate advection vectors from oceanographic models. In this way,
+`marlin` allows users to run anything from a simple Gaussian dispersal kernel up
+to a system governed by species that passively diffuse out from a core habitat
+defined by a dynamic thermal range.
 
 We provide a brief overview of the the general CTMC method here (see
-@thorson2021a for a detailed description). Under this framework,
-movement of individuals from each patch to each other patch in the
-system in a given timestep *t* for life stage *a* of species *s* is
-defined by a movement matrix $\pmb{M}_{t,s,a}$. $\pmb{M}_{t,s,a}$ is
-calculated as a function of diffusion $\pmb{D}$ and taxis $\pmb{\tau}$
-matrices scaled by the width of the time step $\Delta_{t}$ and the
-length of the edge of each patch $\Delta_d$ that the model is running
-under. This parameterization allows users to set the effective area of
-the spatial domain through two avenues; the number of patches, which
-effectively scales the resolution of the model, and the area of each
-patch, which scales the spatial extent of the simulation.
-
-The individual components (*M*) of the movement matrix ($\pmb{M}$) are
-filled based on an adjacency matrix, which defines whether two patches
-are both adjacent and water (as opposed to land or another physical
-barrier), a diffusion rate $D$ defined in units of area of a patch per
-unit of time, and a habitat preference function *H* in units of length
-of a side of a patch per unit time. For example, if we are defining the
-time units as years and the distance units as kilometers, for a tuna $D$
-might be 1,000 $\frac{KM^2}{Year}$. We then use parameters $\Delta_{t}$
-and $\Delta_{d}$ parameters to translate the diffusion rate $D$ to match
-the time step and patch size used in a simulation. For example, if we
-were to run a model on a monthly timestep given time units of years,
-then $\Delta_{t} = 1/12 years$. If one square patch in the simulation
-has an area of 100km^2^, then $\Delta_d = 10KM$. This "scale free"
-parameterization means that appropriate value of $D$ can be identified
-for a species and then set, regardless of the time step or patch size
-used in the simulation model itself. The taxis component of the movement
-process is a function of the difference in habitat quality *H*. The
-habitat preference function itself can take any form the user wishes.
-Exponentiating the difference in the habitat preference function between
-patches turns the taxis matrix into a multiplier of the diffusion rate
-*D*. As such, when creating habitat layers for simulation, users can
-tune the scale of the habitat gradient function to result in realistic
+@thorson2021a for a detailed description). Under this framework, movement of
+individuals from each patch to each other patch in the system in a given
+timestep *t* for life stage *a* of species *s* is defined by a movement matrix
+$\pmb{M}_{t,s,a}$. $\pmb{M}_{t,s,a}$ is calculated as a function of diffusion
+$\pmb{D}$ and taxis $\pmb{\tau}$ matrices scaled by the width of the time step
+(e.g. one year) $\Delta_{t}$ and the length of the edge of each patch (e.g. one
+kilometer) $\Delta_d$ specified by the user. This parameterization allows users
+to set the effective area of the spatial domain through two avenues; the number
+of patches, which effectively scales the resolution of the model, and the area
+of each patch, which scales the spatial extent of the simulation.
+
+The individual components (*M*) of the movement matrix ($\pmb{M}$) are filled
+based on an adjacency matrix, which defines whether two patches are both
+adjacent and water (as opposed to land or another physical barrier), a diffusion
+rate $D$ defined in units of area of a patch per unit of time, and a habitat
+preference function *H* in units of length of a side of a patch per unit time.
+For example, if we are defining the time units as years and the distance units
+as kilometers, for a tuna $D$ might be 1,000 $\frac{KM^2}{Year}$. We then use
+parameters $\Delta_{t}$ and $\Delta_{d}$ parameters to translate the diffusion
+rate $D$ to match the time step and patch size used in a simulation. For
+example, if we were to run a model on a monthly timestep given time units of
+years, then $\Delta_{t} = 1/12 years$. If one square patch in the simulation has
+an area of 100km^2^, then $\Delta_d = 10KM$. This "scale free" parameterization
+means that appropriate value of $D$ can be identified for a species and then
+set, regardless of the time step or patch size used in the simulation model
+itself. The taxis component of the movement process is a function of the
+difference in habitat quality *H*. The habitat preference function itself can
+take any form the user wishes. Exponentiating the difference in the habitat
+preference function between patches turns the taxis matrix into a multiplier of
+the diffusion rate *D*. As such, when creating habitat layers for simulation,
+users can tune the scale of the habitat gradient function to result in realistic
 multipliers of the diffusion rate. This parameterization ensure that the
@@
 For both the diffusion and taxis matrices, we allow for the inclusion of
-physical barriers to movement (i.e. land). Pairs of patches that are
-adjacent but in which one or both patches are a barrier to movement are
-set as non-adjacent. The CTMC model then produces movement dynamics that
-move around barriers rather than over them. This is different than
-setting habitat in a patch to zero; an animal will not preferentially
-move towards a patch with zero habitat all else being equal, but can
-move over a patch with zero habitat towards another better patch. An
-example might be a species that lives and feeds on coastal habitats
-transiting through a patch of open ocean to reach a new feeding ground.
-In contrast, a physical barrier represents something like a peninsula
-that an animal must move around to transit from one side to another.
+physical barriers to movement (i.e. land). Pairs of patches that are adjacent
+but in which one or both patches are a barrier to movement are set as
+non-adjacent. The CTMC model then produces movement dynamics that move around
+barriers rather than over them. This is different than setting habitat in a
+patch to zero; an animal will not preferentially move towards a patch with zero
+habitat all else being equal, but can move over a patch with zero habitat
+towards another better patch. An example might be a species that lives and feeds
+on coastal habitats transiting through a patch of open ocean to reach a new
+feeding ground. In contrast, a physical barrier represents something like a
+peninsula that an animal must move around to transit from one side to another.
 
@@
 multiplication of the pre-movement vector of the number of individuals
-($\pmb{n}$) of species *s* at age *a* in time step *t* across all
-patches *p* times the matrix exponential of the movement matrix
-$\pmb{M}$
+($\pmb{n}$) of species *s* at age *a* in time step *t* across all patches *p*
+times the matrix exponential of the movement matrix $\pmb{M}$
 
@@
 
-While this CTMC approach to movement simulation is to date not commonly
-seen in the marine modeling literature, it has numerous advantages that
-warrant its seeming complexity. First, the parameters of the model have
-interpretable biological meaning (e.g. the diffusion rate *D*). Second,
-when only diffusion is present, the model will generalize to the
-familiar dynamics of a Gaussian dispersal kernel at whatever spatial and
-temporal resolution the simulation is set to. Third, the taxis model
-allow for clearly parameterized active habitat choices by species,
-allowing us to simulate preferences of species in space and time
-efficiently. Lastly, the CTMC form has the advantage that its parameters
-are directly estimable from real data. So, if provided with for example
-spatial abundance data and a tagging study, users can estimate the
-diffusion and taxis movement parameters in the same manner as
-@thorson2021a, and then pass their estimated parameters to `marlin` for
-simulation (so long as the estimating method uses the same functional
-form as the movement model in `marlin`).
+While this CTMC approach to movement simulation is to date not commonly seen in
+the marine modeling literature, it has numerous advantages that warrant its
+seeming complexity. First, the parameters of the model have interpretable
+biological meaning (e.g. the diffusion rate *D*). Second, when only diffusion is
+present, the model will generalize to the familiar dynamics of a Gaussian
+dispersal kernel at whatever spatial and temporal resolution the simulation is
+set to. Third, the taxis model allow for clearly parameterized active habitat
+choices by species, allowing us to simulate preferences of species in space and
+time efficiently. Lastly, the CTMC form has the advantage that its parameters
+are directly estimable from real data. So, if provided with for example spatial
+abundance data and a tagging study, users can estimate the diffusion and taxis
+movement parameters in the same manner as @thorson2021a, and then pass their
+estimated parameters to `marlin` for simulation (so long as the estimating
+method uses the same functional form as the movement model in `marlin`).
 
@@
 
-For the population model, numbers *n* at time *t* for age *a* are a
-function of growth, death, and recruitment
-
-$$
-n_{t,p,s,a} = \begin{cases}
-      = BH(ssb_{t-1,p,s,a}) & \text{if $a = 1$}\\
-     = n_{t-1,p,s,a-1}e^{-(z_{t-1,p,s,a-1}}, & \text{if $1< a < max(age)$}\\
-     =  n_{t-1,p,s,a}e^{-(z_{t-1,p,s,a})} + n_{t-1,a-1}e^{-(z_{t-1,p,s,a-1})}, & \text{if $a = max(a)$}
+For the population model, numbers *N* at time *t* for age *a* are a function of
+growth, death, and recruitment
+
+$$
+N_{t,p,s,a} = \begin{cases}
+      = BH(SSB_{t-1,p,s,a}) & \text{if $a = 1$}\\
+     = N_{t-1,p,s,a-1}e^{-(z_{t-1,p,s,a-1}}, & \text{if $1< a < max(age)$}\\
+     =  N_{t-1,p,s,a}e^{-(z_{t-1,p,s,a})} + N_{t-1,a-1}e^{-(z_{t-1,p,s,a-1})}, & \text{if $a = max(a)$}
   \end{cases}
@@
 
-where *BH* is the Beverton-Holt recruitment function [@beverton1957] and
-*ssb* is spawning-stock-biomass. Per convention, the model allows for a
-"plus group", wherein rather than tracking numbers of every possible
-age, individuals greater than or equal to a given maximum age are
-grouped together.
-
-Spawning stock biomass *ssb* is calculated by converting age to mean
-length, calculating weight at age, maturity at age, and then calculating
-spawning stock biomass as the sum of spawning potential at age in a
-given time step, taking into account the potential for hyperallometry in
-the manner of @marshall2021. Age is converted to length through the von
-Bertalanffy growth equation given parameters asymptotic length
-($l_\infty$), growth (*k*) and theoretical age at length zero ($a0$)
-assuming log-normally distributed variation *u* in the length at age
-with CV $\sigma_s$.
+where *BH* is the Beverton-Holt recruitment function [@beverton1957] and *SSB*
+is spawning-stock-biomass. Per convention, the model allows for a "plus group",
+wherein rather than tracking numbers of every possible age, individuals greater
+than or equal to a given maximum age are grouped together.
+
+Spawning stock biomass *SSB* is calculated by converting age to mean length,
+calculating weight at age, maturity at age, and then calculating spawning stock
+biomass as the sum of spawning potential at age in a given time step, taking
+into account the potential for hyperallometry in the manner of @marshall2021.
+Age is converted to length through the von Bertalanffy growth equation given
+parameters asymptotic length ($l_\infty$), growth (*k*) and theoretical age at
+length zero ($a0$) assuming log-normally distributed variation *u* in the length
+at age with CV $\sigma_s$.
 
@@
 
-Users can manually supply a vector of of natural mortality at age (*m*).
-Or, they can supply one value of natural mortality which is then
-converted into mortality at age through one of two means. Under the
-default behavior, natural mortality at age given a target mean mortality
-across all ages $m_s$ is calculated using a length-inverse mortality
-function [@lorenzen2022].
+Users can manually supply a vector of of natural mortality at age (*m*). Or,
+they can supply one value of natural mortality which is then converted into
+mortality at age through one of two means. Under the default behavior, natural
+mortality at age given a target mean mortality across all ages $m_s$ is
+calculated using a length-inverse mortality function [@lorenzen2022].
 
@@
 
-Biomass *b* at age is then given by the weight at length equation
-governed by a scaling coefficient $\Omega_s$ and an exponent $\Phi_s$
-that controls the rate at which volume scales with length
-
-$$
-  b_{a,s} = _s \times l_{a,s}^{wb_s}
+Biomass *B* at age is then given by the weight at length equation governed by a
+scaling coefficient $\Omega_s$ and an exponent $\Phi_s$ that controls the rate
+at which volume scales with length
+
+$$
+  B_{a,s} = _s \times l_{a,s}^{wb_s}
 $$ {#eq-weight}
 
-The proportion of sexually mature individuals (*mat*) at a given age is
-then calculated as a logistic function where $l_{mat}$ is the length at
-which on average 50% of individuals are sexually mature, and
-$\delta_{mat}$ is the unit of length beyond $l_{mat}$ at which on
-average 95% of fish are sexually mature.
+The proportion of sexually mature individuals (*mat*) at a given age is then
+calculated as a logistic function where $l_{mat}$ is the length at which on
+average 50% of individuals are sexually mature, and $\delta_{mat}$ is the unit
+of length beyond $l_{mat}$ at which on average 95% of fish are sexually mature.
 
@@
 
-Spawning stock biomass at time *t* is then calculated as a function of
-the numbers at age, the maturity at age, and the weight at age raised by
-a parameter $\gamma$. When $\gamma$ is greater than 1, the species is
-said to experience hyperallometric fecundity, i.e. fecundity increases
-faster than weight.
-
-$$
-  ssb_{t,p,s} = \sum_{a=1}^{N_a}w_{a,s,t}^{\gamma_s} mat_{t,a} n_{t,p,s,a}
+Spawning stock biomass at time *t* is then calculated as a function of the
+numbers at age, the maturity at age, and the weight at age raised by a parameter
+$\gamma$. When $\gamma$ is greater than 1, the species is said to experience
+hyperallometric fecundity, i.e. fecundity increases faster than weight.
+
+$$
+  SSB_{t,p,s} = \sum_{a=1}^{N_a}w_{a,s,t}^{\gamma_s} mat_{t,a} N_{t,p,s,a}
 $$ {#eq-ssb}
@@
 
-Recruitment (i.e. the number of age 1 individuals entering the
-population) follows Beverton-Holt dynamics parameterized around
-steepness (*h*) with log-normally distributed recruitment deviates
-$\epsilon$. When steepness is one recruitment is independent of spawning
-biomass. As steepness approaches 0.2 recruitment becomes a linear
-function of spawning biomass. We allow for five variants in the timing
-of density dependent recruitment, building off of @babcock2011 :
-
-1.  Global density dependence: Density dependent recruitment is a
-    function of the sum of spawning biomass across all patches, and
-    recruits are then distributed according to habitat quality
+Recruitment (i.e. the number of age 1 individuals entering the population)
+follows Beverton-Holt dynamics parameterized around steepness (*h*) with
+log-normally distributed recruitment deviates $\epsilon$. When steepness is one
+recruitment is independent of spawning biomass. As steepness approaches 0.2
+recruitment becomes a linear function of spawning biomass. `marlin` allows users
+to specify a target unfinished spawning stock biomass ($SSB0$), which will be
+achieved by tuning the total unfished recruitment ($r0$), given the remaining
+life history parameters and independent of any characteristics of the fishing
+fleets.
+
+We allow for five variants in the timing of density dependent recruitment,
+building off of @babcock2011 :
+
+1.  Global density dependence: Density dependent recruitment is a function of
+    the sum of spawning biomass across all patches, and recruits are then
+    distributed according to habitat quality
 
@@
 \begin{aligned}
-n_{t,p,s,a = 1} = \left(\frac{0.8{\times}\sum_{p=1}^P{r0_{p,s}}\times{h_s}\times\sum_{p=1}^P{ssb_{t-1,p,s}}}{0.2\times{\sum_{p=1}^Pssb0_{p,s}} \times(1 - h_s)+(h_s - 0.2) \times{\sum_{p=1}^Pssb_{t-1,p,s}}}\right) \times \\ {r0_{p,s} / \sum_p^P{r0_{p,s}}\times \epsilon_{t,s}
+N_{t,p,s,a = 1} = \left(\frac{0.8{\times}\sum_{p=1}^P{r0_{p,s}}\times{h_s}\times\sum_{p=1}^P{SSB_{t-1,p,s}}}{0.2\times{\sum_{p=1}^PSSB0_{p,s}} \times(1 - h_s)+(h_s - 0.2) \times{\sum_{p=1}^PSSB_{t-1,p,s}}}\right) \times \\ {r0_{p,s} / \sum_p^P{r0_{p,s}}\times \epsilon_{t,s}
 }
@@
 
-where $r0$ is is a vector of recruits under unfished conditions in a
-given patch.
-
-2.  Local density dependence: Density dependent recruitment occurs
-    independently in each patch and recruits are retained in their home
+where $r0$ is is a vector of recruits under unfished conditions in a given
 patch.
 
-$$
-n_{t,p,s,a = 1} = \left(\frac{0.8{\times}r0_{p,s}\times{h_s}\times{ssb_{t-1,p,s}}}{0.2\times{ssb0_{p,s}}\times(1 - h_s)+(h_s - 0.2)\times{ssb_{t-1,p,s}}}\right) \times \epsilon_{t,s}
+2.  Local density dependence: Density dependent recruitment occurs independently
+    in each patch and recruits are retained in their home patch.
+
+$$
+n_{t,p,s,a = 1} = \left(\frac{0.8{\times}r0_{p,s}\times{h_s}\times{SSB_{t-1,p,s}}}{0.2\times{SSB0_{p,s}}\times(1 - h_s)+(h_s - 0.2)\times{SSB_{t-1,p,s}}}\right) \times \epsilon_{t,s}
 $$ {#eq-dd1}
 
-3.  Local density dependence then disperse: Density dependent
-    recruitment occurs independently in each patch and recruits are then
-    dispersed.
-
-$$
-  n_{t,p,s,a = 1} = \left(\frac{0.8{\times}r0_{p,s}\times{h_s}\times{ssb_{t-1,p,s}}}{0.2\times{ssb0_{p,s}}\times(1 - h_s)+(h_s - 0.2)\times{ssb_{t-1,p,s}}}\right)\times \pmb{d^l_s} \times \epsilon_{t,s}
+3.  Local density dependence then disperse: Density dependent recruitment occurs
+    independently in each patch and recruits are then dispersed.
+
+$$
+  n_{t,p,s,a = 1} = \left(\frac{0.8{\times}r0_{p,s}\times{h_s}\times{SSB_{t-1,p,s}}}{0.2\times{SSB0_{p,s}}\times(1 - h_s)+(h_s - 0.2)\times{SSB_{t-1,p,s}}}\right)\times \pmb{d^l_s} \times \epsilon_{t,s}
 $$ {#eq-dd3}
@@
 
-4.  Post-dispersal density dependence: Larvae are distributed throughout
-    the system, and then density dependent recruitment occurs based on
-    the density of spawning biomass at the destination patch.
-
-$$
-  larv_{t,p,s} = ssb_{t-1,p,s} \times\pmb{d^l_s}
+4.  Post-dispersal density dependence: Larvae are distributed throughout the
+    system, and then density dependent recruitment occurs based on the density
+    of spawning biomass at the destination patch.
+
+$$
+  larv_{t,p,s} = SSB_{t-1,p,s} \times\pmb{d^l_s}
 $$ {#eq-larvmove}
@@
 $$
-n_{t,a = 1,p,s} = \left(\frac{0.8{\times}r0_{p,s}\times{h_s}\times{larv_{t,p,s}}}{0.2\times{ssb0_{p,s}}\times(1 - h_s)+(h_s - 0.2)\times{larv_{t,p,s}}}\right) \times \epsilon_{t,s}
+n_{t,a = 1,p,s} = \left(\frac{0.8{\times}r0_{p,s}\times{h_s}\times{larv_{t,p,s}}}{0.2\times{SSB0_{p,s}}\times(1 - h_s)+(h_s - 0.2)\times{larv_{t,p,s}}}\right) \times \epsilon_{t,s}
 $$ {#eq-dd4}
 
-5.  Global density dependence allocated by spawning biomass: Density
-    dependence is a function of the sum of spawning biomass across all
-    patches, and recruits are then distributed according to the
-    distribution of spawning biomass
+5.  Global density dependence allocated by spawning biomass: Density dependence
+    is a function of the sum of spawning biomass across all patches, and
+    recruits are then distributed according to the distribution of spawning
+    biomass
 
@@
 \begin{aligned}
-n_{t,p,s,a = 1} = \left(\frac{0.8{\times}\sum_{p=1}^P{r0_{p,s}}\times{h_s}\times\sum_{p=1}^P{ssb_{t-1,p,s}}}{0.2\times{\sum_{p=1}^Pssb0_{p,s}}\times(1 - h_s)+(h_s - 0.2)\times{\sum_{p=1}^Pssb_{t-1,p,s}}}\right) \times \\ \frac{ssb_{t-1,p,s}}{\sum_{p=1}^Pssb_{t-1,p,s}} \times \epsilon_{t,s}
+n_{t,p,s,a = 1} = \left(\frac{0.8{\times}\sum_{p=1}^P{r0_{p,s}}\times{h_s}\times\sum_{p=1}^P{SSB_{t-1,p,s}}}{0.2\times{\sum_{p=1}^PSSB0_{p,s}}\times(1 - h_s)+(h_s - 0.2)\times{\sum_{p=1}^PSSB_{t-1,p,s}}}\right) \times \\ \frac{SSB_{t-1,p,s}}{\sum_{p=1}^PSSB_{t-1,p,s}} \times \epsilon_{t,s}
 \end{aligned}
@@
 
-Fisheries management is often concerned with measuring stock status
-relative to maximum sustainable yield (MSY) based reference points,
-though the exact level of stock status relative to MSY reference points
-desired by management or societal objectives may vary widely. MSY based
-reference points present a problem for a multi-fleet and
-spatial-temporal model such as `marlin`. MSY and the fishing mortality
-rate that would produce MSY, $F_{MSY}$, are a function of fishery
-selectivity. Fishery selectivity in this model can vary by fleet, and
-species can be distributed unevenly in both space and time. This means
-that the net effective fishing selectivity on a species can vary
-depending on the dynamics at a given moment, making the definition of an
-equilibrium concept such as MSY challenging.
-
-As such, we do not report MSY based reference points in the model by
-default, with the exception of the spawning stock biomass at MSY
-($SSB_{MSY}$) which tends to be conserved across selectivity patterns as
-it is primarily a function of the mortality and flow of recruits
-relative to the maturity ogive. There are many different strategies for
-estimating reference points in spatially explicit systems [@kapur2021a].
-We leave it to users to define and find relevant reference point as
-required by their specific needs.
+Fisheries management is often concerned with measuring stock status relative to
+maximum sustainable yield (MSY) based reference points, though the exact level
+of stock status relative to MSY reference points desired by management or
+societal objectives may vary widely. MSY based reference points present a
+problem for a multi-fleet and spatial-temporal model such as `marlin`. MSY and
+the fishing mortality rate that would produce MSY, $F_{MSY}$, are a function of
+fishery selectivity. Fishery selectivity in this model can vary by fleet, and
+species can be distributed unevenly in both space and time. This means that the
+net effective fishing selectivity on a species can vary depending on the
+dynamics at a given moment, making the definition of an equilibrium concept such
+as MSY challenging.
+
+As such, we do not report MSY based reference points in the model by default,
+with the exception of the spawning stock biomass at MSY ($SSB_{MSY}$) which
+tends to be conserved across selectivity patterns as it is primarily a function
+of the mortality and flow of recruits relative to the maturity ogive. There are
+many different strategies for estimating reference points in spatially explicit
+systems [@kapur2021a]. We leave it to users to define and find relevant
+reference point as required by their specific needs.
 
@@
 
-We include two examples demonstrating how `marlin` can be used to
-support marine spatial planning. In the first, we show how `marlin` can
-be used to compare the total and distributional impacts of MPAs designed
-in a heavily fished coastal coral reef ecosystem. In the second, we
-demonstrate how `marlin` can be used to assess components of climate
-resilience of alternative MPA design strategies in a pelagic system.
+We include two examples demonstrating how `marlin` can be used to support marine
+spatial planning. In the first, we show how `marlin` can be used to compare the
+total and distributional impacts of MPAs designed in a heavily fished coastal
+coral reef ecosystem. In the second, we demonstrate how `marlin` can be used to
+assess components of climate resilience of alternative MPA design strategies in
+a pelagic system.
 
 Each of the case studies contains too many parameters and options to be
-succinctly presented in the text here, and we are concerned that
-attempting to duplicate all of the parameters of the model within the
-text would simply introduce errors and confusion. As such, and since
-these results are for illustrative purposes only, readers should consult
-the accompanying code to view the precise details of each simulation.
-Targeted tactical applications must carefully consider and document all
-decisions made around model parameters.
+succinctly presented in the text here, and we are concerned that attempting to
+duplicate all of the parameters of the model within the text would simply
+introduce errors and confusion. As such, and since these results are for
+illustrative purposes only, readers should consult the accompanying code to view
+the precise details of each simulation. Targeted tactical applications must
+carefully consider and document all decisions made around model parameters.
 
@@
 1.  *Rate*: MPAs are placed based on the pre-MPA SSB/SSB0 weighted catch
-    relative to the total catch in a patch. So, patches with high rates
-    of catch of depleted species relative to total catch are
-    prioritized.
-
-2.  *Target Fishing*: MPAs locations are prioritized proportional to
-    fishery catches. Patches with high total catches are prioritized
-    over patches with low catches.
-
-3.  *Spawning Aggregation*: MPAs are centered on the location of a known
-    spawning aggregation
-
-To our knowledge, MPAs are never randomly placed; their locations are
-decided based on stakeholder negotiation given available information and
-objectives. In theory, the design of MPA networks can be optimized
-through the use of a modeling framework, and depending on the validity
-of the model, this process is likely to produce better outcomes than
-manually-designed strategies [@rassweiler2012; @rassweiler2014].
-However, designing optimal MPA networks becomes increasingly difficult
-as the range of objectives and the complexity of the model increase.
-Therefore, while `marlin` provides some support for
+    relative to the total catch in a patch. So, patches with high rates of catch
+    of depleted species relative to total catch are prioritized.
+
+2.  *Target Fishing*: MPAs locations are prioritized proportional to fishery
+    catches. Patches with high total catches are prioritized over patches with
+    low catches.
+
+3.  *Spawning Ground*: MPAs are centered on the grounds of a known spawning
+    aggregation
+
+To our knowledge, MPAs are never randomly placed; their locations are decided
+based on stakeholder negotiation given available information and objectives. In
+theory, the design of MPA networks can be optimized through the use of a
+modeling framework, and depending on the validity of the model, this process is
+likely to produce better outcomes than manually-designed strategies
+[@rassweiler2012; @rassweiler2014]. However, designing optimal MPA networks
+becomes increasingly difficult as the range of objectives and the complexity of
+the model increase. Therefore, while `marlin` provides some support for
 [optimization](https://danovando.github.io/marlin/articles/optimize-mpas.html),
-we focus here on the design and performance of these more "rule of
-thumb" design strategies that may be more accessible to a wider range of
-users. We allow all MPAs to be designed in a "mosaic" fashion in these
-examples, but users can easily extend the analysis to compare outcomes
-between "contiguous" (MPA is made up of one continuously connected
-block) and "mosaic" (MPAs can be separated in space) MPA designs
-[@pons2022] by developing and passing their own MPA design algorithms to
-the model.
+we focus here on the design and performance of these more rule of thumb design
+strategies that may be more accessible to a wider range of users. We allow all
+MPAs to be designed in a mosaic fashion in these examples, but users can easily
+extend the analysis to compare outcomes between contiguous (MPA is made up of
+one continuously connected block) and mosaic (MPAs can be separated in space)
+MPA designs [@pons2022] by developing and passing their own MPA design
+algorithms to the model.
 
@@
 
-In our coastal coral reef example, we model the dynamics of four
-tropical Pacific species: a grouper (*Epinephelus fuscoguttatus*), a
-shallow-reef snapper (*Lutjanus malabaricus*), a deep-water snapper
-(*Pristipomoides filamentosus*), and a reef shark (*Carcharhinus
-amblyrhynchos*). The simulated groupers undergo a mass migration to a
-spawning aggregation once per year, followed by the sharks. Shallow-reef
-snappers stay in reefs closer to shore above a steep drop-off
-year-round, while deep-water snappers stay in the deeper reefs past the
+In our coastal coral reef example, we model the dynamics of four tropical
+Pacific species: a grouper (*Epinephelus fuscoguttatus*), a shallow-reef snapper
+(*Lutjanus malabaricus*), a deep-water snapper (*Pristipomoides filamentosus*),
+and a reef shark (*Carcharhinus amblyrhynchos*). The simulated groupers undergo
+a mass migration to a spawning aggregation once per year, followed by the
+sharks. Shallow-reef snappers stay in reefs closer to shore above a steep
+drop-off year-round, while deep-water snappers stay in the deeper reefs past the
 drop-off (@fig-coral).
 
-These species are targeted by two different fleets. Fleet One primarily
-targets the grouper and near-shore snapper populations, but will land
-any incidentally captured sharks. Fleet One has a logistic selectivity
-pattern for all species, as they retain any fish caught for consumption
-or sale. both grouper and shallow-reef snapper due to demand for "plate
-sized" fish , and logistic selectivity for sharks, as any shark caught
-is retained. Fleet One is totally dependent on fishing for their
-livelihood, meaning the local community takes advantage of every
-possible opportunity to fish, and as such we model it as a "constant
-effort" fishery. Do to having less efficient boats, Fleet One has a
-higher cost per distance coefficient than Fleet Two. Fleet One's home
-port is located on the site of the grouper spawning aggregation.
-
-Fleet Two is a more commercial fleet that primarily targets the
-deep-water snapper population. This fleet primarily sells their catch to
-local restaurants and distributors where plate-sized fish are prized,
-and so for both snapper and grouper Fleet Two has a dome-shaped
-selectivity pattern [@kindsvater2017]. While plate-sized deep snapper
-are the primary target of Fleet Two, we model Fleet Two's selectivity
-for deep snapper as logistic due to high levels of discard mortality for
-deep-water snapper resulting from barotrauma. Fleet Two catches
-groupers, though less than Fleet One, and receives no price for sharks
-due to the requirements of a certification program through which they
-sell their deep-water snapper. Accidental captures (bycatch) of sharks
-do occur, which results in mortality. Fleet Two operates under
-open-access dynamics, as fishing is not the only means of subsistence
-for this community; short-term effort expands and contracts in response
-to profitability of the primarily grouper-driven fishery. Fleet two's
-home port is located further north from Fleet One.
+These species are targeted by two different fleets. Fleet One primarily targets
+the grouper and near-shore snapper populations, but will land any incidentally
+captured sharks. Fleet One has a logistic selectivity pattern for all species,
+as they retain any fish caught for consumption or sale. both grouper and
+shallow-reef snapper due to demand for "plate sized" fish , and logistic
+selectivity for sharks, as any shark caught is retained. Fleet One is totally
+dependent on fishing for their livelihood, meaning the local community takes
+advantage of every possible opportunity to fish, and as such we model it as a
+"constant effort" fishery. Due to having less efficient boats, Fleet One has a
+higher cost per distance coefficient than Fleet Two. Fleet One's home port is
+located on the site of the grouper spawning grounds.
+
+Fleet Two is a more commercial fleet that primarily targets the deep-water
+snapper population. This fleet primarily sells their catch to local restaurants
+and distributors where plate-sized fish are prized, and so for both snapper and
+grouper Fleet Two has a dome-shaped selectivity pattern [@kindsvater2017]. While
+plate-sized deep snapper are the primary target of Fleet Two, we model Fleet
+Two's selectivity for deep snapper as logistic due to high levels of discard
+mortality for deep-water snapper resulting from barotrauma. Fleet Two catches
+groupers, though less than Fleet One, and receives no price for sharks due to
+the requirements of a certification program through which they sell their
+deep-water snapper. Accidental captures (bycatch) of sharks do occur, which
+results in mortality. Fleet Two operates under open-access dynamics, as fishing
+is not the only means of subsistence for this community; short-term effort
+expands and contracts in response to profitability of the primarily
+grouper-driven fishery. Fleet two's home port is located further north from
+Fleet One.
 
 We use `marlin` to simulate the outcomes for both food production and
-conservation for each of the species and both of the fleets as a
-function of both MPA size and MPA design strategy. For this exercise,
-MPAs are placed with perfect information and have no design constraints
-for continuity (i.e. the selected MPA design can be as patchwork as
-dictated by the data). We ran the simulation on a quarterly level
-($\Delta_t = 1/4$) and set the area of each patch to be 5km^2^
-($\Delta_p = \sqrt{5}$), using 400 patches for a total area of 2,000
-KM^2^.
+conservation for each of the species and both of the fleets as a function of
+both MPA size and MPA design strategy. For this exercise, MPAs are placed with
+perfect information and have no design constraints for continuity (i.e. the
+selected MPA design can be as patchwork as dictated by the data). We ran the
+simulation for in quarter year time steps for 100 years ($\Delta_t = 1/4$) and
+set the area of each spatch to be 5km^2^ ($\Delta_p = \sqrt{5}$), using 400
+patches for a total area of 2,000 KM^2^.
 
@@
 
-We model our pelagic use case loosely on the characteristics of the
-Western and Central Pacific Ocean (WCPO) tuna fisheries. We simulate
-trajectories of `r n_distinct(names(blue_fauna))` species commonly
-caught in the region, including both the highly abundant skipjack tuna
-and the heavily depleted oceanic whitetip shark (*Carcharhinus
-longimanus*) (@fig-blue-ssb). We use publicly available data on
-catch-per-unit-effort of each of these species from the WCPO as a very
-rough proxy for baseline habitat distributions, noting that where
-possible, fishery-independent abundance indices would be preferable
+We model our pelagic use case loosely on the characteristics of the Western and
+Central Pacific Ocean (WCPO) tuna fisheries. We simulate trajectories of
+`r n_distinct(names(blue_fauna))` species commonly caught in the region,
+including both the highly abundant skipjack tuna and the heavily depleted
+oceanic whitetip shark (*Carcharhinus longimanus*) (@fig-blue-ssb). We use
+publicly available data on catch-per-unit-effort of each of these species from
+the WCPO as a very rough proxy for baseline habitat distributions, noting that
+where possible, fishery-independent abundance indices would be preferable
 (@fig-blue).
 
-These pelagic species are caught by a longline fleet that primarily
-targets large adult tunas such as bigeye and yellowfin for high-grade
-consumption, and a purse seine fishery that primarily targets skipjack
-tunas for bulk canning. Selectivities were modeled as logistic for the
-longline fleet, and dome-shaped for the purse seine fleet, with
-selectivity values set to broadly match the patterns reported in
-relevant [stock
+These pelagic species are caught by a longline fleet that primarily targets
+large adult tunas such as bigeye and yellowfin for high-grade consumption, and a
+purse seine fishery that primarily targets skipjack tunas for bulk canning.
+Selectivities were modeled as logistic for the longline fleet, and dome-shaped
+for the purse seine fleet, with selectivity values set to broadly match the
+patterns reported in relevant [stock
 assessments](https://oceanfish.spc.int/en/ofpsection/sam/sam/210-stock-assessment-input-and-result-files-)
-for each species. Both fleets operate under open-access dynamics with an
-effort cap. The effort cap was set at the level of effort that resulted
-in the desired levels of SSB/SSB0 for each species under open-access
-dynamics (@fig-blue), intended to simulate a scenario where managers
-step in to prevent further expansion of fishing effort in a fully
-developed fishery. For forward-simulation, open-access dynamics can
-result in effort decreasing in response to profitability, but cannot
-result in effort beyond the effort cap set for each fleet.
-
-For this exercise, we focused on using `marlin` to assess resilience of
-the selected Target Fishing and Rate "rule of thumb" MPA design
-strategies to a climate-driven range shift. Specifically, we simulate an
-extreme example where the centroid of each population shifts northward
-at a rate of 0.33 patches per year, over a 20 year time horizon
-(@fig-blue-ssb). We designed MPA networks given the conditions in the
-starting year, and then held that network constant over the years of the
-experiment, running one simulation with and another without the
-climate-drive range shift. We then compared the effects of this range
-shift on food production and conservation outcomes from MPA networks
-designed based on the pre-range shift world. We ran the pelagic
-simulation at a quarterly level ($\Delta_t$ = 1/4 year), and set the
-area of each cell to be roughly 97,000 KM^2^ across 144 patches each
-with a side length of roughly 311 KM, for a total area of 14,000,000
-KM^2^, broadly commensurate with the area of the WCPO.
+for each species. Both fleets operate under open-access dynamics with an effort
+cap. The effort cap was set at the level of effort that resulted in the desired
+levels of SSB/SSB0 for each species under open-access dynamics (@fig-blue),
+intended to simulate a scenario where managers step in to prevent further
+expansion of fishing effort in a fully developed fishery. For
+forward-simulation, open-access dynamics can result in effort decreasing in
+response to profitability, but cannot result in effort beyond the effort cap set
+for each fleet.
+
+For this exercise, we focused on using `marlin` to assess resilience of the
+selected Target Fishing and Rate "rule of thumb" MPA design strategies to a
+climate-driven range shift. Specifically, we simulate an extreme example where
+the centroid of each population shifts northward at a rate of 0.33 patches per
+year, over a 20 year time horizon (@fig-blue-ssb). We designed MPA networks
+given the conditions in the starting year, and then held that network constant
+over the years of the experiment, running one simulation with and another
+without the climate-drive range shift. We then compared the effects of this
+range shift on food production and conservation outcomes from MPA networks
+designed based on the pre-range shift world. We ran the pelagic simulation at a
+quarterly level ($\Delta_t$ = 1/4 year), and set the area of each cell to be
+roughly 97,000 KM^2^ across 144 patches each with a side length of roughly 311
+KM, for a total area of 14,000,000 KM^2^, broadly commensurate with the area of
+the WCPO.
 
@@
 #| label: fig-blue
-#| fig-cap: "Contour plot of fish spawning biomass in space under status quo (blue) and range-shifted (red) conditions for blue-water simulation.
-Packed together lines indicate higher biomass densities, farther spaced lines lower densities.
-The x-axis represents longitude, the y-axis latitude.
-"
+#| fig-cap: "Distribution of unfished spawning biomass in space under status quo (left column) and range-shifted (right column) conditions for blue-water simulation for each species.
+The x-axis represents longitude, the y-axis latitude."
+#| fig-height: 8
+#| fig-width: 6
 
@@
 
-patch_biomass %>%
+sq_habitat <- patch_biomass %>%
   left_join(sci_to_com, by = "critter") %>%
@@
   rename(critter = common_name) %>%
-  filter(step == min(step) | step == max(step)) %>% 
+  filter(step == min(step)) %>%
+  group_by(critter, step) %>%
+  mutate(sbiomass = biomass / max(biomass)) %>%
+  ungroup() %>%
+  ggplot() +
+  geom_tile(aes(x, y, fill = sbiomass),
+            linewidth = .1,
+            show.legend = FALSE) +
+  facet_wrap( ~ critter, ncol = 1) +
+  scale_x_continuous(expand = c(0, 0)) +
+  scale_y_continuous(expand = c(0, 0)) +
+  theme(
+    axis.text = element_blank(),
+    axis.ticks = element_blank(),
+    axis.title = element_blank(),
+    legend.position = "top",
+    axis.line = element_blank(),
+    panel.spacing.y = unit(0, "lines")
+  ) +
+  scale_fill_viridis() +
+  labs(title = "Status Quo")
+
+rs_habitat <- patch_biomass %>%
+  left_join(sci_to_com, by = "critter") %>% 
+  select(-critter) %>% 
+  rename(critter = common_name) %>% 
+  filter(step == max(step)) %>% 
   group_by(critter, step) %>%
@@
   ggplot()+
-  geom_contour(aes(x,y, z = sbiomass, color= factor(step)), linewidth = .1) +
-  facet_wrap(~critter) +
+  geom_tile(aes(x,y, fill = sbiomass), linewidth = .1, show.legend = FALSE) +
+  facet_wrap(~critter, ncol = 1) +
   scale_x_continuous(expand = c(0,0)) +
@@
         legend.position = "top",
-        axis.line = element_blank()) + 
-  scale_color_manual(values = c("#63B8FF","#FF4500"), labels = c("Status Quo","Range Shift"), name = '')
+        axis.line = element_blank(),
+        panel.spacing.y = unit(0, "lines")) + 
+  scale_fill_viridis() + 
+  labs(title = "Range Shift")
+
+range_shift_plot <- sq_habitat | rs_habitat
+
+range_shift_plot
 
@@
 
-For the coastal coral reef example both the *Spawning Ground* and
-*Target fishing* MPA design strategies were capable of producing a
-Pareto-optimal or improving, i.e. potentially win-win, outcome where
-change in total spawning biomass across all species and yield across all
-fleets was greater than or equal to zero for some MPA size. However,
-producing a total increase in SSB/SSB0 greater than 50% came at a net
-cost to fisheries yield across both design strategies. The *Spawning
-Aggregation* strategy produced more immediate trade-offs between food
-security and conservation (@fig-coral-results).
-
-These macro-level changes in total yield and conservation outcomes mask
-interesting internal dynamics. The *Spawning Ground* strategy increased
-yield to Fleet One when MPAs were relatively small, but was able to
-maintain yields near *status quo* levels up to a very large MPA size to
-the dependence of Fleet One on only a relatively small portion of the
-total simulation area. However, *Spawning Ground* MPAs were unable to
-provide increases in yield to Fleet Two. The *Spawning Ground* strategy
-produced the most rapid conservation gains for the snapper population.
-Interestingly, due to effort displacement from the spawning grounds onto
-the fishing areas inhabited by the groupers the rest of the year,
-closure of the spawning aggregation had limited impact on the grouper
-population until the MPA size exceeded 50% of the simulation area.
-Effort displacement from the spawning grounds also resulted in a small
-decrease in the deep snapper population for MPA sizes less than 50%. The
-*Target Fishing* design strategy has similar yield outcomes to the
-*Spawning Ground* strategy, but less dramatic negative impacts on Fleet
-Two, at the expense of slightly fewer benefits for Fleet One. However,
-the *Target Fishing* strategy provided more uniform conservation
-benefits to all species (@fig-coral-results).
-
-MPAs were capable of producing positive outcomes in terms of total
-biomass and yields in the coastal coral reef example, but all the tested
-strategies resulted in trade-offs across at least one dimension.
-`marlin` could be further used by a community such as the one simulated
-here to explore alternative design strategies, or to assess options for
-distributing fishery benefits across communities that may experience
-very different food provision outcomes from the same MPA network.
+For the coastal coral reef example both the *Spawning Ground* and *Target
+fishing* MPA design strategies were capable of producing a Pareto-optimal or
+improving, i.e. potentially win-win, outcome where change in total spawning
+biomass across all species and yield across all fleets was greater than or equal
+to zero for some MPA size. However, producing a total increase in SSB/SSB0
+greater than 50% came at a net cost to fisheries yield across both design
+strategies. The *Spawning Ground* strategy produced more immediate trade-offs
+between food security and conservation (@fig-coral-results).
+
+These macro-level changes in total yield and conservation are the result of
+interesting internal dynamics that `marlin` allows users to explore through
+output tables at the level of individual species and fleets. The *Spawning
+Ground* strategy increased yield to Fleet One when MPAs were relatively small,
+but was able to maintain yields near *status quo* levels up to a very large MPA
+size to the dependence of Fleet One on only a relatively small portion of the
+total simulation area. However, *Spawning Ground* MPAs were unable to provide
+increases in yield to Fleet Two. The *Spawning Ground* strategy produced the
+most rapid conservation gains for the snapper population. Interestingly, due to
+effort displacement from the spawning grounds onto the fishing areas inhabited
+by the groupers the rest of the year, closure of the spawning ground had limited
+impact on the grouper population until the MPA size exceeded 50% of the
+simulation area. Effort displacement from the spawning grounds also resulted in
+a small decrease in the deep snapper population for MPA sizes less than 50%. The
+*Target Fishing* design strategy has similar yield outcomes to the *Spawning
+Ground* strategy, but less dramatic negative impacts on Fleet Two, at the
+expense of slightly fewer benefits for Fleet One. However, the *Target Fishing*
+strategy provided more uniform conservation benefits to all species
+(@fig-coral-results).
+
+MPAs were capable of producing positive outcomes in terms of total biomass and
+yields in the coastal coral reef example, but all the tested strategies resulted
+in trade-offs across at least one dimension. `marlin` could be further used by a
+community such as the one simulated here to explore alternative design
+strategies, or to assess options for distributing fishery benefits across
+communities that may experience very different food provision outcomes from the
+same MPA network.
 
@@
 frontier_labels <- tmp %>%
+  filter(prop_mpa < 0.7) |>
   mutate(delta_50 = (delta_yield - -0.5)^2,
@@
 coral_frontier <- tmp %>%
+  filter(prop_mpa < 0.7) |>
   ggplot(aes(delta_biodiv, delta_yield)) +
   geom_hline(aes(yintercept = 0), linetype = 2) +
-  geom_line(alpha = 0.8, aes(color = placement_strategy),size = 1.1) + 
-  geom_text_repel(data = frontier_labels, aes(label = pmpa, color = placement_strategy), min.segment.length = 0, box.padding = 0, show.legend = FALSE, nudge_x = .25, size = 5) +
-  scale_x_continuous(name = "Change in Total SSB/SSB0", labels = scales::label_percent(accuracy = 1), guide = guide_axis(n.dodge = 2)) +
+  geom_path(alpha = 0.8, aes(color = placement_strategy),size = 1.1) +
+  geom_point(aes(fill = prop_mpa), shape = 21, size = 3) +
+  # geom_text_repel(data = frontier_labels, aes(label = pmpa), min.segment.length = 0, box.padding = 0.5) +
+  scale_x_continuous(name = "Change in Total SSB/SSB0", labels = scales::label_percent(accuracy = 1), guide = guide_axis(n.dodge = 2), limits = c(0,0.3)) +
   scale_y_continuous(name = "Change in Yield", labels = scales::label_percent(accuracy = 1)) +
   scale_color_manual(values = c("tomato", "steelblue"),name = '') +
-  guides(fill = guide_legend(order = 1, override.aes = list(size = 4))) + 
+  # guides(fill = guide_legend(order = 1, override.aes = list(size = 4))) +
+  scale_fill_viridis(name = "% MPA", labels = scales::label_percent(accuracy = 1),guide = guide_colorbar(ticks.colour = "black",frame.colour = "black"), option = "cividis") +
   labs(tag = "C")  +
   theme(
-    legend.position = c(0.7, .85),
+    legend.position = c(0.73, .73),
     legend.background = element_rect(fill = "transparent"),
@@
   )
-
 coral_fleet_yield <- coral_results %>%
@@
         axis.title.x = element_blank(),
-        axis.text.y = element_text(size = 8))  + 
+        axis.text.y = element_text(size = 8),
+        legend.margin = margin())  + 
     scale_color_discrete(name = '') + 
@@
     labs(tag = "B") + 
-  theme(axis.text = element_text(size = 8), axis.text.x = element_text(size = 8))
-
-((coral_fleet_yield / coral_critter_bio) | (coral_frontier + plot_layout(guides = "keep")) ) & theme(strip.text = element_text(size  = 6), plot.margin = margin(.1, .1, .1, .1, "cm"), panel.spacing = unit(0.5, "lines"), legend.text = element_text(size = 6))
+  theme(axis.text = element_text(size = 8), 
+        axis.text.x = element_text(size = 8),
+        legend.margin = margin())
+
+((coral_fleet_yield / coral_critter_bio) | (coral_frontier + plot_layout(guides = "keep")) ) & theme(strip.text = element_text(size  = 6), plot.margin = margin(.05, .05, .05, .05, "cm"), panel.spacing = unit(0.25, "lines"), legend.text = element_text(size = 6))
 
@@
 
-The yield impacts of the proposed MPA strategies varied across each of
-our simulations. Yields of the purse seine fleet were more resilient to
-the range shift under the *Rate* design strategy, while the longline
-fleet had the exact opposite result. Under the *Rate* strategy, longline
-yields were relatively stable across a large MPA size under the *status
-quo* conditions, but declined rapidly under the *range shift*
-conditions. For the purse seine fleet, yields were more stable under the
-*range shift* conditions under the *Target Fishing* design scenario, but
-declined quickly as a function of MPA size under the *status quo*
-conditions (@fig-blue-results). In terms of fisheries yield, the *Target
-Fishing* strategy would produce better yield outcomes for both fishing
-fleets under the *range shift* conditions, but the *Rate* strategy would
-perform best under the *status quo*.
-
-From a conservation perspective, the primary tuna species (bigeye,
-skipjack, and yellowfin) were most sensitive to design strategy and
-climate scenario, with the *Target Fishing* strategies producing for
-example rapid conservation gains for bigeye under the *status quo*,
-while the *Rate* strategy resulted in net conservation losses until the
-MPA became extremely large. However, under the *Range Shift* scenario
-the MPAs had little impact on many of the tuna populations until the MPA
-size became extremely large, do to movement of the primary fishing
-grounds outside of the current hotpots. The 20 year time horizon
-simulated here was not enough to produce substantial gains for any of
-the shark species even with 100% closures except for the more rapidly
-growing blue shark (@fig-blue-results).
+The yield impacts of the proposed MPA strategies varied across each of our
+simulations. Yields of the purse seine fleet were more resilient to the range
+shift under the *Rate* design strategy, while the longline fleet had the exact
+opposite result. Under the *Rate* strategy, longline yields were relatively
+stable across a large MPA size under the *status quo* conditions, but declined
+rapidly under the *range shift* conditions. For the purse seine fleet, yields
+were more stable under the *range shift* conditions under the *Target Fishing*
+design scenario, but declined quickly as a function of MPA size under the
+*status quo* conditions (@fig-blue-results). In terms of fisheries yield, the
+*Target Fishing* strategy would produce better yield outcomes for both fishing
+fleets under the *range shift* conditions, but the *Rate* strategy would perform
+best under the *status quo*.
+
+From a conservation perspective, the primary tuna species (bigeye, skipjack, and
+yellowfin) were most sensitive to design strategy and climate scenario, with the
+*Target Fishing* strategies producing for example rapid conservation gains for
+bigeye under the *status quo*, while the *Rate* strategy resulted in net
+conservation losses until the MPA became extremely large. However, under the
+*Range Shift* scenario the MPAs had little impact on many of the tuna
+populations until the MPA size became extremely large, do to movement of the
+primary fishing grounds outside of the current hotpots. The 20 year time horizon
+simulated here was not enough to produce substantial gains for any of the shark
+species even with 100% closures except for the more rapidly growing blue shark
+(@fig-blue-results).
 
@@
 
-Marine ecosystems are driven by complex social-ecological dynamics.
-Communities must often make decisions on how to manage these systems
-based on limited empirical evidence. Modeling tools such as `marlin` can
-help users answer important scientific questions and design marine
-management policies informed by a better understanding of sensitivities
-to key uncertainties. In this paper we have documented the core system
-of equations underlying `marlin`, and demonstrated how `marlin` can be
-used to support efforts such as MPA planning. The variation in both the
-magnitude and distribution of costs and benefits stemming from
-alternative design choices and states of nature shown in our case study
-results demonstrates the value in our method. For example, our coastal
+Marine ecosystems are driven by complex social-ecological dynamics. Communities
+must often make decisions on how to manage these systems based on limited
+empirical evidence. Modeling tools such as `marlin` can help users answer
+important scientific questions and design marine management policies informed by
+a better understanding of sensitivities to key uncertainties. In this paper we
+have documented the core system of equations underlying `marlin`, and
+demonstrated how `marlin` can be used to support efforts such as MPA planning.
+
+### Insights from Case Studies
+
+The variation in both the magnitude and distribution of costs and benefits
+stemming from alternative design choices and states of nature shown in our case
+study results demonstrates the value in our method. For example, our coastal
 coral reef example illustrates both the potential for MPAs to benefit
-conservation and food production in these systems, and the potential for
-the same MPA to benefit some fleets and species while harming others
-when fishing fleets affected by an MPA do not share the same objectives
-and species are not uniformly distributed. Our pelagic case study
-illustrates how `marlin` can be used to assess the climate resiliency of
-spatial management strategies, and identify strategies that best meet
-the needs of both current and future conditions. Results from this type
-of work could be used to help prioritize communities and species at
-particular risk to climate change impacts.
+conservation and food production in these systems, and the potential for the
+same MPA to benefit some fleets and species while harming others when fishing
+fleets affected by an MPA do not share the same objectives and species are not
+uniformly distributed. Our pelagic case study illustrates how `marlin` can be
+used to assess the climate resiliency of spatial management strategies, and
+identify strategies that best meet the needs of both current and future
+conditions. Results from this type of work could be used to help prioritize
+communities and species at particular risk to climate change impacts.
+
+xx add more details on case study here xx
+
+-   Coral
+
+    -   targeting spawning grounds for MPA provides better / more equitable
+        yield outcomes, but at the expense of much slower conservation gains,
+        and even some negative tradeoffs for species like deep snapper.
+        Targeting fishing provides better gains to fleet two (explain why), and
+        better conservation outcomes, but squarely at the expense of fleet one
+
+-   Blue Water
+
+    -   target-fishing design being much more robust to range shifts than rate
+        design, why?
+
+    -   High concentration of the purseseine fleet means much more sensitive to
+        placement "errors"
+
+    -   Basic life history poses challenge to recovery of severely depleted
+        species e.g. OWT
+
+### Putting `marlin` In Context
 
 The role of this paper is not to conduct a review of the many tools for
-spatial-temporal modeling available in the literature, each of which
-provide useful functionality for specific applications. However, we
-highlight here the specific gaps that we feel `marlin` fills in the
-modeling literature and in the policy support toolbox using some
-selected publications. More end-to-end models such as ATLANTIS
-[@audzijonyte2019] , OSMOSE [@shin2001], Ecopath with Ecosim
-[@christensen2004], POSEIDON [@bailey2019], DISPLACE [@bastardie2013],
-or SEAPODYM [@lehodey2008] are capable of representing tremendous
-amounts of complexity, but can be extremely time consuming to design and
-run. `marlin` allows some of the realism of these end-to-end models
-while being simpler and faster to parameterize and run.
-
-Tools such as virtualspecies [@leroy2016], STEPS [@visintin2020],
-RangeShifter [@bocedi2014], and
+spatial-temporal modeling available in the literature, each of which provide
+useful functionality for specific applications. However, we highlight here the
+specific gaps that we feel `marlin` fills in the modeling literature and in the
+policy support toolbox using some selected publications. More end-to-end models
+such as ATLANTIS [@audzijonyte2019] , OSMOSE [@shin2001], Ecopath with Ecosim
+[@christensen2004], POSEIDON [@bailey2019], DISPLACE [@bastardie2013], or
+SEAPODYM [@lehodey2008] are capable of representing tremendous amounts of
+complexity, but can be extremely time consuming to design and run. `marlin`
+allows some of the realism of these end-to-end models while being simpler and
+faster to parameterize and run.
+
+Tools such as virtualspecies [@leroy2016], STEPS [@visintin2020], RangeShifter
+[@bocedi2014], and
 [SMS](https://www.aquaveo.com/software/sms-surface-water-modeling-system-introduction)
-can be efficiently constructed to model the dispersal and distribution
-of species in space and time as a function of environmental covariates.
-However, these typically require specifying a covariate-response curve
-without explicitly acknowledging how this arises from habitat-specific
-movement and demography (virtualspecies), or model movement using a
-dispersal kernel (RangeShifter, STEPS) or a least-cost path algorithm
-(SMS, STEPS). By contrast, `marlin` uses a continuous-time Markov chain
-movement model, which integrates multiple paths (including their
-path-dependent probability based on intervening habitat types and
-species preferences) while using scale-free parameters that can be
-measured experimentally in laboratory or tagging studies.
-
-Simplified bio-economic models such as those used in @hastings2017,
-@sala2021, and @cabral2019 can be applied at scale and provide
-analytically tractable results, but as a result must abstract over many
-bio-economic dynamics that can be important for more tactical
-applications. Our results demonstrate how nuances in fleet dynamics and
-species distributions can dramatically impact MPA outcomes. `marlin`
-allows for more realistic representations of spatial social-ecological
-systems while maintaining processing speed.
-
-In sum, `marlin` complements the existing spatial marine modeling
-literature by allowing scientists, decision makers, and other
-stakeholders to efficiently examine the impacts of realistic
-bio-economic dynamics on academic and applied problems. For example,
-scientists could use `marlin` to consider the relative importance of
-different biological variables in affecting policy outcomes. Managers
-could use `marlin` to assess the potential of proposed MPA networks to
+can be efficiently constructed to model the dispersal and distribution of
+species in space and time as a function of environmental covariates. However,
+these typically require specifying a covariate-response curve without explicitly
+acknowledging how this arises from habitat-specific movement and demography
+(virtualspecies), or model movement using a dispersal kernel (RangeShifter,
+STEPS) or a least-cost path algorithm (SMS, STEPS). By contrast, `marlin` uses a
+continuous-time Markov chain movement model, which integrates multiple paths
+(including their path-dependent probability based on intervening habitat types
+and species preferences) while using scale-free parameters that can be measured
+experimentally in laboratory or tagging studies.
+
+Simplified bio-economic models such as those used in @hastings2017, @sala2021,
+and @cabral2019 can be applied at scale and provide analytically tractable
+results, but as a result must abstract over many bio-economic dynamics that can
+be important for more tactical applications. Our results demonstrate how nuances
+in fleet dynamics and species distributions can dramatically impact MPA
+outcomes. `marlin` allows for more realistic representations of spatial
+social-ecological systems while maintaining processing speed.
+
+### General Recommendations for Using `marlin`
+
+Marine social-ecological systems are highly complex. The model presented here is
+designed to help users explore the impacts of different variables on policy
+outcomes. But, that freedom comes at a cost, namely that users have a large
+number of options at their disposal that they must decide on. Our recommended
+strategy for dealing with this is for users to try and narrow down a list of
+parameters that they feel are sufficiently "known" and another list of
+parameters that are highly uncertain and / or contentious which the users feel
+may impact results.
+
+For example, a community seeking to a model a well-studied coastal coral reef
+ecosystem might leave as fixed the habitat distribution (as represented by reef
+locations) and general life history of the species in question (growth rates,
+age at maturity, etc). But, within those constraints, the users may wish to test
+the sensitivity of proposed MPA networks to key unknowns such as adult and
+larval dispersal rates or the the economic incentives and gear selectivity of
+the fishing fleets. As an example of this, we ran an alternative version of our
+Coral Reef case study in which Fleet Two was assigned logistic selectivity for
+the snapper and grouper species, rather than the dome-shaped selectivity
+presented in our main results. We found that the impacts of MPAs on conservation
+and food security were relatively insensitive to the form of fishing selectivity
+(Fig.S1). This does not mean that knowing the "correct" form of selectivity may
+not be extremely important in providing accurate assessment results
+[@waterhouse2014], but rather that holding other parameters constant, in this
+case misspecifying the selectivity curve is projected to have little impact on
+the simulated outcomes of MPAs.
+
+Given a set of model runs, the choice of whether to use those results
+"strategically" or "tactically" will depend on the needs of the user. That being
+said, since the model is not fit to data directly it is more easily applied to
+strategic questions. For example, the two case studies presented in this paper
+provide general strategic insights into the sensitivity of outcomes to design
+choices and ... xx expand here xx. But, because `marlin` is not fit to data, the
+degree to which it can represent detailed dynamics of a system depends on the
+quality of the input data; The extent to which users are comfortable using the
+outputs of `marlin` tactically will depend on the confidence they have in their
+parameterization of the model runs relative to the precision policy-makers
+require in order to make a decision.
+
+`marlin` is just one model out of many that can be used to inform marine
+management. Other models may be better suited to address specific forms of
+complexity, such as predatory-prey dynamics (citations) or more complex fleet
+dynamics through agent-based approached [@burgess2020]. Particularly for more
+tactical applications, we would encourage users to explore multiple modeling
+frameworks to help design policies that are likely to be robust to many
+different kinds of complexities.
+
+In sum, `marlin` complements the existing spatial marine modeling literature by
+allowing scientists, decision makers, and other stakeholders to efficiently
+examine the impacts of realistic bio-economic dynamics on academic and applied
+problems. For example, scientists could use `marlin` to consider the relative
+importance of different biological variables in affecting policy outcomes.
+Managers could use `marlin` to assess the potential of proposed MPA networks to
 achieve stated objectives, such as maximizing conservation gains for a
-threatened species while minimizing or removing negative impacts to
-affected fishing fleets.
-
-Beyond the current examples, we envision `marlin` being applicable to
-research on dynamic ocean management, range shifts, management strategy
-evaluation, policy interactions, and spatial stock assessment. In
-addition, the process-based movement model used in `marlin` can directly
-use empirical estimates of movement dynamics derived in the manner of
-@thorson2021a, providing a link between empirical and simulation based
-approaches to marine resource management. Future iterations of the model
-may include life history modules for a broader range of species types
-such as marine mammals, sea birds, and sea turtles.
-
-Even the most complex marine model is a stylized cartoon of the true
-dynamics at play in the ocean. However, for all their limitations,
-models such as `marlin` can help users understand factors that drive the
-performance of marine management strategies. The goal of this tool is to
-empower people to design policies based on evaluation of key
-uncertainties and trade-offs, and in doing so support more effective and
-equitable marine resource management.
+threatened species while minimizing or removing negative impacts to affected
+fishing fleets.
+
+Beyond the current examples, we envision `marlin` being applicable to research
+on dynamic ocean management, range shifts, management strategy evaluation,
+policy interactions, and spatial stock assessment. In addition, the
+process-based movement model used in `marlin` can directly use empirical
+estimates of movement dynamics derived in the manner of @thorson2021a, providing
+a link between empirical and simulation based approaches to marine resource
+management. The model currently focuses on diffusion and taxis, which allows for
+general representation of animals following physical or oceanographic features,
+but research on the incorporation and importance of advective forces would be of
+value going forward. Future iterations of the model may include life history
+modules for a broader range of species types such as marine mammals, sea birds,
+and sea turtles.
+
+## Conclusions
+
+Even the most complex marine model is a stylized cartoon of the true dynamics of
+ocean ecosystems. However, for all their limitations, models such as `marlin`
+can help users understand factors that drive the performance of marine
+management strategies. The goal of this tool is to empower people to design
+policies based on evaluation of key uncertainties and trade-offs, and in doing
+so support more effective and equitable marine resource management.
 
@@
 
-This project was funded by the Waitt Foundation. We thank Dr. Alexa
-Fredston for helpful revisions of this manuscript.
+This project was funded by the Waitt Foundation. We thank Dr. Alexa Fredston for
+helpful revisions of this manuscript.
 
@@
 
-The code, novel code, data, and materials needed to reproduce the
-results and all aspects of this manuscript are publicly available at
-https://github.com/DanOvando/marlin-paper, with supporting novel code
-available at https://github.com/DanOvando/marlin/. Upon acceptance, all
-data, materials, code, and novel code needed to reproduce the results
-and all aspects of this manuscript will be publicly available via GitHub
-at https://github.com/DanOvando/marlin-paper and
+The code, novel code, data, and materials needed to reproduce the results and
+all aspects of this manuscript are publicly available at
+https://github.com/DanOvando/marlin-paper, with supporting novel code available
+at https://github.com/DanOvando/marlin/. Upon acceptance, all data, materials,
+code, and novel code needed to reproduce the results and all aspects of this
+manuscript will be publicly available via GitHub at
+https://github.com/DanOvando/marlin-paper and
 https://github.com/DanOvando/marlin/ and through figshare at

### 2023-09-28 README.Rmd
cleaning up documentation and testing migration to speedy abalone
getting somewhere

@@
 
+## Knitting to PDF
+
+Note that if you want to knit to PDF you'll need to have appropriate LaTeX style software installed. I recommend [`tinytex`](https://yihui.org/tinytex/).
+
 ## Reproducing results on UNIX machines
@@
 
-4. In the terminal, run `RUN_NAME="MyRunName" FORMAT="MyFormat" make` where MyRunName is the name of the folder you want to create to store results (e.g. v0.1) and FORMAT is one of pdf, html, or docx
+4. In the terminal, run `RUN_NAME="MYRUNNAME" FORMAT="FORMAT" make` where MYRUNNAME is the name of the folder you want to create to store results (e.g. v0.1) and FORMAT is one of pdf, html, or docx
 

### 2023-10-07 03_marlin_paper.qmd
almost there

@@
 params:
-  run_name: ["v1.2"]
+  run_name: ["v1.1"]
 header-includes:
@@
 
-## Need to cite..
-
-@bosley2022
-
-@berger2017
-
 ## Introduction
@@
 `marlin` by a set of fishers that have the same fishing skill, prices, and
-selectivity for individual species (each of which we denote as a *metier*). For
-example, both a longline and purse seine fleet may capture bigeye (*Thunnus
-obesus*) and skipjack tunas, but the longline fleet can be made much more likely
-to catch larger bigeye than skipjack, and *vice versa*. Each fishing fleet
-distributes its fishing effort in space according to a specified spatial
-allocation rule (see *Spatial Allocation of Effort* ), for example in proportion
-to profit-per-unit-effort, conditional on management policies in places such as
-quotas and/or the presence of any spatial restrictions such as MPAs.
+contact selectivity [@sampson2014] for individual species (each of which we
+denote as a *metier*). For example, both a longline and purse seine fleet may
+capture bigeye (*Thunnus obesus*) and skipjack tunas, but the longline fleet can
+be made much more likely to catch larger bigeye than skipjack, and *vice versa*.
+Each fishing fleet distributes its fishing effort in space according to a
+specified spatial allocation rule (see *Spatial Allocation of Effort* ), for
+example in proportion to profit-per-unit-effort, conditional on management
+policies in places such as quotas and/or the presence of any spatial
+restrictions such as MPAs. It is important to note that contact selectivity
+reflects the ability of the fishing method in question to capture fish of
+different sizes that come into "contact" with the gear. The contact selectivity
+of each of the specified *metiers* will then interact with the distribution of
+fish sizes and fishing effort in space simulated by `marlin` to produce a net
+"population" selectivity, which may differ from the individual contact
+selectivities of each of the *metiers*.
 
@@
 simulation for in quarter year time steps for 100 years ($\Delta_t = 1/4$) and
-set the area of each spatch to be 5km^2^ ($\Delta_p = \sqrt{5}$), using 400
+set the area of each patch to be 5km^2^ ($\Delta_p = \sqrt{5}$), using 144
 patches for a total area of 2,000 KM^2^.
@@
 
-#### Pelagic Systems
+#### Pelagic Ecosystems
 
@@
   habitat = future_habitat,
-  years = 20,
+  years = length(future_habitat[[1]]),
   manager = list(mpas = list(
@@
 
-For the coastal coral reef example both the *Spawning Ground* and *Target
-fishing* MPA design strategies were capable of producing a Pareto-optimal or
-improving, i.e. potentially win-win, outcome where change in total spawning
-biomass across all species and yield across all fleets was greater than or equal
-to zero for some MPA size. However, producing a total increase in SSB/SSB0
-greater than 50% came at a net cost to fisheries yield across both design
-strategies. The *Spawning Ground* strategy produced more immediate trade-offs
-between food security and conservation (@fig-coral-results).
-
-These macro-level changes in total yield and conservation are the result of
-interesting internal dynamics that `marlin` allows users to explore through
-output tables at the level of individual species and fleets. The *Spawning
-Ground* strategy increased yield to Fleet One when MPAs were relatively small,
-but was able to maintain yields near *status quo* levels up to a very large MPA
-size to the dependence of Fleet One on only a relatively small portion of the
-total simulation area. However, *Spawning Ground* MPAs were unable to provide
-increases in yield to Fleet Two. The *Spawning Ground* strategy produced the
-most rapid conservation gains for the snapper population. Interestingly, due to
-effort displacement from the spawning grounds onto the fishing areas inhabited
-by the groupers the rest of the year, closure of the spawning ground had limited
-impact on the grouper population until the MPA size exceeded 50% of the
-simulation area. Effort displacement from the spawning grounds also resulted in
-a small decrease in the deep snapper population for MPA sizes less than 50%. The
-*Target Fishing* design strategy has similar yield outcomes to the *Spawning
-Ground* strategy, but less dramatic negative impacts on Fleet Two, at the
-expense of slightly fewer benefits for Fleet One. However, the *Target Fishing*
-strategy provided more uniform conservation benefits to all species
-(@fig-coral-results).
-
-MPAs were capable of producing positive outcomes in terms of total biomass and
-yields in the coastal coral reef example, but all the tested strategies resulted
-in trade-offs across at least one dimension. `marlin` could be further used by a
-community such as the one simulated here to explore alternative design
-strategies, or to assess options for distributing fishery benefits across
-communities that may experience very different food provision outcomes from the
-same MPA network.
+```{r}
+#| label: coral-results
+
+coral_mpa_experiments$mpas <- map(coral_mpa_experiments$results, "mpa")
+
+coral_mpa_experiments$obj <- map(coral_mpa_experiments$results, "obj")
+
+coral_results <- coral_mpa_experiments %>%
+  unnest(cols = obj)
+
+coral_fleet_results <- coral_results %>%
+  pivot_longer(starts_with("fleet_"),
+               names_to = "fleet",
+               values_to = "fleet_yield") %>%
+  group_by(prop_mpa, fleet, placement_strategy) %>%
+  summarise(yield = sum(fleet_yield), biodiv = sum(unique(biodiv))) %>%
+  group_by(fleet, placement_strategy) %>%
+  mutate(delta_yield = yield / yield[prop_mpa == 0] - 1) |> 
+  ungroup()
+
+peak_coral_fleet_results <- coral_fleet_results |> 
+  group_by(fleet) |> 
+  filter(delta_yield == max(delta_yield)) |> 
+  mutate(peak_yields = scales::percent(delta_yield))
+
+peak_coral_fleet_results |> filter(fleet == "fleet_two")
+
+tmp <- coral_results %>% 
+  group_by(prop_mpa, placement_strategy) %>% 
+  summarise(biodiv = sum(biodiv), yield = sum(yield)) %>% 
+  group_by(placement_strategy) %>% 
+  mutate(delta_yield = yield / yield[prop_mpa == 0] - 1,
+         delta_biodiv = biodiv / biodiv[prop_mpa == 0] - 1) %>% 
+  ungroup() %>% 
+    mutate(placement_strategy= fct_relabel(placement_strategy, titler))
+
+```
+
+MPAs were capable of producing a range of positive and negative outcomes for
+food security and conservation in the Coral Reef case study depending on the
+design strategy used and the size of MPA implemented. Both of the MPA design
+strategies were capable of increasing fisheries yield for Fleet Two, up to a
+value of
+`r peak_coral_fleet_results$peak_yields[peak_coral_fleet_results$fleet == "fleet_two"]`,
+even when MPAs covered more than 50% of the simulated area. However, Fleet One
+only benefited from MPAs under the *Spawning Ground* design strategy, with a
+maximum increase of
+`r peak_coral_fleet_results$peak_yields[peak_coral_fleet_results$fleet == "fleet_one"]`;
+the *Target Fishing* design strategy produced a roughly linear decrease in
+fishing yields as a function of increasing MPA size (@fig-coral-results A).
+
+On the conservation front, MPAs were uniformly beneficial to the spawning
+biomass of all species under the *Target Fishing* design strategy, with the most
+rapid increases for the Deep Snapper population. The *Spawning Ground* strategy
+primarily benefited the coastal Snapper population, producing little change in
+the Grouper population and decreasing spawning biomass of both the Deep Snapper
+and Reef Shark populations for MPA sizes covering less than 50% of the simulated
+area (@fig-coral-results B).
+
+In total, the *Spawning Ground* strategy was capable or providing net increases
+in fishing yield (summed across both fleets) of up to roughly 7%, with positive
+net yield impacts up to network size of nearly 60% of the area. However, even
+MPAs covering 60% of the area only produced a maximum increase of roughly 5% in
+total SSB/SSB0 when designed around the spawning ground. Conversely, the *Target
+Fishing* design strategy was capable of producing a nearly 25% increase in total
+SSB/SSB0 for the same MPA size, but at a much greater cost to the total food
+production from the system's fisheries (@fig-coral-results C).
+
+<!-- ~~These macro-level changes in total yield and conservation are the result of -->
+
+<!-- interesting internal dynamics that marlin allows users to explore through output -->
+
+<!-- tables at the level of individual species and fleets. The *Spawning Ground* -->
+
+<!-- strategy increased yield to Fleet One when MPAs were relatively small, but was -->
+
+<!-- able to maintain yields near *status quo* levels up to a very large MPA size to -->
+
+<!-- the dependence of Fleet One on only a relatively small portion of the total -->
+
+<!-- simulation area. However, *Spawning Ground* MPAs were unable to provide -->
+
+<!-- increases in yield to Fleet Two. The *Spawning Ground* strategy produced the -->
+
+<!-- most rapid conservation gains for the snapper population. Interestingly, due to -->
+
+<!-- effort displacement from the spawning grounds onto the fishing areas inhabited -->
+
+<!-- by the groupers the rest of the year, closure of the spawning ground had limited -->
+
+<!-- impact on the grouper population until the MPA size exceeded 50% of the -->
+
+<!-- simulation area. Effort displacement from the spawning grounds also resulted in -->
+
+<!-- a small decrease in the deep snapper population for MPA sizes less than 50%. The -->
+
+<!-- *Target Fishing* design strategy has similar yield outcomes to the *Spawning -->
+
+<!-- Ground* strategy, but less dramatic negative impacts on Fleet Two, at the -->
+
+<!-- expense of slightly fewer benefits for Fleet One. However, the *Target Fishing* -->
+
+<!-- strategy provided more uniform conservation benefits to all species -->
+
+<!-- (@fig-coral-results).~~ -->
 
@@
 #| fig-width: 6
-
-coral_mpa_experiments$mpas <- map(coral_mpa_experiments$results, "mpa")
-
-coral_mpa_experiments$obj <- map(coral_mpa_experiments$results, "obj")
-
-coral_results <- coral_mpa_experiments %>%
-  unnest(cols = obj)
 
@@
   facet_wrap(~ fleet, scales = "free_y")
-
-tmp <- coral_results %>% 
-  group_by(prop_mpa, placement_strategy) %>% 
-  summarise(biodiv = sum(biodiv), yield = sum(yield)) %>% 
-  group_by(placement_strategy) %>% 
-  mutate(delta_yield = yield / yield[prop_mpa == 0] - 1,
-         delta_biodiv = biodiv / biodiv[prop_mpa == 0] - 1) %>% 
-  ungroup() %>% 
-    mutate(placement_strategy= fct_relabel(placement_strategy, titler))
 
@@
   )
-coral_fleet_yield <- coral_results %>%
-  pivot_longer(starts_with("fleet_"),
-               names_to = "fleet",
-               values_to = "fleet_yield") %>%
-  group_by(prop_mpa, fleet, placement_strategy) %>%
-  summarise(yield = sum(fleet_yield), biodiv = sum(unique(biodiv))) %>%
-  group_by(fleet, placement_strategy) %>%
-  mutate(delta_yield = yield / yield[prop_mpa == 0] - 1) %>%
-  ungroup() %>% 
+coral_fleet_yield <- coral_fleet_results |> 
   mutate(fleet= fct_relabel(fleet, titler)) %>% 
@@
 
-### Pelagic Systems
+### Pelagic Ecosystems
 
@@
 design scenario, but declined quickly as a function of MPA size under the
-*status quo* conditions (@fig-blue-results). In terms of fisheries yield, the
+*status quo* conditions (@fig-blue-results A). In terms of fisheries yield, the
 *Target Fishing* strategy would produce better yield outcomes for both fishing
@@
 species even with 100% closures except for the more rapidly growing blue shark
-(@fig-blue-results).
+(@fig-blue-results B).
 
@@
 
-xx add more details on case study here xx
-
--   Coral
-
-    -   targeting spawning grounds for MPA provides better / more equitable
-        yield outcomes, but at the expense of much slower conservation gains,
-        and even some negative tradeoffs for species like deep snapper.
-        Targeting fishing provides better gains to fleet two (explain why), and
-        better conservation outcomes, but squarely at the expense of fleet one
-
--   Blue Water
-
-    -   target-fishing design being much more robust to range shifts than rate
-        design, why?
-
-    -   High concentration of the purseseine fleet means much more sensitive to
-        placement "errors"
-
-    -   Basic life history poses challenge to recovery of severely depleted
-        species e.g. OWT
+Examining our coral reef case study, our results indicate how this model could
+be used to examine the sensitivity of conservation and food security outcomes of
+MPAs to alternative design scenarios, holding constant other social-ecological
+variables. The *Spawning Ground* MPA network was able to provide more equitable
+yield outcomes across the two fishing communities, whereas the *Target Fishing*
+MPA network only benefited Fleet Two (@fig-coral-results A). This is because
+Fleet One has two primary fishing grounds; the spawning grounds, and the
+offshore area where the deep snapper live. Fleet Two primarily fishes in the
+north west corner of the simulation grid. Under the *Spawning Ground* strategy,
+while Fleet One quickly losses fishing grounds on the spawning grounds, it can
+compensate for this by moving offshore and fishing the deep snapper population
+harder, while also being able to fish some of the spillover of snapper from the
+spawning ground closure. This is why biomass of deep snapper actually declines
+under the *Spawning Ground* strategy.
+
+Conversely, most of the catch in the fishery comes from the nearshore and deep
+water snapper populations, which overlaps with both of Fleet One's fishing
+grounds. The *Target Fishing* strategy then begins closing both the nearshore
+and offshore snapper fishing grounds, resulting in too much of a loss in fishing
+grounds for Fleet One to be offset by spillover from the MPAs. As a result of
+primarily placing MPAs on Fleet One's fishing grounds, Fleet Two gains spillover
+benefits at little cost to their fishing grounds until the MPAs reach their
+fishing grounds once protection nears 100%. By protecting both the spawning
+ground and the offshore areas, the *Target Fishing* network provides
+conservation benefits to all of the species, in contrast to the *Spawning
+Ground* strategy that only provides meaningful conservation gains to the snapper
+population up until very large MPA sizes. Our result that protection of a
+dedicated spawning ground did not provide substantially greater conservation
+outcomes for the species using that spawning ground (groupers and sharks) than
+an alternative design strategy is supported by other modeling studies that show
+that the impacts of spawning ground protection on conservation and yields may be
+highly variable, and that displacement of high levels fishing effort from the
+spawning grounds can offset potential conservation gains of the protection
+[@grüss2014] .
+
+The two design strategies produced very different outcomes in terms of total
+changes in yield and conservation outcomes (@fig-coral-results C). Both MPA
+network design strategies were capable of producing net "win-win" outcomes in
+which both food security (as measured by change in total fisheries yields) and
+conservation (as measured by change in SSB/SSB0). However, in general for the
+same size MPA network the *Spawning Ground* design strategy produced better food
+security outcomes but worse conservation outcomes, and *vice* *versa* for the
+*Target Fishing* strategy. The coral reef case study shows how the modeling
+framework presented here can help stakeholders explore how different management
+strategies affect outcomes both in total (@fig-coral-results C) and across
+species (@fig-coral-results A) and fishing communities (@fig-coral-results B).
+
+Turning to our Pelagic Ecosystem case study, we see that overall the yield
+outcomes of the purse-seine fleet were much more sensitive to the presence of a
+range shift than the longline fleet, particularly under the *Target Fishing*
+design strategy. This is because the purse-seine fleet primarily targets
+skipjack tuna, which in this simulation are concentrated in a relatively narrow
+latitudinal band (@fig-blue), and the purse-seine fleet makes up a large portion
+of the total catch in the simulated fishery. So, the "Target Fishing" strategy
+starts by closing off the main purse-seine fishing grounds, which while only
+small part of the spatial domain of the model represents a large portion of the
+purse-seine fleet's fishing grounds, resulting both in more rapid conservation
+gains and fishery losses under the *Status Quo* species distributions. The
+*Rate* strategy places more MPAs in areas that are of lesser importance fo the
+purse-seine fleet but overlap more with species such as Oceanic Whitetip Shark.
+Conversely, the purse-seine fleet appears to do have better MPA yield outcomes
+under the Range Shift scenario, not because of rebuilding of the Skipjack
+population, but due to future fishing grounds being essentially unprotected by
+smaller MPAs targeting the current skipjack distribution ( @fig-blue-results ).
+
+The conservation outcomes of the skipjack, bigeye, and yellowfin tunas were
+among the most sensitive to range shifts, whereas even complete closure of the
+region was not sufficient to see significant increases in severely depleted
+species like oceanic whitetip sharks within the 20-year timeline of the
+simulation. Future research could be conducted then to see what sorts of
+timelines might be needed to see recovery of these species, and which design
+strategies result in the fastest recovery at the lowest cost to other objectives
+such as food security.
+
+These case studies are only intended to illustrate the capabilities of the model
+and the importance of considering the bio-economic dynamics represented in
+`marlin`; it is beyond the scope of this paper to provide broader conclusions
+about the performance of MPAs under different contexts or a comprehensive
+comparison of simulation and empirical results of MPAs around the world. That
+being said, the kinds of dynamics resulting in our case studies are well
+supported by both modeling and empirical studies. Our results support conceptual
+[@hilborn2004a; @gaines2010] and empirical [@ban2019] evidence that under the
+right conditions no-take MPAs can benefit fisheries and conservation.
+@rassweiler2014 demonstrated that the kinds of design choices presented in our
+case studies can greatly drive MPA outcomes. @ovando2021a provides empirical
+support for the possibility of even well enforced and large MPAs to have less
+pronounced population-level conservation effects (@fig-coral-results A) than
+might be expected given observed differences in biomass inside and outside of
+protected areas.
+
+MPAs based around coral-reef style ecosystems have been extensively studied
+around the world, with much of the empirical evidence of their performance
+centered on comparing measures such as biomass densities inside protected areas
+relative to outside [@lester2009]. Our model predicts a similar rapid increase
+of the simple ratio of mean biomass inside MPAs relative to outside (Fig.XX,
+acknowledging that designing a proper response ratio would require controlling
+for habitat characteristics and MPA design criteria). However, despite producing
+clear response ratios, The net conservation and fishery outcomes of our
+simulated coral-reef ecosystem MPAs were not nearly as large or clear as the
+response ratio results (@fig-coral-results); this result is supported works such
+as @ferraro2018 an @ovando2021a showing that response ratios alone may be a poor
+indicator of the net causal impacts of MPAs.
+
+Studies such as @gilman2020, @abbott2012, @davies2018 and @pons2022 provide
+empirical support for our case study results showing how MPAs can produce
+trade-offs across the conservation of different species and fleets when the
+spatial-temporal distributions and life history traits of affected species are
+heterogeneous, and MPAs result in some degree of effort displacement or
+concentration in the remaining fishing grounds. @hampton2023 supports our
+pelagic case study result showing that both conservation and fishery impacts of
+MPAs on the highly mobile species of the open oceans can be limited unless
+protected areas are very large. @brown2018 supports our result showing how
+heterogeneity in the behavior and objectives of fishing fleets sharing an
+ecosystem can affect the magnitude and equity of fishery reform outcomes.
+@davies2017 discusses the importance of considering the potential of climate
+driven range shifts when assessing spatial management policies; our work builds
+on this by allowing users to not just simulate species distributions but also
+distribution of biomass and age composition of fish space and time under climate
+change.
 
@@
 
+`marlin` does not represent trophic interactions. Anthropgenic changes in
+species abundance can clearly result in trophic cascades throughout the affected
+food web. However, studies such as @gilman2020, @ovando2021a, @bruno2019 , and
+@malakhoff2021 found no clear signs of MPA driven trophic cascades within the
+first decades of protection. Signals of trophic cascades may be masked by
+variations in the direction and strength of species interactions driven by
+environmental context [@liu2022], or may simply take longer to develop
+detectable effects than the coverage of many time series of MPAs. This is to say
+then that while `marlin` does not incorporate trophic interactions, what
+empirical evidence we have does not suggest that management-mediated trophic
+cascades are so common and clear that they must be incorporated into any
+credible multi-species simulation model. Rather, there is still value in
+capturing other dynamic bio-economic processes, as done here, even if that comes
+at the expense of modeling trophic interactions. However, research on the
+trophic impacts of spatial-temporal management actions is clearly of value and
+models like Atlantis [@audzijonyte2019], Ecopath with Ecosim [@christensen2004],
+and EASI-Fish [@griffiths2019] can help users explore those kinds of processes,
+in the manner of @baskett2007.
+
 ### General Recommendations for Using `marlin`
@@
 parameters that are highly uncertain and / or contentious which the users feel
-may impact results.
+may impact results. Where possible, parameters from locally estimated stock
+assessment models can be used to provide a foundation around which sensitivity
+analyses around specific uncertainties of interest [@berger2017].
 
@@
 the snapper and grouper species, rather than the dome-shaped selectivity
-presented in our main results. We found that the impacts of MPAs on conservation
-and food security were relatively insensitive to the form of fishing selectivity
-(Fig.S1). This does not mean that knowing the "correct" form of selectivity may
-not be extremely important in providing accurate assessment results
-[@waterhouse2014], but rather that holding other parameters constant, in this
-case misspecifying the selectivity curve is projected to have little impact on
-the simulated outcomes of MPAs.
+presented in our main results, reflecting a scenario where users are uncertain
+about the contact selectivity of Fleet Two. We found that the impacts of MPAs on
+conservation and food security were relatively insensitive to the form of
+contact selectivity specified for Fleet Two (Fig.S1). This does not mean that
+knowing the "correct" form of contact selectivity may not be extremely important
+in providing accurate assessment results [@waterhouse2014], but rather that in
+this case misspecifying the contact selectivity curve is projected to have
+little impact on the simulated outcomes of MPAs, conditional on holding other
+parameters constant.
 
@@
 complexity, such as predatory-prey dynamics (citations) or more complex fleet
-dynamics through agent-based approached [@burgess2020]. Particularly for more
+dynamics through agent-based approached [@burgess2020a]. Particularly for more
 tactical applications, we would encourage users to explore multiple modeling
@@
 
-In sum, `marlin` complements the existing spatial marine modeling literature by
-allowing scientists, decision makers, and other stakeholders to efficiently
-examine the impacts of realistic bio-economic dynamics on academic and applied
-problems. For example, scientists could use `marlin` to consider the relative
-importance of different biological variables in affecting policy outcomes.
-Managers could use `marlin` to assess the potential of proposed MPA networks to
-achieve stated objectives, such as maximizing conservation gains for a
-threatened species while minimizing or removing negative impacts to affected
-fishing fleets.
+At this time, `marlin` assumes that all evaluated policies are perfectly
+implemented; e.g. that MPAs are 100% no-take and perfectly enforced, that quotas
+and closed seasons are respected, that there is no discard mortality, etc. In
+reality no policy is perfectly implemented, and users should consider the extent
+to which the policies they simulate are actually feasible to implement. There is
+value though in being able to simulate and compare the outcomes of perfectly
+implmented policies to isolate the concept of the polucy itself from its
+implementation. That being said, future work may add the capability for
+imperfect implmentation to the model.
+
+`marlin` complements the existing spatial marine modeling literature by allowing
+scientists, decision makers, and other stakeholders to efficiently examine the
+impacts of realistic bio-economic dynamics on academic and applied problems. For
+example, scientists could use `marlin` to consider the relative importance of
+different biological variables in affecting policy outcomes. Managers could use
+`marlin` to assess the potential of proposed MPA networks to achieve stated
+objectives, such as maximizing conservation gains for a threatened species while
+minimizing or removing negative impacts to affected fishing fleets.
 
@@
 on dynamic ocean management, range shifts, management strategy evaluation,
-policy interactions, and spatial stock assessment. In addition, the
-process-based movement model used in `marlin` can directly use empirical
-estimates of movement dynamics derived in the manner of @thorson2021a, providing
-a link between empirical and simulation based approaches to marine resource
-management. The model currently focuses on diffusion and taxis, which allows for
-general representation of animals following physical or oceanographic features,
-but research on the incorporation and importance of advective forces would be of
-value going forward. Future iterations of the model may include life history
-modules for a broader range of species types such as marine mammals, sea birds,
-and sea turtles.
+policy interactions, and spatial stock assessment. `marlin` can help researchers
+generate data for further testing of the performance and design of spatially
+explicit integrated population models, in the manner of @bosley2022. In
+addition, the process-based movement model used in `marlin` can directly use
+empirical estimates of movement dynamics derived in the manner of @thorson2021a,
+providing a link between empirical and simulation based approaches to marine
+resource management that has been challenging to implement in spatial
+simulations [@berger2017]. The model currently focuses on diffusion and taxis,
+which allows for general representation of animals following physical or
+oceanographic features, but research on the incorporation and importance of
+advective forces would be of value going forward. Future iterations of the model
+may include life history modules for a broader range of species types such as
+marine mammals, sea birds, and sea turtles.
+
+Fisheries models, assessments, management have often abstracted away many of the
+spatial-temporal complexities of marine social-ecological systems [@ovando2021;
+@berger2017]. The modeling framework described here can help facilitate the
+science and application of spatial fisheries management by supporting the
+simulation of different spatial-temporal dynamics to aid in testing of various
+aspects of the marine resource management process.
 
@@
 
-This project was funded by the Waitt Foundation. We thank Dr. Alexa Fredston for
-helpful revisions of this manuscript.
+This project was funded by the Waitt Foundation. We thank Dr. Alexa Fredston and
+two anonymous reviewers for helpful revisions and comments of this manuscript.
 

### 2023-10-07 04_marlin_paper_si.qmd
almost there

---
title: "Simulating Benefits, Costs, and Tradeoffs of Spatial Management in Marine Social-Ecological Systems"
subtitle: "Supporting Information"
format: 
  pdf:
    toc: false
    fig-pos: 'H'
    mainfont: Times New Roman
    sansfont: Times New Roman
    fontsize: 12pt
    keep-tex: true
    fig-format: png
  html: default
  docx: default
execute:
  echo: false
  message: false
  warning: false
  cache: false
csl: apa-6th-edition.csl
bibliography: references.bib
params:
  run_name: ["v1.1"]
header-includes:
  - \usepackage{setspace}\doublespacing
  - \usepackage{lineno}\linenumbers
  - \usepackage{bm}
editor: 
  markdown: 
    wrap: 80
---

```{r setup}
#| echo: false
#| message: false
#| include: false

source(here::here("00_setup.R"))

theme_set(theme_classic(base_size = 10, base_family = "Helvetica") + theme(strip.background = element_blank()))

if (!file.exists(here("data","sci_to_com.csv"))){
  sci_to_com <- taxize::sci2comm(unique(marlin_inputs$scientific_name),simplify= TRUE,db='ncbi') %>% 
  bind_rows(.id = "critter") %>% 
  pivot_longer(everything(),names_to = "critter",values_to = "common_name") %>% 
  mutate(common_name = stringr::str_to_title(common_name))

  write_csv(sci_to_com, file = here("data","sci_to_com.csv"))

} else{
 sci_to_com <-  read_csv(file = here("data","sci_to_com.csv"))

}

print(results_path)
results_files <- list.files(file.path(results_path))

rdses <- results_files[str_detect(results_files,'.rds')]

walk(rdses,~assign(str_remove(.x,".rds"), read_rds(file.path(results_path,.x)), envir = .GlobalEnv)) #load and assign all .rds objects in the appropriate results folder

coral_mpa_experiments$mpas <- map(coral_mpa_experiments$results, "mpa")

coral_mpa_experiments$obj <- map(coral_mpa_experiments$results, "obj")

```

\renewcommand{\thefigure}{S\arabic{figure}}

\renewcommand{\thetable}{S\arabic{table}}

Author Names: Daniel Ovando^1,2^, Darcy Bradley^3,4,5^, Echelle
Burns^3,4,5^,Lennon Thomas^3,4,5^, James Thorson^6^

Affiliations: ^1^ School of Aquatic and Fishery Sciences, University of
Washington,1122 NE Boat St Box 355020,Seattle, WA, USA ^3^ Marine Science
Institute, University of California, Santa Barbara, CA, USA ^4^ Bren School of
Environmental Science & Management, University of California, Santa Barbara, CA,
USA ^5^ Environmental Markets Lab, University of California, Santa Barbara, CA,
USA ^6^ Habitat and Ecological Processes Research Program, Alaska Fisheries
Science Center, NOAA, NMFS, NOAA, Seattle, WA, USA

```{r}
#| label: fig-coral-fleet
#| fig-cap: "Distribution of annual catches by fishing fleet prior to MPAs under the coral reef case study simulation."
coral_fauna <- coral_fauna_and_fleets$fauna

coral_fleets <- coral_fauna_and_fleets$fleets

resolution <- sqrt((coral_fauna[[1]]$patches))

grid <- expand_grid(x = 1:resolution, y= 1:resolution) %>%
  mutate(patch = 1:nrow(.))

mpa_locations <- expand_grid(x = 1:resolution, y = 1:resolution) %>%
  mutate(mpa = FALSE)

coral_sim <- simmar(
  fauna = coral_fauna,
  fleets = coral_fleets,
  manager = list(mpas = list(
    locations = mpa_locations,
    mpa_year = 1
  )),
  years = 20
)

pcsim <- process_marlin(coral_sim, keep_age = FALSE)

pcsim$fleets |>
  filter(year == max(year)) |>
  group_by(fleet, x, y) |>
  summarise(catch = sum(catch)) |>
  group_by(fleet) |>
  mutate(catch = catch / sum(catch)) |>
  ggplot(aes(x, y, fill = catch)) +
  geom_tile() +
  facet_wrap( ~ fleet) +
  scale_fill_viridis_c(name = "% of Annual Catch", labels = scales::percent) +
  scale_x_continuous(name = "Longitude", expand = expansion()) +
  scale_y_continuous(name = "Latitude", expand = expansion()) +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    strip.text.y = element_text(size = 8)
  )

```

```{r}
#| label: fig-coral-mpas
#| fig-cap: "Location of MPAs under the *Spawning Ground* and *Target Fishing* placement strategies for the coral reef case study simulations. Columns indicate whether MPA was designed around location of spawning grounds, or locations where the most biomass of fish was caught. Rows indicate the percent of the simulation grid placed in a no-take MPA. MPA sizes between 10% and 90% plotted."
mpa_trajectories <- coral_mpa_experiments |>
  select(placement_strategy, prop_mpa, iter, mpas) |>
  unnest(cols = mpas)

mpa_trajectories |>
  filter(prop_mpa %in% seq(0.1,0.9, by = .1)) |> 
  mutate(p_prop_mpa = glue::glue("{prop_mpa * 100}%")) |> 
  mutate(p_prop_mpa = reorder(p_prop_mpa, prop_mpa)) |> 
  ggplot(aes(x,y,fill = mpa)) +
  geom_tile() +
  facet_grid(p_prop_mpa~placement_strategy) + 
  scale_x_continuous(name = "Longitude",expand = expansion()) + 
  scale_y_continuous(name = "Latitude",expand = expansion()) +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        strip.text.y = element_text(size = 8)) + 
  scale_fill_discrete(name = "", labels = c("Fished",'MPA'))

```

```{r}
#| label: fig-select-biodiv
#| fig-cap: "Change in SSB/SSB0 of simulated coral reef species relative to pre-MPA levels under alternative contact selectivity assumptions. 'dome' indicates that the contact selectivity of Fleet Two is assumed to be dome shaped for the snapper and grouper species (the default value presented in the main results). Logistic indicates that the contact selectivity of Fleet Two is assumed to be logistic (asymptotic) for all species."
sel_experiments |>
  group_by(critter, experiment, placement_strategy) |>
  mutate(biodiv = biodiv / biodiv[prop_mpa == 0] -1) |>
  ggplot(aes(prop_mpa, 100 * pmin(2,biodiv), color = experiment)) +
  geom_hline(aes(yintercept = 0), linetype = 2) +
  geom_line() +
  facet_grid(placement_strategy ~ critter) +
  scale_y_continuous(name = "% Change from Baseline SSB/SSB0",
                     breaks = c(-25,seq(0,200, by = 50)),
                     labels = c(as.character(c(-25,seq(0,150, by = 50))),expression("">=200))) + 
  scale_color_discrete("Fleet Two Selectivity ") + 
  scale_x_continuous(name = "MPA Size", labels = scales::percent)

```

```{r}
#| label: fig-select-yield
#| fig-cap: "Change in total yield (biomass of fish caught) of simulated coral reef fishing fleets relative to pre-MPA levels under alternative contact selectivity assumptions. 'dome' indicates that the contact selectivity of Fleet Two is assumed to be dome shaped for the snapper and grouper species (the default value presented in the main results). Logistic indicates that the contact selectivity of Fleet Two is assumed to be logistic (asymptotic) for all species."
sel_experiments |>
  select(critter, prop_mpa, placement_strategy, starts_with("fleet_"), experiment) |> 
  pivot_longer(starts_with("fleet_"), names_to = "fleet", values_to = "catch") |> 
  group_by(prop_mpa, placement_strategy, fleet, experiment) |> 
  summarise(catch = sum(catch)) |> 
  ungroup() |> 
  group_by(fleet, experiment, placement_strategy) |>
  mutate(catch = catch / catch[prop_mpa == 0] -1) |>
  ggplot(aes(prop_mpa, 100 * pmin(2,catch), color = experiment)) +
  geom_hline(aes(yintercept = 0), linetype = 2) +
  geom_line() +
  facet_grid(fleet ~ placement_strategy ) +
  scale_y_continuous(name = "% Change from Baseline Yields") + 
  scale_color_discrete("Fleet Two Selectivity ") + 
  scale_x_continuous(name = "MPA Size", labels = scales::percent)

```

```{r}
#| label: fig-rr
#| fig-cap: "Ratio of mean biomass inside MPAs relative to mean biomass outside MPAs as a function of MPA size. Only *Target Fishing* strategy is shown as biomass is roughly perfectly confounded with MPA location in the *Spawning Ground* design strategy."

coral_mpa_experiments$response_ratios <- map(coral_mpa_experiments$results, "response_ratio")

rrs <- coral_mpa_experiments |>
  select(placement_strategy, prop_mpa, response_ratios) |>
  unnest(cols = response_ratios)

rrs |>
  filter(placement_strategy == "target_fishing", prop_mpa > 0, prop_mpa < 1) |>
  group_by(placement_strategy, prop_mpa) |>
  mutate(rr = biomass[mpa == TRUE] / biomass[mpa == FALSE]) |>
  ungroup() |>
  ggplot(aes(prop_mpa, rr, color = mpa)) +
  geom_line() +
  facet_wrap(~placement_strategy) +
  scale_x_continuous(name = "MPA",labels = scales::label_percent(accuracy = 1), breaks = seq(0,1, by = .1)) +
  scale_y_continuous(limits = c(0, NA), name = "Ratio of biomass inside MPA relative to Outside")

```

### 2023-10-07 01_coral_case_study.R

@@ -19,7 +19,7 @@ seasons <- 4
 
 tune_type <- "depletion"
 
-experiment_workers <- 8
+experiment_workers <- parallel::detectCores() - 2
 
 years <- 100
 
@@ -402,6 +402,13 @@ plrs <- process_marlin(logistic_reef_sim, keep_age = FALSE)
 
 plot_marlin(prs = prs, plrs = plrs, plot_var = "ssb")
 
+prs$fauna |>
+  filter(year == max(year)) |>
+  group_by(critter) |>
+  summarise(b = sum(b)) |>
+  ggplot(aes(reorder(critter, b),b)) +
+  geom_col()
+
 prs$fleets |>
   filter(step == max(step)) |>
   group_by(fleet) |>
@@ -409,6 +416,14 @@ prs$fleets |>
   ggplot(aes(x,y,fill = effort)) +
   geom_tile() +
   facet_wrap(~fleet)
+
+prs$fleets |>
+  filter(step == max(step)) |>
+  group_by(x,y) |>
+  summarise(catch = sum(catch)) |>
+  ggplot(aes(x,y,fill = catch)) +
+  geom_tile()
+
 
 grid <- expand_grid(x = 1:resolution, y = 1:resolution) %>%
   mutate(patch = 1:nrow(.))
@@ -667,6 +682,32 @@ coral_mpa_experiments$mpas <- map(coral_mpa_experiments$results, "mpa")
 
 coral_mpa_experiments$obj <- map(coral_mpa_experiments$results, "obj")
 
+coral_mpa_experiments$response_ratios <- map(coral_mpa_experiments$results, "response_ratio")
+
+rrs <- coral_mpa_experiments |>
+  select(placement_strategy, prop_mpa, response_ratios) |>
+  unnest(cols = response_ratios)
+
+rrs |>
+  filter(placement_strategy == "target_fishing", prop_mpa > 0, prop_mpa < 1) |>
+  group_by(placement_strategy, prop_mpa) |>
+  mutate(rr = biomass[mpa == TRUE] / biomass[mpa == FALSE]) |>
+  ungroup() |>
+  ggplot(aes(prop_mpa, rr, color = mpa)) +
+  geom_line() +
+  facet_wrap(~placement_strategy) +
+  scale_x_continuous(labels = scales::label_percent(accuracy = 1), breaks = seq(0,1, by = .1)) +
[…]

### 2023-10-07 03_marlin_paper.qmd
R1 submitted

@@
 
-Title: Simulating Benefits, Costs, and Tradeoffs of Spatial Management in Marine
-Social-Ecological Systems
+Title: Simulating Benefits, Costs, and Trade-offs of Spatial Management in
+Marine Social-Ecological Systems
 
@@
 fishing fleets respond endogenously to ecosystem changes in ways that depend
-upon ecological and policy context. We present a tool, `marlin`, that can be
-used to efficiently simulate the bio-economic dynamics of marine systems in
-support of both management and research. We demonstrate `marlin’s` capabilities
-by focusing on two case studies on the conservation and food production impacts
-of marine protected areas (MPAs): a coastal coral reef and a pelagic tuna
-fishery. In the coastal coral reef example, we explore how heterogeneity in
-species distributions and fleet preferences can affect distributional outcomes
-of MPAs. In the pelagic case study, we show how our model can be used to assess
-the climate resilience of different MPA design strategies, as well as the
-climate sensitivity of different fishing fleets. This paper demonstrates how
-intermediate complexity simulation of coupled bio-economic dynamics can help
-communities predict and potentially manage trade-offs between conservation,
-fisheries yields, and distributional outcomes of management policies affected by
-spatial bio-economic dynamics.
+upon ecological and policy context. We present a modeling framework, `marlin`,
+that can be used to efficiently simulate the bio-economic dynamics of marine
+systems in support of both management and research. We demonstrate `marlin’s`
+capabilities by focusing on two case studies on the conservation and food
+production impacts of marine protected areas (MPAs): a coastal coral reef and a
+pelagic tuna fishery. In the coastal coral reef example, we explore how
+heterogeneity in species distributions and fleet preferences can affect
+distributional outcomes of MPAs. In the pelagic case study, we show how our
+model can be used to assess the climate resilience of different MPA design
+strategies, as well as the climate sensitivity of different fishing fleets. This
+paper demonstrates how intermediate complexity simulation of coupled
+bio-economic dynamics can help communities predict and potentially manage
+trade-offs between conservation, fisheries yields, and distributional outcomes
+of management policies affected by spatial bio-economic dynamics.
 
@@
 4.  Discussion
+    1.  Insights from Case Studies
+    2.  Putting `marlin` in Context
+    3.  General Recommendations for Use
 5.  Acknowledgements
@@
 across 30% of the world's oceans by the year 2030 [@grorud-colvert2021 and
-references therein]. However , designing spatial management strategies to
-achieve desired objectives -- which may include recovery and resilience of
-overfished species, increased food production and economic well being, and
-equitable distribution of benefits - is not straightforward. Different species
-have different resilience to fishing, distributions in space and time, and value
-to fishing fleets; fishing fleets themselves can have varying ranges of
-incentives and reactions to policy structures. These dynamics pose challenges
-even for single-species assessment and management strategies , which are only
-amplified when we consider management policies specifically designed around
-multiple species, fleets, and spatial features in the oceans [@field2006a].
+references therein]. Recent agreements such as the United Nations (UN) agreement
+on the conservation and sustainable use of marine biodiversity of areas beyond
+national jurisdiction (BBNJ) calls for expanded area-based management in the
+high seas. However, designing spatial management strategies to achieve desired
+objectives -- which may include recovery and resilience of overfished species,
+increased food production and economic well being, and equitable distribution of
+benefits - is not straightforward. Different species have different resilience
+to fishing, distributions in space and time, and value to fishing fleets;
+fishing fleets themselves can have varying ranges of incentives and reactions to
+policy structures. These dynamics pose challenges even for single-species
+assessment and management strategies , which are only amplified when we consider
+management policies specifically designed around multiple species, fleets, and
+spatial features in the oceans [@field2006a].
 
@@
 system. Our results show how interacting ecological, economic, and design
-attributes can affect the degree of benefits, costs, and trade offs between
-conservation and food production outcomes of MPAs. More broadly, this paper both
+attributes can affect the degree of benefits, costs, and trade-offs between
+conservation and food production outcomes of MPAs. More broadly, this paper
 demonstrates the functionality of our model and the critical need for
@@
 
-`marlin` follows a "models of intermediate complexity" philosophy
-[@plagányi2014] targeted specifically towards capturing heterogeneous habitats,
-movement dynamics, and the simultaneous impacts of fishing fleets across
-multiple species (frequently called "technical interactions"). In its general
-form, `marlin` simulates the behavior of populations of spatially explicit
-biologically independent age-structured animal populations affected by one or
-more fishing fleets in time steps and a spatial resolution specified by the
-user.
+The model presented here focuses on representing the dynamics of heterogeneous
+habitats and movement dynamics, along with the simultaneous impacts of fishing
+fleets across multiple species (frequently called "technical interactions"). In
+its general form, `marlin` simulates the behavior of populations of spatially
+explicit biologically independent animal populations with age and subsequent
+size structure affected by one or more fishing fleets in time steps and a
+spatial resolution specified by the user.
 
@@
 the ratio of the *spawning stock biomass* (SSB) in a given time step relative to
-*unfished spawning stock biomass* (SSB0), SSB/SSB0. So, a SSB/SSB0 value of 1
-means that the population is unfished, a SSB/SSB0 value of 0 means the
-population is extinct.
+*unfished spawning stock biomass* (SSB0), SSB/SSB0. An SSB/SSB0 value of 1 means
+that the population is unfished, a SSB/SSB0 value of 0 means the population is
+extinct.
 
@@
 1.  Trade-offs and distributional outcomes for food production and multi-species
-    biomass in a coastal coral-reef fishery
+    biomass in a coastal coral reef fishery
 2.  Implications of climate-driven range shifts for MPA design in a pelagic
@@
 be made much more likely to catch larger bigeye than skipjack, and *vice versa*.
-Each fishing fleet distributes its fishing effort in space according to a
-specified spatial allocation rule (see *Spatial Allocation of Effort* ), for
-example in proportion to profit-per-unit-effort, conditional on management
-policies in places such as quotas and/or the presence of any spatial
-restrictions such as MPAs. It is important to note that contact selectivity
-reflects the ability of the fishing method in question to capture fish of
-different sizes that come into "contact" with the gear. The contact selectivity
-of each of the specified *metiers* will then interact with the distribution of
-fish sizes and fishing effort in space simulated by `marlin` to produce a net
-"population" selectivity, which may differ from the individual contact
-selectivities of each of the *metiers*.
+It is important to note that contact selectivity reflects the ability of the
+fishing method in question to capture fish of different sizes that come into
+"contact" with the gear. The contact selectivity of each of the specified
+*metiers* will then interact with the distribution of fish sizes and fishing
+effort in space simulated by `marlin` to produce a net "population" selectivity,
+which may differ from the individual contact selectivities of each of the
+*metiers* [@waterhouse2014; @sampson2014]. Each fishing fleet distributes its
+fishing effort in space according to a specified spatial allocation rule (see
+*Spatial Allocation of Effort* section), for example in proportion to
+profit-per-unit-effort, conditional on management policies in places such as
+quotas and/or the presence of any spatial restrictions such as MPAs.
 
@@
 must set model parameters themselves based on externally available data and
-their best judgement. On the biological side, at a minimum users must for each
-species they wish to simulate supply the common or scientific name of the
-species in question, a measure of the level of fishing intensity the species is
-experiencing at the start of the simulation, and the diffusion rates for adult
-and larvae. Given a common or scientific name, the model will then input default
-life history parameters based on FishLife [internet connection required,
-@thorson2020], though users are encouraged to check these values and supply
-their own life history parameter values when possible. On the fleet side, users
-must for each fleet they wish to simulate at a minimum specify the selectivity
-curves for each species caught by the fleet, as well relative price per unit
-weights greater than 0 for any target species. Species caught by but not
-targeted by the fleet can be represented by prices of 0 (not targeted) or below
-0 (actively avoided). We provide two case studies here designed to show users
-how the `marlin` model can be used to address the kinds of questions
-stakeholders might face if designing an MPA network for a coastal coral reef or
-a pelagic system.
+their best judgement.
+
+On the biological side, at a minimum users must for each species they wish to
+simulate supply the common or scientific name of the species in question, a
+measure of the level of fishing intensity the species is experiencing at the
+start of the simulation, and the diffusion rates for adult and larvae. Given a
+common or scientific name, the model will then input default life history
+parameters based on FishLife [internet connection required, @thorson2020],
+though users are encouraged to check these values and supply their own life
+history parameter values when possible.
+
+On the fleet side, users must for each fleet they wish to simulate at a minimum
+specify the contact selectivity curves for each species caught by the fleet, as
+well relative price per unit weights greater than 0 for any target species.
+Species caught by but not targeted by the fleet can be represented by prices of
+0 (not targeted) or below 0 (actively avoided).
 
@@
 each species caught and its price ($\Pi$). The amount caught is a function of
-the selectivity at age of each species for each fleet ($\alpha$), the fishing
-efficiency of the fleet for that that species (also called "catchability", *q*),
-the amount of fishing effort of the fleet in question in that patch in a given
-time step (*E*), and the total instantaneous fishing mortality (*u*) at age
-(*a*) (including all other fleets) for that particular species in that patch and
-time step.
+the contact selectivity at age of each species for each fleet ($\alpha$), the
+fishing efficiency of the fleet for that that species (also called
+"catchability", *q*), the amount of fishing effort of the fleet in question in
+that patch in a given time step (*E*), and the total instantaneous fishing
+mortality (*u*) at age (*a*) (including all other fleets) for that particular
+species in that patch and time step.
 
@@
 Total catches or yield *Y* are calculated through the Baranov equation
-[@baranov1918], which appropriately accounts for the total instantaneous
-mortality (both fishing and natural. *z*) and divides the total amount of the
-population biomass (*b*) killed between natural (*m*) and fishing (*u*) sources,
-with the amount of biomass killed through fishing called "yield".
+[@baranov1918], which accounts for the total instantaneous mortality (both
+fishing and natural, *z*) and divides the total amount of the population biomass
+(*b*) killed between natural (*m*) and fishing (*u*) sources, with the amount of
+biomass killed through fishing called "yield".
 
@@
 
-Selectivity at age $\alpha$ is modeled through either a logistic form or a
-dome-shaped form. The logistic form is based on the lengths *l* at which 50% of
-individuals are selected by the fishing gear, $l^{sel}$, and $\delta$ which is
-the difference between the length at 50% selectivity and the length at 95%
+Contact selectivity at age $\alpha$ is modeled through either a logistic form or
+a dome-shaped form. The logistic form is based on the lengths *l* at which 50%
+of individuals are selected by the fishing gear, $l^{sel}$, and $\delta$ which
+is the difference between the length at 50% selectivity and the length at 95%
 selectivity.
@@
 
-Lastly *z* is total mortality, the sum of fishing mortality and natural
+Lastly *z* is total mortality, the sum of fishing mortality (*u*) and natural
 mortality (*m*).
@@
 species is a function of a range of parameters including the total amount of
-effort (*E*), the selectivity ogives ($\alpha$,the curve specifying the
+effort (*E*), the contact selectivity ogives ($\alpha$,the curve specifying the
 proportion of individuals at a given age or size that will be caught by a unit
@@
 this model, assuming that advection is zero, though future extensions could
-easily incorporate advection vectors from oceanographic models. In this way,
-`marlin` allows users to run anything from a simple Gaussian dispersal kernel up
-to a system governed by species that passively diffuse out from a core habitat
-defined by a dynamic thermal range.
+incorporate advection vectors from oceanographic models. In this way, `marlin`
+allows users to run anything from a simple Gaussian dispersal kernel up to a
+system governed by species that passively diffuse out from a core habitat
+defined by a dynamic thermal range. The model currently focuses on diffusion and
+taxis, which allows for general representation of animals following physical or
+oceanographic features, but research on the incorporation and importance of
+advective forces would be of value going forward.
 
@@
 set, regardless of the time step or patch size used in the simulation model
-itself. The taxis component of the movement process is a function of the
-difference in habitat quality *H*. The habitat preference function itself can
-take any form the user wishes. Exponentiating the difference in the habitat
-preference function between patches turns the taxis matrix into a multiplier of
-the diffusion rate *D*. As such, when creating habitat layers for simulation,
-users can tune the scale of the habitat gradient function to result in realistic
-multipliers of the diffusion rate. This parameterization ensure that the
+itself.
+
+The taxis component of the movement process is a function of the difference in
+habitat quality *H*. The habitat preference function itself can take any form
+the user wishes. Exponentiating the difference in the habitat preference
+function between patches turns the taxis matrix into a multiplier of the
+diffusion rate *D*. As such, when creating habitat layers for simulation, users
+can tune the scale of the habitat gradient function to result in realistic
+multipliers of the diffusion rate. This parameterization ensures that the
 off-diagonal elements of the movement matrix $\pmb{M}_{t,s,a}$ are all
@@
 
-Spawning stock biomass *SSB* is calculated by converting age to mean length,
-calculating weight at age, maturity at age, and then calculating spawning stock
-biomass as the sum of spawning potential at age in a given time step, taking
-into account the potential for hyperallometry in the manner of @marshall2021.
-Age is converted to length through the von Bertalanffy growth equation given
-parameters asymptotic length ($l_\infty$), growth (*k*) and theoretical age at
-length zero ($a0$) assuming log-normally distributed variation *u* in the length
-at age with CV $\sigma_s$.
+Spawning stock biomass *SSB* is calculated by converting age to mean length at
+age, calculating weight at age, maturity at age, and then calculating spawning
+stock biomass as the sum of spawning potential at age in a given time step,
+taking into account the potential for hyperallometry in the manner of
+@marshall2021. Age is converted to length through the von Bertalanffy growth
+equation given parameters asymptotic length ($l_\infty$), growth (*k*) and
+theoretical age at length zero ($a0$) assuming log-normally distributed
+variation *u* in the length at age with CV $\sigma_s$.
 
@@
 dynamics at a given moment, making the definition of an equilibrium concept such
-as MSY challenging.
+as MSY challenging [@berger2017 and references therein].
 
@@
 with the exception of the spawning stock biomass at MSY ($SSB_{MSY}$) which
-tends to be conserved across selectivity patterns as it is primarily a function
-of the mortality and flow of recruits relative to the maturity ogive. There are
-many different strategies for estimating reference points in spatially explicit
-systems [@kapur2021a]. We leave it to users to define and find relevant
+tends to be better conserved across selectivity patterns as it is primarily a
+function of the mortality and flow of recruits relative to the maturity ogive.
+There are many different strategies for estimating reference points in spatially
+explicit systems [@kapur2021a]. We leave it to users to define and find relevant
 reference point as required by their specific needs.
-
-<!-- ### MPA Optimization -->
-
-<!-- Our MPA optimization algorithm seems to create an MPA network that -->
-
-<!-- maximizes an objective function, given by -->
-
-<!-- $$ -->
-
-<!-- obj_p = \Phi \times scale(\sum_f^F \pmb{\phi_{f,p}}) + (1 - \Phi)  scale(\pmb{\sum_s^S\frac{ssb_s}{ssb0_{s,p}}}) -->
-
-<!-- $$ {#eq-obj} -->
-
-<!-- Where *scale* rescales the values to be between 0 and 1 such that the -->
-
-<!-- weight parameter is comparable between vectors. The algorithm works by -->
-
-<!-- first assigning every patch in the system to an MPA. For the first -->
-
-<!-- round, the algorithm then calculates the equilibrium objective function -->
-
-<!-- value resulting from sequentially removing each individual patch from -->
-
-<!-- the MPA network. The cell whose removal from the MPA produces the -->
-
-<!-- highest objective function value is then removed from the MPA network, -->
-
-<!-- and the marginal value of the objective function each of the remaining -->
-
-<!-- cells from the last steps are then stored. For the next steps, we -->
-
-<!-- sampled 25% of the remaining MPA cells with probability proportional to -->
-
-<!-- their marginal contribution to the objective function in the prior -->
-
-<!-- iteration, selecting the cell whose removal produces the highest total -->
-
-<!-- objective function as the cell to remove from the network, and updating -->
-
-<!-- the marginal value of the sampled-but-not-selected cells. This in -->
-
-<!-- essence is a sampling-importance-resampling algorithm. The algorithm -->
-
-<!-- continues until no more cells are left in the MPA. This then provides -->
-
-<!-- for a given objective weighting value $\Phi$ the near-optimal MPA -->
-
-<!-- network at every MPA size from 0 to 100%. We say "near" as they is some -->
-
-<!-- path dependency here, in that for example the network that maximizes the -->
-
-<!-- objective function with 40% of the MPAs does not necessarily contain the -->
-
-<!-- optimal network at 30%. However, the reversed nature of the algorithm, -->
-
-<!-- starting at 100% MPA coverage and working better, produces a more stable -->
-
-<!-- network with regards to this problem than starting at 0% and growing to -->
-
-<!-- 100%. -->
 
@@
 illustrative purposes only, readers should consult the accompanying code to view
-the precise details of each simulation. Targeted tactical applications must
-carefully consider and document all decisions made around model parameters.
+the precise details of each simulation. Targeted applications must carefully
+consider and document all decisions made around model parameters.
 
@@
 
-We make use of three potential "rule of thumb" MPA design strategies:
+We make use of three potential "rule of thumb" MPA design strategies in our case
+studies
 
@@
 3.  *Spawning Ground*: MPAs are centered on the grounds of a known spawning
-    aggregation
+    aggregation. This strategy is only used in the coral reef case study.
 
@@
 Pacific species: a grouper (*Epinephelus fuscoguttatus*), a shallow-reef snapper
-(*Lutjanus malabaricus*), a deep-water snapper (*Pristipomoides filamentosus*),
+(*Lutjanus malabaricus*), a deep-reef snapper (*Pristipomoides filamentosus*),
 and a reef shark (*Carcharhinus amblyrhynchos*). The simulated groupers undergo
@@
 captured sharks. Fleet One has a logistic selectivity pattern for all species,
-as they retain any fish caught for consumption or sale. both grouper and
-shallow-reef snapper due to demand for "plate sized" fish , and logistic
-selectivity for sharks, as any shark caught is retained. Fleet One is totally
+as they retain any fish caught for consumption or sale. Fleet One is totally
 dependent on fishing for their livelihood, meaning the local community takes
@@
 higher cost per distance coefficient than Fleet Two. Fleet One's home port is
-located on the site of the grouper spawning grounds.
-
-Fleet Two is a more commercial fleet that primarily targets the deep-water
-snapper population. This fleet primarily sells their catch to local restaurants
-and distributors where plate-sized fish are prized, and so for both snapper and
+located on near the site of the grouper spawning grounds.
+
+Fleet Two is a more commercial fleet that primarily targets the snapper
+populations. This fleet primarily sells their catch to local restaurants and
+distributors where plate-sized fish are prized, and so for both snapper and
 grouper Fleet Two has a dome-shaped selectivity pattern [@kindsvater2017]. While
@@
 expands and contracts in response to profitability of the primarily
-grouper-driven fishery. Fleet two's home port is located further north from
-Fleet One.
-
-We use `marlin` to simulate the outcomes for both food production and
+grouper-driven fishery. Fleet two's home port is located in the northwest corner
+of the simulation space.
+
+We used `marlin` to simulate the outcomes for both food production and
 conservation for each of the species and both of the fleets as a function of
@@
 selected MPA design can be as patchwork as dictated by the data). We ran the
-simulation for in quarter year time steps for 100 years ($\Delta_t = 1/4$) and
+simulation for in quarter year time steps for 20 years ($\Delta_t = 1/4$) and
 set the area of each patch to be 5km^2^ ($\Delta_p = \sqrt{5}$), using 144
@@
 
-#### Pelagic Ecosystems
+#### Pelagic Fishery
 
@@
 
-We model our pelagic use case loosely on the characteristics of the Western and
-Central Pacific Ocean (WCPO) tuna fisheries. We simulate trajectories of
+We model our pelagic case study loosely on the characteristics of the Western
+and Central Pacific Ocean (WCPO) tuna fisheries. We simulate trajectories of
 `r n_distinct(names(blue_fauna))` species commonly caught in the region,
@@
 purse seine fishery that primarily targets skipjack tunas for bulk canning.
-Selectivities were modeled as logistic for the longline fleet, and dome-shaped
-for the purse seine fleet, with selectivity values set to broadly match the
-patterns reported in relevant [stock
+Contact selectivities were modeled as logistic for the longline fleet, and
+dome-shaped for the purse seine fleet, with selectivity values set to broadly
+match the patterns reported in relevant [stock
 assessments](https://oceanfish.spc.int/en/ofpsection/sam/sam/210-stock-assessment-input-and-result-files-)
@@
 For this exercise, we focused on using `marlin` to assess resilience of the
-selected Target Fishing and Rate "rule of thumb" MPA design strategies to a
-climate-driven range shift. Specifically, we simulate an extreme example where
-the centroid of each population shifts northward at a rate of 0.33 patches per
-year, over a 20 year time horizon (@fig-blue-ssb). We designed MPA networks
-given the conditions in the starting year, and then held that network constant
-over the years of the experiment, running one simulation with and another
-without the climate-drive range shift. We then compared the effects of this
-range shift on food production and conservation outcomes from MPA networks
-designed based on the pre-range shift world. We ran the pelagic simulation at a
-quarterly level ($\Delta_t$ = 1/4 year), and set the area of each cell to be
-roughly 97,000 KM^2^ across 144 patches each with a side length of roughly 311
-KM, for a total area of 14,000,000 KM^2^, broadly commensurate with the area of
-the WCPO.
+selected *Target Fishing* and *Rate* MPA design strategies to a climate-driven
+range shift. Specifically, we simulate an extreme example where the centroid of
+each population shifts northward at a rate of 0.2 patches per year over a 20
+year time horizon (@fig-blue-ssb). We designed MPA networks given the conditions
+in the starting year, and then held that network constant over the years of the
+experiment, running one simulation with and another without the climate-drive
+range shift. We then compared the effects of this range shift on food production
+and conservation outcomes from MPA networks designed based on the pre-range
+shift world. We ran the pelagic simulation at a quarterly level ($\Delta_t$ =
+1/4 year), and set the area of each cell to be roughly 97,000 KM^2^ across 144
+patches each with a side length of roughly 311 KM, for a total area of
+14,000,000 KM^2^, broadly commensurate with the area of the WCPO.
 
@@
 
-### Coral Reefs
+### Coastal Coral Reef Fishery
 
@@
 
-peak_coral_fleet_results |> filter(fleet == "fleet_two")
-
 tmp <- coral_results %>% 
@@
 MPAs were capable of producing a range of positive and negative outcomes for
-food security and conservation in the Coral Reef case study depending on the
+food security and conservation in the coral reef case study depending on the
 design strategy used and the size of MPA implemented. Both of the MPA design
-strategies were capable of increasing fisheries yield for Fleet Two, up to a
-value of
+strategies were capable of increasing fisheries yield for the more "industrial"
+Fleet Two, up to a value of
 `r peak_coral_fleet_results$peak_yields[peak_coral_fleet_results$fleet == "fleet_two"]`,
-even when MPAs covered more than 50% of the simulated area. However, Fleet One
-only benefited from MPAs under the *Spawning Ground* design strategy, with a
-maximum increase of
+even when MPAs covered more than 50% of the simulated area. However, the more
+"artisanal" Fleet One only benefited from MPAs under the *Spawning Ground*
+design strategy, with a maximum increase of
 `r peak_coral_fleet_results$peak_yields[peak_coral_fleet_results$fleet == "fleet_one"]`;
@@
 biomass of all species under the *Target Fishing* design strategy, with the most
-rapid increases for the Deep Snapper population. The *Spawning Ground* strategy
-primarily benefited the coastal Snapper population, producing little change in
-the Grouper population and decreasing spawning biomass of both the Deep Snapper
-and Reef Shark populations for MPA sizes covering less than 50% of the simulated
-area (@fig-coral-results B).
+rapid increases in spawning biomass for the deep snapper population. The
+*Spawning Ground* strategy primarily benefited the shallow snapper population,
+producing little change in the grouper population and decreasing spawning
+biomass of both the deep snapper and reef shark populations for MPA sizes
+covering less than 50% of the simulated area (@fig-coral-results B).
 
@@
 Percent change in total yield across both fleets (y-axis) and total SSB/SSB0 across all species (x-axis) and placement strategies (line color) (C).
-Text labels in panel C indicate the percent of the simulation area in an MPA when change in yields is maximized, is 0, and is -50%.
+Color of points along each line in panel C indicates the percent of the simulation area in an MPA.
 All results reflect outcomes after 20 years of simulated MPA protection."
@@
 
-### Pelagic Ecosystems
-
-The yield impacts of the proposed MPA strategies varied across each of our
-simulations. Yields of the purse seine fleet were more resilient to the range
-shift under the *Rate* design strategy, while the longline fleet had the exact
-opposite result. Under the *Rate* strategy, longline yields were relatively
-stable across a large MPA size under the *status quo* conditions, but declined
-rapidly under the *range shift* conditions. For the purse seine fleet, yields
-were more stable under the *range shift* conditions under the *Target Fishing*
-design scenario, but declined quickly as a function of MPA size under the
-*status quo* conditions (@fig-blue-results A). In terms of fisheries yield, the
-*Target Fishing* strategy would produce better yield outcomes for both fishing
-fleets under the *range shift* conditions, but the *Rate* strategy would perform
-best under the *status quo*.
-
-From a conservation perspective, the primary tuna species (bigeye, skipjack, and
-yellowfin) were most sensitive to design strategy and climate scenario, with the
-*Target Fishing* strategies producing for example rapid conservation gains for
-bigeye under the *status quo*, while the *Rate* strategy resulted in net
-conservation losses until the MPA became extremely large. However, under the
-*Range Shift* scenario the MPAs had little impact on many of the tuna
-populations until the MPA size became extremely large, do to movement of the
-primary fishing grounds outside of the current hotpots. The 20 year time horizon
-simulated here was not enough to produce substantial gains for any of the shark
-species even with 100% closures except for the more rapidly growing blue shark
-(@fig-blue-results B).
+### Pelagic Fishery
+
+Yields of the purse-seine fleet were more resilient to the range shift under the
+*Rate* design strategy, while the longline fleet had the opposite result. Under
+the *Rate* strategy, longline yields were relatively stable across a large range
+of MPA sizes under the status quo conditions, but declined rapidly under the
+range shift conditions. For the purse-seine fleet, yields were more stable under
+the range shift conditions under the *Target Fishing* design scenario, but
+declined quickly as a function of MPA size under the status quo conditions
+(@fig-blue-results A). The *Target Fishing* strategy produced better yield
+outcomes for both fishing fleets under the range shift conditions, but the
+*Rate* strategy performed best under the status quo.
+
+The primary tuna species (bigeye, skipjack, and yellowfin) were most sensitive
+to design strategy and climate scenario, with the *Target Fishing* strategies
+producing for example rapid conservation gains for bigeye and skipjack under the
+status quo, while the *Rate* strategy resulted in small net conservation losses
+for both species until the MPA became extremely large. However, under the range
+shift scenario the MPAs had little impact on many of the tuna populations until
+the MPA size became extremely large, due to movement of the primary fishing
+grounds outside of the current hotpots where the MPAs are placed based on status
+quo conditions. The 20 year time horizon simulated here was not enough to
+produce substantial gains for any of the shark species even with 100% closures
+except for the more rapidly growing blue shark (@fig-blue-results B).
 
@@
          purseseine) %>%
-  pivot_longer(c(longline,purseseine), names_to = "fleet", values_to = "yield") %>%
+  rename(`purse-seine` = purseseine) |> 
+  pivot_longer(c(longline,`purse-seine`), names_to = "fleet", values_to = "yield") %>%
   pivot_wider(names_from = "climate", values_from =  "yield") %>%
@@
                values_to = "fleet_yield") %>%
+  mutate(fleet = ifelse(fleet == "purseseine","purse-seine",fleet)) |> 
   group_by(prop_mpa, fleet, placement_strategy, climate) %>%
@@
 must often make decisions on how to manage these systems based on limited
-empirical evidence. Modeling tools such as `marlin` can help users answer
-important scientific questions and design marine management policies informed by
-a better understanding of sensitivities to key uncertainties. In this paper we
+empirical evidence. Modeling tools such as the one presented here can help users
+answer scientific questions and design marine management policies informed by a
+better understanding of sensitivities to key uncertainties. In this paper we
 have documented the core system of equations underlying `marlin`, and
-demonstrated how `marlin` can be used to support efforts such as MPA planning.
+demonstrated how `marlin` can be used to support marine resource management.
 
@@
 stemming from alternative design choices and states of nature shown in our case
-study results demonstrates the value in our method. For example, our coastal
-coral reef example illustrates both the potential for MPAs to benefit
-conservation and food production in these systems, and the potential for the
-same MPA to benefit some fleets and species while harming others when fishing
-fleets affected by an MPA do not share the same objectives and species are not
-uniformly distributed. Our pelagic case study illustrates how `marlin` can be
-used to assess the climate resiliency of spatial management strategies, and
-identify strategies that best meet the needs of both current and future
-conditions. Results from this type of work could be used to help prioritize
-communities and species at particular risk to climate change impacts.
-
-Examining our coral reef case study, our results indicate how this model could
-be used to examine the sensitivity of conservation and food security outcomes of
-MPAs to alternative design scenarios, holding constant other social-ecological
-variables. The *Spawning Ground* MPA network was able to provide more equitable
-yield outcomes across the two fishing communities, whereas the *Target Fishing*
-MPA network only benefited Fleet Two (@fig-coral-results A). This is because
-Fleet One has two primary fishing grounds; the spawning grounds, and the
-offshore area where the deep snapper live. Fleet Two primarily fishes in the
-north west corner of the simulation grid. Under the *Spawning Ground* strategy,
-while Fleet One quickly losses fishing grounds on the spawning grounds, it can
-compensate for this by moving offshore and fishing the deep snapper population
-harder, while also being able to fish some of the spillover of snapper from the
-spawning ground closure. This is why biomass of deep snapper actually declines
-under the *Spawning Ground* strategy.
+studies demonstrates the value in our method. For example, our coastal coral
+reef example illustrates both the potential for MPAs to benefit conservation and
+food production in these systems, and the potential for the same MPA to benefit
+some fleets and species while harming others when fishing fleets affected by an
+MPA do not share the same objectives and species are not uniformly distributed.
+Our pelagic case study illustrates how `marlin` can be used to assess the
+climate resiliency of spatial management strategies, and identify strategies
+that best meet the needs of both current and future conditions. Results from
+this type of work could be used to help prioritize communities and species at
+particular risk to climate change impacts.
+
+Our coastal coral reef case study results show the sensitivity of conservation
+and food security outcomes of MPAs to alternative design scenarios, holding
+constant other social-ecological variables. The *Spawning Ground* MPA network
+was able to provide more equitable yield outcomes across the two fishing
+communities, whereas the *Target Fishing* MPA network only benefited the more
+industrial Fleet Two (@fig-coral-results A). This is because Fleet One has two
+primary fishing grounds; the spawning grounds, and the offshore area where the
+deep snapper live. Fleet Two primarily fishes in the northern portions of the
+simulation grid. Under the *Spawning Ground* strategy, while Fleet One quickly
+losses fishing grounds on the spawning grounds, it can compensate for this by
+moving offshore and fishing the deep snapper population harder, while also being
+able to fish some of the spillover of snapper from the spawning ground closure.
+This displacement of Fleet One's fishing effort is why biomass of deep snapper
+actually declines under the *Spawning Ground* strategy.
 
@@
 benefits at little cost to their fishing grounds until the MPAs reach their
-fishing grounds once protection nears 100%. By protecting both the spawning
-ground and the offshore areas, the *Target Fishing* network provides
-conservation benefits to all of the species, in contrast to the *Spawning
-Ground* strategy that only provides meaningful conservation gains to the snapper
-population up until very large MPA sizes. Our result that protection of a
-dedicated spawning ground did not provide substantially greater conservation
-outcomes for the species using that spawning ground (groupers and sharks) than
-an alternative design strategy is supported by other modeling studies that show
-that the impacts of spawning ground protection on conservation and yields may be
-highly variable, and that displacement of high levels fishing effort from the
-spawning grounds can offset potential conservation gains of the protection
-[@grüss2014] .
+fishing grounds once protection nears 100%.
+
+By protecting both the spawning ground and the offshore areas, the *Target
+Fishing* network provides conservation benefits to all of the species, in
+contrast to the *Spawning Ground* strategy that only provides meaningful
+conservation gains to the snapper population up until very large MPA sizes. Our
+result that protection of a dedicated spawning ground did not provide
+substantially greater conservation outcomes for the species using that spawning
+ground (groupers and sharks) than an alternative design strategy is supported by
+other modeling studies that show that the impacts of spawning ground protection
+on conservation and yields may be highly variable, and that displacement of high
+levels fishing effort from the spawning grounds can offset potential
+conservation gains of the protection [@grüss2014]. That being said there is also
+evidence for the benefits of spawning aggregation protection [@erisman2015] ,
+and our results are not nearly sufficiently resolved to provide any general
+statements as to the relative value of spawning aggregation protection relative
+to other design strategies.
 
 The two design strategies produced very different outcomes in terms of total
-changes in yield and conservation outcomes (@fig-coral-results C). Both MPA
-network design strategies were capable of producing net "win-win" outcomes in
-which both food security (as measured by change in total fisheries yields) and
-conservation (as measured by change in SSB/SSB0). However, in general for the
-same size MPA network the *Spawning Ground* design strategy produced better food
-security outcomes but worse conservation outcomes, and *vice* *versa* for the
-*Target Fishing* strategy. The coral reef case study shows how the modeling
-framework presented here can help stakeholders explore how different management
-strategies affect outcomes both in total (@fig-coral-results C) and across
-species (@fig-coral-results A) and fishing communities (@fig-coral-results B).
-
-Turning to our Pelagic Ecosystem case study, we see that overall the yield
-outcomes of the purse-seine fleet were much more sensitive to the presence of a
-range shift than the longline fleet, particularly under the *Target Fishing*
-design strategy. This is because the purse-seine fleet primarily targets
-skipjack tuna, which in this simulation are concentrated in a relatively narrow
-latitudinal band (@fig-blue), and the purse-seine fleet makes up a large portion
-of the total catch in the simulated fishery. So, the "Target Fishing" strategy
-starts by closing off the main purse-seine fishing grounds, which while only
-small part of the spatial domain of the model represents a large portion of the
-purse-seine fleet's fishing grounds, resulting both in more rapid conservation
-gains and fishery losses under the *Status Quo* species distributions. The
-*Rate* strategy places more MPAs in areas that are of lesser importance fo the
-purse-seine fleet but overlap more with species such as Oceanic Whitetip Shark.
-Conversely, the purse-seine fleet appears to do have better MPA yield outcomes
-under the Range Shift scenario, not because of rebuilding of the Skipjack
-population, but due to future fishing grounds being essentially unprotected by
-smaller MPAs targeting the current skipjack distribution ( @fig-blue-results ).
+changes in yield and conservation outcomes for the coral reef case study
+(@fig-coral-results C). Both MPA network design strategies were capable of
+producing net "win-win" outcomes in which both food security (as measured by
+change in total fisheries yields) and conservation (as measured by change in
+SSB/SSB0). However, in general for the same size MPA network the *Spawning
+Ground* design strategy produced better food security outcomes but worse
+conservation outcomes, and *vice* *versa* for the *Target Fishing* strategy. The
+coral reef case study shows how the modeling framework presented here can help
+stakeholders explore how different management strategies affect outcomes both in
+total (@fig-coral-results C) and across species (@fig-coral-results A) and
+fishing communities (@fig-coral-results B).
+
+Turning to our Pelagic Ecosystem case study, the yield outcomes of the
+purse-seine fleet were much more sensitive to the presence of a range shift than
+the longline fleet, particularly under the *Target Fishing* design strategy.
+This is because the purse-seine fleet primarily targets skipjack tuna, which in
+this simulation are concentrated in a relatively narrow latitudinal band
+(@fig-blue), and the purse-seine fleet makes up a large portion of the total
+catch in the simulated fishery. So, the "Target Fishing" strategy starts by
+closing off the main purse-seine fishing grounds, which while only small part of
+the spatial domain of the model represents a large portion of the purse-seine
+fleet's fishing grounds, resulting both in more rapid conservation gains and
+fishery losses under the status quo species distributions. The *Rate* strategy
+places more MPAs in areas that are of lesser importance fo the purse-seine fleet
+but overlap more with species such as oceanic whitetip shark. Conversely, the
+purse-seine fleet appears to do have better MPA yield outcomes under the range
+shift scenario, not because of rebuilding of the skipjack population, but due to
+future fishing grounds being essentially unprotected by smaller MPAs targeting
+the current skipjack distribution ( @fig-blue-results ).
 
@@
 
-These case studies are only intended to illustrate the capabilities of the model
-and the importance of considering the bio-economic dynamics represented in
-`marlin`; it is beyond the scope of this paper to provide broader conclusions
-about the performance of MPAs under different contexts or a comprehensive
-comparison of simulation and empirical results of MPAs around the world. That
-being said, the kinds of dynamics resulting in our case studies are well
-supported by both modeling and empirical studies. Our results support conceptual
-[@hilborn2004a; @gaines2010] and empirical [@ban2019] evidence that under the
-right conditions no-take MPAs can benefit fisheries and conservation.
-@rassweiler2014 demonstrated that the kinds of design choices presented in our
-case studies can greatly drive MPA outcomes. @ovando2021a provides empirical
-support for the possibility of even well enforced and large MPAs to have less
-pronounced population-level conservation effects (@fig-coral-results A) than
-might be expected given observed differences in biomass inside and outside of
-protected areas.
+These case studies are intended to illustrate the capabilities of the model and
+the importance of considering the bio-economic dynamics represented in `marlin`;
+it is beyond the scope of this paper to provide broader conclusions about the
+performance of MPAs under different contexts or a comprehensive comparison of
+simulation and empirical results of MPAs around the world. That being said, the
+kinds of dynamics resulting in our case studies are well supported by both
+modeling and empirical studies. Our results support conceptual [@hilborn2004a;
+@gaines2010] and empirical [@ban2019] evidence that under the right conditions
+no-take MPAs can benefit fisheries and conservation. @rassweiler2014
+demonstrated that the kinds of design choices presented in our case studies can
+greatly drive MPA outcomes.
 
@@
 around the world, with much of the empirical evidence of their performance
-centered on comparing measures such as biomass densities inside protected areas
-relative to outside [@lester2009]. Our model predicts a similar rapid increase
-of the simple ratio of mean biomass inside MPAs relative to outside (Fig.XX,
-acknowledging that designing a proper response ratio would require controlling
-for habitat characteristics and MPA design criteria). However, despite producing
-clear response ratios, The net conservation and fishery outcomes of our
-simulated coral-reef ecosystem MPAs were not nearly as large or clear as the
-response ratio results (@fig-coral-results); this result is supported works such
-as @ferraro2018 an @ovando2021a showing that response ratios alone may be a poor
-indicator of the net causal impacts of MPAs.
+centered on demonstrating higher metrics such as biomass densities inside
+protected areas relative to fished reference areas ["response ratios",
+@lester2009]. Our model predicts a similar rapid increase of the simple ratio of
+mean biomass inside MPAs relative to outside (Fig. S5, acknowledging that
+designing a proper response ratio would require controlling for habitat
+characteristics and MPA design criteria). However, despite producing clear
+response ratios, the net conservation and fishery outcomes of our simulated
+coral-reef ecosystem MPAs were not nearly as large or clear as the response
+ratio results (@fig-coral-results); this result is supported by works such as
+@ferraro2018 an @ovando2021a showing that response ratios alone may be a poor
+indicator of the net causal impacts of MPAs at the population scale.
 
 Studies such as @gilman2020, @abbott2012, @davies2018 and @pons2022 provide
-empirical support for our case study results showing how MPAs can produce
+empirical support for our case study results showing that MPAs can produce
 trade-offs across the conservation of different species and fleets when the
@@
 complexity, but can be extremely time consuming to design and run. `marlin`
-allows some of the realism of these end-to-end models while being simpler and
-faster to parameterize and run.
+allows some of the realism of these more end-to-end models while being simpler
+and faster to parameterize and run.
 
@@
 and species preferences) while using scale-free parameters that can be measured
-experimentally in laboratory or tagging studies.
+experimentally in laboratory or tagging studies. In addition, `marlin` allows
+for simulating not only the distribution of the species but also biomass, age,
+and length structure in space and time.
 
@@
 `marlin` does not represent trophic interactions. Anthropgenic changes in
-species abundance can clearly result in trophic cascades throughout the affected
-food web. However, studies such as @gilman2020, @ovando2021a, @bruno2019 , and
+species abundance can result in trophic cascades throughout affected food webs.
+However, studies such as @gilman2020, @ovando2021a, @bruno2019 , and
 @malakhoff2021 found no clear signs of MPA driven trophic cascades within the
@@
 environmental context [@liu2022], or may simply take longer to develop
-detectable effects than the coverage of many time series of MPAs. This is to say
-then that while `marlin` does not incorporate trophic interactions, what
-empirical evidence we have does not suggest that management-mediated trophic
-cascades are so common and clear that they must be incorporated into any
-credible multi-species simulation model. Rather, there is still value in
-capturing other dynamic bio-economic processes, as done here, even if that comes
-at the expense of modeling trophic interactions. However, research on the
-trophic impacts of spatial-temporal management actions is clearly of value and
-models like Atlantis [@audzijonyte2019], Ecopath with Ecosim [@christensen2004],
-and EASI-Fish [@griffiths2019] can help users explore those kinds of processes,
-in the manner of @baskett2007.
-
-### General Recommendations for Using `marlin`
+detectable effects than the coverage of many time series of MPAs. While `marlin`
+does not incorporate trophic interactions, what empirical evidence we have does
+not suggest that management-mediated trophic cascades are so common and clear
+that they must be incorporated into any credible multi-species simulation model.
+There is value in capturing other dynamic bio-economic processes, as done here,
+even if that comes at the expense of modeling trophic interactions. However,
+research on the trophic impacts of spatial-temporal management actions is
+clearly of value and models like Atlantis [@audzijonyte2019], Ecopath with
+Ecosim [@christensen2004], and EASI-Fish [@griffiths2019] can help users explore
+those kinds of trophic processes, in the manner of @baskett2007.
+
+### General Recommendations for Use
 
@@
 number of options at their disposal that they must decide on. Our recommended
-strategy for dealing with this is for users to try and narrow down a list of
-parameters that they feel are sufficiently "known" and another list of
-parameters that are highly uncertain and / or contentious which the users feel
-may impact results. Where possible, parameters from locally estimated stock
-assessment models can be used to provide a foundation around which sensitivity
-analyses around specific uncertainties of interest [@berger2017].
+strategy is for users to try and narrow down a list of parameters that they feel
+are sufficiently "known" and another list of parameters that are highly
+uncertain and / or contentious which the users feel may impact results. Where
+possible, parameters from locally estimated stock assessment models can be used
+to provide a foundation around which sensitivity analyses around specific
+uncertainties of interest [@berger2017].
 
@@
 the sensitivity of proposed MPA networks to key unknowns such as adult and
-larval dispersal rates or the the economic incentives and gear selectivity of
+larval dispersal rates or the the economic incentives and contact selectivity of
 the fishing fleets. As an example of this, we ran an alternative version of our
-Coral Reef case study in which Fleet Two was assigned logistic selectivity for
+coral reef case study in which Fleet Two was assigned logistic selectivity for
 the snapper and grouper species, rather than the dome-shaped selectivity
-presented in our main results, reflecting a scenario where users are uncertain
-about the contact selectivity of Fleet Two. We found that the impacts of MPAs on
-conservation and food security were relatively insensitive to the form of
-contact selectivity specified for Fleet Two (Fig.S1). This does not mean that
-knowing the "correct" form of contact selectivity may not be extremely important
-in providing accurate assessment results [@waterhouse2014], but rather that in
-this case misspecifying the contact selectivity curve is projected to have
-little impact on the simulated outcomes of MPAs, conditional on holding other
-parameters constant.
+presented in our main results. We found that the impacts of MPAs on conservation
+and food security were relatively insensitive to the form of contact selectivity
+specified for Fleet Two (Fig.S1). This does not mean that knowing the "correct"
+form of contact selectivity may not be extremely important in providing accurate
+assessment results [@waterhouse2014], but rather that in this case misspecifying
+the contact selectivity curve is projected to have little impact on the
+simulated outcomes of MPAs, conditional on holding other parameters constant.
 
@@
 "strategically" or "tactically" will depend on the needs of the user. That being
-said, since the model is not fit to data directly it is more easily applied to
+said, since. `marlin` is not fit to data directly it is more easily applied to
 strategic questions. For example, the two case studies presented in this paper
 provide general strategic insights into the sensitivity of outcomes to design
-choices and ... xx expand here xx. But, because `marlin` is not fit to data, the
-degree to which it can represent detailed dynamics of a system depends on the
-quality of the input data; The extent to which users are comfortable using the
+choices range shifts. The extent to which users are comfortable using the
 outputs of `marlin` tactically will depend on the confidence they have in their
@@
 management. Other models may be better suited to address specific forms of
-complexity, such as predatory-prey dynamics (citations) or more complex fleet
-dynamics through agent-based approached [@burgess2020a]. Particularly for more
-tactical applications, we would encourage users to explore multiple modeling
-frameworks to help design policies that are likely to be robust to many
-different kinds of complexities.
+complexity. Particularly for more tactical applications, we would encourage
+users to explore multiple modeling frameworks to help design policies that are
+likely to be robust to many different kinds of complexities.
 
@@
 value though in being able to simulate and compare the outcomes of perfectly
-implmented policies to isolate the concept of the polucy itself from its
+implemented policies to isolate the concept of the policy itself from its
 implementation. That being said, future work may add the capability for
-imperfect implmentation to the model.
+imperfect implementation to the model.
+
+Fish populations often exhibit variation in demographic traits such as growth,
+maturation, and mortality across dimensions such as space, time, and sex. Models
+such as Stock Synthesis [@methotjr.2013] capture these processes through the use
+of "morphs". `marlin` does not currently allow for these kinds of processes
+explicitly, though users could simulate different populations of the same
+species with different life history traits as separate stocks. Users should be
+cautious interpreting simulation results from `marlin` for species in which
+these sorts of dynamics are likely to be particularly prevalent. For strongly
+sexually dimorphic species, we would recommend picking the sex most likely to
+drive the outcomes of management policies, which is often females given the
+general prevalence of eggs as the limiting reproductive material in marine
+ecosystems. Users should also proceed with caution using `marlin` for species
+with more complex reproductive biology, such as sex-changing fish
+[@kindsvater2017].
+
+## Conclusions
 
@@
 objectives, such as maximizing conservation gains for a threatened species while
-minimizing or removing negative impacts to affected fishing fleets.
-
-Beyond the current examples, we envision `marlin` being applicable to research
-on dynamic ocean management, range shifts, management strategy evaluation,
-policy interactions, and spatial stock assessment. `marlin` can help researchers
-generate data for further testing of the performance and design of spatially
-explicit integrated population models, in the manner of @bosley2022. In
-addition, the process-based movement model used in `marlin` can directly use
-empirical estimates of movement dynamics derived in the manner of @thorson2021a,
-providing a link between empirical and simulation based approaches to marine
-resource management that has been challenging to implement in spatial
-simulations [@berger2017]. The model currently focuses on diffusion and taxis,
-which allows for general representation of animals following physical or
-oceanographic features, but research on the incorporation and importance of
-advective forces would be of value going forward. Future iterations of the model
-may include life history modules for a broader range of species types such as
-marine mammals, sea birds, and sea turtles.
+minimizing or removing negative impacts to affected fishing fleets. We envision
+`marlin` being applicable to research on dynamic ocean management, range shifts,
+management strategy evaluation, policy interactions, and spatial stock
+assessment. `marlin` can help researchers generate data for further testing of
+the performance and design of spatially explicit integrated population models,
+in the manner of @bosley2022. In addition, the process-based movement model used
+in `marlin` can directly use empirical estimates of movement dynamics derived in
+the manner of @thorson2021a, providing a link between empirical and simulation
+based approaches to marine resource management that has been challenging to
+implement in spatial simulations [@berger2017].
 
@@
 aspects of the marine resource management process.
-
-## Conclusions
 

### 2023-10-24 03_marlin_paper.qmd
pre submission edits

@@
       = BH(SSB_{t-1,p,s,a}) & \text{if $a = 1$}\\
-     = N_{t-1,p,s,a-1}e^{-(z_{t-1,p,s,a-1}}, & \text{if $1< a < max(age)$}\\
+     = N_{t-1,p,s,a-1}e^{-(z_{t-1,p,s,a-1})}, & \text{if $1< a < max(age)$}\\
      =  N_{t-1,p,s,a}e^{-(z_{t-1,p,s,a})} + N_{t-1,a-1}e^{-(z_{t-1,p,s,a-1})}, & \text{if $a = max(a)$}
@@
 :::
+
+## Figure Legends

### 2023-10-25 03_marlin_paper.qmd
first look edits

@@
 
-# rdatas <- results_files[str_detect(results_files,'.Rdata')]
-# 
-# walk(rdatas, ~load(file.path(results_path,.x), envir = .GlobalEnv)) # load results into global environment
+small_width <- 80 # figure width in mm
+
+big_width <- 180 # width in mm
 
@@
 
-Bio-economic modeling, spatio-temporal modeling, marine protected areas,
-fisheries management, marine spatial planning, conservation planning
+bio-economic modeling
+conservation planning
+fisheries
+marine protected areas
+marine spatial planning
+spatio-temporal modeling
 
@@
 7.  References
+8. Figure Legends
 
@@
 distributions, exploitation levels, fleet dynamics, and policy-dependent
-behavior [@reimer2017], all of which may change over time, particularly given
+behavior [@reimer2017], all of which may fluctuate over time, particularly given
 the impacts of climate change. Efforts to effectively design spatial management
@@
 Many models have been developed to support the theory and design of spatial
-management strategies (see @fulton2015 for a review of existing methods).
+management strategies (see @fulton2015 for a review).
 However, these models have tended to be either highly complex and tactical
@@
 representations intended to provide theoretical insights with less relevance for
-specific applications. This lack of accessible models suited for more tactical
-application presents a challenge to stakeholders charged with marine spatial
+specific applications. This lack of accessible models capable of representing reasonable amounts of complexity presents a challenge to stakeholders charged with marine spatial
 planning exercises. It also presents a barrier to the scientific community,
@@
 
-To help address this challenge, we present a novel bio-economic modeling tool
+To help address this challenge, we present a bio-economic modeling tool
 called [`marlin`](https://danovando.github.io/marlin/) that allows for efficient
@@
 contact selectivity [@sampson2014] for individual species (each of which we
-denote as a *metier*). For example, both a longline and purse seine fleet may
-capture bigeye (*Thunnus obesus*) and skipjack tunas, but the longline fleet can
+denote as a *metier*). For example, both a longline and purse-seine fleet may
+capture bigeye (*Thunnus obesus*) and skipjack (*Katsuwonus pelamis*) tunas, but the longline fleet can
 be made much more likely to catch larger bigeye than skipjack, and *vice versa*.
@@
 becomes increasingly difficult as the range of objectives and the complexity of
-the model increase. Therefore, while `marlin` provides some support for
-[optimization](https://danovando.github.io/marlin/articles/optimize-mpas.html),
-we focus here on the design and performance of these more rule of thumb design
+the model increase. Therefore, we focus here on the design and performance of these more rule of thumb design
 strategies that may be more accessible to a wider range of users. We allow all
@@
 
+#### Pelagic Fishery
+
+```{r}
+blue_fauna <- blue_water_fauna_and_fleets$fauna
+
+blue_fleets <- blue_water_fauna_and_fleets$fleets
+
+blue_resolution <- sqrt(blue_fauna[[1]]$patches)
+
+```
+
+We model our pelagic case study loosely on the characteristics of the Western
+and Central Pacific Ocean (WCPO) tuna fisheries.
+Note that this is an illustrative example only and simulated stock status, species distributions, or outcome projections presented here should not be interpreted as a indicative of the current or future state of the WCPO.
+We simulate trajectories of
+`r n_distinct(names(blue_fauna))` species commonly caught in the region,
+including both the highly abundant skipjack tuna and the heavily depleted
+oceanic whitetip shark (*Carcharhinus longimanus*) (@fig-blue-ssb). We use
+publicly available data on catch-per-unit-effort of each of these species from
+the WCPO as a very rough proxy for baseline habitat distributions, noting that
+where possible, fishery-independent abundance indices would be preferable
+(@fig-blue).
+
+These pelagic species are caught by a longline fleet that primarily targets
+large adult tunas such as bigeye and yellowfin for high-grade consumption, and a
+purse-seine fishery that primarily targets skipjack tunas for bulk canning.
+Contact selectivities were modeled as logistic for the longline fleet, and
+dome-shaped for the purse-seine fleet, with selectivity values set to broadly
+match the patterns reported in relevant [stock
+assessments](https://oceanfish.spc.int/en/ofpsection/sam/sam/210-stock-assessment-input-and-result-files-)
+for each species. Both fleets operate under open-access dynamics with an effort
+cap. The effort cap was set at the level of effort that resulted in the desired
+levels of SSB/SSB0 for each species under open-access dynamics (@fig-blue),
+intended to simulate a scenario where managers step in to prevent further
+expansion of fishing effort in a fully developed fishery. For
+forward-simulation, open-access dynamics can result in effort decreasing in
+response to profitability, but cannot result in effort beyond the effort cap set
+for each fleet.
+
+For this exercise, we focused on using `marlin` to assess resilience of the
+selected *Target Fishing* and *Rate* MPA design strategies to a climate-driven
+range shift. Specifically, we simulate an extreme example where the centroid of
+each population shifts northward at a rate of 0.2 patches per year over a 20
+year time horizon (@fig-blue-ssb). We designed MPA networks given the conditions
+in the starting year, and then held that network constant over the years of the
+experiment, running one simulation with and another without the climate-drive
+range shift. We then compared the effects of this range shift on food production
+and conservation outcomes from MPA networks designed based on the pre-range
+shift world. We ran the pelagic simulation at a quarterly level ($\Delta_t$ =
+1/4 year), and set the area of each cell to be roughly 97,000 KM^2^ across 144
+patches each with a side length of roughly 311 KM, for a total area of
+14,000,000 KM^2^, broadly commensurate with the area of the WCPO.
+
+## Results
+
+### Coastal Coral Reef Fishery
+
+```{r}
+#| label: coral-results
+
+coral_mpa_experiments$mpas <- map(coral_mpa_experiments$results, "mpa")
+
+coral_mpa_experiments$obj <- map(coral_mpa_experiments$results, "obj")
+
+coral_results <- coral_mpa_experiments %>%
+  unnest(cols = obj)
+
+coral_fleet_results <- coral_results %>%
+  pivot_longer(starts_with("fleet_"),
+               names_to = "fleet",
+               values_to = "fleet_yield") %>%
+  group_by(prop_mpa, fleet, placement_strategy) %>%
+  summarise(yield = sum(fleet_yield), biodiv = sum(unique(biodiv))) %>%
+  group_by(fleet, placement_strategy) %>%
+  mutate(delta_yield = yield / yield[prop_mpa == 0] - 1) |> 
+  ungroup()
+
+peak_coral_fleet_results <- coral_fleet_results |> 
+  group_by(fleet) |> 
+  filter(delta_yield == max(delta_yield)) |> 
+  mutate(peak_yields = scales::percent(delta_yield))
+
+tmp <- coral_results %>% 
+  group_by(prop_mpa, placement_strategy) %>% 
+  summarise(biodiv = sum(biodiv), yield = sum(yield)) %>% 
+  group_by(placement_strategy) %>% 
+  mutate(delta_yield = yield / yield[prop_mpa == 0] - 1,
+         delta_biodiv = biodiv / biodiv[prop_mpa == 0] - 1) %>% 
+  ungroup() %>% 
+    mutate(placement_strategy= fct_relabel(placement_strategy, titler))
+
+```
+
+MPAs were capable of producing a range of positive and negative outcomes for
+food security and conservation in the coral reef case study depending on the
+design strategy used and the size of MPA implemented. Both of the MPA design
+strategies were capable of increasing fisheries yield for the more "industrial"
+Fleet Two, up to a value of
+`r peak_coral_fleet_results$peak_yields[peak_coral_fleet_results$fleet == "fleet_two"]`,
+even when MPAs covered more than 50% of the simulated area. However, the more
+"artisanal" Fleet One only benefited from MPAs under the *Spawning Ground*
+design strategy, with a maximum increase of
+`r peak_coral_fleet_results$peak_yields[peak_coral_fleet_results$fleet == "fleet_one"]`;
+the *Target Fishing* design strategy produced a roughly linear decrease in
+fishing yields as a function of increasing MPA size (@fig-coral-results A).
+
+On the conservation front, MPAs were uniformly beneficial to the spawning
+biomass of all species under the *Target Fishing* design strategy, with the most
+rapid increases in spawning biomass for the deep snapper population. The
+*Spawning Ground* strategy primarily benefited the shallow snapper population,
+producing little change in the grouper population and decreasing spawning
+biomass of both the deep snapper and reef shark populations for MPA sizes
+covering less than 50% of the simulated area (@fig-coral-results B).
+
+In total, the *Spawning Ground* strategy was capable or providing net increases
+in fishing yield (summed across both fleets) of up to roughly 7%, with positive
+net yield impacts up to network size of nearly 60% of the area. However, even
+MPAs covering 60% of the area only produced a maximum increase of roughly 5% in
+total SSB/SSB0 when designed around the spawning ground. Conversely, the *Target
+Fishing* design strategy was capable of producing a nearly 25% increase in total
+SSB/SSB0 for the same MPA size, but at a much greater cost to the total food
+production from the system's fisheries (@fig-coral-results C).
+
+<!-- ~~These macro-level changes in total yield and conservation are the result of -->
+
+<!-- interesting internal dynamics that marlin allows users to explore through output -->
+
+<!-- tables at the level of individual species and fleets. The *Spawning Ground* -->
+
+<!-- strategy increased yield to Fleet One when MPAs were relatively small, but was -->
+
+<!-- able to maintain yields near *status quo* levels up to a very large MPA size to -->
+
+<!-- the dependence of Fleet One on only a relatively small portion of the total -->
+
+<!-- simulation area. However, *Spawning Ground* MPAs were unable to provide -->
+
+<!-- increases in yield to Fleet Two. The *Spawning Ground* strategy produced the -->
+
+<!-- most rapid conservation gains for the snapper population. Interestingly, due to -->
+
+<!-- effort displacement from the spawning grounds onto the fishing areas inhabited -->
+
+<!-- by the groupers the rest of the year, closure of the spawning ground had limited -->
+
+<!-- impact on the grouper population until the MPA size exceeded 50% of the -->
+
+<!-- simulation area. Effort displacement from the spawning grounds also resulted in -->
+
+<!-- a small decrease in the deep snapper population for MPA sizes less than 50%. The -->
+
+<!-- *Target Fishing* design strategy has similar yield outcomes to the *Spawning -->
+
+<!-- Ground* strategy, but less dramatic negative impacts on Fleet Two, at the -->
+
+<!-- expense of slightly fewer benefits for Fleet One. However, the *Target Fishing* -->
+
+<!-- strategy provided more uniform conservation benefits to all species -->
+
+<!-- (@fig-coral-results).~~ -->
+
+### Pelagic Fishery
+
+Yields of the purse-seine fleet were more resilient to the range shift under the
+*Rate* design strategy, while the longline fleet had the opposite result. Under
+the *Rate* strategy, longline yields were relatively stable across a large range
+of MPA sizes under the status quo conditions, but declined rapidly under the
+range shift conditions. For the purse-seine fleet, yields were more stable under
+the range shift conditions under the *Target Fishing* design scenario, but
+declined quickly as a function of MPA size under the status quo conditions
+(@fig-blue-results A). The *Target Fishing* strategy produced better yield
+outcomes for both fishing fleets under the range shift conditions, but the
+*Rate* strategy performed best under the status quo.
+
+The primary tuna species (bigeye, skipjack, and yellowfin) were most sensitive
+to design strategy and climate scenario, with the *Target Fishing* strategies
+producing for example rapid conservation gains for bigeye and skipjack under the
+status quo, while the *Rate* strategy resulted in small net conservation losses
+for both species until the MPA became extremely large. However, under the range
+shift scenario the MPAs had little impact on many of the tuna populations until
+the MPA size became extremely large, due to movement of the primary fishing
+grounds outside of the current hotpots where the MPAs are placed based on status
+quo conditions. The 20 year time horizon simulated here was not enough to
+produce substantial gains for any of the shark species even with 100% closures
+except for the more rapidly growing blue shark (@fig-blue-results B).
+
+```{r}
+#| include: false
+
+blue_water_experiments$mpas <- map(blue_water_experiments$results, "mpa")
+
+blue_water_experiments$obj <- map(blue_water_experiments$results, "obj")
+
+blue_water_climate_experiments$mpas <- map(blue_water_climate_experiments$results, "mpa")
+
+blue_water_climate_experiments$obj <- map(blue_water_climate_experiments$results, "obj")
+
+static_results <- blue_water_experiments %>%
+  unnest(cols = obj) %>%
+  mutate(climate = FALSE)
+
+climate_results <- blue_water_climate_experiments %>%
+  unnest(cols = obj) %>%
+  mutate(climate = TRUE)
+
+blue_water_results <-static_results %>%
+  bind_rows(climate_results) %>% 
+  left_join(sci_to_com, by = "critter") %>% 
+  select(-critter) %>% 
+  rename(critter = common_name)
+
+fleet_climate_results <- blue_water_results %>%
+  select(placement_strategy,
+         fleet_model,
+         prop_mpa,
+         climate,
+         longline,
+         critter,
+         purseseine) %>%
+  rename(`purse-seine` = purseseine) |> 
+  pivot_longer(c(longline,`purse-seine`), names_to = "fleet", values_to = "yield") %>%
+  pivot_wider(names_from = "climate", values_from =  "yield") %>%
+  mutate(delta = `TRUE` - `FALSE`) %>%
+  group_by(placement_strategy, prop_mpa, fleet) %>%
+  summarise(delta = sum(delta))
+
+fleet_climate_impacts <- fleet_climate_results %>%
+  ggplot(aes(prop_mpa, delta, color = fleet)) +
+  geom_hline(yintercept = 0, linetype = 2) +
+  geom_line() +
+  facet_wrap(~placement_strategy) +
+  scale_y_continuous(name = "Difference in Yield") +
+  theme(axis.text.x = element_blank(), axis.title.x = element_blank())
+
+critter_climate_results <- blue_water_results %>%
+  select(placement_strategy,
+         fleet_model,
+         prop_mpa,
+         climate,
+         critter,
+         biodiv) %>%
+  pivot_wider(names_from = "climate", values_from =  "biodiv") %>%
+  mutate(delta = `TRUE` - `FALSE`)
+
+critter_climate_impacts <- critter_climate_results %>%
+  ggplot(aes(prop_mpa, delta, color = critter)) +
+  geom_line() +
+  facet_wrap(~placement_strategy) +
+  scale_x_continuous(name = "MPA",labels = scales::label_percent()) +
+  scale_y_continuous(name = "Difference in SSB/SSB0")
+
+fleet_climate_impacts / critter_climate_impacts
+
+```
+
+## Discussion
+
+Marine ecosystems are driven by complex social-ecological dynamics. Communities
+must often make decisions on how to manage these systems based on limited
+empirical evidence. Modeling tools such as the one presented here can help users
+answer scientific questions and design marine management policies informed by a
+better understanding of sensitivities to key uncertainties. In this paper we
+have documented the core system of equations underlying `marlin`, and
+demonstrated how `marlin` can be used to support marine resource management.
+
+### Insights from Case Studies
+
+The variation in both the magnitude and distribution of costs and benefits
+stemming from alternative design choices and states of nature shown in our case
+studies demonstrates the value in our method. For example, our coastal coral
+reef example illustrates both the potential for MPAs to benefit conservation and
+food production in these systems, and the potential for the same MPA to benefit
+some fleets and species while harming others when fishing fleets affected by an
+MPA do not share the same objectives and species are not uniformly distributed.
+Our pelagic case study illustrates how `marlin` can be used to assess the
+climate resiliency of spatial management strategies, and identify strategies
+that best meet the needs of both current and future conditions. Results from
+this type of work could be used to help prioritize communities and species at
+particular risk to climate change impacts.
+
+Our coastal coral reef case study results show the sensitivity of conservation
+and food security outcomes of MPAs to alternative design scenarios, holding
+constant other social-ecological variables. The *Spawning Ground* MPA network
+was able to provide more equitable yield outcomes across the two fishing
+communities, whereas the *Target Fishing* MPA network only benefited Fleet Two (@fig-coral-results A). This is because Fleet One has two
+primary fishing grounds; the spawning grounds, and the offshore area where the
+deep snapper live. Fleet Two primarily fishes in the northern portions of the
+simulation grid. Under the *Spawning Ground* strategy, while Fleet One quickly
+losses fishing grounds on the spawning grounds, it can compensate for this by
+moving offshore and fishing the deep snapper population harder, while also being
+able to fish some of the spillover of snapper from the spawning ground closure.
+This displacement of Fleet One's fishing effort is why biomass of deep snapper
+actually declines under the *Spawning Ground* strategy.
+
+Conversely, most of the catch in the fishery comes from the nearshore and deep
+water snapper populations, which overlaps with both of Fleet One's fishing
+grounds. The *Target Fishing* strategy then begins closing both the nearshore
+and offshore snapper fishing grounds, resulting in too much of a loss in fishing
+grounds for Fleet One to be offset by spillover from the MPAs. As a result of
+primarily placing MPAs on Fleet One's fishing grounds, Fleet Two gains spillover
+benefits at little cost to their fishing grounds until the MPAs reach their
+fishing grounds once protection nears 100%.
+
+By protecting both the spawning ground and the offshore areas, the *Target
+Fishing* network provides conservation benefits to all of the species, in
+contrast to the *Spawning Ground* strategy that only provides meaningful
+conservation gains to the snapper population up until very large MPA sizes. Our
+result that protection of a dedicated spawning ground did not provide
+substantially greater conservation outcomes for the species using that spawning
+ground (groupers and sharks) than an alternative design strategy is supported by
+other modeling studies that show that the impacts of spawning ground protection
+on conservation and yields may be highly variable, and that displacement of high
+levels fishing effort from the spawning grounds can offset potential
+conservation gains of the protection [@grüss2014]. That being said there is also
+evidence for the benefits of spawning aggregation protection [@erisman2015] ,
+and our results are not nearly sufficiently resolved to provide any general
+statements as to the relative value of spawning aggregation protection relative
+to other design strategies. Further research could for example alter both the distribution of species in space and the susceptibility of the species to fishing gear when aggregated.
+
+The two design strategies produced very different outcomes in terms of total
+changes in yield and conservation outcomes for the coral reef case study
+(@fig-coral-results C). Both MPA network design strategies were capable of
+producing net "win-win" outcomes in which both food security (as measured by
+change in total fisheries yields) and conservation (as measured by change in
+SSB/SSB0). However, in general for the same size MPA network the *Spawning
+Ground* design strategy produced better food security outcomes but worse
+conservation outcomes, and *vice* *versa* for the *Target Fishing* strategy. The
+coral reef case study shows how the modeling framework presented here can help
+stakeholders explore how different management strategies affect outcomes both in
+total (@fig-coral-results C) and across species (@fig-coral-results A) and
+fishing communities (@fig-coral-results B).
+
+Turning to our Pelagic Ecosystem case study, the yield outcomes of the
+purse-seine fleet were much more sensitive to the presence of a range shift than
+the longline fleet, particularly under the *Target Fishing* design strategy.
+This is because the purse-seine fleet primarily targets skipjack tuna, which in
+this simulation are concentrated in a relatively narrow latitudinal band
+(@fig-blue), and the purse-seine fleet makes up a large portion of the total
+catch in the simulated fishery. So, the "Target Fishing" strategy starts by
+closing off the main purse-seine fishing grounds, which while only small part of
+the spatial domain of the model represents a large portion of the purse-seine
+fleet's fishing grounds, resulting both in more rapid conservation gains and
+fishery losses under the status quo species distributions. The *Rate* strategy
+places more MPAs in areas that are of lesser importance fo the purse-seine fleet
+but overlap more with species such as oceanic whitetip shark. Conversely, the
+purse-seine fleet appears to do have better MPA yield outcomes under the range
+shift scenario, not because of rebuilding of the skipjack population, but due to
+future fishing grounds being essentially unprotected by smaller MPAs targeting
+the current skipjack distribution ( @fig-blue-results ).
+
+The conservation outcomes of the skipjack, bigeye, and yellowfin tunas were
+among the most sensitive to range shifts, whereas even complete closure of the
+region was not sufficient to see significant increases in severely depleted
+species like oceanic whitetip sharks within the 20-year timeline of the
+simulation. Future research could be conducted then to see what sorts of
+timelines might be needed to see recovery of these species, and which design
+strategies result in the fastest recovery at the lowest cost to other objectives
+such as food security.
+
+These case studies are intended to illustrate the capabilities of the model and
+the importance of considering the bio-economic dynamics represented in `marlin`;
+it is beyond the scope of this paper to provide broader conclusions about the
+performance of MPAs under different contexts or a comprehensive comparison of
+simulation and empirical results of MPAs around the world. That being said, the
+kinds of dynamics resulting in our case studies are well supported by both
+modeling and empirical studies. Our results support conceptual [@hilborn2004a;
+@gaines2010] and empirical [@ban2019] evidence that under the right conditions
+no-take MPAs can benefit fisheries and conservation. @rassweiler2014
+demonstrated that the kinds of design choices presented in our case studies can
+greatly drive MPA outcomes.
+
+MPAs based around coral-reef style ecosystems have been extensively studied
+around the world, with much of the empirical evidence of their performance
+centered on demonstrating higher metrics such as biomass densities inside
+protected areas relative to fished reference areas ["response ratios",
+@lester2009]. Our model predicts a similar rapid increase of the simple ratio of
+mean biomass inside MPAs relative to outside (Fig. S5, acknowledging that
+designing a proper response ratio would require controlling for habitat
+characteristics and MPA design criteria). However, despite producing clear
+response ratios, the net conservation and fishery outcomes of our simulated
+coral-reef ecosystem MPAs were not nearly as large or clear as the response
+ratio results (@fig-coral-results); this result is supported by works such as
+@ferraro2018 an @ovando2021a showing that response ratios alone may be a poor
+indicator of the net causal impacts of MPAs at the population scale.
+
+Studies such as @gilman2020, @abbott2012, @davies2018 and @pons2022 provide
+empirical support for our case study results showing that MPAs can produce
+trade-offs across the conservation of different species and fleets when the
+spatial-temporal distributions and life history traits of affected species are
+heterogeneous, and MPAs result in some degree of effort displacement or
+concentration in the remaining fishing grounds. @hampton2023 supports our
+pelagic case study result showing that both conservation and fishery impacts of
+MPAs on the highly mobile species of the open oceans can be limited unless
+protected areas are very large. @brown2018 supports our result showing how
+heterogeneity in the behavior and objectives of fishing fleets sharing an
+ecosystem can affect the magnitude and equity of fishery reform outcomes.
+@davies2017 discusses the importance of considering the potential of climate
+driven range shifts when assessing spatial management policies; our work builds
+on this by allowing users to not just simulate species distributions but also
+distribution of biomass and age composition of fish in space and time under climate
+change.
+
+### Putting `marlin` In Context
+
+The role of this paper is not to conduct a review of the many tools for
+spatial-temporal modeling available in the literature, each of which provide
+useful functionality for specific applications. However, we highlight here the
+specific gaps that we feel `marlin` fills in the modeling literature and in the
+policy support toolbox using some selected publications. More end-to-end models
+such as ATLANTIS [@audzijonyte2019] , OSMOSE [@shin2001], Ecopath with Ecosim
+[@christensen2004], POSEIDON [@bailey2019], DISPLACE [@bastardie2013], or
+SEAPODYM [@lehodey2008] are capable of representing tremendous amounts of
+complexity, but can be extremely time consuming to design and run. `marlin`
+allows some of the realism of these more end-to-end models while being simpler
+and faster to parameterize and run.
+
+Tools such as virtualspecies [@leroy2016], STEPS [@visintin2020], RangeShifter
+[@bocedi2014], and
+[SMS](https://www.aquaveo.com/software/sms-surface-water-modeling-system-introduction)
+can be efficiently constructed to model the dispersal and distribution of
+species in space and time as a function of environmental covariates. However,
+these typically require specifying a covariate-response curve without explicitly
+acknowledging how this arises from habitat-specific movement and demography
+(virtualspecies), or model movement using a dispersal kernel (RangeShifter,
+STEPS) or a least-cost path algorithm (SMS, STEPS). By contrast, `marlin` uses a
+continuous-time Markov chain movement model, which integrates multiple paths
+(including their path-dependent probability based on intervening habitat types
+and species preferences) while using scale-free parameters that can be measured
+experimentally in laboratory or tagging studies. In addition, `marlin` allows
+for simulating not only the distribution of the species but also biomass, age,
+and length structure in space and time.
+
+Simplified bio-economic models such as those used in @hastings2017, @sala2021,
+and @cabral2019 can be applied at scale and provide analytically tractable
+results, but as a result must abstract over many bio-economic dynamics that can
+be important for more tactical applications. Our results demonstrate how nuances
+in fleet dynamics and species distributions can dramatically impact MPA
+outcomes. `marlin` allows for more realistic representations of spatial
+social-ecological systems while maintaining processing speed.
+
+`marlin` does not represent trophic interactions. Anthropgenic changes in
+species abundance can result in trophic cascades.
+However, studies such as @gilman2020, @ovando2021a, @bruno2019 , and
+@malakhoff2021 found no clear signs of MPA driven trophic cascades within the
+first decades of protection. Signals of trophic cascades may be masked by
+variations in the direction and strength of species interactions driven by
+environmental context [@liu2022], or may simply take longer to develop
+detectable effects than the coverage of many time series of MPAs. While `marlin`
+does not incorporate trophic interactions, what empirical evidence we have does
+not suggest that management-mediated trophic cascades are so common and clear
+that they must be incorporated into any credible multi-species simulation model.
+There is value in capturing other dynamic bio-economic processes, as done here,
+even if that comes at the expense of modeling trophic interactions. However,
+research on the trophic impacts of spatial-temporal management actions is
+clearly of value and models like Atlantis [@audzijonyte2019], Ecopath with
+Ecosim [@christensen2004], and EASI-Fish [@griffiths2019] can help users explore
+those kinds of trophic processes, in the manner of @baskett2007.
+
+### General Recommendations for Use
+
+Marine social-ecological systems are highly complex. The model presented here is
+designed to help users explore the impacts of different variables on policy
+outcomes. But, that freedom comes at a cost, namely that users have a large
+number of options at their disposal that they must decide on. Our recommended
+strategy is for users to try and narrow down a list of parameters that they feel
+are sufficiently "known" and another list of parameters that are highly
+uncertain and / or contentious which the users feel may impact results. Where
+possible, parameters from locally estimated stock assessment models can be used
+to provide a foundation around which sensitivity analyses around specific
+uncertainties of interest [@berger2017], though care should be taken interpreting population selectivity curves estimated from stock assessment as contact selectivity curves required as inputs to `marlin` [@sampson2014].
+
+For example, a community seeking to a model a well-studied coastal coral reef
+ecosystem might leave as fixed the habitat distribution (as represented by reef
+locations) and general life history of the species in question (growth rates,
+age at maturity, etc). But, within those constraints, the users may wish to test
+the sensitivity of proposed MPA networks to key unknowns such as adult and
+larval dispersal rates or the the economic incentives and contact selectivity of
+the fishing fleets. As an example of this, we ran an alternative version of our
+coral reef case study in which Fleet Two was assigned logistic selectivity for
+the snapper and grouper species, rather than the dome-shaped selectivity
+presented in our main results. We found that the impacts of MPAs on conservation
+and food security were relatively insensitive to the form of contact selectivity
+specified for Fleet Two (Fig.S1). This does not mean that knowing the "correct"
+form of contact selectivity may not be extremely important in providing accurate
+assessment results [@waterhouse2014], but rather that in this case misspecifying
+the contact selectivity curve is projected to have little impact on the
+simulated outcomes of MPAs, conditional on holding other parameters constant.
+
+Given a set of model runs, the choice of whether to use those results
+"strategically" or "tactically" will depend on the needs of the user. That being
+said, since. `marlin` is not fit to data directly it is more easily applied to
+strategic questions. For example, the two case studies presented in this paper
+provide general strategic insights into the sensitivity of outcomes to design
+choices range shifts. The extent to which users are comfortable using the
+outputs of `marlin` tactically will depend on the confidence they have in their
+parameterization of the model relative to the precision policy-makers
+require in order to make a decision.
+
+`marlin` is just one model out of many that can be used to inform marine
+management. Other models may be better suited to address specific forms of
+complexity. Particularly for more tactical applications, we would encourage
+users to explore multiple modeling frameworks to help design policies that are
+likely to be robust to many different kinds of complexities.
+
+At this time, `marlin` assumes that all evaluated policies are perfectly
+implemented; e.g. that MPAs are 100% no-take and perfectly enforced, that quotas
+and closed seasons are respected, that there is no discard mortality, etc. In
+reality no policy is perfectly implemented, and users should consider the extent
+to which the policies they simulate are actually feasible to implement. There is
+value though in being able to simulate and compare the outcomes of perfectly
+implemented policies to isolate the concept of the policy itself from its
+implementation. That being said, future work may add the capability for
+imperfect implementation to the model.
+
+Fish populations often exhibit variation in demographic traits such as growth,
+maturation, and mortality across dimensions such as space, time, and sex. Models
+such as Stock Synthesis [@methotjr.2013] capture these processes through the use
+of "morphs". `marlin` does not currently allow for these kinds of processes
+explicitly, though users could simulate different populations of the same
+species with different life history traits as separate stocks. Users should be
+cautious interpreting simulation results from `marlin` for species in which
+these sorts of dynamics are likely to be particularly prevalent. For strongly
+sexually dimorphic species, we would recommend picking the sex most likely to
+drive the outcomes of management policies, which is often females given the
+general prevalence of eggs as the limiting reproductive material in marine
+ecosystems. Users should also proceed with caution using `marlin` for species
+with more complex reproductive biology, such as sex-changing fish
+[@kindsvater2017].
+
+## Conclusions
+
+`marlin` complements the existing spatial marine modeling literature by allowing
+scientists, decision makers, and other stakeholders to efficiently examine the
+impacts of realistic bio-economic dynamics on academic and applied problems. For
+example, scientists could use `marlin` to consider the relative importance of
+different biological variables in affecting policy outcomes. Managers could use
+`marlin` to assess the potential of proposed MPA networks to achieve stated
+objectives, such as maximizing conservation gains for a threatened species while
+minimizing or removing negative impacts to affected fishing fleets. We envision
+`marlin` being applicable to research on dynamic ocean management, range shifts,
+management strategy evaluation, policy interactions, and spatial stock
+assessment. `marlin` can help researchers generate data for further testing of
+the performance and design of spatially explicit integrated population models,
+in the manner of @bosley2022. In addition, the process-based movement model used
+in `marlin` can directly use empirical estimates of movement dynamics derived in
+the manner of @thorson2021a, providing a link between empirical and simulation
+based approaches to marine resource management that has been challenging to
+implement in spatial simulations [@berger2017].
+
+Fisheries models, assessments, management have often abstracted away many of the
+spatial-temporal complexities of marine social-ecological systems [@ovando2021;
+@berger2017]. The modeling framework described here can help facilitate the
+science and application of spatial fisheries management by supporting the
+simulation of different spatial-temporal dynamics to aid in testing of various
+aspects of the marine resource management process.
+
+Even the most complex marine model is a stylized cartoon of the true dynamics of
+ocean ecosystems. However, for all their limitations, models
+can help users understand factors that drive the performance of marine
+management strategies. The goal of this tool is to empower people to design
+policies based on evaluation of key uncertainties and trade-offs, and in doing
+so support more effective and equitable marine resource management.
+
+## Acknowledgements
+
+This project was funded by the Waitt Foundation. We thank Dr. Alexa Fredston and
+two anonymous reviewers for helpful revisions and comments of this manuscript.
+
+## Data Availability Statement
+
+The code, novel code, data, and materials needed to reproduce the results and
+all aspects of this manuscript are publicly available at
+https://github.com/DanOvando/marlin-paper, with supporting novel code available
+at https://github.com/DanOvando/marlin/. Upon acceptance, all data, materials,
+code, and novel code needed to reproduce the results and all aspects of this
+manuscript will be publicly available via GitHub at
+https://github.com/DanOvando/marlin-paper and
+https://github.com/DanOvando/marlin/ and through figshare at
+https://figshare.com/articles/preprint/marlin-paper/21843582
+
+## References
+
+::: {#refs}
+:::
+
+## Figure Legends
+
 ```{r}
@@
 Numbers show the approximate location of each fleet's (one or two) port.
-Columns indicate quarterly seasons that repeat each year."
+Columns indicate quarterly seasons that repeat each year.
+"
 
@@
 
- patch_biomass %>%
+ fig_1 <- patch_biomass %>%
   group_by(critter, step) %>%
@@
 
+ ggsave(
+   filename = file.path(results_path, "fig_1.pdf"),
+   fig_1,
+   width = small_width,
+   units = "mm"
+ )
+
+ fig_1
+
 ```
-
-#### Pelagic Fishery
-
-```{r}
-blue_fauna <- blue_water_fauna_and_fleets$fauna
-
-blue_fleets <- blue_water_fauna_and_fleets$fleets
-
-blue_resolution <- sqrt(blue_fauna[[1]]$patches)
-
-```
-
-We model our pelagic case study loosely on the characteristics of the Western
-and Central Pacific Ocean (WCPO) tuna fisheries. We simulate trajectories of
-`r n_distinct(names(blue_fauna))` species commonly caught in the region,
-including both the highly abundant skipjack tuna and the heavily depleted
-oceanic whitetip shark (*Carcharhinus longimanus*) (@fig-blue-ssb). We use
-publicly available data on catch-per-unit-effort of each of these species from
-the WCPO as a very rough proxy for baseline habitat distributions, noting that
-where possible, fishery-independent abundance indices would be preferable
-(@fig-blue).
-
-These pelagic species are caught by a longline fleet that primarily targets
-large adult tunas such as bigeye and yellowfin for high-grade consumption, and a
-purse seine fishery that primarily targets skipjack tunas for bulk canning.
-Contact selectivities were modeled as logistic for the longline fleet, and
-dome-shaped for the purse seine fleet, with selectivity values set to broadly
-match the patterns reported in relevant [stock
-assessments](https://oceanfish.spc.int/en/ofpsection/sam/sam/210-stock-assessment-input-and-result-files-)
-for each species. Both fleets operate under open-access dynamics with an effort
-cap. The effort cap was set at the level of effort that resulted in the desired
-levels of SSB/SSB0 for each species under open-access dynamics (@fig-blue),
-intended to simulate a scenario where managers step in to prevent further
-expansion of fishing effort in a fully developed fishery. For
-forward-simulation, open-access dynamics can result in effort decreasing in
-response to profitability, but cannot result in effort beyond the effort cap set
-for each fleet.
-
-For this exercise, we focused on using `marlin` to assess resilience of the
-selected *Target Fishing* and *Rate* MPA design strategies to a climate-driven
-range shift. Specifically, we simulate an extreme example where the centroid of
-each population shifts northward at a rate of 0.2 patches per year over a 20
-year time horizon (@fig-blue-ssb). We designed MPA networks given the conditions
-in the starting year, and then held that network constant over the years of the
-experiment, running one simulation with and another without the climate-drive
-range shift. We then compared the effects of this range shift on food production
-and conservation outcomes from MPA networks designed based on the pre-range
-shift world. We ran the pelagic simulation at a quarterly level ($\Delta_t$ =
-1/4 year), and set the area of each cell to be roughly 97,000 KM^2^ across 144
-patches each with a side length of roughly 311 KM, for a total area of
-14,000,000 KM^2^, broadly commensurate with the area of the WCPO.
 
@@
 #| label: fig-blue-ssb
-#| fig-cap: "Spawning biomass divided by unfished spawning biomass in the time period prior to implementation of MPAs for the pelagic case study."
+#| fig-cap: "Spawning biomass divided by unfished spawning biomass in the time period prior to implementation of MPAs for the pelagic case study.
+Simulated species include skipjack tuna, yellowfin tuna (*Thunnus albacares*), bigeye tuna, shortfin mako (*Isurus oxyrinchus*), swordfish (*Xiphias gladius*), albacore tuna (*Thunnus alalunga*), blue shark (*Prionace glauca*), silky shark (*Carcharhinus falciformis*), and oceanic whitetip shark (*Carcharhinus longimanus*)."
 
@@
 
-critter_biomass %>%
+fig_2 <- critter_biomass %>%
   ungroup() %>% 
@@
   coord_flip()
+
+ ggsave(
+   filename = file.path(results_path, "fig_2.pdf"),
+   fig_2,
+   width = small_width,
+   units = "mm"
+ )
+
+fig_2
 
@@
 
-range_shift_plot <- sq_habitat | rs_habitat
+range_shift_plot <- fit_3 <- sq_habitat | rs_habitat
+
+ ggsave(
+   filename = file.path(results_path, "fit_3.pdf"),
+   fit_3,
+   width = big_width,
+   units = "mm"
+ )
 
@@
 ```
-
-## Results
-
-### Coastal Coral Reef Fishery
-
-```{r}
-#| label: coral-results
-
-coral_mpa_experiments$mpas <- map(coral_mpa_experiments$results, "mpa")
-
-coral_mpa_experiments$obj <- map(coral_mpa_experiments$results, "obj")
-
-coral_results <- coral_mpa_experiments %>%
-  unnest(cols = obj)
-
-coral_fleet_results <- coral_results %>%
-  pivot_longer(starts_with("fleet_"),
-               names_to = "fleet",
-               values_to = "fleet_yield") %>%
-  group_by(prop_mpa, fleet, placement_strategy) %>%
-  summarise(yield = sum(fleet_yield), biodiv = sum(unique(biodiv))) %>%
-  group_by(fleet, placement_strategy) %>%
-  mutate(delta_yield = yield / yield[prop_mpa == 0] - 1) |> 
-  ungroup()
-
-peak_coral_fleet_results <- coral_fleet_results |> 
-  group_by(fleet) |> 
-  filter(delta_yield == max(delta_yield)) |> 
-  mutate(peak_yields = scales::percent(delta_yield))
-
-tmp <- coral_results %>% 
-  group_by(prop_mpa, placement_strategy) %>% 
-  summarise(biodiv = sum(biodiv), yield = sum(yield)) %>% 
-  group_by(placement_strategy) %>% 
-  mutate(delta_yield = yield / yield[prop_mpa == 0] - 1,
-         delta_biodiv = biodiv / biodiv[prop_mpa == 0] - 1) %>% 
-  ungroup() %>% 
-    mutate(placement_strategy= fct_relabel(placement_strategy, titler))
-
-```
-
-MPAs were capable of producing a range of positive and negative outcomes for
-food security and conservation in the coral reef case study depending on the
-design strategy used and the size of MPA implemented. Both of the MPA design
-strategies were capable of increasing fisheries yield for the more "industrial"
-Fleet Two, up to a value of
-`r peak_coral_fleet_results$peak_yields[peak_coral_fleet_results$fleet == "fleet_two"]`,
-even when MPAs covered more than 50% of the simulated area. However, the more
-"artisanal" Fleet One only benefited from MPAs under the *Spawning Ground*
-design strategy, with a maximum increase of
-`r peak_coral_fleet_results$peak_yields[peak_coral_fleet_results$fleet == "fleet_one"]`;
-the *Target Fishing* design strategy produced a roughly linear decrease in
-fishing yields as a function of increasing MPA size (@fig-coral-results A).
-
-On the conservation front, MPAs were uniformly beneficial to the spawning
-biomass of all species under the *Target Fishing* design strategy, with the most
-rapid increases in spawning biomass for the deep snapper population. The
-*Spawning Ground* strategy primarily benefited the shallow snapper population,
-producing little change in the grouper population and decreasing spawning
-biomass of both the deep snapper and reef shark populations for MPA sizes
-covering less than 50% of the simulated area (@fig-coral-results B).
-
-In total, the *Spawning Ground* strategy was capable or providing net increases
-in fishing yield (summed across both fleets) of up to roughly 7%, with positive
-net yield impacts up to network size of nearly 60% of the area. However, even
-MPAs covering 60% of the area only produced a maximum increase of roughly 5% in
-total SSB/SSB0 when designed around the spawning ground. Conversely, the *Target
-Fishing* design strategy was capable of producing a nearly 25% increase in total
-SSB/SSB0 for the same MPA size, but at a much greater cost to the total food
-production from the system's fisheries (@fig-coral-results C).
-
-<!-- ~~These macro-level changes in total yield and conservation are the result of -->
-
-<!-- interesting internal dynamics that marlin allows users to explore through output -->
-
-<!-- tables at the level of individual species and fleets. The *Spawning Ground* -->
-
-<!-- strategy increased yield to Fleet One when MPAs were relatively small, but was -->
-
-<!-- able to maintain yields near *status quo* levels up to a very large MPA size to -->
-
-<!-- the dependence of Fleet One on only a relatively small portion of the total -->
-
-<!-- simulation area. However, *Spawning Ground* MPAs were unable to provide -->
-
-<!-- increases in yield to Fleet Two. The *Spawning Ground* strategy produced the -->
-
-<!-- most rapid conservation gains for the snapper population. Interestingly, due to -->
-
-<!-- effort displacement from the spawning grounds onto the fishing areas inhabited -->
-
-<!-- by the groupers the rest of the year, closure of the spawning ground had limited -->
-
-<!-- impact on the grouper population until the MPA size exceeded 50% of the -->
-
-<!-- simulation area. Effort displacement from the spawning grounds also resulted in -->
-
-<!-- a small decrease in the deep snapper population for MPA sizes less than 50%. The -->
-
-<!-- *Target Fishing* design strategy has similar yield outcomes to the *Spawning -->
-
-<!-- Ground* strategy, but less dramatic negative impacts on Fleet Two, at the -->
-
-<!-- expense of slightly fewer benefits for Fleet One. However, the *Target Fishing* -->
-
-<!-- strategy provided more uniform conservation benefits to all species -->
-
-<!-- (@fig-coral-results).~~ -->
 
@@
     legend.background = element_rect(fill = "transparent"),
-    legend.text = element_text(size = 8)
+    legend.text = element_text(size = 10)
   )
@@
 
-((coral_fleet_yield / coral_critter_bio) | (coral_frontier + plot_layout(guides = "keep")) ) & theme(strip.text = element_text(size  = 6), plot.margin = margin(.05, .05, .05, .05, "cm"), panel.spacing = unit(0.25, "lines"), legend.text = element_text(size = 6))
-
-```
-
-### Pelagic Fishery
-
-Yields of the purse-seine fleet were more resilient to the range shift under the
-*Rate* design strategy, while the longline fleet had the opposite result. Under
-the *Rate* strategy, longline yields were relatively stable across a large range
-of MPA sizes under the status quo conditions, but declined rapidly under the
-range shift conditions. For the purse-seine fleet, yields were more stable under
-the range shift conditions under the *Target Fishing* design scenario, but
-declined quickly as a function of MPA size under the status quo conditions
-(@fig-blue-results A). The *Target Fishing* strategy produced better yield
-outcomes for both fishing fleets under the range shift conditions, but the
-*Rate* strategy performed best under the status quo.
-
-The primary tuna species (bigeye, skipjack, and yellowfin) were most sensitive
-to design strategy and climate scenario, with the *Target Fishing* strategies
-producing for example rapid conservation gains for bigeye and skipjack under the
-status quo, while the *Rate* strategy resulted in small net conservation losses
-for both species until the MPA became extremely large. However, under the range
-shift scenario the MPAs had little impact on many of the tuna populations until
-the MPA size became extremely large, due to movement of the primary fishing
-grounds outside of the current hotpots where the MPAs are placed based on status
-quo conditions. The 20 year time horizon simulated here was not enough to
-produce substantial gains for any of the shark species even with 100% closures
-except for the more rapidly growing blue shark (@fig-blue-results B).
-
-```{r}
-#| include: false
-
-blue_water_experiments$mpas <- map(blue_water_experiments$results, "mpa")
-
-blue_water_experiments$obj <- map(blue_water_experiments$results, "obj")
-
-blue_water_climate_experiments$mpas <- map(blue_water_climate_experiments$results, "mpa")
-
-blue_water_climate_experiments$obj <- map(blue_water_climate_experiments$results, "obj")
-
-static_results <- blue_water_experiments %>%
-  unnest(cols = obj) %>%
-  mutate(climate = FALSE)
-
-climate_results <- blue_water_climate_experiments %>%
-  unnest(cols = obj) %>%
-  mutate(climate = TRUE)
-
-blue_water_results <-static_results %>%
-  bind_rows(climate_results) %>% 
-  left_join(sci_to_com, by = "critter") %>% 
-  select(-critter) %>% 
-  rename(critter = common_name)
-
-fleet_climate_results <- blue_water_results %>%
-  select(placement_strategy,
-         fleet_model,
-         prop_mpa,
-         climate,
-         longline,
-         critter,
-         purseseine) %>%
-  rename(`purse-seine` = purseseine) |> 
-  pivot_longer(c(longline,`purse-seine`), names_to = "fleet", values_to = "yield") %>%
-  pivot_wider(names_from = "climate", values_from =  "yield") %>%
-  mutate(delta = `TRUE` - `FALSE`) %>%
-  group_by(placement_strategy, prop_mpa, fleet) %>%
-  summarise(delta = sum(delta))
-
-fleet_climate_impacts <- fleet_climate_results %>%
-  ggplot(aes(prop_mpa, delta, color = fleet)) +
-  geom_hline(yintercept = 0, linetype = 2) +
-  geom_line() +
-  facet_wrap(~placement_strategy) +
-  scale_y_continuous(name = "Difference in Yield") +
-  theme(axis.text.x = element_blank(), axis.title.x = element_blank())
-
-critter_climate_results <- blue_water_results %>%
-  select(placement_strategy,
-         fleet_model,
-         prop_mpa,
-         climate,
-         critter,
-         biodiv) %>%
-  pivot_wider(names_from = "climate", values_from =  "biodiv") %>%
-  mutate(delta = `TRUE` - `FALSE`)
-
-critter_climate_impacts <- critter_climate_results %>%
-  ggplot(aes(prop_mpa, delta, color = critter)) +
-  geom_line() +
-  facet_wrap(~placement_strategy) +
-  scale_x_continuous(name = "MPA",labels = scales::label_percent()) +
-  scale_y_continuous(name = "Difference in SSB/SSB0")
-
-fleet_climate_impacts / critter_climate_impacts
+fig_4 <- ((coral_fleet_yield / coral_critter_bio) | (coral_frontier + plot_layout(guides = "keep")) ) & theme(strip.text = element_text(size  = 6), plot.margin = margin(.05, .05, .05, .05, "cm"), panel.spacing = unit(0.25, "lines"), legend.text = element_text(size = 8))
+
+ ggsave(
+   filename = file.path(results_path, "fig_4.pdf"),
+   fig_4,
+   width = big_width,
+   units = "mm"
+ )
+
+fig_4
 
@@
 # blue_frontier / blue_critter_bio + plot_layout(heights = c(1,2))
-(blue_fleet_yield / 
+fig_5 <- (blue_fleet_yield / 
 blue_critter_bio) + plot_layout(heights = c(1.5,2))
+
+ ggsave(
+   filename = file.path(results_path, "fig_5.pdf"),
+   fig_5,
+   width = big_width,
+   units = "mm"
+ )
+
+fig_5
 ```
-
-## Discussion
-
-Marine ecosystems are driven by complex social-ecological dynamics. Communities
-must often make decisions on how to manage these systems based on limited
-empirical evidence. Modeling tools such as the one presented here can help users
-answer scientific questions and design marine management policies informed by a
-better understanding of sensitivities to key uncertainties. In this paper we
-have documented the core system of equations underlying `marlin`, and
-demonstrated how `marlin` can be used to support marine resource management.
-
-### Insights from Case Studies
-
-The variation in both the magnitude and distribution of costs and benefits
-stemming from alternative design choices and states of nature shown in our case
-studies demonstrates the value in our method. For example, our coastal coral
-reef example illustrates both the potential for MPAs to benefit conservation and
-food production in these systems, and the potential for the same MPA to benefit
-some fleets and species while harming others when fishing fleets affected by an
-MPA do not share the same objectives and species are not uniformly distributed.
-Our pelagic case study illustrates how `marlin` can be used to assess the
-climate resiliency of spatial management strategies, and identify strategies
-that best meet the needs of both current and future conditions. Results from
-this type of work could be used to help prioritize communities and species at
-particular risk to climate change impacts.
-
-Our coastal coral reef case study results show the sensitivity of conservation
-and food security outcomes of MPAs to alternative design scenarios, holding
-constant other social-ecological variables. The *Spawning Ground* MPA network
-was able to provide more equitable yield outcomes across the two fishing
-communities, whereas the *Target Fishing* MPA network only benefited the more
-industrial Fleet Two (@fig-coral-results A). This is because Fleet One has two
-primary fishing grounds; the spawning grounds, and the offshore area where the
-deep snapper live. Fleet Two primarily fishes in the northern portions of the
-simulation grid. Under the *Spawning Ground* strategy, while Fleet One quickly
-losses fishing grounds on the spawning grounds, it can compensate for this by
-moving offshore and fishing the deep snapper population harder, while also being
-able to fish some of the spillover of snapper from the spawning ground closure.
-This displacement of Fleet One's fishing effort is why biomass of deep snapper
-actually declines under the *Spawning Ground* strategy.
-
-Conversely, most of the catch in the fishery comes from the nearshore and deep
-water snapper populations, which overlaps with both of Fleet One's fishing
-grounds. The *Target Fishing* strategy then begins closing both the nearshore
-and offshore snapper fishing grounds, resulting in too much of a loss in fishing
-grounds for Fleet One to be offset by spillover from the MPAs. As a result of
-primarily placing MPAs on Fleet One's fishing grounds, Fleet Two gains spillover
-benefits at little cost to their fishing grounds until the MPAs reach their
-fishing grounds once protection nears 100%.
-
-By protecting both the spawning ground and the offshore areas, the *Target
-Fishing* network provides conservation benefits to all of the species, in
-contrast to the *Spawning Ground* strategy that only provides meaningful
-conservation gains to the snapper population up until very large MPA sizes. Our
-result that protection of a dedicated spawning ground did not provide
-substantially greater conservation outcomes for the species using that spawning
-ground (groupers and sharks) than an alternative design strategy is supported by
-other modeling studies that show that the impacts of spawning ground protection
-on conservation and yields may be highly variable, and that displacement of high
-levels fishing effort from the spawning grounds can offset potential
-conservation gains of the protection [@grüss2014]. That being said there is also
-evidence for the benefits of spawning aggregation protection [@erisman2015] ,
-and our results are not nearly sufficiently resolved to provide any general
-statements as to the relative value of spawning aggregation protection relative
-to other design strategies.
-
-The two design strategies produced very different outcomes in terms of total
-changes in yield and conservation outcomes for the coral reef case study
-(@fig-coral-results C). Both MPA network design strategies were capable of
-producing net "win-win" outcomes in which both food security (as measured by
-change in total fisheries yields) and conservation (as measured by change in
-SSB/SSB0). However, in general for the same size MPA network the *Spawning
-Ground* design strategy produced better food security outcomes but worse
-conservation outcomes, and *vice* *versa* for the *Target Fishing* strategy. The
-coral reef case study shows how the modeling framework presented here can help
-stakeholders explore how different management strategies affect outcomes both in
-total (@fig-coral-results C) and across species (@fig-coral-results A) and
-fishing communities (@fig-coral-results B).
-
-Turning to our Pelagic Ecosystem case study, the yield outcomes of the
-purse-seine fleet were much more sensitive to the presence of a range shift than
-the longline fleet, particularly under the *Target Fishing* design strategy.
-This is because the purse-seine fleet primarily targets skipjack tuna, which in
-this simulation are concentrated in a relatively narrow latitudinal band
-(@fig-blue), and the purse-seine fleet makes up a large portion of the total
-catch in the simulated fishery. So, the "Target Fishing" strategy starts by
-closing off the main purse-seine fishing grounds, which while only small part of
-the spatial domain of the model represents a large portion of the purse-seine
-fleet's fishing grounds, resulting both in more rapid conservation gains and
-fishery losses under the status quo species distributions. The *Rate* strategy
-places more MPAs in areas that are of lesser importance fo the purse-seine fleet
-but overlap more with species such as oceanic whitetip shark. Conversely, the
-purse-seine fleet appears to do have better MPA yield outcomes under the range
-shift scenario, not because of rebuilding of the skipjack population, but due to
-future fishing grounds being essentially unprotected by smaller MPAs targeting
-the current skipjack distribution ( @fig-blue-results ).
-
-The conservation outcomes of the skipjack, bigeye, and yellowfin tunas were
-among the most sensitive to range shifts, whereas even complete closure of the
-region was not sufficient to see significant increases in severely depleted
-species like oceanic whitetip sharks within the 20-year timeline of the
-simulation. Future research could be conducted then to see what sorts of
-timelines might be needed to see recovery of these species, and which design
-strategies result in the fastest recovery at the lowest cost to other objectives
-such as food security.
-
-These case studies are intended to illustrate the capabilities of the model and
-the importance of considering the bio-economic dynamics represented in `marlin`;
-it is beyond the scope of this paper to provide broader conclusions about the
-performance of MPAs under different contexts or a comprehensive comparison of
-simulation and empirical results of MPAs around the world. That being said, the
-kinds of dynamics resulting in our case studies are well supported by both
-modeling and empirical studies. Our results support conceptual [@hilborn2004a;
-@gaines2010] and empirical [@ban2019] evidence that under the right conditions
-no-take MPAs can benefit fisheries and conservation. @rassweiler2014
-demonstrated that the kinds of design choices presented in our case studies can
-greatly drive MPA outcomes.
-
-MPAs based around coral-reef style ecosystems have been extensively studied
-around the world, with much of the empirical evidence of their performance
-centered on demonstrating higher metrics such as biomass densities inside
-protected areas relative to fished reference areas ["response ratios",
-@lester2009]. Our model predicts a similar rapid increase of the simple ratio of
-mean biomass inside MPAs relative to outside (Fig. S5, acknowledging that
-designing a proper response ratio would require controlling for habitat
-characteristics and MPA design criteria). However, despite producing clear
-response ratios, the net conservation and fishery outcomes of our simulated
-coral-reef ecosystem MPAs were not nearly as large or clear as the response
-ratio results (@fig-coral-results); this result is supported by works such as
-@ferraro2018 an @ovando2021a showing that response ratios alone may be a poor
-indicator of the net causal impacts of MPAs at the population scale.
-
-Studies such as @gilman2020, @abbott2012, @davies2018 and @pons2022 provide
-empirical support for our case study results showing that MPAs can produce
-trade-offs across the conservation of different species and fleets when the
-spatial-temporal distributions and life history traits of affected species are
-heterogeneous, and MPAs result in some degree of effort displacement or
-concentration in the remaining fishing grounds. @hampton2023 supports our
-pelagic case study result showing that both conservation and fishery impacts of
-MPAs on the highly mobile species of the open oceans can be limited unless
-protected areas are very large. @brown2018 supports our result showing how
-heterogeneity in the behavior and objectives of fishing fleets sharing an
-ecosystem can affect the magnitude and equity of fishery reform outcomes.
-@davies2017 discusses the importance of considering the potential of climate
-driven range shifts when assessing spatial management policies; our work builds
-on this by allowing users to not just simulate species distributions but also
-distribution of biomass and age composition of fish space and time under climate
-change.
-
-### Putting `marlin` In Context
-
-The role of this paper is not to conduct a review of the many tools for
-spatial-temporal modeling available in the literature, each of which provide
-useful functionality for specific applications. However, we highlight here the
-specific gaps that we feel `marlin` fills in the modeling literature and in the
-policy support toolbox using some selected publications. More end-to-end models
-such as ATLANTIS [@audzijonyte2019] , OSMOSE [@shin2001], Ecopath with Ecosim
-[@christensen2004], POSEIDON [@bailey2019], DISPLACE [@bastardie2013], or
-SEAPODYM [@lehodey2008] are capable of representing tremendous amounts of
-complexity, but can be extremely time consuming to design and run. `marlin`
-allows some of the realism of these more end-to-end models while being simpler
-and faster to parameterize and run.
-
-Tools such as virtualspecies [@leroy2016], STEPS [@visintin2020], RangeShifter
-[@bocedi2014], and
-[SMS](https://www.aquaveo.com/software/sms-surface-water-modeling-system-introduction)
-can be efficiently constructed to model the dispersal and distribution of
-species in space and time as a function of environmental covariates. However,
-these typically require specifying a covariate-response curve without explicitly
-acknowledging how this arises from habitat-specific movement and demography
-(virtualspecies), or model movement using a dispersal kernel (RangeShifter,
-STEPS) or a least-cost path algorithm (SMS, STEPS). By contrast, `marlin` uses a
-continuous-time Markov chain movement model, which integrates multiple paths
-(including their path-dependent probability based on intervening habitat types
-and species preferences) while using scale-free parameters that can be measured
-experimentally in laboratory or tagging studies. In addition, `marlin` allows
-for simulating not only the distribution of the species but also biomass, age,
-and length structure in space and time.
-
-Simplified bio-economic models such as those used in @hastings2017, @sala2021,
-and @cabral2019 can be applied at scale and provide analytically tractable
-results, but as a result must abstract over many bio-economic dynamics that can
-be important for more tactical applications. Our results demonstrate how nuances
-in fleet dynamics and species distributions can dramatically impact MPA
-outcomes. `marlin` allows for more realistic representations of spatial
-social-ecological systems while maintaining processing speed.
-
-`marlin` does not represent trophic interactions. Anthropgenic changes in
-species abundance can result in trophic cascades throughout affected food webs.
-However, studies such as @gilman2020, @ovando2021a, @bruno2019 , and
-@malakhoff2021 found no clear signs of MPA driven trophic cascades within the
-first decades of protection. Signals of trophic cascades may be masked by
-variations in the direction and strength of species interactions driven by
-environmental context [@liu2022], or may simply take longer to develop
-detectable effects than the coverage of many time series of MPAs. While `marlin`
-does not incorporate trophic interactions, what empirical evidence we have does
-not suggest that management-mediated trophic cascades are so common and clear
-that they must be incorporated into any credible multi-species simulation model.
-There is value in capturing other dynamic bio-economic processes, as done here,
-even if that comes at the expense of modeling trophic interactions. However,
-research on the trophic impacts of spatial-temporal management actions is
-clearly of value and models like Atlantis [@audzijonyte2019], Ecopath with
-Ecosim [@christensen2004], and EASI-Fish [@griffiths2019] can help users explore
-those kinds of trophic processes, in the manner of @baskett2007.
-
-### General Recommendations for Use
-
-Marine social-ecological systems are highly complex. The model presented here is
-designed to help users explore the impacts of different variables on policy
-outcomes. But, that freedom comes at a cost, namely that users have a large
-number of options at their disposal that they must decide on. Our recommended
-strategy is for users to try and narrow down a list of parameters that they feel
-are sufficiently "known" and another list of parameters that are highly
-uncertain and / or contentious which the users feel may impact results. Where
-possible, parameters from locally estimated stock assessment models can be used
-to provide a foundation around which sensitivity analyses around specific
-uncertainties of interest [@berger2017].
-
-For example, a community seeking to a model a well-studied coastal coral reef
-ecosystem might leave as fixed the habitat distribution (as represented by reef
-locations) and general life history of the species in question (growth rates,
-age at maturity, etc). But, within those constraints, the users may wish to test
-the sensitivity of proposed MPA networks to key unknowns such as adult and
-larval dispersal rates or the the economic incentives and contact selectivity of
-the fishing fleets. As an example of this, we ran an alternative version of our
-coral reef case study in which Fleet Two was assigned logistic selectivity for
-the snapper and grouper species, rather than the dome-shaped selectivity
-presented in our main results. We found that the impacts of MPAs on conservation
-and food security were relatively insensitive to the form of contact selectivity
-specified for Fleet Two (Fig.S1). This does not mean that knowing the "correct"
-form of contact selectivity may not be extremely important in providing accurate
-assessment results [@waterhouse2014], but rather that in this case misspecifying
-the contact selectivity curve is projected to have little impact on the
-simulated outcomes of MPAs, conditional on holding other parameters constant.
-
-Given a set of model runs, the choice of whether to use those results
-"strategically" or "tactically" will depend on the needs of the user. That being
-said, since. `marlin` is not fit to data directly it is more easily applied to
-strategic questions. For example, the two case studies presented in this paper
-provide general strategic insights into the sensitivity of outcomes to design
-choices range shifts. The extent to which users are comfortable using the
-outputs of `marlin` tactically will depend on the confidence they have in their
-parameterization of the model runs relative to the precision policy-makers
-require in order to make a decision.
-
-`marlin` is just one model out of many that can be used to inform marine
-management. Other models may be better suited to address specific forms of
-complexity. Particularly for more tactical applications, we would encourage
-users to explore multiple modeling frameworks to help design policies that are
-likely to be robust to many different kinds of complexities.
-
-At this time, `marlin` assumes that all evaluated policies are perfectly
-implemented; e.g. that MPAs are 100% no-take and perfectly enforced, that quotas
-and closed seasons are respected, that there is no discard mortality, etc. In
-reality no policy is perfectly implemented, and users should consider the extent
-to which the policies they simulate are actually feasible to implement. There is
-value though in being able to simulate and compare the outcomes of perfectly
-implemented policies to isolate the concept of the policy itself from its
-implementation. That being said, future work may add the capability for
-imperfect implementation to the model.
-
-Fish populations often exhibit variation in demographic traits such as growth,
-maturation, and mortality across dimensions such as space, time, and sex. Models
-such as Stock Synthesis [@methotjr.2013] capture these processes through the use
-of "morphs". `marlin` does not currently allow for these kinds of processes
-explicitly, though users could simulate different populations of the same
-species with different life history traits as separate stocks. Users should be
-cautious interpreting simulation results from `marlin` for species in which
-these sorts of dynamics are likely to be particularly prevalent. For strongly
-sexually dimorphic species, we would recommend picking the sex most likely to
-drive the outcomes of management policies, which is often females given the
-general prevalence of eggs as the limiting reproductive material in marine
-ecosystems. Users should also proceed with caution using `marlin` for species
-with more complex reproductive biology, such as sex-changing fish
-[@kindsvater2017].
-
-## Conclusions
-
-`marlin` complements the existing spatial marine modeling literature by allowing
-scientists, decision makers, and other stakeholders to efficiently examine the
-impacts of realistic bio-economic dynamics on academic and applied problems. For
-example, scientists could use `marlin` to consider the relative importance of
-different biological variables in affecting policy outcomes. Managers could use
-`marlin` to assess the potential of proposed MPA networks to achieve stated
-objectives, such as maximizing conservation gains for a threatened species while
-minimizing or removing negative impacts to affected fishing fleets. We envision
-`marlin` being applicable to research on dynamic ocean management, range shifts,
-management strategy evaluation, policy interactions, and spatial stock
-assessment. `marlin` can help researchers generate data for further testing of
-the performance and design of spatially explicit integrated population models,
-in the manner of @bosley2022. In addition, the process-based movement model used
-in `marlin` can directly use empirical estimates of movement dynamics derived in
-the manner of @thorson2021a, providing a link between empirical and simulation
-based approaches to marine resource management that has been challenging to
-implement in spatial simulations [@berger2017].
-
-Fisheries models, assessments, management have often abstracted away many of the
-spatial-temporal complexities of marine social-ecological systems [@ovando2021;
-@berger2017]. The modeling framework described here can help facilitate the
-science and application of spatial fisheries management by supporting the
-simulation of different spatial-temporal dynamics to aid in testing of various
-aspects of the marine resource management process.
-
-Even the most complex marine model is a stylized cartoon of the true dynamics of
-ocean ecosystems. However, for all their limitations, models such as `marlin`
-can help users understand factors that drive the performance of marine
-management strategies. The goal of this tool is to empower people to design
-policies based on evaluation of key uncertainties and trade-offs, and in doing
-so support more effective and equitable marine resource management.
-
-## Acknowledgements
-
-This project was funded by the Waitt Foundation. We thank Dr. Alexa Fredston and
-two anonymous reviewers for helpful revisions and comments of this manuscript.
-
-## Data Availability Statement
-
-The code, novel code, data, and materials needed to reproduce the results and
-all aspects of this manuscript are publicly available at
-https://github.com/DanOvando/marlin-paper, with supporting novel code available
-at https://github.com/DanOvando/marlin/. Upon acceptance, all data, materials,
-code, and novel code needed to reproduce the results and all aspects of this
-manuscript will be publicly available via GitHub at
-https://github.com/DanOvando/marlin-paper and
-https://github.com/DanOvando/marlin/ and through figshare at
-https://figshare.com/articles/preprint/marlin-paper/21843582
-
-## References
-
-::: {#refs}
-:::
-
-## Figure Legends

### 2023-10-25 04_marlin_paper_si.qmd
first look edits

@@
 ```
+
+```{r}
+#| label: fig-coral-ssb
+#| fig-cap: "Spawning biomass divided by unfished spawning biomass in the time period prior to implementation of MPAs for the coral reef case study."
+
+critter_biomass <-
+  map_df(coral_sim, ~ map_df(.x, ~tibble(ssb = rowSums(.x$ssb_p_a), patch = 1:nrow(.x$ssb_p_a)), .id = "critter"), .id = "step") %>%
+  mutate(step = marlin::clean_steps(step)) %>%
+  left_join(grid, by = "patch") %>% 
+  group_by(critter, step) %>% 
+  summarise(ssb = sum(ssb)) %>% 
+  group_by(critter) %>% 
+  mutate(depletion = ssb / ssb[step == 0])
+
+fig_s2 <- critter_biomass %>%
+  ungroup() %>% 
+  filter(step == max(step)) %>% 
+  ggplot(aes(reorder(critter, depletion), depletion)) +
+  geom_col() +
+  scale_y_continuous(limits = c(0, 1), name = "SSB/SSB0", expand = expansion(mult = c(0.0,0.05))) +
+  labs(x = "") + 
+  coord_flip()
+
+fig_s2
+```

### 2023-11-02 03_marlin_paper.qmd
pre-pub edits

@@
   html: default
-  docx: default
+  docx:
+    reference-doc: template.docx
 execute:
@@
 
-bio-economic modeling
-conservation planning
-fisheries
-marine protected areas
-marine spatial planning
-spatio-temporal modeling
+Bio-economic modeling; Conservation planning; Fisheries; Marine protected areas; Marine spatial planning; Spatio-temporal modeling
 
@@
 
+\newpage
+
 ## Introduction
@@
 Communities around the world are increasingly looking to spatial strategies for
-managing marine ecosystems . For example, the "30x30" movement calls for
+managing marine ecosystems. For example, the "30x30" movement calls for
 implementation of marine protected areas (MPAs), a form of spatial management,
@@
 policy structures. These dynamics pose challenges even for single-species
-assessment and management strategies , which are only amplified when we consider
-management policies specifically designed around multiple species, fleets, and
+assessment and management strategies, which are only amplified when we consider
+management policies designed around multiple species, fleets, and
 spatial features in the oceans [@field2006a].
@@
 [@rcoreteam2021] to facilitate use, while the underlying population model is
-written in C++, integrated through the `Rcpp` package [@eddelbuettel2018], for
-speed. Users on standard computers should be able to simulate one age-structured
-population distributed across two-dimensional spatial surface represented by a
-ten by ten grid of cells over 20 years in fractions of a second.
+written in C++, integrated through the `Rcpp` package [@eddelbuettel2018].
+Users on standard computers should be able to simulate one age-structured population distributed across two-dimensional spatial surface represented by a
+ten by ten grid of cells over 20 years in fractions of a second. It is very important to note that the parameters of
+this model cannot be fit to data directly within the `marlin` package; users
+must set model parameters themselves based on externally available data and
+their best judgement.
 
@@
 Below we provide a summary of the `marlin` model, as well as details of the case
-studies. We note that `marlin` is a complex software package with many options
-to control detailed behaviors. Given that these specific features will evolve
-over time, we do not describe in detail every option of the model here. Rather,
-we focus on the equations and methods describing the core foundations of our
-tool for efficiently simulating bio-economic dynamics in aquatic systems.
+studies.
 
@@
 movement towards preferred habitat), experience natural and potentially fishing
-mortality, and if in a spawning season spawn using one of the possible forms of
-density dependence implemented in the model, with the potential to include
+mortality, and potentially spawn using one of the possible forms of
+density dependence implemented in the model, with the ability to include
 auto-correlated stochastic deviations in the amount of offspring produced,
@@
 contact selectivity [@sampson2014] for individual species (each of which we
-denote as a *metier*). For example, both a longline and purse-seine fleet may
-capture bigeye (*Thunnus obesus*) and skipjack (*Katsuwonus pelamis*) tunas, but the longline fleet can
+denote as a *métier*). For example, both a longline and purse-seine fleet may
+capture bigeye (*Thunnus obesus*, Scombridae) and skipjack (*Katsuwonus pelamis*, Scombridae) tunas, but the longline fleet can
 be made much more likely to catch larger bigeye than skipjack, and *vice versa*.
@@
 "contact" with the gear. The contact selectivity of each of the specified
-*metiers* will then interact with the distribution of fish sizes and fishing
+*métiers* will then interact with the distribution of fish sizes and fishing
 effort in space simulated by `marlin` to produce a net "population" selectivity,
 which may differ from the individual contact selectivities of each of the
-*metiers* [@waterhouse2014; @sampson2014]. Each fishing fleet distributes its
+*métiers* [@waterhouse2014; @sampson2014]. Each fishing fleet distributes its
 fishing effort in space according to a specified spatial allocation rule (see
@@
 Unconstrained by management, the total amount of effort exerted by each fleet
-can follow one of two dynamics: open access or constant effort (see *Calculating
-Total Effort*). Under constant effort, the total amount of effort of each fleet
+can follow one of two dynamics: open access or constant effort (see *Calculating Total Effort*). Under constant effort, the total amount of effort of each fleet
 is fixed over time, with the possible exception of attrition due to MPA
@@
 reduced below this cap if required by the profit equation (i.e. the fleet can
-choose to fish less than the quota or effort cap).
-
-`marlin` also allows users to specify closed fishing seasons for one or more
-species in the system. When all species in a simulation share the same closed
-seasons, no fishing occurs in those seasons. When not all species share the same
-closed seasons, fishing still occurs in the closed seasons, but the fleet avoids
-catching the species for which fishing is closed in a given season, assuming no
-discard mortality. When closed seasons are in effect, the amount of effort in
-the next open season is a function of the effort and profits in the last open
-fishing season. Lastly, managers can specify locations of no-take MPAs, which
+choose to fish less than the quota or effort cap). `marlin` also allows users to specify closed fishing seasons for one or more
+species in the system. Lastly, managers can specify locations of no-take MPAs, which
 can change in size and location if needed. When MPAs are implemented, fishing
@@
 
-Users can use the model to explore general properties of spatially explicit
-social-ecological systems, or can parameterize it to explore key uncertainties
-of more tactical applications. It is very important note that the parameters of
-this model cannot be fit to data directly within the `marlin` package; users
-must set model parameters themselves based on externally available data and
-their best judgement.
-
-On the biological side, at a minimum users must for each species they wish to
-simulate supply the common or scientific name of the species in question, a
+On the biological side, at a minimum users must for each species being simulated supply the common or scientific name of the species in question, a
 measure of the level of fishing intensity the species is experiencing at the
@@
 species is a function of a range of parameters including the total amount of
-effort (*E*), the contact selectivity ogives ($\alpha$,the curve specifying the
-proportion of individuals at a given age or size that will be caught by a unit
-of fishing mortality), fishing cost (*C*), the spatial distribution of the
+effort (*E*), the contact selectivity ogives ($\alpha$), fishing cost (*C*), the spatial distribution of the
 species affected by the fleet, the relative prices across species ($\Pi$), and
@@
 revenues. Price data for use in the model can be obtained relatively simply
-through literature reviews, market surveys, or local experts, so long as they
-are converted to the same units of biomass used in the model. However, cost
+through literature reviews, market surveys, or local experts. However, cost
 parameters are more complicated, as translating say cost per day of fishing into
@@
 in question, and `marlin` will tune the cost parameter to achieve this desired
-cost to revenue ratio given the other parameters in the model (e.g. price and
-biomass of the target species). So for example, while users may not have access
-to exact cost parameters, they may have a general sense that when operating
-under ideal conditions a given fleet has a profit margin of 10%, i.e. the cost
-to revenue ratio is 90%. They can input the cost to revenue ratio as 0.9 then
-and allow `marlin` to tune the cost parameters accordingly.
+cost to revenue ratio given the other parameters in the model.
 
@@
 @gillis2003 and references therein). However, analytical solutions to the IFD
-present a number of complications for our model. Our model is capable of
-tracking multiple fleets at once. This means that the IFD for a single fleet
+present a number of complications for our model. The IFD for a single fleet
 would commonly be modeled as a Nash-equilibrium based on the actions of each of
@@
 on the outcomes in the prior timestep, meaning that the impacts of the actions
-of other fishing fleets are somewhat internalized. The timeline required for the
+of other fishing fleets are eventually accounted for. The timeline required for the
 fleets to reach an equilibrium distribution will vary and depend on factors such
-as the population dynamics of the species in question (e.g. movement, habitat
-heterogeneity, growth rates, density dependence) and the number and degree of
+as the population dynamics of the species in question and the number and degree of
 competition across fleets. Users should explore different simulation times to
@@
 in question (scientific preferred), the model will supply default values for
-that species based on the values reported in FishLife [internet connection
-required, @thorson2020]. FishLife provides estimates of core life history
+that species based on the values reported in FishLife [@thorson2020]. FishLife provides estimates of core life history
 parameters for fish species based on a model integrating published values and
@@
 species, but these default values should be not be taken as definitive and users
-should input best available parameters specific to the stock in question if they
-wish to best represent the dynamics of their specific system, and should check
-the plausibility of the default values prior to use. All of the life history
-values can be replaced with user-supplied values as needed and should be for
-more tactical applications; the default values simply ensure that reasonable and
-internally consistent values for the species in question are supplied as a
-starting point.
+should check the default values and input best available parameters specific to the stock in question if they
+wish to best represent the dynamics of their specific system.
 
@@
 non-adjacent. The CTMC model then produces movement dynamics that move around
-barriers rather than over them. This is different than setting habitat in a
-patch to zero; an animal will not preferentially move towards a patch with zero
-habitat all else being equal, but can move over a patch with zero habitat
-towards another better patch. An example might be a species that lives and feeds
-on coastal habitats transiting through a patch of open ocean to reach a new
-feeding ground. In contrast, a physical barrier represents something like a
-peninsula that an animal must move around to transit from one side to another.
+barriers rather than over them.
 
@@
 maximum sustainable yield (MSY) based reference points, though the exact level
-of stock status relative to MSY reference points desired by management or
-societal objectives may vary widely. MSY based reference points present a
+of stock status relative to MSY reference points desired by societal objectives may vary widely. MSY based reference points present a
 problem for a multi-fleet and spatial-temporal model such as `marlin`. MSY and
@@
 
-As such, we do not report MSY based reference points in the model by default,
-with the exception of the spawning stock biomass at MSY ($SSB_{MSY}$) which
-tends to be better conserved across selectivity patterns as it is primarily a
-function of the mortality and flow of recruits relative to the maturity ogive.
+As such, we do not report MSY based reference points in the model by default.
 There are many different strategies for estimating reference points in spatially
 explicit systems [@kapur2021a]. We leave it to users to define and find relevant
-reference point as required by their specific needs.
+reference points as required by their specific needs.
 
@@
 Each of the case studies contains too many parameters and options to be
-succinctly presented in the text here, and we are concerned that attempting to
-duplicate all of the parameters of the model within the text would simply
-introduce errors and confusion. As such, and since these results are for
-illustrative purposes only, readers should consult the accompanying code to view
+succinctly presented in the text here. Readers should consult the accompanying code to view
 the precise details of each simulation. Targeted applications must carefully
@@
 
-To our knowledge, MPAs are never randomly placed; their locations are decided
-based on stakeholder negotiation given available information and objectives. In
-theory, the design of MPA networks can be optimized through the use of a
+In theory, the design of MPA networks can be optimized through the use of a
 modeling framework, and depending on the validity of the model, this process is
@@
 one continuously connected block) and mosaic (MPAs can be separated in space)
-MPA designs [@pons2022] by developing and passing their own MPA design
-algorithms to the model.
+MPA designs [@pons2022].
 
@@
 In our coastal coral reef example, we model the dynamics of four tropical
-Pacific species: a grouper (*Epinephelus fuscoguttatus*), a shallow-reef snapper
-(*Lutjanus malabaricus*), a deep-reef snapper (*Pristipomoides filamentosus*),
-and a reef shark (*Carcharhinus amblyrhynchos*). The simulated groupers undergo
+Pacific species: a grouper (*Epinephelus fuscoguttatus*, Serranidae), a shallow-reef snapper
+(*Lutjanus malabaricus*, Lutjanidae), a deep-reef snapper (*Pristipomoides filamentosus*, Lutjanidae),
+and a reef shark (*Carcharhinus amblyrhynchos*, Carcharhinidae). The simulated groupers undergo
 a mass migration to a spawning aggregation once per year, followed by the
@@
 higher cost per distance coefficient than Fleet Two. Fleet One's home port is
-located on near the site of the grouper spawning grounds.
+located near the site of the grouper spawning grounds.
 
@@
 both MPA size and MPA design strategy. For this exercise, MPAs are placed with
-perfect information and have no design constraints for continuity (i.e. the
-selected MPA design can be as patchwork as dictated by the data). We ran the
+perfect information and have no design constraints for continuity. We ran the
 simulation for in quarter year time steps for 20 years ($\Delta_t = 1/4$) and
@@
 and Central Pacific Ocean (WCPO) tuna fisheries.
-Note that this is an illustrative example only and simulated stock status, species distributions, or outcome projections presented here should not be interpreted as a indicative of the current or future state of the WCPO.
+Note that this is an illustrative example only and simulated stock status, species distributions,  and projections presented here should not be interpreted as a indicative of the current or future state of the WCPO.
 We simulate trajectories of
@@
 including both the highly abundant skipjack tuna and the heavily depleted
-oceanic whitetip shark (*Carcharhinus longimanus*) (@fig-blue-ssb). We use
+oceanic whitetip shark (*Carcharhinus longimanus*, Carcharhinidae) (@fig-blue-ssb). We use
 publicly available data on catch-per-unit-effort of each of these species from
@@
 These pelagic species are caught by a longline fleet that primarily targets
-large adult tunas such as bigeye and yellowfin for high-grade consumption, and a
+large adult tunas such as bigeye and yellowfin (*Thunnus albacares*, Scombridae) for high-grade consumption, and a
 purse-seine fishery that primarily targets skipjack tunas for bulk canning.
 Contact selectivities were modeled as logistic for the longline fleet, and
-dome-shaped for the purse-seine fleet, with selectivity values set to broadly
-match the patterns reported in relevant [stock
-assessments](https://oceanfish.spc.int/en/ofpsection/sam/sam/210-stock-assessment-input-and-result-files-)
-for each species. Both fleets operate under open-access dynamics with an effort
+dome-shaped for the purse-seine fleet. Both fleets operate under open-access dynamics with an effort
 cap. The effort cap was set at the level of effort that resulted in the desired
@@
 range shift. Specifically, we simulate an extreme example where the centroid of
-each population shifts northward at a rate of 0.2 patches per year over a 20
+each population shifts northward at a rate of ~62km per year over a 20
 year time horizon (@fig-blue-ssb). We designed MPA networks given the conditions
@@
 patches each with a side length of roughly 311 KM, for a total area of
-14,000,000 KM^2^, broadly commensurate with the area of the WCPO.
+14e6 KM^2^, broadly commensurate with the area of the WCPO.
 
@@
 design strategy used and the size of MPA implemented. Both of the MPA design
-strategies were capable of increasing fisheries yield for the more "industrial"
+strategies were capable of increasing fisheries yield for
 Fleet Two, up to a value of
 `r peak_coral_fleet_results$peak_yields[peak_coral_fleet_results$fleet == "fleet_two"]`,
-even when MPAs covered more than 50% of the simulated area. However, the more
-"artisanal" Fleet One only benefited from MPAs under the *Spawning Ground*
+even when MPAs covered more than 50% of the simulated area. However Fleet One only benefited from MPAs under the *Spawning Ground*
 design strategy, with a maximum increase of
@@
 
-On the conservation front, MPAs were uniformly beneficial to the spawning
+MPAs were uniformly beneficial to the spawning
 biomass of all species under the *Target Fishing* design strategy, with the most
@@
 answer scientific questions and design marine management policies informed by a
-better understanding of sensitivities to key uncertainties. In this paper we
-have documented the core system of equations underlying `marlin`, and
-demonstrated how `marlin` can be used to support marine resource management.
+better understanding of sensitivities to key uncertainties.
 
@@
 
-The variation in both the magnitude and distribution of costs and benefits
-stemming from alternative design choices and states of nature shown in our case
-studies demonstrates the value in our method. For example, our coastal coral
-reef example illustrates both the potential for MPAs to benefit conservation and
+Our coastal coral reef example illustrates both the potential for MPAs to benefit conservation and
 food production in these systems, and the potential for the same MPA to benefit
@@
 
-Our coastal coral reef case study results show the sensitivity of conservation
-and food security outcomes of MPAs to alternative design scenarios, holding
-constant other social-ecological variables. The *Spawning Ground* MPA network
+Holding constant other social-ecological variables in the coral reef example, the *Spawning Ground* MPA network
 was able to provide more equitable yield outcomes across the two fishing
@@
 
-Conversely, most of the catch in the fishery comes from the nearshore and deep
+Conversely, most of the catch in the coral reef fishery comes from the nearshore and deep
 water snapper populations, which overlaps with both of Fleet One's fishing
@@
 (@fig-coral-results C). Both MPA network design strategies were capable of
-producing net "win-win" outcomes in which both food security (as measured by
-change in total fisheries yields) and conservation (as measured by change in
-SSB/SSB0). However, in general for the same size MPA network the *Spawning
-Ground* design strategy produced better food security outcomes but worse
+producing net "win-win" outcomes in which both food security and conservation.
+However, in general for the same size MPA network the *Spawning Ground* design strategy produced better food security outcomes but worse
 conservation outcomes, and *vice* *versa* for the *Target Fishing* strategy. The
@@
 performance of MPAs under different contexts or a comprehensive comparison of
-simulation and empirical results of MPAs around the world. That being said, the
-kinds of dynamics resulting in our case studies are well supported by both
+simulation and empirical results of MPAs around the world. That being said,
+the kinds of dynamics resulting in our case studies are well supported by both
 modeling and empirical studies. Our results support conceptual [@hilborn2004a;
@@
 SEAPODYM [@lehodey2008] are capable of representing tremendous amounts of
-complexity, but can be extremely time consuming to design and run. `marlin`
+complexity, but can be time consuming to design and run. `marlin`
 allows some of the realism of these more end-to-end models while being simpler
@@
 not suggest that management-mediated trophic cascades are so common and clear
-that they must be incorporated into any credible multi-species simulation model.
-There is value in capturing other dynamic bio-economic processes, as done here,
-even if that comes at the expense of modeling trophic interactions. However,
+that they must be incorporated into any credible multi-species simulation model. However,
 research on the trophic impacts of spatial-temporal management actions is
@@
 
-Marine social-ecological systems are highly complex. The model presented here is
-designed to help users explore the impacts of different variables on policy
-outcomes. But, that freedom comes at a cost, namely that users have a large
+The model presented here is designed to help users explore the impacts of different variables on policy
+outcomes. But, that freedom means that users have a large
 number of options at their disposal that they must decide on. Our recommended
@@
 locations) and general life history of the species in question (growth rates,
-age at maturity, etc). But, within those constraints, the users may wish to test
+age at maturity, etc). From there, users may wish to test
 the sensitivity of proposed MPA networks to key unknowns such as adult and
@@
 Given a set of model runs, the choice of whether to use those results
-"strategically" or "tactically" will depend on the needs of the user. That being
-said, since. `marlin` is not fit to data directly it is more easily applied to
-strategic questions. For example, the two case studies presented in this paper
-provide general strategic insights into the sensitivity of outcomes to design
-choices range shifts. The extent to which users are comfortable using the
+"strategically" or "tactically" will depend on the needs of the user.
+Since `marlin` is not fit to data directly it is more easily applied to strategic questions.
+The extent to which users are comfortable using the
 outputs of `marlin` tactically will depend on the confidence they have in their
 parameterization of the model relative to the precision policy-makers
-require in order to make a decision.
-
-`marlin` is just one model out of many that can be used to inform marine
-management. Other models may be better suited to address specific forms of
+require in order to make a decision. Other models may be better suited to address specific forms of
 complexity. Particularly for more tactical applications, we would encourage
@@
 implemented policies to isolate the concept of the policy itself from its
-implementation. That being said, future work may add the capability for
-imperfect implementation to the model.
-
-Fish populations often exhibit variation in demographic traits such as growth,
-maturation, and mortality across dimensions such as space, time, and sex. Models
+implementation.
+
+Fish populations often exhibit variation in demographic traits across dimensions such as space, time, and sex. Models
 such as Stock Synthesis [@methotjr.2013] capture these processes through the use
 of "morphs". `marlin` does not currently allow for these kinds of processes
-explicitly, though users could simulate different populations of the same
-species with different life history traits as separate stocks. Users should be
+explicitly. Users should be
 cautious interpreting simulation results from `marlin` for species in which
@@
 scientists, decision makers, and other stakeholders to efficiently examine the
-impacts of realistic bio-economic dynamics on academic and applied problems. For
-example, scientists could use `marlin` to consider the relative importance of
-different biological variables in affecting policy outcomes. Managers could use
-`marlin` to assess the potential of proposed MPA networks to achieve stated
-objectives, such as maximizing conservation gains for a threatened species while
-minimizing or removing negative impacts to affected fishing fleets. We envision
+impacts of realistic bio-economic dynamics on academic and applied problems. We envision
 `marlin` being applicable to research on dynamic ocean management, range shifts,
@@
 #| fig-cap: "Spawning biomass divided by unfished spawning biomass in the time period prior to implementation of MPAs for the pelagic case study.
-Simulated species include skipjack tuna, yellowfin tuna (*Thunnus albacares*), bigeye tuna, shortfin mako (*Isurus oxyrinchus*), swordfish (*Xiphias gladius*), albacore tuna (*Thunnus alalunga*), blue shark (*Prionace glauca*), silky shark (*Carcharhinus falciformis*), and oceanic whitetip shark (*Carcharhinus longimanus*)."
+Simulated species include skipjack tuna, yellowfin tuna, bigeye tuna, shortfin mako (*Isurus oxyrinchus*, Lamnidae), swordfish (*Xiphias gladius*, Xiphiidae), albacore tuna (*Thunnus alalunga*, 	Scombridae), blue shark (*Prionace glauca*, Carcharhinidae), silky shark (*Carcharhinus falciformis*, Carcharhinidae), and oceanic whitetip shark."
 

### 2023-11-13 03_marlin_paper.qmd
updating csl

@@
   cache: false
-csl: apa-6th-edition.csl
+csl: fish-and-fisheries.csl
 bibliography: references.bib

### 2023-11-14 03_marlin_paper.qmd
updating .bib file

--- 2023-11-02 03_marlin_paper.qmd
@@
   cache: false
-csl: apa-6th-edition.csl
+csl: fish-and-fisheries.csl
 bibliography: references.bib
 params:
-  run_name: ["v1.1"]
+  run_name: ["v1.2"]
 header-includes:

### 2023-11-27 03_marlin_paper.qmd
fixing recruitment typo and updating csl

--- 2023-11-02 03_marlin_paper.qmd
@@
   cache: false
-csl: apa-6th-edition.csl
+csl: apa.csl
 bibliography: references.bib
 params:
-  run_name: ["v1.1"]
+  run_name: ["v1.2"]
 header-includes:
@@
 $$ 
-log(\epsilon_{t,s}) \sim \begin{cases}
+\upsilon_{t,s} \sim \begin{cases}
        N(0,\sigma_{r,s}), & \text{if $t = 1$}\\
-       \rho_s log(\epsilon_{t-1,s}) + \sqrt{1 - \rho_s^2} N(0,\sigma_{r,s}), & \text{if $t > 1$}
+       \rho_s \upsilon_{t-1,s} + \sqrt{1 - \rho_s^2} N(0,\sigma_{r,s}), & \text{if $t > 1$}
   \end{cases}
@@
 $$
-\epsilon_{t,s} = e^{log(\epsilon_{t,s}) - \sigma_{r,s}^2/2}
+\epsilon_{t,s} = e^{\upsilon_{t,s} - \sigma_{r,s}^2/2}
 $$ {#eq-bias}

### 2024-03-27 03_marlin_paper.qmd
updating to reflect changes to marlin package

@@
 bibliography: references.bib
-params:
-  run_name: ["v1.2"]
 header-includes:
@@
 
+patch_biomass <- bind_cols(patch_biomass,marlin::process_step(patch_biomass$step))
+
  fig_1 <- patch_biomass %>%
@@
   ungroup() %>%
-  filter(between(step,1,2)) %>%
-  mutate(step = fct_recode(as.factor(step), "2" = "1.25", "3" = "1.5", "4" = "1.75")) %>% 
-  mutate(Season = step) %>% 
+  filter(between(year,1,2)) %>%
+  mutate(Season = season) %>% 
   ggplot()+
@@
 
+  ggsave(
+   filename = file.path(results_path, "fig_1.png"),
+   fig_1,
+   width = small_width,
+   units = "mm"
+ )
+
  fig_1
@@
 )
-
 critter_biomass <-
-  map_df(blue_sim, ~ map_df(.x, ~tibble(ssb = rowSums(.x$ssb_p_a), patch = 1:nrow(.x$ssb_p_a)), .id = "critter"), .id = "step") %>%
+  map_df(blue_sim, ~ map_df(.x, ~tibble(ssb = rowSums(.x$ssb_p_a), patch = 1:nrow(.x$ssb_p_a)), .id = "critter"), .id = "step")
+
+critter_biomass <- bind_cols(critter_biomass,marlin::process_step(critter_biomass$step)) %>%
   mutate(step = marlin::clean_steps(step)) %>%
   left_join(grid, by = "patch") %>% 
-  group_by(critter, step) %>% 
+  group_by(critter, year, season,time_step) %>% 
   summarise(ssb = sum(ssb)) %>% 
   group_by(critter) %>% 
-  mutate(depletion = ssb / ssb[step == 0])
+  mutate(depletion = ssb / ssb[time_step == min(time_step)])
 
@@
   rename(critter = common_name) %>%
-  filter(step == max(step)) %>% 
+  filter(time_step == max(time_step)) %>% 
   ggplot(aes(reorder(critter, depletion), depletion)) +
@@
    filename = file.path(results_path, "fig_2.pdf"),
+   fig_2,
+   width = small_width,
+   units = "mm"
+ )
+
+  ggsave(
+   filename = file.path(results_path, "fig_2.png"),
    fig_2,
@@
  ggsave(
-   filename = file.path(results_path, "fit_3.pdf"),
+   filename = file.path(results_path, "fig_3.pdf"),
+   fit_3,
+   width = big_width,
+   units = "mm"
+ )
+
+  ggsave(
+   filename = file.path(results_path, "fig_3.png"),
    fit_3,
@@
 
+ ggsave(
+   filename = file.path(results_path, "fig_4.png"),
+   fig_4,
+   width = big_width,
+   units = "mm"
+ )
+
 fig_4
@@
 
+  ggsave(
+   filename = file.path(results_path, "fig_5.png"),
+   fig_5,
+   width = big_width,
+   units = "mm"
+ )
+
 fig_5

### 2024-03-27 04_marlin_paper_si.qmd
updating to reflect changes to marlin package

--- 2023-10-07 04_marlin_paper_si.qmd
@@
 ```
+
+```{r}
+#| label: fig-coral-ssb
+#| fig-cap: "Spawning biomass divided by unfished spawning biomass in the time period prior to implementation of MPAs for the coral reef case study."
+
+critter_biomass <-
+  map_df(coral_sim, ~ map_df(.x, ~tibble(ssb = rowSums(.x$ssb_p_a), patch = 1:nrow(.x$ssb_p_a)), .id = "critter"), .id = "step")
+
+critter_biomass <- bind_cols(critter_biomass,marlin::process_step(critter_biomass$step)) %>%
+  mutate(step = marlin::clean_steps(step)) %>%
+  left_join(grid, by = "patch") %>% 
+  group_by(critter, year, season,time_step) %>% 
+  summarise(ssb = sum(ssb)) %>% 
+  group_by(critter) %>% 
+  mutate(depletion = ssb / ssb[time_step == min(time_step)])
+
+fig_s2 <- critter_biomass %>%
+  ungroup() %>% 
+  filter(time_step == max(time_step)) %>% 
+  ggplot(aes(reorder(critter, depletion), depletion)) +
+  geom_col() +
+  scale_y_continuous(limits = c(0, 1), name = "SSB/SSB0", expand = expansion(mult = c(0.0,0.05))) +
+  labs(x = "") + 
+  coord_flip()
+
+fig_s2
+```

### 2024-03-27 00_setup.R

@@ -42,7 +42,6 @@ foos <- list.files(here("functions"))
 
 walk(foos, ~ source(here("functions", .x)))
 
-print(Sys.getenv("RUN_NAME"))
 if (Sys.getenv("RUN_NAME") == ''){
   run_name <- "v1.1"
 

### 2024-03-27 functions/run_mpa_experiment.R

@@ -197,18 +197,17 @@ run_mpa_experiment <-
     #
     # run MPA simulation
 
-    # starting_step = as.numeric(last(names(starting_conditions)))
+    starting_step = marlin::clean_steps(last(names(starting_conditions)))
 
-    starting_step = as.numeric(gsub("step_","",last(names(starting_conditions))))
-
+    processed_step <- marlin::process_step(last(names(starting_conditions)))
 
     mpa_sim <- simmar(
       fauna = fauna,
       fleets = fleets,
       years = years,
       manager = list(mpas = list(locations = mpas,
-                  mpa_year = ceiling(starting_step)),
-                  effort_cap = effort_cap),
+                                 mpa_year = processed_step$year + 1),
+                     effort_cap = effort_cap),
       habitat = future_habitat,
       starting_step = starting_step,
       keep_starting_step = FALSE,

### 2024-03-29 03_marlin_paper.qmd
updating to reflect new marlin details

@@
   theme(
-    legend.position = c(0.73, .73),
+    legend.position = c(0.7, .73),
     legend.background = element_rect(fill = "transparent"),

### 2024-03-29 functions/run_mpa_experiment.R

@@ -204,7 +204,7 @@ run_mpa_experiment <-
     mpa_sim <- simmar(
       fauna = fauna,
       fleets = fleets,
-      years = years,
+      steps = years * seasons + 1,
       manager = list(mpas = list(locations = mpas,
                                  mpa_year = processed_step$year + 1),
                      effort_cap = effort_cap),
@@ -253,7 +253,7 @@ run_mpa_experiment <-
 
     final_step <- mpa_sim[[length(mpa_sim)]]
 
-    b_p = map_df(final_step, ~ data.frame(b_p = rowSums(.x$b_p_a), patch = 1:nrow(.x$b_p_a)), .id = "critter")
+    b_p <-  map_df(final_step, ~ data.frame(b_p = rowSums(.x$b_p_a), patch = 1:nrow(.x$b_p_a)), .id = "critter")
 
     response_ratio <- b_p |>
       group_by(patch) |>
